### 2019-06-03 latex/acl2019.tex
Added latex paper (WIP)

%
% File acl2019.tex
%
%% Based on the style files for ACL 2018, NAACL 2018/19, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2019}
\usepackage{times}
\usepackage{latexsym}
\usepackage{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\usepackage{graphicx}
\usepackage[caption=false]{subfig}

% \aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B\textsc{ib}\TeX}

\title{Comparative study of Dask and Spark for neuroimaging pipelines}

\author{Mathieu Dugr{\'e} \\
  Concordia University \\
  \texttt{mathieu.dugre@mail.concordia.ca}\\}

\date{}

\begin{document}
\maketitle
\begin{abstract}

% Topic
As the amount of data increases and is easier to access, Big Data processing becomes critical in neuroimaging. 
% Problem
This is an issue as the currently used frameworks are specialized for neuroimaging and not Big Data concerns. This can lead to performance decrease or limit the research we can perform.
% Relevance
The usage of Big Data framework is beneficial to this problem.
The state-of-the-art general purpose Big Data framework Spark has its code base written in Scala, while our laboratory mostly uses Python. This can be problematic since it is harder to port our pipelines to the framework. Also, theoretically, it leads to performance decrease.

% Approach
We propose to use Dask since it provides native Python parallelism to pipelines while providing support to familiar API from the Python's scientific ecosystem. However there are few comparisons of Spark and Dask; especially in neuroimaging. Moreover, these studies were done when Dask was still an immature framework which make those study unfair nowadays. This is our motivation to compare the latest version of Dask with Spark.
% Methods (one sentence)
To evaluate the frameworks, we focus on their performance and their scheduler.

% Key Impact
Our study demonstrates the potential of using Dask as the framework to build neuroimaging pipelines.

\end{abstract}

\section{Introduction}
% Context
With the increasing number of data available in neuroimaging \citep{ALFAROALMAGRO:18, UKBioBank:18} the processing of Big Data is becoming critical. Frameworks like Nipype \citep{Nipype:11} are usually used to create neuroimaging pipelines however it is worth considering the use of general purpose Big Data frameworks \citep{Hayot-Sasson:17}. In that research, Spark \citep{Spark:16} was a natural choice as it is the state-of-the-art framework for Big Data. However, as our laboratory works mostly with Python, we want to know if Dask \citep{Dask:15}, a Python written framework, could bring better or similar performances while facilitating usage.

%Similarities
Spark and Dask offer in-memory computing, data locality and lazy evaluation; which is common for Big Data framework. Both their scheduler operate dynamically. This is good when the runtimes are not known ahead of time \citep{Dask:15}. Over these similarities, the frameworks are quite different.

% Spark
On the one hand, we have Spark which provides a high-level API. This allows the scheduler to perform more optimizations which makes it well suited for neuroimaging applications that often require lengthy executions. However, Spark's code base is in Scala which theoretically can lead to slow down in execution due to a required serialization. Moreover, while Spark's API is flexible and allows most implementations, it differs from the ones seen in the Python's ecosystem.

% Dask
On the other hand, Dask was created with the purpose to natively parallelize Python pipelines while keeping the syntax of familiar API from the Python's scientific ecosystem. However, Dask is still a young framework with work to be done; it's API does not completely replicate the library it supports. While its lower-level API allows the implementation of more complex algorithms it sacrifice a layer of optimization.
% Related work
Previous work by \citet{Mehta:17} shows that Dask had significant overhead and was hard to debug. 
% What issues your work addresses
Dask was immature at the time and a lot of change was brought to the framework. Therefore we think it is valuable to re-compare it with Spark.

%% Dask API
The Dask API we decide to use for our comparison are Dask bag, delayed and futures. Dask bag offer an easy API to parallelize data. Dask delayed offer a lower-level API that offer more flexibility; this is good to implement more complex task that do not fit in the Dask bag framework. Dask futures is a real time API. Both Dask bag and delayed apply lazy evaluation to tasks while futures trigger them directly.

% Methods used (summary)
The project aims to compare the state-of-the-art general purpose framework Spark with the newcomer Dask. We decide to compare the performance of Spark and Dask on custom incrementation pipeline to simplify the effect of the algorithm on the comparison. Then we access the frameworks on two real-life applications: (1) histogram of the voxels intensity in a 3d images (2) BIDS example.

% Implications of our research
The result from our project help in assessing if Dask is a good choice to build neuroimaging pipelines.

%%%%% MATERIAL AND METHODS %%%%%%
\section{Material and Methods}
%% Material
We compare Dask(\textbf{v1.1.4}) and Spark(\textbf{v2.4.0}) based on two metrics their scheduling and their makespan performance. All scripts used to perform our experiments are available at \href{https://github.com/mathdugre/SOEN-499-Project}{https://github.com/mathdugre/SOEN-499-Project}. For our experiments we used Compute Canada Cloud where each instance has 8 cores and 30GB RAM. The system was dedicated.

%% Incrementation
\subsection{Incrementation}
As our first experiment to benchmark the frameworks we used a simple incrementation pipeline (see Algorithm \ref{alg:incrementation}). It loads chunks from the 75GB BigBrain\citep{ds052:01} image, increment its value by 1 and save the result to an nfs as a Nifti image. The incrementation allows to reduce caching effect of the data. This experiment allows us to study the behaviour of the frameworks when all input are processed independently; i.e. each task in the graph only depend on the previous one (see \ref{fig:tg-inc}).

This experiment refers to  the implementation of the Spark pipelines purposed in \citet{Hayot-Sasson:17}.

\begin{algorithm}\caption{Incrementation}\label{alg:incrementation}
    \begin{algorithmic}
    \Require $x$, a sleep delay in float
    \Require $file$, a file containing a chunk
    \Require $fs$, nfs to save image to.
    \State read $chunk$ from $file$
    \ForEach{$i \in iterations$}
    \ForEach{$chunk \in image$}
        \State $chunk\gets chunk+1$
        \State sleep $x$
    \EndFor
    \EndFor
    \State save $chunk$ to $fs$
\end{algorithmic}
\end{algorithm}

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.125\textwidth, angle=-90]{images/incrementation-task-graph.png}
    \caption{Task graph for Incrementation}
    \label{fig:tg-inc}
\end{figure}

% Histogram
\subsection{Histogram}
For our second experiment we find the frequency of the voxels intensity in BigBrain and save them as a tuple in a file. In this experiment the end results depends on all of the input results (see \ref{fig:tg-histo}). This allow us to study the behaviour of the framework when task are dependent. Also this application require data shuffling thus inter worker communication. In this application we decided to only compare Spark against the Dask bag and array API as they are the more realistic for this use case.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.16\textwidth, angle=-90]{images/histogram-task-graph.png}
    \caption{Task graph for Histogram}
    \label{fig:tg-histo}
\end{figure}

% BIDS Example
\subsection{BIDS Example}
Our last experiment compare both frameworks on the BIDS Example algorithm.
% Specify more details and the data set used

% Note
Note that the code implemented might not be optimal. However we hope to provide a baseline assessment of Dask and Spark.

%% Performance
\subsection{Experiments}
Our experiments have for goal to see how the frameworks react to different environments and tasks types.

%% Incrementation
For the incrementation experiment, we
%% Node
first vary the number of instances in our cluster environment. We test with 1, 2, 4, 8, and 16 instances. From this we can evaluate the level of parallelism offered by both framework.

%% Iterations
Secondly, we modify the number of iterations we perform on the increment task between the experiment. We use 1, 10 and 100 iterations. This allow the study of the schedulers different amount of task are to be performed.

%% Sleep delay
Thirdly, we change the sleep delay in the increment tasks. We fix a 1, 4 and 16 and 64 seconds sleep time for each iterations. This allow us to test the effect of tasks with different time length.

%% Splits
Finally, we split the BigBrain in different number of chunks. We split it in 30, 125 and 750 chunks. From that we can evaluate the framework with task of different size; around 2.5GB, 0.6GB and 0.1GB, respectively.

%% Baseline
Our baseline for all experiment is 10 iterations, 4 sleep delay, 125 chunks running on 8 instances.

%% Histogram
For the histogram experiment, we vary the number of instances, as above. We also change the number of splits as above.

%% BIDS Example
For the BIDS example experiment, we vary the number of instances, again as above.

%%%%% RESULTS %%%%%
\section{Results}

%%% INCREMENTATION %%%
%% Instances
\subsection{Experiment 1: Number of instances}
Figure \ref{fig:inc_node} shows that has the number of node increase the performance decrease. The reason we think this happen is that the number of workers becomes larger than the number of tasks to process. Also, as more worker access the nfs simultaneously this increase the stress put on the system which result in bottleneck, errors and slow down. On the one hand, Dask seems better when using a single worker. On the other hand, Spark makespan decrease faster for low level of parallelism. However Dask quickly catch up at high level of parallelism. This makes us think that the scheduler in Dask might be better

\begin{figure}[p]
    \centering
    \subfloat[Incrementation makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/Incrementation_instances.png}%
    }

    \subfloat[Incrementation total time]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_idle_instances.png}%
    }
    \caption{125 splits, 10 iterations, 4 sec. sleep delay}
    \label{fig:inc_node}
\end{figure}

%% Iterations
\subsection{Experiment 1: Number of iterations}
In Figure \ref{fig:inc_itr} Dask futures grows faster than the other API. This is due to the lack of optimization on the real time processing of tasks. The Dask bag with graph optimization seems to perform worst than other as the number of tasks increase. This might show that over optimizing task graph results in performance decrease. When the task load start to grow slightly Dask bag seems to be the best option. This could be explained as Dask bag is the most optimized Dask's API that we compare. Also Spark serialization could slow down the execution as the number of task increase thus reducing its optimization effects

\begin{figure}[p]
    \centering
    \subfloat[Incrementation makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/Incrementation_iterations.png}%
    }

    \subfloat[Incrementation total time]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_idle_iterations.png}%
    }
    \caption{125 splits, 4 sec. sleep delay, 8 instances}
    \label{fig:inc_itr}
\end{figure}

%% Sleep time
\subsection{Experiment 1: Sleep delay}
As Figure \ref{fig:inc_sleep} shows, Dask is way faster than Spark for any task length. We can consider optimising the graph for shorter task but it seams that normal Dask bag have better optimization policy for longer task. However the results are surprising since the makespan should increase as the sleep delay increase but we see the inverse effect.

\begin{figure}[p]
    \centering
    \subfloat[Incrementation makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/Incrementation_sleeptime.png}%
    }

    \subfloat[Incrementation total time]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleeptime.png}%
    }
    \caption{125 splits, 10 iterations, 8 instances}
    \label{fig:inc_sleep}
\end{figure}

% Chunks
\subsection{Experiment 1: Number of splits}
In Figure \ref{fig:inc_chunk} Spark is not compared for 30 chunks. This is because spark has a 2GB limitation in the task size it can compute. All API have a longer make span at 30 chunks compared to 125 chunks. This is because the 8 instances each have 8 cores thus merely half the resources available are used. At 125 chunks Dask bag are the fastest. This could be explained by Dask bag having the better optimization of the Dask's API and that Spark has to perform serialization to Java. At 750 chunks Spark becomes as good

\begin{figure}[p]
    \centering
    \subfloat[Incrementation makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/Incrementation_splits.png}%
    }

    \subfloat[Incrementation total time]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_idle_splits.png}%
    }
    \caption{10 iterations, 4 sec. sleep delay, 8 instances}
    \label{fig:inc_chunk}
\end{figure}

From the Figures \ref{fig:node} to \ref{fig:chunk}  the Dask bag with optimized graph has high variance on its makespan. Moreover, Dask API seems to always offer a better option than spark RDDs.

%% Baseline gantt chart
\subsection{Experimentation 1: Baseline timeline}
From Figure \ref{fig:dask_gantt} and Figure \ref{fig:spark_gantt} the execution time is similar in Dask and Spark. However Spark takes more time to save the data to disk. This could be due to the FIFO scheduler that tends to have the first set of save consequent thus saturating the nfs. Also serialization could have an effect. It looks like the nfs could be a bottleneck when reading (writing) data from (to) the application.

Dask scheduler seems to have strange behaviour in Figure \ref{fig:dask_gantt}. It often wait before scheduling tasks that we would expect to be scheduled right away. We currently don't have explication for this behaviour. This could be explained by task rescheduled to other workers.

Also, the same length does not seem to be homogeneous. This might be due to caching effect either with the main memory or the nfs.

\begin{figure}[p]
    \centering
    \subfloat[Dask execution timeline]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_dask-bag_baseline_gantt.png}%
    }

    \subfloat[Spark execution timeline]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_spark_baseline_gantt.png}%
    }
    \caption{125 splits, 4 sec. sleep delay, 8 instances}
    \label{fig:inc_gantt}
\end{figure}

%%% HISTOGRAM %%%
\subsection{Experiment 2: Number of instances}

\begin{figure}[p]
    \centering
    \subfloat[Histogram makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/histo_instance.png}%
    }

    \subfloat[Histogram total time]{%
      \includegraphics[clip,width=\columnwidth]{images/histo_idle_instances.png}%
    }
    \caption{8 instances}
    \label{fig:histo_instance}
\end{figure}

\subsection{Experiment 2: Number of splits (task size)}

\begin{figure}[p]
    \centering
    \subfloat[Histogram makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/histo_splits.png}%
    }

    \subfloat[Histogram total time]{%
      \includegraphics[clip,width=\columnwidth]{images/histo_idle_splits.png}%
    }
    \caption{10 iterations, 4 sec. sleep delay, 8 instances}
    \label{fig:histo_splits}
\end{figure}

\subsection{Experiment 2: Number of chunks}

%%% BIDS EXAMPLE %%%
\subsection{Experiment 3: Number of instances}

\subsection{Experiment 3: Number of subjects}

%%%%% DISCUSSION %%%%%
\section{Discussion}
\subsection{Serialization}
We think that serialization could have an important effect on the makespan of the application. This could potentially slow down the application when a lot of tasks are scheduled.

\subsection{NFS}
The nfs seems to be a source of bottleneck. We think that exploring other file system, like Lustre, could give us insight the actual effect of the nfs on the application makespan. We also think that the nfs caching could play an effect on the task of different length.

\subsection{Scheduler}
The Dask scheduler seems to have some bizarre behaviour. It seems like it waits to schedule task. However it could also be due to tasks being scheduled on other workers thus requiring data to be sent over the network; which would explain the stall on the worker. More work would be needed to investigate that issue.

\subsection{Caching}
We think that caching plays an effect on the results. As seen in the gantt chart previously, some of the read tasks are significantly shorter than other while all chunks are of equal size.

%% Future work
\subsection{Future work}
There is definitely a lot of work to be done to assess exactly what is happening at the scheduling level for both Dask and Spark. For the future we plan to test the effect of caching on the system. We would also need to explore caching and saturation of the nfs. We also need to explore different scheduler (e.g. Yarn, Mesos) on both Spark and Dask to see if our result are mostly affected by the scheduler or outside factors. Afterwards, we want to continue the analysis by evaluating the memory consumption of the frameworks. Moreover we want to explore more complex algorithm including some related to neuroimaging. We also plan to investigate which library would be better in which context.

\section{Acknowledgement}
We would like to thanks Compute Canada Cloud for the the infrastructure.

\bibliography{acl2019}
\bibliographystyle{acl_natbib}

\end{document}

### 2019-06-06 latex/acl2019.tex
fixed formatting

@@
 % Topic
-As the amount of data increases and is easier to access, Big Data processing becomes critical in neuroimaging. 
+As the amount of data increases and is easier to access, Big Data processing becomes
+critical in neuroimaging.
 % Problem
-This is an issue as the currently used frameworks are specialized for neuroimaging and not Big Data concerns. This can lead to performance decrease or limit the research we can perform.
+This is an issue as the currently used frameworks are specialized for neuroimaging
+and not Big Data concerns. This can lead to performance decrease or limit the
+research we can perform.
 % Relevance
-The usage of Big Data framework is beneficial to this problem.
-The state-of-the-art general purpose Big Data framework Spark has its code base written in Scala, while our laboratory mostly uses Python.
-This can be problematic since it is harder to port our pipelines to the framework.
-Also, theoretically, it leads to performance decrease.
+The usage of Big Data framework is beneficial to this problem. The state-of-the-art
+general purpose Big Data framework Spark has its code base written in Scala, while
+our laboratory mostly uses Python. This can be problematic since it is harder to port
+our pipelines to the framework. Also, theoretically, it leads to performance
+decrease.
 
 % Approach
-We propose to use Dask since it provides native Python parallelism to pipelines while providing support to familiar API from the Python's scientific ecosystem.
-However there are few comparisons of Spark and Dask; especially in neuroimaging.
-Moreover, these studies were done when Dask was still an immature framework which make those study unfair nowadays.
-This is our motivation to compare the latest version of Dask with Spark.
+We propose to use Dask since it provides native Python parallelism to pipelines while
+providing support to familiar API from the Python's scientific ecosystem. However
+there are few comparisons of Spark and Dask; especially in neuroimaging. Moreover,
+these studies were done when Dask was still an immature framework which make those
+study unfair nowadays. This is our motivation to compare the latest version of Dask
+with Spark.
 % Methods (one sentence)
@@
 % Key Impact
-Our study demonstrates the potential of using Dask as the framework to build neuroimaging pipelines.
+Our study demonstrates the potential of using Dask as the framework to build
+neuroimaging pipelines.
 
@@
 % Context
-With the increasing number of data available in neuroimaging \citep{ALFAROALMAGRO:18, UKBioBank:18} the processing of Big Data is becoming critical.
-Frameworks like Nipype \citep{Nipype:11} are usually used to create neuroimaging pipelines however it is worth considering the use of general purpose Big Data frameworks \citep{Hayot-Sasson:17}.
-In that research, Spark \citep{Spark:16} was a natural choice as it is the state-of-the-art framework for Big Data.
-However, as our laboratory works mostly with Python, we want to know if Dask \citep{Dask:15}, a Python written framework, could bring better or similar performances while facilitating usage.
+With the increasing number of data available in neuroimaging \citep{ALFAROALMAGRO:18,
+UKBioBank:18} the processing of Big Data is becoming critical. Frameworks like Nipype
+\citep{Nipype:11} are usually used to create neuroimaging pipelines however it is
+worth considering the use of general purpose Big Data frameworks
+\citep{Hayot-Sasson:17}. In that research, Spark \citep{Spark:16} was a natural
+choice as it is the state-of-the-art framework for Big Data. However, as our
+laboratory works mostly with Python, we want to know if Dask \citep{Dask:15}, a
+Python written framework, could bring better or similar performances while
+facilitating usage.
 
 %Similarities
-Spark and Dask offer in-memory computing, data locality and lazy evaluation; which is common for Big Data framework.
-Both their scheduler operate dynamically.
-This is good when the runtimes are not known ahead of time \citep{Dask:15}.
-Over these similarities, the frameworks are quite different.
+Spark and Dask offer in-memory computing, data locality and lazy evaluation; which is
+common for Big Data framework. Both their scheduler operate dynamically. This is good
+when the runtimes are not known ahead of time \citep{Dask:15}. Over these
+similarities, the frameworks are quite different.
 
 % Spark
-On the one hand, we have Spark which provides a high-level API.
-This allows the scheduler to perform more optimizations which makes it well suited for neuroimaging applications that often require lengthy executions.
-However, Spark's code base is in Scala which theoretically can lead to slow down in execution due to a required serialization.
-Moreover, while Spark's API is flexible and allows most implementations, it differs from the ones seen in the Python's ecosystem.
+On the one hand, we have Spark which provides a high-level API. This allows the
+scheduler to perform more optimizations which makes it well suited for neuroimaging
+applications that often require lengthy executions. However, Spark's code base is in
+Scala which theoretically can lead to slow down in execution due to a required
+serialization. Moreover, while Spark's API is flexible and allows most
+implementations, it differs from the ones seen in the Python's ecosystem.
 
 % Dask
-On the other hand, Dask was created with the purpose to natively parallelize Python pipelines while keeping the syntax of familiar API from the Python's scientific ecosystem.
-However, Dask is still a young framework with work to be done; it's API does not completely replicate the library it supports.
-While its lower-level API allows the implementation of more complex algorithms it sacrifice a layer of optimization.
+On the other hand, Dask was created with the purpose to natively parallelize Python
+pipelines while keeping the syntax of familiar API from the Python's scientific
+ecosystem. However, Dask is still a young framework with work to be done; it's API
+does not completely replicate the library it supports. While its lower-level API
+allows the implementation of more complex algorithms it sacrifice a layer of
+optimization.
 % Related work
-Previous work by \citet{Mehta:17} shows that Dask had significant overhead and was hard to debug. 
+Previous work by \citet{Mehta:17} shows that Dask had significant overhead and was
+hard to debug.
 % What issues your work addresses
-Dask was immature at the time and a lot of change was brought to the framework. Therefore we think it is valuable to re-compare it with Spark.
+Dask was immature at the time and a lot of change was brought to the framework.
+Therefore we think it is valuable to re-compare it with Spark.
 
@@
 The Dask API we decide to use for our comparison are Dask bag, delayed and futures.
-Dask bag offer an easy API to parallelize data.
-Dask delayed offer a lower-level API that offer more flexibility; this is good to implement more complex task that do not fit in the Dask bag framework.
-Dask futures is a real time API.
-Both Dask bag and delayed apply lazy evaluation to tasks while futures trigger them directly.
+Dask bag offer an easy API to parallelize data. Dask delayed offer a lower-level API
+that offer more flexibility; this is good to implement more complex task that do not
+fit in the Dask bag framework. Dask futures is a real time API. Both Dask bag and
+delayed apply lazy evaluation to tasks while futures trigger them directly.
 
 % Methods used (summary)
-The project aims to compare the state-of-the-art general purpose framework Spark with the newcomer Dask.
-We decide to compare the performance of Spark and Dask on custom incrementation pipeline to simplify the effect of the algorithm on the comparison.
-Then we access the frameworks on two real-life applications: (1) histogram of the voxels intensity in a 3d images (2) BIDS example.
+The project aims to compare the state-of-the-art general purpose framework Spark with
+the newcomer Dask. We decide to compare the performance of Spark and Dask on custom
+incrementation pipeline to simplify the effect of the algorithm on the comparison.
+Then we access the frameworks on two real-life applications: (1) histogram of the
+voxels intensity in a 3d images (2) BIDS example.
 
 % Implications of our research
-The result from our project help in assessing if Dask is a good choice to build neuroimaging pipelines.
+The result from our project help in assessing if Dask is a good choice to build
+neuroimaging pipelines.
 
@@
 %% Material
-We compare Dask(\textbf{v1.1.4}) and Spark(\textbf{v2.4.0}) based on two metrics their scheduling and their makespan performance.
-All scripts used to perform our experiments are available at \href{https://github.com/mathdugre/SOEN-499-Project}{https://github.com/mathdugre/SOEN-499-Project}.
-For our experiments we used Compute Canada Cloud where each instance has 8 cores and 30GB RAM.
-The system was dedicated.
+We compare Dask(\textbf{v1.1.4}) and Spark(\textbf{v2.4.0}) based on two metrics
+their scheduling and their makespan performance. All scripts used to perform our
+experiments are available at
+\href{https://github.com/mathdugre/SOEN-499-Project}{https://github.com/mathdugre/SOEN-499-Project}.
+For our experiments we used Compute Canada Cloud where each instance has 8 cores and
+30GB RAM. The system was dedicated.
 
@@
 \subsection{Incrementation}
-As our first experiment to benchmark the frameworks we used a simple incrementation pipeline (see Algorithm \ref{alg:incrementation}).
-It loads chunks from the 75GB BigBrain\citep{ds052:01} image, increment its value by 1 and save the result to an nfs as a Nifti image.
-The incrementation allows to reduce caching effect of the data.
-This experiment allows us to study the behaviour of the frameworks when all input are processed independently; i.e.
-each task in the graph only depend on the previous one (see \ref{fig:tg-inc}).
+As our first experiment to benchmark the frameworks we used a simple incrementation
+pipeline (see Algorithm \ref{alg:incrementation}). It loads chunks from the 75GB
+BigBrain\citep{ds052:01} image, increment its value by 1 and save the result to an
+nfs as a Nifti image. The incrementation allows to reduce caching effect of the data.
+This experiment allows us to study the behaviour of the frameworks when all input are
+processed independently; i.e. each task in the graph only depend on the previous one
+(see \ref{fig:tg-inc}).
 
-This experiment refers to  the implementation of the Spark pipelines purposed in \citet{Hayot-Sasson:17}.
+This experiment refers to the implementation of the Spark pipelines purposed in
+\citet{Hayot-Sasson:17}.
 
@@
 \subsection{Histogram}
-For our second experiment we find the frequency of the voxels intensity in BigBrain and save them as a tuple in a file.
-In this experiment the end results depends on all of the input results (see \ref{fig:tg-histo}).
-This allow us to study the behaviour of the framework when task are dependent.
-Also this application require data shuffling thus inter worker communication.
-In this application we decided to only compare Spark against the Dask bag and array API as they are the more realistic for this use case.
+For our second experiment we find the frequency of the voxels intensity in BigBrain
+and save them as a tuple in a file. In this experiment the end results depends on all
+of the input results (see \ref{fig:tg-histo}). This allow us to study the behaviour
+of the framework when task are dependent. Also this application require data
+shuffling thus inter worker communication. In this application we decided to only
+compare Spark against the Dask bag and array API as they are the more realistic for
+this use case.
 
@@
 % Note
-Note that the code implemented might not be optimal. However we hope to provide a baseline assessment of Dask and Spark.
+Note that the code implemented might not be optimal. However we hope to provide a
+baseline assessment of Dask and Spark.
 
@@
 \subsection{Experiments}
-Our experiments have for goal to see how the frameworks react to different environments and tasks types.
+Our experiments have for goal to see how the frameworks react to different
+environments and tasks types.
 
@@
 %% Node
-first vary the number of instances in our cluster environment. We test with 1, 2, 4, 8, and 16 instances. From this we can evaluate the level of parallelism offered by both framework.
+first vary the number of instances in our cluster environment. We test with 1, 2, 4,
+8, and 16 instances. From this we can evaluate the level of parallelism offered by
+both framework.
 
 %% Iterations
-Secondly, we modify the number of iterations we perform on the increment task between the experiment.
-We use 1, 10 and 100 iterations.
-This allow the study of the schedulers different amount of task are to be performed.
+Secondly, we modify the number of iterations we perform on the increment task between
+the experiment. We use 1, 10 and 100 iterations. This allow the study of the
+schedulers different amount of task are to be performed.
 
 %% Sleep delay
-Thirdly, we change the sleep delay in the increment tasks. We fix a 1, 4 and 16 and 64 seconds sleep time for each iterations. This allow us to test the effect of tasks with different time length.
+Thirdly, we change the sleep delay in the increment tasks. We fix a 1, 4 and 16 and
+64 seconds sleep time for each iterations. This allow us to test the effect of tasks
+with different time length.
 
 %% Splits
-Finally, we split the BigBrain in different number of chunks.
-We split it in 30, 125 and 750 chunks.
-From that we can evaluate the framework with task of different size; around 2.5GB, 0.6GB and 0.1GB, respectively.
+Finally, we split the BigBrain in different number of chunks. We split it in 30, 125
+and 750 chunks. From that we can evaluate the framework with task of different size;
+around 2.5GB, 0.6GB and 0.1GB, respectively.
 
 %% Baseline
-Our baseline for all experiment is 10 iterations, 4 sleep delay, 125 chunks running on 8 instances.
+Our baseline for all experiment is 10 iterations, 4 sleep delay, 125 chunks running
+on 8 instances.
 
 %% Histogram
-For the histogram experiment, we vary the number of instances, as above. We also change the number of splits as above.
+For the histogram experiment, we vary the number of instances, as above. We also
+change the number of splits as above.
 
@@
 \subsection{Experiment 1: Number of instances}
-Figure \ref{fig:inc_node} shows that has the number of node increase the performance decrease.
-The reason we think this happen is that the number of workers becomes larger than the number of tasks to process.
-Also, as more worker access the nfs simultaneously this increase the stress put on the system which result in bottleneck, errors and slow down.
-On the one hand, Dask seems better when using a single worker.
+Figure \ref{fig:inc_node} shows that has the number of node increase the performance
+decrease. The reason we think this happen is that the number of workers becomes
+larger than the number of tasks to process. Also, as more worker access the nfs
+simultaneously this increase the stress put on the system which result in bottleneck,
+errors and slow down. On the one hand, Dask seems better when using a single worker.
 On the other hand, Spark makespan decrease faster for low level of parallelism.
-However Dask quickly catch up at high level of parallelism.
-This makes us think that the scheduler in Dask might be better
+However Dask quickly catch up at high level of parallelism. This makes us think that
+the scheduler in Dask might be better
 
@@
 \subsection{Experiment 1: Number of iterations}
-In Figure \ref{fig:inc_itr} Dask futures grows faster than the other API.
-This is due to the lack of optimization on the real time processing of tasks.
-The Dask bag with graph optimization seems to perform worst than other as the number of tasks increase.
-This might show that over optimizing task graph results in performance decrease.
-When the task load start to grow slightly Dask bag seems to be the best option.
-This could be explained as Dask bag is the most optimized Dask's API that we compare.
-Also Spark serialization could slow down the execution as the number of task increase thus reducing its optimization effects
+In Figure \ref{fig:inc_itr} Dask futures grows faster than the other API. This is due
+to the lack of optimization on the real time processing of tasks. The Dask bag with
+graph optimization seems to perform worst than other as the number of tasks increase.
+This might show that over optimizing task graph results in performance decrease. When
+the task load start to grow slightly Dask bag seems to be the best option. This could
+be explained as Dask bag is the most optimized Dask's API that we compare. Also Spark
+serialization could slow down the execution as the number of task increase thus
+reducing its optimization effects
 
@@
 \subsection{Experiment 1: Sleep delay}
-As Figure \ref{fig:inc_sleep} shows, Dask is way faster than Spark for any task length.
-We can consider optimising the graph for shorter task but it seams that normal Dask bag have better optimization policy for longer task.
-However the results are surprising since the makespan should increase as the sleep delay increase but we see the inverse effect.
+As Figure \ref{fig:inc_sleep} shows, Dask is way faster than Spark for any task
+length. We can consider optimising the graph for shorter task but it seams that
+normal Dask bag have better optimization policy for longer task. However the results
+are surprising since the makespan should increase as the sleep delay increase but we
+see the inverse effect.
 
@@
 \subsection{Experiment 1: Number of splits}
-In Figure \ref{fig:inc_chunk} Spark is not compared for 30 chunks.
-This is because spark has a 2GB limitation in the task size it can compute.
-All API have a longer make span at 30 chunks compared to 125 chunks.
-This is because the 8 instances each have 8 cores thus merely half the resources available are used.
-At 125 chunks Dask bag are the fastest.
-This could be explained by Dask bag having the better optimization of the Dask's API and that Spark has to perform serialization to Java.
+In Figure \ref{fig:inc_chunk} Spark is not compared for 30 chunks. This is because
+spark has a 2GB limitation in the task size it can compute. All API have a longer
+make span at 30 chunks compared to 125 chunks. This is because the 8 instances each
+have 8 cores thus merely half the resources available are used. At 125 chunks Dask
+bag are the fastest. This could be explained by Dask bag having the better
+optimization of the Dask's API and that Spark has to perform serialization to Java.
 At 750 chunks Spark becomes as good
@@
 
-From the Figures \ref{fig:node} to \ref{fig:chunk}  the Dask bag with optimized graph has high variance on its makespan. Moreover, Dask API seems to always offer a better option than spark RDDs.
+From the Figures \ref{fig:inc_node} to \ref{fig:inc_chunk} the Dask bag with optimized graph
+has high variance on its makespan. Moreover, Dask API seems to always offer a better
+option than spark RDDs.
 
@@
 \subsection{Experimentation 1: Baseline timeline}
-From Figure \ref{fig:dask_gantt} and Figure \ref{fig:spark_gantt} the execution time is similar in Dask and Spark.
-However Spark takes more time to save the data to disk.
-This could be due to the FIFO scheduler that tends to have the first set of save consequent thus saturating the nfs.
-Also serialization could have an effect.
-It looks like the nfs could be a bottleneck when reading (writing) data from (to) the application.
+From Figure \ref{fig:inc_gantt} the execution time
+is similar in Dask and Spark. However Spark takes more time to save the data to disk.
+This could be due to the FIFO scheduler that tends to have the first set of save
+consequent thus saturating the nfs. Also serialization could have an effect. It looks
+like the nfs could be a bottleneck when reading (writing) data from (to) the
+application.
 
-Dask scheduler seems to have strange behaviour in Figure \ref{fig:dask_gantt}.
-It often wait before scheduling tasks that we would expect to be scheduled right away.
-We currently don't have explication for this behaviour.
-This could be explained by task rescheduled to other workers.
+Dask scheduler seems to have strange behaviour in Figure \ref{fig:inc_gantt}. It
+often wait before scheduling tasks that we would expect to be scheduled right away.
+We currently don't have explication for this behaviour. This could be explained by
+task rescheduled to other workers.
 
-Also, the same length does not seem to be homogeneous. This might be due to caching effect either with the main memory or the nfs.
+Also, the same length does not seem to be homogeneous. This might be due to caching
+effect either with the main memory or the nfs.
 
@@
 \subsection{Serialization}
-We think that serialization could have an important effect on the makespan of the application. This could potentially slow down the application when a lot of tasks are scheduled.
+We think that serialization could have an important effect on the makespan of the
+application. This could potentially slow down the application when a lot of tasks are
+scheduled.
 
 \subsection{NFS}
-The nfs seems to be a source of bottleneck.
-We think that exploring other file system, like Lustre, could give us insight the actual effect of the nfs on the application makespan.
-We also think that the nfs caching could play an effect on the task of different length.
+The nfs seems to be a source of bottleneck. We think that exploring other file
+system, like Lustre, could give us insight the actual effect of the nfs on the
+application makespan. We also think that the nfs caching could play an effect on the
+task of different length.
 
 \subsection{Scheduler}
-The Dask scheduler seems to have some bizarre behaviour.
-It seems like it waits to schedule task.
-However it could also be due to tasks being scheduled on other workers thus requiring data to be sent over the network; which would explain the stall on the worker.
-More work would be needed to investigate that issue.
+The Dask scheduler seems to have some bizarre behaviour. It seems like it waits to
+schedule task. However it could also be due to tasks being scheduled on other workers
+thus requiring data to be sent over the network; which would explain the stall on the
+worker. More work would be needed to investigate that issue.
 
 \subsection{Caching}
-We think that caching plays an effect on the results. As seen in the gantt chart previously, some of the read tasks are significantly shorter than other while all chunks are of equal size.
+We think that caching plays an effect on the results. As seen in the gantt chart
+previously, some of the read tasks are significantly shorter than other while all
+chunks are of equal size.
 
@@
 \subsection{Future work}
-There is definitely a lot of work to be done to assess exactly what is happening at the scheduling level for both Dask and Spark.
-For the future we plan to test the effect of caching on the system.
-We would also need to explore caching and saturation of the nfs.
-We also need to explore different scheduler (e.g.
-Yarn, Mesos) on both Spark and Dask to see if our result are mostly affected by the scheduler or outside factors.
-Afterwards, we want to continue the analysis by evaluating the memory consumption of the frameworks.
-Moreover we want to explore more complex algorithm including some related to neuroimaging.
-We also plan to investigate which library would be better in which context.
+There is definitely a lot of work to be done to assess exactly what is happening at
+the scheduling level for both Dask and Spark. For the future we plan to test the
+effect of caching on the system. We would also need to explore caching and saturation
+of the nfs. We also need to explore different scheduler (e.g. Yarn, Mesos) on both
+Spark and Dask to see if our result are mostly affected by the scheduler or outside
+factors. Afterwards, we want to continue the analysis by evaluating the memory
+consumption of the frameworks. Moreover we want to explore more complex algorithm
+including some related to neuroimaging. We also plan to investigate which library
+would be better in which context.
 

### 2019-06-10 latex/acl2019.tex
Modified exp1 - # of instances and # of itr

@@
 of the framework when task are dependent. Also this application require data
-shuffling thus inter worker communication. In this application we decided to only
-compare Spark against the Dask bag and array API as they are the more realistic for
-this use case.
+shuffling thus inter worker communication. In this application we decided to omit the
+dask futures api as it is less realistic for this use case.
 
@@
 % Specify more details and the data set used
-
-% Note
-Note that the code implemented might not be optimal. However we hope to provide a
-baseline assessment of Dask and Spark.
 
@@
 
+% Note
+Note that the code implemented might not be optimal. However we hope to provide a
+baseline assessment of Dask and Spark.
+
 %%%%% RESULTS %%%%%
@@
 \subsection{Experiment 1: Number of instances}
-Figure \ref{fig:inc_node} shows that has the number of node increase the performance
-decrease. The reason we think this happen is that the number of workers becomes
-larger than the number of tasks to process. Also, as more worker access the nfs
-simultaneously this increase the stress put on the system which result in bottleneck,
-errors and slow down. On the one hand, Dask seems better when using a single worker.
-On the other hand, Spark makespan decrease faster for low level of parallelism.
-However Dask quickly catch up at high level of parallelism. This makes us think that
-the scheduler in Dask might be better
+Figure \ref{fig:inc_node}(a) shows that the incease in performance is not
+proportional to the number of worker which differ form what we would expect; it even
+decrease when we reach 8 instances (64 workers). This is explained by an increase in
+IO time corrolated to the number of workers (Figure \ref{fig:inc_node}(b)), as more
+worker access the nfs simultaneously it increase the stress put on the system which
+result in bottleneck, errors and slow down.
 
-\begin{figure}[p]
+On the one hand, Dask seems better when
+using a single worker. On the other hand, Spark makespan decrease faster for low
+level of parallelism. However Dask quickly catch up at high level of parallelism.
+This makes us think that the scheduler in Dask might be better
+
+\begin{figure}[htp]
     \centering
@@
 \subsection{Experiment 1: Number of iterations}
-In Figure \ref{fig:inc_itr} Dask futures grows faster than the other API. This is due
-to the lack of optimization on the real time processing of tasks. The Dask bag with
-graph optimization seems to perform worst than other as the number of tasks increase.
-This might show that over optimizing task graph results in performance decrease. When
-the task load start to grow slightly Dask bag seems to be the best option. This could
-be explained as Dask bag is the most optimized Dask's API that we compare. Also Spark
-serialization could slow down the execution as the number of task increase thus
-reducing its optimization effects
+In Figure \ref{fig:inc_itr} we see that spark and the dask bag api are equivalent in
+term of performance. The dask delayed api seems to perform well with low amount of
+task but grow faster the spark and dask bag. We think this is because dask delayed
+often schedule a block on different worker of the same instance which cause small
+delay everytime due to inter worker communication. This is good when there is fewer
+tasks as is make the io between block go out of sync hence lowering the io bottleneck
+observed earlier however as the number of tasks increases those small delays become
+significant. The dask futures api seems to outperform all the other apis but this is
+most likely because the tasks are not inter dependent thus this application benifits
+from the less optimal but faster scheduling brought by dask futures.
 
-\begin{figure}[p]
+\begin{figure}[htp]
     \centering
@@
 \subsection{Experiment 1: Sleep delay}
-As Figure \ref{fig:inc_sleep} shows, Dask is way faster than Spark for any task
-length. We can consider optimising the graph for shorter task but it seams that
-normal Dask bag have better optimization policy for longer task. However the results
-are surprising since the makespan should increase as the sleep delay increase but we
-see the inverse effect.
+As Figure \ref{fig:inc_sleep}
 
-\begin{figure}[p]
+\begin{figure}[htp]
     \centering
@@
 
-\begin{figure}[p]
+\begin{figure}[htp]
     \centering
@@
 \section{Discussion}
+\subsection{IO Time}
+
 \subsection{Serialization}

### 2019-06-10 histogram/dask-delayed_histogram.py
Improved performance tremendiously

@@ -9,6 +9,7 @@ from time import time
 
 import dask
 from dask.distributed import Client
+import numpy as np
 
 from utils import benchmark, crawl_dir, read_img
 
@@ -48,11 +49,10 @@ if __name__ == "__main__":
     def calculate_histogram(arr):
         start_time = time() - start
 
-        histogram = dict()
-        for x in arr:
-            if x not in histogram:
-                histogram[x] = 0
-            histogram[x] += 1
+        y = np.bincount(arr)
+        ii = np.nonzero(y)[0]
+        out = np.vstack((ii, y[ii])).T
+        histogram = {k: v for k, v in out}
 
         end_time = time() - start
 

### 2019-06-10 histogram/dask-array_histogram.py
Fixed path

@@ -83,7 +83,6 @@ if __name__ == "__main__":
     print(client)
     # Allow workers to use module
     client.upload_file("/nfs/paper-big-data-engines/utils.py")
-    client.upload_file("/nfs/paper-big-data-engines/kmeans/Kmeans.py")
 
     # Read images
     paths = crawl_dir(os.path.abspath(args.bb_dir))

### 2019-06-10 histogram/spark_histogram.py

@@ -38,7 +38,6 @@ if __name__ == "__main__":
     sc = SparkContext.getOrCreate(conf=conf)
 
     sc.addFile("/nfs/paper-big-data-engines/utils.py")
-    sc.addFile("/nfs/paper-big-data-engines/kmeans/Kmeans.py")
     print("Connected")
 
     # Read images

### 2019-06-10 histogram/dask-bag_histogram.py
Fixed path

@@ -42,7 +42,6 @@ if __name__ == "__main__":
     print(client)
     # Allow workers to use module
     client.upload_file("/nfs/paper-big-data-engines/utils.py")
-    client.upload_file("/nfs/paper-big-data-engines/kmeans/Kmeans.py")
 
     # Read images
     paths = crawl_dir(os.path.abspath(args.bb_dir))

### 2019-06-10 histogram/dask-delayed_histogram.py

@@ -1,6 +1,6 @@
 import sys
 
-sys.path.append("/nfs/SOEN-499-Project")
+sys.path.append("/nfs/paper-big-data-engines")
 
 import argparse
 from functools import reduce
@@ -43,8 +43,7 @@ if __name__ == "__main__":
 
     print(client)
     # Allow workers to use module
-    client.upload_file("/nfs/SOEN-499-Project/utils.py")
-    client.upload_file("/nfs/SOEN-499-Project/kmeans/Kmeans.py")
+    client.upload_file("/nfs/paper-big-data-engines/utils.py")
 
     def calculate_histogram(arr):
         start_time = time() - start

### 2019-06-10 latex/acl2019.tex
Modified exp1 - # of splits

@@
 delay everytime due to inter worker communication. This is good when there is fewer
-tasks as is make the io between block go out of sync hence lowering the io bottleneck
-observed earlier however as the number of tasks increases those small delays become
-significant. The dask futures api seems to outperform all the other apis but this is
-most likely because the tasks are not inter dependent thus this application benifits
-from the less optimal but faster scheduling brought by dask futures.
+tasks as is make the IO between block go out of sync hence lowering the IO bottleneck
+however as the number of tasks increases those small delays become significant. The
+dask futures api seems to outperform all the other apis but this is most likely
+because the tasks are not inter dependent thus this application benifits from the
+less optimal but faster scheduling brought by dask futures.
 
@@
 \subsection{Experiment 1: Number of splits}
-In Figure \ref{fig:inc_chunk} Spark is not compared for 30 chunks. This is because
-spark has a 2GB limitation in the task size it can compute. All API have a longer
-make span at 30 chunks compared to 125 chunks. This is because the 8 instances each
-have 8 cores thus merely half the resources available are used. At 125 chunks Dask
-bag are the fastest. This could be explained by Dask bag having the better
-optimization of the Dask's API and that Spark has to perform serialization to Java.
-At 750 chunks Spark becomes as good
+In Figure \ref{fig:inc_splits} Spark is not compared for 30 chunks. This is because
+spark has a 2GB limitation in the task size it can compute. Note from Figure
+\ref{fig:inc_splits}(b) that the compute time increase when there is more splits.
+This is expected because the sleep delay is constant throughout this experiment and
+there are more compute task when the number of splits increases. We also note that in
+Figure \ref{fig:inc_splits}(b) for 30 splits the dask bag api has a much lower
+overhead time. This is because calculate the idle time of active worker only and the
+dask bag api was only using one worker per split in comparison to the dask delayed
+and dask futures api which were offloading the calculations on multiple workers on a
+same instance. From Figure \ref{fig:inc_splits} we observe that when the dataset is
+seperated in more (smaller) splits it saves IO time however it increases the overhead
+by a similar amount; the inverse happens when we reduce teh number of splits.
 
@@
     \caption{10 iterations, 4 sec. sleep delay, 8 instances}
-    \label{fig:inc_chunk}
+    \label{fig:inc_splits}
 \end{figure}

### 2019-06-11 latex/acl2019.tex
Small fix

@@
 With the increasing number of data available in neuroimaging \citep{ALFAROALMAGRO:18,
-UKBioBank:18} the processing of Big Data is becoming critical. Frameworks like Nipype
+UKBioBank:18} the processing of Big Data becomes critical. Frameworks like Nipype
 \citep{Nipype:11} are usually used to create neuroimaging pipelines however it is
@@
 scheduler to perform more optimizations which makes it well suited for neuroimaging
-applications that often require lengthy executions. However, Spark's code base is in
-Scala which theoretically can lead to slow down in execution due to a required
-serialization. Moreover, while Spark's API is flexible and allows most
-implementations, it differs from the ones seen in the Python's ecosystem.
+analysis that often require the usage of pipeline with multiple steps. However,
+Spark's code base is in Scala which theoretically can lead to slow down in execution
+due to a required serialization. Moreover, while Spark's API is flexible and allows
+most implementations, it differs from the ones seen in the Python's ecosystem.
 
@@
 The project aims to compare the state-of-the-art general purpose framework Spark with
-the newcomer Dask. We decide to compare the performance of Spark and Dask on custom
+the newcomer Dask. We decide to compare the performance of Spark and Dask on a custom
 incrementation pipeline to simplify the effect of the algorithm on the comparison.
-Then we access the frameworks on two real-life applications: (1) histogram of the
+Then we assess the frameworks on two real-life applications: (1) histogram of the
 voxels intensity in a 3d images (2) BIDS example.
@@
 % Implications of our research
-The result from our project help in assessing if Dask is a good choice to build
+The result from our project help in deciding if Dask is a good choice to build
 neuroimaging pipelines.
@@
 experiments are available at
-\href{https://github.com/mathdugre/SOEN-499-Project}{https://github.com/mathdugre/SOEN-499-Project}.
-For our experiments we used Compute Canada Cloud where each instance has 8 cores and
-30GB RAM. The system was dedicated.
+\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines}.
+For our experiments we used Compute Canada Cloud where each instance has 8 cores,
+30GB RAM, 10GB bandwidht and is dedicated.
 
 %% Incrementation
-\subsection{Incrementation}
-As our first experiment to benchmark the frameworks we used a simple incrementation
-pipeline (see Algorithm \ref{alg:incrementation}). It loads chunks from the 75GB
-BigBrain\citep{ds052:01} image, increment its value by 1 and save the result to an
-nfs as a Nifti image. The incrementation allows to reduce caching effect of the data.
-This experiment allows us to study the behaviour of the frameworks when all input are
-processed independently; i.e. each task in the graph only depend on the previous one
-(see \ref{fig:tg-inc}).
+\subsection{Incrementation} As our first experiment to benchmark the frameworks we
+used a simple incrementation pipeline (see Algorithm \ref{alg:incrementation}). It
+loads chunks from the 75GB BigBrain\citep{ds052:01} image, increment its value by 1 a
+determined number of time and save the result to an nfs as a Nifti image.
+Incrementting the data reduce the caching effect. This experiment allows us to study
+the behaviour of the frameworks when all input are processed independently; i.e. each
+task in the graph only depend on the previous one (see \ref{Figure fig:tg-inc}).
 
@@
 and save them as a tuple in a file. In this experiment the end results depends on all
-of the input results (see \ref{fig:tg-histo}). This allow us to study the behaviour
+of the input results (see \ref{Figure fig:tg-histo}). This allows us to study the behaviour
 of the framework when task are dependent. Also this application require data
@@
 % Specify more details and the data set used
+% SECTION TO COMPLETE -- Experience not done yet
 
@@
 first vary the number of instances in our cluster environment. We test with 1, 2, 4,
-8, and 16 instances. From this we can evaluate the level of parallelism offered by
+and 8 instances. From this we can evaluate the level of parallelism offered by
 both framework.
@@
 %% Splits
-Finally, we split the BigBrain in different number of chunks. We split it in 30, 125
-and 750 chunks. From that we can evaluate the framework with task of different size;
-around 2.5GB, 0.6GB and 0.1GB, respectively.
+Finally, we seperate the BigBrain in different number of splits. We seperate it in
+30, 125 and 750 splits; around 2.5GB, 0.6GB and 0.1GB, respectively. From that we can
+evaluate the framework with task of different size and IO requirements.
 
 %% Baseline
-Our baseline for all experiment is 10 iterations, 4 sleep delay, 125 chunks running
+Our baseline for all experiment is 10 iterations, 4 sleep delay, 125 splits running
 on 8 instances.
@@
 proportional to the number of worker which differ form what we would expect; it even
-decrease when we reach 8 instances (64 workers). This is explained by an increase in
-IO time corrolated to the number of workers (Figure \ref{fig:inc_node}(b)), as more
-worker access the nfs simultaneously it increase the stress put on the system which
-result in bottleneck, errors and slow down.
+decrease when we reach 8 workers. This is explained by an increase in IO overhead
+corrolated to the number of workers (Figure \ref{fig:inc_node}(b)), as more worker
+access the nfs simultaneously it increase the stress put on the system which result
+in bottleneck, errors and slow down.
 
-On the one hand, Dask seems better when
-using a single worker. On the other hand, Spark makespan decrease faster for low
-level of parallelism. However Dask quickly catch up at high level of parallelism.
-This makes us think that the scheduler in Dask might be better
+On the one hand, Dask seems better when using a single worker. On the other hand,
+Spark makespan decrease faster for higher level of parallelism. However Dask is
+constently faster then spark which make us think it has a better scheduler.
 
@@
 task but grow faster the spark and dask bag. We think this is because dask delayed
-often schedule a block on different worker of the same instance which cause small
-delay everytime due to inter worker communication. This is good when there is fewer
+often schedule a block on different thread of the same worker which cause small delay
+everytime due to inter thread data communication. This is good when there is fewer
 tasks as is make the IO between block go out of sync hence lowering the IO bottleneck
@@
 As Figure \ref{fig:inc_sleep} 
+% Not sure what to conclude.
 
@@
 Figure \ref{fig:inc_splits}(b) for 30 splits the dask bag api has a much lower
-overhead time. This is because calculate the idle time of active worker only and the
-dask bag api was only using one worker per split in comparison to the dask delayed
-and dask futures api which were offloading the calculations on multiple workers on a
-same instance. From Figure \ref{fig:inc_splits} we observe that when the dataset is
+overhead time. This is because we only calculate the idle time of used core and the
+dask bag api was only using one thread per block in comparison to the dask delayed
+and dask futures api which were offloading the calculations on multiple thread of a
+same worker. From Figure \ref{fig:inc_splits} we observe that when the dataset is
 seperated in more (smaller) splits it saves IO time however it increases the overhead

### 2019-06-11 latex/acl2019.tex
Small fix

@@
 \subsection{Experiment 1: Number of instances}
-Figure \ref{fig:inc_node}(a) shows that the incease in performance is not
+Figure \ref{fig:inc_worker}(a) shows that the incease in performance is not
 proportional to the number of worker which differ form what we would expect; it even
 decrease when we reach 8 workers. This is explained by an increase in IO overhead
-corrolated to the number of workers (Figure \ref{fig:inc_node}(b)), as more worker
+corrolated to the number of workers (Figure \ref{fig:inc_worker}(b)), as more worker
 access the nfs simultaneously it increase the stress put on the system which result
@@
     \subfloat[Incrementation makespan]{%
-      \includegraphics[clip,width=\columnwidth]{images/Incrementation_instances.png}%
+      \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
     }
@@
     \subfloat[Incrementation total time]{%
-      \includegraphics[clip,width=\columnwidth]{images/inc_idle_instances.png}%
+      \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
     }
     \caption{125 splits, 10 iterations, 4 sec. sleep delay}
-    \label{fig:inc_node}
+    \label{fig:inc_worker}
 \end{figure}
@@
     \subfloat[Incrementation makespan]{%
-      \includegraphics[clip,width=\columnwidth]{images/Incrementation_iterations.png}%
+      \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
     }
@@
     \subfloat[Incrementation total time]{%
-      \includegraphics[clip,width=\columnwidth]{images/inc_idle_iterations.png}%
+      \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
     }
@@
     \subfloat[Incrementation makespan]{%
-      \includegraphics[clip,width=\columnwidth]{images/Incrementation_sleeptime.png}%
+      \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
     }
@@
     \subfloat[Incrementation total time]{%
-      \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleeptime.png}%
+      \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
     }
@@
 \subsection{Experiment 1: Number of splits}
-In Figure \ref{fig:inc_splits} Spark is not compared for 30 chunks. This is because
+In Figure \ref{fig:inc_chunk} Spark is not compared for 30 chunks. This is because
 spark has a 2GB limitation in the task size it can compute. Note from Figure
-\ref{fig:inc_splits}(b) that the compute time increase when there is more splits.
+\ref{fig:inc_chunk}(b) that the compute time increase when there is more splits.
 This is expected because the sleep delay is constant throughout this experiment and
 there are more compute task when the number of splits increases. We also note that in
-Figure \ref{fig:inc_splits}(b) for 30 splits the dask bag api has a much lower
+Figure \ref{fig:inc_chunk}(b) for 30 splits the dask bag api has a much lower
 overhead time. This is because we only calculate the idle time of used core and the
@@
 and dask futures api which were offloading the calculations on multiple thread of a
-same worker. From Figure \ref{fig:inc_splits} we observe that when the dataset is
+same worker. From Figure \ref{fig:inc_chunk} we observe that when the dataset is
 seperated in more (smaller) splits it saves IO time however it increases the overhead
@@
     \subfloat[Incrementation makespan]{%
-      \includegraphics[clip,width=\columnwidth]{images/Incrementation_splits.png}%
+      \includegraphics[clip,width=\columnwidth]{images/inc_chunk.png}%
     }
@@
     \subfloat[Incrementation total time]{%
-      \includegraphics[clip,width=\columnwidth]{images/inc_idle_splits.png}%
+      \includegraphics[clip,width=\columnwidth]{images/inc_idle_chunk.png}%
     }
     \caption{10 iterations, 4 sec. sleep delay, 8 instances}
-    \label{fig:inc_splits}
+    \label{fig:inc_chunk}
 \end{figure}
 
-From the Figures \ref{fig:inc_node} to \ref{fig:inc_chunk} the Dask bag with optimized graph
+From the Figures \ref{fig:inc_worker} to \ref{fig:inc_chunk} the Dask bag with optimized graph
 has high variance on its makespan. Moreover, Dask API seems to always offer a better
@@
     \centering
-    \subfloat[Dask execution timeline]{%
-      \includegraphics[clip,width=\columnwidth]{images/inc_dask-bag_baseline_gantt.png}%
+    \subfloat[Spark execution timeline]{%
+      \includegraphics[clip,width=\columnwidth]{images/spark_inc_baseline_gantt.png}%
     }
 
-    \subfloat[Spark execution timeline]{%
-      \includegraphics[clip,width=\columnwidth]{images/inc_spark_baseline_gantt.png}%
+    \subfloat[Dask bag execution timeline]{%
+      \includegraphics[clip,width=\columnwidth]{images/dask_bag_inc_baseline_gantt.png}%
+    }
+
+    \subfloat[Dask delayed execution timeline]{%
+      \includegraphics[clip,width=\columnwidth]{images/dask_delayed_inc_baseline_gantt.png}%
+    }
+
+    \subfloat[Dask futures execution timeline]{%
+      \includegraphics[clip,width=\columnwidth]{images/dask_futures_inc_baseline_gantt.png}%
     }
@@
     \caption{8 instances}
-    \label{fig:histo_instance}
+    \label{fig:histo_wroker}
 \end{figure}
 
-\subsection{Experiment 2: Number of splits (task size)}
+\subsection{Experiment 2: Number of chunks}
 
@@
     \caption{10 iterations, 4 sec. sleep delay, 8 instances}
-    \label{fig:histo_splits}
+    \label{fig:histo_chunk}
 \end{figure}
-
-\subsection{Experiment 2: Number of chunks}
 
@@
 \subsection{Experiment 3: Number of instances}
-
-\subsection{Experiment 3: Number of subjects}
 

### 2019-06-12 latex/acl2019.tex
exp1 -- gantt chart

@@
 \subsection{Experimentation 1: Baseline timeline}
-From Figure \ref{fig:inc_gantt} the execution time
-is similar in Dask and Spark. However Spark takes more time to save the data to disk.
-This could be due to the FIFO scheduler that tends to have the first set of save
-consequent thus saturating the nfs. Also serialization could have an effect. It looks
-like the nfs could be a bottleneck when reading (writing) data from (to) the
-application.
+From Table \ref{tb:inc-base} and Figure \ref{fig:inc_gantt} the framework have
+similar makespan however the time spend for each type of task differ significantly.
+Spark tends to spend more time than other framework for IO however it has a much
+lower overhead. dask delayed and futures have a the loowest IO time but a much higher
+overhead. Dask bag stand in the middle for both IO and overhead. Compute time is
+approximately the same for all frameworks.
 
-Dask scheduler seems to have strange behaviour in Figure \ref{fig:inc_gantt}. It
-often wait before scheduling tasks that we would expect to be scheduled right away.
-We currently don't have explication for this behaviour. This could be explained by
-task rescheduled to other workers.
-
-Also, the same length does not seem to be homogeneous. This might be due to caching
-effect either with the main memory or the nfs.
+\begin{table}[h]
+  \resizebox{\columnwidth}{!}{
+    \begin{tabular}{|l|l|l|l|l|l|}
+    \hline
+    Framework    & Read     & Compute & Write    & Overhead & Total    \\ \hline
+    Spark        & 34465.22 & 5117.90 & 22679.15 & 17075.94 & 79338.22 \\ \hline
+    Dask.bag     & 36863.32 & 5120.80 & 15448.91 & 21387.82 & 78820.86 \\ \hline
+    Dask.delayed & 32769.47 & 5119.32 & 11372.86 & 26198.90 & 75460.56 \\ \hline
+    Dask.futures & 31891.29 & 5120.47 & 11269.66 & 27983.72 & 76265.14 \\ \hline
+    \end{tabular}
+  }
+  \caption{Distirbution of the execution time in seconds}
+  \label{tb:inc-base}
+ \end{table}
 

### 2019-06-13 latex/paper.tex
Use the IEEE Tran template and corrected grammar, syntax, and orthograph errors

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{times}
\usepackage{latexsym}
\usepackage{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\usepackage[caption=false]{subfig}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Comparative study of Dask and Spark for neuroimaging pipelines\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Natural Sciences and Engineering Research Council}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Mathieu Dugr\'e}
\IEEEauthorblockA{\textit{Computer Science and Software Engineering} \\
\textit{Concordia University}\\
Montreal, Canada \\
mathieu.dugre@mail.concordia.ca}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{Computer Science and Software Engineering} \\
\textit{Concordia University}\\
Montreal, Canada \\
email address}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{Computer Science and Software Engineering} \\
\textit{Concordia University}\\
Montreal, Canada \\
email address}
}

\maketitle

\begin{abstract}
% Topic
As the amount of data increases and is easier to access, Big Data processing becomes
critical in neuroimaging.
% Problem
This is an issue as the currently used frameworks are specialized for neuroimaging
and not Big Data concerns. This can lead to performance decrease or limit the
research we can perform.
% Relevance
The usage of Big Data framework is beneficial to this problem. The state-of-the-art
general purpose Big Data framework Spark has its code base written in Scala, while
our laboratory mostly uses Python. This can be problematic since it is harder to port
our pipelines to the framework. Also, theoretically, it leads to performance decrease.

% Approach
We propose to use Dask since it provides native Python parallelism to pipelines while
providing support to familiar API from Python's scientific ecosystem. Unfortunately,
there are few comparisons between Spark and Dask; especially in neuroimaging.
Moreover, these studies were done when Dask was still an immature framework which
makes those studies unfair nowadays. This is our motivation to compare the latest
version of Dask with Spark.
% Methods (one sentence)
To evaluate the frameworks, we focus on their performance and their scheduler.

% Key Impact
Our study demonstrates the potential of using Dask as the framework to build
neuroimaging pipelines.
\end{abstract}

\begin{IEEEkeywords}
Big Data, Dask, Spark, performance, neuroimaging
\end{IEEEkeywords}

\section{Introduction}
% Context
With the increasing number of data available in neuroimaging,~\cite{ALFAROALMAGRO:18,
UKBioBank:18} the processing of Big Data becomes critical. Frameworks like
Nipype~\cite{Nipype:11} are usually used to create neuroimaging pipelines however it
is worth considering the use of general purpose Big Data
frameworks~\cite{Hayot-Sasson:17}. In that research, Spark~\cite{Spark:16} was a
natural choice as it is the state-of-the-art framework for Big Data. Since our
laboratory mostly works with Python, we want to know if Dask~\cite{Dask:15}, a Python
written framework, could bring better or similar performances while facilitating
usage.

%Similarities
Spark and Dask offer in-memory computing, data locality, and lazy evaluation; which
is common for Big Data framework. Both their scheduler operates dynamically. This is
good when the runtimes are not known ahead of time~\cite{Dask:15}. Over these
similarities, the frameworks are quite different.

% Spark
On the one hand, we have Spark which provides a high-level API. This allows the
scheduler to perform more optimizations which makes it well suited for neuroimaging
analysis that often requires the usage of a pipeline with multiple steps. Though,
Spark's code base is in Scala which theoretically can lead to slow down in execution
due to a required serialization. Moreover, while Spark's API is flexible and allows
most implementations, it differs from the ones seen in Python's ecosystem.

% Dask
On the other hand, Dask was created with the purpose of natively parallelize Python
pipelines while keeping the syntax of familiar API from Python's scientific
ecosystem. However, Dask is still a young framework with work to be done; it's API
does not completely replicate the library it supports. While its lower-level API
allows the implementation of more complex algorithms it sacrifices a layer of
optimization.
% Related work
Previous work shows that Dask had significant overhead and was
hard to debug~\cite{Mehta:17}.
% What issues your work addresses
Dask was immature at the time and a lot of change was brought to the framework.
Therefore we think it is valuable to re-compare it with Spark.

%% Dask API
The Dask APIs we decide to use for our comparison are Dask bag, delayed and futures.
Dask bag offers an easy API to parallelize data. Dask delayed offer a lower-level API
that offers more flexibility; this is good to implement more complex tasks that do
not fit in the Dask bag framework. Dask futures is a real-time API. Both Dask bag and
delayed apply lazy evaluation to tasks while futures trigger them directly.

% Methods used (summary)
The project aims to compare the state-of-the-art general-purpose framework Spark with
the newcomer Dask. We decide to compare the performance of Spark and Dask on a custom
incrementation pipeline to simplify the effect of the algorithm on the comparison.
Then we assess the frameworks on two real-life applications: (1) histogram of the
voxels intensity in a 3d images (2) BIDS example.

% Implications of our research
The result from our project help in deciding if Dask is a good choice to build
neuroimaging pipelines.

%%%%% MATERIAL AND METHODS %%%%%%
\section{Material and Methods}
%% Material
We compare Dask (\textbf{v1.1.4}) and Spark (\textbf{v2.4.0}) based on two metrics
their scheduling and their makespan performance. All scripts used to perform our
experiments are available at
\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines}.
For our experiments, we used Compute Canada Cloud where each instance has 8 cores,
30GB RAM, 10GB bandwidth and is dedicated.

%% Incrementation
\subsection{Incrementation}
As our first experiment to benchmark the frameworks we used a simple incrementation
pipeline (see Algorithm~\ref{alg:incrementation}). It loads chunks from the 75GB
BigBrain\cite{ds052:01} image, increment its value by 1 a determined number of time
and save the result to an NFS as a Nifti image. Incrementing the data reduce the
caching effect. This experiment allows us to study the behavior of the frameworks
when all input are processed independently; i.e.\ each task in the graph only depends
on the previous one (see~\ref{Figure fig:tg-inc}).

This experiment refers to the implementation of the Spark pipelines purposed
in~\cite{Hayot-Sasson:17}.

\begin{algorithm}[!t]
    \caption{Incrementation}\label{alg:incrementation}
    \begin{algorithmic}
    \Require{\(x\), a sleep delay in float}
    \Require{\(file\), a file containing a chunk}
    \Require{\(fs\), NFS to save image to.}
    \State{read \(chunk\) from \(file\)}
    \ForEach{\(i \in iterations\)}
        \ForEach{\(chunk \in image\)}
            \State{\(chunk\gets chunk+1\)}
            \State{sleep \(x\)}
        \EndFor
    \EndFor
    \State{save \(chunk\) to \(fs\)}
\end{algorithmic}
\end{algorithm}

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.125\textwidth,
    angle=-90]{images/incrementation-task-graph.png}
    \caption{Task graph for Incrementation}\label{fig:tg-inc}
\end{figure}

% Histogram
\subsection{Histogram}
For our second experiment, we find the frequency of the intensity of the voxel in
BigBrain and save them as a tuple in a file. In this experiment, the results depend
on all of the input results (see~\ref{Figure fig:tg-histo}). This allows us to study
the behavior of the framework when tasks are dependent. Also, this application
requires data shuffling thus inter worker communication. We decided to omit the Dask
futures API as it is less realistic for this use case.

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.16\textwidth, angle=-90]{images/histogram-task-graph.png}
    \caption{Task graph for Histogram}\label{fig:tg-histo}
\end{figure}

% BIDS Example
\subsection{BIDS Example}
Our last experiment the BIDS Example algorithm to compare the frameworks.
% Specify more details and the data set used
% SECTION TO COMPLETE -- Experience not done yet

%% Performance
\subsection{Experiments}
Our experiments have for goal to see how the frameworks react to different
environments and task types.

%% Incrementation
%% Node
Firstly, we vary the number of instances in our cluster environment. We test with 1,
2, 4, and 8 instances. From this, we can evaluate the level of parallelism offered by
both frameworks.

%% Iterations
Secondly, we modify the number of iterations we perform on the increment task between
the experiment. We use 1, 10 and 100 iterations. This allows the study of the
schedulers when a different amount of task is to be performed.

%% Sleep delay
Thirdly, we change the sleep delay in the increment tasks. We fix a 1, 4 and 16 and
64 seconds of sleep time for each iteration. This allows us to test the effect of
tasks with different time lengths.

%% Chunks
Finally, we split the BigBrain in a different number of chunks. We divide it in 30,
125 and 750 chunks; around 2.5GB, 0.6GB and 0.1GB, respectively. From that, we can
evaluate the framework with tasks of different size and IO requirements.

%% Baseline
Our baseline for all experiment is 10 iterations, 4 sleep delay, 125 chunks running
on 8 instances.

%% Histogram
For the histogram experiment, we vary the number of instances, as above. We also
change the number of chunks as above.

%% BIDS Example
For the BIDS example experiment, we vary the number of instances, again as above.

% Note
Note that the code implemented might not be optimal. We hope to provide a baseline
assessment of Dask and Spark.

%%%%% RESULTS %%%%%
\section{Results}

%%% INCREMENTATION %%%
%% Instances
\subsection{Experiment 1: Number of instances}
Figure~\ref{fig:inc_worker}(a) shows that the increase in performance is not
proportional to the number of workers which differ from what we would expect; it even
decrease when we reach 8 workers. This is explained by an increase in IO overhead
correlated to the number of workers (Figure~\ref{fig:inc_worker}(b)), as more worker
access the NFS simultaneously it increases the stress put on the system which results
in bottleneck, errors and slow down.

On the one hand, Dask seems better when using a single worker. On the other hand,
Spark makespan decreases faster for a higher level of parallelism. Dask is constantly
faster than Spark which makes us think it has a better scheduler.

\begin{figure}[!t]
    \centering
    \subfloat[Incrementation makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
    }

    \subfloat[Incrementation total time]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
    }
    \caption{125 chunks, 10 iterations, 4 sec.\ sleep delay}\label{fig:inc_worker}
\end{figure}

%% Iterations
\subsection{Experiment 1: Number of iterations}
In Figure~\ref{fig:inc_itr}, Spark and the Dask bag API are equivalent in terms of
performance. The Dask delayed API seems to perform well with a low amount of task but
grow faster the Spark and Dask bag. We think this is because Dask delayed often
schedule a block on a different thread of the same worker which causes small delays
every time due to inter-thread data communication. This is good when there are fewer
tasks as is make the IO between blocks go out of sync hence lowering the IO
bottleneck however as the number of tasks increases those small delays become
significant. The Dask futures API seems to outperform all the other APIs but this is
most likely because the tasks are not interdependent thus this application benefits
from the less optimal but faster scheduling brought by Dask futures.

\begin{figure}[!t]
    \centering
    \subfloat[Incrementation makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
    }

    \subfloat[Incrementation total time]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
    }
    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_itr}
\end{figure}

%% Sleep time
\subsection{Experiment 1: Sleep delay}
As Figure~\ref{fig:inc_sleep}(b) and Table~\ref{tb:inc-sleep} shows Spark has the
smallest overhead all around and it is also the one that reduces the most with longer
task length. In term of overhead, Dask bag trailing slightly behind Spark although is
better than Dask delayed and Dask futures. The opposite happens for the IO time.
% Not sure what to conclude.

%%%% CHANGE NUMBER
\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Distribution of the execution time in seconds for our baseline: 125
    chunks, 4 sec.\ sleep delay, 10 iterations, and 8 workers}\label{tb:inc-sleep}
    \centering
    \begin{tabular}{llllll}
    \hline
                 & Read  & Compute & Write & Overhead & Total \\ \hline
    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
    Dask.bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
    Dask.delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
    Dask.futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
    \end{tabular}
 \end{table}

\begin{figure}[!t]
    \centering
    \subfloat[Incrementation makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
    }

    \subfloat[Incrementation total time]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
    }
    \caption{125 chunks, 10 iterations, 8 instances}
    \label{fig:inc_sleep}
\end{figure}

% Chunks
\subsection{Experiment 1: Number of chunks}
In Figure~\ref{fig:inc_chunk} Spark is not compared for 30 chunks. This is because
Spark has a 2GB limitation in the task size it can compute. Note from
Figure~\ref{fig:inc_chunk}(b) that the compute time increase when there are more
chunks. This is expected because the sleep delay is constant throughout this
experiment and there are more compute tasks when the number of chunks increases. We
also note that in Figure~\ref{fig:inc_chunk}(b) for 30 chunks the Dask bag API has a
much lower overhead time. This is because we only calculate the idle time of the used
core. Dask bag was only using one thread per block in comparison to Dask delayed and
Dask futures which was offloading the calculations on multiple threads of the same
worker. From Figure~\ref{fig:inc_chunk} we observe that when the dataset is separated
in more (smaller) chunks it saves IO time however it increases the overhead by a
similar amount; the inverse happens when we reduce the number of chunks.

\begin{figure}[!t]
    \centering
    \subfloat[Incrementation makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_chunk.png}%
    }

    \subfloat[Incrementation total time]{%
      \includegraphics[clip,width=\columnwidth]{images/inc_idle_chunk.png}%
    }
    \caption{10 iterations, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_chunk}
\end{figure}

%% Baseline gantt chart
\subsection{Experimentation 1: Baseline timeline}
From Table~\ref{tb:inc-base} and Figure~\ref{fig:inc_gantt}, the frameworks have
similar makespan however the time spend for each type of task differs significantly.
Spark tends to spend more time than other frameworks for IO however it has a much
lower overhead. Dask delayed and futures have the lowest IO time but a much higher
overhead. Dask bag stands in the middle for both IO and overhead. Compute time is
approximately the same for all frameworks.

\begin{table}[!t]
    \renewcommand{\arraystretch}{1.3}
    \caption{Distribution of the execution time in seconds for our baseline: 125
    chunks, 4 sec.\ sleep delay, 10 iterations, and 8 workers}\label{tb:inc-base}
    \centering
    \begin{tabular}{llllll}
    \hline
                    & Read  & Compute & Write & Overhead & Total \\ \hline
    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
    Dask.bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
    Dask.delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
    Dask.futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
    \end{tabular}
 \end{table}

\begin{figure}[!t]
    \centering
    \subfloat[Spark execution timeline]{%
      \includegraphics[clip,width=\columnwidth]{images/spark_inc_baseline_gantt.png}%
    }

    \subfloat[Dask bag execution timeline]{%
      \includegraphics[clip,width=\columnwidth]{images/Dask_bag_inc_baseline_gantt.png}%
    }

    \subfloat[Dask delayed execution timeline]{%
      \includegraphics[clip,width=\columnwidth]{images/Dask_delayed_inc_baseline_gantt.png}%
    }

    \subfloat[Dask futures execution timeline]{%
      \includegraphics[clip,width=\columnwidth]{images/Dask_futures_inc_baseline_gantt.png}%
    }
    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}
    \label{fig:inc_gantt}
\end{figure}

%%% HISTOGRAM %%%
\subsection{Experiment 2: Number of instances}

\begin{figure}[!t]
    \centering
    \subfloat[Histogram makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/histo_instance.png}%
    }

    \subfloat[Histogram total time]{%
      \includegraphics[clip,width=\columnwidth]{images/histo_idle_instances.png}%
    }
    \caption{8 instances}\label{fig:histo_wroker}
\end{figure}

\subsection{Experiment 2: Number of chunks}

\begin{figure}[!t]
    \centering
    \subfloat[Histogram makespan]{%
      \includegraphics[clip,width=\columnwidth]{images/histo_splits.png}%
    }

    \subfloat[Histogram total time]{%
      \includegraphics[clip,width=\columnwidth]{images/histo_idle_splits.png}%
    }
    \caption{10 iterations, 4 sec.\ sleep delay, 8 instances}\label{fig:histo_chunk}
\end{figure}

%%% BIDS EXAMPLE %%%
\subsection{Experiment 3: Number of instances}

%%%%% DISCUSSION %%%%%
\section{Discussion}
\subsection{IO Time}

\subsection{Serialization}
We think that serialization could have an important effect on the makespan of the
application. This could potentially slow down the application when a lot of tasks are
scheduled.

\subsection{NFS}
The NFS seems to be a source of the bottleneck. We think that exploring another file
system, like Lustre, could give us insight into the actual effect of the NFS on the
application makespan. We also think that the NFS caching could play an effect on the
task of different lengths.

\subsection{Scheduler}
The Dask scheduler seems to have some bizarre behavior. It seems like it waits to
schedule tasks. It could also be due to tasks being scheduled on other workers
thus requiring data to be sent over the network; which would explain the stall on the
worker. More work would be needed to investigate that issue.

\subsection{Caching}
We think that caching plays an effect on the results. As seen in the Gantt chart
previously, some of the read tasks are significantly shorter than others while all
chunks are of equal size.

\section*{Acknowledgment}

We would like to thanks Compute Canada Cloud for the infrastructure.

\section*{References}
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,reference}

\end{document}

### 2019-06-13 latex/paper.tex
Modified dataset and incrementation sections

@@
 \usepackage{times}
+\usepackage{siunitx}
 \usepackage{latexsym}
@@
 \section{Material and Methods}
-%% Material
-We compare Dask (\textbf{v1.1.4}) and Spark (\textbf{v2.4.0}) based on two metrics
-their scheduling and their makespan performance. All scripts used to perform our
-experiments are available at
-\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines}.
-For our experiments, we used Compute Canada Cloud where each instance has 8 cores,
-30GB RAM, 10GB bandwidth and is dedicated.
+
+\subsection{Engines}
+Spark v2.4.0 and Dask v1.1.4 was used for the experiments.
+
+\textbf{PySpark.RDD \cite{spark-rdd}}
+
+\textbf{Dask.bag \cite{dask-bag}}
+
+\textbf{Dask.delayed \cite{dask-delayed}}
+
+\textbf{Dask.futures \cite{dask-futures}}
+
+\subsection{Dataset}
+For the incrementation and histogram experiment the BigBrain~\cite{Amunts:13} is
+used. It is a three-dimensional image of a brain scanned at \SI{40}{\micro\metre}.
+The data set store the intensity of each voxel and their position. This results in a
+75GB image. It is stored in 125 chunks in the minc~\cite{minc} format. To perform our
+experiment we convert the chunks into the nifti format. Also, using sam~\cite{sam} we
+combine the chunks and split them into 30, 125 and 750 chunks of equal size; 2.5GB,
+0.6GB, and 0.1GB, respectively.
+
+\subsection{Applications}
 
 %% Incrementation
-\subsection{Incrementation}
+\subsubsection{\textbf{Incrementation}}
 As our first experiment to benchmark the frameworks we used a simple incrementation
-pipeline (see Algorithm~\ref{alg:incrementation}). It loads chunks from the 75GB
-BigBrain\cite{ds052:01} image, increment its value by 1 a determined number of time
-and save the result to an NFS as a Nifti image. Incrementing the data reduce the
-caching effect. This experiment allows us to study the behavior of the frameworks
-when all input are processed independently; i.e.\ each task in the graph only depends
-on the previous one (see~\ref{Figure fig:tg-inc}).
+pipeline (see Algorithm~\ref{alg:incrementation}). It reads chunks from the BigBrain
+image, increments the intensity value of each voxel by 1 for a specified amount of
+iterations and write the result to an NFS as a Nifti image. Incrementing the data
+reduce the caching effect. This experiment allows us to study the behavior of the
+frameworks when all input are processed independently; i.e.\ each task in the graph
+only depends on the previous one (see~\ref{Figure fig:tg-inc}).
 
@@
 \end{figure}
+
+\subsubsection{Histogram}
+
+\subsection{Experiment}
+
+\subsection{Infrastructure}
+All experiments were performed on Compute Canda Cloud.
+
+%% Material
+We compare Dask (\textbf{v1.1.4}) and Spark (\textbf{v2.4.0}) based on two metrics
+their scheduling and their makespan performance. All scripts used to perform our
+experiments are available at
+\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines}.
+For our experiments, we used Compute Canada Cloud where each instance has 8 cores,
+30GB RAM, 10GB bandwidth and is dedicated.
 

### 2019-06-13 latex/paper.tex
Modified the spark and dask bag engine sub-section

@@
 
-\textbf{PySpark.RDD \cite{spark-rdd}}
+\subsubsection{PySpark RDD~\cite{spark-rdd}} Spark is a general-purpose big data
+framework. It provides in-memory computing, data locality and lazy evaluation. Spark
+abstract resilient distributed dataset (RDD)~\cite{RDD} making it fault-tolerant. Its
+high-level API provides optimizations on the transformations made to some data. Since
+Spark is written in Scala, this optimization is lessen because it requires a
+serialization from Python to Java. The API offered by Spark is flexible and suit most
+problem however it differ from Python ecosystem. The standalone scheduler is used for
+our experiment. It has a LIFO policy.
 
-\textbf{Dask.bag \cite{dask-bag}}
+\subsubsection{Dask Bag~\cite{dask-bag}} Dask Bag is one of the high-level API
+offered by dask. It also provides in-memory computing, data locality and lazy
+evaluation however it does not have fault resilience. Fundamently it is parallized
+lists. Dask Bag parallelizes computations of large collection of generic python
+objects~\cite{dask-bag}. Also, it offers optimized algorithm that mimic most the
+toolz library. Our experiment use the dask distributed~\cite{dask-distributed}
+scheduler. It has a LIFO policy. In theory it should speed-up the computation by
+finishing task computations before starting new ones.
 
-\textbf{Dask.delayed \cite{dask-delayed}}
+\subsubsection{Dask Delayed~\cite{dask-delayed}} Dask Delayed is a low-level API. It
+does not offer the optimized algorithm as the high-level APIs however it allows the
+user to parallelize custom function that do not fit in the array, dataframe or
+MapReduce framework.
 
-\textbf{Dask.futures \cite{dask-futures}}
+\subsubsection{Dask Futures~\cite{dask-futures}} Dask Futures is also a low-level API
+from dask. It differs from the other APIs as its computations trigger immediately.
 
@@
 frameworks when all input are processed independently; i.e.\ each task in the graph
-only depends on the previous one (see~\ref{Figure fig:tg-inc}).
+only depends on the previous one (see Figure~\ref{Figure fig:tg-inc}). This mimic the
+behavior of analyzing multiple independent subjects in parallel.
 
 This experiment refers to the implementation of the Spark pipelines purposed
-in~\cite{Hayot-Sasson:17}.
+in Ref.~\cite{Hayot-Sasson:17}.
 

### 2019-06-14 latex/paper.tex
Modified the engines section

@@
 \subsubsection{Dask Bag~\cite{dask-bag}} Dask Bag is one of the high-level API
-offered by dask. It also provides in-memory computing, data locality and lazy
+offered by dask. Like Spark, it provides in-memory computing, data locality and lazy
 evaluation however it does not have fault resilience. Fundamently it is parallized
 lists. Dask Bag parallelizes computations of large collection of generic python
-objects~\cite{dask-bag}. Also, it offers optimized algorithm that mimic most the
-toolz library. Our experiment use the dask distributed~\cite{dask-distributed}
-scheduler. It has a LIFO policy. In theory it should speed-up the computation by
-finishing task computations before starting new ones.
+objects~\cite{dask-bag}. Also, it offers optimized algorithm that mimic most of the
+toolz library. Our experiment use the Dask Distributed~\cite{dask-distributed}
+scheduler. It has a LIFO policy. In theory it should speed-up the pipeline excution
+by finishing task computations before starting new ones.
 
@@
 user to parallelize custom function that do not fit in the array, dataframe or
-MapReduce framework.
+MapReduce framework. This is perfect for more complex function which could easily be
+parallelized. Like Dask Bag, it offers in-memory computing, data locality and lack
+fault resilience. The Dask Distributed scheduler.
 
 \subsubsection{Dask Futures~\cite{dask-futures}} Dask Futures is also a low-level API
-from dask. It differs from the other APIs as its computations trigger immediately.
+from Dask. Like the other Dask APIs it benifit of in-memory computing and data
+locatlity. Unlike the other APIs, Dask Futures triggers computation immediately. When
+the user tries to gather data, if available it will return the result otherwise it
+will block until the cimputation is done.
 

### 2019-06-14 histogram/dask-bag_histogram.py
Modified to use a uniform algorithm with numpy

@@ -1,15 +1,12 @@
-import sys
-
-sys.path.append("/nfs/paper-big-data-engines")
-
 import argparse
 import os
+import sys
 from time import time
 
 import dask.bag as db
 from dask.distributed import Client
 
-from utils import benchmark, crawl_dir, read_img
+sys.path.append("/nfs/paper-big-data-engines")
 
 
 if __name__ == "__main__":
@@ -42,34 +39,29 @@ if __name__ == "__main__":
     print(client)
     # Allow workers to use module
     client.upload_file("/nfs/paper-big-data-engines/utils.py")
+    client.upload_file("/nfs/paper-big-data-engines/histogram/Histogram.py")
+    from utils import benchmark, crawl_dir, read_img
+    from histogram import calculate_histogram, combine_histogram
+
 
     # Read images
     paths = crawl_dir(os.path.abspath(args.bb_dir))
     paths = db.from_sequence(paths, npartitions=len(paths))
-    img_rdd = paths.map(lambda p: read_img(p, start=start, args=args))
+    img = paths.map(lambda p: read_img(p, start=start, args=args))
 
-    start_time = time() - start
+    partial_histogram = img.map(
+        lambda x: calculate_histogram(x[1], args=args, start=start, filename=x[0])
+    )
 
-    voxels = img_rdd.map(lambda x: x[1].flatten("F")).flatten()
-    frequency_pair = voxels.frequencies().compute()
-
-    end_time = time() - start
-
-    if args.benchmark:
-        benchmark(
-            start_time,
-            end_time,
-            "all_file",
-            args.output_dir,
-            args.experiment,
-            "find_frequency",
-        )
+    histogram = partial_histogram.fold(
+        lambda x: combine_histogram(x, args=args, start=start)
+    ).compute()
 
     start_time = time() - start
 
     with open(f"{args.output_dir}/histogram.csv", "w") as f_out:
-        for x in frequency_pair:
-            f_out.write(f"{x[0]};{x[1]}\n")
+        for k, v in histogram.items():
+            f_out.write(f"{k};{v}\n")
 
     end_time = time() - start
 

### 2019-06-14 histogram/dask-delayed_histogram.py

@@ -1,17 +1,13 @@
-import sys
-
-sys.path.append("/nfs/paper-big-data-engines")
-
 import argparse
 from functools import reduce
 import os
+import sys
 from time import time
 
 import dask
 from dask.distributed import Client
-import numpy as np
 
-from utils import benchmark, crawl_dir, read_img
+sys.path.append("/nfs/paper-big-data-engines")
 
 
 if __name__ == "__main__":
@@ -44,59 +40,25 @@ if __name__ == "__main__":
     print(client)
     # Allow workers to use module
     client.upload_file("/nfs/paper-big-data-engines/utils.py")
-
-    def calculate_histogram(arr):
-        start_time = time() - start
-
-        y = np.bincount(arr)
-        ii = np.nonzero(y)[0]
-        out = np.vstack((ii, y[ii])).T
-        histogram = {k: v for k, v in out}
-
-        end_time = time() - start
-
-        if args.benchmark:
-            benchmark(
-                start_time,
-                end_time,
-                "all_file",
-                args.output_dir,
-                args.experiment,
-                "calculate_histogram",
-            )
-        return histogram
-
-    def combine_histogram(x, y):
-        start_time = time() - start
-
-        rv = {k: x.get(k, 0) + y.get(k, 0) for k in set(x) | set(y)}
-
-        end_time = time() - start
-
-        if args.benchmark:
-            benchmark(
-                start_time,
-                end_time,
-                "all_file",
-                args.output_dir,
-                args.experiment,
-                "combine_histogram",
-            )
-        return rv
+    client.upload_file("/nfs/paper-big-data-engines/histogram/Histogram.py")
+    from utils import benchmark, crawl_dir, read_img
+    from histogram import calculate_histogram, combine_histogram
 
     # Read images
     paths = crawl_dir(os.path.abspath(args.bb_dir))
 
     partial_histogram = []
     for path in paths:
-        img_rdd = dask.delayed(read_img)(path, start=start, args=args)
+        img = dask.delayed(read_img)(path, start=start, args=args)
 
-        voxels = img_rdd[1].flatten("F")
-
-        partial_histogram.append(dask.delayed(calculate_histogram)(voxels))
+        partial_histogram.append(
+            dask.delayed(calculate_histogram)(
+                img[1], args=args, start=start, filename=img[0]
+            )
+        )
 
     histogram = dask.delayed(reduce)(
-        lambda x, y: combine_histogram(x, y), partial_histogram
[…]

### 2019-06-14 histogram/spark_histogram.py
Modified to use a uniform algorithm with numpy

@@ -1,14 +1,11 @@
-import sys
-
-sys.path.append("/nfs/paper-big-data-engines")
-
 import argparse
 import os
+import sys
 from time import time
 
 from pyspark import SparkConf, SparkContext
 
-from utils import benchmark, crawl_dir, read_img
+sys.path.append("/nfs/paper-big-data-engines")
 
 
 if __name__ == "__main__":
@@ -34,10 +31,14 @@ if __name__ == "__main__":
     args = parser.parse_args()
 
     # Cluster scheduler
-    conf = SparkConf().setAppName("Spark Incrementation")
+    conf = SparkConf().setAppName(f"Spark {args.experiment}")
     sc = SparkContext.getOrCreate(conf=conf)
 
     sc.addFile("/nfs/paper-big-data-engines/utils.py")
+    sc.addFile("/nfs/paper-big-data-engines/histogram/Histogram.py")
+    from utils import benchmark, crawl_dir, read_img
+    from histogram import calculate_histogram, combine_histogram
+
     print("Connected")
 
     # Read images
@@ -45,30 +46,21 @@ if __name__ == "__main__":
     paths = sc.parallelize(paths, len(paths))
     img_rdd = paths.map(lambda p: read_img(p, start=start, args=args))
 
-    start_time = time() - start
+    voxels = img_rdd.map(lambda x: (x[0], x[1].flatten("F")))
 
-    voxels = img_rdd.flatMap(lambda x: x[1].flatten("F"))
-    frequency_pair = (
-        voxels.map(lambda x: (x, 1)).reduceByKey(lambda x, y: x + y).collect()
+    partial_histogram = voxels.map(
+        lambda x: calculate_histogram(x[1], args=args, start=start, filename=x[0])
     )
 
-    end_time = time() - start
-
-    if args.benchmark:
-        benchmark(
-            start_time,
-            end_time,
-            "all_file",
-            args.output_dir,
-            args.experiment,
-            "find_frequency",
-        )
+    histogram = partial_histogram.fold(
+        lambda x: combine_histogram(x, args=args, start=start)
+    ).collect()
 
     start_time = time() - start
 
     with open(f"{args.output_dir}/histogram.csv", "w") as f_out:
-        for x in frequency_pair:
-            f_out.write(f"{x[0]};{x[1]}\n")
+        for k, v in histogram.items():
+            f_out.write(f"{k};{v}\n")
 
     end_time = time() - start
 

### 2019-06-14 experiments-script/run_experiment.py
Modified for histogram with numpy

@@ -3,14 +3,16 @@ from random import shuffle
 import json
 
 
-with open("experiment.json") as f_in:
+with open(
+    "/nfs/paper-big-data-engines/experiments-script/histogram_experiment.json"
+) as f_in:
     experiments = json.load(f_in)
     shuffle(experiments)
     for exp in experiments:
         experiment = exp["experiment"]
         filename = exp["file"]
-        iterations = str(exp["iterations"])
-        delay = str(exp["delay"])
+        # iterations = str(exp["iterations"])
+        # delay = str(exp["delay"])
         chunks = str(exp["chunks"])
 
         subprocess.run(
@@ -29,8 +31,6 @@ with open("experiment.json") as f_in:
                     "/nfs/bb-" + chunks + "chunks",
                     "/nfs/results",
                     experiment,
-                    iterations,
-                    delay,
                     "--benchmark",
                 ]
             )
@@ -43,8 +43,6 @@ with open("experiment.json") as f_in:
                     "/nfs/bb-" + chunks + "chunks",
                     "/nfs/results",
                     experiment,
-                    iterations,
-                    delay,
                     "--benchmark",
                 ]
             )

### 2019-06-14 histogram/Histogram.py
Move function into one file for uniformity

@@ -1,0 +1,48 @@
+from time import time
+
+import numpy as np
+
+from utils import benchmark
+
+
+def calculate_histogram(arr, *, args, start, filename):
+    start_time = time() - start
+
+    arr = arr.flatten('F')
+
+    y = np.bincount(arr)
+    ii = np.nonzero(y)[0]
+    out = np.vstack((ii, y[ii])).T
+    histogram = {k: v for k, v in out}
+
+    end_time = time() - start
+
+    if args.benchmark:
+        benchmark(
+            start_time,
+            end_time,
+            filename,
+            args.output_dir,
+            args.experiment,
+            "calculate_histogram",
+        )
+    return histogram
+
+
+def combine_histogram(x, y, *, args, start):
+    start_time = time() - start
+
+    rv = {k: x.get(k, 0) + y.get(k, 0) for k in set(x) | set(y)}
+
+    end_time = time() - start
+
+    if args.benchmark:
+        benchmark(
+            start_time,
+            end_time,
+            "all_file",
+            args.output_dir,
+            args.experiment,
+            "combine_histogram",
+        )
+    return rv

### 2019-06-17 histogram/dask-bag_histogram.py
modfied import

@@ -41,8 +41,7 @@ if __name__ == "__main__":
     client.upload_file("/nfs/paper-big-data-engines/utils.py")
     client.upload_file("/nfs/paper-big-data-engines/histogram/Histogram.py")
     from utils import benchmark, crawl_dir, read_img
-    from histogram import calculate_histogram, combine_histogram
-
+    from histogram.Histogram import calculate_histogram, combine_histogram
 
     # Read images
     paths = crawl_dir(os.path.abspath(args.bb_dir))

### 2019-06-17 histogram/dask-delayed_histogram.py
modfied import

@@ -42,7 +42,7 @@ if __name__ == "__main__":
     client.upload_file("/nfs/paper-big-data-engines/utils.py")
     client.upload_file("/nfs/paper-big-data-engines/histogram/Histogram.py")
     from utils import benchmark, crawl_dir, read_img
-    from histogram import calculate_histogram, combine_histogram
+    from histogram.Histogram import calculate_histogram, combine_histogram
 
     # Read images
     paths = crawl_dir(os.path.abspath(args.bb_dir))

### 2019-06-17 histogram/spark_histogram.py

@@ -37,7 +37,7 @@ if __name__ == "__main__":
     sc.addFile("/nfs/paper-big-data-engines/utils.py")
     sc.addFile("/nfs/paper-big-data-engines/histogram/Histogram.py")
     from utils import benchmark, crawl_dir, read_img
-    from histogram import calculate_histogram, combine_histogram
+    from histogram.Histogram import calculate_histogram, combine_histogram
 
     print("Connected")
 

### 2019-06-17 histogram/dask-bag_histogram.py
Modified path for server

@@ -6,7 +6,7 @@ from time import time
 import dask.bag as db
 from dask.distributed import Client
 
-sys.path.append("/nfs/paper-big-data-engines")
+sys.path.append("/nfs/paper-big-data-engines/histogram")
 
 
 if __name__ == "__main__":
@@ -41,7 +41,7 @@ if __name__ == "__main__":
     client.upload_file("/nfs/paper-big-data-engines/utils.py")
     client.upload_file("/nfs/paper-big-data-engines/histogram/Histogram.py")
     from utils import benchmark, crawl_dir, read_img
-    from histogram.Histogram import calculate_histogram, combine_histogram
+    from Histogram import calculate_histogram, combine_histogram
 
     # Read images
     paths = crawl_dir(os.path.abspath(args.bb_dir))
@@ -53,7 +53,7 @@ if __name__ == "__main__":
     )
 
     histogram = partial_histogram.fold(
-        lambda x: combine_histogram(x, args=args, start=start)
+        lambda x, y: combine_histogram(x, y, args=args, start=start)
     ).compute()
 
     start_time = time() - start

### 2019-06-17 histogram/dask-delayed_histogram.py

@@ -7,7 +7,7 @@ from time import time
 import dask
 from dask.distributed import Client
 
-sys.path.append("/nfs/paper-big-data-engines")
+sys.path.append("/nfs/papeSr-big-data-engines")
 
 
 if __name__ == "__main__":
@@ -42,7 +42,7 @@ if __name__ == "__main__":
     client.upload_file("/nfs/paper-big-data-engines/utils.py")
     client.upload_file("/nfs/paper-big-data-engines/histogram/Histogram.py")
     from utils import benchmark, crawl_dir, read_img
-    from histogram.Histogram import calculate_histogram, combine_histogram
+    from Histogram import calculate_histogram, combine_histogram
 
     # Read images
     paths = crawl_dir(os.path.abspath(args.bb_dir))

### 2019-06-17 histogram/spark_histogram.py

@@ -5,7 +5,7 @@ from time import time
 
 from pyspark import SparkConf, SparkContext
 
-sys.path.append("/nfs/paper-big-data-engines")
+sys.path.append("/nfs/paper-big-data-engines/histogram")
 
 
 if __name__ == "__main__":
@@ -37,7 +37,7 @@ if __name__ == "__main__":
     sc.addFile("/nfs/paper-big-data-engines/utils.py")
     sc.addFile("/nfs/paper-big-data-engines/histogram/Histogram.py")
     from utils import benchmark, crawl_dir, read_img
-    from histogram.Histogram import calculate_histogram, combine_histogram
+    from Histogram import calculate_histogram, combine_histogram
 
     print("Connected")
 

### 2019-06-17 histogram/dask-delayed_histogram.py
modified path for server

@@ -7,7 +7,7 @@ from time import time
 import dask
 from dask.distributed import Client
 
-sys.path.append("/nfs/papeSr-big-data-engines")
+sys.path.append("/nfs/paper-big-data-engines/histogram")
 
 
 if __name__ == "__main__":

### 2019-06-17 histogram/spark_histogram.py
Fix spark reduction

@@ -53,7 +53,7 @@ if __name__ == "__main__":
     )
 
     histogram = partial_histogram.fold(
-        lambda x: combine_histogram(x, args=args, start=start)
+        {}, lambda x, y: combine_histogram(x, y, args=args, start=start)
     ).collect()
 
     start_time = time() - start

### 2019-06-17 latex/paper.tex
Modified histogram application

@@
 \subsection{Engines}
-Spark v2.4.0 and Dask v1.1.4 was used for the experiments.
+For all our experiments, we used Spark v2.4.0 and Dask v1.1.4.
 
-\subsubsection{PySpark RDD~\cite{spark-rdd}} Spark is a general-purpose big data
+\subsubsection{PySpark RDD~\cite{spark-rdd}} Spark is a general-purpose Big Data
 framework. It provides in-memory computing, data locality and lazy evaluation. Spark
-abstract resilient distributed dataset (RDD)~\cite{RDD} making it fault-tolerant. Its
-high-level API provides optimizations on the transformations made to some data. Since
+abstracts resilient distributed dataset (RDD)~\cite{RDD} making it fault-tolerant. Its
+high-level API provides optimizations for transformations applied data. Since
 Spark is written in Scala, this optimization is lessen because it requires a
-serialization from Python to Java. The API offered by Spark is flexible and suit most
-problem however it differ from Python ecosystem. The standalone scheduler is used for
+serialization from Python to Java. The API offered by Spark is flexible and suits most
+problems however it differs from Python ecosystem. The standalone scheduler is used for
 our experiment. It has a LIFO policy.
@@
 \subsubsection{Dask Bag~\cite{dask-bag}} Dask Bag is one of the high-level API
-offered by dask. Like Spark, it provides in-memory computing, data locality and lazy
-evaluation however it does not have fault resilience. Fundamently it is parallized
+offered by Dask. Like Spark, it provides in-memory computing, data locality and lazy
+evaluation however it does not have fault resilience. Fundamently, it is parallized
 lists. Dask Bag parallelizes computations of large collection of generic python
-objects~\cite{dask-bag}. Also, it offers optimized algorithm that mimic most of the
-toolz library. Our experiment use the Dask Distributed~\cite{dask-distributed}
+objects~\cite{dask-bag}. Also, it offers optimized algorithms that mimic most of the
+toolz~\cite{toolz} library. Our experiment use the Dask Distributed~\cite{dask-distributed}
 scheduler. It has a LIFO policy. In theory it should speed-up the pipeline excution
-by finishing task computations before starting new ones.
+by finishing a branch of task computations before starting new ones.
 
@@
 parallelized. Like Dask Bag, it offers in-memory computing, data locality and lack
-fault resilience. The Dask Distributed scheduler.
+fault resilience. The Dask Distributed scheduler is used.
 
 \subsubsection{Dask Futures~\cite{dask-futures}} Dask Futures is also a low-level API
-from Dask. Like the other Dask APIs it benifit of in-memory computing and data
-locatlity. Unlike the other APIs, Dask Futures triggers computation immediately. When
-the user tries to gather data, if available it will return the result otherwise it
-will block until the cimputation is done.
+from Dask. Like the other Dask APIs, it benifits of in-memory computing and data
+locatlity. However, unlike the other APIs, Dask Futures triggers computation
+immediately. When the user tries to gather data if available it will return the
+result otherwise it will block until the computation is done. Again, the Dask
+Distributed scheduler is used.
 
@@
 The data set store the intensity of each voxel and their position. This results in a
-75GB image. It is stored in 125 chunks in the minc~\cite{minc} format. To perform our
-experiment we convert the chunks into the nifti format. Also, using sam~\cite{sam} we
-combine the chunks and split them into 30, 125 and 750 chunks of equal size; 2.5GB,
-0.6GB, and 0.1GB, respectively.
+75GB image. It is stored in 125 chunks in the \textit{MINC}~\cite{minc} format. To
+perform our experiment we convert the chunks into the \textit{NifTI}~\cite{nifti}
+format. Also, using \textit{sam}~\cite{sam} we combine the chunks and split them into
+30, 125 and 750 chunks of equal size; each of 2.5GB, 0.6GB, and 0.1GB, respectively.
 
 \subsection{Applications}
-
 %% Incrementation
 \subsubsection{\textbf{Incrementation}}
-As our first experiment to benchmark the frameworks we used a simple incrementation
+As our first experiment to benchmark the frameworks, we used a simple incrementation
 pipeline (see Algorithm~\ref{alg:incrementation}). It reads chunks from the BigBrain
 image, increments the intensity value of each voxel by 1 for a specified amount of
-iterations and write the result to an NFS as a Nifti image. Incrementing the data
+iterations, and write the result to an NFS as a NifTI image. Incrementing the data
 reduce the caching effect. This experiment allows us to study the behavior of the
@@
 
-\begin{algorithm}[!t]
+\begin{algorithm}[!b]
     \caption{Incrementation}\label{alg:incrementation}
@@
     \Require{\(file\), a file containing a chunk}
-    \Require{\(fs\), NFS to save image to.}
-    \State{read \(chunk\) from \(file\)}
+    \Require{\(fs\), NFS to write image to.}
+    \State{Read \(chunk\) from \(file\)}
     \ForEach{\(i \in iterations\)}
@@
             \State{\(chunk\gets chunk+1\)}
-            \State{sleep \(x\)}
+            \State{Sleep \(x\)}
         \EndFor
     \EndFor
-    \State{save \(chunk\) to \(fs\)}
+    \State{Write \(chunk\) to \(fs\)}
 \end{algorithmic}
@@
 
-\begin{figure}[!t]
+\begin{figure}[!b]
     \centering
@@
 
-\subsubsection{Histogram}
+\subsubsection{\textbf{Histogram}}
+Our second experiment consist to calculate the histogram of the intensity of the
+voxel in BigBrain (see Algorithm~\ref{alg:histogram}). It reads the images from the
+nfs, calculate inividual image frequency, aggregate the images frequency together,
+and write the result to a file. In this experiment, the result depends on all the
+input (see Fig.~\ref{fig:tg-histo}). This allows to study the behavior of the
+frameworks when tasks are dependent. Also, this application requires data shuffling
+thus inter-worker communication. We decided to omit Dask Futures as it does not had
+any value over Dask Delayed in this application.
+
+\begin{algorithm}[!t]
+    \caption{Histogram}\label{alg:histogram}
+    \begin{algorithmic}
+    \Require{\(files\), files containing BigBrain}
+    \Require{\(fs\), NFS to save image to.}
+    \ForEach{\(file \in files\)}
+        \State{Read \(chunk\) from \(file\)}
+        \State{{Calculate \(frequency\) of \(chunks\)}}
+    \EndFor
+
+    \State{\(histogram\gets\)Aggregate \(frequency\) of each \(file\)}
+
+    \State{Write \(histogram\) to \(fs\)}
+    \end{algorithmic}
+\end{algorithm}
+
+\begin{figure}[!t]
+    \centering
+    \includegraphics[width=0.16\textwidth, angle=-90]{images/histogram-task-graph.png}
+    \caption{Task graph for Histogram}\label{fig:tg-histo}
+\end{figure}
 
@@
 \href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines}.
-For our experiments, we used Compute Canada Cloud where each instance has 8 cores,
+For our experiments, we used Compute Canada Cloud where each instance has 8 vCPUs,
 30GB RAM, 10GB bandwidth and is dedicated.
-
-% Histogram
-\subsection{Histogram}
-For our second experiment, we find the frequency of the intensity of the voxel in
-BigBrain and save them as a tuple in a file. In this experiment, the results depend
-on all of the input results (see~\ref{Figure fig:tg-histo}). This allows us to study
-the behavior of the framework when tasks are dependent. Also, this application
-requires data shuffling thus inter worker communication. We decided to omit the Dask
-futures API as it is less realistic for this use case.
-
-\begin{figure}[!t]
-    \centering
-    \includegraphics[width=0.16\textwidth, angle=-90]{images/histogram-task-graph.png}
-    \caption{Task graph for Histogram}\label{fig:tg-histo}
-\end{figure}
-
-% BIDS Example
-\subsection{BIDS Example}
-Our last experiment the BIDS Example algorithm to compare the frameworks.
-% Specify more details and the data set used
-% SECTION TO COMPLETE -- Experience not done yet
 

### 2019-06-19 latex/paper.tex
Changed figfloat to subfigure. Modified the experiment section

@@
 \usepackage{textcomp}
+\usepackage{enumitem}
 \usepackage{xcolor}
@@
 \usepackage{algorithm}
+\usepackage{caption}
+\usepackage{subcaption}
 \algnewcommand\algorithmicforeach{\textbf{for each}}
 \algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
-\usepackage[caption=false]{subfig}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
 \begin{document}
@@
 frameworks when all input are processed independently; i.e.\ each task in the graph
-only depends on the previous one (see Figure~\ref{Figure fig:tg-inc}). This mimic the
+only depends on the previous one (see Figure~\ref{fig:tg-inc}). This mimic the
 behavior of analyzing multiple independent subjects in parallel.
@@
 \subsection{Experiment}
+The applications are benchmarked according to up to four parameters
+(Table~\ref{tab:param}):
+
+\textbf{Number of workers:} Varying the numbers of workers helps in assessing the
+scaling of the frameworks in a cluster. Also, it helps in determining the level of
+parallelism of the frameworks.
+
+\textbf{Number of chunks:} Evaluating the frameworks on different amount of chunks
+allows us to understand their behavior with different task size and IO requirements.
+
+\textbf{Number of iterations:} We change the amount of iterations for the
+incrementation~\ref{alg:incrementation} to study the effect of the job size.
+
+\textbf{Sleep delay:} We modify the sleep delay in the
+incrementation~\ref{alg:incrementation} to understand the behavior of task of
+different length.
+
+The baseline for our experiments is: 8 workers, 125 chunks, 10 iterations and 4
+seconds sleep delay.
+
+\begin{table}[!b]
+    \renewcommand{\arraystretch}{1.3}
+    \caption{Parameters for the experiments}\label{tab:param}
+    \centering
+    \begin{tabular*}{\columnwidth}{llll}
+    \hline
+                        & Incrementation & Histogram             & BIDS Example          \\ \hline
+    \# of worker        & 1, 2, 4, 8     & 1, 2, 4, 8            & 1, 2, 4, 8            \\
+    \# of chunks        & 30, 125, 750   & 30, 125, 750          & 30, 125, 750          \\
+    \# of iterations    & 1, 10, 100     & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\
+    Sleep delay {[}s{]} & 1, 4, 16, 64   & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\ \hline
+    \end{tabular*}
+    \end{table}
 
 \subsection{Infrastructure}
-All experiments were performed on Compute Canda Cloud.
-
-%% Material
-We compare Dask (\textbf{v1.1.4}) and Spark (\textbf{v2.4.0}) based on two metrics
-their scheduling and their makespan performance. All scripts used to perform our
-experiments are available at
-\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines}.
-For our experiments, we used Compute Canada Cloud where each instance has 8 vCPUs,
-30GB RAM, 10GB bandwidth and is dedicated.
-
-%% Performance
-\subsection{Experiments}
-Our experiments have for goal to see how the frameworks react to different
-environments and task types.
-
-%% Incrementation
-%% Node
-Firstly, we vary the number of instances in our cluster environment. We test with 1,
-2, 4, and 8 instances. From this, we can evaluate the level of parallelism offered by
-both frameworks.
-
-%% Iterations
-Secondly, we modify the number of iterations we perform on the increment task between
-the experiment. We use 1, 10 and 100 iterations. This allows the study of the
-schedulers when a different amount of task is to be performed.
-
-%% Sleep delay
-Thirdly, we change the sleep delay in the increment tasks. We fix a 1, 4 and 16 and
-64 seconds of sleep time for each iteration. This allows us to test the effect of
-tasks with different time lengths.
-
-%% Chunks
-Finally, we split the BigBrain in a different number of chunks. We divide it in 30,
-125 and 750 chunks; around 2.5GB, 0.6GB and 0.1GB, respectively. From that, we can
-evaluate the framework with tasks of different size and IO requirements.
-
-%% Baseline
-Our baseline for all experiment is 10 iterations, 4 sleep delay, 125 chunks running
-on 8 instances.
-
-%% Histogram
-For the histogram experiment, we vary the number of instances, as above. We also
-change the number of chunks as above.
-
-%% BIDS Example
-For the BIDS example experiment, we vary the number of instances, again as above.
-
-% Note
-Note that the code implemented might not be optimal. We hope to provide a baseline
-assessment of Dask and Spark.
+All experiments were performed on Arbutus Cloud Compute Canda. Each node has CentOS
+7.5.1804 as base operating system and have the Linux kernel version
+3.10.0\-862.11.6.el7.x86\_64. The nodes have an Intel Xeon Gold 6130 processor with
+30GB 2666 MHz memory. The networks has a 10GB/s bandwidth. The filesystem is 2TB NFS v4.
 
@@
 %%% INCREMENTATION %%%
-%% Instances
+%% worker
 \subsection{Experiment 1: Number of instances}
@@
     \centering
-    \subfloat[Incrementation makespan]{%
+    \begin{subfigure}[b]{\columnwidth}
         \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
-    }
+        \caption{Incrementation makespan}\label{fig:inc_ms_worker}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
+        \caption{Incrementation total time}\label{fig:inc_tt_worker}
+    \end{subfigure}
+    \caption{125 chunks, 10 iterations, 4 sec.\ sleep delay}\label{fig:inc_worker}
+\end{figure}
 
-    \subfloat[Incrementation total time]{%
-      \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
-    }
-    \caption{125 chunks, 10 iterations, 4 sec.\ sleep delay}\label{fig:inc_worker}
+% Chunks
+\subsection{Experiment 1: Number of chunks}
+In Figure~\ref{fig:inc_chunk} Spark is not compared for 30 chunks. This is because
+Spark has a 2GB limitation in the task size it can compute. Note from
+Figure~\ref{fig:inc_chunk}(b) that the compute time increase when there are more
+chunks. This is expected because the sleep delay is constant throughout this
+experiment and there are more compute tasks when the number of chunks increases. We
+also note that in Figure~\ref{fig:inc_tt_chunk} for 30 chunks the Dask bag API has a
+much lower overhead time. This is because we only calculate the idle time of the used
+core. Dask bag was only using one thread per block in comparison to Dask delayed and
+Dask futures which was offloading the calculations on multiple threads of the same
+worker. From Figure~\ref{fig:inc_chunk} we observe that when the dataset is separated
+in more (smaller) chunks it saves IO time however it increases the overhead by a
+similar amount; the inverse happens when we reduce the number of chunks.
+
+\begin{figure}[!t]
+    \centering
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/inc_chunk.png}%
+        \caption{Incrementation makespan}\label{fig:inc_ms_chunk}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/inc_idle_chunk.png}%
+        \caption{Incrementation total time}\label{fig:inc_tt_chunk}
+    \end{subfigure}
+    \caption{10 iterations, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_chunk}
 \end{figure}
@@
     \centering
-    \subfloat[Incrementation makespan]{%
+    \begin{subfigure}[b]{\columnwidth}
         \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
-    }
-
-    \subfloat[Incrementation total time]{%
+        \caption{Incrementation makespan}\label{fig:inc_ms_chunk}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
         \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
-    }
+        \caption{Incrementation total time}\label{fig:inc_tt_chunk}
+    \end{subfigure}
     \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_itr}
@@
     \centering
-    \begin{tabular}{llllll}
+    \begin{tabular*}{\columnwidth}{llllll}
     \hline
@@
     Dask.futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
-    \end{tabular}
+    \end{tabular*}
  \end{table}
@@
     \centering
-    \subfloat[Incrementation makespan]{%
+    \begin{subfigure}[b]{\columnwidth}
         \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
-    }
-
-    \subfloat[Incrementation total time]{%
+        \caption{Incrementation makespan}\label{fig:inc_ms_chunk}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
         \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
-    }
-    \caption{125 chunks, 10 iterations, 8 instances}
-    \label{fig:inc_sleep}
-\end{figure}
-
-% Chunks
-\subsection{Experiment 1: Number of chunks}
-In Figure~\ref{fig:inc_chunk} Spark is not compared for 30 chunks. This is because
-Spark has a 2GB limitation in the task size it can compute. Note from
-Figure~\ref{fig:inc_chunk}(b) that the compute time increase when there are more
-chunks. This is expected because the sleep delay is constant throughout this
-experiment and there are more compute tasks when the number of chunks increases. We
-also note that in Figure~\ref{fig:inc_chunk}(b) for 30 chunks the Dask bag API has a
-much lower overhead time. This is because we only calculate the idle time of the used
-core. Dask bag was only using one thread per block in comparison to Dask delayed and
-Dask futures which was offloading the calculations on multiple threads of the same
-worker. From Figure~\ref{fig:inc_chunk} we observe that when the dataset is separated
-in more (smaller) chunks it saves IO time however it increases the overhead by a
-similar amount; the inverse happens when we reduce the number of chunks.
-
-\begin{figure}[!t]
-    \centering
-    \subfloat[Incrementation makespan]{%
-      \includegraphics[clip,width=\columnwidth]{images/inc_chunk.png}%
-    }
-
-    \subfloat[Incrementation total time]{%
-      \includegraphics[clip,width=\columnwidth]{images/inc_idle_chunk.png}%
-    }
-    \caption{10 iterations, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_chunk}
+        \caption{Incrementation total time}\label{fig:inc_tt_chunk}
+    \end{subfigure}
+    \caption{125 chunks, 10 iterations, 8 instances}\label{fig:inc_sleep}
 \end{figure}
@@
 \subsection{Experimentation 1: Baseline timeline}
-From Table~\ref{tb:inc-base} and Figure~\ref{fig:inc_gantt}, the frameworks have
+From Table~\ref{tb:inc_base} and Figure~\ref{fig:inc_gantt}, the frameworks have
 similar makespan however the time spend for each type of task differs significantly.
@@
     \caption{Distribution of the execution time in seconds for our baseline: 125
-    chunks, 4 sec.\ sleep delay, 10 iterations, and 8 workers}\label{tb:inc-base}
+    chunks, 4 sec.\ sleep delay, 10 iterations, and 8 workers}\label{tb:inc_base}
     \centering
-    \begin{tabular}{llllll}
+    \begin{tabular*}{\columnwidth}{llllll}
     \hline
@@
     Dask.futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
-    \end{tabular}
+    \end{tabular*}
  \end{table}
 
-\begin{figure}[!t]
+\begin{figure*}[!t]
     \centering
-    \subfloat[Spark execution timeline]{%
-      \includegraphics[clip,width=\columnwidth]{images/spark_inc_baseline_gantt.png}%
-    }
-
-    \subfloat[Dask bag execution timeline]{%
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/spark_inc_baseline_gantt.png}
+        \caption{Spark execution timeline}\label{fig:inc_spark_gantt}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{\columnwidth}
         \includegraphics[clip,width=\columnwidth]{images/Dask_bag_inc_baseline_gantt.png}%
-    }
-
-    \subfloat[Dask delayed execution timeline]{%
+        \caption{Dask bag execution timeline}\label{fig:inc_dask_bag_gantt}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
         \includegraphics[clip,width=\columnwidth]{images/Dask_delayed_inc_baseline_gantt.png}%
-    }
-
-    \subfloat[Dask futures execution timeline]{%
+        \caption{Dask delayed execution timeline}\label{fig:inc_dask_delayed_gantt}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{\columnwidth}
         \includegraphics[clip,width=\columnwidth]{images/Dask_futures_inc_baseline_gantt.png}%
-    }
+        \caption{Dask futures execution timeline}\label{fig:inc_dask_futures_gantt}
+    \end{subfigure}
     \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}
-    \label{fig:inc_gantt}
-\end{figure}
+\end{figure*}
 
@@
     \centering
-    \subfloat[Histogram makespan]{%
+    \begin{subfigure}[b]{\columnwidth}
         \includegraphics[clip,width=\columnwidth]{images/histo_instance.png}%
-    }
-
-    \subfloat[Histogram total time]{%
+        \caption{Histogram makespan}\label{fig:histo_ms_worker}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
         \includegraphics[clip,width=\columnwidth]{images/histo_idle_instances.png}%
-    }
-    \caption{8 instances}\label{fig:histo_wroker}
+        \caption{Histogram total time}\label{fig:histo_tt_worker}
+    \end{subfigure}
+    \caption{125 chunks}
 \end{figure}
@@
     \centering
-    \subfloat[Histogram makespan]{%
+    \begin{subfigure}[b]{\columnwidth}
         \includegraphics[clip,width=\columnwidth]{images/histo_splits.png}%
-    }
-
-    \subfloat[Histogram total time]{%
+        \caption{Histogram makespan}\label{fig:histo_ms_chunk}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
         \includegraphics[clip,width=\columnwidth]{images/histo_idle_splits.png}%
-    }
-    \caption{10 iterations, 4 sec.\ sleep delay, 8 instances}\label{fig:histo_chunk}
+        \caption{Histogram total time}\label{fig:histo_tt_chunk}
+    \end{subfigure}
+    \caption{8 workers}
 \end{figure}

### 2019-06-20 latex/paper.tex
Modified results section for incrementation and fixed typos

@@
 We propose to use Dask since it provides native Python parallelism to pipelines while
-providing support to familiar API from Python's scientific ecosystem. Unfortunately,
-there are few comparisons between Spark and Dask; especially in neuroimaging.
-Moreover, these studies were done when Dask was still an immature framework which
-makes those studies unfair nowadays. This is our motivation to compare the latest
-version of Dask with Spark.
+providing support to familiar API from the Python's scientific ecosystem.
+Unfortunately, there are few comparisons between Spark and Dask; especially in
+neuroimaging. Moreover, these studies were done when Dask was still an immature
+framework which makes those studies unfair nowadays. This is our motivation to
+compare the latest version of Dask with Spark.
 % Methods (one sentence)
@@
 due to a required serialization. Moreover, while Spark's API is flexible and allows
-most implementations, it differs from the ones seen in Python's ecosystem.
+most implementations, it differs from the ones seen in the Python's ecosystem.
 
@@
 On the other hand, Dask was created with the purpose of natively parallelize Python
-pipelines while keeping the syntax of familiar API from Python's scientific
+pipelines while keeping the syntax of familiar API from the Python's scientific
 ecosystem. However, Dask is still a young framework with work to be done; it's API
@@
 %% Dask API
-The Dask APIs we decide to use for our comparison are Dask bag, delayed and futures.
-Dask bag offers an easy API to parallelize data. Dask delayed offer a lower-level API
+The Dask APIs we decide to use for our comparison are Dask Bag, Delayed and Futures.
+Dask Bag offers an easy API to parallelize data. Dask Delayed offer a lower-level API
 that offers more flexibility; this is good to implement more complex tasks that do
-not fit in the Dask bag framework. Dask futures is a real-time API. Both Dask bag and
-delayed apply lazy evaluation to tasks while futures trigger them directly.
+not fit in the Dask Bag framework. Dask Futures is a real-time API. Both Dask Bag and
+Delayed apply lazy evaluation to tasks while Futures trigger them directly.
 
@@
 \subsubsection{PySpark RDD~\cite{spark-rdd}} Spark is a general-purpose Big Data
-framework. It provides in-memory computing, data locality and lazy evaluation. Spark
-abstracts resilient distributed dataset (RDD)~\cite{RDD} making it fault-tolerant. Its
-high-level API provides optimizations for transformations applied data. Since
-Spark is written in Scala, this optimization is lessen because it requires a
-serialization from Python to Java. The API offered by Spark is flexible and suits most
-problems however it differs from Python ecosystem. The standalone scheduler is used for
-our experiment. It has a LIFO policy.
+framework. It provides in-memory computing, data locality, and lazy evaluation. Spark
+abstracts resilient distributed dataset (RDD)~\cite{RDD} making it fault-tolerant.
+Its high-level API provides optimizations for transformations applied data. Since
+Spark is written in Scala, this optimization is lessened because it requires a
+serialization from Python to Java. The API offered by Spark is flexible and suits
+most problems however it differs from the Python ecosystem. The standalone scheduler
+is used for our experiment. It has a LIFO policy.
 
 \subsubsection{Dask Bag~\cite{dask-bag}} Dask Bag is one of the high-level API
-offered by Dask. Like Spark, it provides in-memory computing, data locality and lazy
-evaluation however it does not have fault resilience. Fundamently, it is parallized
-lists. Dask Bag parallelizes computations of large collection of generic python
-objects~\cite{dask-bag}. Also, it offers optimized algorithms that mimic most of the
-toolz~\cite{toolz} library. Our experiment use the Dask Distributed~\cite{dask-distributed}
-scheduler. It has a LIFO policy. In theory it should speed-up the pipeline excution
-by finishing a branch of task computations before starting new ones.
+offered by Dask. Like Spark, it provides in-memory computing, data locality, and lazy
+evaluation however it does not have fault resilience. Fundamentally, it is
+parallelized lists. Dask Bag parallelizes computations across a large collection of
+generic python objects~\cite{dask-bag}. Also, it offers optimized algorithms that
+mimic most of the toolz~\cite{toolz} library. Our experiment uses the Dask
+Distributed~\cite{dask-distributed} scheduler. It has a LIFO policy. In theory, it
+should speed-up the pipeline execution by finishing a branch of task computations
+before starting new ones.
 
@@
 does not offer the optimized algorithm as the high-level APIs however it allows the
-user to parallelize custom function that do not fit in the array, dataframe or
-MapReduce framework. This is perfect for more complex function which could easily be
-parallelized. Like Dask Bag, it offers in-memory computing, data locality and lack
-fault resilience. The Dask Distributed scheduler is used.
+user to parallelize custom function that does not fit in the array, dataframe or
+MapReduce framework. This is perfect for the more complex functions which could
+easily be parallelized. Like Dask Bag, it offers in-memory computing, data locality
+and lacks fault resilience. The Dask Distributed scheduler is used.
 
 \subsubsection{Dask Futures~\cite{dask-futures}} Dask Futures is also a low-level API
-from Dask. Like the other Dask APIs, it benifits of in-memory computing and data
-locatlity. However, unlike the other APIs, Dask Futures triggers computation
+from Dask. Like the other Dask APIs, it benefits of in-memory computing and data
+locality. However, unlike the other APIs, Dask Futures triggers computation
 immediately. When the user tries to gather data if available it will return the
@@
 \subsection{Dataset}
-For the incrementation and histogram experiment the BigBrain~\cite{Amunts:13} is
+For the incrementation and histogram experiment, the BigBrain~\cite{Amunts:13} is
 used. It is a three-dimensional image of a brain scanned at \SI{40}{\micro\metre}.
-The data set store the intensity of each voxel and their position. This results in a
+The dataset stores the intensity of each voxel and its position. This results in a
 75GB image. It is stored in 125 chunks in the \textit{MINC}~\cite{minc} format. To
@@
 image, increments the intensity value of each voxel by 1 for a specified amount of
-iterations, and write the result to an NFS as a NifTI image. Incrementing the data
+iterations, and writes the result to an NFS as a NifTI image. Incrementing the data
 reduce the caching effect. This experiment allows us to study the behavior of the
-frameworks when all input are processed independently; i.e.\ each task in the graph
-only depends on the previous one (see Figure~\ref{fig:tg-inc}). This mimic the
+frameworks when all inputs are processed independently; i.e.\ each task in the graph
+only depends on the previous one (see Figure~\ref{fig:tg-inc}). This mimics the
 behavior of analyzing multiple independent subjects in parallel.
@@
 \subsubsection{\textbf{Histogram}}
-Our second experiment consist to calculate the histogram of the intensity of the
+Our second experiment consists to calculate the histogram of the intensity of the
 voxel in BigBrain (see Algorithm~\ref{alg:histogram}). It reads the images from the
-nfs, calculate inividual image frequency, aggregate the images frequency together,
-and write the result to a file. In this experiment, the result depends on all the
-input (see Fig.~\ref{fig:tg-histo}). This allows to study the behavior of the
-frameworks when tasks are dependent. Also, this application requires data shuffling
-thus inter-worker communication. We decided to omit Dask Futures as it does not had
-any value over Dask Delayed in this application.
+NFS, calculate individual image frequency, aggregate the frequency of the images
+together, and write the result to a file. In this experiment, the result depends on
+all the input (see Fig.~\ref{fig:tg-histo}). This allows the study of the behavior of
+the frameworks when tasks are dependent. Also, this application requires data
+shuffling thus inter-worker communication. We decided to omit Dask Futures as it does
+not have any value over Dask Delayed in this application.
 
@@
 scaling of the frameworks in a cluster. Also, it helps in determining the level of
-parallelism of the frameworks.
+the parallelism of the frameworks.
 
-\textbf{Number of chunks:} Evaluating the frameworks on different amount of chunks
+\textbf{Number of chunks:} Evaluating the frameworks on a different amount of chunks
 allows us to understand their behavior with different task size and IO requirements.
 
-\textbf{Number of iterations:} We change the amount of iterations for the
-incrementation~\ref{alg:incrementation} to study the effect of the job size.
+\textbf{Number of iterations:} We change the number of iterations for the
+incrementation~\ref{alg:incrementation} to study the effect of the job length.
 
 \textbf{Sleep delay:} We modify the sleep delay in the
-incrementation~\ref{alg:incrementation} to understand the behavior of task of
-different length.
+incrementation~\ref{alg:incrementation} to understand the behavior of different task
+lengths.
 
@@
 
-\begin{table}[!b]
+\begin{table}[!t]
     \renewcommand{\arraystretch}{1.3}
@@
 All experiments were performed on Arbutus Cloud Compute Canda. Each node has CentOS
-7.5.1804 as base operating system and have the Linux kernel version
+7.5.1804 as the base operating system and have the Linux kernel version
 3.10.0\-862.11.6.el7.x86\_64. The nodes have an Intel Xeon Gold 6130 processor with
-30GB 2666 MHz memory. The networks has a 10GB/s bandwidth. The filesystem is 2TB NFS v4.
+30GB 2666 MHz memory. The network has a 10GB/s bandwidth. The filesystem is 2TB NFS
+v4.
 
@@
 %% worker
-\subsection{Experiment 1: Number of instances}
-Figure~\ref{fig:inc_worker}(a) shows that the increase in performance is not
-proportional to the number of workers which differ from what we would expect; it even
-decrease when we reach 8 workers. This is explained by an increase in IO overhead
-correlated to the number of workers (Figure~\ref{fig:inc_worker}(b)), as more worker
-access the NFS simultaneously it increases the stress put on the system which results
-in bottleneck, errors and slow down.
+\subsection{Experiment 1: Number of workers}
+Figure~\ref{fig:inc_ms_worker} shows the makespan for the incrementation algorithm on
+the y-axis and the number of workers on the x-axis; the color represents the
+different frameworks. Overall, there is no substantial makespan difference between
+the frameworks. Also, the increase in performance is not proportional to the number
+of workers; performance even decreases when we reach 8 workers.
 
-On the one hand, Dask seems better when using a single worker. On the other hand,
-Spark makespan decreases faster for a higher level of parallelism. Dask is constantly
-faster than Spark which makes us think it has a better scheduler.
+In Figure~\ref{fig:inc_tt_worker}, the y-axis shows the total execution time of the
+incrementation application. The hatch indicates that the time is spent for: read,
+compute, write, and overhead. Again, the color represents the frameworks. The
+computing time stays similar when the number of workers increases. However, the IO
+time and overhead increase proportionally to the number of workers.
 
-\begin{figure}[!t]
+\begin{figure}[!b]
     \centering
@@
 \subsection{Experiment 1: Number of chunks}
-In Figure~\ref{fig:inc_chunk} Spark is not compared for 30 chunks. This is because
-Spark has a 2GB limitation in the task size it can compute. Note from
-Figure~\ref{fig:inc_chunk}(b) that the compute time increase when there are more
-chunks. This is expected because the sleep delay is constant throughout this
-experiment and there are more compute tasks when the number of chunks increases. We
-also note that in Figure~\ref{fig:inc_tt_chunk} for 30 chunks the Dask bag API has a
-much lower overhead time. This is because we only calculate the idle time of the used
-core. Dask bag was only using one thread per block in comparison to Dask delayed and
-Dask futures which was offloading the calculations on multiple threads of the same
-worker. From Figure~\ref{fig:inc_chunk} we observe that when the dataset is separated
-in more (smaller) chunks it saves IO time however it increases the overhead by a
-similar amount; the inverse happens when we reduce the number of chunks.
+In Figure~\ref{fig:inc_chunk}, Spark is not compared for 30 chunks. This is because
+Spark has a 2GB limitation in the task size it can compute.
+
+The Figure~\ref{fig:inc_ms_chunk} shows the makespan evolution when varying the
+number of chunks. The compute time increase when there are more chunks. This is
+because the sleep delay is constant throughout this experiment and there are more
+compute tasks when the number of chunks increases.
+
+In Figure~\ref{fig:inc_tt_chunk}, the total execution time of each function is shown.
+For 30 chunks the Dask Bag API has a much lower overhead time. This is because we
+only calculate the idle time of the used core. Dask Bag was only using one thread per
+block in comparison to Dask Delayed and Dask Futures which was offloading a block
+calculations on multiple threads of the same worker. Furthermore, the variance of the
+overhead increases proportionally to the number of chunks. Also, the IO time reduces
+with more chunks however the overhead time increase by a similar amount. This is not
+observed for 30 chunks as the workers are not used at full capacity; i.e.\ some
+threads are idle.
 
@@
 \subsection{Experiment 1: Number of iterations}
-In Figure~\ref{fig:inc_itr}, Spark and the Dask bag API are equivalent in terms of
-performance. The Dask delayed API seems to perform well with a low amount of task but
-grow faster the Spark and Dask bag. We think this is because Dask delayed often
-schedule a block on a different thread of the same worker which causes small delays
-every time due to inter-thread data communication. This is good when there are fewer
-tasks as is make the IO between blocks go out of sync hence lowering the IO
-bottleneck however as the number of tasks increases those small delays become
-significant. The Dask futures API seems to outperform all the other APIs but this is
-most likely because the tasks are not interdependent thus this application benefits
-from the less optimal but faster scheduling brought by Dask futures.
+Figure~\ref{fig:inc_ms_itr} shows the makespan of the application for the various
+number of iterations. Spark and the Dask Bag API are equivalent in terms of
+performance. The Dask Delayed API is slightly faster with few iterations (tasks) but
+it has a similar makespan to Spark and Dask Bag when the number of iterations
+increases. Overall, Dask Futures seems to be faster but the difference is not
+substantial.
+
+In Figure~\ref{fig:inc_tt_itr}, the execution time of the function type is shown. The
+compute time increase proportionally to the number of iterations. This is because
+each computing task has a constant sleep delay. Also, the overhead increases when the
+number of iterations (tasks) increases while the IO time decreases; however this is
+not substantial.
+
+% We think this is because Dask Delayed often
+% schedule a block on a different thread of the same worker which causes small delays
+% every time due to inter-thread data communication. This is good when there are fewer
+% tasks as is make the IO between blocks go out of sync hence lowering the IO
+% bottleneck however as the number of tasks increases those small delays become
+% significant.
+
+%  The Dask Futures API seems to outperform all the other APIs but this is
+% most likely because the tasks are not interdependent thus this application benefits
+% from the less optimal but faster scheduling brought by Dask Futures.
 
@@
         \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
-        \caption{Incrementation makespan}\label{fig:inc_ms_chunk}
+        \caption{Incrementation makespan}\label{fig:inc_ms_itr}
     \end{subfigure}
@@
         \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
-        \caption{Incrementation total time}\label{fig:inc_tt_chunk}
+        \caption{Incrementation total time}\label{fig:inc_tt_itr}
     \end{subfigure}
-    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_itr}
+    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}
 \end{figure}
@@
 \subsection{Experiment 1: Sleep delay}
-As Figure~\ref{fig:inc_sleep}(b) and Table~\ref{tb:inc-sleep} shows Spark has the
-smallest overhead all around and it is also the one that reduces the most with longer
-task length. In term of overhead, Dask bag trailing slightly behind Spark although is
-better than Dask delayed and Dask futures. The opposite happens for the IO time.
-% Not sure what to conclude.
+Figure~\ref{fig:inc_ms_sleep} shows the makespan of the incrementation algorithm for
+different sleep delays. Spark is initially slower than the Dask APIs, however, it is
+because faster when increasing the sleep time. Also, within the Dask APIs, Dask Bag
+is slower than the other two but it is not considerable.
 
-%%%% CHANGE NUMBER
+Figure~\ref{fig:inc_tt_sleep} shows the execution time for each type of function. On
+one hand, Spark has the smallest overhead. Table~\ref{tab:inc_sleep_overhead} shows
+that it is substantial; especially with a 64 seconds sleep delay. On the other hand,
+Spark has a high IO time. In contrast, Dask Delayed and Futures have the largest
+overhead time but the smallest IO time. Then, Dask Bag stands in the middle for both
+overhead and IO time.
+
 \begin{table}[!t]
+    \caption{Function time for the sleep experiment}
+    \begin{subtable}[b]{\columnwidth}
         \renewcommand{\arraystretch}{1.3}
-    \caption{Distribution of the execution time in seconds for our baseline: 125
-    chunks, 4 sec.\ sleep delay, 10 iterations, and 8 workers}\label{tb:inc-sleep}
+        \caption{Overhead time in second}\label{tab:inc_sleep_overhead}
         \centering
-    \begin{tabular*}{\columnwidth}{llllll}
+        \begin{tabular}{lllll}
         \hline
-                 & Read  & Compute & Write & Overhead & Total \\ \hline
-    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
-    Dask.bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
-    Dask.delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
-    Dask.futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
-    \end{tabular*}
+                     & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
+        Spark        & 14992  & 17076  & 16069   & 9769    \\
+        Dask.Bag     & 20690  & 21388  & 17759   & 18720   \\
+        Dask.Delayed & 27481  & 26199  & 22080   & 24624   \\
+        Dask.Futures & 27069  & 27984  & 23538   & 24762   \\ \hline
+        \end{tabular}
+    \end{subtable}
+    \vskip 0.2cm
+    \begin{subtable}[b]{\columnwidth}
+        \renewcommand{\arraystretch}{1.3}
+        \caption{IO time in second}\label{tab:inc_sleep_io}
+        \centering
+        \begin{tabular}{lllll}
+        \hline
+                     & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
+        Spark        & 63622  & 57144  & 51070   & 57078   \\
+        Dask.Bag     & 54630  & 52312  & 53743   & 53612   \\
+        Dask.Delayed & 45239  & 44142  & 45133   & 46198   \\
+        Dask.Futures & 46537  & 43161  & 44749   & 45846   \\ \hline
+        \end{tabular}
+    \end{subtable}
  \end{table}
 
-\begin{figure}[!t]
+\begin{figure}[!b]
     \centering
@@
         \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
-        \caption{Incrementation makespan}\label{fig:inc_ms_chunk}
+        \caption{Incrementation makespan}\label{fig:inc_ms_sleep}
     \end{subfigure}
@@
         \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
-        \caption{Incrementation total time}\label{fig:inc_tt_chunk}
+        \caption{Incrementation total time}\label{fig:inc_tt_sleep}
     \end{subfigure}
-    \caption{125 chunks, 10 iterations, 8 instances}\label{fig:inc_sleep}
+    \caption{125 chunks, 10 iterations, 8 instances}
 \end{figure}
@@
 \subsection{Experimentation 1: Baseline timeline}
-From Table~\ref{tb:inc_base} and Figure~\ref{fig:inc_gantt}, the frameworks have
+From Table~\ref{tab:inc_base} and Figure, the frameworks have
 similar makespan however the time spend for each type of task differs significantly.
 Spark tends to spend more time than other frameworks for IO however it has a much
-lower overhead. Dask delayed and futures have the lowest IO time but a much higher
-overhead. Dask bag stands in the middle for both IO and overhead. Compute time is
+lower overhead. Dask Delayed and Futures have the lowest IO time but a much higher
+overhead. Dask Bag stands in the middle for both IO and overhead. Compute time is
 approximately the same for all frameworks.
 
-\begin{table}[!t]
-    \renewcommand{\arraystretch}{1.3}
-    \caption{Distribution of the execution time in seconds for our baseline: 125
-    chunks, 4 sec.\ sleep delay, 10 iterations, and 8 workers}\label{tb:inc_base}
-    \centering
-    \begin{tabular*}{\columnwidth}{llllll}
-    \hline
-                    & Read  & Compute & Write & Overhead & Total \\ \hline
-    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
-    Dask.bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
-    Dask.delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
-    Dask.futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
-    \end{tabular*}
- \end{table}
-
-\begin{figure*}[!t]
+\begin{figure*}[!htb]
     \centering
@@
         \includegraphics[clip,width=\columnwidth]{images/Dask_bag_inc_baseline_gantt.png}%
-        \caption{Dask bag execution timeline}\label{fig:inc_dask_bag_gantt}
+        \caption{Dask Bag execution timeline}\label{fig:inc_dask_bag_gantt}
     \end{subfigure}
@@
         \includegraphics[clip,width=\columnwidth]{images/Dask_delayed_inc_baseline_gantt.png}%
-        \caption{Dask delayed execution timeline}\label{fig:inc_dask_delayed_gantt}
+        \caption{Dask Delayed execution timeline}\label{fig:inc_dask_delayed_gantt}
     \end{subfigure}
@@
         \includegraphics[clip,width=\columnwidth]{images/Dask_futures_inc_baseline_gantt.png}%
-        \caption{Dask futures execution timeline}\label{fig:inc_dask_futures_gantt}
+        \caption{Dask Futures execution timeline}\label{fig:inc_dask_futures_gantt}
     \end{subfigure}
@@
 \end{figure*}
+
+\begin{table}[!b]
+    \renewcommand{\arraystretch}{1.3}
+    \caption{Distribution of the execution time in second: 125 chunks, 4 sec.\ sleep
+    delay, 10 iterations, and 8 workers}\label{tab:inc_base}
+    \centering
+    \begin{tabular*}{\columnwidth}{llllll}
+    \hline
+                 & Read  & Compute & Write & Overhead & Total \\ \hline
+    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
+    Dask Bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
+    Dask Delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
+    Dask Futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
+    \end{tabular*}
+ \end{table}
 
@@
 
-\section*{References}
 \bibliographystyle{IEEEtran}

### 2019-06-20 latex/paper.tex
Modified results section

@@
     \end{subtable}
+    \vspace{-3mm}
  \end{table}
 
-\begin{figure}[!b]
+\begin{figure}[!t]
     \centering
@@
 \subsection{Experimentation 1: Baseline timeline}
-From Table~\ref{tab:inc_base} and Figure, the frameworks have
-similar makespan however the time spend for each type of task differs significantly.
-Spark tends to spend more time than other frameworks for IO however it has a much
-lower overhead. Dask Delayed and Futures have the lowest IO time but a much higher
-overhead. Dask Bag stands in the middle for both IO and overhead. Compute time is
-approximately the same for all frameworks.
+Figure\ref{fig:inc_gantt} shows the tasks timeline for each worker thread. On the
+x-axis is the time in seconds and on the y-axis is the workers and their threads. The
+color represents the type of different tasks; note that the color differs from one
+graph to another. The first section of reading tasks is much larger than the
+following ones. This is because, initially, the NFS is saturated by the number of
+workers accessing it. Also, a substantial amount of overhead happens near the IO
+tasks. Moreover, for the computing tasks, there is more overhead for the latter tasks.
+
+% The frameworks have similar makespan however the time spend for each type of
+% task differs considerably (see Table~\ref{tab:inc_base}). Spark tends to spend more
+% time than other frameworks for IO however it has a much lower overhead. Dask Delayed
+% and Futures have the lowest IO time but a much higher overhead. Dask Bag stands in
+% the middle for both IO and overhead. Compute time is approximately the same for all
+% frameworks. Apart from the first IO section, there is a large amount of ovehead
+% everytime time IO is performed. This is true for all frameworks.
 
@@
     \end{subfigure}
-    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}
+    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_gantt}
 \end{figure*}
 
-\begin{table}[!b]
-    \renewcommand{\arraystretch}{1.3}
-    \caption{Distribution of the execution time in second: 125 chunks, 4 sec.\ sleep
-    delay, 10 iterations, and 8 workers}\label{tab:inc_base}
-    \centering
-    \begin{tabular*}{\columnwidth}{llllll}
-    \hline
-                 & Read  & Compute & Write & Overhead & Total \\ \hline
-    Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
-    Dask Bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
-    Dask Delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
-    Dask Futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
-    \end{tabular*}
- \end{table}
+% \begin{table}[!b]
+%     \renewcommand{\arraystretch}{1.3}
+%     \caption{Distribution of the execution time in second: 125 chunks, 4 sec.\ sleep
+%     delay, 10 iterations, and 8 workers}\label{tab:inc_base}
+%     \centering
+%     \begin{tabular*}{\columnwidth}{llllll}
+%     \hline
+%                  & Read  & Compute & Write & Overhead & Total \\ \hline
+%     Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
+%     Dask Bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
+%     Dask Delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
+%     Dask Futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
+%     \end{tabular*}
+%  \end{table}
 

### 2019-06-21 histogram/Histogram.py
Removed numpy implementation

@@ -1,6 +1,5 @@
+from collections import defaultdict
 from time import time
-
-import numpy as np
 
 from utils import benchmark
 
@@ -10,10 +9,9 @@ def calculate_histogram(arr, *, args, start, filename):
 
     arr = arr.flatten('F')
 
-    y = np.bincount(arr)
-    ii = np.nonzero(y)[0]
-    out = np.vstack((ii, y[ii])).T
-    histogram = {k: v for k, v in out}
+    histogram = defaultdict()
+    for x in arr:
+        histogram[x] += 1
 
     end_time = time() - start
 

### 2019-06-21 histogram/Histogram.py
Fixed defaultdict typr

@@ -9,7 +9,7 @@ def calculate_histogram(arr, *, args, start, filename):
 
     arr = arr.flatten('F')
 
-    histogram = defaultdict()
+    histogram = defaultdict(int)
     for x in arr:
         histogram[x] += 1
 

### 2019-06-26 latex/paper.tex

@@
 abstracts resilient distributed dataset (RDD)~\cite{RDD} making it fault-tolerant.
-Its high-level API provides optimizations for transformations applied data. Since
+Its high-level API provides optimization for transformations applied on data. Since
 Spark is written in Scala, this optimization is lessened because it requires a
@@
 MapReduce framework. This is perfect for the more complex functions which could
-easily be parallelized. Like Dask Bag, it offers in-memory computing, data locality
-and lacks fault resilience. The Dask Distributed scheduler is used.
+potentially be parallelized. Like Dask Bag, it offers in-memory computing, data
+locality and lacks fault resilience. The Dask Distributed scheduler is also used.
 
@@
 In Figure~\ref{fig:inc_tt_worker}, the y-axis shows the total execution time of the
-incrementation application. The hatch indicates that the time is spent for: read,
+incrementation application. The hatch indicates the total time spent for: read,
 compute, write, and overhead. Again, the color represents the frameworks. The
@@
 different sleep delays. Spark is initially slower than the Dask APIs, however, it is
-because faster when increasing the sleep time. Also, within the Dask APIs, Dask Bag
-is slower than the other two but it is not considerable.
+faster when increasing the sleep delay. Also, within the Dask APIs, Dask Bag is
+slower than the other two but it is not considerable.
 

### 2019-06-26 bidsApp-examples/dask-delayed.py
BIDS App Example implemented with Dask Delayed

@@ -1,0 +1,118 @@
+import argparse
+from glob import glob
+import os
+import subprocess
+import sys
+from time import time
+
+import dask
+from dask.distributed import Client
+import nibabel
+import numpy
+
+sys.path.append("/nfs/paper-big-data-engines")
+
+
+def run(command, *, start, args):
+
+    start_time = time() - start
+
+    subprocess.run(
+        command, stdout=subprocess.PIPE, stderr=subprocess.STDOUT, shell=True
+    )
+
+    end_time = time() - start
+
+    if args.benchmark:
+        benchmark(
+            start_time,
+            end_time,
+            brain_file,
+            args.output_dir,
+            args.experiment,
+            run.__name__,
+        )
+
+
+def get_data(brain_file, *, start, args):
+
+    start_time = time() - start
+
+    data = nibabel.load(brain_file).get_data()
+
+    end_time = time() - start
+
+    if args.benchmark:
+        benchmark(
+            start_time,
+            end_time,
+            brain_file,
+            args.output_dir,
+            args.experiment,
+            get_data.__name__,
+        )
+    return data
+
+
+if __name__ == "__main__":
+    start = time()
+
+    parser = argparse.ArgumentParser(description="Example BIDS App entrypoint script.")
+    parser.add_argument("scheduler", type=str, help="Scheduler ip and port")
+    parser.add_argument(
+        "bids_dir",
+        help="The directory with the input dataset "
+        "formatted according to the BIDS standard.",
+    )
+    parser.add_argument(
+        "output_dir",
+        help="The directory where the output files "
+        "should be stored. If you are running group level analysis "
+        "this folder should be prepopulated with the results of the"
+        "participant level analysis.",
+    )
+    parser.add_argument(
+        "experiment", type=str, help="Name of the experiment being performed"
+    )
+    parser.add_argument("--benchmark", action="store_true", help="benchmark results")
+
+    args = parser.parse_args()
+
+    subject_dirs = glob(os.path.join(args.bids_dir, "sub-*"))
+    subjects_to_analyze = [subject_dir.split("-")[-1] for subject_dir in subject_dirs]
+
+    # Cluster scheduler
+    cluster = args.scheduler
+    client = Client(cluster)
+
+    print(client)
+    client.upload_file("/nfs/paper-big-data-engines/utils.py")
+    from utils import benchmark
+
+    # running participant level
[…]

### 2019-06-26 latex/paper.tex
Worked on the engine section

@@
 \usepackage{hyperref}
+\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
 \usepackage{algpseudocode}
@@
 \usepackage{subcaption}
+
+\newcommand{\TG}[1]{\color{blue}From Tristan: #1 \color{black}}
+
 \algnewcommand\algorithmicforeach{\textbf{for each}}
@@
 
-\title{Comparative study of Dask and Spark for neuroimaging pipelines\\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-should not be used}
-\thanks{Natural Sciences and Engineering Research Council}
-}
+\title{A Comparative study of Dask and Spark for neuroimaging pipelines}
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Mathieu Dugr\'e}
-\IEEEauthorblockA{\textit{Computer Science and Software Engineering} \\
-\textit{Concordia University}\\
-Montreal, Canada \\
-mathieu.dugre@mail.concordia.ca}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{Computer Science and Software Engineering} \\
-\textit{Concordia University}\\
-Montreal, Canada \\
-email address}
-\and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{Computer Science and Software Engineering} \\
-\textit{Concordia University}\\
-Montreal, Canada \\
-email address}
+\author{Mathieu Dugr\'e, Val\'erie Hayot-Sasson, Tristan Glatard\\
+Department of Computer Science and Software Engineering\\
+Montr\'eal, Qu\'ebec, Canada
 }
@@
 %Similarities
+\TG{Note for later: explain the following concepts in the intro: lazy eval, data locality, in-memory computing.}
 Spark and Dask offer in-memory computing, data locality, and lazy evaluation; which
@@
 \subsection{Engines}
-For all our experiments, we used Spark v2.4.0 and Dask v1.1.4.
 
-\subsubsection{PySpark RDD~\cite{spark-rdd}} Spark is a general-purpose Big Data
-framework. It provides in-memory computing, data locality, and lazy evaluation. Spark
-abstracts resilient distributed dataset (RDD)~\cite{RDD} making it fault-tolerant.
-Its high-level API provides optimization for transformations applied on data. Since
-Spark is written in Scala, this optimization is lessened because it requires a
-serialization from Python to Java. The API offered by Spark is flexible and suits
-most problems however it differs from the Python ecosystem. The standalone scheduler
-is used for our experiment. It has a LIFO policy.
+\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data
+framework. It provides in-memory computing, data locality, and lazy
+evaluation. Spark's main data structure, the Resilient Distributed Dataset
+(RDD)~\cite{RDD}, is a fault-tolerant and parallel collection of data
+elements. 
+Dataframes, Spark's other popular API, are an overlay of RDDs with similar 
+performance features. 
+Spark's core is written in Scala, but it also has an R and a Python API
+called PySpark. We used PySpark, given the popularity of Python in
+neuroscience, although the serialization cost of Python objects to Scala
+might be a substantial source of overhead. We used Spark's default
+scheduler, Spark standalone, although YARN~\cite{yarn} and
+Mesos~\cite{mesos} are also available. In the Spark standalone scheduler, a
+process called the \emph{master} coordinates resources provisioned by
+\emph{workers} on the cluster. The application is submitted to the
+\emph{driver} that in turn requests workers to the master and dispatches
+tasks to them. The Spark standalone scheduler uses a LIFO
+(Last-In-First-Out) task scheduling policy \TG{we should describe it
+briefly and refer to the paper}. The Spark standalone scheduler has two
+execution modes: client mode, where the driver runs in a dedicated  process
+on the cluster, and cluster mode, where the driver runs in a worker. We
+used client mode as PySpark is not available in cluster mode. We used
+Apache Spark v2.4.0.
 
-\subsubsection{Dask Bag~\cite{dask-bag}} Dask Bag is one of the high-level API
-offered by Dask. Like Spark, it provides in-memory computing, data locality, and lazy
-evaluation however it does not have fault resilience. Fundamentally, it is
-parallelized lists. Dask Bag parallelizes computations across a large collection of
-generic python objects~\cite{dask-bag}. Also, it offers optimized algorithms that
-mimic most of the toolz~\cite{toolz} library. Our experiment uses the Dask
-Distributed~\cite{dask-distributed} scheduler. It has a LIFO policy. In theory, it
-should speed-up the pipeline execution by finishing a branch of task computations
-before starting new ones.
+\subsubsection{Dask} Dask is a Big Data engine that is becoming
+increasingly popular in the scientific Python ecosystem. \TG{A few more
+introductory sentences should be added, following Spark's model.}
+Dask represents application tasks as Dask graphs, produced by a
+user-facing API and executed by a scheduler. We used the \href{https://distributed.dask.org/en/latest/index.html}{Dask
+Distributed} scheduler.
+Dask Distributed also has a LIFO policy. In theory, it should speed-up the
+pipeline execution by finishing a branch of task computations before
+starting new ones \TG{This is meant to reduce the memory footprint of the
+application}. \TG{Here we should add a few comparison sentences between
+Spark's and Dask's schedulers.} Dask has five main APIs: Dask Bags, Dask
+Arrays, Dask Futures, Dask Delayed and DataFrames. We compared them all
+except DataFrames as they are not adapted to the type of data we
+considered. All APIs provide in-memory computing, data locality, and lazy
+evaluation, however, none of them has fault tolerance \TG{Check fault
+tolerance in Dask. Fine-grained lineage vs Spark's coarse grained lineage}.
+We used Dask v1.1.4.
 
-\subsubsection{Dask Delayed~\cite{dask-delayed}} Dask Delayed is a low-level API. It
-does not offer the optimized algorithm as the high-level APIs however it allows the
-user to parallelize custom function that does not fit in the array, dataframe or
-MapReduce framework. This is perfect for the more complex functions which could
-potentially be parallelized. Like Dask Bag, it offers in-memory computing, data
-locality and lacks fault resilience. The Dask Distributed scheduler is also used.
-
-\subsubsection{Dask Futures~\cite{dask-futures}} Dask Futures is also a low-level API
-from Dask. Like the other Dask APIs, it benefits of in-memory computing and data
-locality. However, unlike the other APIs, Dask Futures triggers computation
-immediately. When the user tries to gather data if available it will return the
-result otherwise it will block until the computation is done. Again, the Dask
-Distributed scheduler is used.
+\href{https://docs.dask.org/en/latest/bag.html}{Dask Bags}
+are parallel collections of Python objects, similar to Spark's RDD. Dask
+Bags also offers optimized algorithms \TG{``optimized algorithms'' is
+inaccurate: you should explain what they do} that mimic most of the
+toolz~\cite{toolz} library. \href{https://docs.dask.org/en/latest/array.html}{Dask
+Arrays} are used for the
+processing of large arrays. They provide a parallel clone of the popular
+NumPy library. \href{https://docs.dask.org/en/latest/delayed.html}{Dask
+Delayed} supports arbitrary
+tasks that does not fit in the Array, DataFrame or MapReduce \TG{Do you
+mean Bags?} frameworks. Finally, \href{https://docs.dask.org/en/latest/futures.html}{Dask
+Futures} is similar to Dask
+Delayed as it supports arbitrary tasks, but it operates in real-time rather
+than lazily.
 
@@
 
-We would like to thanks Compute Canada Cloud for the infrastructure.
+Mathieu Dugr\'e was funded by a Undergraduate Summer Research Assistant
+award from the National Science and Engineering Research Council. We warmly
+thank Compute Canada and \TG{Westgrid?} for providing the cloud
+infrastructure used in these experiments, and the McGill Center for
+Integrative Neurocience for giving us access to their cloud allocation.
 

### 2019-06-26 latex/paper.tex
Edited section Datasets

@@
 application}. \TG{Here we should add a few comparison sentences between
-Spark's and Dask's schedulers.} Dask has five main APIs: Dask Bags, Dask
-Arrays, Dask Futures, Dask Delayed and DataFrames. We compared them all
-except DataFrames as they are not adapted to the type of data we
-considered. All APIs provide in-memory computing, data locality, and lazy
+Spark's and Dask's schedulers.} Dask has five main APIs and data structures: Bag,
+Array, Futures, Delayed and DataFrame. We used them all
+except DataFrame as it is not adapted to the type of data we
+used. All APIs provide in-memory computing, data locality, and lazy
 evaluation, however, none of them has fault tolerance \TG{Check fault
@@
 
-\href{https://docs.dask.org/en/latest/bag.html}{Dask Bags}
-are parallel collections of Python objects, similar to Spark's RDD. Dask
-Bags also offers optimized algorithms \TG{``optimized algorithms'' is
-inaccurate: you should explain what they do} that mimic most of the
-toolz~\cite{toolz} library. \href{https://docs.dask.org/en/latest/array.html}{Dask
-Arrays} are used for the
-processing of large arrays. They provide a parallel clone of the popular
-NumPy library. \href{https://docs.dask.org/en/latest/delayed.html}{Dask
-Delayed} supports arbitrary
-tasks that does not fit in the Array, DataFrame or MapReduce \TG{Do you
-mean Bags?} frameworks. Finally, \href{https://docs.dask.org/en/latest/futures.html}{Dask
-Futures} is similar to Dask
-Delayed as it supports arbitrary tasks, but it operates in real-time rather
+In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
+collection of Python objects, similar to Spark's RDD. It offers a
+programming abstraction similar to the
+\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}.
+An \href{https://docs.dask.org/en/latest/array.html}{Array} is used for
+the processing of large arrays. It provides a distributed clone of the
+popular
+NumPy library. \href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary
+tasks that do not fit in the Array, DataFrame or Bag frameworks. 
+Finally, \href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to
+Delayed as they support arbitrary tasks, but they operate in real-time rather
 than lazily.
@@
 \subsection{Dataset}
-For the incrementation and histogram experiment, the BigBrain~\cite{Amunts:13} is
-used. It is a three-dimensional image of a brain scanned at \SI{40}{\micro\metre}.
-The dataset stores the intensity of each voxel and its position. This results in a
-75GB image. It is stored in 125 chunks in the \textit{MINC}~\cite{minc} format. To
-perform our experiment we convert the chunks into the \textit{NifTI}~\cite{nifti}
-format. Also, using \textit{sam}~\cite{sam} we combine the chunks and split them into
-30, 125 and 750 chunks of equal size; each of 2.5GB, 0.6GB, and 0.1GB, respectively.
+
+We used BigBrain~\cite{Amunts:13}, a three-dimensional image of a human
+brain. The original data is stored in 125 blocks in the MINC~\cite{minc}
+HDF5-based format, available at
+\url{ftp://bigbrain.loris.ca/BigBrainRelease.2015/3D_Blocks/40um} at the
+resolution of \SI{40}{\micro\metre}. We converted the blocks into the
+\href{https://nifti.nimh.nih.gov/nifti-1}{NifTI} format, a popular format
+in neuroimaging. We left the NifTI blocks uncompressed, resulting in 
+a total data size of 76~GiB \TG{Check if GiB or GB}. 
+To evaluate the effect of block size, we resplit these blocks into 30, 125 and 750 blocks of 
+2.5GB, 0.6GB, and
+0.1GB \TG{check sizes}, using the sam~\cite{sam} library.
+
+We also used the dataset provided by the Consortium for Reliability and
+Reproducibility
+(\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) as
+available on
+\href{http://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
+entire dataset is 408.4~GB, containing anatomical, diffusion and functional
+images of 1,397 subjects acquired in 29 sites. Among these, we only used
+data from \TG{X} subjects, representing \TG{Y}GB overall (\TG{Z}GB by
+subject on average).
 
@@
 As our first experiment to benchmark the frameworks, we used a simple incrementation
-pipeline (see Algorithm~\ref{alg:incrementation}). It reads chunks from the BigBrain
+pipeline (see Algorithm~\ref{alg:incrementation}). It reads blocks from the BigBrain
 image, increments the intensity value of each voxel by 1 for a specified amount of
@@
     \Require{\(x\), a sleep delay in float}
-    \Require{\(file\), a file containing a chunk}
+    \Require{\(file\), a file containing a block}
     \Require{\(fs\), NFS to write image to.}
-    \State{Read \(chunk\) from \(file\)}
+    \State{Read \(block\) from \(file\)}
     \ForEach{\(i \in iterations\)}
-        \ForEach{\(chunk \in image\)}
-            \State{\(chunk\gets chunk+1\)}
+        \ForEach{\(block \in image\)}
+            \State{\(block\gets block+1\)}
             \State{Sleep \(x\)}
@@
     \EndFor
-    \State{Write \(chunk\) to \(fs\)}
+    \State{Write \(block\) to \(fs\)}
 \end{algorithmic}
@@
     \ForEach{\(file \in files\)}
-        \State{Read \(chunk\) from \(file\)}
-        \State{{Calculate \(frequency\) of \(chunks\)}}
+        \State{Read \(block\) from \(file\)}
+        \State{{Calculate \(frequency\) of \(blocks\)}}
     \EndFor
@@
 
-\textbf{Number of chunks:} Evaluating the frameworks on a different amount of chunks
+\textbf{Number of blocks:} Evaluating the frameworks on a different amount of blocks
 allows us to understand their behavior with different task size and IO requirements.
@@
 
-The baseline for our experiments is: 8 workers, 125 chunks, 10 iterations and 4
+The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
 seconds sleep delay.
@@
     \# of worker        & 1, 2, 4, 8     & 1, 2, 4, 8            & 1, 2, 4, 8            \\
-    \# of chunks        & 30, 125, 750   & 30, 125, 750          & 30, 125, 750          \\
+    \# of blocks        & 30, 125, 750   & 30, 125, 750          & 30, 125, 750          \\
     \# of iterations    & 1, 10, 100     & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\
@@
     \end{subfigure}
-    \caption{125 chunks, 10 iterations, 4 sec.\ sleep delay}\label{fig:inc_worker}
+    \caption{125 blocks, 10 iterations, 4 sec.\ sleep delay}\label{fig:inc_worker}
 \end{figure}
 
-% Chunks
-\subsection{Experiment 1: Number of chunks}
-In Figure~\ref{fig:inc_chunk}, Spark is not compared for 30 chunks. This is because
+% blocks
+\subsection{Experiment 1: Number of blocks}
+In Figure~\ref{fig:inc_block}, Spark is not compared for 30 blocks. This is because
 Spark has a 2GB limitation in the task size it can compute.
 
-The Figure~\ref{fig:inc_ms_chunk} shows the makespan evolution when varying the
-number of chunks. The compute time increase when there are more chunks. This is
+The Figure~\ref{fig:inc_ms_block} shows the makespan evolution when varying the
+number of blocks. The compute time increase when there are more blocks. This is
 because the sleep delay is constant throughout this experiment and there are more
-compute tasks when the number of chunks increases.
+compute tasks when the number of blocks increases.
 
-In Figure~\ref{fig:inc_tt_chunk}, the total execution time of each function is shown.
-For 30 chunks the Dask Bag API has a much lower overhead time. This is because we
+In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
+For 30 blocks the Dask Bag API has a much lower overhead time. This is because we
 only calculate the idle time of the used core. Dask Bag was only using one thread per
@@
 calculations on multiple threads of the same worker. Furthermore, the variance of the
-overhead increases proportionally to the number of chunks. Also, the IO time reduces
-with more chunks however the overhead time increase by a similar amount. This is not
-observed for 30 chunks as the workers are not used at full capacity; i.e.\ some
+overhead increases proportionally to the number of blocks. Also, the IO time reduces
+with more blocks however the overhead time increase by a similar amount. This is not
+observed for 30 blocks as the workers are not used at full capacity; i.e.\ some
 threads are idle.
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/inc_chunk.png}%
-        \caption{Incrementation makespan}\label{fig:inc_ms_chunk}
+        \includegraphics[clip,width=\columnwidth]{images/inc_block.png}%
+        \caption{Incrementation makespan}\label{fig:inc_ms_block}
     \end{subfigure}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/inc_idle_chunk.png}%
-        \caption{Incrementation total time}\label{fig:inc_tt_chunk}
+        \includegraphics[clip,width=\columnwidth]{images/inc_idle_block.png}%
+        \caption{Incrementation total time}\label{fig:inc_tt_block}
     \end{subfigure}
-    \caption{10 iterations, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_chunk}
+    \caption{10 iterations, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_block}
 \end{figure}
@@
     \end{subfigure}
-    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}
+    \caption{125 blocks, 4 sec.\ sleep delay, 8 instances}
 \end{figure}
@@
     \end{subfigure}
-    \caption{125 chunks, 10 iterations, 8 instances}
+    \caption{125 blocks, 10 iterations, 8 instances}
 \end{figure}
@@
     \end{subfigure}
-    \caption{125 chunks, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_gantt}
+    \caption{125 blocks, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_gantt}
 \end{figure*}
@@
 %     \renewcommand{\arraystretch}{1.3}
-%     \caption{Distribution of the execution time in second: 125 chunks, 4 sec.\ sleep
+%     \caption{Distribution of the execution time in second: 125 blocks, 4 sec.\ sleep
 %     delay, 10 iterations, and 8 workers}\label{tab:inc_base}
@@
     \end{subfigure}
-    \caption{125 chunks}
+    \caption{125 blocks}
 \end{figure}
 
-\subsection{Experiment 2: Number of chunks}
+\subsection{Experiment 2: Number of blocks}
 
@@
         \includegraphics[clip,width=\columnwidth]{images/histo_splits.png}%
-        \caption{Histogram makespan}\label{fig:histo_ms_chunk}
+        \caption{Histogram makespan}\label{fig:histo_ms_block}
     \end{subfigure}
@@
         \includegraphics[clip,width=\columnwidth]{images/histo_idle_splits.png}%
-        \caption{Histogram total time}\label{fig:histo_tt_chunk}
+        \caption{Histogram total time}\label{fig:histo_tt_block}
     \end{subfigure}
@@
 previously, some of the read tasks are significantly shorter than others while all
-chunks are of equal size.
+blocks are of equal size.
 

### 2019-06-26 latex/paper.tex
Worked in Section II

@@
 Therefore we think it is valuable to re-compare it with Spark.
+\TG{Cite~\cite{hayot2019performance}}
 
@@
 \subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data
-framework. It provides in-memory computing, data locality, and lazy
+engine. It provides in-memory computing, data locality, and lazy
 evaluation. Spark's main data structure, the Resilient Distributed Dataset
@@
 might be a substantial source of overhead. We used Spark's default
-scheduler, Spark standalone, although YARN~\cite{yarn} and
-Mesos~\cite{mesos} are also available. In the Spark standalone scheduler, a
+scheduler, Spark standalone, although YARN~\cite{vavilapalli2013apache} and
+Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler, a
 process called the \emph{master} coordinates resources provisioned by
@@
 introductory sentences should be added, following Spark's model.}
+\TG{Explain workers}
 Dask represents application tasks as Dask graphs, produced by a
@@
 NumPy library. \href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary
-tasks that do not fit in the Array, DataFrame or Bag frameworks. 
+tasks that do not fit in the Array, DataFrame or Bag APIs. 
 Finally, \href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to
@@
 
+\subsection{Infrastructure}
+
+ We used Compute Canada's
+ \href{https://docs.computecanada.ca/wiki/Cloud_resources}{Arbutus Cloud}
+ operated by the \href{https://www.westgrid.ca}{WestGrid} regional
+ organization at the University of Victoria, and running OpenStack
+ \TG{version?}. We used c8-30gb-186 cloud instances with 8 VCPUs, an Intel
+ Xeon Gold 6130 processor, 30~GB of RAM at 2666~MHz, 20~GB of mounted
+ storage, and a base image running CentOS 7.5.1804 with  Linux kernel
+ version
+ 3.10.0\-862.11.6.el7.x86\_64. Instances are connected by a 10Gb/s Ethernet network.
+
+ Cloud instances hosted a single Dask or Spark worker, configured to use 8
+ threads. We used Dask's default configuration that uses all the available
+ memory on the instance. \TG{add a sentence on memory usage policy for Dask, say it's on NFS} We
+ configured Spark to use 1 executor per worker and 25~GB of memory per
+ executor, to leave 5~GB for overhead. \TG{Check spill to disk}
+
+ One cloud instance did not host any worker, and had a 2~TB disk volume
+ shared with the other instances using the Network File System (NFS) v4.
+ This instance was also used for the Spark driver and master, and for the
+ Dask scheduler, and for job monitoring with the Spark and Dask user
+ interfaces. We configured the Spark driver to use 25~GB of memory, and
+ used the default configuration for the master.
+
 \subsection{Dataset}
@@
 We used BigBrain~\cite{Amunts:13}, a three-dimensional image of a human
-brain. The original data is stored in 125 blocks in the MINC~\cite{minc}
-HDF5-based format, available at
-\url{ftp://bigbrain.loris.ca/BigBrainRelease.2015/3D_Blocks/40um} at the
+brain with voxel intensities ranging from 0 to 65,535. The original data is
+stored in 125 blocks in the MINC~\cite{minc} HDF5-based format, available
+at \url{ftp://bigbrain.loris.ca/BigBrainRelease.2015/3D_Blocks/40um} at the
 resolution of \SI{40}{\micro\metre}. We converted the blocks into the
@@
 \subsection{Applications}
+
+We used three neuroimaging applications to evaluate the engines in
+different conditions. The first two ones, incrementation and histogram, are
+simple synthetic benchmarks representing basic map-only and map-reduce applications.
+The third one is a more realistic application representative of popular
+BIDS applications~\cite{gorgolewski2017bids}.
+
 %% Incrementation
-\subsubsection{\textbf{Incrementation}}
-As our first experiment to benchmark the frameworks, we used a simple incrementation
-pipeline (see Algorithm~\ref{alg:incrementation}). It reads blocks from the BigBrain
-image, increments the intensity value of each voxel by 1 for a specified amount of
-iterations, and writes the result to an NFS as a NifTI image. Incrementing the data
-reduce the caching effect. This experiment allows us to study the behavior of the
-frameworks when all inputs are processed independently; i.e.\ each task in the graph
-only depends on the previous one (see Figure~\ref{fig:tg-inc}). This mimics the
-behavior of analyzing multiple independent subjects in parallel.
-
-This experiment refers to the implementation of the Spark pipelines purposed
-in Ref.~\cite{Hayot-Sasson:17}.
+\subsubsection{Incrementation}
+We used the simple image incrementation pipeline used
+in~\cite{hayot2019performance} (see Algorithm~\ref{alg:incrementation}).
+The application reads blocks of the BigBrain image from the shared file
+system, increments the intensity value of each voxel by 1 to avoid caching
+effects, sleeps for a configurable amount of time to emulate a more complex
+processing, repeats this process for a specified amount of iterations, and
+finally writes the result as a NifTI image back to the shared file system.
+This application allows us to study the behavior of the engines when all
+inputs are processed independently, in a map-only way (see
+Figure~\ref{fig:tg-inc}). This mimics the behavior of analyzing multiple
+independent subjects in parallel.
 
 \begin{algorithm}[!b]
-    \caption{Incrementation}\label{alg:incrementation}
+    \caption{Incrementation (adapted from~\cite{hayot2019performance})}\label{alg:incrementation}
     \begin{algorithmic}
@@
     angle=-90]{images/incrementation-task-graph.png}
-    \caption{Task graph for Incrementation}\label{fig:tg-inc}
+    \caption{Task graph for Incrementation with 5 iterations, represented
+    for 3 BigBrain blocks. Circles represent the incrementation and sleep
+    function while rectangles represent image blocks.}\label{fig:tg-inc}
 \end{figure}
 
-\subsubsection{\textbf{Histogram}}
-Our second experiment consists to calculate the histogram of the intensity of the
-voxel in BigBrain (see Algorithm~\ref{alg:histogram}). It reads the images from the
-NFS, calculate individual image frequency, aggregate the frequency of the images
-together, and write the result to a file. In this experiment, the result depends on
-all the input (see Fig.~\ref{fig:tg-histo}). This allows the study of the behavior of
-the frameworks when tasks are dependent. Also, this application requires data
-shuffling thus inter-worker communication. We decided to omit Dask Futures as it does
-not have any value over Dask Delayed in this application.
+\subsubsection{Histogram}
+
+ Our second application calculates the histogram of the BigBrain image (see
+ Algorithm~\ref{alg:histogram}). It reads the image blocks from the shared
+ file system, calculates intensity frequencies, aggregates the frequencies
+ across the blocks, and finally writes the resulting histogram to the
+ shared file system as a single \TG{X}KB file. This application implements
+ a typical map-reduce pattern, where the final result is obtained from all
+ the input blocks (see Fig.~\ref{fig:tg-histo}). This application requires
+ data shuffling thus inter-worker communication. The total amount of shuffled
+ data is however limited to \TG{X}MB as it only consists of image
+ histograms.
+
+ \TG{Add links to the Github repo for applications and any other relevant file.}
 
@@
 
-\subsection{Experiment}
-The applications are benchmarked according to up to four parameters
-(Table~\ref{tab:param}):
+\subsubsection{BIDS app example}
 
-\textbf{Number of workers:} Varying the numbers of workers helps in assessing the
-scaling of the frameworks in a cluster. Also, it helps in determining the level of
-the parallelism of the frameworks.
+We used the \href{https://github.com/BIDS-Apps/example}{BIDS app example}
+application available on GitHub. This application operates in a map-reduce
+way. The map phase, also called participant analysis, extracts the brain
+tissues of a subject's 3D MRI using the popular FMRIB Software Library
+(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the
+resulting image (\TG{X}MB on average) to the shared file system. The reduce
+phase, also called group analysis, computes the volume of each brain and
+returns the average volume, shuffling a total of \TG{X}GB image data.
 
-\textbf{Number of blocks:} Evaluating the frameworks on a different amount of blocks
-allows us to understand their behavior with different task size and IO requirements.
+While Incrementation and Histogram were implemented directly in Python,
+BIDS app example requires an external library and is distributed as a
+Docker container image (bids/base\_fsl on DockerHub). To avoid remote file
+transfers from DockerHub, we preloaded our OpenStack instances with the
+Docker image, and started the Docker engine at startup. \TG{explain run.py modifications if relevant.}
 
-\textbf{Number of iterations:} We change the number of iterations for the
-incrementation~\ref{alg:incrementation} to study the effect of the job length.
+\subsection{Experiments}
 
-\textbf{Sleep delay:} We modify the sleep delay in the
-incrementation~\ref{alg:incrementation} to understand the behavior of different task
-lengths.
+We varied four parameters in our experiments, as shown in
+Table~\ref{tab:param}. We varied (1) the number of workers to assess the
+scalability of the engines, (2) the number of BigBrain blocks in
+Incrementation and Histogram to measure the effect of different IO patterns
+and parallelization degrees, and (3) the number of iterations and sleep
+delay in Incrementation to evaluate the effect of job length.
 
-The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
-seconds sleep delay.
+% The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
+% seconds sleep delay.
 
@@
     \end{table}
-
-\subsection{Infrastructure}
-All experiments were performed on Arbutus Cloud Compute Canda. Each node has CentOS
-7.5.1804 as the base operating system and have the Linux kernel version
-3.10.0\-862.11.6.el7.x86\_64. The nodes have an Intel Xeon Gold 6130 processor with
-30GB 2666 MHz memory. The network has a 10GB/s bandwidth. The filesystem is 2TB NFS
-v4.
 
@@
 \end{figure}
-
+\TG{We decided to omit Dask Futures as it does not have any value over Dask
+Delayed in this application.}
 %%% BIDS EXAMPLE %%%

### 2019-06-27 histogram/spark_histogram.py
Fixed histogram folding

@@ -52,7 +52,7 @@ if __name__ == "__main__":
 
     histogram = partial_histogram.fold(
         {}, lambda x, y: combine_histogram(x, y, args=args, start=start)
-    ).collect()
+    )
 
     start_time = time() - start
 

### 2019-06-27 histogram/Histogram.py
Seperated flatten from calculate_histogram

@@ -4,10 +4,27 @@ from time import time
 from utils import benchmark
 
 
-def calculate_histogram(arr, *, args, start, filename):
+def flatten(arr, *, args, start, filename):
     start_time = time() - start
 
-    arr = arr.flatten('F')
+    arr = arr.flatten("F")
+
+    end_time = time() - start
+
+    if args.benchmark:
+        benchmark(
+            start_time,
+            end_time,
+            filename,
+            args.output_dir,
+            args.experiment,
+            "flatten",
+        )
+    return filename, arr
+
+
+def calculate_histogram(arr, *, args, start, filename):
+    start_time = time() - start
 
     histogram = defaultdict(int)
     for x in arr:

### 2019-06-27 histogram/dask-bag_histogram.py

@@ -41,12 +41,14 @@ if __name__ == "__main__":
     client.upload_file("/nfs/paper-big-data-engines/utils.py")
     client.upload_file("/nfs/paper-big-data-engines/histogram/Histogram.py")
     from utils import benchmark, crawl_dir, read_img
-    from Histogram import calculate_histogram, combine_histogram
+    from Histogram import calculate_histogram, combine_histogram, flatten
 
     # Read images
     paths = crawl_dir(os.path.abspath(args.bb_dir))
     paths = db.from_sequence(paths, npartitions=len(paths))
     img = paths.map(lambda p: read_img(p, start=start, args=args))
+
+    img = img.map(lambda x: flatten(x[1], start=start, args=args, filename=x[0]))
 
     partial_histogram = img.map(
         lambda x: calculate_histogram(x[1], args=args, start=start, filename=x[0])

### 2019-06-27 histogram/dask-delayed_histogram.py

@@ -42,7 +42,7 @@ if __name__ == "__main__":
     client.upload_file("/nfs/paper-big-data-engines/utils.py")
     client.upload_file("/nfs/paper-big-data-engines/histogram/Histogram.py")
     from utils import benchmark, crawl_dir, read_img
-    from Histogram import calculate_histogram, combine_histogram
+    from Histogram import calculate_histogram, combine_histogram, flatten
 
     # Read images
     paths = crawl_dir(os.path.abspath(args.bb_dir))
@@ -50,6 +50,8 @@ if __name__ == "__main__":
     partial_histogram = []
     for path in paths:
         img = dask.delayed(read_img)(path, start=start, args=args)
+
+        img = dask.delayed(flatten)(img[1], start=start, args=args, filename=img[0])
 
         partial_histogram.append(
             dask.delayed(calculate_histogram)(

### 2019-06-27 histogram/spark_histogram.py
Seperated flatten from calculate_histogram

@@ -37,7 +37,7 @@ if __name__ == "__main__":
     sc.addFile("/nfs/paper-big-data-engines/utils.py")
     sc.addFile("/nfs/paper-big-data-engines/histogram/Histogram.py")
     from utils import benchmark, crawl_dir, read_img
-    from Histogram import calculate_histogram, combine_histogram
+    from Histogram import calculate_histogram, combine_histogram, flatten
 
     print("Connected")
 
@@ -45,6 +45,10 @@ if __name__ == "__main__":
     paths = crawl_dir(os.path.abspath(args.bb_dir))
     paths = sc.parallelize(paths, len(paths))
     img_rdd = paths.map(lambda p: read_img(p, start=start, args=args))
+
+    img_rdd = img_rdd.map(
+        lambda x: flatten(x[1], start=start, args=args, filename=x[0])
+    )
 
     partial_histogram = img_rdd.map(
         lambda x: calculate_histogram(x[1], args=args, start=start, filename=x[0])

### 2019-06-28 latex/paper.tex
Modified the engine sub-section

@@
 
-\newcommand{\TG}[1]{\color{blue}From Tristan: #1 \color{black}}
+\newcommand{\TG}[1]{\color{cyan}From Tristan: #1 \color{black}}
+\newcommand{\MD}[1]{\color{magenta}From Mathieu: #1 \color{black}}
 
@@
 
-\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data
-engine. It provides in-memory computing, data locality, and lazy
-evaluation. Spark's main data structure, the Resilient Distributed Dataset
-(RDD)~\cite{RDD}, is a fault-tolerant and parallel collection of data
-elements. 
-Dataframes, Spark's other popular API, are an overlay of RDDs with similar 
-performance features. 
-Spark's core is written in Scala, but it also has an R and a Python API
-called PySpark. We used PySpark, given the popularity of Python in
-neuroscience, although the serialization cost of Python objects to Scala
-might be a substantial source of overhead. We used Spark's default
+\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine. It
+provides in-memory computing, data locality, and lazy evaluation. Spark's main data
+structure, the Resilient Distributed Dataset (RDD)~\cite{RDD}, is a fault-tolerant
+and a parallel collection of data elements. Dataframes, Spark's other popular API, is
+an overlay of RDDs with similar performance features. Spark's core is written in
+Scala, but it also has a Java, an R, and a Python API called PySpark. \MD{Sentence is
+ambiguous? could think that all three APIs are named PySpark} We used PySpark, given
+the popularity of Python in neuroscience, although the serialization cost of Python
+objects to Scala might be a substantial source of overhead. We used Spark's default
 scheduler, Spark standalone, although YARN~\cite{vavilapalli2013apache} and
-Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler, a
-process called the \emph{master} coordinates resources provisioned by
-\emph{workers} on the cluster. The application is submitted to the
-\emph{driver} that in turn requests workers to the master and dispatches
-tasks to them. The Spark standalone scheduler uses a LIFO
-(Last-In-First-Out) task scheduling policy \TG{we should describe it
-briefly and refer to the paper}. The Spark standalone scheduler has two
-execution modes: client mode, where the driver runs in a dedicated  process
-on the cluster, and cluster mode, where the driver runs in a worker. We
-used client mode as PySpark is not available in cluster mode. We used
-Apache Spark v2.4.0.
+Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler,
+a process called the \emph{master} coordinates resources provisioned by
+\emph{workers} on the cluster. The application is submitted to the \emph{driver} that
+in turn requests workers to the master and dispatches jobs to them. A job is divided
+into stages to be processed onto the workers. Each stage's operation is represented
+as a high-level task in the computation graph. The Spark standalone scheduler uses a
+FIFO (First-In-First-Out) job scheduling policy. Processing jobs in order helps in
+increasing the page hit rate. As our applications only have one job, the LIFO
+(Last-In-First-Out) policy used for the stages is more pertinent \MD{Find better
+phrasing} \TG{refer to the paper (\MD{Not sure which one})}. The Spark standalone
+scheduler has two execution modes: client mode, where the driver runs in a dedicated
+process on the cluster, and cluster mode, where the driver runs in a worker. We used
+client mode as PySpark is not available in cluster mode. We used Apache Spark v2.4.0.
 
-\subsubsection{Dask} Dask is a Big Data engine that is becoming
-increasingly popular in the scientific Python ecosystem. \TG{A few more
-introductory sentences should be added, following Spark's model.}
-\TG{Explain workers}
-Dask represents application tasks as Dask graphs, produced by a
-user-facing API and executed by a scheduler. We used the \href{https://distributed.dask.org/en/latest/index.html}{Dask
-Distributed} scheduler.
-Dask Distributed also has a LIFO policy. In theory, it should speed-up the
-pipeline execution by finishing a branch of task computations before
-starting new ones \TG{This is meant to reduce the memory footprint of the
-application}. \TG{Here we should add a few comparison sentences between
-Spark's and Dask's schedulers.} Dask has five main APIs and data structures: Bag,
-Array, Futures, Delayed and DataFrame. We used them all
-except DataFrame as it is not adapted to the type of data we
-used. All APIs provide in-memory computing, data locality, and lazy
-evaluation, however, none of them has fault tolerance \TG{Check fault
-tolerance in Dask. Fine-grained lineage vs Spark's coarse grained lineage}.
-We used Dask v1.1.4.
+\subsubsection{Dask} Dask is a Big Data engine that is becoming increasingly popular
+in the scientific Python ecosystem. Dask has five main APIs and data structures:
+Array, Bag, DataFrame, Delayed, and Futures. We used them all except DataFrame as it
+is not adapted to the type of data we used. Like Spark, all Dask APIs provide
+in-memory computing, data locality, lazy evaluation, and fault tolerance by mean of
+lineage. Dask's core is purely in Python and bypass most of the
+\href{https://docs.python.org/3/glossary.html#term-gil}{GIL} to provide parallelism.
+We used Dask due to the common use of the scientific Python ecosystem in
+neuroscience. Dask represents application tasks as Dask graphs, produced by a
+user-facing API and executed by a scheduler. In Dask, operations generate multiple
+tiny tasks in the computation graph allowing an easier representation of complex
+algorithms. We used the \href{https://distributed.dask.org/en/latest/index.html}{Dask
+Distributed} scheduler, although \href{https://github.com/dask/dask-yarn}{Dask-Yarn}
+and \href{https://github.com/mrocklin/dask-mesos}{Dask-Mesos} is also available. In
+the Dask Distributed scheduler, a process called \textit{dask-scheduler}
+administrates the resources provided by \textit{workers} in the cluster. The
+scheduler receives jobs from clients and assigns tasks to available workers. Like
+Spark scheduler, Dask Distributed also use a LIFO policy, it is meant to reduce the
+memory footprint of the application by finishing the computation of a branch in the
+Dask graph before starting new ones. We used Dask v1.1.4.
 
 In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
-collection of Python objects, similar to Spark's RDD. It offers a
-programming abstraction similar to the
-\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}.
-An \href{https://docs.dask.org/en/latest/array.html}{Array} is used for
-the processing of large arrays. It provides a distributed clone of the
-popular
-NumPy library. \href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary
-tasks that do not fit in the Array, DataFrame or Bag APIs. 
-Finally, \href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to
-Delayed as they support arbitrary tasks, but they operate in real-time rather
-than lazily.
+collection of Python objects, similar to Spark's RDD. It offers a programming
+abstraction similar to the \href{https://toolz.readthedocs.io/en/latest/}{PyToolz
+library}. An \href{https://docs.dask.org/en/latest/array.html}{Array} is used for the
+processing of large arrays. It provides a distributed clone of the popular NumPy
+library. A \href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a
+parallel composition of
+\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas
+Dataframes} used to process a large amount of tabular data.
+\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary tasks
+that do not fit in the Array, DataFrame or Bag APIs. Finally,
+\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to Delayed
+as they support arbitrary tasks, but they operate in real-time rather than lazily.
 
@@
  configured Spark to use 1 executor per worker and 25~GB of memory per
- executor, to leave 5~GB for overhead. \TG{Check spill to disk}
+ executor, to leave 5~GB for off-heap. \TG{Check spill to disk}
 
@@
 %% worker
-\subsection{Experiment 1: Number of workers}
+\subsection{Incrementation: Number of workers}
 Figure~\ref{fig:inc_ms_worker} shows the makespan for the incrementation algorithm on
@@
 % blocks
-\subsection{Experiment 1: Number of blocks}
+\subsection{Incrementation: Number of blocks}
 In Figure~\ref{fig:inc_block}, Spark is not compared for 30 blocks. This is because
@@
 %% Iterations
-\subsection{Experiment 1: Number of iterations}
+\subsection{Incrementation: Number of iterations}
 Figure~\ref{fig:inc_ms_itr} shows the makespan of the application for the various
@@
 %% Sleep time
-\subsection{Experiment 1: Sleep delay}
+\subsection{Incrementation: Sleep delay}
 Figure~\ref{fig:inc_ms_sleep} shows the makespan of the incrementation algorithm for

### 2019-06-28 latex/paper.tex
some comments on methods

@@
 \newcommand{\MD}[1]{\color{magenta}From Mathieu: #1 \color{black}}
+\newcommand{\VHS}[1]{\color{green}From Valerie: #1 \color{black}}
 
@@
 structure, the Resilient Distributed Dataset (RDD)~\cite{RDD}, is a fault-tolerant
-and a parallel collection of data elements. Dataframes, Spark's other popular API, is
-an overlay of RDDs with similar performance features. Spark's core is written in
+and a parallel collection of data elements. Dataframes, Spark's API for tabular data, is
+an overlay of RDDs with similar performance features \VHS{It's probably fine, but the SparkSQL Engine available for Scala and Java do bring an added peformance value}.\VHS{Spark has another API (Datasets).
+i know you used 'popular' for DataFrames, but it's not like Spark has a plethora of APIs.
+I think the APIs should be mentioned after the API languages such that you only list the APIs relevant to Python (RDD and Dataframes)} Spark's core is written in
 Scala, but it also has a Java, an R, and a Python API called PySpark. \MD{Sentence is
@@
 \emph{workers} on the cluster. The application is submitted to the \emph{driver} that
-in turn requests workers to the master and dispatches jobs to them. A job is divided
+in turn requests workers to the master and dispatches jobs to them\VHS{maybe say tasks instead of jobs, otherwise it sounds like 1 job/worker}. A job is divided
 into stages to be processed onto the workers. Each stage's operation is represented
@@
 (Last-In-First-Out) policy used for the stages is more pertinent \MD{Find better
-phrasing} \TG{refer to the paper (\MD{Not sure which one})}. The Spark standalone
+phrasing} \TG{refer to the paper (\MD{Not sure which one})} \VHS{I wouldn't explicitly state LIFO here.
+It's a bit complicated to be frank.
+Basically both Spark and Dask opt to complete a chain of dependencies rather than start a bunch of ``new chains''.
+And in this context, LIFO is great for the page hit rate, so you might want to reword the prior sentence as well}.
+The Spark standalone
 scheduler has two execution modes: client mode, where the driver runs in a dedicated
 process on the cluster, and cluster mode, where the driver runs in a worker. We used
-client mode as PySpark is not available in cluster mode. We used Apache Spark v2.4.0.
+client mode as cluster mode in not available in PySpark. We used Apache Spark v2.4.0.
 
@@
 in the scientific Python ecosystem. Dask has five main APIs and data structures:
-Array, Bag, DataFrame, Delayed, and Futures. We used them all except DataFrame as it
-is not adapted to the type of data we used. Like Spark, all Dask APIs provide
+Array, Bag, DataFrame, Delayed, and Futures. All APIs were used in our experiments,
+with the exception of Dask Dataframes, which is intended for use with tabular data.
+Like Spark, all Dask APIs provide
 in-memory computing, data locality, lazy evaluation, and fault tolerance by mean of
-lineage. Dask's core is purely in Python and bypass most of the
+lineage. Dask's core is purely in Python and bypasses most of the
 \href{https://docs.python.org/3/glossary.html#term-gil}{GIL} to provide parallelism.
 We used Dask due to the common use of the scientific Python ecosystem in
-neuroscience. Dask represents application tasks as Dask graphs, produced by a
+neuroscience\VHS{If you add Dask Arrays to your experiments here, i'd mention that their
+APIs might be particularly well suited for the processing of high resolution neuroimaging data}. Dask represents application tasks as Dask graphs, produced by a
 user-facing API and executed by a scheduler. In Dask, operations generate multiple
@@
 as they support arbitrary tasks, but they operate in real-time rather than lazily.
+\VHS{maybe mention which one use threads and which use processes}
 

### 2019-06-28 latex/paper.tex
fixed typo
Worker on infracstructure sectcion

@@
 \usepackage{times}
-\usepackage{siunitx}
+\usepackage[binary-units=true]{siunitx}
 \usepackage{latexsym}
@@
 structure, the Resilient Distributed Dataset (RDD)~\cite{RDD}, is a fault-tolerant
-and a parallel collection of data elements. Dataframes, Spark's API for tabular data, is
-an overlay of RDDs with similar performance features \VHS{It's probably fine, but the SparkSQL Engine available for Scala and Java do bring an added peformance value}.\VHS{Spark has another API (Datasets).
-i know you used 'popular' for DataFrames, but it's not like Spark has a plethora of APIs.
-I think the APIs should be mentioned after the API languages such that you only list the APIs relevant to Python (RDD and Dataframes)} Spark's core is written in
-Scala, but it also has a Java, an R, and a Python API called PySpark. \MD{Sentence is
-ambiguous? could think that all three APIs are named PySpark} We used PySpark, given
-the popularity of Python in neuroscience, although the serialization cost of Python
-objects to Scala might be a substantial source of overhead. We used Spark's default
-scheduler, Spark standalone, although YARN~\cite{vavilapalli2013apache} and
+and a parallel collection of data elements. Dataframes, Spark's API for tabular data,
+is an overlay of RDDs with similar performance features \VHS{It's probably fine, but
+the SparkSQL Engine available for Scala and Java do bring an added peformance
+value}.\VHS{Spark has another API (Datasets). i know you used 'popular' for
+DataFrames, but it's not like Spark has a plethora of APIs. I think the APIs should
+be mentioned after the API languages such that you only list the APIs relevant to
+Python (RDD and Dataframes)} Spark's core is written in Scala, but it also has a
+Java, an R, and a Python API called PySpark. \MD{Sentence is ambiguous? could think
+that all three APIs are named PySpark} We used PySpark, given the popularity of
+Python in neuroscience, although the serialization cost of Python objects to Scala
+might be a substantial source of overhead. We used Spark's default scheduler, Spark
+standalone, although YARN~\cite{vavilapalli2013apache} and
 Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler,
@@
 \emph{workers} on the cluster. The application is submitted to the \emph{driver} that
-in turn requests workers to the master and dispatches jobs to them\VHS{maybe say tasks instead of jobs, otherwise it sounds like 1 job/worker}. A job is divided
-into stages to be processed onto the workers. Each stage's operation is represented
-as a high-level task in the computation graph. The Spark standalone scheduler uses a
-FIFO (First-In-First-Out) job scheduling policy. Processing jobs in order helps in
-increasing the page hit rate. As our applications only have one job, the LIFO
-(Last-In-First-Out) policy used for the stages is more pertinent \MD{Find better
-phrasing} \TG{refer to the paper (\MD{Not sure which one})} \VHS{I wouldn't explicitly state LIFO here.
-It's a bit complicated to be frank.
-Basically both Spark and Dask opt to complete a chain of dependencies rather than start a bunch of ``new chains''.
-And in this context, LIFO is great for the page hit rate, so you might want to reword the prior sentence as well}.
-The Spark standalone
+in turn requests workers to the master and dispatches tasks to them. A job is divided
+into stages to be excuted in a different process onto the workers. Each stage's
+operation is represented as a high-level task in the computation graph. The Spark
+standalone scheduler uses a FIFO (First-In-First-Out) job scheduling policy.
+Processing jobs in order helps in increasing the page hit rate. As our applications
+only have one job, the LIFO (Last-In-First-Out) policy used for the stages is more
+pertinent \MD{Find better phrasing} \TG{refer to the paper (\MD{Not sure which one})}
+\VHS{I wouldn't explicitly state LIFO here. It's a bit complicated to be frank.
+Basically both Spark and Dask opt to complete a chain of dependencies rather than
+start a bunch of ``new chains''. And in this context, LIFO is great for the page hit
+rate, so you might want to reword the prior sentence as well}. The Spark standalone
 scheduler has two execution modes: client mode, where the driver runs in a dedicated
 process on the cluster, and cluster mode, where the driver runs in a worker. We used
-client mode as cluster mode in not available in PySpark. We used Apache Spark v2.4.0.
+client mode as cluster mode is not available in PySpark. We used Apache Spark v2.4.0.
 
@@
 In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
-collection of Python objects, similar to Spark's RDD. It offers a programming
-abstraction similar to the \href{https://toolz.readthedocs.io/en/latest/}{PyToolz
-library}. An \href{https://docs.dask.org/en/latest/array.html}{Array} is used for the
-processing of large arrays. It provides a distributed clone of the popular NumPy
-library. A \href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a
-parallel composition of
+collection of Python objects, similar to Spark's RDD. It use multi-process to enable
+parallelism. It offers a programming abstraction similar to the
+\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}. An
+\href{https://docs.dask.org/en/latest/array.html}{Array} is used for the processing
+of large arrays. It provides a distributed clone of the popular NumPy library. A
+\href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a parallel
+composition of
 \href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas
@@
 
- Cloud instances hosted a single Dask or Spark worker, configured to use 8
- threads. We used Dask's default configuration that uses all the available
- memory on the instance. \TG{add a sentence on memory usage policy for Dask, say it's on NFS} We
- configured Spark to use 1 executor per worker and 25~GB of memory per
- executor, to leave 5~GB for off-heap. \TG{Check spill to disk}
+ Cloud instances hosted a single Dask or Spark worker, configured to use 8 CPUs. We
+ used Dask's default configuration that uses all the available memory on the
+ instance. Its default heuristic is to: target a 60\% memory load, spill to disk at
+ 70\%, pause the worker at 80\%, and terminate the worker at 95\%. We configured
+ Spark to use 1 executor per worker and 25~GB of memory per executor, to leave 5~GB
+ for off-heap. We configured the Spark driver to use 25~GB of memory, and used the
+ default configuration for the master. We used the default confirguration for worker
+ memory management: at 60\% it spill data to disk, and 50\% of that amount is
+ reserved for storage and is immuned from eviction.
 
@@
  Dask scheduler, and for job monitoring with the Spark and Dask user
- interfaces. We configured the Spark driver to use 25~GB of memory, and
- used the default configuration for the master.
+ interfaces. For both Spark and Dask, spilled data is evicted to the NFS.
 
@@
 in neuroimaging. We left the NifTI blocks uncompressed, resulting in 
-a total data size of 76~GiB \TG{Check if GiB or GB}. 
+a total data size of \SI{81}{\giga\byte}. 
 To evaluate the effect of block size, we resplit these blocks into 30, 125 and 750 blocks of 
-2.5GB, 0.6GB, and
-0.1GB \TG{check sizes}, using the sam~\cite{sam} library.
+\SI{2.7}{\giga\byte}, \SI{0.648}{\giga\byte}, and
+\SI{0.108}{\giga\byte}, using the sam~\cite{sam} library.
 
@@
 subject on average).
+\MD{X, Y, Z values TBD after testing the makespan with different number of subjects.}
 

### 2019-06-28 latex/paper.tex
Worked on Applications section

@@
 
-We used three neuroimaging applications to evaluate the engines in
-different conditions. The first two ones, incrementation and histogram, are
-simple synthetic benchmarks representing basic map-only and map-reduce applications.
-The third one is a more realistic application representative of popular
-BIDS applications~\cite{gorgolewski2017bids}.
+We used three neuroimaging applications to evaluate the engines in different
+conditions. The first two ones, incrementation and histogram, are simple synthetic
+benchmarks representing basic map-only and map-reduce applications. The third one is
+a realistic application representative of popular BIDS
+applications~\cite{gorgolewski2017bids}.
 
@@
 \subsubsection{Incrementation}
-We used the simple image incrementation pipeline used
+We used an adaptation of the simple image incrementation pipeline used
 in~\cite{hayot2019performance} (see Algorithm~\ref{alg:incrementation}).
@@
     angle=-90]{images/incrementation-task-graph.png}
-    \caption{Task graph for Incrementation with 5 iterations, represented
-    for 3 BigBrain blocks. Circles represent the incrementation and sleep
-    function while rectangles represent image blocks.}\label{fig:tg-inc}
+    \caption{Task graph for Incrementation with 5 iterations and 3 BigBrain blocks.
+    Circles represent the incrementation and sleep function while rectangles
+    represent stages of the image blocks.}\label{fig:tg-inc}
 \end{figure}
@@
  Our second application calculates the histogram of the BigBrain image (see
- Algorithm~\ref{alg:histogram}). It reads the image blocks from the shared
- file system, calculates intensity frequencies, aggregates the frequencies
- across the blocks, and finally writes the resulting histogram to the
- shared file system as a single \TG{X}KB file. This application implements
- a typical map-reduce pattern, where the final result is obtained from all
- the input blocks (see Fig.~\ref{fig:tg-histo}). This application requires
- data shuffling thus inter-worker communication. The total amount of shuffled
- data is however limited to \TG{X}MB as it only consists of image
- histograms.
+ Algorithm~\ref{alg:histogram}). It reads the image blocks from the shared file
+ system, calculates intensity frequencies, aggregates the frequencies across the
+ blocks, and finally writes the resulting histogram to the shared file system as a
+ single \SI{766}{\kilo\byte} file. This application implements a typical map-reduce
+ pattern, where the final result is obtained from all the input blocks (see
+ Fig.~\ref{fig:tg-histo}). This application requires data shuffling thus inter-worker
+ communication. The total amount of shuffled data is however limited to
+ \SI{2.62}{\mega\byte} per blocks as it only consists of image histograms.
 
@@
     \includegraphics[width=0.16\textwidth, angle=-90]{images/histogram-task-graph.png}
-    \caption{Task graph for Histogram}\label{fig:tg-histo}
+    \caption{Task graph for Histogram with 3 BigBrain blocks. Circles represent the
+    functions while rectangles represent stages of the image blocks.
+    }\label{fig:tg-histo}
 \end{figure}
@@
 returns the average volume, shuffling a total of \TG{X}GB image data.
+\MD{Determine values when dataset is determined for the application.}
 
@@
 delay in Incrementation to evaluate the effect of job length.
+We run the experiments in a non-deterministic order and clear the cache of each
+worker before each benchmark.
 

### 2019-07-02 bidsApp-examples/Example.py
Implementation of Bids App
Modified crawler to analyze only one site

@@ -1,0 +1,64 @@
+from glob import glob
+import subprocess
+from time import time
+
+from utils import benchmark
+
+
+def subject_crawler(path):
+    return [
+        (site_dir, subj_dir.split("/")[-2].split("-")[-1])
+        for site_dir in glob(f"{path}")
+        for subj_dir in glob(f"{site_dir}/*/")
+    ]
+
+
+def run_participant(*, subject_dir, start, args, input_dir, output_dir):
+    start_time = time() - start
+
+    path = "../"
+    subprocess.run(
+        [
+            f"{path}/bids_example.img",
+            input_dir,
+            output_dir,
+            "participant",
+            "--participant_label",
+            subject_dir,
+        ],
+        shell=True,
+    )
+
+    end_time = time() - start
+
+    if args.benchmark:
+        benchmark(
+            start_time,
+            end_time,
+            subject_dir,
+            args.output_dir,
+            args.experiment,
+            run_participant.__name__,
+        )
+
+
+def run_group(*, start, args, file_, output_dir):
+    start_time = time() - start
+
+    path = "../"
+    subprocess.run(
+        [f"{path}/bids_example.img", "/nfs/bids-data/RawDataBIDS", output_dir, "group"],
+        shell=True,
+    )
+
+    end_time = time() - start
+
+    if args.benchmark:
+        benchmark(
+            start_time,
+            end_time,
+            file_,
+            args.output_dir,
+            args.experiment,
+            run_group.__name__,
+        )

### 2019-07-02 bidsApp-examples/spark.py

@@ -2,8 +2,10 @@ import argparse
 import sys
 from time import time
 
-import dask
-from dask.distributed import Client, fire_and_forget
+
+from pyspark import SparkConf, SparkContext
+
+from Example import run_group, run_participant, subject_crawler
 
 sys.path.append("/nfs/paper-big-data-engines")
 
@@ -33,30 +35,21 @@ if __name__ == "__main__":
     args = parser.parse_args()
 
     # Cluster scheduler
-    cluster = args.scheduler
-    client = Client(cluster)
-
-    print(client)
-    client.upload_file("/nfs/paper-big-data-engines/Example.py")
-    from Example import run_group, run_participant, subject_crawler
+    conf = SparkConf().setAppName("Spark Incrementation")
+    sc = SparkContext.getOrCreate(conf=conf)
+    print("Connected")
 
     # Retrieve all subject path
-    subjects = subject_crawler(args.bids_dir)
+    subjects_to_analyze = sc.parallelize(subject_crawler(args.bids_dir))
 
-    results = list()
-    for subject in subjects:
-        results.append(
-            dask.delayed(run_participant)(
-                subject_dir=subject[1],
-                start=start,
-                args=args,
-                input_dir=subject[0],
-                output_dir=args.output_dir,
-            )
+    subjects_to_analyze.map(
+        lambda x: run_participant(
+            subject_dir=x[1],
+            start=start,
+            args=args,
+            input_dir=x[0],
+            output_dir=args.output_dir,
         )
+    )
 
-    client.scatter(results)
-    futures = client.compute(results)
-    client.gather(futures)
-
-    dask.delayed(run_group)(start=start, args=args, output_dir=args.output_dir)
+    run_group(start=start, args=args, output_dir=args.output_dir)

### 2019-07-02 bidsApp-examples/dask-bag.py
Implementation of Bids App
Modified crawler to analyze only one site

@@ -2,8 +2,8 @@ import argparse
 import sys
 from time import time
 
-import dask
-from dask.distributed import Client, fire_and_forget
+import dask.bag as db
+from dask.distributed import Client
 
 sys.path.append("/nfs/paper-big-data-engines")
 
@@ -41,22 +41,20 @@ if __name__ == "__main__":
     from Example import run_group, run_participant, subject_crawler
 
     # Retrieve all subject path
-    subjects = subject_crawler(args.bids_dir)
+    subjects_to_analyze = db.from_sequence(subject_crawler(args.bids_dir))
 
-    results = list()
-    for subject in subjects:
-        results.append(
-            dask.delayed(run_participant)(
-                subject_dir=subject[1],
-                start=start,
-                args=args,
-                input_dir=subject[0],
-                output_dir=args.output_dir,
-            )
+    subjects_to_analyze.map(
+        lambda x: run_participant(
+            subject_dir=x[1],
+            start=start,
+            args=args,
+            input_dir=x[0],
+            output_dir=args.output_dir,
         )
+    )
 
-    client.scatter(results)
-    futures = client.compute(results)
-    client.gather(futures)
-
-    dask.delayed(run_group)(start=start, args=args, output_dir=args.output_dir)
+    run_group(
+        start=start,
+        args=args,
+        output_dir=args.output_dir,
+    )

### 2019-07-02 bidsApp-examples/dask-futures.py

@@ -42,11 +42,13 @@ if __name__ == "__main__":
 
     # Retrieve all subject path
     subjects = subject_crawler(args.bids_dir)
+    client.scatter(subjects)
 
     results = list()
     for subject in subjects:
         results.append(
-            dask.delayed(run_participant)(
+            client.submit(
+                run_participant,
                 subject_dir=subject[1],
                 start=start,
                 args=args,
@@ -55,8 +57,7 @@ if __name__ == "__main__":
             )
         )
 
-    client.scatter(results)
-    futures = client.compute(results)
-    client.gather(futures)
+    client.gather(results)
 
-    dask.delayed(run_group)(start=start, args=args, output_dir=args.output_dir)
+    group = client.submit(run_group, start=start, args=args, output_dir=args.output_dir)
+    client.gather(group)

### 2019-07-02 latex/paper.tex
some changes in intro

@@
 % Context
-With the increasing number of data available in neuroimaging,~\cite{ALFAROALMAGRO:18,
-UKBioBank:18} the processing of Big Data becomes critical. Frameworks like
-Nipype~\cite{Nipype:11} are usually used to create neuroimaging pipelines however it
-is worth considering the use of general purpose Big Data
-frameworks~\cite{Hayot-Sasson:17}. In that research, Spark~\cite{Spark:16} was a
-natural choice as it is the state-of-the-art framework for Big Data. Since our
-laboratory mostly works with Python, we want to know if Dask~\cite{Dask:15}, a Python
-written framework, could bring better or similar performances while facilitating
-usage.
+\VHS{it's not the the processing of BD becomes critical, it's that the results obtained
+from processing more data should be more reliable\dots and the data is there too}
+The recent rise in data sharing and improved data collection strategies
+have brought neuroimaging to the Big Data era ~\cite{ALFAROALMAGRO:18,
+UKBioBank:18}. Existing neuroimaging frameworks, such as Nipype~\cite{Nipype:11},
+are well suited for processing the standard compute-intensive neuroimaging pipelines,
+but lack incorporation Big Data strategies (i.e. in-memory computing,
+data locality and lazy-evaluation) to improve performance of the increasingly
+prevalent data-intensive pipelines. As was noted in ~\cite{Hayot-Sasson:17}, 
+in-memory computing, coupled with data locality, can bring significant performance
+improvements to data intensive neuroimaging pipelines. We extend this work by studying
+the differences between two Big Data frameworks, Dask~\cite{Dask:15} and Apache Spark~\cite{Spark:16},
+for their suitability in the processing neuroimaging pipelines.
 

### 2019-07-02 latex/paper.tex
Comments and fixes on III.A

--- 2019-06-28 latex/paper.tex
@@
 delay in Incrementation to evaluate the effect of job length.
-We run the experiments in a non-deterministic order and clear the cache of each
-worker before each benchmark.
+It should be noted that increasing the number of blocks or iterations also
+increases the total compute time of the application for a given sleep
+delay. To avoid any potential external bias such as background load on the
+network, we ran the experiments in a randomized order and cleared the page
+cache of each worker before each execution.
 
@@
 \subsection{Incrementation: Number of workers}
-Figure~\ref{fig:inc_ms_worker} shows the makespan for the incrementation algorithm on
-the y-axis and the number of workers on the x-axis; the color represents the
-different frameworks. Overall, there is no substantial makespan difference between
-the frameworks. Also, the increase in performance is not proportional to the number
-of workers; performance even decreases when we reach 8 workers.
+Figure~\ref{fig:inc_ms_worker} shows the makespan of the Incrementation
+application for different numbers of workers and for different engines.
+Overall, there is no substantial makespan difference between the engines.
+Dask seems to have a slight advantage over Spark, \TG{X (toutes les Dask vs Spark)} seconds on
+average, with Delayed and Futures being slightly better than Bags.
 
-In Figure~\ref{fig:inc_tt_worker}, the y-axis shows the total execution time of the
-incrementation application. The hatch indicates the total time spent for: read,
-compute, write, and overhead. Again, the color represents the frameworks. The
-computing time stays similar when the number of workers increases. However, the IO
-time and overhead increase proportionally to the number of workers.
+For all engines, the makespan is far from decreasing linearly with the
+number of workers. Note that there are 8 threads per worker. The makespan
+even increases between 4 and 8 workers. This is due to the high impact of
+data transfers and engine overhead on the application.
+
+Figure~\ref{fig:inc_tt_worker} shows the total execution time of the
+Incrementation application, broken down into data transfer (read and
+write), compute (sleep), and overhead time. As expected, the computing time
+stays similar when the number of workers increases. However, the data
+transfer time and overhead increase proportionally to the number of
+workers \TG{check if indeed proprotional, add regression slopes}.
+
+On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
+slightly lower than Dask's in particular as the number of workers increase.
+Within Dask, Delayed and Futures have a higher overhead than Bags. However,
+overhead differences are compensated by an increase of data transfer time,
+as a reduced overhead increases the concurrency between data transfers.
 
@@
     \end{subfigure}
-    \caption{125 blocks, 10 iterations, 4 sec.\ sleep delay}\label{fig:inc_worker}
+    \caption{125 blocks, 10 iterations, 4-s delay, 8 threads/worker}\label{fig:inc_worker}
 \end{figure}
@@
 \subsection{Incrementation: Number of blocks}
-In Figure~\ref{fig:inc_block}, Spark is not compared for 30 blocks. This is because
-Spark has a 2GB limitation in the task size it can compute.
 
-The Figure~\ref{fig:inc_ms_block} shows the makespan evolution when varying the
-number of blocks. The compute time increase when there are more blocks. This is
-because the sleep delay is constant throughout this experiment and there are more
-compute tasks when the number of blocks increases.
+Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when
+varying the number of image blocks for a constant BigBrain image size. We
+were not able to run Spark for 30 blocks due to its 2GB limitation in the
+task size it can compute. Once again, we do not observe any significant
+difference among the engines.
 
@@
     \end{subfigure}
-    \caption{10 iterations, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_block}
+    \caption{10 iterations, 4-s delay, 8 workers, 8 threads/worker}\label{fig:inc_block}
 \end{figure}
@@
 \section{Discussion}
+
+\TG{Talk about engine overhead: (1) where is it coming from? (2) 
+explain the impact on data transfers with a graph.}
+
 \subsection{IO Time}

### 2019-07-02 latex/paper.tex
Went through Section III

@@
 
-\title{A Comparative study of Dask and Spark for neuroimaging pipelines}
+\title{A Comparative study of Dask and Spark for data-intensive neuroimaging pipelines}
 
@@
 neuroimaging pipelines.
+
+\TG{Say that it is often said that Spark has a larger overhead but scales 
+worse than Dask. In the dicussion, get back to it and mention that (1) we don't
+always observe that, (2) for data-intensive applications, overheads don't really matter
+as reducing them increases data trasnfers.}
 
@@
 
+For each run, we measured the application makespan as well as the cumulative 
+data read, compute, data write, and engine overhead across all application tasks. 
+We measured the overhead by \TG{\ldots, maybe add a figure}.
+
+\TG{explain that number of iterations may impact futures.}
+
 % The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
@@
 application for different numbers of workers and for different engines. 
+The bars show the average makespan over 5 repetitions while the error bars 
+are the standard deviation \TG{check this}.
 Overall, there is no substantial makespan difference between the engines.
@@
 task size it can compute. Once again, we do not observe any significant
-difference among the engines.
+difference among the engines. For all engines, makespan variability
+increases with the number of blocks, however, engines scale very well in
+general.
 
 In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
-For 30 blocks the Dask Bag API has a much lower overhead time. This is because we
-only calculate the idle time of the used core. Dask Bag was only using one thread per
-block in comparison to Dask Delayed and Dask Futures which was offloading a block
-calculations on multiple threads of the same worker. Furthermore, the variance of the
-overhead increases proportionally to the number of blocks. Also, the IO time reduces
-with more blocks however the overhead time increase by a similar amount. This is not
-observed for 30 blocks as the workers are not used at full capacity; i.e.\ some
-threads are idle.
+For 30 blocks the Dask Bag API has a much lower overhead time. This is because Dask Bag was only using one thread per
+block in comparison to Dask Delayed and Dask Futures which were offloading a block
+calculations on multiple threads of the same worker.  Once again, the data transfer time reduces
+with more blocks however the overhead time increases by a similar amount. This is not
+observed for 30 blocks as the workers are not used at full capacity, i.e., some
+threads are idle. Finally, the variability of the
+overhead increases with the number of blocks, which explains the makespan
+variability mentioned previously.
 
@@
 \subsection{Incrementation: Number of iterations}
-Figure~\ref{fig:inc_ms_itr} shows the makespan of the application for the various
-number of iterations. Spark and the Dask Bag API are equivalent in terms of
-performance. The Dask Delayed API is slightly faster with few iterations (tasks) but
-it has a similar makespan to Spark and Dask Bag when the number of iterations
-increases. Overall, Dask Futures seems to be faster but the difference is not
-substantial.
+Figure~\ref{fig:inc_ms_itr} shows the makespan of the application while
+varying the number of iterations. Overall,  Spark and Dask APIs are once
+again equivalent, although Dask Delayed and Futures are slightly faster
+than Bags and Spark for 1 and 10 few iterations, and Fuures are faster than
+Delayed, Bags and Spark for 100 iterations. Differences remain minor
+though.
 
-In Figure~\ref{fig:inc_tt_itr}, the execution time of the function type is shown. The
-compute time increase proportionally to the number of iterations. This is because
-each computing task has a constant sleep delay. Also, the overhead increases when the
-number of iterations (tasks) increases while the IO time decreases; however this is
-not substantial.
+In Figure~\ref{fig:inc_tt_itr}, the total execution time breakdown is
+shown. We observe good scalability of all the engines with the number of
+iterations.
 
@@
     \end{subfigure}
-    \caption{125 blocks, 4 sec.\ sleep delay, 8 instances}
+    \caption{125 blocks, 4-s delay, 8 workers, 8 threads/worker}
 \end{figure}
@@
 \subsection{Incrementation: Sleep delay}
-Figure~\ref{fig:inc_ms_sleep} shows the makespan of the incrementation algorithm for
-different sleep delays. Spark is initially slower than the Dask APIs, however, it is
-faster when increasing the sleep delay. Also, within the Dask APIs, Dask Bag is
-slower than the other two but it is not considerable.
+Figure~\ref{fig:inc_ms_sleep} shows the makespan of the Incrementation
+application for different sleep delays. Overall, all engines again perform the same, and scale well with task duration.
+ Spark is initially slower than the
+Dask APIs, however, it is faster when increasing the sleep delay. Also,
+within the Dask APIs, Dask Bag is slower than the other two but it is not
+considerable.
 
-Figure~\ref{fig:inc_tt_sleep} shows the execution time for each type of function. On
-one hand, Spark has the smallest overhead. Table~\ref{tab:inc_sleep_overhead} shows
-that it is substantial; especially with a 64 seconds sleep delay. On the other hand,
-Spark has a high IO time. In contrast, Dask Delayed and Futures have the largest
-overhead time but the smallest IO time. Then, Dask Bag stands in the middle for both
-overhead and IO time.
+Figure~\ref{fig:inc_tt_sleep} shows the total execution time breakdown. On
+one hand, Spark has the smallest overhead. As previously observed,
+variations in overhead time are almost exactly compensated by variations in data
+transfer time.
 
-\begin{table}[!t]
-    \caption{Function time for the sleep experiment}
-    \begin{subtable}[b]{\columnwidth}
-        \renewcommand{\arraystretch}{1.3}
-        \caption{Overhead time in second}\label{tab:inc_sleep_overhead}
-        \centering
-        \begin{tabular}{lllll}
-        \hline
-                     & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
-        Spark        & 14992  & 17076  & 16069   & 9769    \\
-        Dask.Bag     & 20690  & 21388  & 17759   & 18720   \\
-        Dask.Delayed & 27481  & 26199  & 22080   & 24624   \\
-        Dask.Futures & 27069  & 27984  & 23538   & 24762   \\ \hline
-        \end{tabular}
-    \end{subtable}
-    \vskip 0.2cm
-    \begin{subtable}[b]{\columnwidth}
-        \renewcommand{\arraystretch}{1.3}
-        \caption{IO time in second}\label{tab:inc_sleep_io}
-        \centering
-        \begin{tabular}{lllll}
-        \hline
-                     & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
-        Spark        & 63622  & 57144  & 51070   & 57078   \\
-        Dask.Bag     & 54630  & 52312  & 53743   & 53612   \\
-        Dask.Delayed & 45239  & 44142  & 45133   & 46198   \\
-        Dask.Futures & 46537  & 43161  & 44749   & 45846   \\ \hline
-        \end{tabular}
-    \end{subtable}
-    \vspace{-3mm}
- \end{table}
+% \begin{table}[!t]
+%     \caption{Function time for the sleep experiment}
+%     \begin{subtable}[b]{\columnwidth}
+%         \renewcommand{\arraystretch}{1.3}
+%         \caption{Overhead time in second}\label{tab:inc_sleep_overhead}
+%         \centering
+%         \begin{tabular}{lllll}
+%         \hline
+%                      & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
+%         Spark        & 14992  & 17076  & 16069   & 9769    \\
+%         Dask.Bag     & 20690  & 21388  & 17759   & 18720   \\
+%         Dask.Delayed & 27481  & 26199  & 22080   & 24624   \\
+%         Dask.Futures & 27069  & 27984  & 23538   & 24762   \\ \hline
+%         \end{tabular}
+%     \end{subtable}
+%     \vskip 0.2cm
+%     \begin{subtable}[b]{\columnwidth}
+%         \renewcommand{\arraystretch}{1.3}
+%         \caption{IO time in second}\label{tab:inc_sleep_io}
+%         \centering
+%         \begin{tabular}{lllll}
+%         \hline
+%                      & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
+%         Spark        & 63622  & 57144  & 51070   & 57078   \\
+%         Dask.Bag     & 54630  & 52312  & 53743   & 53612   \\
+%         Dask.Delayed & 45239  & 44142  & 45133   & 46198   \\
+%         Dask.Futures & 46537  & 43161  & 44749   & 45846   \\ \hline
+%         \end{tabular}
+%     \end{subtable}
+%     \vspace{-3mm}
+%  \end{table}
 
@@
     \end{subfigure}
-    \caption{125 blocks, 10 iterations, 8 instances}
+    \caption{125 blocks, 10 iterations, 8 workers, 8 threads/worker}
 \end{figure}
@@
 %% Baseline gantt chart
-\subsection{Experimentation 1: Baseline timeline}
-Figure\ref{fig:inc_gantt} shows the tasks timeline for each worker thread. On the
-x-axis is the time in seconds and on the y-axis is the workers and their threads. The
-color represents the type of different tasks; note that the color differs from one
-graph to another. The first section of reading tasks is much larger than the
-following ones. This is because, initially, the NFS is saturated by the number of
-workers accessing it. Also, a substantial amount of overhead happens near the IO
-tasks. Moreover, for the computing tasks, there is more overhead for the latter tasks.
+\subsection{Incrementation: Gantt chart}
 
+Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine
+and API, for a baseline execution with 125 blocks, a 4-second sleep delay
+and 8 workers \TG{Make the colors uniform}. Gantt charts are structured in 
+batches of up to 64 read-compute-write concurrent sequences.
+File reads in the first batch are much longer than in the following ones:
+this is due to the high synchronization of data transfers that leads to a
+high saturation of the shared file system. We also note that overhead,
+represented in white, is concentrated around the data transfer tasks
+\TG{find out why}, and the computing tasks that run concurrently with data
+transfers. 
 % The frameworks have similar makespan however the time spend for each type of
@@
     \end{subfigure}
-    \caption{125 blocks, 4 sec.\ sleep delay, 8 instances}\label{fig:inc_gantt}
+    \caption{125 blocks, 4 sec.\ sleep delay, 8 workers}\label{fig:inc_gantt}
 \end{figure*}
@@
 %%% HISTOGRAM %%%
-\subsection{Experiment 2: Number of instances}
+\subsection{Experiment 2: Number of workers}
 
@@
 %%% BIDS EXAMPLE %%%
-\subsection{Experiment 3: Number of instances}
+\subsection{Experiment 3: Number of workers}
 
@@
 explain the impact on data transfers with a graph.}
+
+\TG{In the dicussion, get back to it and mention that (1) we don't
+always observe that, (2) for data-intensive applications, overheads don't really matter
+as reducing them increases data trasnfers.}
 

### 2019-07-03 bidsApp-examples/Example.py
Modified for one site

@@ -7,24 +7,27 @@ from utils import benchmark
 
 def subject_crawler(path):
     return [
-        (site_dir, subj_dir.split("/")[-2].split("-")[-1])
+        (site_dir, subj_id.split("/")[-2].split("-")[-1])
         for site_dir in glob(f"{path}")
-        for subj_dir in glob(f"{site_dir}/*/")
+        for subj_id in glob(f"{site_dir}/*/")
     ]
 
 
-def run_participant(*, subject_dir, start, args, input_dir, output_dir):
+def run_participant(*, subject_id, start, args):
     start_time = time() - start
 
-    path = "../"
     subprocess.run(
         [
-            f"{path}/bids_example.img",
-            input_dir,
-            output_dir,
-            "participant",
-            "--participant_label",
-            subject_dir,
+            "singularity",
+            "exec",
+            "-B",
+            "/nfs/singularity-image:/run",
+            "bids_example.simg",
+            "bash",
+            "/run/participant.sh",
+            args.bids_dir,
+            args.output_dir,
+            subject_id,
         ],
         shell=True,
     )
@@ -35,19 +38,29 @@ def run_participant(*, subject_dir, start, args, input_dir, output_dir):
         benchmark(
             start_time,
             end_time,
-            subject_dir,
+            subject_id,
             args.output_dir,
             args.experiment,
             run_participant.__name__,
         )
 
 
-def run_group(*, start, args, file_, output_dir):
+def run_group(*, start, args):
     start_time = time() - start
 
-    path = "../"
     subprocess.run(
-        [f"{path}/bids_example.img", "/nfs/bids-data/RawDataBIDS", output_dir, "group"],
+        [
+            "singularity",
+            "exec",
+            "-B",
+            "/nfs/singularity-image:/run",
+            "bids_example.simg",
+            "bash",
+            "/run/group.sh",
+            "/nfs/bids-data",
+            args.bids_dir,
+            args.output_dir,
+        ],
         shell=True,
     )
 
@@ -57,7 +70,7 @@ def run_group(*, start, args, file_, output_dir):
         benchmark(
             start_time,
             end_time,
-            file_,
+            "all_file",
             args.output_dir,
             args.experiment,
             run_group.__name__,

### 2019-07-03 bidsApp-examples/dask-futures.py

@@ -37,7 +37,7 @@ if __name__ == "__main__":
     client = Client(cluster)
 
     print(client)
-    client.upload_file("/nfs/paper-big-data-engines/Example.py")
+    client.upload_file("/nfs/paper-big-data-engines/bidsApp-examples/Example.py")
     from Example import run_group, run_participant, subject_crawler
 
     # Retrieve all subject path
@@ -48,16 +48,11 @@ if __name__ == "__main__":
     for subject in subjects:
         results.append(
             client.submit(
-                run_participant,
-                subject_dir=subject[1],
-                start=start,
-                args=args,
-                input_dir=subject[0],
-                output_dir=args.output_dir,
+                run_participant, subject_id=subject[1], start=start, args=args
             )
         )
 
     client.gather(results)
 
-    group = client.submit(run_group, start=start, args=args, output_dir=args.output_dir)
+    group = client.submit(run_group, start=start, args=args)
     client.gather(group)

### 2019-07-03 bidsApp-examples/spark.py

@@ -5,8 +5,6 @@ from time import time
 
 from pyspark import SparkConf, SparkContext
 
-from Example import run_group, run_participant, subject_crawler
-
 sys.path.append("/nfs/paper-big-data-engines")
 
 
@@ -39,17 +37,14 @@ if __name__ == "__main__":
     sc = SparkContext.getOrCreate(conf=conf)
     print("Connected")
 
+    sc.addFile("/nfs/paper-big-data-engines/bidsApp-examples/Example.py")
+    from Example import run_group, run_participant, subject_crawler
+
     # Retrieve all subject path
     subjects_to_analyze = sc.parallelize(subject_crawler(args.bids_dir))
 
     subjects_to_analyze.map(
-        lambda x: run_participant(
-            subject_dir=x[1],
-            start=start,
-            args=args,
-            input_dir=x[0],
-            output_dir=args.output_dir,
-        )
-    )
+        lambda x: run_participant(subject_id=x[1], start=start, args=args)
+    ).collect()
 
-    run_group(start=start, args=args, output_dir=args.output_dir)
+    run_group(start=start, args=args)

### 2019-07-03 experiments-script/run_experiment.py

@@ -3,15 +3,15 @@ from random import shuffle
 import json
 
 
-with open("experiment.json") as f_in:
+with open("/nfs/paper-big-data-engines/experiments-script/bids.json") as f_in:
     experiments = json.load(f_in)
     shuffle(experiments)
     for exp in experiments:
         experiment = exp["experiment"]
         filename = exp["file"]
-        iterations = str(exp["iterations"])
-        delay = str(exp["delay"])
-        chunks = str(exp["chunks"])
+        # iterations = str(exp["iterations"])
+        # delay = str(exp["delay"])
+        # chunks = str(exp["chunks"])
 
         subprocess.run(
             ["sh", "/nfs/paper-big-data-engines/experiments-script/clear-cache.sh"]
@@ -25,12 +25,12 @@ with open("experiment.json") as f_in:
                     "spark://192.168.73.10:7077",
                     "--executor-memory",
                     "25G",
-                    "/nfs/paper-big-data-engines/incrementation/" + filename,
-                    "/nfs/bb-" + chunks + "chunks",
-                    "/nfs/results",
+                    "--driver-memory",
+                    "25G",
+                    "/nfs/paper-big-data-engines/bidsApp_examples/" + filename,
+                    "/nfs/bids-data/RawDataBIDS/NYU_2",
+                    "/nfs/bids/outputs",
                     experiment,
-                    iterations,
-                    delay,
                     "--benchmark",
                 ]
             )
@@ -38,13 +38,11 @@ with open("experiment.json") as f_in:
             subprocess.run(
                 [
                     "python",
-                    "/nfs/paper-big-data-engines/incrementation/" + filename,
+                    "/nfs/paper-big-data-engines/bidsApp_examples/" + filename,
                     "192.168.73.10:8786",
-                    "/nfs/bb-" + chunks + "chunks",
-                    "/nfs/results",
+                    "/nfs/bids-data/RawDataBIDS/NYU_2",
+                    "/nfs/bids/outputs",
                     experiment,
-                    iterations,
-                    delay,
                     "--benchmark",
                 ]
             )

### 2019-07-03 bidsApp-examples/dask-bag.py
Modified for one site

@@ -37,24 +37,14 @@ if __name__ == "__main__":
     client = Client(cluster)
 
     print(client)
-    client.upload_file("/nfs/paper-big-data-engines/Example.py")
+    client.upload_file("/nfs/paper-big-data-engines/bidsApp-examples/Example.py")
     from Example import run_group, run_participant, subject_crawler
 
     # Retrieve all subject path
     subjects_to_analyze = db.from_sequence(subject_crawler(args.bids_dir))
 
     subjects_to_analyze.map(
-        lambda x: run_participant(
-            subject_dir=x[1],
-            start=start,
-            args=args,
-            input_dir=x[0],
-            output_dir=args.output_dir,
-        )
-    )
+        lambda x: run_participant(subject_id=x[1], start=start, args=args)
+    ).compute()
 
-    run_group(
-        start=start,
-        args=args,
-        output_dir=args.output_dir,
-    )
+    run_group(start=start, args=args)

### 2019-07-03 bidsApp-examples/dask-delayed.py

@@ -37,7 +37,7 @@ if __name__ == "__main__":
     client = Client(cluster)
 
     print(client)
-    client.upload_file("/nfs/paper-big-data-engines/Example.py")
+    client.upload_file("/nfs/paper-big-data-engines/bidsApp-examples/Example.py")
     from Example import run_group, run_participant, subject_crawler
 
     # Retrieve all subject path
@@ -46,17 +46,14 @@ if __name__ == "__main__":
     results = list()
     for subject in subjects:
         results.append(
-            dask.delayed(run_participant)(
-                subject_dir=subject[1],
-                start=start,
-                args=args,
-                input_dir=subject[0],
-                output_dir=args.output_dir,
-            )
+            dask.delayed(run_participant)(subject_id=subject[1], start=start, args=args)
         )
 
     client.scatter(results)
     futures = client.compute(results)
     client.gather(futures)
 
-    dask.delayed(run_group)(start=start, args=args, output_dir=args.output_dir)
+    group = dask.delayed(run_group)(start=start, args=args)
+    client.scatter(group)
+    futures = client.compute(group)
+    client.gather(futures)

### 2019-07-11 latex/paper.tex
Wroked on comments

@@
  We used Compute Canada's
- \href{https://docs.computecanada.ca/wiki/Cloud_resources}{Arbutus Cloud}
- operated by the \href{https://www.westgrid.ca}{WestGrid} regional
- organization at the University of Victoria, and running OpenStack
- \TG{version?}. We used c8-30gb-186 cloud instances with 8 VCPUs, an Intel
- Xeon Gold 6130 processor, 30~GB of RAM at 2666~MHz, 20~GB of mounted
- storage, and a base image running CentOS 7.5.1804 with  Linux kernel
- version
- 3.10.0\-862.11.6.el7.x86\_64. Instances are connected by a 10Gb/s Ethernet network.
+ \href{https://docs.computecanada.ca/wiki/Cloud_resources}{Arbutus Cloud} operated by
+ the \href{https://www.westgrid.ca}{WestGrid} regional organization at the University
+ of Victoria, and running OpenStack \TG{version?}. We used c8-30gb-186 cloud
+ instances with 8 VCPUs, an Intel Xeon Gold 6130 processor, \SI{30}{\giga\byte} of
+ RAM at \SI{2666}{\mega\hertz}, \SI{20}{\giga\byte} of mounted storage, and a base
+ image running CentOS 7.5.1804 with Linux kernel version
+ 3.10.0\-862.11.6.el7.x86\_64. Instances are connected by a
+ \SI{10}{\giga\bit/\second} Ethernet network.
 
@@
  70\%, pause the worker at 80\%, and terminate the worker at 95\%. We configured
- Spark to use 1 executor per worker and 25~GB of memory per executor, to leave 5~GB
+ Spark to use 1 executor per worker and \SI{25}{\giga\byte} of memory per executor, to leave 5~GB
  for off-heap. We configured the Spark driver to use 25~GB of memory, and used the
@@
 
- One cloud instance did not host any worker, and had a 2~TB disk volume
+ One cloud instance did not host any worker, and had a \SI{2}{\tera\byte} disk volume
  shared with the other instances using the Network File System (NFS) v4.
@@
 We also used the dataset provided by the Consortium for Reliability and
-Reproducibility
-(\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) as
-available on
-\href{http://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
-entire dataset is 408.4~GB, containing anatomical, diffusion and functional
-images of 1,397 subjects acquired in 29 sites. Among these, we only used
-data from \TG{X} subjects, representing \TG{Y}GB overall (\TG{Z}GB by
-subject on average).
-\MD{X, Y, Z values TBD after testing the makespan with different number of subjects.}
+Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) as
+available on \href{http://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
+entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and
+functional images of 1,397 subjects acquired in 29 sites. We used all 3491 anatomical
+images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} by image on
+average).
 
@@
 a realistic application representative of popular BIDS
-applications~\cite{gorgolewski2017bids}.
+applications~\cite{gorgolewski2017bids}. All script used for our experiment is availble
+GitHub at \url{https://github.com/big-data-lab-team/paper-big-data-engines}
 
@@
 
- \TG{Add links to the Github repo for applications and any other relevant file.}
-
 \begin{algorithm}[!t]
@@
 
-We used the \href{https://github.com/BIDS-Apps/example}{BIDS app example}
-application available on GitHub. This application operates in a map-reduce
-way. The map phase, also called participant analysis, extracts the brain
-tissues of a subject's 3D MRI using the popular FMRIB Software Library
-(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the
-resulting image (\TG{X}MB on average) to the shared file system. The reduce
-phase, also called group analysis, computes the volume of each brain and
-returns the average volume, shuffling a total of \TG{X}GB image data.
-\MD{Determine values when dataset is determined for the application.}
+We used the \href{https://github.com/BIDS-Apps/example}{BIDS app example} application
+available on GitHub. This application operates in a map-reduce way. The map phase,
+also called participant analysis, extracts the brain tissues of a subject's 3D MRI
+using the popular FMRIB Software Library
+(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the resulting image
+(\SI{2.47}{\mega\byte} on average) to the shared file system. The reduce phase, also
+called group analysis, computes the volume of each brain and returns the average
+volume, shuffling a total of \SI{8.6}{\giga\byte} image data.
 
@@
 transfers from DockerHub, we preloaded our OpenStack instances with the
-Docker image, and started the Docker engine at startup. \TG{explain run.py modifications if relevant.}
-
+Docker image, and started the Docker engine at startup.
 \subsection{Experiments}
@@
     \# of iterations    & 1, 10, 100     & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\
-    Sleep delay {[}s{]} & 1, 4, 16, 64   & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\ \hline
+    Sleep delay {[}\SI{}{\second}{]} & 1, 4, 16, 64   & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\ \hline
     \end{tabular*}
@@
 \subsection{Incrementation: Number of workers}
-Figure~\ref{fig:inc_ms_worker} shows the makespan of the Incrementation
-application for different numbers of workers and for different engines. 
-The bars show the average makespan over 5 repetitions while the error bars 
-are the standard deviation \TG{check this}.
-Overall, there is no substantial makespan difference between the engines.
-Dask seems to have a slight advantage over Spark, \TG{X (toutes les Dask vs Spark)} seconds on
-average, with Delayed and Futures being slightly better than Bags.
+Figure~\ref{fig:inc_ms_worker} shows the makespan of the Incrementation application
+for different numbers of workers and for different engines. The bars show the average
+makespan over 5 repetitions while the error bars are the standard deviation. Overall,
+there is no substantial makespan difference between the engines. Dask seems to have a
+slight advantage over Spark, \SI{83.61}{\second} on average,
+with Delayed and Futures being slightly better than Bags.
 
@@
     \end{subfigure}
-    \caption{125 blocks, 10 iterations, 4-s delay, 8 threads/worker}\label{fig:inc_worker}
+    \caption{125 blocks, 10 iterations, \SI{4}{\second} delay, 8 threads/worker}\label{fig:inc_worker}
 \end{figure}
@@
 
-Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when
-varying the number of image blocks for a constant BigBrain image size. We
-were not able to run Spark for 30 blocks due to its 2GB limitation in the
-task size it can compute. Once again, we do not observe any significant
-difference among the engines. For all engines, makespan variability
-increases with the number of blocks, however, engines scale very well in
-general.
+Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when varying the
+number of image blocks for a constant BigBrain image size. We were not able to run
+Spark for 30 blocks due to its \SI{2}{\giga\byte} limitation in the task size it can
+compute. Once again, we do not observe any significant difference among the engines.
+For all engines, makespan variability increases with the number of blocks, however,
+engines scale very well in general.
 
 In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
-For 30 blocks the Dask Bag API has a much lower overhead time. This is because Dask Bag was only using one thread per
-block in comparison to Dask Delayed and Dask Futures which were offloading a block
-calculations on multiple threads of the same worker.  Once again, the data transfer time reduces
-with more blocks however the overhead time increases by a similar amount. This is not
-observed for 30 blocks as the workers are not used at full capacity, i.e., some
-threads are idle. Finally, the variability of the
-overhead increases with the number of blocks, which explains the makespan
-variability mentioned previously.
+For 30 blocks the Dask Bag API has a much lower overhead time. This is because Dask
+Bag was only using one thread per block in comparison to Dask Delayed and Dask
+Futures which were offloading a block calculations on multiple threads of the same
+worker. Once again, the data transfer time reduces with more blocks however the
+overhead time increases by a similar amount. This is not observed for 30 blocks as
+the workers are not used at full capacity, i.e., some threads are idle. Finally, the
+variability of the overhead increases with the number of blocks, which explains the
+makespan variability mentioned previously.
 
@@
     \end{subfigure}
-    \caption{10 iterations, 4-s delay, 8 workers, 8 threads/worker}\label{fig:inc_block}
+    \caption{10 iterations, \SI{4}{\second} delay, 8 workers, 8 threads/worker}\label{fig:inc_block}
 \end{figure}
@@
     \end{subfigure}
-    \caption{125 blocks, 4-s delay, 8 workers, 8 threads/worker}
+    \caption{125 blocks, \SI{4}{\second} delay, 8 workers, 8 threads/worker}
 \end{figure}
@@
 
-Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine
-and API, for a baseline execution with 125 blocks, a 4-second sleep delay
-and 8 workers \TG{Make the colors uniform}. Gantt charts are structured in 
-batches of up to 64 read-compute-write concurrent sequences.
-File reads in the first batch are much longer than in the following ones:
-this is due to the high synchronization of data transfers that leads to a
-high saturation of the shared file system. We also note that overhead,
-represented in white, is concentrated around the data transfer tasks
-\TG{find out why}, and the computing tasks that run concurrently with data
-transfers. 
+Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine and API,
+for a baseline execution with 125 blocks, a \SI{4}{\second} sleep delay and 8
+workers. Gantt charts are structured in batches of up to 64 read-compute-write
+concurrent sequences. File reads in the first batch are much longer than in the
+following ones: this is due to the high synchronization of data transfers that leads
+to a high saturation of the shared file system. We also note that overhead,
+represented in white, is concentrated around the data transfer tasks \TG{find out
+why}, and the computing tasks that run concurrently with data transfers. 
 % The frameworks have similar makespan however the time spend for each type of
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/spark_inc_baseline_gantt.png}
+        \includegraphics[clip,width=\columnwidth]{images/spark_inc_gantt.png}
         \caption{Spark execution timeline}\label{fig:inc_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/Dask_bag_inc_baseline_gantt.png}%
+        \includegraphics[clip,width=\columnwidth]{images/bag_inc_gantt.png}%
         \caption{Dask Bag execution timeline}\label{fig:inc_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/Dask_delayed_inc_baseline_gantt.png}%
+        \includegraphics[clip,width=\columnwidth]{images/delayed_inc_gantt.png}%
         \caption{Dask Delayed execution timeline}\label{fig:inc_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/Dask_futures_inc_baseline_gantt.png}%
+        \includegraphics[clip,width=\columnwidth]{images/futures_inc_gantt.png}%
         \caption{Dask Futures execution timeline}\label{fig:inc_dask_futures_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_instance.png}%
+        \includegraphics[clip,width=\columnwidth]{images/histo_worker.png}%
         \caption{Histogram makespan}\label{fig:histo_ms_worker}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_idle_instances.png}%
+        \includegraphics[clip,width=\columnwidth]{images/histo_idle_worker.png}%
         \caption{Histogram total time}\label{fig:histo_tt_worker}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_splits.png}%
+        \includegraphics[clip,width=\columnwidth]{images/histo_block.png}%
         \caption{Histogram makespan}\label{fig:histo_ms_block}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_idle_splits.png}%
+        \includegraphics[clip,width=\columnwidth]{images/histo_idle_block.png}%
         \caption{Histogram total time}\label{fig:histo_tt_block}
@@
 
-\subsection{IO Time}
-
-\subsection{Serialization}
-We think that serialization could have an important effect on the makespan of the
-application. This could potentially slow down the application when a lot of tasks are
-scheduled.
-
-\subsection{NFS}
-The NFS seems to be a source of the bottleneck. We think that exploring another file
-system, like Lustre, could give us insight into the actual effect of the NFS on the
-application makespan. We also think that the NFS caching could play an effect on the
-task of different lengths.
-
-\subsection{Scheduler}
-The Dask scheduler seems to have some bizarre behavior. It seems like it waits to
-schedule tasks. It could also be due to tasks being scheduled on other workers
-thus requiring data to be sent over the network; which would explain the stall on the
-worker. More work would be needed to investigate that issue.
-
-\subsection{Caching}
-We think that caching plays an effect on the results. As seen in the Gantt chart
-previously, some of the read tasks are significantly shorter than others while all
-blocks are of equal size.
+% \subsection{IO Time}
+% 
+% 
+% \subsection{Serialization}
+% We think that serialization could have an important effect on the makespan of the
+% application. This could potentially slow down the application when a lot of tasks are
+% scheduled.
+% 
+% \subsection{NFS}
+% The NFS seems to be a source of the bottleneck. We think that exploring another file
+% system, like Lustre, could give us insight into the actual effect of the NFS on the
+% application makespan. We also think that the NFS caching could play an effect on the
+% task of different lengths.
+% 
+% \subsection{Scheduler}
+% The Dask scheduler seems to have some bizarre behavior. It seems like it waits to
+% schedule tasks. It could also be due to tasks being scheduled on other workers
+% thus requiring data to be sent over the network; which would explain the stall on the
+% worker. More work would be needed to investigate that issue.
+% 
+% 
+% \subsection{Caching}
+% We think that caching plays an effect on the results. As seen in the Gantt chart
+% previously, some of the read tasks are significantly shorter than others while all
+% blocks are of equal size.
 

### 2019-07-11 latex/paper.tex
Fix plot y-axis

@@
         \includegraphics[clip,width=\columnwidth]{images/spark_inc_gantt.png}
-        \caption{Spark execution timeline}\label{fig:inc_spark_gantt}
+        \caption{Spark}\label{fig:inc_spark_gantt}
     \end{subfigure}
@@
         \includegraphics[clip,width=\columnwidth]{images/bag_inc_gantt.png}%
-        \caption{Dask Bag execution timeline}\label{fig:inc_dask_bag_gantt}
+        \caption{Dask Bag}\label{fig:inc_dask_bag_gantt}
     \end{subfigure}
@@
         \includegraphics[clip,width=\columnwidth]{images/delayed_inc_gantt.png}%
-        \caption{Dask Delayed execution timeline}\label{fig:inc_dask_delayed_gantt}
+        \caption{Dask Delayed}\label{fig:inc_dask_delayed_gantt}
     \end{subfigure}
@@
         \includegraphics[clip,width=\columnwidth]{images/futures_inc_gantt.png}%
-        \caption{Dask Futures execution timeline}\label{fig:inc_dask_futures_gantt}
+        \caption{Dask Futures}\label{fig:inc_dask_futures_gantt}
     \end{subfigure}
-    \caption{125 blocks, 4 sec.\ sleep delay, 8 workers}\label{fig:inc_gantt}
+    \caption{Gantt chart with 125 blocks, \SI{4}{\second} delay, 8 workers}\label{fig:inc_gantt}
 \end{figure*}
@@
 %%% HISTOGRAM %%%
-\subsection{Experiment 2: Number of workers}
+\subsection{Histogram: Number of workers}
 
@@
 
-\subsection{Experiment 2: Number of blocks}
+\subsection{Histogram: Number of blocks}
 
@@
 Delayed in this application.}
+
 %%% BIDS EXAMPLE %%%
-\subsection{Experiment 3: Number of workers}
+\subsection{BidsApp example: Number of workers}
+
+\begin{figure}[!t]
+    \centering
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/bids_worker.png}%
+        \caption{Histogram makespan}\label{fig:bids_ms_worker}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/bids_idle_worker.png}%
+        \caption{Histogram total time}\label{fig:bids_tt_worker}
+    \end{subfigure}
+    \caption{Variation of the amount of worker, 8 threads/worker}
+\end{figure}
+
+%% Baseline gantt chart
+\subsection{BidsApp example: Gantt chart}
+
+% Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine and API,
+% for a baseline execution with 125 blocks, a \SI{4}{\second} sleep delay and 8
+% workers. Gantt charts are structured in batches of up to 64 read-compute-write
+% concurrent sequences. File reads in the first batch are much longer than in the
+% following ones: this is due to the high synchronization of data transfers that leads
+% to a high saturation of the shared file system. We also note that overhead,
+% represented in white, is concentrated around the data transfer tasks \TG{find out
+% why}, and the computing tasks that run concurrently with data transfers. 
+% The frameworks have similar makespan however the time spend for each type of
+% task differs considerably (see Table~\ref{tab:inc_base}). Spark tends to spend more
+% time than other frameworks for IO however it has a much lower overhead. Dask Delayed
+% and Futures have the lowest IO time but a much higher overhead. Dask Bag stands in
+% the middle for both IO and overhead. Compute time is approximately the same for all
+% frameworks. Apart from the first IO section, there is a large amount of ovehead
+% everytime time IO is performed. This is true for all frameworks.
+
+\begin{figure*}[!htb]
+    \centering
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/spark_bids_gantt.png}
+        \caption{Spark}\label{fig:bids_spark_gantt}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/bag_bids_gantt.png}%
+        \caption{Dask Bag}\label{fig:bids_dask_bag_gantt}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/delayed_bids_gantt.png}%
+        \caption{Dask Delayed}\label{fig:bids_dask_delayed_gantt}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/futures_bids_gantt.png}%
+        \caption{Dask Futures}\label{fig:bids_dask_futures_gantt}
+    \end{subfigure}
+    \caption{Gantt chart with 8 workers}\label{fig:bids_gantt}
+\end{figure*}
 

### 2019-07-11 latex/paper.tex
Worked on BidsApp example restuls section

@@
     \end{subfigure}
-    \caption{125 blocks, 10 iterations, \SI{4}{\second} delay, 8 threads/worker}\label{fig:inc_worker}
+    \caption{125 blocks, 10 iterations, \SI{4}{\second} delay, 8
+    threads/worker}\label{fig:inc_worker}
 \end{figure}
@@
     \end{subfigure}
-    \caption{10 iterations, \SI{4}{\second} delay, 8 workers, 8 threads/worker}\label{fig:inc_block}
+    \caption{10 iterations, \SI{4}{\second} delay, 8 workers, 8
+    threads/worker}\label{fig:inc_block}
 \end{figure}
@@
 \subsection{Incrementation: Sleep delay}
-Figure~\ref{fig:inc_ms_sleep} shows the makespan of the Incrementation
-application for different sleep delays. Overall, all engines again perform the same, and scale well with task duration.
- Spark is initially slower than the
-Dask APIs, however, it is faster when increasing the sleep delay. Also,
-within the Dask APIs, Dask Bag is slower than the other two but it is not
-considerable.
+Figure~\ref{fig:inc_ms_sleep} shows the makespan of the Incrementation application
+for different sleep delays. Overall, all engines again perform the same, and scale
+well with task duration. Spark is initially slower than the Dask APIs, however, it is
+faster when increasing the sleep delay. Also, within the Dask APIs, Dask Bag is
+slower than the other two but it is not considerable.
 
@@
 
-Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine and API,
-for a baseline execution with 125 blocks, a \SI{4}{\second} sleep delay and 8
-workers. Gantt charts are structured in batches of up to 64 read-compute-write
-concurrent sequences. File reads in the first batch are much longer than in the
-following ones: this is due to the high synchronization of data transfers that leads
-to a high saturation of the shared file system. We also note that overhead,
-represented in white, is concentrated around the data transfer tasks \TG{find out
-why}, and the computing tasks that run concurrently with data transfers. 
+Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine and API.
+Gantt charts are structured in batches of up to 64 read-compute-write concurrent
+sequences. File reads in the first batch are much longer than in the following ones:
+this is due to the high synchronization of data transfers that leads to a high
+saturation of the shared file system. We also note that overhead, represented in
+white, is concentrated around the data transfer tasks \TG{find out why}, and the
+computing tasks that run concurrently with data transfers. 
 % The frameworks have similar makespan however the time spend for each type of
@@
     \end{subfigure}
-    \caption{Gantt chart with 125 blocks, \SI{4}{\second} delay, 8 workers}\label{fig:inc_gantt}
+    \caption{Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second} delay, 8
+    workers, 8 threads/worker}\label{fig:inc_gantt}
 \end{figure*}
@@
     \end{subfigure}
-    \caption{125 blocks}
+    \caption{125 blocks, 8 threads/worker}
 \end{figure}
@@
     \end{subfigure}
-    \caption{8 workers}
+    \caption{8 workers, 8 threads/worker}
 \end{figure}
@@
 \subsection{BidsApp example: Number of workers}
+Figure~\ref{fig:bids_ms_worker} shows the makespan of the application while varying
+the amount of worker. Overall, there is no substantial difference between the
+engines however Futures are always slightly faster. The performance difference between
+Futures and the other APIs increase with the number of workers. The makespan scales
+well with the increase in workers however the scaling reduces considerably when
+reaching 8 workers. This due to the IO requirement increasing with the number of
+workers, since more tasks are read concurrently, and the application becomes more data
+intensive.
+
+In Figure~\ref{fig:bids_tt_worker}, the total execution time of each function is
+shown. \MD{Maybe mention why Spark overhead is so big.} Overall, the computation time
+for the group analysis is equal for all engines and numbers of workers. Bag and
+Delayed takes much longer to compute the participant analysis with 8 workers,
+otherwise, the makespan is similar for all engines and numbers of workers. The overhead
+increases as the number of workers increases. \MD{Check regression}
+
+\MD{Depending on how the overhead is calculated spark overhead is either minimal or
+gigantic. It is hard to determine which method should be used to calculate the
+overhead for this application. For this application, since Spark has 2 job this
+causes Spark to have different worker process for the different job and thus some job
+have overlaping process ID while other are uniques (See
+Figure~\ref{fig:bids_spark_gantt}).}
 
@@
 \subsection{BidsApp example: Gantt chart}
-
-% Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine and API,
-% for a baseline execution with 125 blocks, a \SI{4}{\second} sleep delay and 8
-% workers. Gantt charts are structured in batches of up to 64 read-compute-write
-% concurrent sequences. File reads in the first batch are much longer than in the
-% following ones: this is due to the high synchronization of data transfers that leads
-% to a high saturation of the shared file system. We also note that overhead,
-% represented in white, is concentrated around the data transfer tasks \TG{find out
-% why}, and the computing tasks that run concurrently with data transfers. 
-% The frameworks have similar makespan however the time spend for each type of
-% task differs considerably (see Table~\ref{tab:inc_base}). Spark tends to spend more
-% time than other frameworks for IO however it has a much lower overhead. Dask Delayed
-% and Futures have the lowest IO time but a much higher overhead. Dask Bag stands in
-% the middle for both IO and overhead. Compute time is approximately the same for all
-% frameworks. Apart from the first IO section, there is a large amount of ovehead
-% everytime time IO is performed. This is true for all frameworks.
+Figure~\ref{fig:bids_gantt} shows the Gantt charts obtained for each engine and API.
+The Gantt charts are structured in three main parts: participant analysis (orange),
+group analysis (blue), and the transition in between.\MD{rephrase} The Gantt chart
+for Spark seems to have more workers however it does not. This is due to the
+computation of the participant and group analysis on two different jobs with their
+worker processes; usually with different process id. The tasks differ considerably in
+length. This is due to the high IO requirement of the application. During the
+participant analysis phase, Spark and Futures have less overhead than Delayed and
+Bag. At the end of the participant analysis phase, Futures and Delayed has less
+overhead than Bag while Spark has way more than Dask engine. The group analysis is
+similar for all engines and APIs.
 
@@
 
-Mathieu Dugr\'e was funded by a Undergraduate Summer Research Assistant
-award from the National Science and Engineering Research Council. We warmly
-thank Compute Canada and \TG{Westgrid?} for providing the cloud
-infrastructure used in these experiments, and the McGill Center for
-Integrative Neurocience for giving us access to their cloud allocation.
+Mathieu Dugr\'e was funded by a Undergraduate Student Research Assistant award from
+the National Science and Engineering Research Council of Canada. We warmly thank
+Compute Canada and \TG{Westgrid?} for providing the cloud infrastructure used in
+these experiments, and the McGill Center for Integrative Neurocience for giving us
+access to their cloud allocation.
 

### 2019-07-12 results/histogram/data-1/merger.py
Added data

@@ -1,19 +1,0 @@
-import subprocess
-import os
-
-directories = []
-for root, sub, filename in os.walk('.'):
-    if root != '.' and root[2:][0] != '.':
-        directories.append(root)
-        print('Crawling through ', root)
-
-processes = []
-for directory in directories:
-    cmd = 'cat {}/* >> /nfs/results/merged-node-1/results-{}.csv'\
-        .format(directory, directory[2:])
-    print('Merging results for', directory)
-    p = subprocess.Popen(cmd, shell=True)
-    processes.append(p)
-
-for p in processes:
-    p.wait()

### 2019-07-15 latex/paper.tex
Made correction to comment + Modified Bids App to talk about singularity

@@
 
-\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine. It
-provides in-memory computing, data locality, and lazy evaluation. Spark's main data
-structure, the Resilient Distributed Dataset (RDD)~\cite{RDD}, is a fault-tolerant
-and a parallel collection of data elements. Dataframes, Spark's API for tabular data,
-is an overlay of RDDs with similar performance features \VHS{It's probably fine, but
-the SparkSQL Engine available for Scala and Java do bring an added peformance
-value}.\VHS{Spark has another API (Datasets). i know you used 'popular' for
-DataFrames, but it's not like Spark has a plethora of APIs. I think the APIs should
-be mentioned after the API languages such that you only list the APIs relevant to
-Python (RDD and Dataframes)} Spark's core is written in Scala, but it also has a
-Java, an R, and a Python API called PySpark. \MD{Sentence is ambiguous? could think
-that all three APIs are named PySpark} We used PySpark, given the popularity of
-Python in neuroscience, although the serialization cost of Python objects to Scala
-might be a substantial source of overhead. We used Spark's default scheduler, Spark
-standalone, although YARN~\cite{vavilapalli2013apache} and
-Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler,
-a process called the \emph{master} coordinates resources provisioned by
-\emph{workers} on the cluster. The application is submitted to the \emph{driver} that
-in turn requests workers to the master and dispatches tasks to them. A job is divided
-into stages to be excuted in a different process onto the workers. Each stage's
-operation is represented as a high-level task in the computation graph. The Spark
-standalone scheduler uses a FIFO (First-In-First-Out) job scheduling policy.
-Processing jobs in order helps in increasing the page hit rate. As our applications
-only have one job, the LIFO (Last-In-First-Out) policy used for the stages is more
-pertinent \MD{Find better phrasing} \TG{refer to the paper (\MD{Not sure which one})}
-\VHS{I wouldn't explicitly state LIFO here. It's a bit complicated to be frank.
-Basically both Spark and Dask opt to complete a chain of dependencies rather than
-start a bunch of ``new chains''. And in this context, LIFO is great for the page hit
-rate, so you might want to reword the prior sentence as well}. The Spark standalone
-scheduler has two execution modes: client mode, where the driver runs in a dedicated
-process on the cluster, and cluster mode, where the driver runs in a worker. We used
-client mode as cluster mode is not available in PySpark. We used Apache Spark v2.4.0.
+\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine.
+Spark's core is written in Scala, but it also has a Java, an R, and a Python API
+called PySpark. We used PySpark, given the popularity of Python in neuroscience,
+although the serialization cost of Python objects to Scala might be a substantial
+source of overhead. Spark provides in-memory computing, data locality, and lazy
+evaluation. PySpark's fundamental data structure, the Resilient Distributed Dataset
+(RDD)~\cite{RDD}, is a fault-tolerant and a parallel collection of data elements.
+Dataframes, PySpark's API for tabular data, is an overlay of RDDs with similar
+performance features. We used Spark's default scheduler, Spark standalone, although
+YARN~\cite{vavilapalli2013apache} and Mesos~\cite{hindman2011mesos} are also
+available. In the Spark standalone scheduler, a process called the \emph{master}
+coordinates resources provisioned by \emph{worker} processes on the cluster. The
+application is submitted to the \emph{driver} that in turn requests workers to the
+master and dispatches tasks to them. A job is divided into stages to be excuted in a
+different process onto the workers. Each stage's operation is represented as a
+high-level task in the computation graph. The Spark standalone scheduler uses a FIFO
+(First-In-First-Out) job scheduling policy. Processing jobs in order helps in
+increasing the page hit rate. As our applications only have one job, the LIFO
+(Last-In-First-Out) policy used for the stages is more pertinent \MD{Find better
+phrasing} \TG{refer to the paper (\MD{Not sure which one})} \VHS{I wouldn't
+explicitly state LIFO here. It's a bit complicated to be frank. Basically both Spark
+and Dask opt to complete a chain of dependencies rather than start a bunch of ``new
+chains''. And in this context, LIFO is great for the page hit rate, so you might want
+to reword the prior sentence as well}. The Spark standalone scheduler has two
+execution modes: client mode, where the driver runs in a dedicated process on the
+cluster, and cluster mode, where the driver runs in a worker. We used client mode as
+cluster mode is not available in PySpark. We used Apache Spark v2.4.0.
 
@@
 as they support arbitrary tasks, but they operate in real-time rather than lazily.
-\VHS{maybe mention which one use threads and which use processes}
 
@@
 
-While Incrementation and Histogram were implemented directly in Python,
-BIDS app example requires an external library and is distributed as a
-Docker container image (bids/base\_fsl on DockerHub). To avoid remote file
-transfers from DockerHub, we preloaded our OpenStack instances with the
-Docker image, and started the Docker engine at startup.
+While Incrementation and Histogram were implemented directly in Python, BIDS app
+example requires an external library and is distributed as a Docker container image
+(bids/base\_fsl on DockerHub). We converted the Docker image to a Singularity image,
+given its popularity for HPC and reproducibility, using
+\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}.
+The image was preloaded on the nfs and singularity version 3.2.1\-1.el7 was installed
+on all instances. Due to a bug in the conversion from Docker to Singularity, we had to
+export the path for fsl.
+
 \subsection{Experiments}
@@
 data read, compute, data write, and engine overhead across all application tasks. 
-We measured the overhead by \TG{\ldots, maybe add a figure}.
+We measured the overhead by \TG{\ldots, maybe add a figure}. \MD{Done but might need
+adjustment depending on how the overhead is calculate (idle vs wasted)}
 

### 2019-07-16 latex/paper.tex
typos

@@
 engines however Futures are always slightly faster. The performance difference between
-Futures and the other APIs increase with the number of workers. The makespan scales
+Futures and the other APIs increases with the number of workers. The makespan scales
 well with the increase in workers however the scaling reduces considerably when
-reaching 8 workers. This due to the IO requirement increasing with the number of
-workers, since more tasks are read concurrently, and the application becomes more data
-intensive.
+reaching 8 workers. This is due to the IO requirement increasing with the number of
+workers, since more tasks are read concurrently, and the application becomes more
+data-intensive.
 

### 2019-07-22 latex/paper.tex
Added university name in author section

@@
 Department of Computer Science and Software Engineering\\
-Montr\'eal, Qu\'ebec, Canada
+Concordia University, Montr\'eal, Qu\'ebec, Canada
 }

### 2019-07-22 latex/paper.tex
Updated figures and started working on histogram section

@@
 \subsection{Histogram: Number of workers}
+Figure~\ref{fig:histo_ms_worker} shows the makespan of the application for various
+amounts of workers. Spark is tremendiously faster than Dask APIs. The difference
+narrows as the number of workers increases. Between Dask APIs there is no substantial
+difference, however, Bag is slightly faster on average.
+
+From Figure~\ref{fig:histo_tt_worker} the total time spend in each function is shown.
+The computation time is significantly larger in Dask than Spark. This is potentially
+due to Python's GIL preventing Dask to parallelize the computation on multiple thread.
+Overall, the IO and overhead is comparable for all engines.
 
@@
 \subsection{Histogram: Number of blocks}
+Figure~\ref{fig:histo_ms_block}
 
@@
         \includegraphics[clip,width=\columnwidth]{images/bids_worker.png}%
-        \caption{Histogram makespan}\label{fig:bids_ms_worker}
+        \caption{BIDS App example makespan}\label{fig:bids_ms_worker}
     \end{subfigure}
@@
         \includegraphics[clip,width=\columnwidth]{images/bids_idle_worker.png}%
-        \caption{Histogram total time}\label{fig:bids_tt_worker}
+        \caption{BIDS App example total time}\label{fig:bids_tt_worker}
     \end{subfigure}
@@
 
-\begin{figure*}[!htb]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/spark_bids_gantt.png}
-        \caption{Spark}\label{fig:bids_spark_gantt}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/bag_bids_gantt.png}%
-        \caption{Dask Bag}\label{fig:bids_dask_bag_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/delayed_bids_gantt.png}%
-        \caption{Dask Delayed}\label{fig:bids_dask_delayed_gantt}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/futures_bids_gantt.png}%
-        \caption{Dask Futures}\label{fig:bids_dask_futures_gantt}
-    \end{subfigure}
-    \caption{Gantt chart with 8 workers}\label{fig:bids_gantt}
-\end{figure*}
+% \begin{figure*}[!htb]
+%     \centering
+%     \begin{subfigure}[b]{\columnwidth}
+%         \includegraphics[clip,width=\columnwidth]{images/spark_bids_gantt.png}
+%         \caption{Spark}\label{fig:bids_spark_gantt}
+%     \end{subfigure}
+%     \hfill
+%     \begin{subfigure}[b]{\columnwidth}
+%         \includegraphics[clip,width=\columnwidth]{images/bag_bids_gantt.png}%
+%         \caption{Dask Bag}\label{fig:bids_dask_bag_gantt}
+%     \end{subfigure}
+%     \\
+%     \begin{subfigure}[b]{\columnwidth}
+%         \includegraphics[clip,width=\columnwidth]{images/delayed_bids_gantt.png}%
+%         \caption{Dask Delayed}\label{fig:bids_dask_delayed_gantt}
+%     \end{subfigure}
+%     \hfill
+%     \begin{subfigure}[b]{\columnwidth}
+%         \includegraphics[clip,width=\columnwidth]{images/futures_bids_gantt.png}%
+%         \caption{Dask Futures}\label{fig:bids_dask_futures_gantt}
+%     \end{subfigure}
+%     \caption{Gantt chart with 8 workers}\label{fig:bids_gantt}
+% \end{figure*}
 

### 2019-07-24 latex/paper.tex
Edited the introduction

--- 2019-07-16 latex/paper.tex
@@
 
-\title{A Comparative study of Dask and Spark for data-intensive neuroimaging pipelines}
+\title{A performance comparison of Dask and Spark for data-intensive neuroimaging pipelines}
 
@@
 % Context
-\VHS{it's not the the processing of BD becomes critical, it's that the results obtained
-from processing more data should be more reliable\dots and the data is there too}
 The recent rise in data sharing and improved data collection strategies
 have brought neuroimaging to the Big Data era ~\cite{ALFAROALMAGRO:18,
-UKBioBank:18}. Existing neuroimaging frameworks, such as Nipype~\cite{Nipype:11},
-are well suited for processing the standard compute-intensive neuroimaging pipelines,
-but lack incorporation Big Data strategies (i.e. in-memory computing,
-data locality and lazy-evaluation) to improve performance of the increasingly
-prevalent data-intensive pipelines. As was noted in ~\cite{Hayot-Sasson:17}, 
-in-memory computing, coupled with data locality, can bring significant performance
-improvements to data intensive neuroimaging pipelines. We extend this work by studying
-the differences between two Big Data frameworks, Dask~\cite{Dask:15} and Apache Spark~\cite{Spark:16},
-for their suitability in the processing neuroimaging pipelines.
+UKBioBank:18}. Existing neuroimaging workflow engines, such as
+Nipype~\cite{Nipype:11}, are well suited for processing the standard
+compute-intensive neuroimaging pipelines, but lack incorporated Big Data
+strategies (i.e. in-memory computing, data locality and lazy-evaluation) to
+improve performance of the increasingly prevalent data-intensive pipelines.
+As was noted in ~\cite{hayot2019performance}, in-memory computing, coupled
+with data locality, can bring significant performance improvements to data
+intensive neuroimaging pipelines. We extend this work by studying the
+differences between two Big Data frameworks, Dask~\cite{Dask:15} and Apache
+Spark~\cite{Spark:16}, for their suitability in the processing of neuroimaging
+pipelines. Our goal is to test whether Spark or Dask has a clear
+performance advantage to process Big neuroimaging Data.
 
 %Similarities
-\TG{Note for later: explain the following concepts in the intro: lazy eval, data locality, in-memory computing.}
-Spark and Dask offer in-memory computing, data locality, and lazy evaluation; which
-is common for Big Data framework. Both their scheduler operates dynamically. This is
-good when the runtimes are not known ahead of time~\cite{Dask:15}. Over these
-similarities, the frameworks are quite different.
+Spark and Dask both offer in-memory computing, data locality, and lazy
+evaluation; which is common for Big Data frameworks. Both their schedulers
+operate dynamically, which is good when task runtimes are not known ahead of
+time~\cite{Dask:15}. They also provide rich, high-level programming APIs,
+and support a variety of infrastructure schedulers, such as Mesos (Spark
+only), YARN, or HPC clusters (Dask only). Over these similarities, the
+frameworks have differences.
 
-% Spark
-On the one hand, we have Spark which provides a high-level API. This allows the
-scheduler to perform more optimizations which makes it well suited for neuroimaging
-analysis that often requires the usage of a pipeline with multiple steps. Though,
-Spark's code base is in Scala which theoretically can lead to slow down in execution
-due to a required serialization. Moreover, while Spark's API is flexible and allows
-most implementations, it differs from the ones seen in the Python's ecosystem.
+% Differences
+First and foremost, Spark is written in Scala while Dask is in Python.
+Given the popularity of Python in scientific communities, this arguably
+gives an edge to Dask due to data serialization costs from Python to Scala.
+On the other hand, Python's Global Interpreter Lock (GIL) might reduce
+parallelism in some cases. This difference in programming languages also
+has qualitative implications. As part of the \href{http://scipy.org}{SciPy}
+ecosystem, Dask provides almost transparent parallelization of applications
+manipulating Numpy arrays or Pandas data frames. On the other hand, Spark's
+Java, R and Python APIs allow to easily parallelize analyses that combine
+these languages, with reduced performance loss.
 
-% Dask
-On the other hand, Dask was created with the purpose of natively parallelize Python
-pipelines while keeping the syntax of familiar API from the Python's scientific
-ecosystem. However, Dask is still a young framework with work to be done; it's API
-does not completely replicate the library it supports. While its lower-level API
-allows the implementation of more complex algorithms it sacrifices a layer of
-optimization.
 % Related work
-Previous work shows that Dask had significant overhead and was
-hard to debug~\cite{Mehta:17}.
-% What issues your work addresses
-Dask was immature at the time and a lot of change was brought to the framework.
-Therefore we think it is valuable to re-compare it with Spark.
-\TG{Cite~\cite{hayot2019performance}}
+Spark and Dask were both included in the evaluation reported
+in~\cite{Mehta:17} that used a neuroimaging and an astronomy application
+processing approximately 100~GB of data each. In this work, Dask was
+reported to have a slight performance advantage over Spark. Overall, Dask's
+end-to-end time (makespan) was measured to be up to 14\% faster than Spark,
+due to ``more efficient pipelining'' %(Fig 10) 
+and serialization time to
+Python. % (Fig. 12-a)
+Dask, however, was reported to have a larger startup
+time than Spark. % (Fig. 12-c)
+The analysis remains at a quite high level
+though, leaving most of the observed performance difference unexplained. In
+comparison, our study will provide a detailed analysis of performance
+differences and similarities between Spark and Dask.
 
-%% Dask API
-The Dask APIs we decide to use for our comparison are Dask Bag, Delayed and Futures.
-Dask Bag offers an easy API to parallelize data. Dask Delayed offer a lower-level API
-that offers more flexibility; this is good to implement more complex tasks that do
-not fit in the Dask Bag framework. Dask Futures is a real-time API. Both Dask Bag and
-Delayed apply lazy evaluation to tasks while Futures trigger them directly.
-
-% Methods used (summary)
-The project aims to compare the state-of-the-art general-purpose framework Spark with
-the newcomer Dask. We decide to compare the performance of Spark and Dask on a custom
-incrementation pipeline to simplify the effect of the algorithm on the comparison.
-Then we assess the frameworks on two real-life applications: (1) histogram of the
-voxels intensity in a 3d images (2) BIDS example.
-
-% Implications of our research
-The result from our project help in deciding if Dask is a good choice to build
-neuroimaging pipelines.
-
-\TG{Say that it is often said that Spark has a larger overhead but scales 
-worse than Dask. In the dicussion, get back to it and mention that (1) we don't
-always observe that, (2) for data-intensive applications, overheads don't really matter
-as reducing them increases data trasnfers.}
+The next section details the design of our experiments. We consider two
+data-intensive neuroimaging applications: high-resolution imaging,
+represented by the Big Brain data~\cite{Amunts:13}, and large functional
+MRI studies, represented by data from the consortium for reliability and
+reproducibility (CoRR~\cite{zuo2014open}). We test application pipelines
+involving different patterns (map-only, map-reduce), and different types of
+implementations (plain Python, command-line,
+containerized). We evaluate performance on a dedicated cluster
+representative of the ones used in today's data-intensive neuroimaging
+studies, using the main Dask and Spark APIs. The following sections present
+our results, discussion and conclusions.
 
@@
 
-\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine.
-Spark's core is written in Scala, but it also has a Java, an R, and a Python API
-called PySpark. We used PySpark, given the popularity of Python in neuroscience,
-although the serialization cost of Python objects to Scala might be a substantial
-source of overhead. Spark provides in-memory computing, data locality, and lazy
-evaluation. PySpark's fundamental data structure, the Resilient Distributed Dataset
-(RDD)~\cite{RDD}, is a fault-tolerant and a parallel collection of data elements.
-Dataframes, PySpark's API for tabular data, is an overlay of RDDs with similar
-performance features. We used Spark's default scheduler, Spark standalone, although
-YARN~\cite{vavilapalli2013apache} and Mesos~\cite{hindman2011mesos} are also
-available. In the Spark standalone scheduler, a process called the \emph{master}
-coordinates resources provisioned by \emph{worker} processes on the cluster. The
-application is submitted to the \emph{driver} that in turn requests workers to the
-master and dispatches tasks to them. A job is divided into stages to be excuted in a
-different process onto the workers. Each stage's operation is represented as a
-high-level task in the computation graph. The Spark standalone scheduler uses a FIFO
-(First-In-First-Out) job scheduling policy. Processing jobs in order helps in
-increasing the page hit rate. As our applications only have one job, the LIFO
-(Last-In-First-Out) policy used for the stages is more pertinent \MD{Find better
-phrasing} \TG{refer to the paper (\MD{Not sure which one})} \VHS{I wouldn't
-explicitly state LIFO here. It's a bit complicated to be frank. Basically both Spark
-and Dask opt to complete a chain of dependencies rather than start a bunch of ``new
-chains''. And in this context, LIFO is great for the page hit rate, so you might want
-to reword the prior sentence as well}. The Spark standalone scheduler has two
-execution modes: client mode, where the driver runs in a dedicated process on the
-cluster, and cluster mode, where the driver runs in a worker. We used client mode as
-cluster mode is not available in PySpark. We used Apache Spark v2.4.0.
+\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine. It
+provides in-memory computing, data locality, and lazy evaluation. Spark's main data
+structure, the Resilient Distributed Dataset (RDD)~\cite{RDD}, is a fault-tolerant
+and a parallel collection of data elements. Dataframes, Spark's API for tabular data,
+is an overlay of RDDs with similar performance features \VHS{It's probably fine, but
+the SparkSQL Engine available for Scala and Java do bring an added peformance
+value}.\VHS{Spark has another API (Datasets). i know you used 'popular' for
+DataFrames, but it's not like Spark has a plethora of APIs. I think the APIs should
+be mentioned after the API languages such that you only list the APIs relevant to
+Python (RDD and Dataframes)} Spark's core is written in Scala, but it also has a
+Java, an R, and a Python API called PySpark. \MD{Sentence is ambiguous? could think
+that all three APIs are named PySpark} We used PySpark, given the popularity of
+Python in neuroscience, although the serialization cost of Python objects to Scala
+might be a substantial source of overhead. We used Spark's default scheduler, Spark
+standalone, although YARN~\cite{vavilapalli2013apache} and
+Mesos~\cite{hindman2011mesos} are also available. In the Spark standalone scheduler,
+a process called the \emph{master} coordinates resources provisioned by
+\emph{workers} on the cluster. The application is submitted to the \emph{driver} that
+in turn requests workers to the master and dispatches tasks to them. A job is divided
+into stages to be excuted in a different process onto the workers. Each stage's
+operation is represented as a high-level task in the computation graph. The Spark
+standalone scheduler uses a FIFO (First-In-First-Out) job scheduling policy.
+Processing jobs in order helps in increasing the page hit rate. As our applications
+only have one job, the LIFO (Last-In-First-Out) policy used for the stages is more
+pertinent \MD{Find better phrasing} \TG{refer to the paper (\MD{Not sure which one})}
+\VHS{I wouldn't explicitly state LIFO here. It's a bit complicated to be frank.
+Basically both Spark and Dask opt to complete a chain of dependencies rather than
+start a bunch of ``new chains''. And in this context, LIFO is great for the page hit
+rate, so you might want to reword the prior sentence as well}. The Spark standalone
+scheduler has two execution modes: client mode, where the driver runs in a dedicated
+process on the cluster, and cluster mode, where the driver runs in a worker. We used
+client mode as cluster mode is not available in PySpark. We used Apache Spark v2.4.0.
 
@@
 
-While Incrementation and Histogram were implemented directly in Python, BIDS app
-example requires an external library and is distributed as a Docker container image
-(bids/base\_fsl on DockerHub). We converted the Docker image to a Singularity image,
-given its popularity for HPC and reproducibility, using
-\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}.
-The image was preloaded on the nfs and singularity version 3.2.1\-1.el7 was installed
-on all instances. Due to a bug in the conversion from Docker to Singularity, we had to
-export the path for fsl.
-
+While Incrementation and Histogram were implemented directly in Python,
+BIDS app example requires an external library and is distributed as a
+Docker container image (bids/base\_fsl on DockerHub). To avoid remote file
+transfers from DockerHub, we preloaded our OpenStack instances with the
+Docker image, and started the Docker engine at startup.
 \subsection{Experiments}
@@
 data read, compute, data write, and engine overhead across all application tasks. 
-We measured the overhead by \TG{\ldots, maybe add a figure}. \MD{Done but might need
-adjustment depending on how the overhead is calculate (idle vs wasted)}
+We measured the overhead by \TG{\ldots, maybe add a figure}.
 
@@
 engines however Futures are always slightly faster. The performance difference between
-Futures and the other APIs increases with the number of workers. The makespan scales
+Futures and the other APIs increase with the number of workers. The makespan scales
 well with the increase in workers however the scaling reduces considerably when
-reaching 8 workers. This is due to the IO requirement increasing with the number of
-workers, since more tasks are read concurrently, and the application becomes more
-data-intensive.
+reaching 8 workers. This due to the IO requirement increasing with the number of
+workers, since more tasks are read concurrently, and the application becomes more data
+intensive.
 
@@
 
+% partitioning issue
+
 % \subsection{IO Time}

### 2019-07-24 latex/paper.tex
Added new data and updated figures

@@
 %Similarities
-Spark and Dask both offer in-memory computing, data locality, and lazy
-evaluation; which is common for Big Data frameworks. Both their schedulers
-operate dynamically, which is good when task runtimes are not known ahead of
-time~\cite{Dask:15}. They also provide rich, high-level programming APIs,
-and support a variety of infrastructure schedulers, such as Mesos (Spark
-only), YARN, or HPC clusters (Dask only). Over these similarities, the
-frameworks have differences.
+Spark and Dask both offer in-memory computing, data locality, and lazy evaluation;
+which is common for Big Data frameworks. Both their schedulers operate dynamically,
+which is good when task runtimes are not known ahead of time~\cite{Dask:15}. They
+also provide rich, high-level programming APIs, and support a variety of
+infrastructure schedulers, such as Mesos, YARN, or HPC clusters (Dask only). Over
+these similarities, the frameworks have differences.
 
@@
 in~\cite{Mehta:17} that used a neuroimaging and an astronomy application
-processing approximately 100~GB of data each. In this work, Dask was
+processing approximately \SI{100}{\giga\byte} of data each. In this work, Dask was
 reported to have a slight performance advantage over Spark. Overall, Dask's
@@
     \# of worker        & 1, 2, 4, 8     & 1, 2, 4, 8            & 1, 2, 4, 8            \\
-    \# of blocks        & 30, 125, 750   & 30, 125, 750          & 30, 125, 750          \\
+    \# of blocks        & 30, 125, 750   & 30, 125, 750          & \multicolumn{1}{c}{-} \\
     \# of iterations    & 1, 10, 100     & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\

### 2019-07-25 latex/paper.tex
Worked on results :: histogram section

@@
 for different numbers of workers and for different engines. The bars show the average
-makespan over 5 repetitions while the error bars are the standard deviation. Overall,
+makespan over 3 repetitions while the error bars are the standard deviation. Overall,
 there is no substantial makespan difference between the engines. Dask seems to have a
@@
 again equivalent, although Dask Delayed and Futures are slightly faster
-than Bags and Spark for 1 and 10 few iterations, and Fuures are faster than
+than Bags and Spark for 1 and 10 few iterations, and Futures are faster than
 Delayed, Bags and Spark for 100 iterations. Differences remain minor
@@
     \end{subfigure}
-    \caption{Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second} delay, 8
-    workers, 8 threads/worker}\label{fig:inc_gantt}
+    \caption{Incrementation: Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second}
+delay, 8 workers, 8 threads/worker}\label{fig:inc_gantt}
 \end{figure*}
@@
 Overall, the IO and overhead is comparable for all engines.
+
+On Figure~\ref{fig:histo_tt_worker}, Dask engines have similar total execution time
+when their number of workers varies, except for 8 worker where it incresase slightly.
+On the other hand, Spark total execution time keeps increasing; especially at 8
+workers. This is because Dask engines benefit more from additional workers.
 
@@
 
+\subsection{Histogram (NumPy): Number of workers}
+Figure~\ref{fig:histo_np_ms_worker}
+
+From Figure~\ref{fig:histo_np_tt_worker}
+
+\begin{figure}[!t]
+    \centering
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/histo_np_worker.png}%
+        \caption{Histogram (NumPy) makespan}\label{fig:histo_np_ms_worker}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/histo_idle_np_worker.png}%
+        \caption{Histogram (NumPy) total time}\label{fig:histo_np_tt_worker}
+    \end{subfigure}
+    \caption{125 blocks, 8 threads/worker}
+\end{figure}
+
+\subsection{Histogram (NumPy): Number of blocks}
+Figure~\ref{fig:histo_np_ms_block}
+
+\begin{figure}[!t]
+    \centering
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/histo_np_block.png}%
+        \caption{Histogram (NumPy) makespan}\label{fig:histo_np_ms_block}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/histo_idle_np_block.png}%
+        \caption{Histogram (NumPy) total time}\label{fig:histo_np_tt_block}
+    \end{subfigure}
+    \caption{8 workers, 8 threads/worker}
+\end{figure}
+
 %%% BIDS EXAMPLE %%%
@@
 the amount of worker. Overall, there is no substantial difference between the
-engines however Futures are always slightly faster. The performance difference between
-Futures and the other APIs increases with the number of workers. The makespan scales
-well with the increase in workers however the scaling reduces considerably when
-reaching 8 workers. This is due to the IO requirement increasing with the number of
-workers, since more tasks are read concurrently, and the application becomes more
-data-intensive.
+engines. The makespan scales well with the increase in workers however the scaling
+reduces considerably when reaching 8 workers. This is due to the IO requirement
+increasing with the number of workers --- more tasks are read concurrently and
+the application becomes more data-intensive.
 
 In Figure~\ref{fig:bids_tt_worker}, the total execution time of each function is
-shown. \MD{Maybe mention why Spark overhead is so big.} Overall, the computation time
-for the group analysis is equal for all engines and numbers of workers. Bag and
-Delayed takes much longer to compute the participant analysis with 8 workers,
-otherwise, the makespan is similar for all engines and numbers of workers. The overhead
-increases as the number of workers increases. \MD{Check regression}
-
-\MD{Depending on how the overhead is calculated spark overhead is either minimal or
-gigantic. It is hard to determine which method should be used to calculate the
-overhead for this application. For this application, since Spark has 2 job this
-causes Spark to have different worker process for the different job and thus some job
-have overlaping process ID while other are uniques (See
-Figure~\ref{fig:bids_spark_gantt}).}
+shown. Overall, there is no significant difference between the engines. Each engines
+have similar total time independently of their number of worker. This is because the
+engines scale almost linearly. The overhead increase slowly as the number of workers
+increases. This is due to more communication between the scheduler and workers as
+well as inter-worker communication.
 
@@
 
-% \begin{figure*}[!htb]
-%     \centering
-%     \begin{subfigure}[b]{\columnwidth}
-%         \includegraphics[clip,width=\columnwidth]{images/spark_bids_gantt.png}
-%         \caption{Spark}\label{fig:bids_spark_gantt}
-%     \end{subfigure}
-%     \hfill
-%     \begin{subfigure}[b]{\columnwidth}
-%         \includegraphics[clip,width=\columnwidth]{images/bag_bids_gantt.png}%
-%         \caption{Dask Bag}\label{fig:bids_dask_bag_gantt}
-%     \end{subfigure}
-%     \\
-%     \begin{subfigure}[b]{\columnwidth}
-%         \includegraphics[clip,width=\columnwidth]{images/delayed_bids_gantt.png}%
-%         \caption{Dask Delayed}\label{fig:bids_dask_delayed_gantt}
-%     \end{subfigure}
-%     \hfill
-%     \begin{subfigure}[b]{\columnwidth}
-%         \includegraphics[clip,width=\columnwidth]{images/futures_bids_gantt.png}%
-%         \caption{Dask Futures}\label{fig:bids_dask_futures_gantt}
-%     \end{subfigure}
-%     \caption{Gantt chart with 8 workers}\label{fig:bids_gantt}
-% \end{figure*}
+\begin{figure*}[!htb]
+    \centering
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/spark_bids_gantt.png}
+        \caption{Spark}\label{fig:bids_spark_gantt}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/bag_bids_gantt.png}%
+        \caption{Dask Bag}\label{fig:bids_dask_bag_gantt}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/delayed_bids_gantt.png}%
+        \caption{Dask Delayed}\label{fig:bids_dask_delayed_gantt}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/futures_bids_gantt.png}%
+        \caption{Dask Futures}\label{fig:bids_dask_futures_gantt}
+    \end{subfigure}
+    \caption{BIDS App example: Gantt chart with 8 workers}\label{fig:bids_gantt}
+\end{figure*}
 

### 2019-07-25 latex/paper.tex
Worked on results section for  histo, hitso (NumPy) and modified results section for Bids

@@
 iterations.
+% TODO
 
@@
 The computation time is significantly larger in Dask than Spark. This is potentially
-due to Python's GIL preventing Dask to parallelize the computation on multiple thread.
+due to Python's GIL preventing Dask to parallelize the computation on multiple threads.
 Overall, the IO and overhead is comparable for all engines.
@@
 \subsection{Histogram: Number of blocks}
-Figure~\ref{fig:histo_ms_block}
+Figure~\ref{fig:histo_ms_block} shows the makespan for different block size.
+Again, Spark is significantly faster then Dask engines and Bag is slightly faster
+than Delayed. Overall, the engines do not react to change in block size.
+
+In figure~\ref{fig:histo_tt_block} 
+%TODO
 
@@
 
+\subsection{Histogram: Gantt chart}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/spark_histo_gantt.png}
+        \caption{Spark}\label{fig:histo_spark_gantt}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/bag_histo_gantt.png}%
+        \caption{Dask Bag}\label{fig:histo_dask_bag_gantt}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/delayed_histo_gantt.png}%
+        \caption{Dask Delayed}\label{fig:histo_dask_delayed_gantt}
+    \end{subfigure}
+    \caption{Histogram: Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second}
+delay, 8 workers, 8 threads/worker}\label{fig:histo_gantt}
+\end{figure}
+
+%%% HISTOGRAM (NumPy) %%%
 \subsection{Histogram (NumPy): Number of workers}
-Figure~\ref{fig:histo_np_ms_worker}
+Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount fo workers.
+Overall, the makespan for all engines does not change considerably with an increase 
+in worker. This is beacuse the engines do not scale well.
 
-From Figure~\ref{fig:histo_np_tt_worker}
+On figure~\ref{fig:histo_np_ms_worker}, there is no substantial difference between
+ all engines, althought, Delayed is slightly faster.
+
+From Figure~\ref{fig:histo_np_tt_worker}, the total execution time spent in each
+function is shown. The overhead and read time increase proportionally to the number
+of workers thus increasing the total time. This is because (1) most worker read at
+the same time due to a low computation-IO ratio and (2) \MD{Determine why the
+overhead is so large compared to the pure Python implementation}
 
@@
 \subsection{Histogram (NumPy): Number of blocks}
-Figure~\ref{fig:histo_np_ms_block}
+Figure~\ref{fig:histo_np_ms_block} shows the makespan of the application for
+different block size. Note that the workers can process up to 64 tasks in paralle thus
+they are not used at full capacity with 30 blocks. Overall, there is no substantial
+difference for all engines and block size.
+
+In figure~\ref{fig:histo_np_tt_block}, shows the total time spent in each function.
+Again, the overhead for 30 blocks is erroneous due a lower amount of task than thread
+which result in higher overhead for thread based engines. Considering the overhead
+error and that at 30 blocks the worker only use half their core, the  total time is
+substantially the same for all block size. This is due lower block sizes having a
+lower IO time, however, it is balanced by the resulting increase in overhead.
 
@@
     \caption{8 workers, 8 threads/worker}
+\end{figure}
+
+\subsection{Histogram (NumPy): Gantt chart}
+
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/spark_histo_np_gantt.png}
+        \caption{Spark}\label{fig:histo_np_spark_gantt}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/bag_histo_np_gantt.png}%
+        \caption{Dask Bag}\label{fig:histo_np_dask_bag_gantt}
+    \end{subfigure}
+    \\
+    \begin{subfigure}[b]{\columnwidth}
+        \includegraphics[clip,width=\columnwidth]{images/delayed_histo_np_gantt.png}%
+        \caption{Dask Delayed}\label{fig:histo_np_dask_delayed_gantt}
+    \end{subfigure}
+    \caption{Histogram (NumPy): Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second}
+delay, 8 workers, 8 threads/worker}\label{fig:histo_np_gantt}
 \end{figure}
@@
 Figure~\ref{fig:bids_gantt} shows the Gantt charts obtained for each engine and API.
-The Gantt charts are structured in three main parts: participant analysis (orange),
-group analysis (blue), and the transition in between.\MD{rephrase} The Gantt chart
-for Spark seems to have more workers however it does not. This is due to the
-computation of the participant and group analysis on two different jobs with their
-worker processes; usually with different process id. The tasks differ considerably in
-length. This is due to the high IO requirement of the application. During the
-participant analysis phase, Spark and Futures have less overhead than Delayed and
-Bag. At the end of the participant analysis phase, Futures and Delayed has less
-overhead than Bag while Spark has way more than Dask engine. The group analysis is
-similar for all engines and APIs.
+The Gantt charts are structured in two main parts: participant analysis (orange) and
+group analysis (blue). The participant analysis tasks differ greatly in lenght. This
+is due to inqual amount of sessions per subject to process. The group analysis is
+similar for all engines and APIs. Overall, most of the overhead encountered results
+from the transition between the two analysis. This is because the group analysis is
+requires the results of the participant analysis to start.
 

### 2019-07-25 latex/paper.tex
Worked on histogram results

@@
 
-In figure~\ref{fig:histo_tt_block} 
+In figure~\ref{fig:histo_tt_block} the total time for each function is shown.
+Note that at 30 blocks the worker's resources are not fully used since the workers
+can process up to 64 task in parallel. Moreover, the overhead for 30 blocks is
+erroneous due to a lower amount of block than available threads. This leads to thread
+based engines to have higher overhead. Given that,
+
+Spark has a much lower compute time which makes it significantly faster than Dask
+engines. Overall, for all engines, lower block size results in lower IO time but
+increased overhead time. Again, this is due to the overhead desynchronizing task thus
+reducing the NFS bottle neck.
 %TODO
@@
 \subsection{Histogram: Gantt chart}
+Figure~\ref{fig:histo_gantt} shows the Gantt chart of the baseline Histogram
+experiment. Spark overhead is distributed through all the execution while Bag
+overhead happens mostly when a thread read subsequent blocks and Delayed overhead is
+concentrated after reading all the blocks.
+
+The reduce is done differently on the engines. While Spark and Delayed compute all
+the reduce task on one process (thread) Spark computes it on the driver process. On
+the other hand, Bag computes the reduce tasks on different workers and sends the
+informations to the driver process afterwards.
 
@@
 Figure~\ref{fig:histo_np_ms_block} shows the makespan of the application for
-different block size. Note that the workers can process up to 64 tasks in paralle thus
+different block size. Note that the workers can process up to 64 tasks in parallel thus
 they are not used at full capacity with 30 blocks. Overall, there is no substantial
@@
 In figure~\ref{fig:histo_np_tt_block}, shows the total time spent in each function.
-Again, the overhead for 30 blocks is erroneous due a lower amount of task than thread
-which result in higher overhead for thread based engines. Considering the overhead
-error and that at 30 blocks the worker only use half their core, the  total time is
-substantially the same for all block size. This is due lower block sizes having a
-lower IO time, however, it is balanced by the resulting increase in overhead.
+Again, the overhead for 30 blocks is erroneous due a lower amount of task than
+available thread which result in higher overhead for thread based engines.
+Considering the overhead error and that at 30 blocks the worker only use half their
+core, the total time is substantially the same for all block size. This is due lower
+block sizes having a lower IO time, however, it is balanced by the resulting increase
+in overhead.
 
@@
 \subsection{Histogram (NumPy): Gantt chart}
+Figure~\ref{fig:histo_np_gantt} shows the Gantt chart of the baseline Histogram
+(NumPy) experiment. The overhead for Spark and Bag is mostly located between a read
+and compute task or compute and read task. On the other hand, Delayed overhead is
+dispersed between all type of tasks. This cause Delayed to overlap more compute and
+read tasks amongs the different workers.
+
+In figure~\ref{fig:histo_np_gantt}, The read tasks of all engines are similar initially, however, when computation starts
+the subsequent read tasks are much shorter. This is because read tasks for Delayed
+become desynchronized hence it significanlty reduces the IO bottle neck.
 

### 2019-07-25 latex/paper.tex
Typos

@@
 Within Dask, Delayed and Futures have a higher overhead than Bags. However,
-overhead differences are compensated by an increase of data transfer time,
+overhead differences are compensated by an increase in data transfer time,
 as a reduced overhead increases the concurrency between data transfers.
@@
 Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when varying the
-number of image blocks for a constant BigBrain image size. We were not able to run
+number of image blocks for constant BigBrain image size. We were not able to run
 Spark for 30 blocks due to its \SI{2}{\giga\byte} limitation in the task size it can
@@
 In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
-For 30 blocks the Dask Bag API has a much lower overhead time. This is because Dask
-Bag was only using one thread per block in comparison to Dask Delayed and Dask
-Futures which were offloading a block calculations on multiple threads of the same
-worker. Once again, the data transfer time reduces with more blocks however the
-overhead time increases by a similar amount. This is not observed for 30 blocks as
-the workers are not used at full capacity, i.e., some threads are idle. Finally, the
-variability of the overhead increases with the number of blocks, which explains the
-makespan variability mentioned previously.
+For 30 blocks the Dask Bag API has a much lower overhead time. This is because Bag
+was only using one thread per block in comparison to Delayed and Futures that were
+offloading block calculations on multiple threads of the same worker. Once again,
+the data transfer time reduces with more blocks however the overhead time increases
+by a similar amount. This is not observed for 30 blocks as the workers are not used
+at full capacity, i.e., some threads are idle. Finally, the variability of the
+overhead increases with the number of blocks, which explains the makespan variability
+mentioned previously.
 
@@
 again equivalent, although Dask Delayed and Futures are slightly faster
-than Bags and Spark for 1 and 10 few iterations, and Futures are faster than
+than Bags and, Spark for 1 and 10 few iterations, and Futures are faster than
 Delayed, Bags and Spark for 100 iterations. Differences remain minor
@@
 In Figure~\ref{fig:inc_tt_itr}, the total execution time breakdown is
-shown. We observe good scalability of all the engines with the number of
+shown. We observe the good scalability of all the engines with the number of
 iterations.
@@
 Figure~\ref{fig:inc_ms_sleep} shows the makespan of the Incrementation application
-for different sleep delays. Overall, all engines again perform the same, and scale
+for different sleep delays. Overall, all engines again perform the same and scale
 well with task duration. Spark is initially slower than the Dask APIs, however, it is
@@
 Figure~\ref{fig:histo_ms_worker} shows the makespan of the application for various
-amounts of workers. Spark is tremendiously faster than Dask APIs. The difference
+amounts of workers. Spark is tremendously faster than Dask APIs. The difference
 narrows as the number of workers increases. Between Dask APIs there is no substantial
@@
 
-From Figure~\ref{fig:histo_tt_worker} the total time spend in each function is shown.
+From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is shown.
 The computation time is significantly larger in Dask than Spark. This is potentially
 due to Python's GIL preventing Dask to parallelize the computation on multiple threads.
-Overall, the IO and overhead is comparable for all engines.
+Overall, the IO and overhead are comparable for all engines.
 
 On Figure~\ref{fig:histo_tt_worker}, Dask engines have similar total execution time
-when their number of workers varies, except for 8 worker where it incresase slightly.
+when their number of workers varies, except for 8 workers where it increases slightly.
 On the other hand, Spark total execution time keeps increasing; especially at 8
@@
 Note that at 30 blocks the worker's resources are not fully used since the workers
-can process up to 64 task in parallel. Moreover, the overhead for 30 blocks is
+can process up to 64 tasks in parallel. Moreover, the overhead for 30 blocks is
 erroneous due to a lower amount of block than available threads. This leads to thread
@@
 engines. Overall, for all engines, lower block size results in lower IO time but
-increased overhead time. Again, this is due to the overhead desynchronizing task thus
-reducing the NFS bottle neck.
+increases the overhead time. Again, this is due to the overhead desynchronizing task
+thus reducing the NFS bottleneck.
 %TODO
@@
 The reduce is done differently on the engines. While Spark and Delayed compute all
-the reduce task on one process (thread) Spark computes it on the driver process. On
+the reduce tasks on one process (thread) Spark computes it on the driver process. On
 the other hand, Bag computes the reduce tasks on different workers and sends the
-informations to the driver process afterwards.
+information to the driver process afterward.
 
@@
 \subsection{Histogram (NumPy): Number of workers}
-Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount fo workers.
+Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount of workers.
 Overall, the makespan for all engines does not change considerably with an increase
-in worker. This is beacuse the engines do not scale well.
+in the number of workers. This is because the engines do not scale well.
 
 On figure~\ref{fig:histo_np_ms_worker}, there is no substantial difference between
- all engines, althought, Delayed is slightly faster.
+ all engines, although, Delayed is slightly faster.
 
@@
 function is shown. The overhead and read time increase proportionally to the number
-of workers thus increasing the total time. This is because (1) most worker read at
+of workers thus increasing the total time. This is because (1) most workers read at
 the same time due to a low computation-IO ratio and (2) \MD{Determine why the
@@
 Figure~\ref{fig:histo_np_ms_block} shows the makespan of the application for
-different block size. Note that the workers can process up to 64 tasks in parallel thus
-they are not used at full capacity with 30 blocks. Overall, there is no substantial
-difference for all engines and block size.
+different block size. Note that the workers can process up to 64 tasks in parallel
+thus they are not used at full capacity with 30 blocks. Overall, there is no
+substantial difference for all engines and block size.
 
 In figure~\ref{fig:histo_np_tt_block}, shows the total time spent in each function.
-Again, the overhead for 30 blocks is erroneous due a lower amount of task than
-available thread which result in higher overhead for thread based engines.
-Considering the overhead error and that at 30 blocks the worker only use half their
-core, the total time is substantially the same for all block size. This is due lower
-block sizes having a lower IO time, however, it is balanced by the resulting increase
-in overhead.
+Again, the overhead for 30 blocks is erroneous due to a lower amount of task than
+available threads which result in higher overhead for thread-based engines.
+Considering the overhead error and that at 30 blocks the workers only use half their
+core, the total time is substantially the same for all block size. This is due to
+lower block sizes having a lower IO time, however, it is balanced by the resulting
+increase in overhead.
 
@@
 and compute task or compute and read task. On the other hand, Delayed overhead is
-dispersed between all type of tasks. This cause Delayed to overlap more compute and
-read tasks amongs the different workers.
+dispersed between all types of tasks. This causes Delayed to overlap more compute and
+read tasks among the different workers.
 
@@
 the subsequent read tasks are much shorter. This is because read tasks for Delayed
-become desynchronized hence it significanlty reduces the IO bottle neck.
+become desynchronized hence it significantly reduces the IO bottleneck.
 
@@
 In Figure~\ref{fig:bids_tt_worker}, the total execution time of each function is
-shown. Overall, there is no significant difference between the engines. Each engines
-have similar total time independently of their number of worker. This is because the
-engines scale almost linearly. The overhead increase slowly as the number of workers
-increases. This is due to more communication between the scheduler and workers as
-well as inter-worker communication.
+shown. Overall, there is no significant difference between the engines. Each engine
+has a similar total time independently of its number of workers. This is because of
+the, almost, linearly scale of the engines. The overhead increases slowly as the
+number of workers increases. This is due to more communication between the scheduler
+and workers as well as inter-worker communication.
 
@@
 The Gantt charts are structured in two main parts: participant analysis (orange) and
-group analysis (blue). The participant analysis tasks differ greatly in lenght. This
-is due to inqual amount of sessions per subject to process. The group analysis is
-similar for all engines and APIs. Overall, most of the overhead encountered results
-from the transition between the two analysis. This is because the group analysis is
-requires the results of the participant analysis to start.
+group analysis (blue). The participant analysis tasks differ greatly in length. This
+is due to the unequal amount of sessions per subject to process. The group analysis
+is similar for all engines and APIs. Overall, most of the overhead encountered
+results from the transition between the two analysis. This is because the group
+analysis requires the results of the participant analysis to start.
 
@@
 
-Mathieu Dugr\'e was funded by a Undergraduate Student Research Assistant award from
+Mathieu Dugr\'e was funded by an Undergraduate Student Research Assistant award from
 the National Science and Engineering Research Council of Canada. We warmly thank
 Compute Canada and \TG{Westgrid?} for providing the cloud infrastructure used in
-these experiments, and the McGill Center for Integrative Neurocience for giving us
+these experiments, and the McGill Center for Integrative Neuroscience for giving us
 access to their cloud allocation.

### 2019-07-25 latex/paper.tex
fixed spark section

--- 2019-07-25 latex/paper.tex
@@
 \subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine.
-Spark's core is written in Scala, but it also has a Java, an R, and a Python API
-called PySpark. We used PySpark, given the popularity of Python in neuroscience,
-although the serialization cost of Python objects to Scala might be a substantial
-source of overhead. Spark provides in-memory computing, data locality, and lazy
-evaluation. PySpark's fundamental data structure, the Resilient Distributed Dataset
-(RDD)~\cite{RDD}, is a fault-tolerant and a parallel collection of data elements.
-Dataframes, PySpark's API for tabular data, is an overlay of RDDs with similar
-performance features. We used Spark's default scheduler, Spark standalone, although
-YARN~\cite{vavilapalli2013apache} and Mesos~\cite{hindman2011mesos} are also
-available. In the Spark standalone scheduler, a process called the \emph{master}
-coordinates resources provisioned by \emph{worker} processes on the cluster. The
-application is submitted to the \emph{driver} that in turn requests workers to the
-master and dispatches tasks to them. A job is divided into stages to be excuted in a
-different process onto the workers. Each stage's operation is represented as a
-high-level task in the computation graph. The Spark standalone scheduler uses a FIFO
-(First-In-First-Out) job scheduling policy. Processing jobs in order helps in
-increasing the page hit rate. As our applications only have one job, the LIFO
-(Last-In-First-Out) policy used for the stages is more pertinent \MD{Find better
-phrasing} \TG{refer to the paper (\MD{Not sure which one})} \VHS{I wouldn't
-explicitly state LIFO here. It's a bit complicated to be frank. Basically both Spark
-and Dask opt to complete a chain of dependencies rather than start a bunch of ``new
-chains''. And in this context, LIFO is great for the page hit rate, so you might want
-to reword the prior sentence as well}. The Spark standalone scheduler has two
-execution modes: client mode, where the driver runs in a dedicated process on the
-cluster, and cluster mode, where the driver runs in a worker. We used client mode as
-cluster mode is not available in PySpark. We used Apache Spark v2.4.0.
+While its core is written in Scala, Spark provides additional APIs in Java, R
+and Python (PySpark). Its main abstraction, the Resilient Distributed Dataset
+(RDD)~\cite{RDD}, is a fault-tolerant, parallel collection of data elements. This
+data structure is the basis of Spark's other data structures, namely, DataFrames
+and Datasets. Datasets are similar to RDDs but additionally use Spark SQL's
+optimized execution to further improve performance. DataFrames, used to process
+tabular data, Datasets where the data is organized into named-columns. While 
+the DataFrame API exists in all the available language APIs, Datasets are
+limited to Scala and Java.
+
+As data transfers in BigData workflows are an important source of performance
+bottlenecks, Spark incorporated
+data locality, and additionally, introduced in-memory computing to the Big Data 
+frameworks. Data locality, introduced by MapReduce~\cite{mapreduce}, schedules
+tasks as close as possible to where the data is storage. In-memory computing ensures
+data is maintained in memory whenever possible, as writting large amounts data to
+slower storage devices may be costly. To reduce any unecessary communication
+and computation, Spark also included lazy evaluation, which builds the entire
+task graph prior to execution to determine what needs to be computed.
+
+Spark is compatible with three different schedulers: Spark
+standalone, YARN~\cite{vavilapalli2013apache} and Mesos~\cite{hindman2011mesos}. 
+The Spark standalone is a simple default scheduler built into Spark. YARN is a
+more complex scheduler primarily designed to schedule Hadoop-based workflows. 
+Mesos, on the other hand, is not limited to Hadoop and can be used to schedule
+a variety of different workflows. As scientific researchers would likely be
+executing their workflows in HPC environments with neither YARN or Mesos
+installed, we limit our focus to Spark's Standalone scheduler.
+
+The Spark Standalone scheduler is composed of three main processes: the
+\emph{master}, the \emph{workers} (\emph{slaves}, in Spark) and the
+\emph{driver}. The \emph{master} coordinates resources provisioned by
+\emph{worker} processes on the cluster. The application is submitted to the
+\emph{driver} that in turn requests workers to the master and dispatches tasks
+to them. A job is divided into stages to be excuted in a different process onto
+the workers. Each stage's operation is represented as a high-level task in the computation
+graph. The Spark standalone scheduler uses a FIFO (First-In-First-Out) job
+scheduling policy. The Spark standalone scheduler has two
+execution modes: client mode, where the driver runs in a dedicated process outside
+of the Spark cluster, and cluster mode, where the driver runs in a worker. We used
+client mode as cluster mode is not available in PySpark. We used Apache Spark v2.4.0.
+
+Python is commonly selected as the programming language of choice in scientific
+communities, especially for our use case of neuroscience, where numerous specialized
+Python libraries exist to study the data. While serialization of 
+Python to Java may lead to significant overheads, we chose to focus on Spark 
+Python API due to its suitability for neuroimaging research.
 
 \subsubsection{Dask} Dask is a Big Data engine that is becoming increasingly popular
-in the scientific Python ecosystem. Dask has five main APIs and data structures:
+in the scientific Python ecosystem. Dask has five main data structure APIs:
 Array, Bag, DataFrame, Delayed, and Futures. All APIs were used in our experiments,
@@
 for different numbers of workers and for different engines. The bars show the average
-makespan over 3 repetitions while the error bars are the standard deviation. Overall,
+makespan over 5 repetitions while the error bars are the standard deviation. Overall,
 there is no substantial makespan difference between the engines. Dask seems to have a
@@
 again equivalent, although Dask Delayed and Futures are slightly faster
-than Bags and Spark for 1 and 10 few iterations, and Futures are faster than
+than Bags and Spark for 1 and 10 few iterations, and Fuures are faster than
 Delayed, Bags and Spark for 100 iterations. Differences remain minor
@@
 iterations.
-% TODO
 
@@
     \end{subfigure}
-    \caption{Incrementation: Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second}
-delay, 8 workers, 8 threads/worker}\label{fig:inc_gantt}
+    \caption{Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second} delay, 8
+    workers, 8 threads/worker}\label{fig:inc_gantt}
 \end{figure*}
@@
 The computation time is significantly larger in Dask than Spark. This is potentially
-due to Python's GIL preventing Dask to parallelize the computation on multiple threads.
+due to Python's GIL preventing Dask to parallelize the computation on multiple thread.
 Overall, the IO and overhead is comparable for all engines.
-
-On Figure~\ref{fig:histo_tt_worker}, Dask engines have similar total execution time
-when their number of workers varies, except for 8 worker where it incresase slightly.
-On the other hand, Spark total execution time keeps increasing; especially at 8
-workers. This is because Dask engines benefit more from additional workers.
 
@@
 \subsection{Histogram: Number of blocks}
-Figure~\ref{fig:histo_ms_block} shows the makespan for different block size.
-Again, Spark is significantly faster then Dask engines and Bag is slightly faster
-than Delayed. Overall, the engines do not react to change in block size.
-
-In figure~\ref{fig:histo_tt_block} 
-%TODO
+Figure~\ref{fig:histo_ms_block}
 
@@
 
-\subsection{Histogram: Gantt chart}
-
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/spark_histo_gantt.png}
-        \caption{Spark}\label{fig:histo_spark_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/bag_histo_gantt.png}%
-        \caption{Dask Bag}\label{fig:histo_dask_bag_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/delayed_histo_gantt.png}%
-        \caption{Dask Delayed}\label{fig:histo_dask_delayed_gantt}
-    \end{subfigure}
-    \caption{Histogram: Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second}
-delay, 8 workers, 8 threads/worker}\label{fig:histo_gantt}
-\end{figure}
-
-%%% HISTOGRAM (NumPy) %%%
-\subsection{Histogram (NumPy): Number of workers}
-Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount fo workers.
-Overall, the makespan for all engines does not change considerably with an increase 
-in worker. This is beacuse the engines do not scale well.
-
-On figure~\ref{fig:histo_np_ms_worker}, there is no substantial difference between
- all engines, althought, Delayed is slightly faster.
-
-From Figure~\ref{fig:histo_np_tt_worker}, the total execution time spent in each
-function is shown. The overhead and read time increase proportionally to the number
-of workers thus increasing the total time. This is because (1) most worker read at
-the same time due to a low computation-IO ratio and (2) \MD{Determine why the
-overhead is so large compared to the pure Python implementation}
-
-\begin{figure}[!t]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_np_worker.png}%
-        \caption{Histogram (NumPy) makespan}\label{fig:histo_np_ms_worker}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_idle_np_worker.png}%
-        \caption{Histogram (NumPy) total time}\label{fig:histo_np_tt_worker}
-    \end{subfigure}
-    \caption{125 blocks, 8 threads/worker}
-\end{figure}
-
-\subsection{Histogram (NumPy): Number of blocks}
-Figure~\ref{fig:histo_np_ms_block} shows the makespan of the application for
-different block size. Note that the workers can process up to 64 tasks in paralle thus
-they are not used at full capacity with 30 blocks. Overall, there is no substantial
-difference for all engines and block size.
-
-In figure~\ref{fig:histo_np_tt_block}, shows the total time spent in each function.
-Again, the overhead for 30 blocks is erroneous due a lower amount of task than thread
-which result in higher overhead for thread based engines. Considering the overhead
-error and that at 30 blocks the worker only use half their core, the  total time is
-substantially the same for all block size. This is due lower block sizes having a
-lower IO time, however, it is balanced by the resulting increase in overhead.
-
-\begin{figure}[!t]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_np_block.png}%
-        \caption{Histogram (NumPy) makespan}\label{fig:histo_np_ms_block}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_idle_np_block.png}%
-        \caption{Histogram (NumPy) total time}\label{fig:histo_np_tt_block}
-    \end{subfigure}
-    \caption{8 workers, 8 threads/worker}
-\end{figure}
-
-\subsection{Histogram (NumPy): Gantt chart}
-
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/spark_histo_np_gantt.png}
-        \caption{Spark}\label{fig:histo_np_spark_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/bag_histo_np_gantt.png}%
-        \caption{Dask Bag}\label{fig:histo_np_dask_bag_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/delayed_histo_np_gantt.png}%
-        \caption{Dask Delayed}\label{fig:histo_np_dask_delayed_gantt}
-    \end{subfigure}
-    \caption{Histogram (NumPy): Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second}
-delay, 8 workers, 8 threads/worker}\label{fig:histo_np_gantt}
-\end{figure}
-
 %%% BIDS EXAMPLE %%%
@@
 the amount of worker. Overall, there is no substantial difference between the
-engines. The makespan scales well with the increase in workers however the scaling
-reduces considerably when reaching 8 workers. This is due to the IO requirement
-increasing with the number of workers --- more tasks are read concurrently and
-the application becomes more data-intensive.
+engines however Futures are always slightly faster. The performance difference between
+Futures and the other APIs increases with the number of workers. The makespan scales
+well with the increase in workers however the scaling reduces considerably when
+reaching 8 workers. This is due to the IO requirement increasing with the number of
+workers, since more tasks are read concurrently, and the application becomes more
+data-intensive.
 
 In Figure~\ref{fig:bids_tt_worker}, the total execution time of each function is
-shown. Overall, there is no significant difference between the engines. Each engines
-have similar total time independently of their number of worker. This is because the
-engines scale almost linearly. The overhead increase slowly as the number of workers
-increases. This is due to more communication between the scheduler and workers as
-well as inter-worker communication.
+shown. \MD{Maybe mention why Spark overhead is so big.} Overall, the computation time
+for the group analysis is equal for all engines and numbers of workers. Bag and
+Delayed takes much longer to compute the participant analysis with 8 workers,
+otherwise, the makespan is similar for all engines and numbers of workers. The overhead
+increases as the number of workers increases. \MD{Check regression}
+
+\MD{Depending on how the overhead is calculated spark overhead is either minimal or
+gigantic. It is hard to determine which method should be used to calculate the
+overhead for this application. For this application, since Spark has 2 job this
+causes Spark to have different worker process for the different job and thus some job
+have overlaping process ID while other are uniques (See
+Figure~\ref{fig:bids_spark_gantt}).}
 
@@
 Figure~\ref{fig:bids_gantt} shows the Gantt charts obtained for each engine and API.
-The Gantt charts are structured in two main parts: participant analysis (orange) and
-group analysis (blue). The participant analysis tasks differ greatly in lenght. This
-is due to inqual amount of sessions per subject to process. The group analysis is
-similar for all engines and APIs. Overall, most of the overhead encountered results
-from the transition between the two analysis. This is because the group analysis is
-requires the results of the participant analysis to start.
+The Gantt charts are structured in three main parts: participant analysis (orange),
+group analysis (blue), and the transition in between.\MD{rephrase} The Gantt chart
+for Spark seems to have more workers however it does not. This is due to the
+computation of the participant and group analysis on two different jobs with their
+worker processes; usually with different process id. The tasks differ considerably in
+length. This is due to the high IO requirement of the application. During the
+participant analysis phase, Spark and Futures have less overhead than Delayed and
+Bag. At the end of the participant analysis phase, Futures and Delayed has less
+overhead than Bag while Spark has way more than Dask engine. The group analysis is
+similar for all engines and APIs.
 
-\begin{figure*}[!htb]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/spark_bids_gantt.png}
-        \caption{Spark}\label{fig:bids_spark_gantt}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/bag_bids_gantt.png}%
-        \caption{Dask Bag}\label{fig:bids_dask_bag_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/delayed_bids_gantt.png}%
-        \caption{Dask Delayed}\label{fig:bids_dask_delayed_gantt}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/futures_bids_gantt.png}%
-        \caption{Dask Futures}\label{fig:bids_dask_futures_gantt}
-    \end{subfigure}
-    \caption{BIDS App example: Gantt chart with 8 workers}\label{fig:bids_gantt}
-\end{figure*}
+% \begin{figure*}[!htb]
+%     \centering
+%     \begin{subfigure}[b]{\columnwidth}
+%         \includegraphics[clip,width=\columnwidth]{images/spark_bids_gantt.png}
+%         \caption{Spark}\label{fig:bids_spark_gantt}
+%     \end{subfigure}
+%     \hfill
+%     \begin{subfigure}[b]{\columnwidth}
+%         \includegraphics[clip,width=\columnwidth]{images/bag_bids_gantt.png}%
+%         \caption{Dask Bag}\label{fig:bids_dask_bag_gantt}
+%     \end{subfigure}
+%     \\
+%     \begin{subfigure}[b]{\columnwidth}
+%         \includegraphics[clip,width=\columnwidth]{images/delayed_bids_gantt.png}%
+%         \caption{Dask Delayed}\label{fig:bids_dask_delayed_gantt}
+%     \end{subfigure}
+%     \hfill
+%     \begin{subfigure}[b]{\columnwidth}
+%         \includegraphics[clip,width=\columnwidth]{images/futures_bids_gantt.png}%
+%         \caption{Dask Futures}\label{fig:bids_dask_futures_gantt}
+%     \end{subfigure}
+%     \caption{Gantt chart with 8 workers}\label{fig:bids_gantt}
+% \end{figure*}
 

### 2019-07-25 latex/paper.tex
made fixes to Dask and the remainder of the methods section

--- 2019-07-25 latex/paper.tex
@@
 and Python (PySpark). Its main abstraction, the Resilient Distributed Dataset
-(RDD)~\cite{RDD}, is a fault-tolerant, parallel collection of data elements. This
-data structure is the basis of Spark's other data structures, namely, DataFrames
+(RDD)~\cite{RDD}, is a fault-tolerant, parallel collection of data elements. It 
+achieves fault tolerance through recording the data lineage, the sequence of 
+operations required to produce the data. The RDD is the basis of Spark's other
+data structures, namely, DataFrames
 and Datasets. Datasets are similar to RDDs but additionally use Spark SQL's
@@
 \subsubsection{Dask} Dask is a Big Data engine that is becoming increasingly popular
-in the scientific Python ecosystem. Dask has five main data structure APIs:
-Array, Bag, DataFrame, Delayed, and Futures. All APIs were used in our experiments,
-with the exception of Dask Dataframes, which is intended for use with tabular data.
-Like Spark, all Dask APIs provide
-in-memory computing, data locality, lazy evaluation, and fault tolerance by mean of
-lineage. Dask's core is purely in Python and bypasses most of the
-\href{https://docs.python.org/3/glossary.html#term-gil}{GIL} to provide parallelism.
-We used Dask due to the common use of the scientific Python ecosystem in
-neuroscience\VHS{If you add Dask Arrays to your experiments here, i'd mention that their
-APIs might be particularly well suited for the processing of high resolution neuroimaging data}. Dask represents application tasks as Dask graphs, produced by a
-user-facing API and executed by a scheduler. In Dask, operations generate multiple
-tiny tasks in the computation graph allowing an easier representation of complex
-algorithms. We used the \href{https://distributed.dask.org/en/latest/index.html}{Dask
-Distributed} scheduler, although \href{https://github.com/dask/dask-yarn}{Dask-Yarn}
-and \href{https://github.com/mrocklin/dask-mesos}{Dask-Mesos} is also available. In
-the Dask Distributed scheduler, a process called \textit{dask-scheduler}
-administrates the resources provided by \textit{workers} in the cluster. The
-scheduler receives jobs from clients and assigns tasks to available workers. Like
-Spark scheduler, Dask Distributed also use a LIFO policy, it is meant to reduce the
-memory footprint of the application by finishing the computation of a branch in the
-Dask graph before starting new ones. We used Dask v1.1.4.
+in the scientific Python ecosystem. Like Spark, Dask provides
+in-memory computing, data locality, lazy evaluation, and fault tolerance by
+means of data lineage. In contrast, Dask's core is purely in Python, eliminating
+the cost of serialization to Java. Furthermore, it is lightweight and modular,
+allowing users to only install the components they require.
 
-In Dask, a \href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
-collection of Python objects, similar to Spark's RDD. It use multi-process to enable
-parallelism. It offers a programming abstraction similar to the
-\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}. An
-\href{https://docs.dask.org/en/latest/array.html}{Array} is used for the processing
-of large arrays. It provides a distributed clone of the popular NumPy library. A
+Dask has five main data structure APIs:
+Array, Bag, DataFrame, Delayed, and Futures. A Dask
+\href{https://docs.dask.org/en/latest/array.html}{Array} is used for the
+processing of large arrays. It provides a distributed clone of the popular NumPy
+library. Similar to Spark's RDD, a Dask
+\href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
+collection of Python objects. It offers a programming abstraction similar to the
+\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}. A
 \href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a parallel
@@
 as they support arbitrary tasks, but they operate in real-time rather than lazily.
+All Dask data structure APIs, with the exception of Dask Arrays and Dask
+DataFrames, were used in our experiments.
+
+While Python multithreading is limited by the
+\href{https://docs.python.org/3/glossary.html#term-gil}{Global Interpreter Lock (GIL)}
+as a result of CPython's non thread-safe memory management, Dask attempts to
+bypass the GIL whenever possible and use multithreading to provide parallelism.
+With the exception of Dask Bags, all APIs, by default, use the local multithreaded
+scheduler. Dask Bags instead relies on the local multiprocessing scheduler.
+
+The Dask graph is the internal representation of Dask application to be executed 
+by the scheduler. API Operations generate multiple
+small tasks in the computation graph allowing an easier representation of complex
+algorithms.
+
+The Dask framework is compatible with numerous distributed schedulers, including 
+\href{https://github.com/dask/dask-yarn}{YARN} and
+\href{https://github.com/mrocklin/dask-mesos}{Mesos}, similarly to Spark. Dask 
+also provides its own distributed scheduler, known as the Dask Distributed
+scheduler. Although Dask is compatible with schedulers commonly found in HPC
+clusters, we chose to use Dask's Distributed scheduler, to keep the environment
+balanced between both frameworks.
+
+ In the Dask Distributed scheduler, a process called \textit{dask-scheduler}
+administrates the resources provided by \textit{workers} in the cluster. The
+scheduler receives jobs from clients and assigns tasks to available workers. Similarly
+to Spark's scheduler, Dask Distributed completes the processing of a graph branch
+before moving along to the next one.
+
+In our experiments, we used Dask v1.1.4.
 
@@
  We used Compute Canada's
- \href{https://docs.computecanada.ca/wiki/Cloud_resources}{Arbutus Cloud} operated by
+ \href{https://docs.computecanada.ca/wiki/Cloud\_resources}{Arbutus Cloud} operated by
  the \href{https://www.westgrid.ca}{WestGrid} regional organization at the University
- of Victoria, and running OpenStack \TG{version?}. We used c8-30gb-186 cloud
+ of Victoria, running OpenStack Queens release. We used c8-30gb-186 cloud
  instances with 8 VCPUs, an Intel Xeon Gold 6130 processor, \SI{30}{\giga\byte} of
@@
  memory management: at 60\% it spill data to disk, and 50\% of that amount is
- reserved for storage and is immuned from eviction.
+ reserved for storage and is immune to eviction.
 
- One cloud instance did not host any worker, and had a \SI{2}{\tera\byte} disk volume
+ One cloud instance did not host any worker and had a \SI{2}{\tera\byte} disk volume
  shared with the other instances using the Network File System (NFS) v4.
@@
 stored in 125 blocks in the MINC~\cite{minc} HDF5-based format, available
-at \url{ftp://bigbrain.loris.ca/BigBrainRelease.2015/3D_Blocks/40um} at the
+at \url{ftp://bigbrain.loris.ca/BigBrainRelease.2015/3D\_Blocks/40um} at a
 resolution of \SI{40}{\micro\metre}. We converted the blocks into the
@@
 
-We used three neuroimaging applications to evaluate the engines in different
-conditions. The first two ones, incrementation and histogram, are simple synthetic
-benchmarks representing basic map-only and map-reduce applications. The third one is
-a realistic application representative of popular BIDS
-applications~\cite{gorgolewski2017bids}. All script used for our experiment is availble
-GitHub at \url{https://github.com/big-data-lab-team/paper-big-data-engines}
+Three neuroimaging applications to evaluate the engines in different
+conditions were used. The first two, incrementation and histogram, are simple 
+synthetic applications representing map-only and map-reduce applications,
+respectively. The third one is a real neuroimaging application representative
+of popular BIDS applications~\cite{gorgolewski2017bids}. All scripts used for our
+experiments are available at
+\url{https://github.com/big-data-lab-team/paper-big-data-engines}
 
@@
 \subsubsection{Incrementation}
-We used an adaptation of the simple image incrementation pipeline used
+We used an adaptation of the image incrementation pipeline used
 in~\cite{hayot2019performance} (see Algorithm~\ref{alg:incrementation}).
@@
 system, increments the intensity value of each voxel by 1 to avoid caching
-effects, sleeps for a configurable amount of time to emulate a more complex
-processing, repeats this process for a specified amount of iterations, and
+effects, sleeps for a configurable amount of time to emulate more
+compute-intensive processing, repeats this process for a specified number of
+iterations, and
 finally writes the result as a NifTI image back to the shared file system.
 This application allows us to study the behavior of the engines when all
-inputs are processed independently, in a map-only way (see
+inputs are processed independently, in a map-only scenario (see
 Figure~\ref{fig:tg-inc}). This mimics the behavior of analyzing multiple
@@
  single \SI{766}{\kilo\byte} file. This application implements a typical map-reduce
- pattern, where the final result is obtained from all the input blocks (see
- Fig.~\ref{fig:tg-histo}). This application requires data shuffling thus inter-worker
- communication. The total amount of shuffled data is however limited to
+ program, where the final result is obtained from all the input blocks (see
+ Fig.~\ref{fig:tg-histo}). This application requires data shuffling, incurring 
+ additional performance costs compared to Algorithm~\ref{alg:incrementation}.
+ The total amount of shuffled data is, however, limited to
  \SI{2.62}{\mega\byte} per blocks as it only consists of image histograms.
@@
 We used the \href{https://github.com/BIDS-Apps/example}{BIDS app example} application
-available on GitHub. This application operates in a map-reduce way. The map phase,
+available on GitHub. This application can also be broken down into a map and
+reduce phase. The map phase,
 also called participant analysis, extracts the brain tissues of a subject's 3D MRI

### 2019-07-26 latex/paper.tex
Modified results section + Positioned the figures

@@
 \usepackage{subcaption}
+\usepackage{dblfloatfix}
 
@@
 bypass the GIL whenever possible and use multithreading to provide parallelism.
-With the exception of Dask Bags, all APIs, by default, use the local multithreaded
-scheduler. Dask Bags instead relies on the local multiprocessing scheduler.
+With the exception of Dask Bag, all APIs, by default, use the local multithreaded
+scheduler. Dask Bag instead relies on the local multiprocessing scheduler.
 
@@
     \centering
-    \includegraphics[width=0.125\textwidth,
+    \includegraphics[width=2cm, height=\columnwidth,
     angle=-90]{images/incrementation-task-graph.png}
@@
     Circles represent the incrementation and sleep function while rectangles
-    represent stages of the image blocks.}\label{fig:tg-inc}
+    represent stages of the blocks.}\label{fig:tg-inc}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.16\textwidth, angle=-90]{images/histogram-task-graph.png}
+    \includegraphics[width=2cm, height=\columnwidth,
+    angle=-90]{images/histogram-task-graph.png}
     \caption{Task graph for Histogram with 3 BigBrain blocks. Circles represent the
-    functions while rectangles represent stages of the image blocks.
+    functions while rectangles represent stages of the blocks.
     }\label{fig:tg-histo}
@@
 
-While Incrementation and Histogram were implemented directly in Python, BIDS app
+While Incrementation and Histogram were implemented directly in Python, BIDS App
 example requires an external library and is distributed as a Docker container image
@@
 data read, compute, data write, and engine overhead across all application tasks. 
-We measured the overhead by \TG{\ldots, maybe add a figure}. \MD{Done but might need
-adjustment depending on how the overhead is calculate (idle vs wasted)}
+We measured the overhead by
 
-\TG{explain that number of iterations may impact futures.}
+% \TG{\ldots, maybe add a figure}. \MD{Done but might need
+% adjustment depending on how the overhead is calculate (idle vs wasted)}
 
-% The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
-% seconds sleep delay.
+% \TG{explain that number of iterations may impact futures.}
+
+The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
+seconds sleep delay.
 
@@
 Figure~\ref{fig:inc_ms_worker} shows the makespan of the Incrementation application
-for different numbers of workers and for different engines. The bars show the average
+for different numbers of workers and engines. The bars show the average
 makespan over 3 repetitions while the error bars are the standard deviation. Overall,
@@
 slight advantage over Spark, \SI{83.61}{\second} on average,
-with Delayed and Futures being slightly better than Bags.
+with Delayed and Futures being slightly better than Bag.
 
 For all engines, the makespan is far from decreasing linearly with the
-number of workers. Note that there are 8 threads per worker. The makespan
+number of workers. Note that each worker has 8 CPUs. The makespan
 even increases between 4 and 8 workers. This is due to the high impact of
@@
 transfer time and overhead increase proportionally to the number of
-workers \TG{check if indeed proprotional, add regression slopes}.
+workers with regression lines:
+
+\hspace{1cm}\(\hat{y}=2336.9499\cdot{x} + 325.00574\) (Spark)
+
+\hspace{1cm}\(\hat{y}=2650.13505\cdot{x} + 1585.37218\) (Bag)
+
+\hspace{1cm}\(\hat{y}=3250.53406\cdot{x} + 1769.50351\) (Delayed)
+
+\hspace{1cm}\(\hat{y}=3570.39758\cdot{x} + 1062.23031\) (Futures)
 
@@
 slightly lower than Dask's in particular as the number of workers increase.
-Within Dask, Delayed and Futures have a higher overhead than Bags. However,
+Within Dask, Delayed and Futures have a higher overhead than Bag. However,
 overhead differences are compensated by an increase in data transfer time,
@@
     \caption{125 blocks, 10 iterations, \SI{4}{\second} delay, 8
-    threads/worker}\label{fig:inc_worker}
+    CPUs/worker}\label{fig:inc_worker}
 \end{figure}
@@
 In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
-For 30 blocks the Dask Bag API has a much lower overhead time. This is because Bag
-was only using one thread per block in comparison to Delayed and Futures that were
-offloading block calculations on multiple threads of the same worker. Once again,
-the data transfer time reduces with more blocks however the overhead time increases
-by a similar amount. This is not observed for 30 blocks as the workers are not used
-at full capacity, i.e., some threads are idle. Finally, the variability of the
-overhead increases with the number of blocks, which explains the makespan variability
-mentioned previously.
+For 30 blocks, Delayed and Futres have much higher overhead. This is because they are
+thread-based and use multiple threads to compute a block which affect the calculation
+of the overhead. Once again, the data transfer time reduces with more blocks however
+the overhead time increases by a similar amount. This is not observed for 30 blocks
+as the workers are not used at full capacity, i.e., some threads are idle. Finally,
+the variability of the overhead increases with the number of blocks, which explains
+the makespan variability mentioned previously.
 
@@
     \caption{10 iterations, \SI{4}{\second} delay, 8 workers, 8
-    threads/worker}\label{fig:inc_block}
+    CPUs/worker}\label{fig:inc_block}
 \end{figure}
@@
 varying the number of iterations. Overall,  Spark and Dask APIs are once
-again equivalent, although Dask Delayed and Futures are slightly faster
-than Bags and, Spark for 1 and 10 few iterations, and Futures are faster than
-Delayed, Bags and Spark for 100 iterations. Differences remain minor
+again equivalent, although Delayed and Futures are slightly faster
+than Bag and, Spark for 1 and 10 iterations, and Futures are faster than
+Delayed, Bag and Spark for 100 iterations. Differences remain minor
 though.
@@
 iterations.
-% TODO
-
-% We think this is because Dask Delayed often
-% schedule a block on a different thread of the same worker which causes small delays
-% every time due to inter-thread data communication. This is good when there are fewer
-% tasks as is make the IO between blocks go out of sync hence lowering the IO
-% bottleneck however as the number of tasks increases those small delays become
-% significant.
-
-%  The Dask Futures API seems to outperform all the other APIs but this is
-% most likely because the tasks are not interdependent thus this application benefits
-% from the less optimal but faster scheduling brought by Dask Futures.
 
@@
     \end{subfigure}
-    \caption{125 blocks, \SI{4}{\second} delay, 8 workers, 8 threads/worker}
+    \caption{125 blocks, \SI{4}{\second} delay, 8 workers, 8 CPUs/worker}
 \end{figure}
@@
 
-Figure~\ref{fig:inc_tt_sleep} shows the total execution time breakdown. On
-one hand, Spark has the smallest overhead. As previously observed,
-variations in overhead time are almost exactly compensated by variations in data
-transfer time.
+Figure~\ref{fig:inc_tt_sleep} shows the total execution time breakdown. Spark has the
+smallest overhead. As previously observed, variations in overhead time are almost
+exactly compensated by variations in data transfer time.
 
-% \begin{table}[!t]
-%     \caption{Function time for the sleep experiment}
-%     \begin{subtable}[b]{\columnwidth}
-%         \renewcommand{\arraystretch}{1.3}
-%         \caption{Overhead time in second}\label{tab:inc_sleep_overhead}
-%         \centering
-%         \begin{tabular}{lllll}
-%         \hline
-%                      & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
-%         Spark        & 14992  & 17076  & 16069   & 9769    \\
-%         Dask.Bag     & 20690  & 21388  & 17759   & 18720   \\
-%         Dask.Delayed & 27481  & 26199  & 22080   & 24624   \\
-%         Dask.Futures & 27069  & 27984  & 23538   & 24762   \\ \hline
-%         \end{tabular}
-%     \end{subtable}
-%     \vskip 0.2cm
-%     \begin{subtable}[b]{\columnwidth}
-%         \renewcommand{\arraystretch}{1.3}
-%         \caption{IO time in second}\label{tab:inc_sleep_io}
-%         \centering
-%         \begin{tabular}{lllll}
-%         \hline
-%                      & 1 sec. & 4 sec. & 16 sec. & 64 sec. \\ \hline
-%         Spark        & 63622  & 57144  & 51070   & 57078   \\
-%         Dask.Bag     & 54630  & 52312  & 53743   & 53612   \\
-%         Dask.Delayed & 45239  & 44142  & 45133   & 46198   \\
-%         Dask.Futures & 46537  & 43161  & 44749   & 45846   \\ \hline
-%         \end{tabular}
-%     \end{subtable}
-%     \vspace{-3mm}
-%  \end{table}
-
-\begin{figure}[!t]
+\begin{figure}[!b]
     \centering
@@
     \end{subfigure}
-    \caption{125 blocks, 10 iterations, 8 workers, 8 threads/worker}
+    \caption{125 blocks, 10 iterations, 8 workers, 8 CPUs/worker}
 \end{figure}
@@
 
-Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine and API.
-Gantt charts are structured in batches of up to 64 read-compute-write concurrent
-sequences. File reads in the first batch are much longer than in the following ones:
+Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine. Gantt
+charts are structured in batches of up to 64 read-compute-write concurrent sequences.
+Files read in the first batch are much longer than the ones in the following batches:
 this is due to the high synchronization of data transfers that leads to a high
 saturation of the shared file system. We also note that overhead, represented in
-white, is concentrated around the data transfer tasks \TG{find out why}, and the
+white, is concentrated around the data transfer tasks, and the
 computing tasks that run concurrently with data transfers.
-% The frameworks have similar makespan however the time spend for each type of
-% task differs considerably (see Table~\ref{tab:inc_base}). Spark tends to spend more
-% time than other frameworks for IO however it has a much lower overhead. Dask Delayed
-% and Futures have the lowest IO time but a much higher overhead. Dask Bag stands in
-% the middle for both IO and overhead. Compute time is approximately the same for all
-% frameworks. Apart from the first IO section, there is a large amount of ovehead
-% everytime time IO is performed. This is true for all frameworks.
 
-\begin{figure*}[!htb]
+\begin{figure}[!b]
     \centering
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/spark_inc_gantt.png}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/spark_inc_gantt.png}
         \caption{Spark}\label{fig:inc_spark_gantt}
     \end{subfigure}
-    \hfill
+    \\
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/bag_inc_gantt.png}%
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/bag_inc_gantt.png}%
         \caption{Dask Bag}\label{fig:inc_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/delayed_inc_gantt.png}%
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/delayed_inc_gantt.png}%
         \caption{Dask Delayed}\label{fig:inc_dask_delayed_gantt}
     \end{subfigure}
-    \hfill
+    \\
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/futures_inc_gantt.png}%
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/futures_inc_gantt.png}%
         \caption{Dask Futures}\label{fig:inc_dask_futures_gantt}
     \end{subfigure}
-    \caption{Incrementation: Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second}
-delay, 8 workers, 8 threads/worker}\label{fig:inc_gantt}
-\end{figure*}
-
-% \begin{table}[!b]
-%     \renewcommand{\arraystretch}{1.3}
-%     \caption{Distribution of the execution time in second: 125 blocks, 4 sec.\ sleep
-%     delay, 10 iterations, and 8 workers}\label{tab:inc_base}
-%     \centering
-%     \begin{tabular*}{\columnwidth}{llllll}
-%     \hline
-%                  & Read  & Compute & Write & Overhead & Total \\ \hline
-%     Spark        & 34465 & 5118    & 22679 & 17076    & 79338 \\
-%     Dask Bag     & 36863 & 5121    & 15449 & 21388    & 78821 \\
-%     Dask Delayed & 32769 & 5119    & 11373 & 26199    & 75461 \\
-%     Dask Futures & 31891 & 5120    & 11270 & 27984    & 76265 \\ \hline
-%     \end{tabular*}
-%  \end{table}
+    \caption{Incrementation: 125 blocks, 10 iterations, \SI{4}{\second} delay, 8 workers,
+8 CPUs/worker}\label{fig:inc_gantt}
+\end{figure}
 
@@
 amounts of workers. Spark is tremendously faster than Dask APIs. The difference
-narrows as the number of workers increases. Between Dask APIs there is no substantial
-difference, however, Bag is slightly faster on average.
+narrows as the number of workers increases but stays significant. Between Dask APIs
+there is no substantial difference, however, Bag is slightly faster on average.
 
-From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is shown.
-The computation time is significantly larger in Dask than Spark. This is potentially
-due to Python's GIL preventing Dask to parallelize the computation on multiple threads.
-Overall, the IO and overhead are comparable for all engines.
+From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is
+shown. The computation time is significantly larger for Dask than for Spark. This is
+potentially due to Python's GIL preventing Dask to parallelize the computation on
+multiple threads. Overall, the IO and overhead are comparable for all engines.
 
@@
     \end{subfigure}
-    \caption{125 blocks, 8 threads/worker}
+    \caption{125 blocks, 8 CPUs/worker}
 \end{figure}
@@
 \subsection{Histogram: Number of blocks}
-Figure~\ref{fig:histo_ms_block} shows the makespan for different block size.
-Again, Spark is significantly faster then Dask engines and Bag is slightly faster
+Figure~\ref{fig:histo_ms_block} shows the makespan for different block size. Again,
+Spark is not compared at 30 blocks due to its \SI{2}{\giga\byte} task limitation.
+Once more, Spark is significantly faster then Dask engines and Bag is slightly faster
 than Delayed. Overall, the engines do not react to change in block size.
@@
 engines. Overall, for all engines, lower block size results in lower IO time but
-increases the overhead time. Again, this is due to the overhead desynchronizing task
+increases the overhead time. Again, this is due to large overhead desynchronizing task
 thus reducing the NFS bottleneck.
-%TODO
 
@@
     \end{subfigure}
-    \caption{8 workers, 8 threads/worker}
+    \caption{8 workers, 8 CPUs/worker}
 \end{figure}
-\TG{We decided to omit Dask Futures as it does not have any value over Dask
-Delayed in this application.}
+% \TG{We decided to omit Dask Futures as it does not have any value over Dask
+% Delayed in this application.}
 
@@
 
-The reduce is done differently on the engines. While Spark and Delayed compute all
-the reduce tasks on one process (thread) Spark computes it on the driver process. On
-the other hand, Bag computes the reduce tasks on different workers and sends the
-information to the driver process afterward.
+The reduce is done differently for the engines. Spark and Delayed compute all the
+reduce tasks on one process (thread), however, Spark only computes it on the driver
+process. On the other hand, Bag computes the reduce tasks on different workers and
+sends the information to the driver process afterward.
 
-\begin{figure}[!htb]
+\begin{figure}[!t]
     \centering
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/spark_histo_gantt.png}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/spark_histo_gantt.png}
         \caption{Spark}\label{fig:histo_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/bag_histo_gantt.png}%
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/bag_histo_gantt.png}%
         \caption{Dask Bag}\label{fig:histo_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/delayed_histo_gantt.png}%
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/delayed_histo_gantt.png}%
         \caption{Dask Delayed}\label{fig:histo_dask_delayed_gantt}
     \end{subfigure}
-    \caption{Histogram: Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second}
-delay, 8 workers, 8 threads/worker}\label{fig:histo_gantt}
+    \caption{Histogram: 125 blocks, 8 workers, 8
+    CPUs/worker}\label{fig:histo_gantt}
 \end{figure}
@@
 Overall, the makespan for all engines does not change considerably with an increase
-in the number of workers. This is because the engines do not scale well.
+in the number of workers. This is because the engines do not scale well. There is no
+substantial difference between all engines, although, Delayed is slightly faster.
 
-On figure~\ref{fig:histo_np_ms_worker}, there is no substantial difference between
- all engines, although, Delayed is slightly faster.
+From Figure~\ref{fig:histo_np_tt_worker}, the total execution time breakdown is
+shown. The overhead and read time increase proportionally to the number of workers
+thus increasing the total time. This is because (1) most workers read at the same
+time due to a low computation-IO ratio and (2) there is more intra-worker and
+inter-worker communication.
 
-From Figure~\ref{fig:histo_np_tt_worker}, the total execution time spent in each
-function is shown. The overhead and read time increase proportionally to the number
-of workers thus increasing the total time. This is because (1) most workers read at
-the same time due to a low computation-IO ratio and (2) \MD{Determine why the
-overhead is so large compared to the pure Python implementation}
-
-\begin{figure}[!t]
+\begin{figure}[!b]
     \centering
@@
 Figure~\ref{fig:histo_np_ms_block} shows the makespan of the application for
-different block size. Note that the workers can process up to 64 tasks in parallel
-thus they are not used at full capacity with 30 blocks. Overall, there is no
-substantial difference for all engines and block size.
+different block size. Again, Spark is not compared at 30 blocks due to its
+\SI{2}{\giga\byte} task limitation. Overall, there is no substantial difference for
+all engines and block sizes.
 
 In figure~\ref{fig:histo_np_tt_block}, shows the total time spent in each function.
-Again, the overhead for 30 blocks is erroneous due to a lower amount of task than
-available threads which result in higher overhead for thread-based engines.
-Considering the overhead error and that at 30 blocks the workers only use half their
-core, the total time is substantially the same for all block size. This is due to
-lower block sizes having a lower IO time, however, it is balanced by the resulting
-increase in overhead.
+Once more, since Delayed is thread-based its overhead at 30 blocks is seems much
+higher that it really is. Considering the overhead error and that at 30 blocks the
+workers only use half their core, the total time is substantially the same for all
+block size. This is due to lower block sizes having a lower IO time, however, it
+balances out by the resulting increase in overhead.
 
@@
     \end{subfigure}
-    \caption{8 workers, 8 threads/worker}
+    \caption{8 workers, 8 CPUs/worker}
 \end{figure}
@@
 \subsection{Histogram (NumPy): Gantt chart}
-Figure~\ref{fig:histo_np_gantt} shows the Gantt chart of the baseline Histogram
-(NumPy) experiment. The overhead for Spark and Bag is mostly located between a read
-and compute task or compute and read task. On the other hand, Delayed overhead is
-dispersed between all types of tasks. This causes Delayed to overlap more compute and
-read tasks among the different workers.
+Figure~\ref{fig:histo_np_gantt} shows the Gantt chart of the baseline for the
+Histogram (NumPy) experiment. The overhead for Spark and Bag is mostly located
+between a read and compute task or compute and read task. On the other hand, Delayed
+overhead is dispersed between all types of tasks. This causes Delayed to overlap more
+compute and read tasks simultaneously among the different workers.
 
-In figure~\ref{fig:histo_np_gantt}, The read tasks of all engines are similar initially, however, when computation starts
-the subsequent read tasks are much shorter. This is because read tasks for Delayed
-become desynchronized hence it significantly reduces the IO bottleneck.
+In figure~\ref{fig:histo_np_gantt}, read tasks are initially similar for all engines.
+However, when computation starts the subsequent read tasks are much shorter for
+Delayed. This is due the desynchronization of the task reducing the IO bottleneck.
 
-\begin{figure}[!htb]
+\begin{figure}[!tb]
     \centering
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/spark_histo_np_gantt.png}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/spark_histo_np_gantt.png}
         \caption{Spark}\label{fig:histo_np_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/bag_histo_np_gantt.png}%
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/bag_histo_np_gantt.png}%
         \caption{Dask Bag}\label{fig:histo_np_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/delayed_histo_np_gantt.png}%
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/delayed_histo_np_gantt.png}%
         \caption{Dask Delayed}\label{fig:histo_np_dask_delayed_gantt}
     \end{subfigure}
-    \caption{Histogram (NumPy): Gantt chart with 125 blocks, 10 iterations, \SI{4}{\second}
-delay, 8 workers, 8 threads/worker}\label{fig:histo_np_gantt}
+    \caption{Histogram (NumPy): 125 blocks, 8 workers, 8
+    CPUs/worker}\label{fig:histo_np_gantt}
 \end{figure}
@@
 \subsection{BidsApp example: Number of workers}
-Figure~\ref{fig:bids_ms_worker} shows the makespan of the application while varying
-the amount of worker. Overall, there is no substantial difference between the
-engines. The makespan scales well with the increase in workers however the scaling
-reduces considerably when reaching 8 workers. This is due to the IO requirement
-increasing with the number of workers --- more tasks are read concurrently and
-the application becomes more data-intensive.
+Figure~\ref{fig:bids_ms_worker} shows the makespan of the application when varying
+the amount of workers. Overall, there is no substantial difference between the
+engines. The makespan scales well with the increase in workers, however, the scaling
+reduces considerably when reaching 8 workers. This is due to an increase in overhead
+and of the IO requirement increasing with the number of workers --- more tasks are
+read concurrently and the application becomes more data-intensive.
 
 In Figure~\ref{fig:bids_tt_worker}, the total execution time of each function is
-shown. Overall, there is no significant difference between the engines. Each engine
-has a similar total time independently of its number of workers. This is because of
-the, almost, linearly scale of the engines. The overhead increases slowly as the
-number of workers increases. This is due to more communication between the scheduler
-and workers as well as inter-worker communication.
+shown. Each engine has a similar total time independently of its number of workers.
+This is due to the, almost, linear scaling of the engines. The overhead increases
+proportionally with the number of workers (see regression lines below). This is due
+to more communication between the scheduler and workers as well as intra-worker and
+inter-worker communication.
+
+\hspace{1cm}\(\hat{y}=223.53586\cdot{x} + 91.616964\) (Spark)
+
+\hspace{1cm}\(\hat{y}=182.82659\cdot{x} - 11.43834\) (Bag)
+
+\hspace{1cm}\(\hat{y}=172.54183\cdot + 400.69433\) (Delayed)
+
+\hspace{1cm}\(\hat{y}=132.41481\cdot{x} + 90.32813\) (Futures)
 
@@
     \end{subfigure}
-    \caption{Variation of the amount of worker, 8 threads/worker}
+    \caption{Variation of the amount of worker, 8 CPUs/worker}
 \end{figure}
@@
 \subsection{BidsApp example: Gantt chart}
-Figure~\ref{fig:bids_gantt} shows the Gantt charts obtained for each engine and API.
-The Gantt charts are structured in two main parts: participant analysis (orange) and
+Figure~\ref{fig:bids_gantt} shows the Gantt chart obtained for each engine.
+Gantt charts are structured in two parts: participant analysis (orange) and
 group analysis (blue). The participant analysis tasks differ greatly in length. This
-is due to the unequal amount of sessions per subject to process. The group analysis
+is due to the unequal amount of sessions per subject processed. The group analysis
 is similar for all engines and APIs. Overall, most of the overhead encountered
@@
 
-\begin{figure*}[!htb]
+\begin{figure}[!htb]
     \centering
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/spark_bids_gantt.png}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/spark_bids_gantt.png}
         \caption{Spark}\label{fig:bids_spark_gantt}
     \end{subfigure}
-    \hfill
+    \\
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/bag_bids_gantt.png}%
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/bag_bids_gantt.png}%
         \caption{Dask Bag}\label{fig:bids_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/delayed_bids_gantt.png}%
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/delayed_bids_gantt.png}%
         \caption{Dask Delayed}\label{fig:bids_dask_delayed_gantt}
     \end{subfigure}
-    \hfill
+    \\
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/futures_bids_gantt.png}%
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/futures_bids_gantt.png}%
         \caption{Dask Futures}\label{fig:bids_dask_futures_gantt}
     \end{subfigure}
-    \caption{BIDS App example: Gantt chart with 8 workers}\label{fig:bids_gantt}
-\end{figure*}
+    \caption{BIDS App example: 8 workers}\label{fig:bids_gantt}
+\end{figure}
 
@@
 \section{Discussion}
+
+\subsection{Effect of Network File System}
+
+\subsection{Effect of data transfer}
+
+\subsection{Effect of the Global Interpreter Lock}
+
+\subsection{Effect of the number of partitions}
 
@@
 as reducing them increases data trasnfers.}
-
-% partitioning issue
-
-% \subsection{IO Time}
-% 
-% 
-% \subsection{Serialization}
-% We think that serialization could have an important effect on the makespan of the
-% application. This could potentially slow down the application when a lot of tasks are
-% scheduled.
-% 
-% \subsection{NFS}
-% The NFS seems to be a source of the bottleneck. We think that exploring another file
-% system, like Lustre, could give us insight into the actual effect of the NFS on the
-% application makespan. We also think that the NFS caching could play an effect on the
-% task of different lengths.
-% 
-% \subsection{Scheduler}
-% The Dask scheduler seems to have some bizarre behavior. It seems like it waits to
-% schedule tasks. It could also be due to tasks being scheduled on other workers
-% thus requiring data to be sent over the network; which would explain the stall on the
-% worker. More work would be needed to investigate that issue.
-% 
-% 
-% \subsection{Caching}
-% We think that caching plays an effect on the results. As seen in the Gantt chart
-% previously, some of the read tasks are significantly shorter than others while all
-% blocks are of equal size.
 

### 2019-07-27 latex/paper.tex

@@
 \usepackage{subcaption}
-\usepackage{dblfloatfix}
 
@@
  \SI{2.62}{\mega\byte} per blocks as it only consists of image histograms.
+Two implementation are studied: pure Python and NumPy.
 
@@
 data read, compute, data write, and engine overhead across all application tasks. 
-We measured the overhead by
+The overhead calculation for each CPU is the end time of its last processed task
+minus the start time of the application; the summation of those results is the total
+overhead.
 
-% \TG{\ldots, maybe add a figure}. \MD{Done but might need
-% adjustment depending on how the overhead is calculate (idle vs wasted)}
-
-% \TG{explain that number of iterations may impact futures.}
-
-The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
-seconds sleep delay.
+% The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
+% seconds sleep delay.
 

### 2019-07-27 latex/paper.tex
typo
Worked on Discussion section

@@
 Spark is not compared at 30 blocks due to its \SI{2}{\giga\byte} task limitation.
-Once more, Spark is significantly faster then Dask engines and Bag is slightly faster
+Once more, Spark is significantly faster than Dask engines and Bag is slightly faster
 than Delayed. Overall, the engines do not react to change in block size.
@@
 \subsection{Effect of Network File System}
+We used an NFS as it is a common approach to store a large amount of data. However,
+it limits the available bandwidth since all workers interact with one disk. This
+causes data-intensive applications to take a significant amount of time for IO. As we
+expect, this effect is amplified when too many tasks simultaneously try to operate
+with the NFS at the same time as
+Fig.~\ref{fig:inc_gantt},\ref{fig:histo_gantt},\ref{fig:histo_np_gantt},\ref{fig:bids_gantt}
+shows. We could use a parallel shared file system, like Lustre \MD{cite lustre if not
+done already}, to reduce the impact of operations with the disk, however, we would
+quickly be limited by the network's bandwidth.
 
-\subsection{Effect of data transfer}
+\subsection{Effect of data-intensive applications}
+In our experiments, Incrementation, Histogram (NumPy) and Histogram for Spark have
+low compute time while BIDS App example and Histogram for Dask have significantly
+longer compute time. When adding additional workers, we observe that data-intensive
+applications scale poorly while compute-intensive applications scale almost linearly
+This is due to the limitation from the NFS, as mentioned above.
+
+Data transfer accounts for the majority of the execution time for data-intensive
+applications (see
+Fig.~\ref{fig:inc_gantt},\ref{fig:histo_spark_gantt},\ref{fig:histo_np_gantt}). We
+see that when a low number of IO tasks are overlapping, this diminishes the bandwidth
+limitation from the NFS. We want to explore solutions to desynchronize IO tasks in
+the aim to reduce the makespan of data-intensive applications.
 
 \subsection{Effect of the Global Interpreter Lock}
+Python's GIL prevents multi-threading in Python. Dask tries to minimize the
+consequences of the GIL by using external Python extensions. We found that Dask has
+issues to parallelize Python functions while Spark did not (see
+Fig.~\ref{fig:histo_tt_worker}). However, Dask was easily able to parallelize
+application with functions in Numpy (see
+Fig.~\ref{fig:inc_tt_worker},\ref{fig:histo_np_tt_worker}) and command-line
+applications (see Fig.~\ref{fig:bids_tt_worker}). This is because, for Dask, the
+instances are configured to have 1 worker process with 8 threads. Since the Histogram
+implementation does not release the GIL it can only use one of the thread at a given
+time thus it slows down the computation significantly.
 
-\subsection{Effect of the number of partitions}
+One should be careful to configure the workers properly depending on the type
+of functions executed in an application.
+
+\subsection{Effect of tasks' size}
+We measured the effect of various block sizes for constant BigBrain image for
+Incrementation, Histogram, and Histogram (NumPy). We observe that larger block size
+results in more IO compensated by lower overhead time. On the other hand, smaller
+block size results in higher overhead time compensated by reduced IO time.
+
+The number of partitions also has an impact on the size of the tasks. We set a
+partition for each block or subject. There is a trade-off for the number of
+partitions. On the one hand, a low amount of partition results in larger tasks thus
+in lower scheduling overhead. On the other hand, a higher number of partition
+results in smaller tasks hence more tasks fit in the workers' memory insuring that
+they always have something to work on.
+
+Varying the block size is equivalent to varying the size of the partitions. We
+observed no difference when varying the task size while using an NFS. However, if
+one uses a parallel shared file system, we suppose it would be preferable to use a
+larger block size as the IO time should be reduced substantially. It is worth noting
+that the Dask developers suggest, in
+\href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to
+use a partition size such that workers have 2-3 times as many chunks available to
+work on. More experiments would have to be performed to gain insight on an optimal
+number of partitions to use for a given set of data-intensive applications.
+
+\subsection{Other comments}
+Spark performance should be reduced by the serialization of Python functions to Java,
+however, our results show no significant impact of it.
 

### 2019-07-27 histogram-np/dask-array_histogram.py
Implementation of Histogram (NumPy)

@@ -1,0 +1,129 @@
+import sys
+
+sys.path.append("/nfs/paper-big-data-engines")
+
+import argparse
+from io import BytesIO
+import os
+from time import time
+
+import dask
+import dask.array as da
+from dask.distributed import Client
+import nibabel as nib
+import numpy as np
+
+from utils import benchmark, crawl_dir
+
+
+def get_voxels(filename, start, args):
+    """Retrieve voxel intensity of a Nifti image as a byte stream.
+
+    Parameters
+    ----------
+    filename: str
+        Representation of the path for the input file.
+
+    Returns
+    -------
+    data : da.array
+        Data of the nifti imaeg read.
+    """
+    start_time = time() - start
+
+    img = None
+    with open(filename, "rb") as f_in:
+        fh = nib.FileHolder(fileobj=BytesIO(f_in.read()))
+        img = nib.Nifti1Image.from_file_map({"header": fh, "image": fh})
+    data = img.get_fdata(caching="unchanged")
+    data = nib.casting.float_to_int(data, np.int16)
+
+    end_time = time() - start
+
+    if args.benchmark:
+        benchmark(
+            start_time,
+            end_time,
+            filename,
+            args.output_dir,
+            args.experiment,
+            get_voxels.__name__,
+        )
+
+    return da.from_array(data)
+
+
+if __name__ == "__main__":
+
+    start = time()  # Start time of the pipeline
+
+    parser = argparse.ArgumentParser(description="BigBrain Kmeans")
+    parser.add_argument("scheduler", type=str, help="Scheduler ip and port")
+    parser.add_argument(
+        "bb_dir",
+        type=str,
+        help=("The folder containing BigBrain NIfTI images" "(local fs only)"),
+    )
+    parser.add_argument(
+        "output_dir",
+        type=str,
+        help=("the folder to save incremented images to" "(local fs only)"),
+    )
+    parser.add_argument(
+        "experiment", type=str, help="Name of the experiment being performed"
+    )
+    parser.add_argument("--benchmark", action="store_true", help="benchmark results")
+
+    args = parser.parse_args()
+
+    # Cluster scheduler
+    cluster = args.scheduler
+    client = Client(cluster)
+
+    print(client)
+    # Allow workers to use module
+    client.upload_file("/nfs/paper-big-data-engines/utils.py")
+
+    # Read images
+    paths = crawl_dir(os.path.abspath(args.bb_dir))
+
[…]

### 2019-07-27 histogram-np/dask-bag_histogram.py
Implementation of Histogram (NumPy)

@@ -1,0 +1,66 @@
+import argparse
+import os
+import sys
+from time import time
+
+import dask.bag as db
+from dask.distributed import Client
+
+sys.path.append("/nfs/paper-big-data-engines/histogram")
+
+
+if __name__ == "__main__":
+
+    start = time()  # Start time of the pipeline
+
+    parser = argparse.ArgumentParser(description="BigBrain Kmeans")
+    parser.add_argument("scheduler", type=str, help="Scheduler ip and port")
+    parser.add_argument(
+        "bb_dir",
+        type=str,
+        help=("The folder containing BigBrain NIfTI images" "(local fs only)"),
+    )
+    parser.add_argument(
+        "output_dir",
+        type=str,
+        help=("the folder to save incremented images to" "(local fs only)"),
+    )
+    parser.add_argument(
+        "experiment", type=str, help="Name of the experiment being performed"
+    )
+    parser.add_argument("--benchmark", action="store_true", help="benchmark results")
+
+    args = parser.parse_args()
+
+    # Cluster scheduler
+    cluster = args.scheduler
+    client = Client(cluster)
+
+    print(client)
+    # Allow workers to use module
+    client.upload_file("/nfs/paper-big-data-engines/utils.py")
+    client.upload_file("/nfs/paper-big-data-engines/histogram/Histogram.py")
+    from utils import benchmark, crawl_dir, read_img
+    from Histogram import (
+        calculate_histogram,
+        combine_histogram,
+        flatten,
+        save_histogram,
+    )
+
+    # Read images
+    paths = crawl_dir(os.path.abspath(args.bb_dir))
+    paths = db.from_sequence(paths, npartitions=len(paths))
+    img = paths.map(lambda p: read_img(p, start=start, args=args))
+
+    img = img.map(lambda x: flatten(x[1], start=start, args=args, filename=x[0]))
+
+    partial_histogram = img.map(
+        lambda x: calculate_histogram(x[1], args=args, start=start, filename=x[0])
+    )
+
+    histogram = partial_histogram.fold(
+        lambda x, y: combine_histogram(x, y, args=args, start=start)
+    ).compute()
+
+    save_histogram(histogram, args=args, start=start)

### 2019-07-27 histogram-np/dask-delayed_histogram.py

@@ -1,0 +1,76 @@
+import argparse
+from functools import reduce
+import os
+import sys
+from time import time
+
+import dask
+from dask.distributed import Client
+
+sys.path.append("/nfs/paper-big-data-engines/histogram")
+
+
+if __name__ == "__main__":
+
+    start = time()  # Start time of the pipeline
+
+    parser = argparse.ArgumentParser(description="BigBrain Kmeans")
+    parser.add_argument("scheduler", type=str, help="Scheduler ip and port")
+    parser.add_argument(
+        "bb_dir",
+        type=str,
+        help=("The folder containing BigBrain NIfTI images" "(local fs only)"),
+    )
+    parser.add_argument(
+        "output_dir",
+        type=str,
+        help=("the folder to save incremented images to" "(local fs only)"),
+    )
+    parser.add_argument(
+        "experiment", type=str, help="Name of the experiment being performed"
+    )
+    parser.add_argument("--benchmark", action="store_true", help="benchmark results")
+
+    args = parser.parse_args()
+
+    # Cluster scheduler
+    cluster = args.scheduler
+    client = Client(cluster)
+
+    print(client)
+    # Allow workers to use module
+    client.upload_file("/nfs/paper-big-data-engines/utils.py")
+    client.upload_file("/nfs/paper-big-data-engines/histogram/Histogram.py")
+    from utils import benchmark, crawl_dir, read_img
+    from Histogram import (
+        calculate_histogram,
+        combine_histogram,
+        flatten,
+        save_histogram,
+    )
+
+    # Read images
+    paths = crawl_dir(os.path.abspath(args.bb_dir))
+
+    partial_histogram = []
+    for path in paths:
+        img = dask.delayed(read_img)(path, start=start, args=args)
+
+        img = dask.delayed(flatten)(img[1], start=start, args=args, filename=img[0])
+
+        partial_histogram.append(
+            dask.delayed(calculate_histogram)(
+                img[1], args=args, start=start, filename=img[0]
+            )
+        )
+
+    histogram = dask.delayed(reduce)(
+        lambda x, y: combine_histogram(x, y, args=args, start=start), partial_histogram
+    )
+
+    future = client.compute(histogram)
+    histogram = client.gather(future)
+
+    save_histogram(histogram, args=args, start=start)
+
+    client.close()

### 2019-07-27 histogram-np/spark_histogram.py

@@ -1,0 +1,68 @@
+import argparse
+import os
+import sys
+from time import time
+
+import numpy as np
+from pyspark import SparkConf, SparkContext
+
+sys.path.append("/nfs/paper-big-data-engines/histogram")
+
+
+if __name__ == "__main__":
+
+    start = time()  # Start time of the pipeline
+
+    parser = argparse.ArgumentParser(description="BigBrain Kmeans")
+    parser.add_argument(
+        "bb_dir",
+        type=str,
+        help=("The folder containing BigBrain NIfTI images" "(local fs only)"),
+    )
+    parser.add_argument(
+        "output_dir",
+        type=str,
+        help=("the folder to save incremented images to" "(local fs only)"),
+    )
+    parser.add_argument(
+        "experiment", type=str, help="Name of the experiment being performed"
+    )
+    parser.add_argument("--benchmark", action="store_true", help="benchmark results")
+
+    args = parser.parse_args()
+
+    # Cluster scheduler
+    conf = SparkConf().setAppName(f"Spark {args.experiment}")
+    sc = SparkContext.getOrCreate(conf=conf)
+
+    sc.addFile("/nfs/paper-big-data-engines/utils.py")
+    sc.addFile("/nfs/paper-big-data-engines/histogram/Histogram.py")
+    from utils import benchmark, crawl_dir, read_img
+    from Histogram import (
+        calculate_histogram,
+        combine_histogram,
+        flatten,
+        save_histogram,
+    )
+
+    print("Connected")
+
+    # Read images
+    paths = crawl_dir(os.path.abspath(args.bb_dir))
+    paths = sc.parallelize(paths, len(paths))
+    img_rdd = paths.map(lambda p: read_img(p, start=start, args=args))
+
+    img_rdd = img_rdd.map(
+        lambda x: flatten(x[1], start=start, args=args, filename=x[0])
+    )
+
+    partial_histogram = img_rdd.map(
+        lambda x: calculate_histogram(x[1], args=args, start=start, filename=x[0])
+    )
+
+    histogram = partial_histogram.fold(
+        np.array([0] * (2**16 - 1)),
+        lambda x, y: combine_histogram(x, y, args=args, start=start),
+    )
+
+    save_histogram(histogram, args=args, start=start)

### 2019-07-28 latex/paper.tex

@@
 \usepackage{subcaption}
+
+%\setlength{\textfloatsep}{7pt}
 
@@
 % Topic
-As the amount of data increases and is easier to access, Big Data processing becomes
-critical in neuroimaging.
+Recent increases in data volume and improved accessibility have 
+led neuroimaging into the Big Data era. 
 % Problem
-This is an issue as the currently used frameworks are specialized for neuroimaging
-and not Big Data concerns. This can lead to performance decrease or limit the
-research we can perform.
+While existing neuroimaging pipeline frameworks typically do not attempt to
+limit data transfers, Big Data frameworks remain seldom used.
 % Relevance
-The usage of Big Data framework is beneficial to this problem. The state-of-the-art
-general purpose Big Data framework Spark has its code base written in Scala, while
-our laboratory mostly uses Python. This can be problematic since it is harder to port
-our pipelines to the framework. Also, theoretically, it leads to performance decrease.
-
+In this paper, we compare two popular Big Data frameworks with Python APIs,
+Apache Spark and Dask, for their runtime performance in processing neuroimaging
+pipelines.
 % Approach
-We propose to use Dask since it provides native Python parallelism to pipelines while
-providing support to familiar API from the Python's scientific ecosystem.
-Unfortunately, there are few comparisons between Spark and Dask; especially in
-neuroimaging. Moreover, these studies were done when Dask was still an immature
-framework which makes those studies unfair nowadays. This is our motivation to
-compare the latest version of Dask with Spark.
+To evaluate performance differences we used two synthetic, Incrementation
+(map-only) and Histogram (map-reduce), pipelines and real neuroimaging map-reduce pipeline (BIDS App Example). For both our 
+synthetic pipelines we varied the number of workers and block size using the 
+76~GB BigBrain image. We additionally varied the number of iterations and task
+duration for our Incrementation pipeline. 39~GB of anatomical images from the
+CoRR were used for the BIDS App Example pipeline.
 % Methods (one sentence)
-To evaluate the frameworks, we focus on their performance and their scheduler.
-
+All experiments were executed on 8 instances of Compute Canada's Arbutus cloud
+with 8~vCPUs and 30 GB of RAM all connected to a 2~TB shared NFS. We evaluated
+PySpark's RDD against Dask's Bag, Delayed and Futures.
 % Key Impact
-Our study demonstrates the potential of using Dask as the framework to build
-neuroimaging pipelines.
+Our result show that despite the differences between the two frameworks, both
+frameworks perform the comparably. However, Dask pipelines risk being limited by
+Python's GIL depending on task type and cluster configuration. In all cases, 
+the major limiting factor was data transfer. While either framework is suitable
+for neuroimaging pipelines, more effort needs to be placed in reducing data 
+transfer time.
 \end{abstract}
@@
 also provide rich, high-level programming APIs, and support a variety of
-infrastructure schedulers, such as Mesos, YARN, or HPC clusters (Dask only). Over
+infrastructure schedulers, such as Mesos~\cite{hindman2011mesos}, YARN~\cite{vavilapalli2013apache}, or HPC clusters (Dask only). Over
 these similarities, the frameworks have differences.
@@
 
-\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data engine.
-While its core is written in Scala, Spark provides additional APIs in Java, R
-and Python (PySpark). Its main abstraction, the Resilient Distributed Dataset
+\subsubsection{Apache Spark} Apache Spark is a widely-used 
+general-purpose Big Data engine.
+Its main abstraction, the Resilient Distributed Dataset
 (RDD)~\cite{RDD}, is a fault-tolerant, parallel collection of data elements. It 
-achieves fault tolerance through recording the data lineage, the sequence of 
-operations required to produce the data. The RDD is the basis of Spark's other
-data structures, namely, DataFrames
-and Datasets. Datasets are similar to RDDs but additionally use Spark SQL's
+achieves fault tolerance through the recording of data lineage, the
+sequence of operations used to modify the original data. The RDD is the basis
+of Spark's other data structures, namely, Datasets
+and DataFrames. Datasets are similar to RDDs but additionally use Spark SQL's
 optimized execution to further improve performance. DataFrames, used to process
-tabular data, Datasets where the data is organized into named-columns. While 
-the DataFrame API exists in all the available language APIs, Datasets are
+tabular data, are Datasets where the data is organized into named-columns.
+While the DataFrame API exists in all the available language APIs, Datasets are
 limited to Scala and Java.
 
-As data transfers in BigData workflows are an important source of performance
-bottlenecks, Spark incorporated
-data locality, and additionally, introduced in-memory computing to the Big Data 
-frameworks. Data locality, introduced by MapReduce~\cite{mapreduce}, schedules
-tasks as close as possible to where the data is storage. In-memory computing ensures
-data is maintained in memory whenever possible, as writting large amounts data to
-slower storage devices may be costly. To reduce any unecessary communication
-and computation, Spark also included lazy evaluation, which builds the entire
-task graph prior to execution to determine what needs to be computed.
+As data transfers in Big Data workflows are an important source of performance
+bottlenecks, Spark incorporates the concepts of data locality and in-memory 
+computing. Data locality, introduced by MapReduce~\cite{dean2008mapreduce},
+schedules
+tasks as close as possible to where the data is storage. In-memory computing
+ensures data is maintained in memory whenever possible, as writing large 
+amounts data to slower storage devices may be costly. To reduce any unnecessary
+communication and computation, Spark also included lazy evaluation, which
+builds the entire task graph prior to execution to determine what needs to be
+computed.
 
+% Didn't mention Kubernetes as it's still experimental
 Spark is compatible with three different schedulers: Spark
-standalone, YARN~\cite{vavilapalli2013apache} and Mesos~\cite{hindman2011mesos}. 
-The Spark standalone is a simple default scheduler built into Spark. YARN is a
-more complex scheduler primarily designed to schedule Hadoop-based workflows. 
-Mesos, on the other hand, is not limited to Hadoop and can be used to schedule
+standalone, YARN and Mesos.
+The Spark standalone scheduler is a simple default scheduler built into Spark.
+In contrast, the YARN scheduler is primarily designed to schedule Hadoop-based 
+workflows, whereas Mesos can be used to schedule
 a variety of different workflows. As scientific researchers would likely be
@@
 \emph{driver} that in turn requests workers to the master and dispatches tasks
-to them. A job is divided into stages to be excuted in a different process onto
-the workers. Each stage's operation is represented as a high-level task in the computation
-graph. The Spark standalone scheduler uses a FIFO (First-In-First-Out) job
-scheduling policy. The Spark standalone scheduler has two
-execution modes: client mode, where the driver runs in a dedicated process outside
-of the Spark cluster, and cluster mode, where the driver runs in a worker. We used
-client mode as cluster mode is not available in PySpark. We used Apache Spark v2.4.0.
+to them. A job is divided into stages to be executed in a different process onto
+the workers. Each stage's operation is represented as a high-level task in the
+computation graph. The Spark standalone scheduler uses a FIFO 
+(First-In-First-Out) job scheduling policy. The Spark standalone scheduler has
+two execution modes: client mode, where the driver runs in a dedicated process
+outside of the Spark cluster, and cluster mode, where the driver runs within
+a worker process. Our experiments use client mode as cluster mode is not
+available in PySpark.
 
 Python is commonly selected as the programming language of choice in scientific
-communities, especially for our use case of neuroscience, where numerous specialized
-Python libraries exist to study the data. While serialization of 
-Python to Java may lead to significant overheads, we chose to focus on Spark 
-Python API due to its suitability for neuroimaging research.
+communities, and in particular, for our use case of neuroscience, where
+numerous specialized Python libraries exist to study the data.
+While serialization of Python to Java may lead to significant overheads, we
+chose to focus on Spark Python API due to its suitability for neuroimaging
+research.
 
-\subsubsection{Dask} Dask is a Big Data engine that is becoming increasingly popular
-in the scientific Python ecosystem. Like Spark, Dask provides
-in-memory computing, data locality, lazy evaluation, and fault tolerance by
-means of data lineage. In contrast, Dask's core is purely in Python, eliminating
-the cost of serialization to Java. Furthermore, it is lightweight and modular,
+We used Apache Spark v2.4.0.
+
+\subsubsection{Dask} Dask is a Python-based Big Data engine that is becoming
+increasingly popular in the scientific Python ecosystem. Like Spark, Dask avoids data transfers and needless computation and communication through
+in-memory computing, data locality, lazy evaluation. Dask worklflows can 
+further reduce data transfer costs by leveraging multithreading whenever 
+communication is not bounded by Python's GIL. Dask relies on data 
+lineage to achieve fault tolerance, however, unlike Spark, does not require
+operations to be coarse grained. Furthermore, Dask is lightweight and modular,
 allowing users to only install the components they require.
@@
 composition of
-\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas
-Dataframes} used to process a large amount of tabular data.
-\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports arbitrary tasks
+\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas Dataframes} used to process a large amount of tabular data.
+\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports
+arbitrary tasks
 that do not fit in the Array, DataFrame or Bag APIs. Finally,
-\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to Delayed
-as they support arbitrary tasks, but they operate in real-time rather than lazily.
-All Dask data structure APIs, with the exception of Dask Arrays and Dask
-DataFrames, were used in our experiments.
+\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to
+Delayed in that they support arbitrary tasks, but they operate in real-time
+rather than lazily. With the exception of Dask Bag, all APIs, by default, use
+the local multithreaded scheduler. Dask Bag, instead, relies on the local
+multiprocessing scheduler.  All Dask data structure APIs, with the exception of
+Dask Array and Dask DataFrame, were used in our experiments.
 
-While Python multithreading is limited by the
-\href{https://docs.python.org/3/glossary.html#term-gil}{Global Interpreter Lock (GIL)}
-as a result of CPython's non thread-safe memory management, Dask attempts to
-bypass the GIL whenever possible and use multithreading to provide parallelism.
-With the exception of Dask Bag, all APIs, by default, use the local multithreaded
-scheduler. Dask Bag instead relies on the local multiprocessing scheduler.
-
-The Dask graph is the internal representation of Dask application to be executed 
-by the scheduler. API Operations generate multiple
-small tasks in the computation graph allowing an easier representation of complex
+The Dask graph is the internal representation of Dask application to be
+executed  by the scheduler. API Operations generate multiple small tasks in
+the computation graph allowing an easier representation of complex
 algorithms.
@@
 administrates the resources provided by \textit{workers} in the cluster. The
-scheduler receives jobs from clients and assigns tasks to available workers. Similarly
-to Spark's scheduler, Dask Distributed completes the processing of a graph branch
-before moving along to the next one.
+scheduler receives jobs from clients and assigns tasks to available workers.
+Similarly to Spark's scheduler, Dask Distributed completes the processing of a
+graph branch before moving along to the next one.
 
@@
  for off-heap. We configured the Spark driver to use 25~GB of memory, and used the
- default configuration for the master. We used the default confirguration for worker
- memory management: at 60\% it spill data to disk, and 50\% of that amount is
- reserved for storage and is immune to eviction.
+ default configuration for the master. We used the default configuration for worker
+ memory management: at 60\% it spills data to disk, and 50\% of that amount is
+ reserved for storage that is immune to eviction.
 
@@
  shared with the other instances using the Network File System (NFS) v4.
- This instance was also used for the Spark driver and master, and for the
+ This instance was also used for the Spark driver and master, for the
  Dask scheduler, and for job monitoring with the Spark and Dask user
@@
 functional images of 1,397 subjects acquired in 29 sites. We used all 3491 anatomical
-images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} by image on
+images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} per image on
 average).
@@
 finally writes the result as a NifTI image back to the shared file system.
-This application allows us to study the behavior of the engines when all
+This application allows us to study the behaviour of the engines when all
 inputs are processed independently, in a map-only scenario (see
-Figure~\ref{fig:tg-inc}). This mimics the behavior of analyzing multiple
+Figure~\ref{fig:tg-inc}). This mimics the behaviour of analyzing multiple
 independent subjects in parallel.
@@
  blocks, and finally writes the resulting histogram to the shared file system as a
- single \SI{766}{\kilo\byte} file. This application implements a typical map-reduce
- program, where the final result is obtained from all the input blocks (see
+ single \SI{766}{\kilo\byte} file. It follows the map-reduce
+ paradigm, in which the final result is obtained from all the input blocks (see
  Fig.~\ref{fig:tg-histo}). This application requires data shuffling, incurring 
@@
 
-We used the \href{https://github.com/BIDS-Apps/example}{BIDS app example} application
-available on GitHub. This application can also be broken down into a map and
-reduce phase. The map phase,
-also called participant analysis, extracts the brain tissues of a subject's 3D MRI
-using the popular FMRIB Software Library
-(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the resulting image
-(\SI{2.47}{\mega\byte} on average) to the shared file system. The reduce phase, also
-called group analysis, computes the volume of each brain and returns the average
+For this experiment, we used the
+\href{https://github.com/BIDS-Apps/example}{BIDS app example} map-reduce style 
+application. The map phase, also called participant analysis, extracts the brain
+tissues of a subject's 3D MRI using the FMRIB Software Library
+(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the resulting
+image (\SI{2.47}{\mega\byte} on average) to the NFS. The reduce phase, or group
+analysis, computes the volume of each brain and returns the average
 volume, shuffling a total of \SI{8.6}{\giga\byte} image data.
 
-While Incrementation and Histogram were implemented directly in Python, BIDS App
-example requires an external library and is distributed as a Docker container image
+While Incrementation and Histogram were implemented directly in Python, the 
+BIDS App example requires an external library and is distributed as a Docker container image
 (bids/base\_fsl on DockerHub). We converted the Docker image to a Singularity image,
@@
 \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}.
-The image was preloaded on the nfs and singularity version 3.2.1\-1.el7 was installed
+The image was preloaded on the NFS and singularity version 3.2.1\-1.el7 was installed
 on all instances. Due to a bug in the conversion from Docker to Singularity, we had to
-export the path for fsl.
+export the path for FSL.
 
@@
 
-We varied four parameters in our experiments, as shown in
+The four parameters were varied in our experiments, as shown in
 Table~\ref{tab:param}. We varied (1) the number of workers to assess the
@@
 and parallelization degrees, and (3) the number of iterations and sleep
-delay in Incrementation to evaluate the effect of job length.
+delay in Incrementation to evaluate the effect of job length and number of
+tasks.
 It should be noted that increasing the number of blocks or iterations also
@@
 For each run, we measured the application makespan as well as the cumulative 
-data read, compute, data write, and engine overhead across all application tasks. 
+data read, compute, data write, and engine overhead across all application
+tasks. 
 The overhead calculation for each CPU is the end time of its last processed task
-minus the start time of the application; the summation of those results is the total
-overhead.
+minus the start time of the application. The summation of those results is the
+total overhead.
 
@@
 For all engines, the makespan is far from decreasing linearly with the
-number of workers. Note that each worker has 8 CPUs. The makespan
+number of workers. The makespan
 even increases between 4 and 8 workers. This is due to the high impact of
@@
 On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
-slightly lower than Dask's in particular as the number of workers increase.
+slightly lower than Dask's, particularly when the number of workers increase.
 Within Dask, Delayed and Futures have a higher overhead than Bag. However,
@@
 In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
-For 30 blocks, Delayed and Futres have much higher overhead. This is because they are
+For 30 blocks, Delayed and Futures have much higher overhead. This is because they are
 thread-based and use multiple threads to compute a block which affect the calculation
@@
 well with task duration. Spark is initially slower than the Dask APIs, however, it is
-faster when increasing the sleep delay. Also, within the Dask APIs, Dask Bag is
-slower than the other two but it is not considerable.
+faster with an increased sleep delay. Also, within the Dask APIs, Dask Bag is
+slower than the other two, but it is not considerable.
 
@@
 Figure~\ref{fig:histo_ms_worker} shows the makespan of the application for various
-amounts of workers. Spark is tremendously faster than Dask APIs. The difference
+amounts of workers. Spark is significantly faster than Dask APIs. The difference
 narrows as the number of workers increases but stays significant. Between Dask APIs
@@
 From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is
-shown. The computation time is significantly larger for Dask than for Spark. This is
-potentially due to Python's GIL preventing Dask to parallelize the computation on
+shown. The compute time is significantly larger for Dask than for Spark. This is
+potentially due to Python's GIL preventing Dask to parallelize the computation
+over
 multiple threads. Overall, the IO and overhead are comparable for all engines.
@@
 On the other hand, Spark total execution time keeps increasing; especially at 8
-workers. This is because Dask engines benefit more from additional workers.
+workers. This is because Dask engines benefit more from additional workers, in
+this GIL-bounded scenario.
 
@@
 \subsection{Histogram: Number of blocks}
-Figure~\ref{fig:histo_ms_block} shows the makespan for different block size. Again,
-Spark is not compared at 30 blocks due to its \SI{2}{\giga\byte} task limitation.
-Once more, Spark is significantly faster than Dask engines and Bag is slightly faster
-than Delayed. Overall, the engines do not react to change in block size.
+Figure~\ref{fig:histo_ms_block} shows the makespan for different block sizes. 
+Once more, Spark is significantly faster than Dask engines and Bag is slightly
+faster than Delayed. Overall, the engines do not react to changes in block size.
 
 In figure~\ref{fig:histo_tt_block} the total time for each function is shown.
-Note that at 30 blocks the worker's resources are not fully used since the workers
-can process up to 64 tasks in parallel. Moreover, the overhead for 30 blocks is
-erroneous due to a lower amount of block than available threads. This leads to thread
-based engines to have higher overhead. Given that,
-
-Spark has a much lower compute time which makes it significantly faster than Dask
-engines. Overall, for all engines, lower block size results in lower IO time but
-increases the overhead time. Again, this is due to large overhead desynchronizing task
-thus reducing the NFS bottleneck.
+Note that at 30 blocks the workers' resources are not fully used since the
+workers can process up to 64 tasks in parallel. Moreover, the overhead for
+30 blocks is erroneous due to a lower amount of block than available threads.
+This leads to thread based engines to have higher overhead. Given that, 
+Spark has a much lower compute time which makes it significantly faster
+than Dask engines. Overall, for all engines, lower block size results in lower
+IO time but increases the overhead time. Again, this is due to large overhead
+desynchronizing task thus reducing the NFS bottleneck.
 
@@
 experiment. Spark overhead is distributed through all the execution while Bag
-overhead happens mostly when a thread read subsequent blocks and Delayed overhead is
-concentrated after reading all the blocks.
+overhead happens mostly when a thread reads subsequent blocks and
+Delayed overhead is concentrated after reading all the blocks.
 
-The reduce is done differently for the engines. Spark and Delayed compute all the
-reduce tasks on one process (thread), however, Spark only computes it on the driver
-process. On the other hand, Bag computes the reduce tasks on different workers and
-sends the information to the driver process afterward.
+% remove? - vhs
+The reduce is done differently for the engines. Spark and Delayed compute
+all the reduce tasks on one process (thread), however, Spark only computes it
+on the driver process. On the other hand, Bag computes the reduce tasks on 
+different workers and sends the information to the driver process afterward.
 
@@
 Overall, the makespan for all engines does not change considerably with an increase
-in the number of workers. This is because the engines do not scale well. There is no
+in the number of workers. This is likely a result of using a non-distributed
+shared file system. There is no
 substantial difference between all engines, although, Delayed is slightly faster.
@@
 From Figure~\ref{fig:histo_np_tt_worker}, the total execution time breakdown is
-shown. The overhead and read time increase proportionally to the number of workers
-thus increasing the total time. This is because (1) most workers read at the same
-time due to a low computation-IO ratio and (2) there is more intra-worker and
-inter-worker communication.
+shown. The overhead and read time increase proportionally to the number of
+workers thus increasing the total time. This is because (1) most workers read
+at the same time due to a low computation-IO ratio and (2) there is more
+intra-worker and inter-worker communication.
 
@@
 Figure~\ref{fig:histo_np_ms_block} shows the makespan of the application for
-different block size. Again, Spark is not compared at 30 blocks due to its
-\SI{2}{\giga\byte} task limitation. Overall, there is no substantial difference for
+different block sizes. Overall, there is no substantial difference for
 all engines and block sizes.
 
-In figure~\ref{fig:histo_np_tt_block}, shows the total time spent in each function.
+Figure~\ref{fig:histo_np_tt_block} shows the total time spent in each function.
 Once more, since Delayed is thread-based its overhead at 30 blocks is seems much
-higher that it really is. Considering the overhead error and that at 30 blocks the
-workers only use half their core, the total time is substantially the same for all
-block size. This is due to lower block sizes having a lower IO time, however, it
-balances out by the resulting increase in overhead.
+higher that it really is. Considering the overhead error and that at 30 blocks
+the workers only use half their core, the total time is substantially the same
+for all block size. This is due to lower block sizes having a lower IO time,
+however, it balances out by the resulting increase in overhead.
 
@@
 Histogram (NumPy) experiment. The overhead for Spark and Bag is mostly located
-between a read and compute task or compute and read task. On the other hand, Delayed
-overhead is dispersed between all types of tasks. This causes Delayed to overlap more
-compute and read tasks simultaneously among the different workers.
+between read and compute tasks. On the other hand, Delayed
+overhead is dispersed between all types of tasks. This causes Delayed to
+overlap more compute and read tasks simultaneously amongst different workers.
 
-In figure~\ref{fig:histo_np_gantt}, read tasks are initially similar for all engines.
-However, when computation starts the subsequent read tasks are much shorter for
-Delayed. This is due the desynchronization of the task reducing the IO bottleneck.
+In figure~\ref{fig:histo_np_gantt}, read tasks are initially similar for all
+engines. However, when computation starts the subsequent read tasks are much
+shorter for Delayed. This is due the desynchronization of the task reducing the
+IO bottleneck.
 
@@
 \subsection{BidsApp example: Number of workers}
-Figure~\ref{fig:bids_ms_worker} shows the makespan of the application when varying
-the amount of workers. Overall, there is no substantial difference between the
-engines. The makespan scales well with the increase in workers, however, the scaling
-reduces considerably when reaching 8 workers. This is due to an increase in overhead
-and of the IO requirement increasing with the number of workers --- more tasks are
-read concurrently and the application becomes more data-intensive.
+Figure~\ref{fig:bids_ms_worker} shows the makespan of the application when
+varying the amount of workers. Overall, there is no substantial difference
+between the engines. The makespan scales well with the increase in workers,
+however, the scaling reduces considerably when reaching 8 workers. This is due
+to an increase in overhead and an IO bottleneck caused parallel accesses to 
+a non-distributed file system.
 
@@
 Fig.~\ref{fig:inc_gantt},\ref{fig:histo_gantt},\ref{fig:histo_np_gantt},\ref{fig:bids_gantt}
-shows. We could use a parallel shared file system, like Lustre \MD{cite lustre if not
-done already}, to reduce the impact of operations with the disk, however, we would
-quickly be limited by the network's bandwidth.
+shows. A parallel file system, such as Lustre~\cite{lustre}, would likely improve
+IO times of parallel applications simultaneously writing to shared storage.
 

### 2019-07-29 latex/paper.tex

@@
 
-\TG{Talk about engine overhead: (1) where is it coming from? (2) 
-explain the impact on data transfers with a graph.}
+Previous work showed that Dask has a the largest start-up overhead~\cite{Mehta:17}
+when processing few subjects. Our Gantt charts show that Spark has a higher start-up
+overhead. We think this is due to the larger amount of subjects we process.
 
-\TG{In the dicussion, get back to it and mention that (1) we don't
-always observe that, (2) for data-intensive applications, overheads don't really matter
-as reducing them increases data trasnfers.}
+The same study showed that Dask has a lower end-to-end runtime than Spark in a
+neuroimaging analysis application. In our results, there is no significant difference
+between the engines. We think this is because the applications we chose are highly
+data-intensive while the one used in their study is compute intensive.
+
+% \TG{Talk about engine overhead: (1) where is it coming from? (2) 
+% explain the impact on data transfers with a graph.}
+
+% \TG{In the dicussion, get back to it and mention that (1) we don't
+% always observe that, (2) for data-intensive applications, overheads don't really matter
+% as reducing them increases data trasnfers.}
 

### 2019-07-29 latex/paper.tex
Worked on conclusion

@@
 low compute time while BIDS App example and Histogram for Dask have significantly
-longer compute time. When adding additional workers, we observe that data-intensive
-applications scale poorly while compute-intensive applications scale almost linearly
-This is due to the limitation from the NFS, as mentioned above.
+longer compute time. When adding additional workers, we observe that the highly
+data-intensive applications scale poorly while compute-intensive applications scale
+almost linearly This is due to the limitation from the NFS, as mentioned above.
 
@@
 
+\section{Conclusion}
+We presented a comparison of two Big Data engines: Spark and Dask. We studied the
+engines on three data-intensive neuroimaging applications. Our results show no
+substantial difference between the engines. Moreover, the majority of the execution
+time was spent on data transfer and overhead. We also observed a possible correlation
+between IO and overhead time: an higher overhead is compensated by a lower IO time and
+a low overhead is compensated by an high IO time. This would be due to concurrent IO
+tasks saturating the bandwidth. We suggest that current schedulers policies would benefit
+from knowing the task types, either IO or compute, it is processing concurrently.
+
 % \TG{Talk about engine overhead: (1) where is it coming from? (2) 

### 2019-07-29 latex/paper.tex

@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_inc_gantt.png}
+        \href{https://zenodo.org/record/3354434/files/spark-inc-baseline.html?download=1}{
+            \includegraphics[clip,width=\columnwidth,
+            height=0.15\textheight]{images/spark_inc_gantt.png}}
         \caption{Spark}\label{fig:inc_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_inc_gantt.png}%
+        \href{https://zenodo.org/record/3354434/files/dask-bag-inc-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/bag_inc_gantt.png}}
         \caption{Dask Bag}\label{fig:inc_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_inc_gantt.png}%
+        \href{https://zenodo.org/record/3354434/files/dask-delayed-inc-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/delayed_inc_gantt.png}}
         \caption{Dask Delayed}\label{fig:inc_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/futures_inc_gantt.png}%
+        \href{https://zenodo.org/record/3354434/files/dask-futures-inc-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/futures_inc_gantt.png}}
         \caption{Dask Futures}\label{fig:inc_dask_futures_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_histo_gantt.png}
+        \href{https://zenodo.org/record/3354434/files/spark-histo-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/spark_histo_gantt.png}}
         \caption{Spark}\label{fig:histo_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_histo_gantt.png}%
+        \href{https://zenodo.org/record/3354434/files/dask-bag-histo-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/bag_histo_gantt.png}}
         \caption{Dask Bag}\label{fig:histo_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_histo_gantt.png}%
+        \href{https://zenodo.org/record/3354434/files/dask-delayed-histo-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/delayed_histo_gantt.png}}
         \caption{Dask Delayed}\label{fig:histo_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_histo_np_gantt.png}
+        \href{https://zenodo.org/record/3354434/files/spark-histo_np-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/spark_histo_np_gantt.png}}
         \caption{Spark}\label{fig:histo_np_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_histo_np_gantt.png}%
+        \href{https://zenodo.org/record/3354434/files/dask-bag-histo_np-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/bag_histo_np_gantt.png}}
         \caption{Dask Bag}\label{fig:histo_np_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_histo_np_gantt.png}%
+        \href{https://zenodo.org/record/3354434/files/dask-delayed-histo_np-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/delayed_histo_np_gantt.png}}
         \caption{Dask Delayed}\label{fig:histo_np_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_bids_gantt.png}
+        \href{https://zenodo.org/record/3354434/files/spark-bids-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/spark_bids_gantt.png}}
         \caption{Spark}\label{fig:bids_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_bids_gantt.png}%
+        \href{https://zenodo.org/record/3354434/files/dask-bag-bids-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/bag_bids_gantt.png}}
         \caption{Dask Bag}\label{fig:bids_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_bids_gantt.png}%
+        \href{https://zenodo.org/record/3354434/files/dask-delayed-bids-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/delayed_bids_gantt.png}}
         \caption{Dask Delayed}\label{fig:bids_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/futures_bids_gantt.png}%
+        \href{https://zenodo.org/record/3354434/files/dask-futures-bids-baseline.html?download=1}{
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/futures_bids_gantt.png}}
         \caption{Dask Futures}\label{fig:bids_dask_futures_gantt}

### 2019-07-29 latex/paper.tex
Updates in abstract

--- 2019-07-29 latex/paper.tex
@@
 
-\title{A performance comparison of Dask and Spark for data-intensive neuroimaging pipelines}
+\title{A performance comparison of Dask and Apache Spark for data-intensive neuroimaging pipelines}
 
@@
 \begin{abstract}
-% Topic
-Recent increases in data volume and improved accessibility have 
-led neuroimaging into the Big Data era. 
-% Problem
-While existing neuroimaging pipeline frameworks typically do not attempt to
-limit data transfers, Big Data frameworks remain seldom used.
-% Relevance
-In this paper, we compare two popular Big Data frameworks with Python APIs,
-Apache Spark and Dask, for their runtime performance in processing neuroimaging
-pipelines.
-% Approach
-To evaluate performance differences we used two synthetic, Incrementation
-(map-only) and Histogram (map-reduce), pipelines and real neuroimaging map-reduce pipeline (BIDS App Example). For both our 
-synthetic pipelines we varied the number of workers and block size using the 
-76~GB BigBrain image. We additionally varied the number of iterations and task
-duration for our Incrementation pipeline. 39~GB of anatomical images from the
-CoRR were used for the BIDS App Example pipeline.
-% Methods (one sentence)
-All experiments were executed on 8 instances of Compute Canada's Arbutus cloud
-with 8~vCPUs and 30 GB of RAM all connected to a 2~TB shared NFS. We evaluated
-PySpark's RDD against Dask's Bag, Delayed and Futures.
-% Key Impact
-Our result show that despite the differences between the two frameworks, both
-frameworks perform the comparably. However, Dask pipelines risk being limited by
-Python's GIL depending on task type and cluster configuration. In all cases, 
-the major limiting factor was data transfer. While either framework is suitable
-for neuroimaging pipelines, more effort needs to be placed in reducing data 
-transfer time.
+
+In the past few years, neuroimaging has entered the Big Data era due to the
+joint increase in image resolution, data sharing, and study sizes. However,
+no particular Big Data framework has emerged in this field, and several
+alternatives remain available. We compare two popular Big Data frameworks
+with Python APIs, Apache Spark and Dask, for their runtime performance in
+processing neuroimaging pipelines. Our evaluation uses two synthetic
+pipelines processing the 76~GB Big Brain image, and a real pipeline
+processing anatomical data from more than 1,000 subjects. We benchmark these pipelines using
+ various combinations of task durations, data sizes, and numbers of workers,
+deployed on a 64-node compute cluster in Compute Canada's Arbutus
+cloud. We evaluate PySpark's RDD API against Dask's Bag, Delayed and Futures.
+Results show that despite slight differences between Spark and Dask, both
+frameworks perform comparably. However, Dask pipelines risk being limited
+by Python's GIL depending on task type and cluster configuration. In all
+cases, the major limiting factor was data transfer. While either framework
+is suitable for neuroimaging pipelines, more effort needs to be placed in
+reducing data transfer time.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Big Data, Dask, Spark, performance, neuroimaging
+Dask, Apache Spark, performance, neuroimaging
 \end{IEEEkeywords}

### 2019-07-29 latex/paper.tex
Reduce wording

--- 2019-07-29 latex/paper.tex
@@
 
-%Similarities
 Spark and Dask both offer in-memory computing, data locality, and lazy evaluation;
@@
 
-% Differences
 First and foremost, Spark is written in Scala while Dask is in Python.
@@
 
-% Related work
 Spark and Dask were both included in the evaluation reported
@@
 Incrementation, Histogram, and Histogram (NumPy). We observe that larger block size
-results in more IO compensated by lower overhead time. On the other hand, smaller
-block size results in higher overhead time compensated by reduced IO time.
+results in more IO compensated by lower overhead time and vice-versa.
 
@@
 
-Varying the block size is equivalent to varying the size of the partitions. We
-observed no difference when varying the task size while using an NFS. However, if
-one uses a parallel shared file system, we suppose it would be preferable to use a
-larger block size as the IO time should be reduced substantially. It is worth noting
-that the Dask developers suggest, in
+Varying the block size or the size of the partitions is equivalent. No difference is
+observed when varying the task size while using an NFS. However, if one uses a
+parallel shared file system, we suppose it would be preferable to use a larger block
+size as the IO time should be reduced substantially. It is worth noting that the Dask
+developers suggest, in
 \href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to

### 2019-07-29 latex/paper.tex
Reduce wording

@@
 
-% remove? - vhs
-The reduce is done differently for the engines. Spark and Delayed compute
-all the reduce tasks on one process (thread), however, Spark only computes it
-on the driver process. On the other hand, Bag computes the reduce tasks on 
-different workers and sends the information to the driver process afterward.
+% % remove? - vhs
+% The reduce is done differently for the engines. Spark and Delayed compute
+% all the reduce tasks on one process (thread), however, Spark only computes it
+% on the driver process. On the other hand, Bag computes the reduce tasks on 
+% different workers and sends the information to the driver process afterward.
 
@@
 observed when varying the task size while using an NFS. However, if one uses a
-parallel shared file system, we suppose it would be preferable to use a larger block
+parallel shared file system, it would be preferable to use a larger block
 size as the IO time should be reduced substantially. It is worth noting that the Dask

### 2019-07-29 latex/paper.tex
Minor edits in introduction

--- 2019-07-29 latex/paper.tex
@@
 Concordia University, Montr\'eal, Qu\'ebec, Canada 
+\vspace*{0.8cm}
 }
@@
 have brought neuroimaging to the Big Data era ~\cite{ALFAROALMAGRO:18,
-UKBioBank:18}. Existing neuroimaging workflow engines, such as
+van2014human}. Existing neuroimaging workflow engines, such as
 Nipype~\cite{Nipype:11}, are well suited for processing the standard
@@
 
-Spark and Dask both offer in-memory computing, data locality, and lazy evaluation;
+%Similarities
+Spark and Dask both offer in-memory computing, data locality, and lazy evaluation,
 which is common for Big Data frameworks. Both their schedulers operate dynamically,
@@
 
+% Differences
 First and foremost, Spark is written in Scala while Dask is in Python.
@@
 Java, R and Python APIs allow to easily parallelize analyses that combine
-these languages, with reduced performance loss.
+these languages, with reduced performance loss. Our study focuses on
+performance, although we recognize that other factors will also impact the
+choice of a framework in practice.
 
+% Related work
 Spark and Dask were both included in the evaluation reported
-in~\cite{Mehta:17} that used a neuroimaging and an astronomy application
-processing approximately \SI{100}{\giga\byte} of data each. In this work, Dask was
+in~\cite{Mehta:17}, where a neuroimaging application
+processed approximately \SI{100}{\giga\byte} of data. In this work, Dask was
 reported to have a slight performance advantage over Spark. Overall, Dask's
@@
 time than Spark. % (Fig. 12-c)
-The analysis remains at a quite high level
+The analysis remained at a quite high level
 though, leaving most of the observed performance difference unexplained. In
 comparison, our study will provide a detailed analysis of performance
-differences and similarities between Spark and Dask.
+differences and similarities between Spark and Dask for neuroimaging data processing.
 
-The next section details the design of our experiments. We consider two
+The next section details the design of our benchmarking experiments. We consider two
 data-intensive neuroimaging applications: high-resolution imaging,
@@
 representative of the ones used in today's data-intensive neuroimaging
-studies, using the main Dask and Spark APIs. The following sections present
+studies, using the main Dask and Spark APIs. The other sections present
 our results, discussion and conclusions.
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/spark-inc-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-            height=0.15\textheight]{images/spark_inc_gantt.png}}
+        height=0.15\textheight]{images/spark_inc_gantt.png}
         \caption{Spark}\label{fig:inc_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-bag-inc-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_inc_gantt.png}}
+        height=0.15\textheight]{images/bag_inc_gantt.png}%
         \caption{Dask Bag}\label{fig:inc_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-inc-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_inc_gantt.png}}
+        height=0.15\textheight]{images/delayed_inc_gantt.png}%
         \caption{Dask Delayed}\label{fig:inc_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-futures-inc-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/futures_inc_gantt.png}}
+        height=0.15\textheight]{images/futures_inc_gantt.png}%
         \caption{Dask Futures}\label{fig:inc_dask_futures_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/spark-histo-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_histo_gantt.png}}
+        height=0.15\textheight]{images/spark_histo_gantt.png}
         \caption{Spark}\label{fig:histo_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-bag-histo-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_histo_gantt.png}}
+        height=0.15\textheight]{images/bag_histo_gantt.png}%
         \caption{Dask Bag}\label{fig:histo_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-histo-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_histo_gantt.png}}
+        height=0.15\textheight]{images/delayed_histo_gantt.png}%
         \caption{Dask Delayed}\label{fig:histo_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/spark-histo_np-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_histo_np_gantt.png}}
+        height=0.15\textheight]{images/spark_histo_np_gantt.png}
         \caption{Spark}\label{fig:histo_np_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-bag-histo_np-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_histo_np_gantt.png}}
+        height=0.15\textheight]{images/bag_histo_np_gantt.png}%
         \caption{Dask Bag}\label{fig:histo_np_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-histo_np-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_histo_np_gantt.png}}
+        height=0.15\textheight]{images/delayed_histo_np_gantt.png}%
         \caption{Dask Delayed}\label{fig:histo_np_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/spark-bids-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_bids_gantt.png}}
+        height=0.15\textheight]{images/spark_bids_gantt.png}
         \caption{Spark}\label{fig:bids_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-bag-bids-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_bids_gantt.png}}
+        height=0.15\textheight]{images/bag_bids_gantt.png}%
         \caption{Dask Bag}\label{fig:bids_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-bids-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_bids_gantt.png}}
+        height=0.15\textheight]{images/delayed_bids_gantt.png}%
         \caption{Dask Delayed}\label{fig:bids_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-futures-bids-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/futures_bids_gantt.png}}
+        height=0.15\textheight]{images/futures_bids_gantt.png}%
         \caption{Dask Futures}\label{fig:bids_dask_futures_gantt}
@@
 low compute time while BIDS App example and Histogram for Dask have significantly
-longer compute time. When adding additional workers, we observe that the highly
-data-intensive applications scale poorly while compute-intensive applications scale
-almost linearly This is due to the limitation from the NFS, as mentioned above.
+longer compute time. When adding additional workers, we observe that data-intensive
+applications scale poorly while compute-intensive applications scale almost linearly
+This is due to the limitation from the NFS, as mentioned above.
 
@@
 Incrementation, Histogram, and Histogram (NumPy). We observe that larger block size
-results in more IO compensated by lower overhead time and vice-versa.
+results in more IO compensated by lower overhead time. On the other hand, smaller
+block size results in higher overhead time compensated by reduced IO time.
 
@@
 
-Varying the block size or the size of the partitions is equivalent. No difference is
-observed when varying the task size while using an NFS. However, if one uses a
-parallel shared file system, we suppose it would be preferable to use a larger block
-size as the IO time should be reduced substantially. It is worth noting that the Dask
-developers suggest, in
+Varying the block size is equivalent to varying the size of the partitions. We
+observed no difference when varying the task size while using an NFS. However, if
+one uses a parallel shared file system, we suppose it would be preferable to use a
+larger block size as the IO time should be reduced substantially. It is worth noting
+that the Dask developers suggest, in
 \href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to
@@
 data-intensive while the one used in their study is compute intensive.
-
-\section{Conclusion}
-We presented a comparison of two Big Data engines: Spark and Dask. We studied the
-engines on three data-intensive neuroimaging applications. Our results show no
-substantial difference between the engines. Moreover, the majority of the execution
-time was spent on data transfer and overhead. We also observed a possible correlation
-between IO and overhead time: an higher overhead is compensated by a lower IO time and
-a low overhead is compensated by an high IO time. This would be due to concurrent IO
-tasks saturating the bandwidth. We suggest that current schedulers policies would benefit
-from knowing the task types, either IO or compute, it is processing concurrently.
 

### 2019-07-29 latex/paper.tex
Modified regression slopes wording

--- 2019-07-29 latex/paper.tex
@@
 transfer time and overhead increase proportionally to the number of
-workers with regression lines:
-
-\hspace{1cm}\(\hat{y}=2336.9499\cdot{x} + 325.00574\) (Spark)
-
-\hspace{1cm}\(\hat{y}=2650.13505\cdot{x} + 1585.37218\) (Bag)
-
-\hspace{1cm}\(\hat{y}=3250.53406\cdot{x} + 1769.50351\) (Delayed)
-
-\hspace{1cm}\(\hat{y}=3570.39758\cdot{x} + 1062.23031\) (Futures)
+workers with regression slopes: Spark (\SI{2337}{\second/task}), Bag
+(\SI{2650}{\second/task}), Delayed (\SI{3251}{\second/task}), Futures
+(\SI{3570}{\second/task}).
 
@@
 This is due to the, almost, linear scaling of the engines. The overhead increases
-proportionally with the number of workers (see regression lines below). This is due
-to more communication between the scheduler and workers as well as intra-worker and
+proportionally to the number of workers with regression slopes: Spark
+(\SI{224}{\second/task}), Bag (\SI{183}{\second/task}), Delayed
+(\SI{173}{\second/task}), Futures (\SI{132}{\second/task}). This is due to more
+communication between the scheduler and workers as well as intra-worker and
 inter-worker communication.
-
-\hspace{1cm}\(\hat{y}=223.53586\cdot{x} + 91.616964\) (Spark)
-
-\hspace{1cm}\(\hat{y}=182.82659\cdot{x} - 11.43834\) (Bag)
-
-\hspace{1cm}\(\hat{y}=172.54183\cdot + 400.69433\) (Delayed)
-
-\hspace{1cm}\(\hat{y}=132.41481\cdot{x} + 90.32813\) (Futures)
 

### 2019-07-29 latex/paper.tex

--- 2019-07-29 latex/paper.tex
@@
 Java, R and Python APIs allow to easily parallelize analyses that combine
-these languages, with reduced performance loss.
-
+these languages, with reduced performance loss. Our study focuses on
+performance, although we recognize that other factors will also impact the
+choice of a framework in practice.
+
+% Related work
 Spark and Dask were both included in the evaluation reported
-in~\cite{Mehta:17} that used a neuroimaging and an astronomy application
-processing approximately \SI{100}{\giga\byte} of data each. In this work, Dask was
+in~\cite{Mehta:17}, where a neuroimaging application
+processed approximately \SI{100}{\giga\byte} of data. In this work, Dask was
 reported to have a slight performance advantage over Spark. Overall, Dask's
@@
 time than Spark. % (Fig. 12-c)
-The analysis remains at a quite high level
+The analysis remained at a quite high level
 though, leaving most of the observed performance difference unexplained. In
 comparison, our study will provide a detailed analysis of performance
-differences and similarities between Spark and Dask.
-
-The next section details the design of our experiments. We consider two
+differences and similarities between Spark and Dask for neuroimaging data processing.
+
+The next section details the design of our benchmarking experiments. We consider two
 data-intensive neuroimaging applications: high-resolution imaging,
@@
 
-% % remove? - vhs
-% The reduce is done differently for the engines. Spark and Delayed compute
-% all the reduce tasks on one process (thread), however, Spark only computes it
-% on the driver process. On the other hand, Bag computes the reduce tasks on 
-% different workers and sends the information to the driver process afterward.
+% remove? - vhs
+The reduce is done differently for the engines. Spark and Delayed compute
+all the reduce tasks on one process (thread), however, Spark only computes it
+on the driver process. On the other hand, Bag computes the reduce tasks on 
+different workers and sends the information to the driver process afterward.
 
@@
 
-Varying the block size or the size of the partitions is equivalent. No difference is
-observed when varying the task size while using an NFS. However, if one uses a
-parallel shared file system, it would be preferable to use a larger block
-size as the IO time should be reduced substantially. It is worth noting that the Dask
-developers suggest, in
+Varying the block size is equivalent to varying the size of the partitions. We
+observed no difference when varying the task size while using an NFS. However, if
+one uses a parallel shared file system, we suppose it would be preferable to use a
+larger block size as the IO time should be reduced substantially. It is worth noting
+that the Dask developers suggest, in
 \href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to

### 2019-07-29 latex/paper.tex

--- 2019-07-29 latex/paper.tex
@@
 Java, R and Python APIs allow to easily parallelize analyses that combine
-these languages, with reduced performance loss.
+these languages, with reduced performance loss. Our study focuses on
+performance, although we recognize that other factors will also impact the
+choice of a framework in practice.
 
 Spark and Dask were both included in the evaluation reported
-in~\cite{Mehta:17} that used a neuroimaging and an astronomy application
-processing approximately \SI{100}{\giga\byte} of data each. In this work, Dask was
+in~\cite{Mehta:17}, where a neuroimaging application
+processed approximately \SI{100}{\giga\byte} of data. In this work, Dask was
 reported to have a slight performance advantage over Spark. Overall, Dask's

### 2019-07-29 latex/paper.tex
Merge concflict+ Slight changes in abstract

@@
 
-In the past few years, neuroimaging has entered the Big Data era due to the
-joint increase in image resolution, data sharing, and study sizes. However,
-no particular Big Data framework has emerged in this field, and several
-alternatives remain available. We compare two popular Big Data frameworks
-with Python APIs, Apache Spark and Dask, for their runtime performance in
-processing neuroimaging pipelines. Our evaluation uses two synthetic
-pipelines processing the 76~GB Big Brain image, and a real pipeline
-processing anatomical data from more than 1,000 subjects. We benchmark these pipelines using
- various combinations of task durations, data sizes, and numbers of workers,
-deployed on a 64-node compute cluster in Compute Canada's Arbutus
-cloud. We evaluate PySpark's RDD API against Dask's Bag, Delayed and Futures.
-Results show that despite slight differences between Spark and Dask, both
-frameworks perform comparably. However, Dask pipelines risk being limited
-by Python's GIL depending on task type and cluster configuration. In all
-cases, the major limiting factor was data transfer. While either framework
-is suitable for neuroimaging pipelines, more effort needs to be placed in
-reducing data transfer time.
+In the past few years, neuroimaging has entered the Big Data era due to the joint
+increase in image resolution, data sharing, and study sizes. However, no particular
+Big Data engines has emerged in this field, and several alternatives remain
+available. We compare two popular Big Data engines with Python APIs, Apache Spark and
+Dask, for their runtime performance in processing neuroimaging pipelines. Our
+evaluation uses two synthetic pipelines processing the \SI{81}{\giga\byte} Big Brain
+image, and a real pipeline processing anatomical data from more than 1,000 subjects.
+We benchmark these pipelines using various combinations of task durations, data
+sizes, and numbers of workers, deployed on a 8-node (8 cores ea.) compute cluster in
+Compute Canada's Arbutus cloud. We evaluate PySpark's RDD API against Dask's Bag,
+Delayed and Futures. Results show that despite slight differences between Spark and
+Dask, both frameworks perform comparably. However, Dask pipelines risk being limited
+by Python's GIL depending on task type and cluster configuration. In all cases, the
+major limiting factor was data transfer. While either framework is suitable for
+neuroimaging pipelines, more effort needs to be placed in reducing data transfer time.
 \end{abstract}

### 2019-07-29 latex/paper.tex
Fixes and comment in Methods

--- 2019-07-29 latex/paper.tex
@@
 Concordia University, Montr\'eal, Qu\'ebec, Canada 
-\vspace*{0.8cm}
+\vspace*{0.8cm} % to avoid weird spacing of 1st page by Latex.
 }
@@
 processing neuroimaging pipelines. Our evaluation uses two synthetic
-pipelines processing the 76~GB Big Brain image, and a real pipeline
+pipelines processing the 81~GB Big Brain image, and a real pipeline
 processing anatomical data from more than 1,000 subjects. We benchmark these pipelines using
@@
 bottlenecks, Spark incorporates the concepts of data locality and in-memory 
-computing. Data locality, introduced by MapReduce~\cite{dean2008mapreduce},
+computing. Data locality, popularized by MapReduce~\cite{dean2008mapreduce},
 schedules
-tasks as close as possible to where the data is storage. In-memory computing
+tasks as close as possible to where the data is stored. In-memory computing
 ensures data is maintained in memory whenever possible, as writing large 
@@
 workflows, whereas Mesos can be used to schedule
-a variety of different workflows. As scientific researchers would likely be
+a variety of different workflows. As researchers would likely be
 executing their workflows in HPC environments with neither YARN or Mesos
@@
 computation graph. The Spark standalone scheduler uses a FIFO 
-(First-In-First-Out) job scheduling policy. The Spark standalone scheduler has
+(First-In-First-Out) job scheduling policy, and it has
 two execution modes: client mode, where the driver runs in a dedicated process
@@
 increasingly popular in the scientific Python ecosystem. Like Spark, Dask avoids data transfers and needless computation and communication through
-in-memory computing, data locality, lazy evaluation. Dask worklflows can 
+in-memory computing, data locality, and lazy evaluation. Dask worklflows can 
 further reduce data transfer costs by leveraging multithreading whenever 
@@
 
-Dask has five main data structure APIs:
-Array, Bag, DataFrame, Delayed, and Futures. A Dask
-\href{https://docs.dask.org/en/latest/array.html}{Array} is used for the
+Dask has five main data structures:
+\href{https://docs.dask.org/en/latest/array.html}{Array}, \href{https://docs.dask.org/en/latest/bag.html}{Bag},
+\href{https://docs.dask.org/en/latest/dataframe.html}{DataFrame},
+ \href{https://docs.dask.org/en/latest/delayed.html}{Delayed}, 
+ and \href{https://docs.dask.org/en/latest/futures.html}{Futures}. A Dask
+Array is used for the
 processing of large arrays. It provides a distributed clone of the popular NumPy
 library. Similar to Spark's RDD, a Dask
-\href{https://docs.dask.org/en/latest/bag.html}{Bag} is a parallel
+Bag is a parallel
 collection of Python objects. It offers a programming abstraction similar to the
-\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}. A
-\href{https://docs.dask.org/en/latest/dataframe.html}{Dataframe} is a parallel
+\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}. A Dask
+Dataframe is a parallel
 composition of
-\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas Dataframes} used to process a large amount of tabular data.
-\href{https://docs.dask.org/en/latest/delayed.html}{Delayed} supports
+\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas Dataframes},
+ used to process a large amount of tabular data.
+Dask Delayed supports
 arbitrary tasks
 that do not fit in the Array, DataFrame or Bag APIs. Finally,
-\href{https://docs.dask.org/en/latest/futures.html}{Futures} are similar to
+Dask Futures are similar to
 Delayed in that they support arbitrary tasks, but they operate in real-time
@@
 the local multithreaded scheduler. Dask Bag, instead, relies on the local
-multiprocessing scheduler.  All Dask data structure APIs, with the exception of
+multiprocessing scheduler.  All Dask data structures, with the exception of
 Dask Array and Dask DataFrame, were used in our experiments.
 
-The Dask graph is the internal representation of Dask application to be
-executed  by the scheduler. API Operations generate multiple small tasks in
-the computation graph allowing an easier representation of complex
+The Dask graph is the internal representation of a Dask application to be
+executed  by the scheduler. API operations generate multiple small tasks in
+the computation graph, allowing an easier representation of complex
 algorithms.
@@
  image running CentOS 7.5.1804 with Linux kernel version
- 3.10.0\-862.11.6.el7.x86\_64. Instances are connected by a
+ 3.10.0\-862.11.6.el7.x86\_64. Instances were connected by a
  \SI{10}{\giga\bit/\second} Ethernet network.
@@
  Dask scheduler, and for job monitoring with the Spark and Dask user
- interfaces. For both Spark and Dask, spilled data is evicted to the NFS.
+ interfaces. For both Spark and Dask, spilled data was evicted to the NFS.
 
@@
 \SI{2.7}{\giga\byte}, \SI{0.648}{\giga\byte}, and
-\SI{0.108}{\giga\byte}, using the sam~\cite{sam} library.
+\SI{0.108}{\giga\byte}, using the \href{https://github.com/big-data-lab-team/sam}{sam} library.
 
 We also used the dataset provided by the Consortium for Reliability and
-Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) as
+Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}~\cite{zuo2014open}) as
 available on \href{http://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
 entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and
-functional images of 1,397 subjects acquired in 29 sites. We used all 3491 anatomical
+functional images of 1,397 subjects acquired in 29 sites. We used all 3,491 anatomical
 images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} per image on
@@
     \Require{\(x\), a sleep delay in float}
-    \Require{\(file\), a file containing a block}
-    \Require{\(fs\), NFS to write image to.}
+    \Require{\(file\), a file containing a BigBrain block}
+    \Require{\(fs\), NFS mount to write image to}
     \State{Read \(block\) from \(file\)}
@@
     \centering
-    \includegraphics[width=2cm, height=\columnwidth,
+    \includegraphics[height=\columnwidth,
     angle=-90]{images/incrementation-task-graph.png}
@@
     Circles represent the incrementation and sleep function while rectangles
-    represent stages of the blocks.}\label{fig:tg-inc}
+    represent data elements.}\label{fig:tg-inc}
 \end{figure}
@@
  Our second application calculates the histogram of the BigBrain image (see
- Algorithm~\ref{alg:histogram}). It reads the image blocks from the shared file
- system, calculates intensity frequencies, aggregates the frequencies across the
- blocks, and finally writes the resulting histogram to the shared file system as a
- single \SI{766}{\kilo\byte} file. It follows the map-reduce
- paradigm, in which the final result is obtained from all the input blocks (see
- Fig.~\ref{fig:tg-histo}). This application requires data shuffling, incurring 
- additional performance costs compared to Algorithm~\ref{alg:incrementation}.
- The total amount of shuffled data is, however, limited to
- \SI{2.62}{\mega\byte} per blocks as it only consists of image histograms.
-Two implementation are studied: pure Python and NumPy.
+ Algorithm~\ref{alg:histogram}). It reads the image blocks from the shared
+ file system, calculates intensity frequencies, aggregates the frequencies
+ across the blocks, and finally writes the resulting histogram to the
+ shared file system as a single \SI{766}{\kilo\byte} file. It follows the
+ map-reduce paradigm, in which the final result is obtained from all the
+ input blocks (see Fig.~\ref{fig:tg-histo}). This application requires data
+ shuffling, incurring additional performance costs compared to
+ Algorithm~\ref{alg:incrementation}. The total amount of shuffled data is,
+ however, limited to \SI{2.62}{\mega\byte} per blocks as it only consists
+ of image histograms. Two implementation are studied: a pure Python one,
+ where frequencies are computed through Python dictionary manipulations, and one based
+ on the \href{https://github.com/numpy}{NumPy} library, that implements array computations in C.
 
@@
     \begin{algorithmic}
-    \Require{\(files\), files containing BigBrain}
-    \Require{\(fs\), NFS to save image to.}
+    \Require{\(files\), files containing BigBrain blocks}
+    \Require{\(fs\), NFS mount to save image to}
     \ForEach{\(file \in files\)}
         \State{Read \(block\) from \(file\)}
-        \State{{Calculate \(frequency\) of \(blocks\)}}
+        \State{{Calculate \(frequencies\) in \(block\)}}
     \EndFor
 
-    \State{\(histogram\gets\)Aggregate \(frequency\) of each \(file\)}
+    \State{\(histogram\gets\)Aggregate \(frequencies\) of each \(file\)}
 
@@
     \centering
-    \includegraphics[width=2cm, height=\columnwidth,
+    \includegraphics[height=\columnwidth,
     angle=-90]{images/histogram-task-graph.png}
     \caption{Task graph for Histogram with 3 BigBrain blocks. Circles represent the
-    functions while rectangles represent stages of the blocks.
+    functions while rectangles represent data elements.
     }\label{fig:tg-histo}
@@
 
-For this experiment, we used the
+We used the
 \href{https://github.com/BIDS-Apps/example}{BIDS app example} map-reduce style 
 application. The map phase, also called participant analysis, extracts the brain
-tissues of a subject's 3D MRI using the FMRIB Software Library
+tissues of a subject's 3D MRI from the CoRR dataset using the FMRIB Software Library
 (\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the resulting
@@
 While Incrementation and Histogram were implemented directly in Python, the 
-BIDS App example requires an external library and is distributed as a Docker container image
-(bids/base\_fsl on DockerHub). We converted the Docker image to a Singularity image,
-given its popularity for HPC and reproducibility, using
+BIDS App example requires an external set of command-line tools distributed as a Docker container image
+(bids/base\_fsl on DockerHub). We converted the Docker image to a Singularity image for use in 
+an HPC environment, using
 \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}.
-The image was preloaded on the NFS and singularity version 3.2.1\-1.el7 was installed
-on all instances. Due to a bug in the conversion from Docker to Singularity, we had to
-export the path for FSL.
+The image was preloaded on the NFS and Singularity version 3.2.1\-1.el7 was installed
+on all instances.
 
@@
 
-The four parameters were varied in our experiments, as shown in
-Table~\ref{tab:param}. We varied (1) the number of workers to assess the
+Four parameters were varied in our experiments, as shown in
+Table~\ref{tab:param}. We varied (1) the number of workers, to assess the
 scalability of the engines, (2) the number of BigBrain blocks in
-Incrementation and Histogram to measure the effect of different IO patterns
+Incrementation and Histogram, to measure the effect of different I/O patterns
 and parallelization degrees, and (3) the number of iterations and sleep
-delay in Incrementation to evaluate the effect of job length and number of
+delay in Incrementation, to evaluate the effect of job length and number of
 tasks.
@@
 For each run, we measured the application makespan as well as the cumulative 
-data read, compute, data write, and engine overhead across all application
+data read, compute, data write, and engine overhead times across all application
 tasks. 
@@
 minus the start time of the application. The summation of those results is the
-total overhead.
+total overhead. \TG{This last sentence looks wrong to me.}
 
@@
 for different numbers of workers and engines. The bars show the average
-makespan over 3 repetitions while the error bars are the standard deviation. Overall,
+makespan over 3 repetitions while the error bars are the standard deviations. Overall,
 there is no substantial makespan difference between the engines. Dask seems to have a
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/spark-inc-baseline.html?download=1}{
-            \includegraphics[clip,width=\columnwidth,
-            height=0.15\textheight]{images/spark_inc_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/spark_inc_gantt.png}
         \caption{Spark}\label{fig:inc_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-bag-inc-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_inc_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/bag_inc_gantt.png}%
         \caption{Dask Bag}\label{fig:inc_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-inc-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_inc_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/delayed_inc_gantt.png}%
         \caption{Dask Delayed}\label{fig:inc_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-futures-inc-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/futures_inc_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/futures_inc_gantt.png}%
         \caption{Dask Futures}\label{fig:inc_dask_futures_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/spark-histo-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_histo_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/spark_histo_gantt.png}
         \caption{Spark}\label{fig:histo_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-bag-histo-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_histo_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/bag_histo_gantt.png}%
         \caption{Dask Bag}\label{fig:histo_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-histo-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_histo_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/delayed_histo_gantt.png}%
         \caption{Dask Delayed}\label{fig:histo_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/spark-histo_np-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_histo_np_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/spark_histo_np_gantt.png}
         \caption{Spark}\label{fig:histo_np_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-bag-histo_np-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_histo_np_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/bag_histo_np_gantt.png}%
         \caption{Dask Bag}\label{fig:histo_np_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-histo_np-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_histo_np_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/delayed_histo_np_gantt.png}%
         \caption{Dask Delayed}\label{fig:histo_np_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/spark-bids-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_bids_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/spark_bids_gantt.png}
         \caption{Spark}\label{fig:bids_spark_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-bag-bids-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_bids_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/bag_bids_gantt.png}%
         \caption{Dask Bag}\label{fig:bids_dask_bag_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-bids-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_bids_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/delayed_bids_gantt.png}%
         \caption{Dask Delayed}\label{fig:bids_dask_delayed_gantt}
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-futures-bids-baseline.html?download=1}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/futures_bids_gantt.png}}
+        \includegraphics[clip,width=\columnwidth,
+        height=0.15\textheight]{images/futures_bids_gantt.png}%
         \caption{Dask Futures}\label{fig:bids_dask_futures_gantt}
@@
 
-\section{Conclusion}
-We presented a comparison of two Big Data engines: Spark and Dask. We studied the
-engines on three data-intensive neuroimaging applications. Our results show no
-substantial difference between the engines. Moreover, the majority of the execution
-time was spent on data transfer and overhead. We also observed a possible correlation
-between IO and overhead time: an higher overhead is compensated by a lower IO time and
-a low overhead is compensated by an high IO time. This would be due to concurrent IO
-tasks saturating the bandwidth. We suggest that current schedulers policies would benefit
-from knowing the task types, either IO or compute, it is processing concurrently.
-
 % \TG{Talk about engine overhead: (1) where is it coming from? (2) 

### 2019-07-29 latex/paper.tex

@@
 
-In the past few years, neuroimaging has entered the Big Data era due to the
-joint increase in image resolution, data sharing, and study sizes. However,
-no particular Big Data framework has emerged in this field, and several
-alternatives remain available. We compare two popular Big Data frameworks
-with Python APIs, Apache Spark and Dask, for their runtime performance in
-processing neuroimaging pipelines. Our evaluation uses two synthetic
-pipelines processing the 81~GB Big Brain image, and a real pipeline
-processing anatomical data from more than 1,000 subjects. We benchmark these pipelines using
- various combinations of task durations, data sizes, and numbers of workers,
-deployed on a 64-node compute cluster in Compute Canada's Arbutus
-cloud. We evaluate PySpark's RDD API against Dask's Bag, Delayed and Futures.
-Results show that despite slight differences between Spark and Dask, both
-frameworks perform comparably. However, Dask pipelines risk being limited
-by Python's GIL depending on task type and cluster configuration. In all
-cases, the major limiting factor was data transfer. While either framework
-is suitable for neuroimaging pipelines, more effort needs to be placed in
-reducing data transfer time.
+In the past few years, neuroimaging has entered the Big Data era due to the joint
+increase in image resolution, data sharing, and study sizes. However, no particular
+Big Data engines has emerged in this field, and several alternatives remain
+available. We compare two popular Big Data engines with Python APIs, Apache Spark and
+Dask, for their runtime performance in processing neuroimaging pipelines. Our
+evaluation uses two synthetic pipelines processing the \SI{81}{\giga\byte} Big Brain
+image, and a real pipeline processing anatomical data from more than 1,000 subjects.
+We benchmark these pipelines using various combinations of task durations, data
+sizes, and numbers of workers, deployed on a 8-node (8 cores ea.) compute cluster in
+Compute Canada's Arbutus cloud. We evaluate PySpark's RDD API against Dask's Bag,
+Delayed and Futures. Results show that despite slight differences between Spark and
+Dask, both frameworks perform comparably. However, Dask pipelines risk being limited
+by Python's GIL depending on task type and cluster configuration. In all cases, the
+major limiting factor was data transfer. While either framework is suitable for
+neuroimaging pipelines, more effort needs to be placed in reducing data transfer time.
 \end{abstract}
@@
 
-Varying the block size is equivalent to varying the size of the partitions. We
-observed no difference when varying the task size while using an NFS. However, if
-one uses a parallel shared file system, we suppose it would be preferable to use a
-larger block size as the IO time should be reduced substantially. It is worth noting
-that the Dask developers suggest, in
+Varying the block size or the size of the partitions is equivalent. No difference is
+observed when varying the task size while using an NFS. However, if one uses a
+parallel shared file system, it would be preferable to use a larger block
+size as the IO time should be reduced substantially. It is worth noting that the Dask
+developers suggest, in
 \href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to

### 2019-07-29 latex/paper.tex
Renamed framework to engine

@@
 Delayed and Futures. Results show that despite slight differences between Spark and
-Dask, both frameworks perform comparably. However, Dask pipelines risk being limited
+Dask, both engines perform comparably. However, Dask pipelines risk being limited
 by Python's GIL depending on task type and cluster configuration. In all cases, the
-major limiting factor was data transfer. While either framework is suitable for
+major limiting factor was data transfer. While either engine is suitable for
 neuroimaging pipelines, more effort needs to be placed in reducing data transfer time.
@@
 intensive neuroimaging pipelines. We extend this work by studying the
-differences between two Big Data frameworks, Dask~\cite{Dask:15} and Apache
+differences between two Big Data engines, Dask~\cite{Dask:15} and Apache
 Spark~\cite{Spark:16}, for their suitability in the processing of neuroimaging
@@
 Spark and Dask both offer in-memory computing, data locality, and lazy evaluation,
-which is common for Big Data frameworks. Both their schedulers operate dynamically,
+which is common for Big Data engines. Both their schedulers operate dynamically,
 which is good when task runtimes are not known ahead of time~\cite{Dask:15}. They
@@
 infrastructure schedulers, such as Mesos~\cite{hindman2011mesos}, YARN~\cite{vavilapalli2013apache}, or HPC clusters (Dask only). Over
-these similarities, the frameworks have differences.
+these similarities, the engines have differences.
 
@@
 performance, although we recognize that other factors will also impact the
-choice of a framework in practice.
+choice of an engine in practice.
 
@@
 and DataFrames. Datasets are similar to RDDs but additionally use Spark SQL's
-optimized execution to further improve performance. DataFrames, used to process
+optimized execution engine to further improve performance. DataFrames, used to process
 tabular data, are Datasets where the data is organized into named-columns.
@@
 
-The Dask framework is compatible with numerous distributed schedulers, including 
+The Dask engine is compatible with numerous distributed schedulers, including 
 \href{https://github.com/dask/dask-yarn}{YARN} and
@@
 clusters, we chose to use Dask's Distributed scheduler, to keep the environment
-balanced between both frameworks.
+balanced between both engines.
 

### 2019-07-29 latex/paper.tex
typo

@@
 While serialization of Python to Java may lead to significant overheads, we
-chose to focus on Spark Python API due to its suitability for neuroimaging
+chose to focus on PySpark API due to its suitability for neuroimaging
 research.

### 2019-07-29 latex/paper.tex
Edits in results

--- 2019-07-29 latex/paper.tex
@@
 
-In the past few years, neuroimaging has entered the Big Data era due to the joint
-increase in image resolution, data sharing, and study sizes. However, no particular
-Big Data engines has emerged in this field, and several alternatives remain
-available. We compare two popular Big Data engines with Python APIs, Apache Spark and
-Dask, for their runtime performance in processing neuroimaging pipelines. Our
-evaluation uses two synthetic pipelines processing the \SI{81}{\giga\byte} Big Brain
-image, and a real pipeline processing anatomical data from more than 1,000 subjects.
-We benchmark these pipelines using various combinations of task durations, data
-sizes, and numbers of workers, deployed on a 8-node (8 cores ea.) compute cluster in
-Compute Canada's Arbutus cloud. We evaluate PySpark's RDD API against Dask's Bag,
-Delayed and Futures. Results show that despite slight differences between Spark and
-Dask, both frameworks perform comparably. However, Dask pipelines risk being limited
-by Python's GIL depending on task type and cluster configuration. In all cases, the
-major limiting factor was data transfer. While either framework is suitable for
-neuroimaging pipelines, more effort needs to be placed in reducing data transfer time.
+In the past few years, neuroimaging has entered the Big Data era due to the
+joint increase in image resolution, data sharing, and study sizes. However,
+no particular Big Data framework has emerged in this field, and several
+alternatives remain available. We compare two popular Big Data frameworks
+with Python APIs, Apache Spark and Dask, for their runtime performance in
+processing neuroimaging pipelines. Our evaluation uses two synthetic
+pipelines processing the 81~GB Big Brain image, and a real pipeline
+processing anatomical data from more than 1,000 subjects. We benchmark these pipelines using
+ various combinations of task durations, data sizes, and numbers of workers,
+deployed on a 64-node compute cluster in Compute Canada's Arbutus
+cloud. We evaluate PySpark's RDD API against Dask's Bag, Delayed and Futures.
+Results show that despite slight differences between Spark and Dask, both
+frameworks perform comparably. However, Dask pipelines risk being limited
+by Python's GIL depending on task type and cluster configuration. In all
+cases, the major limiting factor was data transfer. While either framework
+is suitable for neuroimaging pipelines, more effort needs to be placed in
+reducing data transfer time.
 \end{abstract}
@@
 
+% Related work
 Spark and Dask were both included in the evaluation reported
@@
 stays similar when the number of workers increases. However, the data
-transfer time and overhead increase proportionally to the number of
-workers with regression slopes: Spark (\SI{2337}{\second/task}), Bag
-(\SI{2650}{\second/task}), Delayed (\SI{3251}{\second/task}), Futures
-(\SI{3570}{\second/task}).
+transfer time and overhead increase substantially with the number of
+workers.
+
+% \hspace{1cm}\(\hat{y}=2336.9499\cdot{x} + 325.00574\) (Spark)
+
+% \hspace{1cm}\(\hat{y}=2650.13505\cdot{x} + 1585.37218\) (Bag)
+
+% \hspace{1cm}\(\hat{y}=3250.53406\cdot{x} + 1769.50351\) (Delayed)
+
+% \hspace{1cm}\(\hat{y}=3570.39758\cdot{x} + 1062.23031\) (Futures)
 
 On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
-slightly lower than Dask's, particularly when the number of workers increase.
-Within Dask, Delayed and Futures have a higher overhead than Bag. However,
-overhead differences are compensated by an increase in data transfer time,
-as a reduced overhead increases the concurrency between data transfers.
+slightly lower than Dask's, particularly when the number of workers
+increase. Moreover, Delayed and Futures have a higher overhead than Bag.
+However, overhead differences are compensated by an increase in data
+transfer time, as a reduced overhead increases the concurrency between data
+transfers and the application saturates the 10Gbits/s network bandwidth.
 
@@
 number of image blocks for constant BigBrain image size. We were not able to run
-Spark for 30 blocks due to its \SI{2}{\giga\byte} limitation in the task size it can
-compute. Once again, we do not observe any significant difference among the engines.
+Spark for 30 blocks due to its \SI{2}{\giga\byte} limitation in the task size.
+ Once again, we do not observe any substantial difference among the engines.
 For all engines, makespan variability increases with the number of blocks, however,
@@
 In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
-For 30 blocks, Delayed and Futures have much higher overhead. This is because they are
-thread-based and use multiple threads to compute a block which affect the calculation
-of the overhead. Once again, the data transfer time reduces with more blocks however
+For 30 blocks, the increased overhead of Delayed and Futures is a calculation artifact 
+coming from the fact that only 30 of the 64 available threads can be used concurrently. Once again, the data transfer time reduces with more blocks but 
 the overhead time increases by a similar amount. This is not observed for 30 blocks
@@
 again equivalent, although Delayed and Futures are slightly faster
-than Bag and, Spark for 1 and 10 iterations, and Futures are faster than
+than Bag and Spark for 1 and 10 iterations, and Futures are faster than
 Delayed, Bag and Spark for 100 iterations. Differences remain minor
@@
 charts are structured in batches of up to 64 read-compute-write concurrent sequences.
-Files read in the first batch are much longer than the ones in the following batches:
+File reads in the first batch are much longer than the ones in the following batches:
 this is due to the high synchronization of data transfers that leads to a high
 saturation of the shared file system. We also note that overhead, represented in
-white, is concentrated around the data transfer tasks, and the
+white, is concentrated around the data transfer tasks and the
 computing tasks that run concurrently with data transfers.
@@
 %%% HISTOGRAM %%%
-\subsection{Histogram: Number of workers}
-Figure~\ref{fig:histo_ms_worker} shows the makespan of the application for various
+\subsection{Histogram (Python): Number of workers}
+Figure~\ref{fig:histo_ms_worker} shows the makespan of the Histogram application for various
 amounts of workers. Spark is significantly faster than Dask APIs. The difference
-narrows as the number of workers increases but stays significant. Between Dask APIs
+narrows as the number of workers increases but it remains significant. Between Dask APIs
 there is no substantial difference, however, Bag is slightly faster on average.
@@
 shown. The compute time is significantly larger for Dask than for Spark. This is
-potentially due to Python's GIL preventing Dask to parallelize the computation
+presumably due to Python's GIL preventing Dask to parallelize the computation
 over
-multiple threads. Overall, the IO and overhead are comparable for all engines.
+multiple threads. Overall, the I/O and overhead times are comparable for all engines.
 
-On Figure~\ref{fig:histo_tt_worker}, Dask engines have similar total execution time
-when their number of workers varies, except for 8 workers where it increases slightly.
+Figure~\ref{fig:histo_tt_worker} shows that Dask engines have a similar
+total execution time when their number of workers varies, except for 8
+workers where it increases slightly due to increased read time.
 On the other hand, Spark total execution time keeps increasing; especially at 8
@@
         \includegraphics[clip,width=\columnwidth]{images/histo_worker.png}%
-        \caption{Histogram makespan}\label{fig:histo_ms_worker}
+        \caption{Histogram (Python) makespan}\label{fig:histo_ms_worker}
     \end{subfigure}
@@
         \includegraphics[clip,width=\columnwidth]{images/histo_idle_worker.png}%
-        \caption{Histogram total time}\label{fig:histo_tt_worker}
+        \caption{Histogram (Python) total time}\label{fig:histo_tt_worker}
     \end{subfigure}
@@
 
-\subsection{Histogram: Number of blocks}
+\subsection{Histogram (Python): Number of blocks}
 Figure~\ref{fig:histo_ms_block} shows the makespan for different block sizes. 
@@
 
-In figure~\ref{fig:histo_tt_block} the total time for each function is shown.
+In Figure~\ref{fig:histo_tt_block} the total time for each function is shown.
 Note that at 30 blocks the workers' resources are not fully used since the
@@
         \includegraphics[clip,width=\columnwidth]{images/histo_block.png}%
-        \caption{Histogram makespan}\label{fig:histo_ms_block}
+        \caption{Histogram (Python) makespan}\label{fig:histo_ms_block}
     \end{subfigure}
@@
         \includegraphics[clip,width=\columnwidth]{images/histo_idle_block.png}%
-        \caption{Histogram total time}\label{fig:histo_tt_block}
+        \caption{Histogram (Python) total time}\label{fig:histo_tt_block}
     \end{subfigure}
@@
 
-\subsection{Histogram: Gantt chart}
-Figure~\ref{fig:histo_gantt} shows the Gantt chart of the baseline Histogram
-experiment. Spark overhead is distributed through all the execution while Bag
-overhead happens mostly when a thread reads subsequent blocks and
-Delayed overhead is concentrated after reading all the blocks.
+\subsection{Histogram (Python): Gantt chart}
+Figure~\ref{fig:histo_gantt} shows the Gantt chart of the baseline
+Histogram experiment implemented in Python. Spark overhead is distributed
+through all the execution while Bag overhead happens mostly when a thread
+reads subsequent blocks and Delayed overhead is concentrated after reading
+all the blocks.
 
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-histo-baseline.html?download=1}{
         \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_histo_gantt.png}}
+        height=0.15\textheight]{images/delayed_histo_gantt.png}%
         \caption{Dask Delayed}\label{fig:histo_dask_delayed_gantt}
     \end{subfigure}
-    \caption{Histogram: 125 blocks, 8 workers, 8
+    \caption{Histogram (Python): 125 blocks, 8 workers, 8
     CPUs/worker}\label{fig:histo_gantt}
@@
 shared file system. There is no
-substantial difference between all engines, although, Delayed is slightly faster.
-
-From Figure~\ref{fig:histo_np_tt_worker}, the total execution time breakdown is
-shown. The overhead and read time increase proportionally to the number of
-workers thus increasing the total time. This is because (1) most workers read
-at the same time due to a low computation-IO ratio and (2) there is more
-intra-worker and inter-worker communication.
+substantial difference between all engines although Delayed is slightly faster.
+Figure~\ref{fig:histo_np_tt_worker} shows the total execution time
+breakdown. The overhead and read time increase proportionally to the number
+of workers thus increasing the total time.
 
@@
 
-In figure~\ref{fig:histo_np_gantt}, read tasks are initially similar for all
+In Figure~\ref{fig:histo_np_gantt}, read tasks are initially similar for all
 engines. However, when computation starts the subsequent read tasks are much
@@
 between the engines. The makespan scales well with the increase in workers,
-however, the scaling reduces considerably when reaching 8 workers. This is due
-to an increase in overhead and an IO bottleneck caused parallel accesses to 
-a non-distributed file system.
+however, the scaling reduces slightly when reaching 8 workers. This is due
+to an increase in I/O time caused by parallel accesses to 
+the shared file system.
 
@@
 shown. Each engine has a similar total time independently of its number of workers.
-This is due to the, almost, linear scaling of the engines. The overhead increases
-proportionally to the number of workers with regression slopes: Spark
-(\SI{224}{\second/task}), Bag (\SI{183}{\second/task}), Delayed
-(\SI{173}{\second/task}), Futures (\SI{132}{\second/task}). This is due to more
-communication between the scheduler and workers as well as intra-worker and
-inter-worker communication.
+Consistently with previous observations, the overhead increases substantially with the 
+number of workers.
+
+% The overhead increases
+% proportionally with the number of workers (see regression lines below). This is due
+% to more communication between the scheduler and workers as well as intra-worker and
+% inter-worker communication.
+
+% \hspace{1cm}\(\hat{y}=223.53586\cdot{x} + 91.616964\) (Spark)
+
+% \hspace{1cm}\(\hat{y}=182.82659\cdot{x} - 11.43834\) (Bag)
+
+% \hspace{1cm}\(\hat{y}=172.54183\cdot + 400.69433\) (Delayed)
+
+% \hspace{1cm}\(\hat{y}=132.41481\cdot{x} + 90.32813\) (Futures)
 

### 2019-07-29 latex/paper.tex
Convention formatting

--- 2019-07-29 latex/paper.tex
@@
  70\%, pause the worker at 80\%, and terminate the worker at 95\%. We configured
- Spark to use 1 executor per worker and \SI{25}{\giga\byte} of memory per executor, to leave 5~GB
- for off-heap. We configured the Spark driver to use 25~GB of memory, and used the
+ Spark to use 1 executor per worker and \SI{25}{\giga\byte} of memory per executor, to leave \SI{5}{\giga\byte}
+ for off-heap. We configured the Spark driver to use \SI{25}{\giga\byte} of memory, and used the
  default configuration for the master. We used the default configuration for worker
@@
 resolution of \SI{40}{\micro\metre}. We converted the blocks into the
-\href{https://nifti.nimh.nih.gov/nifti-1}{NifTI} format, a popular format
-in neuroimaging. We left the NifTI blocks uncompressed, resulting in 
+\href{https://nifti.nimh.nih.gov/nifti-1}{NIfTI} format, a popular format
+in neuroimaging. We left the NIfTI blocks uncompressed, resulting in 
 a total data size of \SI{81}{\giga\byte}. 
@@
 iterations, and
-finally writes the result as a NifTI image back to the shared file system.
+finally writes the result as a NIfTI image back to the shared file system.
 This application allows us to study the behaviour of the engines when all

### 2019-07-29 latex/paper.tex
Fixed comment in Acknowledgments

--- 2019-07-29 latex/paper.tex
@@
 the National Science and Engineering Research Council of Canada. We warmly thank
-Compute Canada and \TG{Westgrid?} for providing the cloud infrastructure used in
+Compute Canada and its regional center Westgrid for providing the cloud infrastructure used in
 these experiments, and the McGill Center for Integrative Neuroscience for giving us

### 2019-07-29 latex/paper.tex

@@
 \section{Conclusion}
-We presented a comparison of two Big Data engines: Spark and Dask. We studied the
-engines on three data-intensive neuroimaging applications. Our results show no
-substantial difference between the engines. Moreover, the majority of the execution
-time was spent on data transfer and overhead. We also observed a possible correlation
-between IO and overhead time: an higher overhead is compensated by a lower IO time and
-a low overhead is compensated by an high IO time. This would be due to concurrent IO
-tasks saturating the bandwidth. We suggest that current schedulers policies would benefit
-from knowing the task types, either IO or compute, it is processing concurrently.
-
-% \TG{Talk about engine overhead: (1) where is it coming from? (2) 
-% explain the impact on data transfers with a graph.}
-
-% \TG{In the dicussion, get back to it and mention that (1) we don't
-% always observe that, (2) for data-intensive applications, overheads don't really matter
-% as reducing them increases data trasnfers.}
+We presented a comparison of two Big Data engines, Apache Spark and Dask.
+We studied the engines on three data-intensive neuroimaging applications
+representative of common use cases. Our results show no substantial
+performance difference between the engines, except for compute-intensive
+Python applications where Dask is slower than Spark due to parallelism loss
+coming from Python's Global Interpreter Lock. Interestingly, differences in
+engine overheads do not impact performance due to their impact on data
+transfer times: higher overheads are almost exactly compensated by a lower
+transfer times when data transfers saturate the bandwidth. These results
+suggest that future research should focus on strategies to reduce the
+impact of data transfers on applications.
 

### 2019-07-29 latex/paper.tex
Edited discussion

@@
 
-\subsection{Effect of Network File System}
-We used an NFS as it is a common approach to store a large amount of data. However,
-it limits the available bandwidth since all workers interact with one disk. This
-causes data-intensive applications to take a significant amount of time for IO. As we
-expect, this effect is amplified when too many tasks simultaneously try to operate
-with the NFS at the same time as
-Fig.~\ref{fig:inc_gantt},\ref{fig:histo_gantt},\ref{fig:histo_np_gantt},\ref{fig:bids_gantt}
-shows. A parallel file system, such as Lustre~\cite{lustre}, would likely improve
-IO times of parallel applications simultaneously writing to shared storage.
-
-\subsection{Effect of data-intensive applications}
-In our experiments, Incrementation, Histogram (NumPy) and Histogram for Spark have
-low compute time while BIDS App example and Histogram for Dask have significantly
-longer compute time. When adding additional workers, we observe that the highly
-data-intensive applications scale poorly while compute-intensive applications scale
-almost linearly This is due to the limitation from the NFS, as mentioned above.
-
-Data transfer accounts for the majority of the execution time for data-intensive
-applications (see
-Fig.~\ref{fig:inc_gantt},\ref{fig:histo_spark_gantt},\ref{fig:histo_np_gantt}). We
-see that when a low number of IO tasks are overlapping, this diminishes the bandwidth
-limitation from the NFS. We want to explore solutions to desynchronize IO tasks in
-the aim to reduce the makespan of data-intensive applications.
-
-\subsection{Effect of the Global Interpreter Lock}
-Python's GIL prevents multi-threading in Python. Dask tries to minimize the
-consequences of the GIL by using external Python extensions. We found that Dask has
-issues to parallelize Python functions while Spark did not (see
-Fig.~\ref{fig:histo_tt_worker}). However, Dask was easily able to parallelize
-application with functions in Numpy (see
-Fig.~\ref{fig:inc_tt_worker},\ref{fig:histo_np_tt_worker}) and command-line
-applications (see Fig.~\ref{fig:bids_tt_worker}). This is because, for Dask, the
-instances are configured to have 1 worker process with 8 threads. Since the Histogram
-implementation does not release the GIL it can only use one of the thread at a given
-time thus it slows down the computation significantly.
-
-One should be careful to configure the workers properly depending on the type
-of functions executed in an application.
-
-\subsection{Effect of tasks' size}
-We measured the effect of various block sizes for constant BigBrain image for
-Incrementation, Histogram, and Histogram (NumPy). We observe that larger block size
-results in more IO compensated by lower overhead time. On the other hand, smaller
-block size results in higher overhead time compensated by reduced IO time.
-
-The number of partitions also has an impact on the size of the tasks. We set a
-partition for each block or subject. There is a trade-off for the number of
-partitions. On the one hand, a low amount of partition results in larger tasks thus
-in lower scheduling overhead. On the other hand, a higher number of partition
-results in smaller tasks hence more tasks fit in the workers' memory insuring that
-they always have something to work on.
-
-Varying the block size is equivalent to varying the size of the partitions. We
-observed no difference when varying the task size while using an NFS. However, if
-one uses a parallel shared file system, we suppose it would be preferable to use a
-larger block size as the IO time should be reduced substantially. It is worth noting
-that the Dask developers suggest, in
-\href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to
-use a partition size such that workers have 2-3 times as many chunks available to
-work on. More experiments would have to be performed to gain insight on an optimal
-number of partitions to use for a given set of data-intensive applications.
-
-\subsection{Other comments}
-Spark performance should be reduced by the serialization of Python functions to Java,
-however, our results show no significant impact of it.
-
-Previous work showed that Dask has a the largest start-up overhead~\cite{Mehta:17}
-when processing few subjects. Our Gantt charts show that Spark has a higher start-up
-overhead. We think this is due to the larger amount of subjects we process.
-
-The same study showed that Dask has a lower end-to-end runtime than Spark in a
-neuroimaging analysis application. In our results, there is no significant difference
-between the engines. We think this is because the applications we chose are highly
-data-intensive while the one used in their study is compute intensive.
+\subsection{Comparison with previous studies}
+
+Overall, our experiments did not show any substantial performance
+difference between Spark and Dask. In contradiction with the experiments
+of~\cite{Mehta:17}, we did not observe consequential overhead differences
+due to serialization or startup costs. This apparent contradiction is most
+likely coming from differences in applications: while the applications in
+~\cite{Mehta:17} were mostly compute-intensive, ours were more data-intensive.
+Consequently, in our case, potential
+overhead differences were compensated by reverse trends in data transfer
+time. Since our focus are Big Data applications, this leads us to conclude that 
+engines are equivalent. However, this conclusion doesn't invalidate the results of~\cite{Mehta:17} that 
+would apply to more compute-intensive applications.
+
+\subsection{Overhead / data transfer compensation}
+
+The observed almost exact compensation between differences in engine
+overheads and in data transfer times (see for instance
+Figure~\ref{fig:inc_tt_worker}) is not surprising. Indeed, when the transfer
+bandwidth is saturated, and in absence of service thrashing,
+desynchronizing file transfers usually does not reduce the makespan of data
+transfers. To take an extreme example, the makespan of $n$ concurrent file
+transfers of size $F$ on a system of bandwidth $B$ would be $nF/B$, as $n$
+transfers would share the bandwidth, which equals to the time required
+to transfer the $n$ files sequentially. In other words, saturated bandwidths
+give room for extra engine overhead as long as it desynchronizes data
+transfers. This could be an interesting idea to explore in data transfer optimization.
+
+\subsection{Effect of the shared file system}
+
+We used NFS as our shared file system as it is a common and
+easy-to-configure solution to share data between compute nodes. The
+performance of NFS, however, degrades substantially when accessed by
+concurrent tasks, as shown in the Gantt charts in
+Figures~\ref{fig:inc_gantt}, ~\ref{fig:histo_gantt}, ~\ref{fig:histo_np_gantt} 
+and~\ref{fig:bids_gantt}. This is due to the limitations imposed by network and disk
+bandwidths. While network bandwidth can hardly be overcome, disk bandwidth
+can be increased by storing data to multiple disks addressed by a parallel
+file system such a Lustre~\cite{lustre}. Such a configuration would likely
+reduce the impact of data transfers, although the main observed patterns
+should remain.
+
+\subsection{Effect of Python's Global Interpreter Lock}
+
+Although Dask reduces the impact of Python's GIL on parallelism by using external
+Python extensions, our Python histogram experiment (Figure
+~\ref{fig:histo_tt_worker}) showed that issues remain in Dask to
+parallelize Python functions. However, Dask was easily able to parallelize
+the same application when implemented with Numpy
+(Figure~\ref{fig:histo_np_tt_worker}) because in this case the computation
+happened in C, outside of the GIL. Similarly, the computation in
+command-line applications happens in a sub-process, hence outside of the
+GIL, which explains why the BIDS apps example application was not impacted
+by this issue. Although implemented in Python, Incrementation was not
+impacted either, as computation was only emulated through sleep time.
+
+This behavior is also the consequence of our configuration of Dask that
+created 1 worker process with 8 threads. Since the Histogram Python
+implementation does not release the GIL it can only use one of the thread
+at a given time thus it slows down the computation significantly. This
+behavior highlights the need to configure the workers properly depending on
+the type of functions executed in an application.
+
+\subsection{Effect of data partitioning}
+
+The number of data partitions also has an impact on application
+performance. In our experiments, we set a partition for each block or
+subject, resulting in a maximum of 750 partitions for the Big Brain
+applications, and 1,397 partitions for the BIDS app. The number of
+partitions obeys a classical trade-off: small numbers of partitions reduce 
+overhead, while large numbers of
+partitions increase parallelism. Since both Spark and Dask can
+work with custom partition sizes, we do not expect substantial differences
+coming from variations of this parameter.
+
+% Varying the block size is equivalent to varying the size of the partitions. We
+% observed no difference when varying the task size while using an NFS. However, if
+% one uses a parallel shared file system, we suppose it would be preferable to use a
+% larger block size as the IO time should be reduced substantially. It is worth noting
+% that the Dask developers suggest, in
+% \href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to
+% use a partition size such that workers have 2-3 times as many chunks available to
+% work on. More experiments would have to be performed to gain insight on an optimal
+% number of partitions to use for a given set of data-intensive applications.
 
@@
 We studied the engines on three data-intensive neuroimaging applications
-representative of common use cases. Our results show no substantial
-performance difference between the engines, except for compute-intensive
-Python applications where Dask is slower than Spark due to parallelism loss
-coming from Python's Global Interpreter Lock. Interestingly, differences in
+representative of common use cases. Overall, our results show no substantial
+performance difference between the engines. Interestingly, differences in
 engine overheads do not impact performance due to their impact on data

### 2019-07-29 latex/paper.tex

--- 2019-07-29 latex/paper.tex
@@
 
-In the past few years, neuroimaging has entered the Big Data era due to the
-joint increase in image resolution, data sharing, and study sizes. However,
-no particular Big Data framework has emerged in this field, and several
-alternatives remain available. We compare two popular Big Data frameworks
-with Python APIs, Apache Spark and Dask, for their runtime performance in
-processing neuroimaging pipelines. Our evaluation uses two synthetic
-pipelines processing the 81~GB Big Brain image, and a real pipeline
-processing anatomical data from more than 1,000 subjects. We benchmark these pipelines using
- various combinations of task durations, data sizes, and numbers of workers,
-deployed on a 64-node compute cluster in Compute Canada's Arbutus
-cloud. We evaluate PySpark's RDD API against Dask's Bag, Delayed and Futures.
-Results show that despite slight differences between Spark and Dask, both
-frameworks perform comparably. However, Dask pipelines risk being limited
-by Python's GIL depending on task type and cluster configuration. In all
-cases, the major limiting factor was data transfer. While either framework
-is suitable for neuroimaging pipelines, more effort needs to be placed in
-reducing data transfer time.
+In the past few years, neuroimaging has entered the Big Data era due to the joint
+increase in image resolution, data sharing, and study sizes. However, no particular
+Big Data engines has emerged in this field, and several alternatives remain
+available. We compare two popular Big Data engines with Python APIs, Apache Spark and
+Dask, for their runtime performance in processing neuroimaging pipelines. Our
+evaluation uses two synthetic pipelines processing the \SI{81}{\giga\byte} Big Brain
+image, and a real pipeline processing anatomical data from more than 1,000 subjects.
+We benchmark these pipelines using various combinations of task durations, data
+sizes, and numbers of workers, deployed on a 8-node (8 cores ea.) compute cluster in
+Compute Canada's Arbutus cloud. We evaluate PySpark's RDD API against Dask's Bag,
+Delayed and Futures. Results show that despite slight differences between Spark and
+Dask, both engines perform comparably. However, Dask pipelines risk being limited
+by Python's GIL depending on task type and cluster configuration. In all cases, the
+major limiting factor was data transfer. While either engine is suitable for
+neuroimaging pipelines, more effort needs to be placed in reducing data transfer time.
 \end{abstract}
@@
 infrastructure schedulers, such as Mesos~\cite{hindman2011mesos}, YARN~\cite{vavilapalli2013apache}, or HPC clusters (Dask only). Over
-these similarities, the frameworks have differences.
-
-% Differences
+these similarities, the engines have differences.
+
 First and foremost, Spark is written in Scala while Dask is in Python.
@@
 performance, although we recognize that other factors will also impact the
-choice of a framework in practice.
-
-% Related work
+choice of an engine in practice.
+
 Spark and Dask were both included in the evaluation reported
@@
 stays similar when the number of workers increases. However, the data
-transfer time and overhead increase substantially with the number of
-workers.
-
-% \hspace{1cm}\(\hat{y}=2336.9499\cdot{x} + 325.00574\) (Spark)
-
-% \hspace{1cm}\(\hat{y}=2650.13505\cdot{x} + 1585.37218\) (Bag)
-
-% \hspace{1cm}\(\hat{y}=3250.53406\cdot{x} + 1769.50351\) (Delayed)
-
-% \hspace{1cm}\(\hat{y}=3570.39758\cdot{x} + 1062.23031\) (Futures)
+transfer time and overhead increase proportionally to the number of
+workers with regression slopes: Spark (\SI{2337}{\second/task}), Bag
+(\SI{2650}{\second/task}), Delayed (\SI{3251}{\second/task}), Futures
+(\SI{3570}{\second/task}).
 
@@
 than Dask engines. Overall, for all engines, lower block size results in lower
-IO time but increases the overhead time. Again, this is due to large overhead
+data transfer time but increases the overhead time. Again, this is due to large overhead
 desynchronizing task thus reducing the NFS bottleneck.
@@
 the workers only use half their core, the total time is substantially the same
-for all block size. This is due to lower block sizes having a lower IO time,
+for all block size. This is due to lower block sizes having a lower data transfer time,
 however, it balances out by the resulting increase in overhead.
@@
 shorter for Delayed. This is due the desynchronization of the task reducing the
-IO bottleneck.
+data transfer bottleneck.
 
@@
 shown. Each engine has a similar total time independently of its number of workers.
-Consistently with previous observations, the overhead increases substantially with the 
-number of workers.
-
-% The overhead increases
-% proportionally with the number of workers (see regression lines below). This is due
-% to more communication between the scheduler and workers as well as intra-worker and
-% inter-worker communication.
-
-% \hspace{1cm}\(\hat{y}=223.53586\cdot{x} + 91.616964\) (Spark)
-
-% \hspace{1cm}\(\hat{y}=182.82659\cdot{x} - 11.43834\) (Bag)
-
-% \hspace{1cm}\(\hat{y}=172.54183\cdot + 400.69433\) (Delayed)
-
-% \hspace{1cm}\(\hat{y}=132.41481\cdot{x} + 90.32813\) (Futures)
+This is due to the, almost, linear scaling of the engines. The overhead increases
+proportionally to the number of workers with regression slopes: Spark
+(\SI{224}{\second/task}), Bag (\SI{183}{\second/task}), Delayed
+(\SI{173}{\second/task}), Futures (\SI{132}{\second/task}). This is due to more
+communication between the scheduler and workers as well as intra-worker and
+inter-worker communication.
 
@@
 it limits the available bandwidth since all workers interact with one disk. This
-causes data-intensive applications to take a significant amount of time for IO. As we
+causes data-intensive applications to take a significant amount of time for data transfer. As we
 expect, this effect is amplified when too many tasks simultaneously try to operate
@@
 shows. A parallel file system, such as Lustre~\cite{lustre}, would likely improve
-IO times of parallel applications simultaneously writing to shared storage.
+data transfer times of parallel applications simultaneously writing to shared storage.
 
@@
 Fig.~\ref{fig:inc_gantt},\ref{fig:histo_spark_gantt},\ref{fig:histo_np_gantt}). We
-see that when a low number of IO tasks are overlapping, this diminishes the bandwidth
-limitation from the NFS. We want to explore solutions to desynchronize IO tasks in
+see that when a low number of data transfer tasks are overlapping, this diminishes the bandwidth
+limitation from the NFS. We want to explore solutions to desynchronize data transfer tasks in
 the aim to reduce the makespan of data-intensive applications.
@@
 Incrementation, Histogram, and Histogram (NumPy). We observe that larger block size
-results in more IO compensated by lower overhead time. On the other hand, smaller
-block size results in higher overhead time compensated by reduced IO time.
+results in more data transfer compensated by lower overhead time and vice-versa.
 
@@
 
-Varying the block size is equivalent to varying the size of the partitions. We
-observed no difference when varying the task size while using an NFS. However, if
-one uses a parallel shared file system, we suppose it would be preferable to use a
-larger block size as the IO time should be reduced substantially. It is worth noting
-that the Dask developers suggest, in
+Varying the block size or the size of the partitions is equivalent. No difference is
+observed when varying the task size while using an NFS. However, if one uses a
+parallel shared file system, it would be preferable to use a larger block
+size as the data transfer time should be reduced substantially. It is worth noting that the Dask
+developers suggest, in
 \href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to

### 2019-07-30 latex/paper.tex
reword

--- 2019-07-29 latex/paper.tex
@@
 due to serialization or startup costs. This apparent contradiction is most
-likely coming from differences in applications: while the applications in
-~\cite{Mehta:17} were mostly compute-intensive, ours were more data-intensive.
-Consequently, in our case, potential
-overhead differences were compensated by reverse trends in data transfer
-time. Since our focus are Big Data applications, this leads us to conclude that 
-engines are equivalent. However, this conclusion doesn't invalidate the results of~\cite{Mehta:17} that 
-would apply to more compute-intensive applications.
+likely coming from differences in applications, engine parametrizations, or
+infrastructure characteristics. In our case, potential overhead differences were
+compensated by reverse trends in data transfer time. Since our focus are
+Big Data applications, this leads us to conclude that engines are
+equivalent.
 

### 2019-07-30 latex/paper.tex

@@
 The overhead calculation for each CPU is the end time of its last processed task
-minus the start time of the application. The summation of those results is the
-total overhead. \TG{This last sentence looks wrong to me.}
+minus its runtime. The summation of those results is the
+total overhead.
 

### 2019-07-30 latex/paper.tex
Changed url for interactive figures

@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/spark-inc-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/spark-inc-baseline.html}{
             \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-bag-inc-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/dask-bag-inc-baseline.html}{
         \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-inc-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/dask-delayed-inc-baseline.html}{
         \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-futures-inc-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/dask-futures-inc-baseline.html}{
         \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/spark-histo-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/spark-histo-baseline.html}{
         \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-bag-histo-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/dask-bag-histo-baseline.html}{
         \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-histo-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/dask-delayed-histo-baseline.html}{
         \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/spark-histo_np-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/spark-histo_np-baseline.html}{
         \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-bag-histo_np-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/dask-bag-histo_np-baseline.html}{
         \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-histo_np-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/dask-delayed-histo_np-baseline.html}{
         \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/spark-bids-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/spark-bids-baseline.html}{
         \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-bag-bids-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/dask-bag-bids-baseline.html}{
         \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-delayed-bids-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/dask-delayed-bids-baseline.html}{
         \includegraphics[clip,width=\columnwidth,
@@
     \begin{subfigure}[b]{\columnwidth}
-        \href{https://zenodo.org/record/3354434/files/dask-futures-bids-baseline.html?download=1}{
+        \href{https://mathdugre.github.io/paper-big-data-engines/dask-futures-bids-baseline.html}{
         \includegraphics[clip,width=\columnwidth,

### 2019-07-30 latex/paper.tex
fixed minor typos in abstract to methods

--- 2019-07-30 latex/paper.tex
@@
 Dask, for their runtime performance in processing neuroimaging pipelines. Our
-evaluation uses two synthetic pipelines processing the \SI{81}{\giga\byte} Big Brain
+evaluation uses two synthetic pipelines processing the \SI{81}{\giga\byte} BigBrain
 image, and a real pipeline processing anatomical data from more than 1,000 subjects.
@@
 data-intensive neuroimaging applications: high-resolution imaging,
-represented by the Big Brain data~\cite{Amunts:13}, and large functional
+represented by the BigBrain data~\cite{Amunts:13}, and large functional
 MRI studies, represented by data from the consortium for reliability and
@@
 of Spark's other data structures, namely, Datasets
-and DataFrames. Datasets are similar to RDDs but additionally use Spark SQL's
+and DataFrames. Datasets are similar to RDDs, but additionally use Spark SQL's
 optimized execution engine to further improve performance. DataFrames, used to process
@@
 Spark is compatible with three different schedulers: Spark
-standalone, YARN and Mesos.
-The Spark standalone scheduler is a simple default scheduler built into Spark.
+Standalone, YARN and Mesos.
+The Spark Standalone scheduler is a simple default scheduler built into Spark.
 In contrast, the YARN scheduler is primarily designed to schedule Hadoop-based 
@@
 increasingly popular in the scientific Python ecosystem. Like Spark, Dask avoids data transfers and needless computation and communication through
-in-memory computing, data locality, and lazy evaluation. Dask worklflows can 
+in-memory computing, data locality, and lazy evaluation. Dask workflows can 
 further reduce data transfer costs by leveraging multithreading whenever 
@@
 
-Three neuroimaging applications to evaluate the engines in different
-conditions were used. The first two, incrementation and histogram, are simple 
+We used three neuroimaging applications to evaluate the engines in different
+conditions. The first two, incrementation and histogram, are simple 
 synthetic applications representing map-only and map-reduce applications,
-respectively. The third one is a real neuroimaging application representative
+respectively. The third, is a real neuroimaging application representative
 of popular BIDS applications~\cite{gorgolewski2017bids}. All scripts used for our
@@
 performance. In our experiments, we set a partition for each block or
-subject, resulting in a maximum of 750 partitions for the Big Brain
+subject, resulting in a maximum of 750 partitions for the BigBrain
 applications, and 1,397 partitions for the BIDS app. The number of

### 2019-07-30 latex/paper.tex
typos

@@
 increase in image resolution, data sharing, and study sizes. However, no particular
-Big Data engines has emerged in this field, and several alternatives remain
+Big Data engines have emerged in this field, and several alternatives remain
 available. We compare two popular Big Data engines with Python APIs, Apache Spark and
@@
 We benchmark these pipelines using various combinations of task durations, data
-sizes, and numbers of workers, deployed on a 8-node (8 cores ea.) compute cluster in
+sizes, and numbers of workers, deployed on an 8-node (8 cores ea.) compute cluster in
 Compute Canada's Arbutus cloud. We evaluate PySpark's RDD API against Dask's Bag,
@@
 The recent rise in data sharing and improved data collection strategies
-have brought neuroimaging to the Big Data era ~\cite{ALFAROALMAGRO:18,
+have brought neuroimaging to the Big Data era~\cite{ALFAROALMAGRO:18,
 van2014human}. Existing neuroimaging workflow engines, such as
@@
 compute-intensive neuroimaging pipelines, but lack incorporated Big Data
-strategies (i.e. in-memory computing, data locality and lazy-evaluation) to
+strategies (i.e. in-memory computing, data locality, and lazy-evaluation) to
 improve performance of the increasingly prevalent data-intensive pipelines.
-As was noted in ~\cite{hayot2019performance}, in-memory computing, coupled
+As was noted in~\cite{hayot2019performance}, in-memory computing, coupled
 with data locality, can bring significant performance improvements to data
@@
 studies, using the main Dask and Spark APIs. The other sections present
-our results, discussion and conclusions.
+our results, discussion, and conclusions.
 
@@
 \subsubsection{Apache Spark} Apache Spark is a widely-used 
-general-purpose Big Data engine.
-Its main abstraction, the Resilient Distributed Dataset
-(RDD)~\cite{RDD}, is a fault-tolerant, parallel collection of data elements. It 
-achieves fault tolerance through the recording of data lineage, the
-sequence of operations used to modify the original data. The RDD is the basis
-of Spark's other data structures, namely, Datasets
-and DataFrames. Datasets are similar to RDDs, but additionally use Spark SQL's
-optimized execution engine to further improve performance. DataFrames, used to process
-tabular data, are Datasets where the data is organized into named-columns.
-While the DataFrame API exists in all the available language APIs, Datasets are
-limited to Scala and Java.
+general-purpose Big Data engine. Its main abstraction, the Resilient Distributed
+Dataset (RDD)~\cite{RDD}, is a fault-tolerant, parallel collection of data elements.
+It achieves fault tolerance through the recording of data lineage, the sequence of
+operations used to modify the original data. The RDD is the basis of Spark's other
+data structures, namely, Datasets and DataFrames. Datasets are similar to RDDs, but
+additionally use Spark SQL's optimized execution engine to further improve
+performance. DataFrames, used to process tabular data, are Datasets where the data is
+organized into named-columns. While the DataFrame API exists in all the available
+language APIs, Datasets are limited to Scala and Java.
 
@@
 computing. Data locality, popularized by MapReduce~\cite{dean2008mapreduce},
-schedules
-tasks as close as possible to where the data is stored. In-memory computing
-ensures data is maintained in memory whenever possible, as writing large 
-amounts data to slower storage devices may be costly. To reduce any unnecessary
-communication and computation, Spark also included lazy evaluation, which
-builds the entire task graph prior to execution to determine what needs to be
-computed.
+schedules tasks as close as possible to where the data is stored. In-memory computing
+ensures data is maintained in memory whenever possible, as writing large amounts of
+data to slower storage devices may be costly. To reduce any unnecessary communication
+and computation, Spark also included lazy evaluation, which builds the entire task
+graph prior to execution to determine what needs to be computed.
 
@@
 communication is not bounded by Python's GIL. Dask relies on data 
-lineage to achieve fault tolerance, however, unlike Spark, does not require
-operations to be coarse grained. Furthermore, Dask is lightweight and modular,
+lineage to achieve fault tolerance, however, unlike Spark, it does not require
+operations to be coarse-grained. Furthermore, Dask is lightweight and modular,
 allowing users to only install the components they require.
@@
 Delayed in that they support arbitrary tasks, but they operate in real-time
-rather than lazily. With the exception of Dask Bag, all APIs, by default, use
+rather than lazily. Except for Dask Bag, all APIs, by default, use
 the local multithreaded scheduler. Dask Bag, instead, relies on the local
-multiprocessing scheduler.  All Dask data structures, with the exception of
+multiprocessing scheduler.  All Dask data structures, except for
 Dask Array and Dask DataFrame, were used in our experiments.
@@
 finally writes the result as a NIfTI image back to the shared file system.
-This application allows us to study the behaviour of the engines when all
+This application allows us to study the behavior of the engines when all
 inputs are processed independently, in a map-only scenario (see
-Figure~\ref{fig:tg-inc}). This mimics the behaviour of analyzing multiple
+Figure~\ref{fig:tg-inc}). This mimics the behavior of analyzing multiple
 independent subjects in parallel.
@@
  Algorithm~\ref{alg:incrementation}. The total amount of shuffled data is,
- however, limited to \SI{2.62}{\mega\byte} per blocks as it only consists
- of image histograms. Two implementation are studied: a pure Python one,
+ however, limited to \SI{2.62}{\mega\byte} per block as it only consists
+ of image histograms. Two implementations are studied: a pure Python one,
  where frequencies are computed through Python dictionary manipulations, and one based
@@
 The image was preloaded on the NFS and Singularity version 3.2.1\-1.el7 was installed
-on all instances.
+for all instances.
 
@@
 slightly lower than Dask's, particularly when the number of workers
-increase. Moreover, Delayed and Futures have a higher overhead than Bag.
+increases. Moreover, Delayed and Futures have a higher overhead than Bag.
 However, overhead differences are compensated by an increase in data
@@
 From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is
-shown. The compute time is significantly larger for Dask than for Spark. This is
+shown. The computing time is significantly larger for Dask than for Spark. This is
 presumably due to Python's GIL preventing Dask to parallelize the computation
@@
 Figure~\ref{fig:histo_np_tt_block} shows the total time spent in each function.
-Once more, since Delayed is thread-based its overhead at 30 blocks is seems much
-higher that it really is. Considering the overhead error and that at 30 blocks
+Once more, since Delayed is thread-based its overhead at 30 blocks is much
+higher. Considering the overhead error and that at 30 blocks
 the workers only use half their core, the total time is substantially the same
@@
 engines. However, when computation starts the subsequent read tasks are much
-shorter for Delayed. This is due the desynchronization of the task reducing the
+shorter for Delayed. This is due to the desynchronization of the task reducing the
 data transfer bottleneck.
@@
 Figure~\ref{fig:bids_ms_worker} shows the makespan of the application when
-varying the amount of workers. Overall, there is no substantial difference
+varying the number of workers. Overall, there is no substantial difference
 between the engines. The makespan scales well with the increase in workers,
@@
 Overall, our experiments did not show any substantial performance
-difference between Spark and Dask. In contradiction with the experiments
+difference between Spark and Dask. In contradiction to the experiments
 of~\cite{Mehta:17}, we did not observe consequential overhead differences
@@
 infrastructure characteristics. In our case, potential overhead differences were
-compensated by reverse trends in data transfer time. Since our focus are
+compensated by reverse trends in data transfer time. Since our focus is
 Big Data applications, this leads us to conclude that engines are
@@
 Figure~\ref{fig:inc_tt_worker}) is not surprising. Indeed, when the transfer
-bandwidth is saturated, and in absence of service thrashing,
+bandwidth is saturated, and in the absence of service thrashing,
 desynchronizing file transfers usually does not reduce the makespan of data
@@
 engine overheads do not impact performance due to their impact on data
-transfer times: higher overheads are almost exactly compensated by a lower
-transfer times when data transfers saturate the bandwidth. These results
+transfer time: higher overheads are almost exactly compensated by a lower
+transfer time when data transfers saturate the bandwidth. These results
 suggest that future research should focus on strategies to reduce the

### 2019-07-30 latex/paper.tex
Final edits

@@
 The overhead calculation for each CPU is the end time of its last processed task
-minus its runtime. The summation of those results is the
+minus the total runtime of the tasks it executed. The summation of those results is the
 total overhead.
@@
     \end{subfigure}
-    \caption{Incrementation: 125 blocks, 10 iterations, \SI{4}{\second} delay, 8 workers,
-8 CPUs/worker}\label{fig:inc_gantt}
+    \caption{Incrementation Gantt chart: 125 blocks, 10 iterations, \SI{4}{\second} delay, 8 workers,
+8 CPUs/worker. All Gantt charts are clickable and link to interactive figures with additional information.}\label{fig:inc_gantt}
 \end{figure}
@@
     \end{subfigure}
-    \caption{Histogram (Python): 125 blocks, 8 workers, 8
+    \caption{Histogram (Python) Gantt chart: 125 blocks, 8 workers, 8
     CPUs/worker}\label{fig:histo_gantt}
@@
     \end{subfigure}
-    \caption{Histogram (NumPy): 125 blocks, 8 workers, 8
+    \caption{Histogram (NumPy) Gantt chart: 125 blocks, 8 workers, 8
     CPUs/worker}\label{fig:histo_np_gantt}
@@
     \end{subfigure}
-    \caption{BIDS App example: 8 workers}\label{fig:bids_gantt}
+    \caption{BIDS App example Gantt chart: 8 workers}\label{fig:bids_gantt}
 \end{figure}
@@
 Overall, our experiments did not show any substantial performance
-difference between Spark and Dask. In contradiction to the experiments
+difference between Spark and Dask. In contradiction with the experiments
 of~\cite{Mehta:17}, we did not observe consequential overhead differences
 due to serialization or startup costs. This apparent contradiction is most
-likely coming from differences in applications, engine parametrizations, or
-infrastructure characteristics. In our case, potential overhead differences were
+likely coming from differences in applications, engine parametrization, or
+infrastructure characteristics between the two studies. In our case, potential overhead differences were
 compensated by reverse trends in data transfer time. Since our focus is

### 2019-09-26 latex/paper.tex
Positioned figures
Added figure number to GIL paragraph

@@
     \end{subfigure}
-    \caption{125 blocks, 8 CPUs/worker}
+    \caption{125 blocks, 8 CPUs/worker}\label{fig:histo_worker}
 \end{figure}
@@
     \end{subfigure}
-    \caption{8 workers, 8 CPUs/worker}
+    \caption{8 workers, 8 CPUs/worker}\label{fig:histo_block}
 \end{figure}
@@
 
-\begin{figure}[!b]
+\begin{figure}[!t]
     \centering
@@
     \end{subfigure}
-    \caption{125 blocks, 8 CPUs/worker}
+    \caption{125 blocks, 8 CPUs/worker}\label{fig:histo_np_worker}
 \end{figure}
@@
     \end{subfigure}
-    \caption{8 workers, 8 CPUs/worker}
+    \caption{8 workers, 8 CPUs/worker}\label{fig:histo_np_block}
 \end{figure}
@@
 Python extensions, our Python histogram experiment (Figure
-~\ref{fig:histo_tt_worker}) showed that issues remain in Dask to
+~\ref{fig:histo_worker} and \ref{fig:histo_block}) showed that issues remain in Dask to
 parallelize Python functions. However, Dask was easily able to parallelize
 the same application when implemented with Numpy
-(Figure~\ref{fig:histo_np_tt_worker}) because in this case the computation
+(Figure~\ref{fig:histo_np_worker} and \ref{fig:histo_np_block}) because in this case the computation
 happened in C, outside of the GIL. Similarly, the computation in

### 2019-09-26 latex/paper.tex
Modified Introduction

@@
 % Context
-The recent rise in data sharing and improved data collection strategies
-have brought neuroimaging to the Big Data era~\cite{ALFAROALMAGRO:18,
-van2014human}. Existing neuroimaging workflow engines, such as
-Nipype~\cite{Nipype:11}, are well suited for processing the standard
-compute-intensive neuroimaging pipelines, but lack incorporated Big Data
-strategies (i.e. in-memory computing, data locality, and lazy-evaluation) to
-improve performance of the increasingly prevalent data-intensive pipelines.
-As was noted in~\cite{hayot2019performance}, in-memory computing, coupled
-with data locality, can bring significant performance improvements to data
-intensive neuroimaging pipelines. We extend this work by studying the
+The recent rise in data sharing and improved data collection strategies have brought
+neuroimaging to the Big Data era~\cite{ALFAROALMAGRO:18, van2014human}. Existing
+neuroimaging workflow engines, such as Nipype~\cite{Nipype:11}, are well suited for
+processing the standard compute-intensive neuroimaging pipelines, but lack
+incorporated Big Data strategies (i.e. in-memory computing, data locality, and
+lazy-evaluation) to improve performance of the increasingly prevalent data-intensive
+pipelines. As was noted in~\cite{hayot2019performance}, in-memory computing, coupled
+with data locality, can bring significant performance improvements to data intensive
+neuroimaging pipelines. We are interested in the impact on performance when
+implementing these Big Data strategies in different engines. Our work studies the
 differences between two Big Data engines, Dask~\cite{Dask:15} and Apache
 Spark~\cite{Spark:16}, for their suitability in the processing of neuroimaging
-pipelines. Our goal is to test whether Spark or Dask has a clear
-performance advantage to process Big neuroimaging Data.
+pipelines. Our goal is to test whether Spark or Dask has a clear performance
+advantage to process Big neuroimaging Data.
 
@@
 transfer time, as a reduced overhead increases the concurrency between data
-transfers and the application saturates the 10Gbits/s network bandwidth.
+transfers and the application saturates the disk bandwidth of the shared file system.
 

### 2019-09-27 latex/paper.tex
Added paragraph on Impact of the NFS of the different applications

@@
 Department of Computer Science and Software Engineering\\
-Concordia University, Montr\'eal, Qu\'ebec, Canada 
+Concordia University, Montr\'eal, Qu\'ebec, Canada\\
+\{mathieu.dugre, valerie.hayot-sasson, tristan.glatard\}@concordia.ca
 \vspace*{0.8cm} % to avoid weird spacing of 1st page by Latex.
@@
 ecosystem, Dask provides almost transparent parallelization of applications
-manipulating Numpy arrays or Pandas data frames. On the other hand, Spark's
+manipulating NumPy arrays or Pandas data frames. On the other hand, Spark's
 Java, R and Python APIs allow to easily parallelize analyses that combine
@@
 
+Throughout the applications, the use of large data blocks amplified the consequences
+of using an NFS. Moreover, the bandwidth saturation almost neglects the scaling, from
+additional workers, for data-intensive applications: Incrementation and Histogram
+(NumPy). On the other hand, an almost linear scaling shows that it has no substantial
+effect on compute-intensive applications: Histogram (Python) and BIDS app example.
+
+We suppose that using a parallel file system would allow applications to have almost
+linear scaling as long as they do not saturate the bandwidth of the system.
+
 \subsection{Effect of Python's Global Interpreter Lock}
@@
 parallelize Python functions. However, Dask was easily able to parallelize
-the same application when implemented with Numpy
+the same application when implemented with NumPy
 (Figure~\ref{fig:histo_np_worker} and \ref{fig:histo_np_block}) because in this case the computation

### 2019-09-30 latex/paper.tex
Modified intro
    Integrated sugestion from Tirstan comment.

@@
 with data locality, can bring significant performance improvements to data intensive
-neuroimaging pipelines. We are interested in the impact on performance when
-implementing these Big Data strategies in different engines. Our work studies the
-differences between two Big Data engines, Dask~\cite{Dask:15} and Apache
-Spark~\cite{Spark:16}, for their suitability in the processing of neuroimaging
-pipelines. Our goal is to test whether Spark or Dask has a clear performance
-advantage to process Big neuroimaging Data.
+neuroimaging pipelines. In this work, we investigate how performance benchmarks
+generalize across Big Data engines. In particular, we study the differences between
+Dask~\cite{Dask:15} and Apache Spark~\cite{Spark:16}, for their suitability in the
+processing of neuroimaging pipelines. Our goal is to test whether Spark or Dask has a
+clear performance advantage to process Big neuroimaging Data.
 

### 2019-10-01 latex/paper.tex
Modified GIL paragraph
        Expanded on our suggested configuration to avoid undesirable performance
    behaviour as in the Histogram (Python) application

@@
 implementation does not release the GIL it can only use one of the thread
-at a given time thus it slows down the computation significantly. This
-behavior highlights the need to configure the workers properly depending on
-the type of functions executed in an application.
+at a given time thus it slows down the computation significantly. 
+% This behavior highlights the need to configure the workers properly depending on
+% the type of functions executed in an application. 
+Our results suggest configuring Dask with multi-threaded workers for application
+releasing the GIL and multi-process workers while the default configuration of Spark
+workers brings stable results across different types of applications.
 

### 2021-01-15 latex/paper.tex
Partial draft for the paper.

--- 2019-09-30 latex/paper.tex
@@
 
-\title{A performance comparison of Dask and Apache Spark for data-intensive neuroimaging pipelines}
+\title{Impact of engines scheduler on the performance of Big Data pipelines in HPC}
 
@@
 \begin{abstract}
-
-In the past few years, neuroimaging has entered the Big Data era due to the joint
-increase in image resolution, data sharing, and study sizes. However, no particular
-Big Data engines have emerged in this field, and several alternatives remain
-available. We compare two popular Big Data engines with Python APIs, Apache Spark and
-Dask, for their runtime performance in processing neuroimaging pipelines. Our
-evaluation uses two synthetic pipelines processing the \SI{81}{\giga\byte} BigBrain
-image, and a real pipeline processing anatomical data from more than 1,000 subjects.
-We benchmark these pipelines using various combinations of task durations, data
-sizes, and numbers of workers, deployed on an 8-node (8 cores ea.) compute cluster in
-Compute Canada's Arbutus cloud. We evaluate PySpark's RDD API against Dask's Bag,
-Delayed and Futures. Results show that despite slight differences between Spark and
-Dask, both engines perform comparably. However, Dask pipelines risk being limited
-by Python's GIL depending on task type and cluster configuration. In all cases, the
-major limiting factor was data transfer. While either engine is suitable for
-neuroimaging pipelines, more effort needs to be placed in reducing data transfer time.
+	% TODO
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Dask, Apache Spark, performance, neuroimaging
+	Performance, Scheduler, Big Data, Dask, Spark, Ray, Neuroimaging
 \end{IEEEkeywords}
+\MD{Discuss which keyword to use for the paper.}
 
 \section{Introduction}
-% Context
-The recent rise in data sharing and improved data collection strategies have brought
-neuroimaging to the Big Data era~\cite{ALFAROALMAGRO:18, van2014human}. Existing
-neuroimaging workflow engines, such as Nipype~\cite{Nipype:11}, are well suited for
-processing the standard compute-intensive neuroimaging pipelines, but lack
-incorporated Big Data strategies (i.e. in-memory computing, data locality, and
-lazy-evaluation) to improve performance of the increasingly prevalent data-intensive
-pipelines. As was noted in~\cite{hayot2019performance}, in-memory computing, coupled
-with data locality, can bring significant performance improvements to data intensive
-neuroimaging pipelines. In this work, we investigate how performance benchmarks
-generalize across Big Data engines. In particular, we study the differences between
-Dask~\cite{Dask:15} and Apache Spark~\cite{Spark:16}, for their suitability in the
-processing of neuroimaging pipelines. Our goal is to test whether Spark or Dask has a
-clear performance advantage to process Big neuroimaging Data.
 
-%Similarities
-Spark and Dask both offer in-memory computing, data locality, and lazy evaluation,
-which is common for Big Data engines. Both their schedulers operate dynamically,
-which is good when task runtimes are not known ahead of time~\cite{Dask:15}. They
-also provide rich, high-level programming APIs, and support a variety of
-infrastructure schedulers, such as Mesos~\cite{hindman2011mesos}, YARN~\cite{vavilapalli2013apache}, or HPC clusters (Dask only). Over
-these similarities, the engines have differences.
+\section{Background}
+\subsection{Lustre} % TODO
 
-First and foremost, Spark is written in Scala while Dask is in Python.
-Given the popularity of Python in scientific communities, this arguably
-gives an edge to Dask due to data serialization costs from Python to Scala.
-On the other hand, Python's Global Interpreter Lock (GIL) might reduce
-parallelism in some cases. This difference in programming languages also
-has qualitative implications. As part of the \href{http://scipy.org}{SciPy}
-ecosystem, Dask provides almost transparent parallelization of applications
-manipulating NumPy arrays or Pandas data frames. On the other hand, Spark's
-Java, R and Python APIs allow to easily parallelize analyses that combine
-these languages, with reduced performance loss. Our study focuses on
-performance, although we recognize that other factors will also impact the
-choice of an engine in practice.
+\subsection{Dask}
+Dask is a Python-based Big Data engine with growing popularity in the scientific Python ecosystem.
+To mitigate the data transfer bottleneck in Big Data workflows, Dask was designed with data locality and in-memory in mind.
+Data locality, popularized by Map-Reduce \cite{dean2008mapreduce}, schedules tasks where the data reside.
+In-memory computing minimize the overhead of transfering data to disk by keeping data in memory when possible.
+Dask uses lazy evalutation to reduce unnecessary communication and computation.
+The engine builds a dynamic graph prior to execution allowing to determine which task to compute.
+Dask workflows can further reduce data transfer by using  leveraging mulithreading whenever the application is not bounded by Python's GIL.
+Fault-tolerance is achieved by recording of data lineage, the sequence of operations used on to modify the initial data.
 
-Spark and Dask were both included in the evaluation reported
-in~\cite{Mehta:17}, where a neuroimaging application
-processed approximately \SI{100}{\giga\byte} of data. In this work, Dask was
-reported to have a slight performance advantage over Spark. Overall, Dask's
-end-to-end time (makespan) was measured to be up to 14\% faster than Spark,
-due to ``more efficient pipelining'' %(Fig 10) 
-and serialization time to
-Python. % (Fig. 12-a)
-Dask, however, was reported to have a larger startup
-time than Spark. % (Fig. 12-c)
-The analysis remained at a quite high level
-though, leaving most of the observed performance difference unexplained. In
-comparison, our study will provide a detailed analysis of performance
-differences and similarities between Spark and Dask for neuroimaging data processing.
+Dask offers five data structures: Array, Bag, DataFrame, Delayed, and Futures. \MD{Insert link to the API.}
+The Dask Array offers a clone of NumPy API for distributed processing of large arrays.
+A Dask Bag is a distributed collection of Python object that offers a programming abstraction similar to PyToolz. \MD{Link to PyToolz.}
+Dask Dataframe is a parallel composition of Pandas Dataframes used to process large amount of tabular data.
+Dask Delayed offers an API for distributing abrirtary function that do not fit in the above frameworks.
+Lastly, Dask Futures can also execute abrirtary functions, however, it launch computation immediately rather than lazily.
+Dask modularity allow users to install only required components making it lightweigth.
 
-The next section details the design of our benchmarking experiments. We consider two
-data-intensive neuroimaging applications: high-resolution imaging,
-represented by the BigBrain data~\cite{Amunts:13}, and large functional
-MRI studies, represented by data from the consortium for reliability and
-reproducibility (CoRR~\cite{zuo2014open}). We test application pipelines
-involving different patterns (map-only, map-reduce), and different types of
-implementations (plain Python, command-line,
-containerized). We evaluate performance on a dedicated cluster
-representative of the ones used in today's data-intensive neuroimaging
-studies, using the main Dask and Spark APIs. The other sections present
-our results, discussion, and conclusions.
+In Dask a scheduler decides where and when to execute tasks using the Dask graph.
+API operations generate multiple fine-coarsed tasks in the computation graph allwing an easier representatin of complex algorithms.
 
-%%%%% MATERIAL AND METHODS %%%%%%
-\section{Material and Methods}
+The Dask engine is compatible to numerous distributed schedulers, including YARN and Mesos.
+Dask also provide its own \textit{Dask Distributed scheduler}.
+Althought Dask is compatible with schedulers available in HPC clusters,
+We chose to use Dask Distributed scheduler, to keep the environment balanced between the engines.
 
-\subsection{Engines}
+In Dask Distributed scheduler, a \textit{dask-scheduler} process admistrates the resource provided by \textit{dask-worker} in the cluster.
+The scheduler receives jobs from clients and assigns tasks to available workers.
+Dask scheduler use a FIFO (First-In-First-Out) job scheduling policy.
+That is, completely process a branch of the Dask graph before proceding with the next one.
 
-\subsubsection{Apache Spark} Apache Spark is a widely-used 
-general-purpose Big Data engine. Its main abstraction, the Resilient Distributed
-Dataset (RDD)~\cite{RDD}, is a fault-tolerant, parallel collection of data elements.
-It achieves fault tolerance through the recording of data lineage, the sequence of
-operations used to modify the original data. The RDD is the basis of Spark's other
-data structures, namely, Datasets and DataFrames. Datasets are similar to RDDs, but
-additionally use Spark SQL's optimized execution engine to further improve
-performance. DataFrames, used to process tabular data, are Datasets where the data is
-organized into named-columns. While the DataFrame API exists in all the available
-language APIs, Datasets are limited to Scala and Java.
+Dask offers multiple ways to deploy a cluster including: with SSH configs, Kubernetes, SLURM, PBS, and others.
+For our experiments, we used the \href{https://jobqueue.dask.org/en/latest/generated/dask_jobqueue.SLURMCluster.html}{Dask SLURM cluster} API.
 
-As data transfers in Big Data workflows are an important source of performance
-bottlenecks, Spark incorporates the concepts of data locality and in-memory
-computing. Data locality, popularized by MapReduce~\cite{dean2008mapreduce},
-schedules tasks as close as possible to where the data is stored. In-memory computing
-ensures data is maintained in memory whenever possible, as writing large amounts of
-data to slower storage devices may be costly. To reduce any unnecessary communication
-and computation, Spark also included lazy evaluation, which builds the entire task
-graph prior to execution to determine what needs to be computed.
+Dask \MD{Add version when after locking it for experiments.} was used.
 
-% Didn't mention Kubernetes as it's still experimental
-Spark is compatible with three different schedulers: Spark
-Standalone, YARN and Mesos.
-The Spark Standalone scheduler is a simple default scheduler built into Spark.
-In contrast, the YARN scheduler is primarily designed to schedule Hadoop-based 
-workflows, whereas Mesos can be used to schedule
-a variety of different workflows. As researchers would likely be
-executing their workflows in HPC environments with neither YARN or Mesos
-installed, we limit our focus to Spark's Standalone scheduler.
+\subsection{Ray}
 
-The Spark Standalone scheduler is composed of three main processes: the
-\emph{master}, the \emph{workers} (\emph{slaves}, in Spark) and the
-\emph{driver}. The \emph{master} coordinates resources provisioned by
-\emph{worker} processes on the cluster. The application is submitted to the
-\emph{driver} that in turn requests workers to the master and dispatches tasks
-to them. A job is divided into stages to be executed in a different process onto
-the workers. Each stage's operation is represented as a high-level task in the
-computation graph. The Spark standalone scheduler uses a FIFO 
-(First-In-First-Out) job scheduling policy, and it has
-two execution modes: client mode, where the driver runs in a dedicated process
-outside of the Spark cluster, and cluster mode, where the driver runs within
-a worker process. Our experiments use client mode as cluster mode is not
-available in PySpark.
+\subsection{Apache Spark}
+Apache Spark is a widely-used general pupose Big Data engine.
+Like Dask, it aims at reducing data transfer cost by incorporating data locality and in-memory computing, and lazy evaluation.
 
-Python is commonly selected as the programming language of choice in scientific
-communities, and in particular, for our use case of neuroscience, where
-numerous specialized Python libraries exist to study the data.
-While serialization of Python to Java may lead to significant overheads, we
-chose to focus on PySpark API due to its suitability for neuroimaging
-research.
+Spark primary data structure is Resilient Distributed Dataset (RDD)\cite{RDD}, a fault-tolerant, parallel collection of data elements.
+RDDs is the basis of the other Spark data structure: Datasets and DataFrames.
+Datasets are similar to RDD but benefits from additional performance by leveraging the Spark SQL's optimized execution engine.
+The DataFrame API are Datasets organized into named-columns and used to process tabular data.
+While the DataFrame API is available in all supported language, DAtasets are limited to Sacla and Java.
 
-We used Apache Spark v2.4.0.
+Python is a common programming language in the scientific community.
+This is particulary true in the neuroimaging field with numerous specialized Python libraries to operate on data.
+While serialization from Python to Java create overhead, we found it to be minimal. \MD{cite SC19 paper.}
+We chose to focus on PySpark API to have a more balance environment between the different engines and due to its suitability to neuroimaging.
 
-\subsubsection{Dask} Dask is a Python-based Big Data engine that is becoming
-increasingly popular in the scientific Python ecosystem. Like Spark, Dask avoids data transfers and needless computation and communication through
-in-memory computing, data locality, and lazy evaluation. Dask workflows can 
-further reduce data transfer costs by leveraging multithreading whenever 
-communication is not bounded by Python's GIL. Dask relies on data 
-lineage to achieve fault tolerance, however, unlike Spark, it does not require
-operations to be coarse-grained. Furthermore, Dask is lightweight and modular,
-allowing users to only install the components they require.
-
-Dask has five main data structures:
-\href{https://docs.dask.org/en/latest/array.html}{Array}, \href{https://docs.dask.org/en/latest/bag.html}{Bag},
-\href{https://docs.dask.org/en/latest/dataframe.html}{DataFrame},
- \href{https://docs.dask.org/en/latest/delayed.html}{Delayed}, 
- and \href{https://docs.dask.org/en/latest/futures.html}{Futures}. A Dask
-Array is used for the
-processing of large arrays. It provides a distributed clone of the popular NumPy
-library. Similar to Spark's RDD, a Dask
-Bag is a parallel
-collection of Python objects. It offers a programming abstraction similar to the
-\href{https://toolz.readthedocs.io/en/latest/}{PyToolz library}. A Dask
-Dataframe is a parallel
-composition of
-\href{http://pandas.pydata.org/pandas-docs/stable/reference/api/pandas.DataFrame.html}{Pandas Dataframes},
- used to process a large amount of tabular data.
-Dask Delayed supports
-arbitrary tasks
-that do not fit in the Array, DataFrame or Bag APIs. Finally,
-Dask Futures are similar to
-Delayed in that they support arbitrary tasks, but they operate in real-time
-rather than lazily. Except for Dask Bag, all APIs, by default, use
-the local multithreaded scheduler. Dask Bag, instead, relies on the local
-multiprocessing scheduler.  All Dask data structures, except for
-Dask Array and Dask DataFrame, were used in our experiments.
-
-The Dask graph is the internal representation of a Dask application to be
-executed  by the scheduler. API operations generate multiple small tasks in
-the computation graph, allowing an easier representation of complex
-algorithms.
-
-The Dask engine is compatible with numerous distributed schedulers, including 
-\href{https://github.com/dask/dask-yarn}{YARN} and
-\href{https://github.com/mrocklin/dask-mesos}{Mesos}, similarly to Spark. Dask 
-also provides its own distributed scheduler, known as the Dask Distributed
-scheduler. Although Dask is compatible with schedulers commonly found in HPC
-clusters, we chose to use Dask's Distributed scheduler, to keep the environment
-balanced between both engines.
-
-In the Dask Distributed scheduler, a process called \textit{dask-scheduler}
-administrates the resources provided by \textit{workers} in the cluster. The
-scheduler receives jobs from clients and assigns tasks to available workers.
-Similarly to Spark's scheduler, Dask Distributed completes the processing of a
-graph branch before moving along to the next one.
-
-In our experiments, we used Dask v1.1.4.
-
-\subsection{Infrastructure}
-
- We used Compute Canada's
- \href{https://docs.computecanada.ca/wiki/Cloud\_resources}{Arbutus Cloud} operated by
- the \href{https://www.westgrid.ca}{WestGrid} regional organization at the University
- of Victoria, running OpenStack Queens release. We used c8-30gb-186 cloud
- instances with 8 VCPUs, an Intel Xeon Gold 6130 processor, \SI{30}{\giga\byte} of
- RAM at \SI{2666}{\mega\hertz}, \SI{20}{\giga\byte} of mounted storage, and a base
- image running CentOS 7.5.1804 with Linux kernel version
- 3.10.0\-862.11.6.el7.x86\_64. Instances were connected by a
- \SI{10}{\giga\bit/\second} Ethernet network.
-
- Cloud instances hosted a single Dask or Spark worker, configured to use 8 CPUs. We
- used Dask's default configuration that uses all the available memory on the
- instance. Its default heuristic is to: target a 60\% memory load, spill to disk at
- 70\%, pause the worker at 80\%, and terminate the worker at 95\%. We configured
- Spark to use 1 executor per worker and \SI{25}{\giga\byte} of memory per executor, to leave \SI{5}{\giga\byte}
- for off-heap. We configured the Spark driver to use \SI{25}{\giga\byte} of memory, and used the
- default configuration for the master. We used the default configuration for worker
- memory management: at 60\% it spills data to disk, and 50\% of that amount is
- reserved for storage that is immune to eviction.
-
- One cloud instance did not host any worker and had a \SI{2}{\tera\byte} disk volume
- shared with the other instances using the Network File System (NFS) v4.
- This instance was also used for the Spark driver and master, for the
- Dask scheduler, and for job monitoring with the Spark and Dask user
- interfaces. For both Spark and Dask, spilled data was evicted to the NFS.
-
-\subsection{Dataset}
-
-We used BigBrain~\cite{Amunts:13}, a three-dimensional image of a human
-brain with voxel intensities ranging from 0 to 65,535. The original data is
-stored in 125 blocks in the MINC~\cite{minc} HDF5-based format, available
-at \url{ftp://bigbrain.loris.ca/BigBrainRelease.2015/3D\_Blocks/40um} at a
-resolution of \SI{40}{\micro\metre}. We converted the blocks into the
-\href{https://nifti.nimh.nih.gov/nifti-1}{NIfTI} format, a popular format
-in neuroimaging. We left the NIfTI blocks uncompressed, resulting in 
-a total data size of \SI{81}{\giga\byte}. 
-To evaluate the effect of block size, we resplit these blocks into 30, 125 and 750 blocks of 
-\SI{2.7}{\giga\byte}, \SI{0.648}{\giga\byte}, and
-\SI{0.108}{\giga\byte}, using the \href{https://github.com/big-data-lab-team/sam}{sam} library.
-
-We also used the dataset provided by the Consortium for Reliability and
-Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}~\cite{zuo2014open}) as
-available on \href{http://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
-entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and
-functional images of 1,397 subjects acquired in 29 sites. We used all 3,491 anatomical
-images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} per image on
-average).
+\section{Methods}
+\subsection{Infrastructure} % TODO
+\subsection{Dataset} % TODO
 
 \subsection{Applications}
+\subsubsection{Increment}
+We adapted the increment pipeline used in \cite{hayot2019performance}.
+This synthetic application read blocks of the BigBrain from Lustre and sleeps for specified period.
+To simulate intermediate results, we repeat the sleep process for a configurable amount of time.
+We prevent data caching of the blocks by incrementing their voxels value by 1 after each sleep operation.
+Finally, we write the resulting NIfTI image back to Lustre.
+This application allow us the study engines when the inputs are processed independently.
+The map-only scenario of this application mimics the processing of multiple independt subjects in parralel.
+\MD{TODO: Add the Figure for the application graph.}
 
-We used three neuroimaging applications to evaluate the engines in different
-conditions. The first two, incrementation and histogram, are simple 
-synthetic applications representing map-only and map-reduce applications,
-respectively. The third, is a real neuroimaging application representative
-of popular BIDS applications~\cite{gorgolewski2017bids}. All scripts used for our
-experiments are available at
-\url{https://github.com/big-data-lab-team/paper-big-data-engines}
-
-%% Incrementation
-\subsubsection{Incrementation}
-We used an adaptation of the image incrementation pipeline used
-in~\cite{hayot2019performance} (see Algorithm~\ref{alg:incrementation}).
-The application reads blocks of the BigBrain image from the shared file
-system, increments the intensity value of each voxel by 1 to avoid caching
-effects, sleeps for a configurable amount of time to emulate more
-compute-intensive processing, repeats this process for a specified number of
-iterations, and
-finally writes the result as a NIfTI image back to the shared file system.
-This application allows us to study the behavior of the engines when all
-inputs are processed independently, in a map-only scenario (see
-Figure~\ref{fig:tg-inc}). This mimics the behavior of analyzing multiple
-independent subjects in parallel.
-
-\begin{algorithm}[!b]
-    \caption{Incrementation (adapted from~\cite{hayot2019performance})}\label{alg:incrementation}
-    \begin{algorithmic}
-    \Require{\(x\), a sleep delay in float}
-    \Require{\(file\), a file containing a BigBrain block}
-    \Require{\(fs\), NFS mount to write image to}
-    \State{Read \(block\) from \(file\)}
-    \ForEach{\(i \in iterations\)}
-        \ForEach{\(block \in image\)}
-            \State{\(block\gets block+1\)}
-            \State{Sleep \(x\)}
-        \EndFor
-    \EndFor
-    \State{Write \(block\) to \(fs\)}
-\end{algorithmic}
-\end{algorithm}
-
-\begin{figure}[!b]
-    \centering
-    \includegraphics[height=\columnwidth,
-    angle=-90]{images/incrementation-task-graph.png}
-    \caption{Task graph for Incrementation with 5 iterations and 3 BigBrain blocks.
-    Circles represent the incrementation and sleep function while rectangles
-    represent data elements.}\label{fig:tg-inc}
-\end{figure}
+\subsubsection{Multi-Increment}
+Our second application is a revision of the increment pipeline.
+A major difference is that, at each iteration, a random BigBrain blocks is used to as the increment value.
+This change allows the multi-increment application to have inter-worker communication while remaining simplistic.
+\MD{TODO: Add the Figure for the application graph.}
 
 \subsubsection{Histogram}
+Our third application calculatest the histogram of the BigBrain image.
+It reads the BigBrain blocks from Lustre, calculate frequency of each intensity, then write the aggregated result back on Lustre.
+This map-reduce application has a very high read over write ratio.
+Moreover, this application requires data shuffling; although little.
+This place it in between the increment and multi-increment pipeline in terms of inter-worker communication.
+\MD{TODO: Add the Figure for the application graph.}
 
- Our second application calculates the histogram of the BigBrain image (see
- Algorithm~\ref{alg:histogram}). It reads the image blocks from the shared
- file system, calculates intensity frequencies, aggregates the frequencies
- across the blocks, and finally writes the resulting histogram to the
- shared file system as a single \SI{766}{\kilo\byte} file. It follows the
- map-reduce paradigm, in which the final result is obtained from all the
- input blocks (see Fig.~\ref{fig:tg-histo}). This application requires data
- shuffling, incurring additional performance costs compared to
- Algorithm~\ref{alg:incrementation}. The total amount of shuffled data is,
- however, limited to \SI{2.62}{\mega\byte} per block as it only consists
- of image histograms. Two implementations are studied: a pure Python one,
- where frequencies are computed through Python dictionary manipulations, and one based
- on the \href{https://github.com/numpy}{NumPy} library, that implements array computations in C.
-
-\begin{algorithm}[!t]
-    \caption{Histogram}\label{alg:histogram}
-    \begin{algorithmic}
-    \Require{\(files\), files containing BigBrain blocks}
-    \Require{\(fs\), NFS mount to save image to}
-    \ForEach{\(file \in files\)}
-        \State{Read \(block\) from \(file\)}
-        \State{{Calculate \(frequencies\) in \(block\)}}
-    \EndFor
-
-    \State{\(histogram\gets\)Aggregate \(frequencies\) of each \(file\)}
-
-    \State{Write \(histogram\) to \(fs\)}
-    \end{algorithmic}
-\end{algorithm}
-
-\begin{figure}[!t]
-    \centering
-    \includegraphics[height=\columnwidth,
-    angle=-90]{images/histogram-task-graph.png}
-    \caption{Task graph for Histogram with 3 BigBrain blocks. Circles represent the
-    functions while rectangles represent data elements.
-    }\label{fig:tg-histo}
-\end{figure}
+\subsubsection{Kmeans} % TODO
+\MD{TODO: Write about it once implemented.}
 
 \subsubsection{BIDS app example}
+Our last application is BIDS App example: a neuroimaging pipeline to extract brain volume.
+For this application, the CoRR dataset was used.
+It extracts the brain volume of each participant, then computes the average for each group of participants.
+Unlike the other application, BIDS App example is a set of command-line distributed in a Docker image (bids/example on DockerHub).
+We converted the Docker image to a Singularity image for use in HPC environment, \MD{Cite paper on reason why this is done.}
+using \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
+\MD{Need to discuss if we preload the image on the worker nodes.}
+\MD{TODO: Add the Figure for the application graph.}
 
-We used the
-\href{https://github.com/BIDS-Apps/example}{BIDS app example} map-reduce style 
-application. The map phase, also called participant analysis, extracts the brain
-tissues of a subject's 3D MRI from the CoRR dataset using the FMRIB Software Library
-(\href{https://fsl.fmrib.ox.ac.uk/fsl/fslwiki}{FSL}), and writes the resulting
-image (\SI{2.47}{\mega\byte} on average) to the NFS. The reduce phase, or group
-analysis, computes the volume of each brain and returns the average
-volume, shuffling a total of \SI{8.6}{\giga\byte} image data.
+\subsection{Experiments} % TODO
 
-While Incrementation and Histogram were implemented directly in Python, the 
-BIDS App example requires an external set of command-line tools distributed as a Docker container image
-(bids/base\_fsl on DockerHub). We converted the Docker image to a Singularity image for use in 
-an HPC environment, using
-\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}.
-The image was preloaded on the NFS and Singularity version 3.2.1\-1.el7 was installed
-for all instances.
+\section{Results} % TODO
 
-\subsection{Experiments}
+\section{Discussion} % TODO
 
-Four parameters were varied in our experiments, as shown in
-Table~\ref{tab:param}. We varied (1) the number of workers, to assess the
-scalability of the engines, (2) the number of BigBrain blocks in
-Incrementation and Histogram, to measure the effect of different I/O patterns
-and parallelization degrees, and (3) the number of iterations and sleep
-delay in Incrementation, to evaluate the effect of job length and number of
-tasks.
-It should be noted that increasing the number of blocks or iterations also
-increases the total compute time of the application for a given sleep
-delay. To avoid any potential external bias such as background load on the
-network, we ran the experiments in a randomized order and cleared the page
-cache of each worker before each execution.
+\section{Conclusion} % TODO
 
-For each run, we measured the application makespan as well as the cumulative 
-data read, compute, data write, and engine overhead times across all application
-tasks. 
-The overhead calculation for each CPU is the end time of its last processed task
-minus the total runtime of the tasks it executed. The summation of those results is the
-total overhead.
-
-% The baseline for our experiments is: 8 workers, 125 blocks, 10 iterations and 4
-% seconds sleep delay.
-
-\begin{table}[!t]
-    \renewcommand{\arraystretch}{1.3}
-    \caption{Parameters for the experiments}\label{tab:param}
-    \centering
-    \begin{tabular*}{\columnwidth}{llll}
-    \hline
-                        & Incrementation & Histogram             & BIDS Example          \\ \hline
-    \# of worker        & 1, 2, 4, 8     & 1, 2, 4, 8            & 1, 2, 4, 8            \\
-    \# of blocks        & 30, 125, 750   & 30, 125, 750          & \multicolumn{1}{c}{-} \\
-    \# of iterations    & 1, 10, 100     & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\
-    Sleep delay {[}\SI{}{\second}{]} & 1, 4, 16, 64   & \multicolumn{1}{c}{-} & \multicolumn{1}{c}{-} \\ \hline
-    \end{tabular*}
-    \end{table}
-
-%%%%% RESULTS %%%%%
-\section{Results}
-
-%%% INCREMENTATION %%%
-%% worker
-\subsection{Incrementation: Number of workers}
-Figure~\ref{fig:inc_ms_worker} shows the makespan of the Incrementation application
-for different numbers of workers and engines. The bars show the average
-makespan over 3 repetitions while the error bars are the standard deviations. Overall,
-there is no substantial makespan difference between the engines. Dask seems to have a
-slight advantage over Spark, \SI{83.61}{\second} on average,
-with Delayed and Futures being slightly better than Bag.
-
-For all engines, the makespan is far from decreasing linearly with the
-number of workers. The makespan
-even increases between 4 and 8 workers. This is due to the high impact of
-data transfers and engine overhead on the application.
-
-Figure~\ref{fig:inc_tt_worker} shows the total execution time of the
-Incrementation application, broken down into data transfer (read and
-write), compute (sleep), and overhead time. As expected, the computing time
-stays similar when the number of workers increases. However, the data
-transfer time and overhead increase proportionally to the number of
-workers with regression slopes: Spark (\SI{2337}{\second/task}), Bag
-(\SI{2650}{\second/task}), Delayed (\SI{3251}{\second/task}), Futures
-(\SI{3570}{\second/task}).
-
-On Figure~\ref{fig:inc_tt_worker}, we also note that Spark's overhead is
-slightly lower than Dask's, particularly when the number of workers
-increases. Moreover, Delayed and Futures have a higher overhead than Bag.
-However, overhead differences are compensated by an increase in data
-transfer time, as a reduced overhead increases the concurrency between data
-transfers and the application saturates the disk bandwidth of the shared file system.
-
-\begin{figure}[!b]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/inc_worker.png}%
-        \caption{Incrementation makespan}\label{fig:inc_ms_worker}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/inc_idle_worker.png}%
-        \caption{Incrementation total time}\label{fig:inc_tt_worker}
-    \end{subfigure}
-    \caption{125 blocks, 10 iterations, \SI{4}{\second} delay, 8
-    CPUs/worker}\label{fig:inc_worker}
-\end{figure}
-
-% blocks
-\subsection{Incrementation: Number of blocks}
-
-Figure~\ref{fig:inc_ms_block} shows the Incrementation makespan when varying the
-number of image blocks for constant BigBrain image size. We were not able to run
-Spark for 30 blocks due to its \SI{2}{\giga\byte} limitation in the task size.
- Once again, we do not observe any substantial difference among the engines.
-For all engines, makespan variability increases with the number of blocks, however,
-engines scale very well in general.
-
-In Figure~\ref{fig:inc_tt_block}, the total execution time of each function is shown.
-For 30 blocks, the increased overhead of Delayed and Futures is a calculation artifact 
-coming from the fact that only 30 of the 64 available threads can be used concurrently. Once again, the data transfer time reduces with more blocks but 
-the overhead time increases by a similar amount. This is not observed for 30 blocks
-as the workers are not used at full capacity, i.e., some threads are idle. Finally,
-the variability of the overhead increases with the number of blocks, which explains
-the makespan variability mentioned previously.
-
-\begin{figure}[!t]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/inc_block.png}%
-        \caption{Incrementation makespan}\label{fig:inc_ms_block}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/inc_idle_block.png}%
-        \caption{Incrementation total time}\label{fig:inc_tt_block}
-    \end{subfigure}
-    \caption{10 iterations, \SI{4}{\second} delay, 8 workers, 8
-    CPUs/worker}\label{fig:inc_block}
-\end{figure}
-
-%% Iterations
-\subsection{Incrementation: Number of iterations}
-Figure~\ref{fig:inc_ms_itr} shows the makespan of the application while
-varying the number of iterations. Overall, Spark and Dask APIs are once
-again equivalent, although Delayed and Futures are slightly faster
-than Bag and Spark for 1 and 10 iterations, and Futures are faster than
-Delayed, Bag and Spark for 100 iterations. Differences remain minor
-though.
-
-In Figure~\ref{fig:inc_tt_itr}, the total execution time breakdown is
-shown. We observe the good scalability of all the engines with the number of
-iterations.
-
-\begin{figure}[!t]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/inc_itr.png}%
-        \caption{Incrementation makespan}\label{fig:inc_ms_itr}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/inc_idle_itr.png}%
-        \caption{Incrementation total time}\label{fig:inc_tt_itr}
-    \end{subfigure}
-    \caption{125 blocks, \SI{4}{\second} delay, 8 workers, 8 CPUs/worker}
-\end{figure}
-
-%% Sleep time
-\subsection{Incrementation: Sleep delay}
-Figure~\ref{fig:inc_ms_sleep} shows the makespan of the Incrementation application
-for different sleep delays. Overall, all engines again perform the same and scale
-well with task duration. Spark is initially slower than the Dask APIs, however, it is
-faster with an increased sleep delay. Also, within the Dask APIs, Dask Bag is
-slower than the other two, but it is not considerable.
-
-Figure~\ref{fig:inc_tt_sleep} shows the total execution time breakdown. Spark has the
-smallest overhead. As previously observed, variations in overhead time are almost
-exactly compensated by variations in data transfer time.
-
-\begin{figure}[!b]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/inc_sleep.png}%
-        \caption{Incrementation makespan}\label{fig:inc_ms_sleep}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/inc_idle_sleep.png}%
-        \caption{Incrementation total time}\label{fig:inc_tt_sleep}
-    \end{subfigure}
-    \caption{125 blocks, 10 iterations, 8 workers, 8 CPUs/worker}
-\end{figure}
-
-%% Baseline gantt chart
-\subsection{Incrementation: Gantt chart}
-
-Figure~\ref{fig:inc_gantt} shows the Gantt chart obtained for each engine. Gantt
-charts are structured in batches of up to 64 read-compute-write concurrent sequences.
-File reads in the first batch are much longer than the ones in the following batches:
-this is due to the high synchronization of data transfers that leads to a high
-saturation of the shared file system. We also note that overhead, represented in
-white, is concentrated around the data transfer tasks and the
-computing tasks that run concurrently with data transfers.
-
-\begin{figure}[!b]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/spark-inc-baseline.html}{
-            \includegraphics[clip,width=\columnwidth,
-            height=0.15\textheight]{images/spark_inc_gantt.png}}
-        \caption{Spark}\label{fig:inc_spark_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/dask-bag-inc-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_inc_gantt.png}}
-        \caption{Dask Bag}\label{fig:inc_dask_bag_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/dask-delayed-inc-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_inc_gantt.png}}
-        \caption{Dask Delayed}\label{fig:inc_dask_delayed_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/dask-futures-inc-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/futures_inc_gantt.png}}
-        \caption{Dask Futures}\label{fig:inc_dask_futures_gantt}
-    \end{subfigure}
-    \caption{Incrementation Gantt chart: 125 blocks, 10 iterations, \SI{4}{\second} delay, 8 workers,
-8 CPUs/worker. All Gantt charts are clickable and link to interactive figures with additional information.}\label{fig:inc_gantt}
-\end{figure}
-
-%%% HISTOGRAM %%%
-\subsection{Histogram (Python): Number of workers}
-Figure~\ref{fig:histo_ms_worker} shows the makespan of the Histogram application for various
-amounts of workers. Spark is significantly faster than Dask APIs. The difference
-narrows as the number of workers increases but it remains significant. Between Dask APIs
-there is no substantial difference, however, Bag is slightly faster on average.
-
-From Figure~\ref{fig:histo_tt_worker}, the total time spent in each function is
-shown. The computing time is significantly larger for Dask than for Spark. This is
-presumably due to Python's GIL preventing Dask to parallelize the computation
-over
-multiple threads. Overall, the I/O and overhead times are comparable for all engines.
-
-Figure~\ref{fig:histo_tt_worker} shows that Dask engines have a similar
-total execution time when their number of workers varies, except for 8
-workers where it increases slightly due to increased read time.
-On the other hand, Spark total execution time keeps increasing; especially at 8
-workers. This is because Dask engines benefit more from additional workers, in
-this GIL-bounded scenario.
-
-\begin{figure}[!b]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_worker.png}%
-        \caption{Histogram (Python) makespan}\label{fig:histo_ms_worker}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_idle_worker.png}%
-        \caption{Histogram (Python) total time}\label{fig:histo_tt_worker}
-    \end{subfigure}
-    \caption{125 blocks, 8 CPUs/worker}\label{fig:histo_worker}
-\end{figure}
-
-\subsection{Histogram (Python): Number of blocks}
-Figure~\ref{fig:histo_ms_block} shows the makespan for different block sizes. 
-Once more, Spark is significantly faster than Dask engines and Bag is slightly
-faster than Delayed. Overall, the engines do not react to changes in block size.
-
-In Figure~\ref{fig:histo_tt_block} the total time for each function is shown.
-Note that at 30 blocks the workers' resources are not fully used since the
-workers can process up to 64 tasks in parallel. Moreover, the overhead for
-30 blocks is erroneous due to a lower amount of block than available threads.
-This leads to thread based engines to have higher overhead. Given that, 
-Spark has a much lower compute time which makes it significantly faster
-than Dask engines. Overall, for all engines, lower block size results in lower
-data transfer time but increases the overhead time. Again, this is due to large overhead
-desynchronizing task thus reducing the NFS bottleneck.
-
-\begin{figure}[!b]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_block.png}%
-        \caption{Histogram (Python) makespan}\label{fig:histo_ms_block}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_idle_block.png}%
-        \caption{Histogram (Python) total time}\label{fig:histo_tt_block}
-    \end{subfigure}
-    \caption{8 workers, 8 CPUs/worker}\label{fig:histo_block}
-\end{figure}
-% \TG{We decided to omit Dask Futures as it does not have any value over Dask
-% Delayed in this application.}
-
-\subsection{Histogram (Python): Gantt chart}
-Figure~\ref{fig:histo_gantt} shows the Gantt chart of the baseline
-Histogram experiment implemented in Python. Spark overhead is distributed
-through all the execution while Bag overhead happens mostly when a thread
-reads subsequent blocks and Delayed overhead is concentrated after reading
-all the blocks.
-
-% % remove? - vhs
-% The reduce is done differently for the engines. Spark and Delayed compute
-% all the reduce tasks on one process (thread), however, Spark only computes it
-% on the driver process. On the other hand, Bag computes the reduce tasks on 
-% different workers and sends the information to the driver process afterward.
-
-\begin{figure}[!b]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/spark-histo-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_histo_gantt.png}}
-        \caption{Spark}\label{fig:histo_spark_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/dask-bag-histo-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_histo_gantt.png}}
-        \caption{Dask Bag}\label{fig:histo_dask_bag_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/dask-delayed-histo-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_histo_gantt.png}}
-        \caption{Dask Delayed}\label{fig:histo_dask_delayed_gantt}
-    \end{subfigure}
-    \caption{Histogram (Python) Gantt chart: 125 blocks, 8 workers, 8
-    CPUs/worker}\label{fig:histo_gantt}
-\end{figure}
-
-%%% HISTOGRAM (NumPy) %%%
-\subsection{Histogram (NumPy): Number of workers}
-Figure~\ref{fig:histo_np_ms_worker} shows the makespan for various amount of workers.
-Overall, the makespan for all engines does not change considerably with an increase
-in the number of workers. This is likely a result of using a non-distributed
-shared file system. There is no
-substantial difference between all engines although Delayed is slightly faster.
-Figure~\ref{fig:histo_np_tt_worker} shows the total execution time
-breakdown. The overhead and read time increase proportionally to the number
-of workers thus increasing the total time.
-
-\begin{figure}[!t]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_np_worker.png}%
-        \caption{Histogram (NumPy) makespan}\label{fig:histo_np_ms_worker}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_idle_np_worker.png}%
-        \caption{Histogram (NumPy) total time}\label{fig:histo_np_tt_worker}
-    \end{subfigure}
-    \caption{125 blocks, 8 CPUs/worker}\label{fig:histo_np_worker}
-\end{figure}
-
-\subsection{Histogram (NumPy): Number of blocks}
-Figure~\ref{fig:histo_np_ms_block} shows the makespan of the application for
-different block sizes. Overall, there is no substantial difference for
-all engines and block sizes.
-
-Figure~\ref{fig:histo_np_tt_block} shows the total time spent in each function.
-Once more, since Delayed is thread-based its overhead at 30 blocks is much
-higher. Considering the overhead error and that at 30 blocks
-the workers only use half their core, the total time is substantially the same
-for all block size. This is due to lower block sizes having a lower data transfer time,
-however, it balances out by the resulting increase in overhead.
-
-\begin{figure}[!t]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_np_block.png}%
-        \caption{Histogram (NumPy) makespan}\label{fig:histo_np_ms_block}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/histo_idle_np_block.png}%
-        \caption{Histogram (NumPy) total time}\label{fig:histo_np_tt_block}
-    \end{subfigure}
-    \caption{8 workers, 8 CPUs/worker}\label{fig:histo_np_block}
-\end{figure}
-
-\subsection{Histogram (NumPy): Gantt chart}
-Figure~\ref{fig:histo_np_gantt} shows the Gantt chart of the baseline for the
-Histogram (NumPy) experiment. The overhead for Spark and Bag is mostly located
-between read and compute tasks. On the other hand, Delayed
-overhead is dispersed between all types of tasks. This causes Delayed to
-overlap more compute and read tasks simultaneously amongst different workers.
-
-In Figure~\ref{fig:histo_np_gantt}, read tasks are initially similar for all
-engines. However, when computation starts the subsequent read tasks are much
-shorter for Delayed. This is due to the desynchronization of the task reducing the
-data transfer bottleneck.
-
-\begin{figure}[!tb]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/spark-histo_np-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_histo_np_gantt.png}}
-        \caption{Spark}\label{fig:histo_np_spark_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/dask-bag-histo_np-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_histo_np_gantt.png}}
-        \caption{Dask Bag}\label{fig:histo_np_dask_bag_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/dask-delayed-histo_np-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_histo_np_gantt.png}}
-        \caption{Dask Delayed}\label{fig:histo_np_dask_delayed_gantt}
-    \end{subfigure}
-    \caption{Histogram (NumPy) Gantt chart: 125 blocks, 8 workers, 8
-    CPUs/worker}\label{fig:histo_np_gantt}
-\end{figure}
-
-%%% BIDS EXAMPLE %%%
-\subsection{BidsApp example: Number of workers}
-Figure~\ref{fig:bids_ms_worker} shows the makespan of the application when
-varying the number of workers. Overall, there is no substantial difference
-between the engines. The makespan scales well with the increase in workers,
-however, the scaling reduces slightly when reaching 8 workers. This is due
-to an increase in I/O time caused by parallel accesses to 
-the shared file system.
-
-In Figure~\ref{fig:bids_tt_worker}, the total execution time of each function is
-shown. Each engine has a similar total time independently of its number of workers.
-This is due to the, almost, linear scaling of the engines. The overhead increases
-proportionally to the number of workers with regression slopes: Spark
-(\SI{224}{\second/task}), Bag (\SI{183}{\second/task}), Delayed
-(\SI{173}{\second/task}), Futures (\SI{132}{\second/task}). This is due to more
-communication between the scheduler and workers as well as intra-worker and
-inter-worker communication.
-
-\begin{figure}[!t]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/bids_worker.png}%
-        \caption{BIDS App example makespan}\label{fig:bids_ms_worker}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \includegraphics[clip,width=\columnwidth]{images/bids_idle_worker.png}%
-        \caption{BIDS App example total time}\label{fig:bids_tt_worker}
-    \end{subfigure}
-    \caption{Variation of the amount of worker, 8 CPUs/worker}
-\end{figure}
-
-%% Baseline gantt chart
-\subsection{BidsApp example: Gantt chart}
-Figure~\ref{fig:bids_gantt} shows the Gantt chart obtained for each engine.
-Gantt charts are structured in two parts: participant analysis (orange) and
-group analysis (blue). The participant analysis tasks differ greatly in length. This
-is due to the unequal amount of sessions per subject processed. The group analysis
-is similar for all engines and APIs. Overall, most of the overhead encountered
-results from the transition between the two analysis. This is because the group
-analysis requires the results of the participant analysis to start.
-
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/spark-bids-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/spark_bids_gantt.png}}
-        \caption{Spark}\label{fig:bids_spark_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/dask-bag-bids-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/bag_bids_gantt.png}}
-        \caption{Dask Bag}\label{fig:bids_dask_bag_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/dask-delayed-bids-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/delayed_bids_gantt.png}}
-        \caption{Dask Delayed}\label{fig:bids_dask_delayed_gantt}
-    \end{subfigure}
-    \\
-    \begin{subfigure}[b]{\columnwidth}
-        \href{https://mathdugre.github.io/paper-big-data-engines/dask-futures-bids-baseline.html}{
-        \includegraphics[clip,width=\columnwidth,
-        height=0.15\textheight]{images/futures_bids_gantt.png}}
-        \caption{Dask Futures}\label{fig:bids_dask_futures_gantt}
-    \end{subfigure}
-    \caption{BIDS App example Gantt chart: 8 workers}\label{fig:bids_gantt}
-\end{figure}
-
-%%%%% DISCUSSION %%%%%
-\section{Discussion}
-
-\subsection{Comparison with previous studies}
-
-Overall, our experiments did not show any substantial performance
-difference between Spark and Dask. In contradiction with the experiments
-of~\cite{Mehta:17}, we did not observe consequential overhead differences
-due to serialization or startup costs. This apparent contradiction is most
-likely coming from differences in applications, engine parametrization, or
-infrastructure characteristics between the two studies. In our case, potential overhead differences were
-compensated by reverse trends in data transfer time. Since our focus is
-Big Data applications, this leads us to conclude that engines are
-equivalent.
-
-\subsection{Overhead / data transfer compensation}
-
-The observed almost exact compensation between differences in engine
-overheads and in data transfer times (see for instance
-Figure~\ref{fig:inc_tt_worker}) is not surprising. Indeed, when the transfer
-bandwidth is saturated, and in the absence of service thrashing,
-desynchronizing file transfers usually does not reduce the makespan of data
-transfers. To take an extreme example, the makespan of $n$ concurrent file
-transfers of size $F$ on a system of bandwidth $B$ would be $nF/B$, as $n$
-transfers would share the bandwidth, which equals to the time required
-to transfer the $n$ files sequentially. In other words, saturated bandwidths
-give room for extra engine overhead as long as it desynchronizes data
-transfers. This could be an interesting idea to explore in data transfer optimization.
-
-\subsection{Effect of the shared file system}
-
-We used NFS as our shared file system as it is a common and
-easy-to-configure solution to share data between compute nodes. The
-performance of NFS, however, degrades substantially when accessed by
-concurrent tasks, as shown in the Gantt charts in
-Figures~\ref{fig:inc_gantt}, ~\ref{fig:histo_gantt}, ~\ref{fig:histo_np_gantt} 
-and~\ref{fig:bids_gantt}. This is due to the limitations imposed by network and disk
-bandwidths. While network bandwidth can hardly be overcome, disk bandwidth
-can be increased by storing data to multiple disks addressed by a parallel
-file system such a Lustre~\cite{lustre}. Such a configuration would likely
-reduce the impact of data transfers, although the main observed patterns
-should remain.
-
-Throughout the applications, the use of large data blocks amplified the consequences
-of using an NFS. Moreover, the bandwidth saturation almost neglects the scaling, from
-additional workers, for data-intensive applications: Incrementation and Histogram
-(NumPy). On the other hand, an almost linear scaling shows that it has no substantial
-effect on compute-intensive applications: Histogram (Python) and BIDS app example.
-
-We suppose that using a parallel file system would allow applications to have almost
-linear scaling as long as they do not saturate the bandwidth of the system.
-
-\subsection{Effect of Python's Global Interpreter Lock}
-
-Although Dask reduces the impact of Python's GIL on parallelism by using external
-Python extensions, our Python histogram experiment (Figure
-~\ref{fig:histo_worker} and \ref{fig:histo_block}) showed that issues remain in Dask to
-parallelize Python functions. However, Dask was easily able to parallelize
-the same application when implemented with NumPy
-(Figure~\ref{fig:histo_np_worker} and \ref{fig:histo_np_block}) because in this case the computation
-happened in C, outside of the GIL. Similarly, the computation in
-command-line applications happens in a sub-process, hence outside of the
-GIL, which explains why the BIDS apps example application was not impacted
-by this issue. Although implemented in Python, Incrementation was not
-impacted either, as computation was only emulated through sleep time.
-
-This behavior is also the consequence of our configuration of Dask that
-created 1 worker process with 8 threads. Since the Histogram Python
-implementation does not release the GIL it can only use one of the thread
-at a given time thus it slows down the computation significantly. This
-behavior highlights the need to configure the workers properly depending on
-the type of functions executed in an application.
-
-\subsection{Effect of data partitioning}
-
-The number of data partitions also has an impact on application
-performance. In our experiments, we set a partition for each block or
-subject, resulting in a maximum of 750 partitions for the BigBrain
-applications, and 1,397 partitions for the BIDS app. The number of
-partitions obeys a classical trade-off: small numbers of partitions reduce 
-overhead, while large numbers of
-partitions increase parallelism. Since both Spark and Dask can
-work with custom partition sizes, we do not expect substantial differences
-coming from variations of this parameter.
-
-% Varying the block size is equivalent to varying the size of the partitions. We
-% observed no difference when varying the task size while using an NFS. However, if
-% one uses a parallel shared file system, we suppose it would be preferable to use a
-% larger block size as the IO time should be reduced substantially. It is worth noting
-% that the Dask developers suggest, in
-% \href{https://docs.dask.org/en/latest/best-practices.html}{Dask's best practices}, to
-% use a partition size such that workers have 2-3 times as many chunks available to
-% work on. More experiments would have to be performed to gain insight on an optimal
-% number of partitions to use for a given set of data-intensive applications.
-
-\section{Conclusion}
-We presented a comparison of two Big Data engines, Apache Spark and Dask.
-We studied the engines on three data-intensive neuroimaging applications
-representative of common use cases. Overall, our results show no substantial
-performance difference between the engines. Interestingly, differences in
-engine overheads do not impact performance due to their impact on data
-transfer time: higher overheads are almost exactly compensated by a lower
-transfer time when data transfers saturate the bandwidth. These results
-suggest that future research should focus on strategies to reduce the
-impact of data transfers on applications.
-
-\section*{Acknowledgment}
-
-Mathieu Dugr\'e was funded by an Undergraduate Student Research Assistant award from
-the National Science and Engineering Research Council of Canada. We warmly thank
-Compute Canada and its regional center Westgrid for providing the cloud infrastructure used in
-these experiments, and the McGill Center for Integrative Neuroscience for giving us
-access to their cloud allocation.
+\section*{Acknowledgment} % TODO
 

### 2021-01-16 src/app/commons/kmeans.py
Implement kmeans in dask.

@@ -1,0 +1,135 @@
+import os
+import time
+
+import nibabel as nib
+import numpy as np
+import numexpr as ne
+
+from ..utils import log
+
+
+def eucledian_distance(arr: np.array, c: float):
+    """Returns the eucledian distance between x and y.
+
+    Parameters
+    ----------
+    arr : float
+        1-D array of voxel
+    c : float
+        centroid
+
+    Returns
+    -------
+        The Eucledian distance between x and y.
+    """
+    # TODO benchmark
+    return np.linalg.norm(arr[:, None] - c, axis=-1)
+
+
+def closest_centroids(x, centroids):
+    """Returns the index of the closest centroids.
+
+    Parameters
+    ----------
+    x : np.array
+        1-D array of voxels.
+    centroids : list(T)
+        List of current centroids.
+
+    Returns
+    -------
+    closest_centroid : T
+    """
+    # TODO benchmark
+
+    dist = np.matrix([eucledian_distance(x, c) for c in centroids])
+    return np.squeeze(np.asarray(dist.T.argmin(1)))
+
+
+def classify_block(block, centroids):
+    filename = block[0]
+    img = block[1]
+    metadata = block[2]
+
+    img = np.array([np.absolute(img - centroid) for centroid in centroids]).argmin(0)
+    # TODO benchmark
+
+    return filename, img, metadata
+
+
+def add_component_wise(total, x):
+    """Binary operator to reduce
+
+    Parameters
+    ----------
+    total : (float, int)
+        Total value of the combination and number of element combined to obtain it.
+    x : int, (float, int)
+        Total value of the combination and number of element combined to obtain it.
+    """
+    return total[0] + x[0], (total[1][0] + x[1][0], total[1][1] + x[1][1])
+
+
+def get_voxels(voxels):
+    """Return the list of voxels for the NiftiImage.
+
+    Parameters
+    ----------
+    voxels : 2D np.array
+        Voxels from the nifti image.
+
+    Returns
+    -------
+    np.array
+        The list of voxels for the NiftiImage.
+    """
+
+    return voxels.flatten("F")
+
+
+def dump(img_rdd, *, benchmark, start, output_folder, experiment):
+    """Save a Nifti image.
+
+    Parameters
+    ----------
+    img_rdd: (str, np.array, (np.array, np.array))
+        Filename, image, and image header and affine.
+    assignment: (float, (int, int))
+        Voxel's class, intensity and frequency.
+    start : float
[…]

### 2021-01-16 src/app/dask/kmeans.py

@@ -1,0 +1,102 @@
+import glob
+import time
+
+import dask
+import dask.array as da
+import dask.bag as db
+from dask.distributed import Client
+import nibabel as nib
+import numpy as np
+
+from ..commons.kmeans import classify_block, closest_centroids, dump
+from ..utils import load, log, merge_logs
+
+
+def run(
+    input_folder: str,
+    output_folder: str,
+    scheduler: str,
+    benchmark: bool,
+    *,
+    iterations,
+) -> None:
+    experiment = f"dask:kmeans:iterations={iterations}"
+    start_time = time.time()
+    common_args = {
+        "benchmark": benchmark,
+        "start": start_time,
+        "output_folder": output_folder,
+        "experiment": experiment,
+    }
+
+    client = Client(scheduler)
+
+    blocks = [
+        dask.delayed(load)(
+            filename,
+            **common_args,
+        )
+        for filename in glob.glob(input_folder + "/*.nii")
+    ]
+
+    sample = blocks[0].compute()[1]  # Compute a sample block to obtain dtype and shape.
+    voxels = (
+        da.concatenate(
+            [
+                da.from_delayed(block[1], dtype=sample.dtype, shape=sample.shape)
+                for block in blocks
+            ]
+        )
+        .reshape(-1)
+        .rechunk(16_000_000, balance=True)
+    )
+
+    # Pick random initial centroids
+    # TODO benchmark
+    centroids = da.random.choice(
+        da.unique(voxels).compute(), size=3, replace=False
+    ).compute()
+
+    for _ in range(0, iterations):  # Disregard convergence.
+        start = time.time() - start_time
+
+        centroid_index = voxels.map_blocks(
+            lambda block: closest_centroids(block, centroids)
+        )
+
+        # Find centroid (total, count) => total/count = centroid
+        centroids = np.array(
+            [da.mean(voxels[centroid_index == i]) for i in range(len(centroids))]
+        )
+        print(f"{centroids=}")
+
+        end_time = time.time() - start
+
+        if benchmark:
+            log(
+                start,
+                end_time,
+                "all_file",
+                output_folder,
+                experiment,
+                "update_centroids",
+            )
+
+    results = []
+    for block in blocks:
+        block = dask.delayed(classify_block)(block, centroids)
+        results.append(
+            dask.delayed(dump)(
+                block,
+                **common_args,
+            )
+        )
+
+    futures = client.compute(results)
+    client.gather(futures)
+
[…]

### 2021-01-21 latex/paper.tex
[REVIEW] Methods section

@@
 \begin{IEEEkeywords}
-	Performance, Scheduler, Big Data, Dask, Spark, Ray, Neuroimaging
+	Performance, Scheduler, Big Data, Dask, Spark, Neuroimaging
 \end{IEEEkeywords}
@@
 
-\subsection{Ray}
-
 \subsection{Apache Spark}
@@
 We adapted the increment pipeline used in \cite{hayot2019performance}.
-This synthetic application read blocks of the BigBrain from Lustre and sleeps for specified period.
+This synthetic application reads blocks of the BigBrain from Lustre and simulates computation by sleeping for a specified period.
 To simulate intermediate results, we repeat the sleep process for a configurable amount of time.
-We prevent data caching of the blocks by incrementing their voxels value by 1 after each sleep operation.
+We prevent data caching of the blocks by incrementing their voxels value by one after each sleep operation.
 Finally, we write the resulting NIfTI image back to Lustre.
-This application allow us the study engines when the inputs are processed independently.
-The map-only scenario of this application mimics the processing of multiple independt subjects in parralel.
+This application allows us to study the engines when their inputs are processed independently.
+The map-only scenario of this application mimics the processing of multiple independent subjects in parallel.
 \MD{TODO: Add the Figure for the application graph.}
@@
 Our second application is a revision of the increment pipeline.
-A major difference is that, at each iteration, a random BigBrain blocks is used to as the increment value.
+A significant difference is that, at each iteration, it uses a random BigBrain block as the increment value. 
 This change allows the multi-increment application to have inter-worker communication while remaining simplistic.
@@
 \subsubsection{Histogram}
-Our third application calculatest the histogram of the BigBrain image.
-It reads the BigBrain blocks from Lustre, calculate frequency of each intensity, then write the aggregated result back on Lustre.
-This map-reduce application has a very high read over write ratio.
-Moreover, this application requires data shuffling; although little.
-This place it in between the increment and multi-increment pipeline in terms of inter-worker communication.
+As our third application, we calculate the histogram of the BigBrain image. 
+It reads the BigBrain blocks from Lustre, calculates each intensity's frequency, and then writes the aggregated result back on Lustre.
+This map-reduce application has a very high read overwrite ratio.
+Moreover, this application requires data shuffling, although little. 
+The amount of inter-worker communication is in-between the increment and multi-increment pipelines.
 \MD{TODO: Add the Figure for the application graph.}
@@
 \subsubsection{Kmeans} % TODO
-\MD{TODO: Write about it once implemented.}
+For our fourth application, we use Kmeans to cluster the BigBrain image by voxel intensity.
+We set the number of clusters to 3 to segment the white and grey matter and the noise.
+The application starts by reading the image blocks, combining all voxels in a 1-D array, and choosing initial centroids using the min, max, and intermediate values.
+It assigns each voxel to the centroids it is the closest and updates each centroid by computing the average of the voxels associated with it. 
+It repeats the assignment and update steps for a configurable amount of time. 
+Finally, the voxels of the image blocks are classified and write back on the file system.
+Updating the centroids involves a lot of data communication between the worker.
+
+Spark and Dask's implementation differs slightly for this application to take advantage of the" most suited" API from both engines.
+Spark implementation uses the Map-Reduce paradigm, while Dask implementation uses array programming.
+\MD{TODO: Add the Figure for the application graph.}
 
@@
 Our last application is BIDS App example: a neuroimaging pipeline to extract brain volume.
-For this application, the CoRR dataset was used.
+For this application, we use the CoRR dataset.
 It extracts the brain volume of each participant, then computes the average for each group of participants.
-Unlike the other application, BIDS App example is a set of command-line distributed in a Docker image (bids/example on DockerHub).
-We converted the Docker image to a Singularity image for use in HPC environment, \MD{Cite paper on reason why this is done.}
+Unlike the other application, BIDS App example is a command-line set in a Docker image (bids/example on DockerHub).
+We converted the Docker image to a Singularity image for use in HPC environments, \MD{Cite paper on reason why this is done.}
 using \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}

### 2021-01-21 latex/paper.tex
[ADD] link for dask and citations.

@@
 
-Dask offers five data structures: Array, Bag, DataFrame, Delayed, and Futures. \MD{Insert link to the API.}
+Dask offers five data structures:
+\href{https://docs.dask.org/en/latest/array.html}{Array},
+\href{https://docs.dask.org/en/latest/bag.html}{Bag},
+\href{https://docs.dask.org/en/latest/dataframe.html}{DataFrame},
+\href{https://docs.dask.org/en/latest/delayed.html}{Delayed},
+and \href{https://docs.dask.org/en/latest/futures.html}{Futures}.
 The Dask Array offers a clone of NumPy API for distributed processing of large arrays.
-A Dask Bag is a distributed collection of Python object that offers a programming abstraction similar to PyToolz. \MD{Link to PyToolz.}
+A Dask Bag is a distributed collection of Python object that offers a programming abstraction similar to \href{https://toolz.readthedocs.io/en/latest/}{PyToolz}.
 Dask Dataframe is a parallel composition of Pandas Dataframes used to process large amount of tabular data.
@@
 This is particulary true in the neuroimaging field with numerous specialized Python libraries to operate on data.
-While serialization from Python to Java create overhead, we found it to be minimal. \MD{cite SC19 paper.}
+While serialization from Python to Java create overhead, we found it to be minimal \cite{8943502}.
 We chose to focus on PySpark API to have a more balance environment between the different engines and due to its suitability to neuroimaging.

### 2021-01-21 latex/paper.tex
First write for the datasets section

@@
 \subsection{Infrastructure} % TODO
-\subsection{Dataset} % TODO
+
+\subsection{Dataset}
+We used BigBrain\cite{Amunts:13}, a 3-D image of the human brain with voxel intensity ranging between 0 and 65,535.
+We converted the blocks into the NIfTI format, a popular format in neuroimaging.
+We left the NIfTI blocks uncompressed, resulting in a total data size of \SI{00}{\giga\byte}\MD{Verify the size of the dataset when Lustre back on}.
+To evaluate the effect of block size, we resplit these blocks into 30, 125 and 750 blocks of \MD{size for each block segment}, using the \href{https://github.com/big-data-lab-team/sam}{sam} library.
+
+We also used the dataset provided by the Consortium for Reliability and Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) \cite{zuo2014open}.
+This dataset is freely available on \href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{datalad}.
+The entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and functional images of 1,397 subjects acquiredin 29 sites.
+We used all 3,491 anatomical images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} per image on average).
 

### 2021-01-21 latex/paper.tex
Drafted the infrastructure part (partial)

@@
 \section{Methods}
-\subsection{Infrastructure} % TODO
+\subsection{Infrastructure}
+For our experiments we used the SlashBin cluster at Concordia University.
+Each compute node has 2 16-cores CPUs,
+\SI{256}{\giga\byte} memory, \MD{find out what's the memory speed and \#-channel supported.}
+\SI{2.88}{\tera\byte} SSDs of mounted storage.
+The nodes are interconnected with a dedicated \SI{10}{\giga\bit/\second} Ethernet connection.
+Centos \MD{Verify version} and Linux kernel \MD{Verify kernel version} is installed on the nodes.
+
+\MD{Write about the worker/leader setup once the dataset size is determined. Need to tune so that 2-3 blocks can fit in the excutors memmory}
 

### 2021-01-21 latex/paper.tex
Finished first writeo of the engine section

@@
 Dask is a Python-based Big Data engine with growing popularity in the scientific Python ecosystem.
-To mitigate the data transfer bottleneck in Big Data workflows, Dask was designed with data locality and in-memory in mind.
+Dask was designed with data locality and in-memory in mind to mitigate the data transfer bottleneck in Big Data workflows.
 Data locality, popularized by Map-Reduce \cite{dean2008mapreduce}, schedules tasks where the data reside.
-In-memory computing minimize the overhead of transfering data to disk by keeping data in memory when possible.
-Dask uses lazy evalutation to reduce unnecessary communication and computation.
-The engine builds a dynamic graph prior to execution allowing to determine which task to compute.
-Dask workflows can further reduce data transfer by using leveraging mulithreading whenever the application is not bounded by Python's GIL.
-Fault-tolerance is achieved by recording of data lineage, the sequence of operations used on to modify the initial data.
+In-memory computing minimizes the overhead of transferring data to disk by keeping data in memory when possible.
+Dask uses lazy evaluation to reduce unnecessary communication and computation.
+The engine builds a dynamic graph before execution, allowing it to determine which task to compute.
+Dask workflows can further reduce data transfer by leveraging multithreading whenever Python's GIL does not restrict it.
+Fault-tolerance is achieved by recording data lineage: the sequence of operations used to modify the initial data.
 
@@
 and \href{https://docs.dask.org/en/latest/futures.html}{Futures}.
-The Dask Array offers a clone of NumPy API for distributed processing of large arrays.
-A Dask Bag is a distributed collection of Python object that offers a programming abstraction similar to \href{https://toolz.readthedocs.io/en/latest/}{PyToolz}.
-Dask Dataframe is a parallel composition of Pandas Dataframes used to process large amount of tabular data.
-Dask Delayed offers an API for distributing abrirtary function that do not fit in the above frameworks.
-Lastly, Dask Futures can also execute abrirtary functions, however, it launch computation immediately rather than lazily.
-Dask modularity allow users to install only required components making it lightweigth.
+Arrays offer a clone of NumPy API for distributed processing of large arrays.
+Bags are a distributed collection of Python object that offers a programming abstraction similar to \href{https://toolz.readthedocs.io/en/latest/}{PyToolz}.
+Dataframes are a parallel composition of \href{https://pandas.pydata.org/}{Pandas} Dataframes used to process a large amount of tabular data.
+Dask Delayed offers an API for distributing arbitrary functions that do not fit in the above frameworks.
+Lastly, Dask Futures can also execute arbitrary functions; however, it launches computation immediately rather than lazily.
+Dask modularity allows users to install only required components making it lightweight.
 
-In Dask a scheduler decides where and when to execute tasks using the Dask graph.
-API operations generate multiple fine-coarsed tasks in the computation graph allwing an easier representatin of complex algorithms.
+In Dask, a scheduler decides where and when to execute tasks using the Dask graph.
+API operations generate multiple fine-coarse tasks in the computation graph, allowing a more straightforward representation of complex algorithms.
 
-The Dask engine is compatible to numerous distributed schedulers, including YARN and Mesos.
-Dask also provide its own \textit{Dask Distributed scheduler}.
-Althought Dask is compatible with schedulers available in HPC clusters,
-We chose to use Dask Distributed scheduler, to keep the environment balanced between the engines.
+The Dask engine is compatible with multiple distributed schedulers, including YARN and Mesos.
+Dask also provides its own \textit{Dask Distributed scheduler}.
+We chose to use Dask Distributed scheduler to keep the environment balanced between the engines.
 
-In Dask Distributed scheduler, a \textit{dask-scheduler} process admistrates the resource provided by \textit{dask-worker} in the cluster.
+In the Dask Distributed scheduler, a \textit{dask-scheduler} process administrates the resource provided by  \textit{dask-worker}s in the cluster.
 The scheduler receives jobs from clients and assigns tasks to available workers.
-Dask scheduler use a FIFO (First-In-First-Out) job scheduling policy.
-That is, completely process a branch of the Dask graph before proceding with the next one.
+Task scheduler uses a FIFO (First-In-First-Out) job scheduling policy.
+That is an utter process branch of the Dask graph before proceeding with the next one.
 
-Dask offers multiple ways to deploy a cluster including: with SSH configs, Kubernetes, SLURM, PBS, and others.
+Dask offers multiple ways to deploy a cluster, including, but not limited to, SSH configs, Kubernetes, SLURM, PBS.
 For our experiments, we used the \href{https://jobqueue.dask.org/en/latest/generated/dask_jobqueue.SLURMCluster.html}{Dask SLURM cluster} API.
@@
 \subsection{Apache Spark}
-Apache Spark is a widely-used general pupose Big Data engine.
-Like Dask, it aims at reducing data transfer cost by incorporating data locality and in-memory computing, and lazy evaluation.
+Apache Spark is a widely-used general-purpose Big Data engine.
+Like Dask, it aims at reducing data transfer costs by incorporating data locality, in-memory computing, and lazy evaluation.
 
-Spark primary data structure is Resilient Distributed Dataset (RDD)\cite{RDD}, a fault-tolerant, parallel collection of data elements.
-RDDs is the basis of the other Spark data structure: Datasets and DataFrames.
-Datasets are similar to RDD but benefits from additional performance by leveraging the Spark SQL's optimized execution engine.
-The DataFrame API are Datasets organized into named-columns and used to process tabular data.
-While the DataFrame API is available in all supported language, DAtasets are limited to Sacla and Java.
+Spark offers three options to schedule jobs: Spark Standalone, Mesos, and YARN.
+The Spark standalone is a simple built-in scheduler.
+YARN is mainly used to schedule Hadoop-based workflows, while Mesos can be used for various workflows.
+We limit our focus to Spark Standalone scheduler, as researchers are likely to execute their workflows in an HPC environment where, usually, neither YARN nor Mesos is available.
 
-Python is a common programming language in the scientific community.
-This is particulary true in the neuroimaging field with numerous specialized Python libraries to operate on data.
-While serialization from Python to Java create overhead, we found it to be minimal \cite{8943502}.
-We chose to focus on PySpark API to have a more balance environment between the different engines and due to its suitability to neuroimaging.
+In the Spark Standalone scheduler, a \textit{leader} coordinates the resource provisioned by \textit{workers} in the cluster.
+A \textit{driver}  process receives jobs from clients and requests workers from the leader.
+Jobs are divided into stages to be executed onto workers.
+Each operation in a stage is represented by a high-level task in the computation graph.
+Like Dask, Spark Standalone scheduler uses a FIFO policy to schedule tasks.
+Spark Standalone has two execution modes: (1) the client mode, where the driver process runs in a dedicated process,
+and (2) the cluster mode, where the driver runs within a worker process.
+Our experiments use the client mode since cluster mode is not available in PySpark.
+
+Spark's primary data structure is Resilient Distributed Dataset (RDD)\cite{RDD}, a fault-tolerant, parallel collection of data elements.
+RDDs are the basis of the other Spark data structure: Datasets and DataFrames.
+Datasets are similar to RDD but benefit additional performance by leveraging the Spark SQL's optimized execution engine. 
+The DataFrames are Datasets organized into named-columns and are used to process tabular data. 
+While the DataFrame API is available in all supported languages, Datasets are limited to Scala and Java.
+
+Python is a standard programming language in the scientific community.
+This is particularly true in the neuroimaging field, with numerous specialized Python libraries to operate on data.
+While serialization from Python to Java creates overhead, we found it to be minimal \cite{8943502}.
+We focus on PySpark API to have a more balanced environment between the different engines and for its suitability to neuroimaging.
 

### 2021-01-22 latex/paper.tex
comments and edits

@@
 
-\title{Impact of engines scheduler on the performance of Big Data pipelines in HPC}
+\title{Impact of engines scheduler on the performance of Big Data pipelines in HPC \TG{Suggested title: Performance comparison of Dask and Apache Spark on HPC systems}}
 
@@
 Dask is a Python-based Big Data engine with growing popularity in the scientific Python ecosystem.
-Dask was designed with data locality and in-memory in mind to mitigate the data transfer bottleneck in Big Data workflows.
+Dask was designed with data locality and in-memory computing in mind, to mitigate the data transfer bottleneck in Big Data workflows.
 Data locality, popularized by Map-Reduce \cite{dean2008mapreduce}, schedules tasks where the data reside.
@@
 
-Dask \MD{Add version when after locking it for experiments.} was used.
+Dask \MD{Add version when after locking it for experiments.} was used. \TG{should be in methods}
 
@@
 Spark offers three options to schedule jobs: Spark Standalone, Mesos, and YARN.
-The Spark standalone is a simple built-in scheduler.
+Spark Standalone is a simple built-in scheduler.
 YARN is mainly used to schedule Hadoop-based workflows, while Mesos can be used for various workflows.
@@
 
-In the Spark Standalone scheduler, a \textit{leader} coordinates the resource provisioned by \textit{workers} in the cluster.
+In the Spark Standalone scheduler, a \textit{leader} \TG{I am all in favor of inclusive terminology, but if Spark didn't update their wording we shouldn't do it for them} coordinates the resource provisioned by \textit{workers} in the cluster.
 A \textit{driver}  process receives jobs from clients and requests workers from the leader.
@@
 
-Python is a standard programming language in the scientific community.
-This is particularly true in the neuroimaging field, with numerous specialized Python libraries to operate on data.
-While serialization from Python to Java creates overhead, we found it to be minimal \cite{8943502}.
+Python is a standard programming language in the scientific community, offering numerous data processing libraries.
+While serialization from Python to Java, an operation required when using Spark's Python API, creates overhead, we found it minimal \cite{8943502}.
 We focus on PySpark API to have a more balanced environment between the different engines and for its suitability to neuroimaging.
@@
 \subsubsection{Increment}
-We adapted the increment pipeline used in \cite{hayot2019performance}.
+We adapted the increment application used in \cite{hayot2019performance}.
 This synthetic application reads blocks of the BigBrain from Lustre and simulates computation by sleeping for a specified period.
@@
 \subsubsection{Multi-Increment}
-Our second application is a revision of the increment pipeline.
+Our second application is an adaptation of the increment application.
 A significant difference is that, at each iteration, it uses a random BigBrain block as the increment value. 
-This change allows the multi-increment application to have inter-worker communication while remaining simplistic.
+This change allows the multi-increment application to have inter-worker communication while remaining simple.
 \MD{TODO: Add the Figure for the application graph.}
@@
 As our third application, we calculate the histogram of the BigBrain image. 
-It reads the BigBrain blocks from Lustre, calculates each intensity's frequency, and then writes the aggregated result back on Lustre.
+The application reads the BigBrain blocks from Lustre, calculates each intensity's frequency, and then writes the aggregated result back on Lustre.
 This map-reduce application has a very high read overwrite ratio.
-Moreover, this application requires data shuffling, although little. 
-The amount of inter-worker communication is in-between the increment and multi-increment pipelines.
+Moreover, this application requires shuffling, albeit of a limited amoutn of data. 
+The amount of inter-worker communication is in-between the increment and multi-increment applications.
 \MD{TODO: Add the Figure for the application graph.}
@@
 \subsubsection{Kmeans} % TODO
-For our fourth application, we use Kmeans to cluster the BigBrain image by voxel intensity.
-We set the number of clusters to 3 to segment the white and grey matter and the noise.
+For our fourth application, we apply Kmeans clustering to the voxel intensities of the BigBrain image.
+We set the number of clusters to 3, to segment the white and grey matter and the noise.
 The application starts by reading the image blocks, combining all voxels in a 1-D array, and choosing initial centroids using the min, max, and intermediate values.
@@
 It repeats the assignment and update steps for a configurable amount of time. 
-Finally, the voxels of the image blocks are classified and write back on the file system.
-Updating the centroids involves a lot of data communication between the worker.
+Finally, the voxels of the image blocks are classified and written back to the file system.
+Updating the centroids involves substantial data communication between the workers.
 
-Spark and Dask's implementation differs slightly for this application to take advantage of the" most suited" API from both engines.
-Spark implementation uses the Map-Reduce paradigm, while Dask implementation uses array programming.
+For this application, the Spark and Dask implementations differ slightly,  to take advantage of the best-suited API from both engines.
+The Spark implementation uses the Map-Reduce paradigm, while the Dask one uses array programming.
 \MD{TODO: Add the Figure for the application graph.}
@@
 \subsubsection{BIDS app example}
-Our last application is BIDS App example: a neuroimaging pipeline to extract brain volume.
+Our last application is BIDS App example: a neuroimaging pipeline to measure the brain volume from MRIs.
 For this application, we use the CoRR dataset.
-It extracts the brain volume of each participant, then computes the average for each group of participants.
-Unlike the other application, BIDS App example is a command-line set in a Docker image (bids/example on DockerHub).
+The application extracts the brain volume of each participant, then computes the average for each group of participants.
+Unlike the other applications, BIDS App example is a command-line executed in a Docker image (bids/example on DockerHub).
 We converted the Docker image to a Singularity image for use in HPC environments, \MD{Cite paper on reason why this is done.}

### 2021-01-22 src/app/dask/bids.py
[FIX] argument spawning n-workers

@@ -13,7 +13,7 @@ def run(
     input_folder: str,
     output_folder: str,
     scheduler: str,
-    n_workers: int,
+    n_worker: int,
     benchmark: bool,
     container_path: str,
 ) -> None:
@@ -31,7 +31,7 @@ def run(
     if scheduler.lower() == "slurm":
         cluster = SLURMCluster()
         client = Client(cluster)
-        cluster.scale(n_workers)
+        cluster.scale(n_worker)
     else:
         client = Client(scheduler)
 

### 2021-01-22 src/app/dask/histogram.py

@@ -19,7 +19,7 @@ def run(
     input_folder: str,
     output_folder: str,
     scheduler: str,
-    n_workers: int,
+    n_worker: int,
     benchmark: bool,
 ) -> None:
     experiment = f"dask:histogram"
@@ -34,7 +34,7 @@ def run(
     if scheduler.lower() == "slurm":
         cluster = SLURMCluster()
         client = Client(cluster)
-        cluster.scale(n_workers)
+        cluster.scale(n_worker)
     else:
         client = Client(scheduler)
 

### 2021-01-22 src/app/dask/increment.py

@@ -13,7 +13,7 @@ def run(
     input_folder: str,
     output_folder: str,
     scheduler: str,
-    n_workers: int,
+    n_worker: int,
     benchmark: bool,
     *,
     iterations: int,
@@ -31,7 +31,7 @@ def run(
     if scheduler.lower() == "slurm":
         cluster = SLURMCluster()
         client = Client(cluster)
-        cluster.scale(n_workers)
+        cluster.scale(n_worker)
     else:
         client = Client(scheduler)
 

### 2021-01-22 src/app/dask/kmeans.py

@@ -16,7 +16,7 @@ def run(
     input_folder: str,
     output_folder: str,
     scheduler: str,
-    n_workers: int,
+    n_worker: int,
     benchmark: bool,
     *,
     iterations,
@@ -33,7 +33,7 @@ def run(
     if scheduler.lower() == "slurm":
         cluster = SLURMCluster()
         client = Client(cluster)
-        cluster.scale(n_workers)
+        cluster.scale(n_worker)
     else:
         client = Client(scheduler)
 

### 2021-01-22 src/app/dask/multi_increment.py

@@ -14,7 +14,7 @@ def run(
     input_folder: str,
     output_folder: str,
     scheduler: str,
-    n_workers: int,
+    n_worker: int,
     benchmark: bool,
     *,
     iterations: int,
@@ -35,7 +35,7 @@ def run(
     if scheduler.lower() == "slurm":
         cluster = SLURMCluster()
         client = Client(cluster)
-        cluster.scale(n_workers)
+        cluster.scale(n_worker)
     else:
         client = Client(scheduler)
 

### 2021-01-22 src/app/spark/bids.py

@@ -10,7 +10,7 @@ def run(
     input_folder: str,
     output_folder: str,
     scheduler: str,
-    n_workers: int,
+    n_worker: int,
     benchmark: bool,
     container_path: str,
 ) -> None:

### 2021-01-22 src/app/spark/histogram.py

@@ -17,7 +17,7 @@ def run(
     input_folder: str,
     output_folder: str,
     scheduler: str,
-    n_workers: int,
+    n_worker: int,
     benchmark: bool,
 ) -> None:
     experiment = f"spark:histogram"

### 2021-01-22 src/app/spark/increment.py

@@ -11,7 +11,7 @@ def run(
     input_folder: str,
     output_folder: str,
     scheduler: str,
-    n_workers: int,
+    n_worker: int,
     benchmark: bool,
     *,
     iterations: int,

### 2021-01-22 src/app/spark/kmeans.py

@@ -12,7 +12,7 @@ def run(
     input_folder: str,
     output_folder: str,
     scheduler: str,
-    n_workers: int,
+    n_worker: int,
     benchmark: bool,
     *,
     iterations,

### 2021-01-22 src/app/spark/multi_increment.py

@@ -12,7 +12,7 @@ def run(
     input_folder: str,
     output_folder: str,
     scheduler: str,
-    n_workers: int,
+    n_worker: int,
     benchmark: bool,
     *,
     iterations: int,

### 2021-01-22 src/cli.py

@@ -24,7 +24,7 @@ def import_from(module, name):
     type=click.Path(exists=True, dir_okay=True, writable=True),
 )
 @click.option("-s", "--scheduler", required=True)
-@click.option("-n", "--n-worker", type=int, required=False)
+@click.option("-n", "--n-worker", type=int, default=1, required=False)
 @click.option("--benchmark", is_flag=True)
 @engine_choices
 @click.pass_context

### 2021-01-22 src/app/dask/bids.py
[ADD] print dask client at startup

@@ -35,6 +35,8 @@ def run(
     else:
         client = Client(scheduler)
 
+    print(client)
+
     subjects = subject_crawler(input_folder)
     sites = site_crawler(input_folder)
 

### 2021-01-22 src/app/dask/histogram.py

@@ -38,6 +38,8 @@ def run(
     else:
         client = Client(scheduler)
 
+    print(client)
+
     blocks = [
         dask.delayed(load)(
             filename,

### 2021-01-22 src/app/dask/increment.py

@@ -35,6 +35,8 @@ def run(
     else:
         client = Client(scheduler)
 
+    print(client)
+    
     blocks = [
         dask.delayed(load)(
             filename,

### 2021-01-22 src/app/dask/kmeans.py

@@ -37,6 +37,8 @@ def run(
     else:
         client = Client(scheduler)
 
+    print(client)
+    
     blocks = [
         dask.delayed(load)(
             filename,

### 2021-01-22 src/app/dask/multi_increment.py

@@ -39,6 +39,8 @@ def run(
     else:
         client = Client(scheduler)
 
+    print(client)
+
     blocks = [
         dask.delayed(load)(
             filename,

### 2021-01-22 src/app/dask/bids.py
[FIX] set hostname for the scheduler at when launching slurm cluster

@@ -29,7 +29,8 @@ def run(
     }
 
     if scheduler.lower() == "slurm":
-        cluster = SLURMCluster()
+        hostname = os.environ["HOSTNAME"]
+        cluster = SLURMCluster(scheduler_options={"host": hostname})
         client = Client(cluster)
         cluster.scale(n_worker)
     else:

### 2021-01-22 src/app/dask/histogram.py

@@ -1,5 +1,6 @@
 from functools import reduce
 import glob
+import os
 import time
 
 import dask
@@ -32,7 +33,8 @@ def run(
     }
 
     if scheduler.lower() == "slurm":
-        cluster = SLURMCluster()
+        hostname = os.environ["HOSTNAME"]
+        cluster = SLURMCluster(scheduler_options={"host": hostname})
         client = Client(cluster)
         cluster.scale(n_worker)
     else:

### 2021-01-22 src/app/dask/increment.py

@@ -1,4 +1,5 @@
 import glob
+import os
 import time
 
 import dask
@@ -29,14 +30,15 @@ def run(
     }
 
     if scheduler.lower() == "slurm":
-        cluster = SLURMCluster()
+        hostname = os.environ["HOSTNAME"]
+        cluster = SLURMCluster(scheduler_options={"host": hostname})
         client = Client(cluster)
         cluster.scale(n_worker)
     else:
         client = Client(scheduler)
 
     print(client)
-    
+
     blocks = [
         dask.delayed(load)(
             filename,

### 2021-01-22 src/app/dask/kmeans.py

@@ -1,4 +1,5 @@
 import glob
+import os
 import time
 
 import dask
@@ -31,14 +32,15 @@ def run(
     }
 
     if scheduler.lower() == "slurm":
-        cluster = SLURMCluster()
+        hostname = os.environ["HOSTNAME"]
+        cluster = SLURMCluster(scheduler_options={"host": hostname})
         client = Client(cluster)
         cluster.scale(n_worker)
     else:
         client = Client(scheduler)
 
     print(client)
-    
+
     blocks = [
         dask.delayed(load)(
             filename,
@@ -78,7 +80,10 @@ def run(
         ).persist()
 
         centroids = np.array(
-            [da.mean(voxels[centroid_index == c]).persist() for c in range(len(centroids))]
+            [
+                da.mean(voxels[centroid_index == c]).persist()
+                for c in range(len(centroids))
+            ]
         )
         print(f"{centroids=}")
 
@@ -93,7 +98,7 @@ def run(
                 experiment,
                 "update_centroids",
             )
-    
+
     del voxels
     del centroid_index
 

### 2021-01-22 src/app/dask/multi_increment.py

@@ -1,4 +1,5 @@
 import glob
+import os
 import random
 import time
 
@@ -33,7 +34,8 @@ def run(
     }
 
     if scheduler.lower() == "slurm":
-        cluster = SLURMCluster()
+        hostname = os.environ["HOSTNAME"]
+        cluster = SLURMCluster(scheduler_options={"host": hostname})
         client = Client(cluster)
         cluster.scale(n_worker)
     else:

### 2021-01-22 src/app/dask/bids.py
[FIX] only merge logs if benchmark flag is passed

@@ -60,7 +60,8 @@ def run(
     )
     client.gather(futures)
     client.close()
-    merge_logs(
-        output_folder=output_folder,
-        experiment=experiment,
-    )
+    if benchmark:
+        merge_logs(
+            output_folder=output_folder,
+            experiment=experiment,
+        )

### 2021-01-22 src/app/dask/histogram.py

@@ -84,7 +84,8 @@ def run(
     )
 
     client.close()
-    merge_logs(
-        output_folder=output_folder,
-        experiment=experiment,
-    )
+    if benchmark:
+        merge_logs(
+            output_folder=output_folder,
+            experiment=experiment,
+        )

### 2021-01-22 src/app/dask/increment.py

@@ -67,7 +67,8 @@ def run(
     client.gather(futures)
 
     client.close()
-    merge_logs(
-        output_folder=output_folder,
-        experiment=experiment,
-    )
+    if benchmark:
+        merge_logs(
+            output_folder=output_folder,
+            experiment=experiment,
+        )

### 2021-01-22 src/app/dask/kmeans.py

@@ -115,7 +115,8 @@ def run(
     dask.compute(*results)
 
     client.close()
-    merge_logs(
-        output_folder=output_folder,
-        experiment=experiment,
-    )
+    if benchmark:
+        merge_logs(
+            output_folder=output_folder,
+            experiment=experiment,
+        )

### 2021-01-22 src/app/dask/multi_increment.py

@@ -73,7 +73,8 @@ def run(
     client.gather(futures)
 
     client.close()
-    merge_logs(
-        output_folder=output_folder,
-        experiment=experiment,
-    )
+    if benchmark:
+        merge_logs(
+            output_folder=output_folder,
+            experiment=experiment,
+        )

### 2021-01-22 src/app/spark/bids.py

@@ -36,7 +36,9 @@ def run(
 
     sites = sc.parallelize(site_crawler(input_folder), 512)
     sites.map(lambda x: run_group(site=x, **common_args)).collect()
-    merge_logs(
-        output_folder=output_folder,
-        experiment=experiment,
-    )
+    
+    if benchmark:
+        merge_logs(
+            output_folder=output_folder,
+            experiment=experiment,
+        )

### 2021-01-22 src/app/spark/histogram.py

@@ -48,7 +48,9 @@ def run(
     )
 
     save_histogram(histogram, **common_args)
-    merge_logs(
-        output_folder=output_folder,
-        experiment=experiment,
-    )
+    
+    if benchmark:
+        merge_logs(
+            output_folder=output_folder,
+            experiment=experiment,
+        )

### 2021-01-22 src/app/spark/increment.py

@@ -39,7 +39,9 @@ def run(
     img_rdd = img_rdd.map(lambda x: dump(x, **common_args))
 
     img_rdd.collect()
-    merge_logs(
-        output_folder=output_folder,
-        experiment=experiment,
-    )
+    
+    if benchmark:
+        merge_logs(
+            output_folder=output_folder,
+            experiment=experiment,
+        )

### 2021-01-22 src/app/spark/kmeans.py

@@ -85,7 +85,9 @@ def run(
     ).collect()
 
     sc.stop()
-    merge_logs(
-        output_folder=output_folder,
-        experiment=experiment,
-    )
+    
+    if benchmark:
+        merge_logs(
+            output_folder=output_folder,
+            experiment=experiment,
+        )

### 2021-01-22 src/app/spark/multi_increment.py

@@ -47,7 +47,9 @@ def run(
     img_rdd = img_rdd.map(lambda x: dump(x, **common_args))
 
     img_rdd.collect()
-    merge_logs(
-        output_folder=output_folder,
-        experiment=experiment,
-    )
+
+    if benchmark:
+        merge_logs(
+            output_folder=output_folder,
+            experiment=experiment,
+        )

### 2021-01-22 src/app/dask/bids.py
Revert "[ADD] print dask client at startup"
This reverts commit 965016e819d41c5f1178b8c4636bdcfe5667850b.

@@ -36,8 +36,6 @@ def run(
     else:
         client = Client(scheduler)
 
-    print(client)
-
     subjects = subject_crawler(input_folder)
     sites = site_crawler(input_folder)
 

### 2021-01-22 src/app/dask/histogram.py

@@ -40,8 +40,6 @@ def run(
     else:
         client = Client(scheduler)
 
-    print(client)
-
     blocks = [
         dask.delayed(load)(
             filename,

### 2021-01-22 src/app/dask/increment.py

@@ -37,8 +37,6 @@ def run(
     else:
         client = Client(scheduler)
 
-    print(client)
-
     blocks = [
         dask.delayed(load)(
             filename,

### 2021-01-22 src/app/dask/kmeans.py

@@ -39,8 +39,6 @@ def run(
     else:
         client = Client(scheduler)
 
-    print(client)
-
     blocks = [
         dask.delayed(load)(
             filename,

### 2021-01-22 src/app/dask/multi_increment.py

@@ -41,8 +41,6 @@ def run(
     else:
         client = Client(scheduler)
 
-    print(client)
-
     blocks = [
         dask.delayed(load)(
             filename,

### 2021-06-04 latex/paper.tex
[REWORK] methods section.

@@
 
-\title{Impact of engines scheduler on the performance of Big Data pipelines in HPC \TG{Suggested title: Performance comparison of Dask and Apache Spark on HPC systems}}
+\title{Performance comparison of Dask and Apache Spark on HPC systems}
 
@@
 \begin{IEEEkeywords}
-	Performance, Scheduler, Big Data, Dask, Spark, Neuroimaging
+	Performance, Big Data, Dask, Spark, Neuroimaging
 \end{IEEEkeywords}
@@
 The scheduler receives jobs from clients and assigns tasks to available workers.
-Task scheduler uses a FIFO (First-In-First-Out) job scheduling policy.
+Task scheduler uses a LIFO (Last-In-First-Out) job scheduling policy.
 That is an utter process branch of the Dask graph before proceeding with the next one.
@@
 For our experiments, we used the \href{https://jobqueue.dask.org/en/latest/generated/dask_jobqueue.SLURMCluster.html}{Dask SLURM cluster} API.
-
-Dask \MD{Add version when after locking it for experiments.} was used. \TG{should be in methods}
 
@@
 Each operation in a stage is represented by a high-level task in the computation graph.
-Like Dask, Spark Standalone scheduler uses a FIFO policy to schedule tasks.
+Like Dask, Spark Standalone scheduler uses a LIFO policy to schedule tasks.
 Spark Standalone has two execution modes: (1) the client mode, where the driver process runs in a dedicated process,
@@
 Each compute node has 2 16-cores CPUs,
-\SI{256}{\giga\byte} memory, \MD{find out what's the memory speed and \#-channel supported.}
-\SI{2.88}{\tera\byte} SSDs of mounted storage.
+\SI{256}{\giga\byte} \SI{2666}{\mega\hertz} memory in 8-channel,
+and \SI{2.88}{\tera\byte} SSDs of mounted storage.
 The nodes are interconnected with a dedicated \SI{10}{\giga\bit/\second} Ethernet connection.
-Centos \MD{Verify version} and Linux kernel \MD{Verify kernel version} is installed on the nodes.
+Centos 8 with Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64} installed on the compute nodes.
 
-\MD{Write about the worker/leader setup once the dataset size is determined. Need to tune so that 2-3 blocks can fit in the excutors memmory}
+Both Spark and Dask are configured to have 8 worker processes each with 8 threads.
+Each worker is allocated \SI{31.5}{\giga\byte} of memory.
+A new cluster is spinned up and teared down for each experiment.
+
+Dask ?? and Spark ?? \MD{Add version when after locking it for experiments.} was used for our experiments.
 
@@
 We converted the blocks into the NIfTI format, a popular format in neuroimaging.
-We left the NIfTI blocks uncompressed, resulting in a total data size of \SI{00}{\giga\byte}\MD{Verify the size of the dataset when Lustre back on}.
-To evaluate the effect of block size, we resplit these blocks into 30, 125 and 750 blocks of \MD{size for each block segment}, using the \href{https://github.com/big-data-lab-team/sam}{sam} library.
+We left the NIfTI blocks uncompressed, resulting in a total data size of \SI{648}{\giga\byte}.
+To evaluate the effect of block size, we resplit these blocks into 1000, 2500, and 5000 blocks of \SI{648}{\mega\byte}, \SI{259.2}{\mega\byte}, and \SI{129.6}{\mega\byte}, respectively.
+\href{https://github.com/big-data-lab-team/sam}{sam} library was used to resplit the image.
 
-We also used the dataset provided by the Consortium for Reliability and Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) \cite{zuo2014open}.
-This dataset is freely available on \href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{datalad}.
+We also used the dataset provided by the Consortium for Reliability and Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) \cite{zuo2014open}, freely available on \href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{datalad}.
 The entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and functional images of 1,397 subjects acquiredin 29 sites.
@@
 The map-only scenario of this application mimics the processing of multiple independent subjects in parallel.
-\MD{TODO: Add the Figure for the application graph.}
 
+\begin{figure}[!ht]
+	\centering
+	\includegraphics[height=\columnwidth,
+	angle=0]{figures/increment.png}
+	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
+	\label{fig:graph-increment}
+\end{figure}
+
+% Currently ommitted due to issue with memory.
+% 
 \subsubsection{Multi-Increment}
@@
 This change allows the multi-increment application to have inter-worker communication while remaining simple.
-\MD{TODO: Add the Figure for the application graph.}
+
+\begin{figure}[!hb]
+	\centering
+	\includegraphics[height=\columnwidth,
+	angle=0]{figures/multi-increment.png}
+	\caption{Task graph for Multi-Incrementation with 3 iterations and 3 BigBrain blocks.}
+	\label{fig:graph-muti-increment}
+\end{figure}
 
@@
 The amount of inter-worker communication is in-between the increment and multi-increment applications.
-\MD{TODO: Add the Figure for the application graph.}
 
-\subsubsection{Kmeans} % TODO
+\begin{figure}[!hb]
+	\centering
+	\includegraphics[height=\columnwidth,
+	angle=0]{figures/histogram.png}
+	\caption{Task graph for Histogram with 3 BigBrain blocks.}
+	\label{fig:graph-histogram}
+\end{figure}
+
+\subsubsection{Kmeans}
 For our fourth application, we apply Kmeans clustering to the voxel intensities of the BigBrain image.
@@
 The Spark implementation uses the Map-Reduce paradigm, while the Dask one uses array programming.
-\MD{TODO: Add the Figure for the application graph.}
 
-\subsubsection{BIDS app example}
-Our last application is BIDS App example: a neuroimaging pipeline to measure the brain volume from MRIs.
+\subsubsection{BIDS App example}
+Our fifth application is BIDS App example: a neuroimaging pipeline to measure the brain volume from MRIs.
 For this application, we use the CoRR dataset.
@@
 using \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
-\MD{Need to discuss if we preload the image on the worker nodes.}
-\MD{TODO: Add the Figure for the application graph.}
 
-\subsection{Experiments} % TODO
+\subsubsection{BIDS App MRIQC}
+Our last application is BIDS App MRIQC: a neuroimaging pipeline to perfom quality control of a brain image.
+For this application, we use the CoRR dataset. \MD{Validate if this one was used after running the experiments. Maybe another dataset will be required; e.g. ADHD200}
+This application verifies the quality of images on a per-sbuject basis.
+Like the BIDS App example, this application is runned using a command-line tool.
+We use the same method to convert the Docker image to a Singularity image.
+
+\subsection{Experiments}
+Table~\ref{table:parameters} the four parameters that are varied throughout the experiments.
+We varied (1) the number of workers to assess the scalibility of the scheduler for the engine,
+(2) the BigBrain block size in Increment, Multi-Increment, and Histogram to measure the effect of the different I/O pattern and parallelization degrees,
+(3) the number of iterations to evaluate the effect of number of task,
+and (4) the sleep delay to study the effect of task duration.
+It should note that increasing the number of iterations for a given sleep delay also increases the total compute time of an application.
+
+To avoid potential external bias such as caching, background process and network load, we ran the applications in randomized order and cleared the page cache in between every experiments.
+Each benchmark was run ten times.
+
+For each run, we measure the makespan of the application as well as the cumulative time spent in the different functions for read, processing, and writing data.
+The overhead calculation for each CPU thread is the end time of the last processed task minus the total runtime of the tasks ran for this thread.
+Summing those results gives the total overhead for the application.
+
+\begin{table*}[t]
+	\renewcommand{\arraystretch}{1.5}
+	\caption{Parameters for the experiments}\label{table:parameters}
+	\centering
+	\begin{tabular}{|l|c|c|c|c|c|c|}
+		\hline & Increment & Multi-Increment & Kmeans & Histogram & BIDS Example & BIDS MRIQC \\\hline
+		\# of Nodes & \multicolumn{6}{c|}{2, 4, 8} \\\hline
+		\# of Workers & \multicolumn{6}{c|}{16, 32 , 64} \\ \hline
+		Block resplit & \multicolumn{4}{c|}{1000, 2500, 5000}  & \multicolumn{2}{c|}{n/a} \\\hline
+		Block Size {[}\SI{}{\mega\byte}{]} & \multicolumn{4}{c|}{648, 259.2, 129.6} & \multicolumn{2}{c|}{n/a} \\\hline
+		% \# of Subjects & \multicolumn{4}{c|}{n/a} & \multicolumn{2}{c|}{1397} \\ \hline
+		\# of Iterations & \multicolumn{3}{c|}{1, 8, 64}                 & \multicolumn{3}{c|}{n/a} \\\hline
+		Sleep Delay {[}\SI{}{\second}{]} & \multicolumn{2}{c|}{0.25, 1, 4, 16} & \multicolumn{4}{c|}{n/a} \\\hline
+	\end{tabular}
+\end{table*}
+\MD{Decide which row to keep for (\# fo nodes / \# of workers) and (Block resplit / Block size)}
 
 \section{Results} % TODO
-
 \section{Discussion} % TODO
-
 \section{Conclusion} % TODO
-
 \section*{Acknowledgment} % TODO
@@
 \bibliography{IEEEabrv,reference}
-
 \end{document}

### 2021-07-13 launch_experiments.py
[REFACTOR] change start time of the application after cluster spin up.
Also update the requirements to have latest version of dask.

@@ -4,19 +4,18 @@ import random
 import subprocess
 import time
 
-from tqdm import tqdm
-
 
 REPETITIONS = 10
 
 n_nodes = [2, 4, 8]
 n_iterations = [1, 5, 25]
-sleep_time = [0.33, 1, 3, 9]
+n_iterations_small = [1, 3, 9]
+BB_blocks = [1000, 2500, 5000]
 
 default = {
     "node": n_nodes[1],
     "itr": n_iterations[1],
-    "sleep": sleep_time[1],
+    "sleep": 1,
 }
 
 cmd_templates = [
@@ -27,20 +26,13 @@ cmd_templates = [
 benchmark_folder = os.path.join("/", "home", "mathdugre", "ccpe-output")
 os.makedirs(benchmark_folder, exist_ok=True)
 
-BB_5000 = os.path.join(
-    "/", "mnt", "lustre", "mathdugre", "datasets", "bigbrain", "nii", "5000_blocks"
+BB_path = lambda x: os.path.join(
+    "/", "mnt", "lustre", "mathdugre", "datasets", "bigbrain", "nii", f"{x}_blocks"
 )
-BB_2500 = os.path.join(
-    "/", "mnt", "lustre", "mathdugre", "datasets", "bigbrain", "nii", "2500_blocks"
+BB_sample_path = lambda x: os.path.join(
+    "/", "mnt", "lustre", "mathdugre", "datasets", "bigbrain", "nii", f"{x}_blocks-sample"
 )
-BB_1000 = os.path.join(
-    "/", "mnt", "lustre", "mathdugre", "datasets", "bigbrain", "nii", "1000_blocks"
-)
-BB_blocks = {
-    "5000": BB_5000,
-    "2500": BB_2500,
-    "1000": BB_1000,
-}
+
 
 container = os.path.join("/", "home", "mathdugre", "containers", "bids.sif")
 CoRR = os.path.join("/", "mnt", "lustre", "mathdugre", "datasets", "CoRR")
@@ -51,7 +43,7 @@ for cmd_template in cmd_templates:
     for x in n_nodes:
         cmds.append(
             cmd_template.format(
-                BB_5000,
+                BB_path(5000),
                 "increment",
                 f"{5000} {default['itr']} {default['sleep']}",
                 x,
@@ -60,7 +52,7 @@ for cmd_template in cmd_templates:
         )
         cmds.append(
             cmd_template.format(
-                BB_5000,
+                BB_path(5000),
                 "multi-increment",
[…]

### 2021-07-13 src/app/dask/bids.py

@@ -19,15 +19,6 @@ def run(
     container_path: str,
 ) -> None:
     experiment = os.path.join(f"dask:bids:{n_worker=}", str(uuid.uuid1()))
-    start_time = time.time()
-    common_args = {
-        "benchmark_folder": benchmark_folder,
-        "start": start_time,
-        "input_folder": input_folder,
-        "output_folder": output_folder,
-        "experiment": experiment,
-        "container_path": container_path,
-    }
 
     SLURM = scheduler.lower() == "slurm"
     if SLURM:
@@ -38,6 +29,16 @@ def run(
 
     else:
         client = Client(scheduler)
+
+    start_time = time.time()
+    common_args = {
+        "benchmark_folder": benchmark_folder,
+        "start": start_time,
+        "input_folder": input_folder,
+        "output_folder": output_folder,
+        "experiment": experiment,
+        "container_path": container_path,
+    }
 
     subjects = subject_crawler(input_folder)
     sites = site_crawler(input_folder)

### 2021-07-13 src/app/dask/histogram.py

@@ -28,13 +28,6 @@ def run(
     experiment = os.path.join(
         f"dask:histogram:{n_worker=}:{block_size=}", str(uuid.uuid1())
     )
-    start_time = time.time()
-    common_args = {
-        "benchmark_folder": benchmark_folder,
-        "start": start_time,
-        "output_folder": output_folder,
-        "experiment": experiment,
-    }
 
     SLURM = scheduler.lower() == "slurm"
     if SLURM:
@@ -44,6 +37,14 @@ def run(
         cluster.scale(jobs=n_worker)
     else:
         client = Client(scheduler)
+
+    start_time = time.time()
+    common_args = {
+        "benchmark_folder": benchmark_folder,
+        "start": start_time,
+        "output_folder": output_folder,
+        "experiment": experiment,
+    }
 
     filenames = glob.glob(input_folder + "/*.nii")
     paths = db.from_sequence(filenames, npartitions=len(filenames))

### 2021-07-13 src/app/dask/kmeans.py

@@ -14,7 +14,6 @@ from ..commons.kmeans import classify_block, dump
 from ..utils import load, log, merge_logs
 
 
-
 @numba.njit(nogil=True, fastmath=True)
 def _centers_dense(X, labels, n_clusters):
     centers = np.zeros(n_clusters, dtype=np.uint16)
@@ -44,13 +43,6 @@ def run(
     experiment = os.path.join(
         f"dask:kmeans:{n_worker=}:{block_size=}:{iterations=}", str(uuid.uuid1())
     )
-    start_time = time.time()
-    common_args = {
-        "benchmark_folder": benchmark_folder,
-        "start": start_time,
-        "output_folder": output_folder,
-        "experiment": experiment,
-    }
 
     SLURM = scheduler.lower() == "slurm"
     if SLURM:
@@ -60,6 +52,14 @@ def run(
         cluster.scale(jobs=n_worker)
     else:
         client = Client(scheduler)
+
+    start_time = time.time()
+    common_args = {
+        "benchmark_folder": benchmark_folder,
+        "start": start_time,
+        "output_folder": output_folder,
+        "experiment": experiment,
+    }
 
     blocks = [
         dask.delayed(load)(
@@ -121,7 +121,7 @@ def run(
         (centroids,) = dask.compute(new_centers)
 
         print(f"{centroids=}")
-        # del centroid_index
+        del labels
 
         end_time = time.time() - start_time
 

### 2021-07-13 src/app/dask/multi_increment.py

@@ -28,13 +28,6 @@ def run(
         f"dask:multi-increment:{n_worker=}:{block_size=}:{iterations=}:{delay=}:{seed=}",
         str(uuid.uuid1()),
     )
-    start_time = time.time()
-    common_args = {
-        "benchmark_folder": benchmark_folder,
-        "start": start_time,
-        "output_folder": output_folder,
-        "experiment": experiment,
-    }
 
     SLURM = scheduler.lower() == "slurm"
     if SLURM:
@@ -44,6 +37,14 @@ def run(
         cluster.scale(jobs=n_worker)
     else:
         client = Client(scheduler)
+
+    start_time = time.time()
+    common_args = {
+        "benchmark_folder": benchmark_folder,
+        "start": start_time,
+        "output_folder": output_folder,
+        "experiment": experiment,
+    }
 
     blocks = [
         dask.delayed(load)(

### 2021-07-13 src/app/spark/kmeans.py

@@ -23,6 +23,13 @@ def run(
     experiment = os.path.join(
         f"spark:kmeans:{n_worker=}:{block_size=}:{iterations=}", str(uuid.uuid1())
     )
+
+    if scheduler.lower() == "slurm":
+        scheduler = os.environ["MASTER_URL"]
+
+    conf = SparkConf().setMaster(scheduler).setAppName(experiment)
+    sc = SparkContext.getOrCreate(conf=conf)
+
     start_time = time.time()
     common_args = {
         "benchmark_folder": benchmark_folder,
@@ -30,12 +37,6 @@ def run(
         "output_folder": output_folder,
         "experiment": experiment,
     }
-
-    if scheduler.lower() == "slurm":
-        scheduler = os.environ["MASTER_URL"]
-
-    conf = SparkConf().setMaster(scheduler).setAppName(experiment)
-    sc = SparkContext.getOrCreate(conf=conf)
 
     paths = sc.parallelize(glob.glob(input_folder + "/*.nii"))
     blocks = paths.map(lambda p: load(p, **common_args)).cache()

### 2021-07-13 src/app/spark/multi_increment.py

@@ -26,6 +26,13 @@ def run(
         f"spark:multi-increment:{n_worker=}:{block_size=}:{iterations=}:{delay=}:{seed=}",
         str(uuid.uuid1()),
     )
+
+    if scheduler.lower() == "slurm":
+        scheduler = os.environ["MASTER_URL"]
+
+    conf = SparkConf().setMaster(scheduler).setAppName(experiment)
+    sc = SparkContext.getOrCreate(conf=conf)
+
     start_time = time.time()
     common_args = {
         "benchmark_folder": benchmark_folder,
@@ -33,12 +40,6 @@ def run(
         "output_folder": output_folder,
         "experiment": experiment,
     }
-
-    if scheduler.lower() == "slurm":
-        scheduler = os.environ["MASTER_URL"]
-
-    conf = SparkConf().setMaster(scheduler).setAppName(experiment)
-    sc = SparkContext.getOrCreate(conf=conf)
 
     filenames = glob.glob(input_folder + "/*.nii")
     paths = sc.parallelize(filenames, len(filenames))

### 2021-07-13 latex/paper.tex
[ADD] part of intro + results.

@@
 
-\algnewcommand\algorithmicforeach{\textbf{for each}}
-\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-  T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
 \begin{document}
@@
 \section{Introduction}
+The raise in data sharing coupled with improved data collection technologies lead neuroimaging into the Big Data era\cite{ALFAROALMAGRO:18, van2014human}. \MD{cite CONP portal}
+While common neuroimaging workflow engines, such as Nipype\cite{Nipype:11}, are well rounded to process standard compute-intensive pipeline,
+they lack Big Data strategies (i.e. in-memory computing, data locality, and lazy-evaluation) to improve the performance of increasingly prevalent data-intensive pipelines.
+A previous study\cite{hayot2019performance} notes the importantce of in-memory computing and data locality to improve the performance of data-intensive pipelines.
+In this work, we study how performance benchmarks generalize accross Big Data engines in an HPC environement.
+In particular, we study the performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16}, for their sutability to process neuroimaging pipelines.
+Our objective is to assess whether Spark or Dask has substantial performance benefits to process data-intensive pipeline in HPC clusters.
+
+Spark and Dask both provide in-memory computing, data locality, and lazy-evlaution, which is common for Big Data engines.
+Both of their scheduler operate dynamically, which benefit applications with runtime unknown ahead of time\cite{Dask:15}.
+They also provide rich high-level API and support a variety of scyeduler and deployment environment, such as Mesos\cite{hindman2011mesos}, YARN\cite{vavilapalli2013apache}, Kubernetes, and HPC clusters.
+Although sharing similarities, these engines have differences.
+
+% Overview of dask and spark
+First and foremost, Spark is written in Scala while Dask is in Python.
+Given the popularity of Python in the scientific community, this arguably gives a edge to Dask due to the serialization cost between Python and Scala.
+However, if not careful, the Python Global Intepreter Lock (GIL) can significatly reduce the parallelism of an application, in some cases.
+The difference in programmming language also provides different benefits.
+On the one hand, as part of the Scipy ecosystem, Dask provides almost transparent integration with APIs such as Numpy arrays, Pandas dataframe, or RAPIDS GPU accelerated code framework.
+On the other hand, Spark's Java, Scala, R, and Python APIs allow to easily parallelize pipelines with minimal performance loss; thanks to the JVM.
+Our study is restrained to performance, although, we recognize that, in practice, other factors affect the choice of an engine.
+
+% Why we chose Lustre TODO
+While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructure to perform studies.
+This paper focuses on the performance of Big Data engine in HPCs environments.
+
+% Related work TODO
+
+The next section introduce the Lustre file system and both Big Data engines used: Dask and Apache Spark.
+Following, we present the design of our benchmarking experiments.
+We consider two type of data-intensive neuroimaging pipelines: high-resolution imaging and large functional MRI studies.
+The pipelines we chose involve different compute and I/O patterns; e.g. map-reduce or map-only and requiring partial data in memory or the whole dataset.
+Our experiments were executed on a dedicated cluster representative of an HPC environment used to process state-of-the-art data-intensive neuroimaging studies.
+Further sections present our results, dicsussion, and concusion.
 
 \section{Background}
-\subsection{Lustre} % TODO
+\subsection{Lustre}
+% TODO
 
@@
 \subsection{Infrastructure}
-For our experiments we used the SlashBin cluster at Concordia University.
-Each compute node has 2 16-cores CPUs,
+For our experiments we used a dedicated cluster.
+Each compute node has 2 16-cores CPUs (64 threads total),
 \SI{256}{\giga\byte} \SI{2666}{\mega\hertz} memory in 8-channel,
@@
 We also used the dataset provided by the Consortium for Reliability and Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) \cite{zuo2014open}, freely available on \href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{datalad}.
-The entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and functional images of 1,397 subjects acquiredin 29 sites.
+The entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and functional images of 1,397 subjects acquired in 29 sites.
 We used all 3,491 anatomical images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} per image on average).
@@
 Finally, we write the resulting NIfTI image back to Lustre.
-This application allows us to study the engines when their inputs are processed independently.
+Using this application we study the engines when their inputs are processed independently.
 The map-only scenario of this application mimics the processing of multiple independent subjects in parallel.
@@
 
-\subsubsection{BIDS App MRIQC}
-Our last application is BIDS App MRIQC: a neuroimaging pipeline to perfom quality control of a brain image.
-For this application, we use the CoRR dataset. \MD{Validate if this one was used after running the experiments. Maybe another dataset will be required; e.g. ADHD200}
-This application verifies the quality of images on a per-sbuject basis.
-Like the BIDS App example, this application is runned using a command-line tool.
-We use the same method to convert the Docker image to a Singularity image.
-
 \subsection{Experiments}
@@
 	\centering
-	\begin{tabular}{|l|c|c|c|c|c|c|}
-		\hline & Increment & Multi-Increment & Kmeans & Histogram & BIDS Example & BIDS MRIQC \\\hline
-		\# of Nodes & \multicolumn{6}{c|}{2, 4, 8} \\\hline
-		\# of Workers & \multicolumn{6}{c|}{16, 32 , 64} \\ \hline
-		Block resplit & \multicolumn{4}{c|}{1000, 2500, 5000}  & \multicolumn{2}{c|}{n/a} \\\hline
-		Block Size {[}\SI{}{\mega\byte}{]} & \multicolumn{4}{c|}{648, 259.2, 129.6} & \multicolumn{2}{c|}{n/a} \\\hline
-		% \# of Subjects & \multicolumn{4}{c|}{n/a} & \multicolumn{2}{c|}{1397} \\ \hline
-		\# of Iterations & \multicolumn{3}{c|}{1, 8, 64}                 & \multicolumn{3}{c|}{n/a} \\\hline
-		Sleep Delay {[}\SI{}{\second}{]} & \multicolumn{2}{c|}{0.25, 1, 4, 16} & \multicolumn{4}{c|}{n/a} \\\hline
+	\begin{tabular}{|l|c|c|c|c|c|}
+		\hline & Increment & Multi-Increment & Kmeans & Histogram & BIDS Example \\\hline
+		\# of Nodes & \multicolumn{5}{c|}{2, 4, 8} \\\hline
+		Dataset size [\SI{}{\giga\byte}] & 648 &\multicolumn{2}{c|}{486} & 648 & \multicolumn{1}{c|}{39} \\\hline
+		Number of file & \multicolumn{4}{c|}{1000, 2500, 5000}  & \multicolumn{1}{c|}{3491} \\\hline
+		\# of Iterations & 1, 5, 25 & \multicolumn{2}{c|}{1, 3, 9}                 & \multicolumn{2}{c|}{-} \\\hline
 	\end{tabular}
 \end{table*}
-\MD{Decide which row to keep for (\# fo nodes / \# of workers) and (Block resplit / Block size)}
 
-\section{Results} % TODO
-\section{Discussion} % TODO
-\section{Conclusion} % TODO
-\section*{Acknowledgment} % TODO
+\section{Results}
+\subsection{Increment}
+Figure \ref{fig:increment_worker} depicts the total execution time spent in each code segment of the \textit{Increment} application.
+The bars are broken down into Idle, Load, Dump, and other functions representing the overhead from the engine, the read and write time, and the aplication function, respectilvely.
+The mean of each bar was calulated from 10 repetitions, with the error bar representing the standard deviation.
+As expected, the compute time, from \textit{Increment}, remains the same when we vary the number of nodes.
+However, the \textit{Load} and \textit{Dump} time increase with the number of nodes for both engines.
+This is explained by an I/O bottleneck on the Lustre file system.
+
+From Figure \ref{fig:increment_worker}, we also observe that Dask's overhead is higher than Spark's.
+While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
+The difference in total execution time between the engines is explained from this overhead difference.
+\begin{figure}[!h]
+	\centering
+	\includegraphics[clip,width=\columnwidth]{figures/stacked_increment_worker.jpg}
+	\caption{Increment: varying nodes -- 5000 block resplit, 5 iterations, \SI{1}{\second} delay}
+	\label{fig:increment_worker}
+\end{figure}
+
+% Block resplit
+\begin{figure}[!h]
+	\centering
+	\includegraphics[clip,width=\columnwidth]{figures/stacked_increment_block.jpg}
+	\caption{Increment: varying resplit -- 4 nodes, 5 iterations, \SI{1}{\second} delay}
+	\label{fig:increment_block}
+\end{figure}
+
+%  # of iterations
+\begin{figure}[!h]
+	\centering
+	\includegraphics[clip,width=\columnwidth]{figures/stacked_increment_itr.jpg}
+	\caption{Increment: varying iterations -- 4 nodes, 5000 block resplit, \SI{1}{\second} delay}
+	\label{fig:increment_itr}
+\end{figure}
+
+%  Sleep delay
+% \begin{figure}[!h]
+% 	\centering
+% 	\includegraphics[clip,width=\columnwidth]{figures/stacked_increment_delay.jpg}
+% 	\caption{Increment: varying sleep delay -- 4 nodes, 5000 block resplit, 5 iterations}
+% 	\label{fig:increment_delay}
+% \end{figure}
+
+\subsection{Multi-Increment}
+% TODO: experiment
+
+\subsection{Kmeans}
+% TODO: experiement
+% Memory issues
+
+\subsection{Histogram}
+Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} when varying the number of nodes.
+% # of nodes
+\begin{figure}[!h]
+	\centering
+	\includegraphics[clip,width=\columnwidth]{figures/stacked_histogram_worker.jpg}
+	\caption{Histogram: varying nodes -- 5000 block resplit, 5 iterations, \SI{1}{\second} delay}
+	\label{fig:histogram_worker}
+\end{figure}
+
+Figure \ref{fig:histogram_worker} shows the execution time for the \textit{Histogram} when varying the number of resplit.
+% Block resplit
+\begin{figure}[!h]
+	\centering
+	\includegraphics[clip,width=\columnwidth]{figures/stacked_histogram_block.jpg}
+	\caption{Histogram: varying resplit -- 4 nodes, 5 iterations, \SI{1}{\second} delay}
+	\label{fig:histogram_block}
+\end{figure}
+
+\subsection{BIDS Example}
+Figure \ref{fig:bids} shows the total execution time for the \textit{BIDS App example} application when varying the number of nodes.
+We observe that Dask has significantly lower execution time than Spark; between 10\% to 15\%.
+The main difference comes from a increased amount of stagger tasks for Spark, thus increasing the idle time of the application.
+Moreover, a shorther cluster deployement time for Dask reinforce the difference in idle time.
+As with other experiments, the I/O from Lustre affects the scaling of both engines as we increase the number of nodes.
+\begin{figure}[!h]
+	\centering
+	\includegraphics[clip,width=\columnwidth]{figures/stacked_bids.jpg}
+	\caption{BIDS App Example: varying nodes}
+	\label{fig:bids}
+\end{figure}
+
+\section{Discussion}
+% TODO
+
+\section{Conclusion}
+% TODO
+
+\section*{Acknowledgment}
+% TODO
 

### 2021-07-13 launch_experiments.py
[FIX] syntax error.

@@ -160,7 +160,7 @@ random.shuffle(cmds)
 
 total_cmds = len(cmds)
 for i, cmd in enumerate(cmds, 1):
-    print(f"[{datetime.now()}] -- ({i}/{len(total_cmds)}) -- Running : {cmd}")
+    print(f"[{datetime.now()}] -- ({i}/{total_cmds}) -- Running : {cmd}")
     subprocess.run("sudo /home/shared/dropcache_compute.sh", shell=True)
     subprocess.run(cmd, shell=True)
     time.sleep(10)

### 2021-07-13 launch_experiments.py
[FIX] wrong dataset path for some kmeans and multi-increment experiements.

@@ -52,7 +52,7 @@ for cmd_template in cmd_templates:
         )
         cmds.append(
             cmd_template.format(
-                BB_path(5000),
+                BB_sample_path(5000),
                 "multi-increment",
                 f"{5000} {default['itr']} {default['sleep']}",
                 x,
@@ -61,7 +61,7 @@ for cmd_template in cmd_templates:
         )
         cmds.append(
             cmd_template.format(
-                BB_path(5000),
+                BB_sample_path(5000),
                 "kmeans",
                 f"{5000} {default['itr']}",
                 x,
@@ -138,7 +138,7 @@ for cmd_template in cmd_templates:
     for x in n_iterations_small:
         cmds.append(
             cmd_template.format(
-                BB_path(5000),
+                BB_sample_path(5000),
                 "multi-increment",
                 f"{5000} {x} {default['sleep']}",
                 default["node"],
@@ -147,7 +147,7 @@ for cmd_template in cmd_templates:
         )
         cmds.append(
             cmd_template.format(
-                BB_path(5000),
+                BB_sample_path(5000),
                 "kmeans",
                 f"{5000} {x}",
                 default["node"],

### 2021-07-13 src/app/dask/kmeans.py
[FIX] deleted labels var twice.

@@ -136,7 +136,6 @@ def run(
             )
 
     del voxels
-    del labels
 
     results = []
     for block in blocks:

### 2021-07-13 launch_experiments.py
[FIX] launch params.

@@ -15,6 +15,7 @@ BB_blocks = [1000, 2500, 5000]
 default = {
     "node": n_nodes[1],
     "itr": n_iterations[1],
+    "small_itr": n_iterations_small[1],
     "sleep": 1,
 }
 
@@ -54,7 +55,7 @@ for cmd_template in cmd_templates:
             cmd_template.format(
                 BB_sample_path(5000),
                 "multi-increment",
-                f"{5000} {default['itr']} {default['sleep']}",
+                f"{5000} {default['small_itr']} {default['sleep']}",
                 x,
                 benchmark_folder,
             )
@@ -63,7 +64,7 @@ for cmd_template in cmd_templates:
             cmd_template.format(
                 BB_sample_path(5000),
                 "kmeans",
-                f"{5000} {default['itr']}",
+                f"{5000} {default['small_itr']}",
                 x,
                 benchmark_folder,
             )
@@ -101,7 +102,7 @@ for cmd_template in cmd_templates:
             cmd_template.format(
                 BB_sample_path(n),
                 "multi-increment",
-                f"{n} {default['itr']} {default['sleep']}",
+                f"{n} {default['small_itr']} {default['sleep']}",
                 default["node"],
                 benchmark_folder,
             )
@@ -110,7 +111,7 @@ for cmd_template in cmd_templates:
             cmd_template.format(
                 BB_sample_path(n),
                 "kmeans",
-                f"{n} {default['itr']}",
+                f"{n} {default['small_itr']}",
                 default["node"],
                 benchmark_folder,
             )

### 2021-07-13 latex/paper.tex
edits and comments in methods

--- 2021-06-04 latex/paper.tex
@@
 \subsection{Infrastructure}
-For our experiments we used the SlashBin cluster at Concordia University.
-Each compute node has 2 16-cores CPUs,
-\SI{256}{\giga\byte} \SI{2666}{\mega\hertz} memory in 8-channel,
-and \SI{2.88}{\tera\byte} SSDs of mounted storage.
-The nodes are interconnected with a dedicated \SI{10}{\giga\bit/\second} Ethernet connection.
-Centos 8 with Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64} installed on the compute nodes.
+We used the ``slashbin" cluster at Concordia University.
+The cluster has 8 compute nodes, 4 storage nodes, 1 login node, and 1
+Lustre metadata node. Each
+compute node has 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
+@ 2.10GHz, 256~GiB of RAM, 6 $\times$ SSDs of 450~GiB each with the XFS
+file system (no RAID enabled), 378~GiB of tmpfs, 126~GiB of devtmpfs file system,
+CentOS~8.1 and Linux kernel
+\textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
 
-Both Spark and Dask are configured to have 8 worker processes each with 8 threads.
-Each worker is allocated \SI{31.5}{\giga\byte} of memory.
-A new cluster is spinned up and teared down for each experiment.
+\TG{describe the Lustre installation: data nodes and lustre config}
+\TG{Use GB or GiB consistently}
+
+Both Spark and Dask were configured to have 8 worker processes each with 8
+threads. Each worker was allocated \SI{31.5}{\giga\byte} of memory \TG{,
+determined as the amount of memory that ...}. \TG{any other Spark or Dask
+param? Location of log files, etc} A new Dask or Spark cluster was started and
+teared down for each experiment.
 
@@
 \subsection{Dataset}
-We used BigBrain\cite{Amunts:13}, a 3-D image of the human brain with voxel intensity ranging between 0 and 65,535.
-We converted the blocks into the NIfTI format, a popular format in neuroimaging.
-We left the NIfTI blocks uncompressed, resulting in a total data size of \SI{648}{\giga\byte}.
-To evaluate the effect of block size, we resplit these blocks into 1000, 2500, and 5000 blocks of \SI{648}{\mega\byte}, \SI{259.2}{\mega\byte}, and \SI{129.6}{\mega\byte}, respectively.
-\href{https://github.com/big-data-lab-team/sam}{sam} library was used to resplit the image.
+We used BigBrain\cite{Amunts:13}, a 3-D image of the human brain with voxel
+intensities ranging from 0 to 65,535. We converted the blocks into the
+NIfTI format, a popular format in neuroimaging. We left the NIfTI blocks
+uncompressed, resulting in a total data size of \SI{648}{\giga\byte}. To
+evaluate the effect of block size, we resplit these blocks into 1000, 2500,
+and 5000 blocks of \SI{648}{\mega\byte}, \SI{259.2}{\mega\byte}, and
+\SI{129.6}{\mega\byte}, respectively. The
+\href{https://github.com/big-data-lab-team/sam}{sam} library was used to
+resplit the image.
 
-We also used the dataset provided by the Consortium for Reliability and Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) \cite{zuo2014open}, freely available on \href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{datalad}.
-The entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and functional images of 1,397 subjects acquiredin 29 sites.
-We used all 3,491 anatomical images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} per image on average).
+We also used the dataset provided by the Consortium for Reliability and
+Reproducibility
+(\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR})
+\cite{zuo2014open}, freely available on
+\href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
+entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion
+and functional images of 1,397 subjects acquired in 29 sites. We used all
+3,491 anatomical images, representing \SI{39}{\giga\byte} overall
+(\SI{11.17}{\mega\byte} per image on average).
 
 \subsection{Applications}
+
 \subsubsection{Increment}
+
 We adapted the increment application used in \cite{hayot2019performance}.
-This synthetic application reads blocks of the BigBrain from Lustre and simulates computation by sleeping for a specified period.
-To simulate intermediate results, we repeat the sleep process for a configurable amount of time.
-We prevent data caching of the blocks by incrementing their voxels value by one after each sleep operation.
-Finally, we write the resulting NIfTI image back to Lustre.
-This application allows us to study the engines when their inputs are processed independently.
-The map-only scenario of this application mimics the processing of multiple independent subjects in parallel.
+This synthetic application reads blocks of the BigBrain from Lustre and
+simulates computation by sleeping for a specified period. To simulate
+intermediate results, we repeat the sleep process for a configurable amount
+of time. We prevent data caching of the blocks by incrementing their voxels
+value by one after each sleep operation. Finally, we write the resulting
+NIfTI image back to Lustre. This application allows us to study the engines
+when their inputs are processed independently. The map-only pattern
+ mimics the processing of multiple independent subjects in
+parallel.
 
@@
 	angle=0]{figures/increment.png}
-	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
+	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.
+	\TG{nitpick: I'd prefer using ``read and write" than ``load and dump''}}
 	\label{fig:graph-increment}
@@
 \subsubsection{Multi-Increment}
-Our second application is an adaptation of the increment application.
-A significant difference is that, at each iteration, it uses a random BigBrain block as the increment value. 
-This change allows the multi-increment application to have inter-worker communication while remaining simple.
+Our second application is an adaptation of the increment application. A
+significant difference is that, at each iteration, it uses a random
+BigBrain block as the increment value. This change allows the
+multi-increment application to have inter-worker communication while
+remaining simple.
 
@@
 This map-reduce application has a very high read overwrite ratio.
-Moreover, this application requires shuffling, albeit of a limited amoutn of data. 
+Moreover, this application requires shuffling, albeit of a limited amount of data. 
 The amount of inter-worker communication is in-between the increment and multi-increment applications.
@@
 \subsubsection{Kmeans}
-For our fourth application, we apply Kmeans clustering to the voxel intensities of the BigBrain image.
-We set the number of clusters to 3, to segment the white and grey matter and the noise.
-The application starts by reading the image blocks, combining all voxels in a 1-D array, and choosing initial centroids using the min, max, and intermediate values.
-It assigns each voxel to the centroids it is the closest and updates each centroid by computing the average of the voxels associated with it. 
-It repeats the assignment and update steps for a configurable amount of time. 
-Finally, the voxels of the image blocks are classified and written back to the file system.
-Updating the centroids involves substantial data communication between the workers.
+For our fourth application, we apply Kmeans clustering to the voxel
+intensities of the BigBrain image. We set the number of clusters to 3. The
+application starts by reading the image blocks, combining all voxels in a
+1-D array, and choosing initial centroids using the min, max, and
+intermediate values. It assigns each voxel to its closest centroid and
+updates each centroid with the average of its assigned voxels. It repeats
+the assignment and update steps for a configurable number of iterations.
+Finally, the voxels of the image blocks are classified and written back to
+the file system. Updating the centroids involves substantial data
+communication between the workers.
 
-For this application, the Spark and Dask implementations differ slightly,  to take advantage of the best-suited API from both engines.
-The Spark implementation uses the Map-Reduce paradigm, while the Dask one uses array programming.
+For this application, the Spark and Dask implementations differ slightly,
+to take advantage of the best-suited API from both engines. The Spark
+implementation uses the Map-Reduce paradigm, while the Dask one uses array
+programming.
 
 \subsubsection{BIDS App example}
-Our fifth application is BIDS App example: a neuroimaging pipeline to measure the brain volume from MRIs.
-For this application, we use the CoRR dataset.
-The application extracts the brain volume of each participant, then computes the average for each group of participants.
-Unlike the other applications, BIDS App example is a command-line executed in a Docker image (bids/example on DockerHub).
-We converted the Docker image to a Singularity image for use in HPC environments, \MD{Cite paper on reason why this is done.}
-using \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
+Our fifth application is BIDS App example, a neuroimaging pipeline to
+measure the brain volume from MRIs. For this application, we use the CoRR
+dataset. The application extracts the brain volume of each participant,
+then computes the average for each group of participants. Unlike the other
+applications, BIDS App example is a command-line executed in a Docker image
+(bids/example on DockerHub). We converted the Docker image to a Singularity
+image for use in HPC environments, \MD{Cite paper on reason why this is
+done.} using
+\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
 
@@
 
+\TG{You should explain how the applications were implemented in Dask and
+Spark, that you reused the same code in both  cases, which APIs were used 
+and why, etc}
+
 \subsection{Experiments}
-Table~\ref{table:parameters} the four parameters that are varied throughout the experiments.
-We varied (1) the number of workers to assess the scalibility of the scheduler for the engine,
-(2) the BigBrain block size in Increment, Multi-Increment, and Histogram to measure the effect of the different I/O pattern and parallelization degrees,
-(3) the number of iterations to evaluate the effect of number of task,
-and (4) the sleep delay to study the effect of task duration.
-It should note that increasing the number of iterations for a given sleep delay also increases the total compute time of an application.
+Table~\ref{table:parameters} shows the parameters that were varied
+throughout the experiments. We varied (1) the number of workers to assess
+the scalibility of the engine scheduler, (2) the BigBrain
+block size in Increment, Multi-Increment, and Histogram to measure the
+effect of different I/O patterns and parallelization degrees, (3) the
+number of iterations to evaluate the effect of number of task, and (4) the
+sleep delay to study the effect of task duration. It should note that
+increasing the number of iterations for a given sleep delay also increases
+the total compute time of an application.
 
-To avoid potential external bias such as caching, background process and network load, we ran the applications in randomized order and cleared the page cache in between every experiments.
-Each benchmark was run ten times.
+To avoid potential external biases due to caching, background processes and
+network load, we ran the applications in randomized order and cleared the
+page cache in between every experiments. Each benchmark was run ten times.
 
-For each run, we measure the makespan of the application as well as the cumulative time spent in the different functions for read, processing, and writing data.
-The overhead calculation for each CPU thread is the end time of the last processed task minus the total runtime of the tasks ran for this thread.
-Summing those results gives the total overhead for the application.
+For each run, we measured the makespan of the application as well as the
+cumulative time spent in the different functions for read, processing, and
+writing data. The overhead calculation for each CPU thread is the end time
+of the last processed task minus the total runtime of the tasks ran for
+this thread \TG{unclear definition. Don't refer to CPU threads as a given task may be executed on more than 1. Maybe refer to engine executors?}. Summing those results gives the total overhead for the
+application.
 

### 2021-07-13 latex/paper.tex

--- 2021-07-13 latex/paper.tex
@@
 \subsection{Infrastructure}
-For our experiments we used a dedicated cluster.
-Each compute node has 2 16-cores CPUs (64 threads total),
-\SI{256}{\giga\byte} \SI{2666}{\mega\hertz} memory in 8-channel,
-and \SI{2.88}{\tera\byte} SSDs of mounted storage.
-The nodes are interconnected with a dedicated \SI{10}{\giga\bit/\second} Ethernet connection.
-Centos 8 with Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64} installed on the compute nodes.
+We used the ``slashbin" cluster at Concordia University.
+The cluster has 8 compute nodes, 4 storage nodes, 1 login node, and 1
+Lustre metadata node. Each
+compute node has 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
+@ 2.10GHz, 256~GiB of RAM, 6 $\times$ SSDs of 450~GiB each with the XFS
+file system (no RAID enabled), 378~GiB of tmpfs, 126~GiB of devtmpfs file system,
+CentOS~8.1 and Linux kernel
+\textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
 
-Both Spark and Dask are configured to have 8 worker processes each with 8 threads.
-Each worker is allocated \SI{31.5}{\giga\byte} of memory.
-A new cluster is spinned up and teared down for each experiment.
+\TG{describe the Lustre installation: data nodes and lustre config}
+\TG{Use GB or GiB consistently}
+
+Both Spark and Dask were configured to have 8 worker processes each with 8
+threads. Each worker was allocated \SI{31.5}{\giga\byte} of memory \TG{,
+determined as the amount of memory that ...}. \TG{any other Spark or Dask
+param? Location of log files, etc} A new Dask or Spark cluster was started and
+teared down for each experiment.
 
@@
 \subsection{Dataset}
-We used BigBrain\cite{Amunts:13}, a 3-D image of the human brain with voxel intensity ranging between 0 and 65,535.
-We converted the blocks into the NIfTI format, a popular format in neuroimaging.
-We left the NIfTI blocks uncompressed, resulting in a total data size of \SI{648}{\giga\byte}.
-To evaluate the effect of block size, we resplit these blocks into 1000, 2500, and 5000 blocks of \SI{648}{\mega\byte}, \SI{259.2}{\mega\byte}, and \SI{129.6}{\mega\byte}, respectively.
-\href{https://github.com/big-data-lab-team/sam}{sam} library was used to resplit the image.
+We used BigBrain\cite{Amunts:13}, a 3-D image of the human brain with voxel
+intensities ranging from 0 to 65,535. We converted the blocks into the
+NIfTI format, a popular format in neuroimaging. We left the NIfTI blocks
+uncompressed, resulting in a total data size of \SI{648}{\giga\byte}. To
+evaluate the effect of block size, we resplit these blocks into 1000, 2500,
+and 5000 blocks of \SI{648}{\mega\byte}, \SI{259.2}{\mega\byte}, and
+\SI{129.6}{\mega\byte}, respectively. The
+\href{https://github.com/big-data-lab-team/sam}{sam} library was used to
+resplit the image.
 
-We also used the dataset provided by the Consortium for Reliability and Reproducibility (\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR}) \cite{zuo2014open}, freely available on \href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{datalad}.
-The entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion and functional images of 1,397 subjects acquired in 29 sites.
-We used all 3,491 anatomical images, representing \SI{39}{\giga\byte} overall (\SI{11.17}{\mega\byte} per image on average).
+We also used the dataset provided by the Consortium for Reliability and
+Reproducibility
+(\href{http://fcon_1000.projects.nitrc.org/indi/CoRR/html/}{CoRR})
+\cite{zuo2014open}, freely available on
+\href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
+entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion
+and functional images of 1,397 subjects acquired in 29 sites. We used all
+3,491 anatomical images, representing \SI{39}{\giga\byte} overall
+(\SI{11.17}{\mega\byte} per image on average).
 
 \subsection{Applications}
+
 \subsubsection{Increment}
+
 We adapted the increment application used in \cite{hayot2019performance}.
-This synthetic application reads blocks of the BigBrain from Lustre and simulates computation by sleeping for a specified period.
-To simulate intermediate results, we repeat the sleep process for a configurable amount of time.
-We prevent data caching of the blocks by incrementing their voxels value by one after each sleep operation.
-Finally, we write the resulting NIfTI image back to Lustre.
-Using this application we study the engines when their inputs are processed independently.
-The map-only scenario of this application mimics the processing of multiple independent subjects in parallel.
+This synthetic application reads blocks of the BigBrain from Lustre and
+simulates computation by sleeping for a specified period. To simulate
+intermediate results, we repeat the sleep process for a configurable amount
+of time. We prevent data caching of the blocks by incrementing their voxels
+value by one after each sleep operation. Finally, we write the resulting
+NIfTI image back to Lustre. This application allows us to study the engines
+when their inputs are processed independently. The map-only pattern
+ mimics the processing of multiple independent subjects in
+parallel.
 
@@
 \subsubsection{Kmeans}
-For our fourth application, we apply Kmeans clustering to the voxel intensities of the BigBrain image.
-We set the number of clusters to 3, to segment the white and grey matter and the noise.
-The application starts by reading the image blocks, combining all voxels in a 1-D array, and choosing initial centroids using the min, max, and intermediate values.
-It assigns each voxel to the centroids it is the closest and updates each centroid by computing the average of the voxels associated with it. 
-It repeats the assignment and update steps for a configurable amount of time. 
-Finally, the voxels of the image blocks are classified and written back to the file system.
-Updating the centroids involves substantial data communication between the workers.
+For our fourth application, we apply Kmeans clustering to the voxel
+intensities of the BigBrain image. We set the number of clusters to 3. The
+application starts by reading the image blocks, combining all voxels in a
+1-D array, and choosing initial centroids using the min, max, and
+intermediate values. It assigns each voxel to its closest centroid and
+updates each centroid with the average of its assigned voxels. It repeats
+the assignment and update steps for a configurable number of iterations.
+Finally, the voxels of the image blocks are classified and written back to
+the file system. Updating the centroids involves substantial data
+communication between the workers.
 
-For this application, the Spark and Dask implementations differ slightly,  to take advantage of the best-suited API from both engines.
-The Spark implementation uses the Map-Reduce paradigm, while the Dask one uses array programming.
+For this application, the Spark and Dask implementations differ slightly,
+to take advantage of the best-suited API from both engines. The Spark
+implementation uses the Map-Reduce paradigm, while the Dask one uses array
+programming.
 
 \subsubsection{BIDS App example}
-Our fifth application is BIDS App example: a neuroimaging pipeline to measure the brain volume from MRIs.
-For this application, we use the CoRR dataset.
-The application extracts the brain volume of each participant, then computes the average for each group of participants.
-Unlike the other applications, BIDS App example is a command-line executed in a Docker image (bids/example on DockerHub).
-We converted the Docker image to a Singularity image for use in HPC environments, \MD{Cite paper on reason why this is done.}
-using \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
+Our fifth application is BIDS App example, a neuroimaging pipeline to
+measure the brain volume from MRIs. For this application, we use the CoRR
+dataset. The application extracts the brain volume of each participant,
+then computes the average for each group of participants. Unlike the other
+applications, BIDS App example is a command-line executed in a Docker image
+(bids/example on DockerHub). We converted the Docker image to a Singularity
+image for use in HPC environments, \MD{Cite paper on reason why this is
+done.} using
+\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
+
+\TG{You should explain how the applications were implemented in Dask and
+Spark, that you reused the same code in both  cases, which APIs were used 
+and why, etc}
 
 \subsection{Experiments}
-Table~\ref{table:parameters} the four parameters that are varied throughout the experiments.
-We varied (1) the number of workers to assess the scalibility of the scheduler for the engine,
-(2) the BigBrain block size in Increment, Multi-Increment, and Histogram to measure the effect of the different I/O pattern and parallelization degrees,
-(3) the number of iterations to evaluate the effect of number of task,
-and (4) the sleep delay to study the effect of task duration.
-It should note that increasing the number of iterations for a given sleep delay also increases the total compute time of an application.
+Table~\ref{table:parameters} shows the parameters that were varied
+throughout the experiments. We varied (1) the number of workers to assess
+the scalibility of the engine scheduler, (2) the BigBrain
+block size in Increment, Multi-Increment, and Histogram to measure the
+effect of different I/O patterns and parallelization degrees, (3) the
+number of iterations to evaluate the effect of number of task, and (4) the
+sleep delay to study the effect of task duration. It should note that
+increasing the number of iterations for a given sleep delay also increases
+the total compute time of an application.
 
-To avoid potential external bias such as caching, background process and network load, we ran the applications in randomized order and cleared the page cache in between every experiments.
-Each benchmark was run ten times.
+To avoid potential external biases due to caching, background processes and
+network load, we ran the applications in randomized order and cleared the
+page cache in between every experiments. Each benchmark was run ten times.
 
-For each run, we measure the makespan of the application as well as the cumulative time spent in the different functions for read, processing, and writing data.
-The overhead calculation for each CPU thread is the end time of the last processed task minus the total runtime of the tasks ran for this thread.
-Summing those results gives the total overhead for the application.
+For each run, we measured the makespan of the application as well as the
+cumulative time spent in the different functions for read, processing, and
+writing data. The overhead calculation for each CPU thread is the end time
+of the last processed task minus the total runtime of the tasks ran for
+this thread \TG{unclear definition. Don't refer to CPU threads as a given task may be executed on more than 1. Maybe refer to engine executors?}. Summing those results gives the total overhead for the
+application.
 

### 2021-07-28 latex/paper.tex
[UPDATE] Modified intro, background, methods, and results.
- Intro: Add related work + clarify goal.
- Application: Add a introduction explaining the overall code base setup.
- Background: Add Lustre background.
- Results: Add increment + histogram.

@@
 \begin{IEEEkeywords}
-	Performance, Big Data, Dask, Spark, Neuroimaging
+	Performance, Big Data, Dask, Spark, Neuroimaging, HPC
 \end{IEEEkeywords}
-\MD{Discuss which keyword to use for the paper.}
 
@@
 In this work, we study how performance benchmarks generalize accross Big Data engines in an HPC environement.
-In particular, we study the performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16}, for their sutability to process neuroimaging pipelines.
+In particular, we study the detailed performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16}, for their suitability to process neuroimaging pipelines.
 Our objective is to assess whether Spark or Dask has substantial performance benefits to process data-intensive pipeline in HPC clusters.
 
-Spark and Dask both provide in-memory computing, data locality, and lazy-evlaution, which is common for Big Data engines.
+Spark and Dask both provide in-memory computing, data locality, and lazy-evaluation, which is common for Big Data engines.
 Both of their scheduler operate dynamically, which benefit applications with runtime unknown ahead of time\cite{Dask:15}.
@@
 
-% Overview of dask and spark
 First and foremost, Spark is written in Scala while Dask is in Python.
@@
 
-% Why we chose Lustre TODO
 While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructure to perform studies.
 This paper focuses on the performance of Big Data engine in HPCs environments.
+We installed the Lustre file system \MD{Not sure which paper to cite} on a dedicated cluster to mimic an HPC system.
+Lustre is a powerful and scalable file system for HPC environement offering parallel file access.
 
-% Related work TODO
+In our previous study\cite{8943502} we found that their was no significant performance between Apache Spark and Dask for data-intensive application.
+However, due to the use of a network file-system (NFS) the performance of the application were limited by an I/O bottleneck.
+Another study \cite{8588652} also reported network and I/O bottleneck when having too many nodes.
+This same study found that the startup overhead time for Spark was larger than Dask.
+On the opposite, \cite{Mehta:17} claims that Dask startup overhead is larger than Spark.
+This suggests that the startup overhead might vary depending on cluster configuration.
+\cite{Mehta:17} processed a \SI{100}{\giga\byte} dataset using a neuroimaging pipeline and found Dask to be up to 14\% faster than Spark due to ``more efficient pipelining" and serialization time to Python.
+On the contrary, \cite{10.1145/3225058.3225128} shows that with large datasets, Spark provides better speedup than Dask.
+However, that study used the Dask Bag API which is know to have performance limitation compared to other Dask APIs.
+The above studies compared the engines performance at a high-level.
+In comparison, our study provides a detailed analysis of the performance differences and similarities between Dask and Spark for data-intensive applications.
 
@@
 \subsection{Lustre}
-% TODO
+The Lustre file-system is a parallel file system well suited for HPC environment.
+Lustre offers a POSIX compliant interface with the file system.
+Powering some of the most powerful supercomputer, Lustre is designed to scale to thousands of nodes, petabytes of storage, and hundreds of gigabit per second of I/O throughput.
+
+At the core of Lustre are the \textit{management servers} (MGSs) and \textit{metatdata server} (MDSs) in charge of the files operations and metadata,
+the \textit{object storage targets} (OSTs) to store file data, and the \textit{obejct sotrage server} (OSSs) to handle the OSTs.
+Lustre clients will then make requests to the MGSs to access files on the OSTs.
+To handle failure in the system, the MGSs, MDSs, and OSTs have failover capabilities.
+To connect the different component of the Lustre file system, it is possible to use InfiniBand network, Ethernet Network, or both.
 
@@
 uncompressed, resulting in a total data size of \SI{648}{\giga\byte}. To
-evaluate the effect of block size, we resplit these blocks into 1000, 2500,
-and 5000 blocks of \SI{648}{\mega\byte}, \SI{259.2}{\mega\byte}, and
+evaluate the effect of block size, we split the image into 1000, 2500,
+and 5000 files of \SI{648}{\mega\byte}, \SI{259.2}{\mega\byte}, and
 \SI{129.6}{\mega\byte}, respectively. The
 \href{https://github.com/big-data-lab-team/sam}{sam} library was used to
-resplit the image.
+split the image.
 
@@
 \subsection{Applications}
+In this section, we describe the applications used to benchmark Dask and Spark.
+We use three simple synthetic applications to have a deep understanding of the 
+underlying behavior of the applications with simple I/O pattern; namely
+\textit{Increment}, \textit{Multi-Increment}, and \textit{Histogram}.
+Also, two neuroimaging applications are used to study more realistic 
+applications; namely \textit{Kmeans} and \textit{BIDS App Example}.
+
+To have a fair comparison between the engines, the code base for the core functions of the applications is the same for both engines.
+This allow us to reduce the performance difference from different implementations.
+Note that this is not the case for the \textit{Kmeans} application as the best performing API for Dask and Spark differed largely in term of implementation.
 
 \subsubsection{Increment}
-
 We adapted the increment application used in \cite{hayot2019performance}.
@@
 when their inputs are processed independently. The map-only pattern
- mimics the processing of multiple independent subjects in
-parallel.
+mimics the processing of multiple independent subjects in parallel.
 
@@
 	angle=0]{figures/increment.png}
-	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.
-	\TG{nitpick: I'd prefer using ``read and write" than ``load and dump''}}
+	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
 	\label{fig:graph-increment}
@@
 % 
-\subsubsection{Multi-Increment}
-Our second application is an adaptation of the increment application. A
-significant difference is that, at each iteration, it uses a random
-BigBrain block as the increment value. This change allows the
-multi-increment application to have inter-worker communication while
-remaining simple.
+% \subsubsection{Multi-Increment}
+% Our second application is an adaptation of the increment application. A
+% significant difference is that, at each iteration, it uses a random
+% BigBrain block as the increment value. This change allows the
+% multi-increment application to have inter-worker communication while
+% remaining simple.
 
-\begin{figure}[!hb]
-	\centering
-	\includegraphics[height=\columnwidth,
-	angle=0]{figures/multi-increment.png}
-	\caption{Task graph for Multi-Incrementation with 3 iterations and 3 BigBrain blocks.}
-	\label{fig:graph-muti-increment}
-\end{figure}
+% \begin{figure}[!hb]
+% 	\centering
+% 	\includegraphics[height=\columnwidth,
+% 	angle=0]{figures/multi-increment.png}
+% 	\caption{Task graph for Multi-Incrementation with 3 iterations and 3 BigBrain blocks.}
+% 	\label{fig:graph-muti-increment}
+% \end{figure}
 
@@
 \TG{You should explain how the applications were implemented in Dask and
-Spark, that you reused the same code in both  cases, which APIs were used 
-and why, etc}
+Spark which APIs were used and why, etc}
 
@@
 
+%  TODO Fix table for the final parameters/applications used.
 \begin{table*}[t]
@@
 	\begin{tabular}{|l|c|c|c|c|c|}
-		\hline & Increment & Multi-Increment & Kmeans & Histogram & BIDS Example \\\hline
+		\hline & Increment & Multi-Increment & Kmeans & Histogram & BIDS App Example \\\hline
 		\# of Nodes & \multicolumn{5}{c|}{2, 4, 8} \\\hline
@@
 Figure \ref{fig:increment_worker} depicts the total execution time spent in each code segment of the \textit{Increment} application.
-The bars are broken down into Idle, Load, Dump, and other functions representing the overhead from the engine, the read and write time, and the aplication function, respectilvely.
+The bars are broken down into Idle, Read, Write, and other application specific functions.
 The mean of each bar was calulated from 10 repetitions, with the error bar representing the standard deviation.
@@
 While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
-The difference in total execution time between the engines is explained from this overhead difference.
+Althought Dask has higher overhead, the total time for both engine is similar; with a maximum makespan difference of \SI{7}{\second} for 2 nodes.
+This is due to a tradeoff between disk bandwidth and compute/idle time.
+When a multiple threads access Lustre concurrently, they shared the disk bandwitdth resulting in longer I/O time.
+However, when most threads are either idle or performing computations, it reduce the contention on the disk bandwitdh.
 \begin{figure}[!h]
@@
 	\includegraphics[clip,width=\columnwidth]{figures/stacked_increment_worker.jpg}
-	\caption{Increment: varying nodes -- 5000 block resplit, 5 iterations, \SI{1}{\second} delay}
+	\caption{Increment: varying nodes -- 5000 files, 5 iterations, \SI{1}{\second} delay}
 	\label{fig:increment_worker}
@@
 
-% Block resplit
-\begin{figure}[!h]
-	\centering
-	\includegraphics[clip,width=\columnwidth]{figures/stacked_increment_block.jpg}
-	\caption{Increment: varying resplit -- 4 nodes, 5 iterations, \SI{1}{\second} delay}
-	\label{fig:increment_block}
-\end{figure}
+Figure \ref{fig:increment_itr} shows the detailed execution time for the \textit{Increment} application while varying the number of iterations.
+As expected, the idle time for both engines increases with the number of iterations.
+However, idle time for Dask increase significantly faster that Spark; resulting in a \SI{38}{\second} makespan difference for 25 iterations.
 
+Figure \ref{fig:increment_itr} also depicts that as the total computation time increases the total I/O time decreases.
+This is explained by an I/O bottleneck on the Lustre file system.
+When threads are busy with computations they reduce the contention on Lustre, thus improving I/O speed for individual accesses.
 %  # of iterations
@@
 	\includegraphics[clip,width=\columnwidth]{figures/stacked_increment_itr.jpg}
-	\caption{Increment: varying iterations -- 4 nodes, 5000 block resplit, \SI{1}{\second} delay}
+	\caption{Increment: varying iterations -- 4 nodes, 5000 files, \SI{1}{\second} delay}
 	\label{fig:increment_itr}
@@
 
-%  Sleep delay
-% \begin{figure}[!h]
-% 	\centering
-% 	\includegraphics[clip,width=\columnwidth]{figures/stacked_increment_delay.jpg}
-% 	\caption{Increment: varying sleep delay -- 4 nodes, 5000 block resplit, 5 iterations}
-% 	\label{fig:increment_delay}
-% \end{figure}
+Figure \ref{fig:increment_block} shows an increase in \textit{Increment} time.
+This is because the increment is performed on each file.
+Thus increasing the number of file increase the computation time.
+Similarly to Fig. \ref{fig:increment_itr} increasing the total compute time of the application reduce the total I/O time.
+Again, this is explain by the I/O tradeoff between compute time and contention of disk bandwidth.
 
-\subsection{Multi-Increment}
+From figure \ref{fig:increment_block} we also observe that Dask has significantly higher idle time than Spark.
+However, the total of the application is similar due to the tradeoff between I/O and compute/idle time.
+
+% # of files
+\begin{figure}[!h]
+	\centering
+	\includegraphics[clip,width=\columnwidth]{figures/stacked_increment_block.jpg}
+	\caption{Increment: varying \# of files -- 4 nodes, 5 iterations, \SI{1}{\second} delay}
+	\label{fig:increment_block}
+\end{figure}
+
+% \subsection{Multi-Increment}
 % TODO: experiment
+% Memory issue with Spark (Heap space runs out of memory)
 
 \subsection{Kmeans}
-% TODO: experiement
-% Memory issues
 
@@
 Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} when varying the number of nodes.
+Overall, Spark has upt to 5\% lower total execution time with the difference comming from a shorter execution time for \textit{Calculate\_histogram}.
+This is unexpected as the code base for both engine is the same; with only the calling APIs differing.
+% This happened for both Bags and Delayed. There might be a higher serialization cost.
+As with previous experiments, the overhead for Spark is less and scale better, although, compensated by a longer total I/O time.
 % # of nodes
@@
 	\includegraphics[clip,width=\columnwidth]{figures/stacked_histogram_worker.jpg}
-	\caption{Histogram: varying nodes -- 5000 block resplit, 5 iterations, \SI{1}{\second} delay}
+	\caption{Histogram: varying nodes -- 5000 files, 5 iterations, \SI{1}{\second} delay}
 	\label{fig:histogram_worker}
@@
 
-Figure \ref{fig:histogram_worker} shows the execution time for the \textit{Histogram} when varying the number of resplit.
-% Block resplit
+Figure \ref{fig:histogram_worker} shows the execution time for the \textit{Histogram} when varying the number of files.
+The results are similar to varying the number of nodes with Spark having faster compute time for \textit{Calculate\_histogram}.
+However, the makespan difference is much smaller due to larger I/O time for both engine.
+%  # of files
 \begin{figure}[!h]
@@
 	\includegraphics[clip,width=\columnwidth]{figures/stacked_histogram_block.jpg}
-	\caption{Histogram: varying resplit -- 4 nodes, 5 iterations, \SI{1}{\second} delay}
+	\caption{Histogram: varying \# of files -- 4 nodes, 5 iterations, \SI{1}{\second} delay}
 	\label{fig:histogram_block}
@@
 
-\subsection{BIDS Example}
+\subsection{BIDS App Example}
 Figure \ref{fig:bids} shows the total execution time for the \textit{BIDS App example} application when varying the number of nodes.

### 2021-07-28 plot-script/plot.py

@@ -335,56 +335,8 @@ def stacked_bar(
         framework = experiment.split(":")[0]
         results[framework][experiment].append(x)
 
-    # Application makespan
-    for summary_xs, (framework, experiments) in enumerate(sorted(results.items())):
-        data = [
-            [
-                pd.read_csv(file_, names=col_names).end.max()
-                for file_ in experiments[key]
-            ]
-            for key in sorted(
-                experiments,
-                key=lambda k: float(re.search(f"{freedom}=(\d+\.?\d*)", k).group(1)),
-            )
-        ]
-
-        stats = {
-            "mean": list(map(np.mean, data)),
-            "std": list(map(np.std, data)),
-        }
-        y_offset = max(stats["mean"]) * 0.01
-        totals = stats["mean"]
-        color = Colorblind8[summary_xs % len(Colorblind8)]
-
-        ############
-        # Plotting #
-        ############
-        ax.bar(
-            x_pos(summary_xs),
-            stats["mean"],
-            yerr=stats["std"],
-            color=color,
-            width=bar_width,
-            edgecolor="black",
-            alpha=1,
-            #             label=framework.capitalize(),
-        )
-
-        # Annotate bar height
-        for x, total in zip(x_pos(summary_xs), totals):
-            ax.text(
-                x,
-                total + y_offset,
-                round(total),
-                ha="center",
-                weight="bold",
-                fontsize=14,
-                color=color,
-            )
-
     # Application detailed time spent
-    ax2 = ax.twinx()
-    for detailed_xs, (framework, experiments) in enumerate(sorted(results.items())):
+    for xs, (framework, experiments) in enumerate(sorted(results.items())):
         data = [
             np.array(
                 [
@@ -409,11 +361,11 @@ def stacked_bar(
         ############
         y_offset = stats["mean"].max() * 0.01
         totals = stats["mean"].sum(axis=0)
-        color = Colorblind8[detailed_xs % len(Colorblind8)]
+        color = Colorblind8[xs % len(Colorblind8)]
         # Plot idle time as a special case since it always exist.
         # Thus we want to having consistent hatch and position.
-        ax2.bar(
-            x_pos(detailed_xs + summary_xs + 1),
+        ax.bar(
+            x_pos(xs),
             stats["mean"][0],
[…]

### 2021-07-29 latex/paper.tex

@@
 \section{Introduction}
-The raise in data sharing coupled with improved data collection technologies lead neuroimaging into the Big Data era\cite{ALFAROALMAGRO:18, van2014human}. \MD{cite CONP portal}
+The rise in data sharing coupled with improved data collection technologies lead neuroimaging into the Big Data era\cite{ALFAROALMAGRO:18, van2014human}. \MD{cite CONP portal}
 While common neuroimaging workflow engines, such as Nipype\cite{Nipype:11}, are well rounded to process standard compute-intensive pipeline,
 they lack Big Data strategies (i.e. in-memory computing, data locality, and lazy-evaluation) to improve the performance of increasingly prevalent data-intensive pipelines.
-A previous study\cite{hayot2019performance} notes the importantce of in-memory computing and data locality to improve the performance of data-intensive pipelines.
+A previous study\cite{hayot2019performance} noted the importantce of in-memory computing and data locality to improve the performance of data-intensive pipelines.
 In this work, we study how performance benchmarks generalize accross Big Data engines in an HPC environement.
@@
 Both of their scheduler operate dynamically, which benefit applications with runtime unknown ahead of time\cite{Dask:15}.
-They also provide rich high-level API and support a variety of scyeduler and deployment environment, such as Mesos\cite{hindman2011mesos}, YARN\cite{vavilapalli2013apache}, Kubernetes, and HPC clusters.
+They also provide rich high-level API and support a variety of schedulers and deployment environment, such as Mesos\cite{hindman2011mesos}, YARN\cite{vavilapalli2013apache}, Kubernetes, and HPC clusters.
 Although sharing similarities, these engines have differences.
@@
 On the one hand, as part of the Scipy ecosystem, Dask provides almost transparent integration with APIs such as Numpy arrays, Pandas dataframe, or RAPIDS GPU accelerated code framework.
-On the other hand, Spark's Java, Scala, R, and Python APIs allow to easily parallelize pipelines with minimal performance loss; thanks to the JVM.
-Our study is restrained to performance, although, we recognize that, in practice, other factors affect the choice of an engine.
+On the other hand, Spark's Java, Scala, R, and Python APIs allow to easily parallelize pipelines with minimal performance loss, thanks to the Java Virtual Machine \TG{you could add a few words to explain why the JVM is useful for parallelization ("thanks to the JVM that ...")}.
+Our study is restrained to performance, although, we recognize that other factors affect the choice of an engine in practice.
 
-While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructure to perform studies.
-This paper focuses on the performance of Big Data engine in HPCs environments.
-We installed the Lustre file system \MD{Not sure which paper to cite} on a dedicated cluster to mimic an HPC system.
-Lustre is a powerful and scalable file system for HPC environement offering parallel file access.
+While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructures.
+This paper focuses on the performance of Big Data engines in HPCs environments.
+We installed the Slurm scheduler and Lustre file system~\cite{braam2019lustre} on a dedicated cluster to mimic a typical HPC system.
+Lustre is a powerful and scalable file system for HPC environement offering parallel file accesses.
 
-In our previous study\cite{8943502} we found that their was no significant performance between Apache Spark and Dask for data-intensive application.
-However, due to the use of a network file-system (NFS) the performance of the application were limited by an I/O bottleneck.
-Another study \cite{8588652} also reported network and I/O bottleneck when having too many nodes.
-This same study found that the startup overhead time for Spark was larger than Dask.
-On the opposite, \cite{Mehta:17} claims that Dask startup overhead is larger than Spark.
+In a previous study\cite{8943502} we found that their was no significant performance difference between Apache Spark and Dask for data-intensive application.
+However, due to the use of the network file-system (NFS) the performance of the applications was limited by an I/O bottleneck.
+Another study \cite{8588652} also reported network and I/O bottlenecks when concurrency increased.
+This same study found that the startup overhead time for Spark was larger than for Dask.
+On the contrary, the work in~\cite{Mehta:17} claims that the startup overhead is larger for Dask than for Spark.
 This suggests that the startup overhead might vary depending on cluster configuration.
-\cite{Mehta:17} processed a \SI{100}{\giga\byte} dataset using a neuroimaging pipeline and found Dask to be up to 14\% faster than Spark due to ``more efficient pipelining" and serialization time to Python.
-On the contrary, \cite{10.1145/3225058.3225128} shows that with large datasets, Spark provides better speedup than Dask.
+The latter study processed a \SI{100}{\giga\byte} dataset using a neuroimaging pipeline and found Dask to be up to 14\% faster than Spark due to ``more efficient pipelining" and serialization time to Python.
+On the contrary, the work in \cite{10.1145/3225058.3225128} shows that with large datasets, Spark provides better speedup than Dask.
 However, that study used the Dask Bag API which is know to have performance limitation compared to other Dask APIs.
@@
 
-The next section introduce the Lustre file system and both Big Data engines used: Dask and Apache Spark.
-Following, we present the design of our benchmarking experiments.
-We consider two type of data-intensive neuroimaging pipelines: high-resolution imaging and large functional MRI studies.
+The next section introduce the Lustre file system and the Big Data engines.
+Next, we present the design of our benchmarking experiments.
+We consider two types of data-intensive neuroimaging pipelines: high-resolution imaging and large functional MRI studies.
 The pipelines we chose involve different compute and I/O patterns; e.g. map-reduce or map-only and requiring partial data in memory or the whole dataset.
-Our experiments were executed on a dedicated cluster representative of an HPC environment used to process state-of-the-art data-intensive neuroimaging studies.
-Further sections present our results, dicsussion, and concusion.
+Further sections present our results, discussion, and conclusion.
 
@@
 \subsection{Lustre}
-The Lustre file-system is a parallel file system well suited for HPC environment.
+The Lustre file-system is a parallel file system well suited for HPC environments.
 Lustre offers a POSIX compliant interface with the file system.
-Powering some of the most powerful supercomputer, Lustre is designed to scale to thousands of nodes, petabytes of storage, and hundreds of gigabit per second of I/O throughput.
+Installed in some of the most powerful supercomputer, Lustre is designed to scale to thousands of nodes, petabytes of storage, and hundreds of gigabits per second of I/O throughput.
 
@@
 the \textit{object storage targets} (OSTs) to store file data, and the \textit{obejct sotrage server} (OSSs) to handle the OSTs.
-Lustre clients will then make requests to the MGSs to access files on the OSTs.
-To handle failure in the system, the MGSs, MDSs, and OSTs have failover capabilities.
-To connect the different component of the Lustre file system, it is possible to use InfiniBand network, Ethernet Network, or both.
+Lustre clients make requests to the MGSs to access files on the OSTs.
+To handle failures in the system, the MGSs, MDSs, and OSTs have failover capabilities.
+InfiniBand, Ethernet, or a combination of both can be used to interconnect these different components.
 
@@
 The engine builds a dynamic graph before execution, allowing it to determine which task to compute.
-Dask workflows can further reduce data transfer by leveraging multithreading whenever Python's GIL does not restrict it.
-Fault-tolerance is achieved by recording data lineage: the sequence of operations used to modify the initial data.
+Dask workflows can further reduce data transfer by leveraging multithreading whenever the Python GIL does not restrict it.
+Fault-tolerance is achieved by recording data lineage, the sequence of operations used to modify the initial data.
 
@@
 Lastly, Dask Futures can also execute arbitrary functions; however, it launches computation immediately rather than lazily.
-Dask modularity allows users to install only required components making it lightweight.
+Dask allows users to install only required components making it lightweight.
 
@@
 In the Dask Distributed scheduler, a \textit{dask-scheduler} process administrates the resource provided by  \textit{dask-worker}s in the cluster.
-The scheduler receives jobs from clients and assigns tasks to available workers.
-Task scheduler uses a LIFO (Last-In-First-Out) job scheduling policy.
-That is an utter process branch of the Dask graph before proceeding with the next one.
+The scheduler receives jobs from clients and assigns tasks to available workers using a LIFO (Last-In-First-Out) job scheduling policy.
+That is an utter process branch of the Dask graph before proceeding with the next one \TG{No comprendo, please revise syntax :)}.
 
@@
 
-In the Spark Standalone scheduler, a \textit{leader} \TG{I am all in favor of inclusive terminology, but if Spark didn't update their wording we shouldn't do it for them} coordinates the resource provisioned by \textit{workers} in the cluster.
+In the Spark Standalone scheduler, a \textit{leader} (a.k.a master) coordinates the resource provisioned by \textit{workers} in the cluster.
 A \textit{driver}  process receives jobs from clients and requests workers from the leader.

### 2021-09-02 latex/paper.tex
First final draft; minus abstract and acknoledgments.

@@
 \section{Introduction}
-The rise in data sharing coupled with improved data collection technologies lead neuroimaging into the Big Data era\cite{ALFAROALMAGRO:18, van2014human}. \MD{cite CONP portal}
-While common neuroimaging workflow engines, such as Nipype\cite{Nipype:11}, are well rounded to process standard compute-intensive pipeline,
-they lack Big Data strategies (i.e. in-memory computing, data locality, and lazy-evaluation) to improve the performance of increasingly prevalent data-intensive pipelines.
-A previous study\cite{hayot2019performance} noted the importantce of in-memory computing and data locality to improve the performance of data-intensive pipelines.
-In this work, we study how performance benchmarks generalize accross Big Data engines in an HPC environement.
-In particular, we study the detailed performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16}, for their suitability to process neuroimaging pipelines.
-Our objective is to assess whether Spark or Dask has substantial performance benefits to process data-intensive pipeline in HPC clusters.
+The rise in data sharing coupled with improved data collection technologies leads neuroimaging into the Big Data era\cite{ALFAROALMAGRO:18, van2014human, ConpPortal}.
+While standard neuroimaging workflow engines, such as Nipype\cite{Nipype:11}, are well rounded to process standard compute-intensive pipelines,
+they lack Big Data strategies (i.e., in-memory computing, data locality, and lazy-evaluation) to improve the performance of increasingly prevalent data-intensive pipelines.
+A previous study\cite{hayot2019performance} noted the importance of in-memory computing and data locality to improve the performance of data-intensive pipelines.
+In this work, we study how performance benchmarks generalize across Big Data engines in an HPC environment.
+In particular, we study in-depth the performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16} for their suitability to process neuroimaging pipelines.
+Our objective is to assess whether Spark or Dask has substantial performance benefits in processing data-intensive HPC clusters pipelines.
 
-Spark and Dask both provide in-memory computing, data locality, and lazy-evaluation, which is common for Big Data engines.
-Both of their scheduler operate dynamically, which benefit applications with runtime unknown ahead of time\cite{Dask:15}.
-They also provide rich high-level API and support a variety of schedulers and deployment environment, such as Mesos\cite{hindman2011mesos}, YARN\cite{vavilapalli2013apache}, Kubernetes, and HPC clusters.
+Spark and Dask both provide in-memory computing, data locality, and lazy-evaluation, typical for Big Data engines.
+The scheduler for both engines operates dynamically, which benefits applications with runtime unknown ahead of time\cite{Dask:15}.
+They also provide rich high-level API and support a variety of schedulers and deployment environments, such as Mesos\cite{hindman2011mesos}, YARN\cite{vavilapalli2013apache}, Kubernetes, and HPC clusters.
 Although sharing similarities, these engines have differences.
@@
 First and foremost, Spark is written in Scala while Dask is in Python.
-Given the popularity of Python in the scientific community, this arguably gives a edge to Dask due to the serialization cost between Python and Scala.
-However, if not careful, the Python Global Intepreter Lock (GIL) can significatly reduce the parallelism of an application, in some cases.
-The difference in programmming language also provides different benefits.
+The popularity of Python in the scientific community arguably gives an edge to Dask due to the serialization cost between Python and Scala.
+However, if not careful, the Python Global Interpreter Lock (GIL) can significantly reduce the parallelism of an application in some cases.
+The difference in programming languages also provides different benefits.
 On the one hand, as part of the Scipy ecosystem, Dask provides almost transparent integration with APIs such as Numpy arrays, Pandas dataframe, or RAPIDS GPU accelerated code framework.
-On the other hand, Spark's Java, Scala, R, and Python APIs allow to easily parallelize pipelines with minimal performance loss, thanks to the Java Virtual Machine \TG{you could add a few words to explain why the JVM is useful for parallelization ("thanks to the JVM that ...")}.
-Our study is restrained to performance, although, we recognize that other factors affect the choice of an engine in practice.
+On the other hand, Spark's Java, Scala, R, and Python APIs efficiently parallelize pipelines with minimal performance loss, thanks to the Java Virtual Machine.
+\TG{you could add a few words to explain why the JVM is useful for parallelization ("thanks to the JVM that ...")}
+Our study is restrained to performance, although we recognize that other factors affect the choice of an engine in practice.
 
@@
 We installed the Slurm scheduler and Lustre file system~\cite{braam2019lustre} on a dedicated cluster to mimic a typical HPC system.
-Lustre is a powerful and scalable file system for HPC environement offering parallel file accesses.
+Lustre is a powerful and scalable file system for HPC environments offering parallel file accesses.
 
-In a previous study\cite{8943502} we found that their was no significant performance difference between Apache Spark and Dask for data-intensive application.
-However, due to the use of the network file-system (NFS) the performance of the applications was limited by an I/O bottleneck.
+In a previous study\cite{8943502} we found that there was no significant performance difference between Apache Spark and Dask for data-intensive applications.
+However, using a network file system (NFS) led to an I/O bottleneck limiting the performance of the applications.
 Another study \cite{8588652} also reported network and I/O bottlenecks when concurrency increased.
-This same study found that the startup overhead time for Spark was larger than for Dask.
-On the contrary, the work in~\cite{Mehta:17} claims that the startup overhead is larger for Dask than for Spark.
-This suggests that the startup overhead might vary depending on cluster configuration.
+This same study found that the startup overhead time for Spark was more prominent than for Dask.
+On the contrary, the work in~\cite{Mehta:17} claims that the startup overhead is more significant for Dask than for Spark.
+This contradiction suggests that the startup overhead might vary depending on cluster configuration.
 The latter study processed a \SI{100}{\giga\byte} dataset using a neuroimaging pipeline and found Dask to be up to 14\% faster than Spark due to ``more efficient pipelining" and serialization time to Python.
 On the contrary, the work in \cite{10.1145/3225058.3225128} shows that with large datasets, Spark provides better speedup than Dask.
-However, that study used the Dask Bag API which is know to have performance limitation compared to other Dask APIs.
-The above studies compared the engines performance at a high-level.
+However, that study used the Dask Bag API, which has performance limitations compared to other Dask APIs.
+
+The above studies compared the performance of the engine at a high level.
 In comparison, our study provides a detailed analysis of the performance differences and similarities between Dask and Spark for data-intensive applications.
 
-The next section introduce the Lustre file system and the Big Data engines.
+The following section introduces the Lustre file system and the Big Data engines.
 Next, we present the design of our benchmarking experiments.
-We consider two types of data-intensive neuroimaging pipelines: high-resolution imaging and large functional MRI studies.
-The pipelines we chose involve different compute and I/O patterns; e.g. map-reduce or map-only and requiring partial data in memory or the whole dataset.
+We consider two types of data-intensive neuroimaging pipelines: high-resolution imaging and functional MRI studies for a cohort of subjects.
+The pipelines we chose involve different compute and I/O patterns, e.g., map-reduce or map-only, requiring partial data in memory or the whole dataset.
 Further sections present our results, discussion, and conclusion.
@@
 \subsection{Lustre}
-The Lustre file-system is a parallel file system well suited for HPC environments.
-Lustre offers a POSIX compliant interface with the file system.
-Installed in some of the most powerful supercomputer, Lustre is designed to scale to thousands of nodes, petabytes of storage, and hundreds of gigabits per second of I/O throughput.
+The Lustre file system is a parallel file system well suited for HPC environments.
+Lustre offers a POSIX-compliant interface with the file system.
+Installed in some of the most powerful supercomputers, the design of Lustre allows
+scaling to thousands of nodes, petabytes of storage, and hundreds of gigabits per second of I/O throughput.
 
 At the core of Lustre are the \textit{management servers} (MGSs) and \textit{metatdata server} (MDSs) in charge of the files operations and metadata,
-the \textit{object storage targets} (OSTs) to store file data, and the \textit{obejct sotrage server} (OSSs) to handle the OSTs.
+the \textit{object storage targets} (OSTs) to store file data, and the \textit{object sotrage server} (OSSs) to handle the OSTs.
 Lustre clients make requests to the MGSs to access files on the OSTs.
@@
 Dask is a Python-based Big Data engine with growing popularity in the scientific Python ecosystem.
-Dask was designed with data locality and in-memory computing in mind, to mitigate the data transfer bottleneck in Big Data workflows.
+Dask was designed with data locality and in-memory computing in mind to mitigate the data transfer bottleneck in Big Data workflows.
 Data locality, popularized by Map-Reduce \cite{dean2008mapreduce}, schedules tasks where the data reside.
-In-memory computing minimizes the overhead of transferring data to disk by keeping data in memory when possible.
+In-memory computing minimizes the overhead of transferring data to disk sby keeping data in memory when possible.
 Dask uses lazy evaluation to reduce unnecessary communication and computation.
@@
 Dask workflows can further reduce data transfer by leveraging multithreading whenever the Python GIL does not restrict it.
-Fault-tolerance is achieved by recording data lineage, the sequence of operations used to modify the initial data.
+It achieves Fault tolerance by recording data lineage, the sequence of operations used to modify the initial data.
 
@@
 The Dask engine is compatible with multiple distributed schedulers, including YARN and Mesos.
-Dask also provides its own \textit{Dask Distributed scheduler}.
+It also provides a \textit{Dask Distributed scheduler}.
 We chose to use Dask Distributed scheduler to keep the environment balanced between the engines.
@@
 In the Dask Distributed scheduler, a \textit{dask-scheduler} process administrates the resource provided by \textit{dask-worker}s in the cluster.
-The scheduler receives jobs from clients and assigns tasks to available workers using a LIFO (Last-In-First-Out) job scheduling policy.
-That is an utter process branch of the Dask graph before proceeding with the next one \TG{No comprendo, please revise syntax :)}.
+The scheduler receives jobs from clients and assigns tasks to available workers using a global FIFO (First-In-First-Out) job scheduling policy.
+However, the Dask scheduler attempts to complete immediate task dependencies using a LIFO (Last-in-First-Out) policy to reduce the memory footprint of the workers.
 
@@
 YARN is mainly used to schedule Hadoop-based workflows, while Mesos can be used for various workflows.
-We limit our focus to Spark Standalone scheduler, as researchers are likely to execute their workflows in an HPC environment where, usually, neither YARN nor Mesos is available.
+We limit our focus to Spark Standalone scheduler, as researchers are likely to 
+execute their workflows in an HPC environment where neither YARN nor Mesos is usually available.
 
@@
 A \textit{driver}  process receives jobs from clients and requests workers from the leader.
-Jobs are divided into stages to be executed onto workers.
-Each operation in a stage is represented by a high-level task in the computation graph.
+Jobs are divided into stages to execute onto workers.
+Each operation in a stage is represents by a high-level task in the computation graph.
 Like Dask, Spark Standalone scheduler uses a LIFO policy to schedule tasks.
@@
 RDDs are the basis of the other Spark data structure: Datasets and DataFrames.
-Datasets are similar to RDD but benefit additional performance by leveraging the Spark SQL's optimized execution engine. 
-The DataFrames are Datasets organized into named-columns and are used to process tabular data. 
+Datasets are similar to RDD but benefit additional performance by leveraging Spark SQL's optimized execution engine. 
+The DataFrames are Datasets organized into named columns and are used to process tabular data. 
 While the DataFrame API is available in all supported languages, Datasets are limited to Scala and Java.
@@
 While serialization from Python to Java, an operation required when using Spark's Python API, creates overhead, we found it minimal \cite{8943502}.
-We focus on PySpark API to have a more balanced environment between the different engines and for its suitability to neuroimaging.
+We focus on PySpark API to have a more balanced environment between the different engines and its suitability for neuroimaging.
 
@@
 We used the ``slashbin" cluster at Concordia University.
-The cluster has 8 compute nodes, 4 storage nodes, 1 login node, and 1
-Lustre metadata node. Each
-compute node has 2 $\times$ 16-core Intel(R) Xeon(R) Gold 6130 CPU
-@ 2.10GHz, 256~GiB of RAM, 6 $\times$ SSDs of 450~GiB each with the XFS
-file system (no RAID enabled), 378~GiB of tmpfs, 126~GiB of devtmpfs file system,
-CentOS~8.1 and Linux kernel
-\textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
+The cluster has eight compute nodes, four storage nodes, one login node, and one Lustre metadata node.
+Each compute node has two 16-core Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz,
+\SI{250}{\gibi\byte} of RAM, \SI{126}{\gibi\byte} of tmpfs,
+% 6 $\times$ SSDs of \SI{447}{\gibi\byte} each with the XFS file system (no RAID enabled),
+CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
 
-\TG{describe the Lustre installation: data nodes and lustre config}
-\TG{Use GB or GiB consistently}
+We configured Lustre to have 1 MGS and 4 OSS.
+Each OSS has 11 OSTs (disks) of \SI{8.8}{\tebi\byte}.
+The OSTs are configured to have a maximum of \SI{1000}{\mebi\byte} of dirty data written into the client page cache.
+Both the MDT and OSTs are configured to a maximum of 64 concurrent RPCs in flight.
+A timeout of \SI{100}{\second} is set for Lustre, and OSTs checksum are disabled.
+Lustre \textit{2.14.0-1.el8} is installed on both the Lustre clients and servers.
 
-Both Spark and Dask were configured to have 8 worker processes each with 8
-threads. Each worker was allocated \SI{31.5}{\giga\byte} of memory \TG{,
-	determined as the amount of memory that ...}. \TG{any other Spark or Dask
-param? Location of log files, etc} A new Dask or Spark cluster was started and
-teared down for each experiment.
-
-Dask ?? and Spark ?? \MD{Add version when after locking it for experiments.} was used for our experiments.
+Both Spark and Dask are configured to have eight worker processes, each with eight threads.
+Each worker was allocated \SI{31.5}{\gibi\byte} of memory.
+For Spark, the JVM heap space uses 10\% () of workers' memory, and the location of the log is set on an NFS.
+The worker workspace for both engines is located on Lustre due to the limited fast storage space on the worker nodes.
+Other Dask configurations are left to default.
+The Spark network timeout is set to \SI{600}{\second} and the executor heartbeat to \SI{120}{\second}
+A new Dask or Spark cluster was started and tore down for each experiment.
+Dask \textit{2021.7.0} and Spark \textit{3.1.2} was used for our experiments.
 
@@
 NIfTI format, a popular format in neuroimaging. We left the NIfTI blocks
-uncompressed, resulting in a total data size of \SI{648}{\giga\byte}. To
+uncompressed, resulting in total data size of \SI{606}{\gibi\byte}. To
 evaluate the effect of block size, we split the image into 1000, 2500,
-and 5000 files of \SI{648}{\mega\byte}, \SI{259.2}{\mega\byte}, and
-\SI{129.6}{\mega\byte}, respectively. The
-\href{https://github.com/big-data-lab-team/sam}{sam} library was used to
-split the image.
+and 5000 files of \SI{606}{\mebi\byte}, \SI{242.4}{\mebi\byte}, and
+\SI{121.2}{\mebi\byte}, respectively.
+We used the \href{https://github.com/big-data-lab-team/sam}{sam} library to split the image.
 
@@
 \cite{zuo2014open}, freely available on
-\href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}. The
-entire dataset is \SI{408.4}{\giga\byte}, containing anatomical, diffusion
-and functional images of 1,397 subjects acquired in 29 sites. We used all
-3,491 anatomical images, representing \SI{39}{\giga\byte} overall
-(\SI{11.17}{\mega\byte} per image on average).
+\href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}.
+The entire dataset is \SI{379.83}{\gibi\byte}, containing anatomical, diffusion,
+and functional images of 1,654 subjects acquired in 35 sites.
+We used all 3,491 anatomical images, representing \SI{26.67}{\gibi\byte} overall
+(\SI{7.82}{\mebi\byte} per image on average).
 
@@
 We use three simple synthetic applications to have a deep understanding of the 
-underlying behavior of the applications with simple I/O pattern; namely
+underlying behavior of the applications with simple I/O pattern, namely
 \textit{Increment}, \textit{Multi-Increment}, and \textit{Histogram}.
@@
 To have a fair comparison between the engines, the code base for the core functions of the applications is the same for both engines.
-This allow us to reduce the performance difference from different implementations.
-Note that this is not the case for the \textit{Kmeans} application as the best performing API for Dask and Spark differed largely in term of implementation.
+This allows us to reduce the performance difference from different implementations.
 
@@
 mimics the processing of multiple independent subjects in parallel.
-
 \begin{figure}[!ht]
@@
 
-% Currently ommitted due to issue with memory.
-% 
-% \subsubsection{Multi-Increment}
-% Our second application is an adaptation of the increment application. A
-% significant difference is that, at each iteration, it uses a random
-% BigBrain block as the increment value. This change allows the
-% multi-increment application to have inter-worker communication while
-% remaining simple.
-
-% \begin{figure}[!hb]
-% 	\centering
-% 	\includegraphics[height=\columnwidth,
-% 	angle=0]{figures/multi-increment.png}
-% 	\caption{Task graph for Multi-Incrementation with 3 iterations and 3 BigBrain blocks.}
-% 	\label{fig:graph-muti-increment}
-% \end{figure}
-
 \subsubsection{Histogram}
@@
 The amount of inter-worker communication is in-between the increment and multi-increment applications.
-
 \begin{figure}[!hb]
@@
 communication between the workers.
-
-For this application, the Spark and Dask implementations differ slightly,
-to take advantage of the best-suited API from both engines. The Spark
-implementation uses the Map-Reduce paradigm, while the Dask one uses array
-programming.
+The computation to find the closest centroids for each voxels scales with the number of centroids.
+It can be computed in parallel at the cost of higher memory or sequentially to reduce memory usage.
+We decided to compute the nearest centroids in parallel since we use only three centroids.
 
 \subsubsection{BIDS App example}
-Our fifth application is BIDS App example, a neuroimaging pipeline to
+Our fifth application is the BIDS App example, a neuroimaging pipeline to
 measure the brain volume from MRIs. For this application, we use the CoRR
@@
 then computes the average for each group of participants. Unlike the other
-applications, BIDS App example is a command-line executed in a Docker image
+applications, the BIDS App example is a command-line executed in a Docker image
 (bids/example on DockerHub). We converted the Docker image to a Singularity
-image for use in HPC environments, \MD{Cite paper on reason why this is
+image for use in HPC environments, \MD{Cite paper on the reason why this is
 done.} using
 \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
-
-\TG{You should explain how the applications were implemented in Dask and
-Spark which APIs were used and why, etc}
 
@@
 throughout the experiments. We varied (1) the number of workers to assess
-the scalibility of the engine scheduler, (2) the BigBrain
+the scalability of the engine scheduler, (2) the BigBrain
 block size in Increment, Multi-Increment, and Histogram to measure the
 effect of different I/O patterns and parallelization degrees, (3) the
-number of iterations to evaluate the effect of number of task, and (4) the
+number of iterations to evaluate the effect of the number of tasks, and (4) the
 sleep delay to study the effect of task duration. It should note that
@@
 
-To avoid potential external biases due to caching, background processes and
+To avoid potential external biases due to caching, background processes and,
 network load, we ran the applications in randomized order and cleared the
-page cache in between every experiments. Each benchmark was run ten times.
+page cache between every experiment.
+Moreover, we ran each benchmark ten times.
 
-For each run, we measured the makespan of the application as well as the
-cumulative time spent in the different functions for read, processing, and
-writing data. The overhead calculation for each CPU thread is the end time
-of the last processed task minus the total runtime of the tasks ran for
-this thread \TG{unclear definition. Don't refer to CPU threads as a given task may be executed on more than 1. Maybe refer to engine executors?}. Summing those results gives the total overhead for the
-application.
+For each execution, we measured the time spent in the different functions of the
+application to read, process, and write data.
+We calculate the overhead for each CPU thread by subtracting the total time
+spent in the reading, writing, and processing functions to the makespan of the application.
+Summing those results gives the total overhead for the application.
 
-%  TODO Fix table for the final parameters/applications used.
 \begin{table*}[t]
@@
 	\begin{tabular}{|l|c|c|c|c|c|}
-		\hline & Increment & Multi-Increment & Kmeans & Histogram & BIDS App Example \\\hline
-		\# of Nodes & \multicolumn{5}{c|}{2, 4, 8} \\\hline
-		Dataset size [\SI{}{\giga\byte}] & 648 &\multicolumn{2}{c|}{486} & 648 & \multicolumn{1}{c|}{39} \\\hline
-		Number of file & \multicolumn{4}{c|}{1000, 2500, 5000}  & \multicolumn{1}{c|}{3491} \\\hline
-		\# of Iterations & 1, 5, 25 & \multicolumn{2}{c|}{1, 3, 9}                 & \multicolumn{2}{c|}{-} \\\hline
+		\hline           & Increment & Histogram & Kmeans  & BIDS App Example \\\hline
+		\# of Nodes & \multicolumn{4}{c|}{2, 4, 8} \\\hline
+		Dataset size [\SI{}{\gibi\byte}] &\multicolumn{2}{c|}{606} & 303 & \multicolumn{1}{c|}{26.67} \\\hline
+		Number of file & \multicolumn{2}{c|}{1000, 2500, 5000} & - & \multicolumn{1}{c|}{3491} \\\hline
+		\# of Iterations & 1, 5, 25  & -         & 1, 3, 9 & -                \\\hline
 	\end{tabular}
@@
 Figure \ref{fig:increment_worker} depicts the total execution time spent in each code segment of the \textit{Increment} application.
-The bars are broken down into Idle, Read, Write, and other application specific functions.
-The mean of each bar was calulated from 10 repetitions, with the error bar representing the standard deviation.
+The bars are broken down into Idle, Read, Write, and other application-specific functions.
+The mean of each bar is calculated from ten repetitions, with the error bar representing the standard deviation.
 As expected, the compute time, from \textit{Increment}, remains the same when we vary the number of nodes.
-However, the \textit{Load} and \textit{Dump} time increase with the number of nodes for both engines.
+However, the \textit{Read} and \textit{Write} time increase with the number of nodes for both engines.
 This is explained by an I/O bottleneck on the Lustre file system.
@@
 While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
-Althought Dask has higher overhead, the total time for both engine is similar; with a maximum makespan difference of \SI{7}{\second} for 2 nodes.
+Although Dask has a higher overhead, the total time for both engines is similar; with a maximum makespan difference of \SI{7}{\second} for two nodes.
 This is due to a tradeoff between disk bandwidth and compute/idle time.
-When a multiple threads access Lustre concurrently, they shared the disk bandwitdth resulting in longer I/O time.
-However, when most threads are either idle or performing computations, it reduce the contention on the disk bandwitdh.
-\begin{figure}[!h]
+On the one hand, multiple threads access Lustre concurrently, thus sharing the disk bandwidth resulting in longer I/O time.
+On the other hand, when most threads are either idle or performing computations, it reduces the contention on the disk bandwidth resulting in shorter I/O time.
+\begin{figure}[!t]
 	\centering
@@
 As expected, the idle time for both engines increases with the number of iterations.
-However, idle time for Dask increase significantly faster that Spark; resulting in a \SI{38}{\second} makespan difference for 25 iterations.
+However, idle time for Dask increases significantly faster than Spark, resulting in a \SI{38}{\second} makespan difference for 25 iterations.
 
-Figure \ref{fig:increment_itr} also depicts that as the total computation time increases the total I/O time decreases.
+Figure \ref{fig:increment_itr} also depicts that the total computation time increases as the total I/O time decreases.
 This is explained by an I/O bottleneck on the Lustre file system.
-When threads are busy with computations they reduce the contention on Lustre, thus improving I/O speed for individual accesses.
-%  # of iterations
-\begin{figure}[!h]
+When threads are busy with computations, they reduce the contention on Lustre, thus improving I/O speed for individual accesses.
+\begin{figure}[!t]
 	\centering
@@
 Figure \ref{fig:increment_block} shows an increase in \textit{Increment} time.
-This is because the increment is performed on each file.
-Thus increasing the number of file increase the computation time.
-Similarly to Fig. \ref{fig:increment_itr} increasing the total compute time of the application reduce the total I/O time.
-Again, this is explain by the I/O tradeoff between compute time and contention of disk bandwidth.
+This is because the application performs an incrementation for each file, thus increasing the computation time.
+Similarly to Figure \ref{fig:increment_itr}, increasing the total compute time of the application reduces the total I/O time.
+Again, this is explained by the I/O tradeoff between compute time and contention of disk bandwidth.
 
 From figure \ref{fig:increment_block} we also observe that Dask has significantly higher idle time than Spark.
-However, the total of the application is similar due to the tradeoff between I/O and compute/idle time.
-
-% # of files
-\begin{figure}[!h]
+However, the total execution time of the application is similar due to the tradeoff between I/O and compute/idle time.
+\begin{figure}[!t]
 	\centering
@@
 
-% \subsection{Multi-Increment}
-% TODO: experiment
-% Memory issue with Spark (Heap space runs out of memory)
-
-\subsection{Kmeans}
-
 \subsection{Histogram}
 Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} when varying the number of nodes.
-Overall, Spark has upt to 5\% lower total execution time with the difference comming from a shorter execution time for \textit{Calculate\_histogram}.
-This is unexpected as the code base for both engine is the same; with only the calling APIs differing.
-% This happened for both Bags and Delayed. There might be a higher serialization cost.
-As with previous experiments, the overhead for Spark is less and scale better, although, compensated by a longer total I/O time.
-% # of nodes
-\begin{figure}[!h]
+Overall, Spark has up to 5\% lower total execution time, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
+This is unexpected as the code base for both engines is the same, with only the calling APIs differing.
+It is most likely due to the use of Dask Bags, which are slower than the other Dask APIs.
+As with previous experiments, the overhead for Spark is less and scale better, although compensated by a longer total I/O time.
+\begin{figure}[!t]
 	\centering
@@
 
-Figure \ref{fig:histogram_worker} shows the execution time for the \textit{Histogram} when varying the number of files.
+Figure \ref{fig:histogram_block} shows the execution time for the \textit{Histogram} when varying the number of files.
 The results are similar to varying the number of nodes with Spark having faster compute time for \textit{Calculate\_histogram}.
-However, the makespan difference is much smaller due to larger I/O time for both engine.
-%  # of files
-\begin{figure}[!h]
+\begin{figure}[!t]
 	\centering
@@
 
+\subsection{Kmeans}
+Figure \ref{fig:kmeans_worker} depicts the total execution time for the \textit{Kmeans} application when varying the number of workers.
+We observe that Dask is multifold faster than Spark.
+Specifically, Spark has much longer idle, \textit{get\_labels} and read time than Dask.
+This difference arises from Spark workers restarting due to lack of executor memory, leading to recomputation of tasks.
+
+For Spark, we reduced the number of processes per node from 64 to 32 while keeping the memory the same and found similar execution times.
+However, the Spark worker did not restart with a lower process count and no longer recomputed tasks.
+
+We omit the figure when varying the number of iterations as the results are similar to varying the number of workers.
+\begin{figure}[!t]
+	\centering
+	\includegraphics[clip,width=\columnwidth]{figures/stacked_kmeans_worker.jpg}
+	\caption{Kmeans: varying nodes -- 5000 files, 3 iterations, \SI{1}{\second} delay}
+	\label{fig:kmeans_worker}
+\end{figure}
+
 \subsection{BIDS App Example}
 Figure \ref{fig:bids} shows the total execution time for the \textit{BIDS App example} application when varying the number of nodes.
-We observe that Dask has significantly lower execution time than Spark; between 10\% to 15\%.
-The main difference comes from a increased amount of stagger tasks for Spark, thus increasing the idle time of the application.
-Moreover, a shorther cluster deployement time for Dask reinforce the difference in idle time.
+We observe that Dask has a significantly lower execution time than Spark, between 10\% to 15\%.
+The main difference comes from an increased amount of stagger tasks for Spark, thus increasing the idle time of the application.
+Moreover, although minimal, a shorter cluster deployment time for Dask reinforces the difference in idle time.
 As with other experiments, the I/O from Lustre affects the scaling of both engines as we increase the number of nodes.
-\begin{figure}[!h]
+\begin{figure}[!t]
 	\centering
@@
 \section{Discussion}
-% TODO
+\subsection{I/O bottleneck}
+We used Lustre for our share file system as it is a standard solution in HPCs.
+The performance of Lustre degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
+In a previous study \cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
+In this paper, Lustre shows some scaling when increasing the number of nodes, although not linear.
+This suggests that increasing the parallelism level of the file system might help alleviate the I/O bottleneck observed.
+
+As in our previous study \cite{8943502}, we observe that the I/O bottleneck from the file system has a lower impact on more compute-intensive applications, such as BIDS App Example.
+However, it significantly affects data-intensive applications, such as Increment and Histogram.
+
+These results suggest that increasing the parallelism of the file system is critical when scaling the compute resources to process a data-intensive application.
+Otherwise, the compute resource might be underutilized.
+
+\subsection{Tradeoff between overhead and data transfer}
+Overall, except for Kmeans, the overhead for Dask is larger than for Spark.
+However, the total execution time for both engines is comparable.
+This almost exact compensation between the engine overhead and the data transfer time (see for example Figure \ref{fig:increment_worker}) is not surprising.
+When there is high contention on the bandwidth of the file system, and in the absence of trashing, desynchronizing the file transfers generally does not reduce the total data transfer time.
+To take an extreme example, the makespan of $n$ concurrent file transfer of size $F$ from a file system with a bandwidth $B$ would be $nF/B$,
+as the $n$ transfer would share the bandwidth $B$, which is equivalent to sequentially transfer the $n$ files.
+i.e., when the bandwidth of the file system is saturated, engine overhead does not increase the makespan as long as it desynchronizes the data transfers.
+In a previous study \cite{8943502}, we also observed this tradeoff between overhead and data transfer when the file system was saturated.
+It could be interesting to explore strategies that take advantage of this tradeoff to reduce the application makespan.
+
+\subsection{Memory management}
+Our Kmeans experiment shows a significant difference between Dask and Spark.
+The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks
+This is unexpected, given that Dask and Spark are configured similarly and use the same amount of resources.
+We think those issues arise from a combination of
+(1) the reserved memory for the JVM heap space,
+(2) the serialization of Python to Java kept in memory,
+(3) Spark only using process compared to Dask using threads, thus limiting shared memory,
+and (4) different policies to persist data to disk when running out of memory.
+
+When reducing the number of processes from 64 to 32 for Spark (effectively doubling the memory of each process), the total execution time was similar.
+However, with only 32, and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
+We suppose that given more memory per thread/process, Dask and Spark would then perform similarly.
+
+\subsection{Comparison to other studies}
+Overall, except for Kmeans, our experiments did not show substantial performance differences between Spark and Dask.
+In contradiction with the experiments of \cite{Mehta:17}, we found that Dask had a larger overhead than Spark, especially when scaling the number of workers.
+However, this was compensated by the overhead and data transfer tradeoff previously.
+Since our focus is on data-intensive applications, we conclude that the performance of the engine is equivalent.
+
+The experiments from \cite{Mehta:17} found that Dask has a larger startup overhead than Spark, while another study \cite{8588652} found the opposite.
+We did not find significant differences in startup overhead between the engines in this study nor our previous study \cite{8943502}.
+Those apparent contractions most likely come from differences in infrastructure and configuration of the engines.
+
+Our results from the Histogram experiments show that the Dask Bags was slower than Spark.
+The experiments from \cite{10.1145/3225058.3225128} also found that Dask Bags performed slower than Spark.
+We conclude that when using Dask, the other Dask APIs should be favored unless Dask Bags are strictly required.
+
+Our results showed that, for data-intensive applications, increasing the number of workers should only be done if the bandwidth of the file system is not saturated.
+Two other studies support this \cite{8943502, 8588652}.
+Therefore, it is critical to assure enough bandwidth to the file system before scaling the compute resources for data-intensive applications.
+Otherwise, those extras resources might be underused.
 
 \section{Conclusion}
-% TODO
+We present a detailed comparison of two Big Data engines, Apache Spark and Dask.
+We studied the engines using three synthetic data-intensive applications and a data-intensive neuroimaging pipeline.
+Overall, our results showed no substantial difference between the engines.
+Although Spark overhead is lower than Dask, Dask performs I/O faster due to a tradeoff between overhead and I/O when the file system is saturated.
+For one application, our results showed that Spark used more memory, causing issues with its workers resulting in a significantly faster runtime for Dask.
+However, using an infrastructure offering more memory per worker would most likely solve this issue.
+
+These results suggest that future research should focus on strategies to:
+\begin{itemize}
+	\item Limit the impact of data transfer for data-intensive applications
+	\item Reduce the footprint of application memory
+	\item Manage the memory of workers more efficiently
+\end{itemize}
 
@@
 % TODO
+Mathieu Dugre\'e was funded by an Alexander Graham Bell scholarship from the Natural Sciences and Engineering Research Council of Canada.
+\MD{Add other things here..?}
 

### 2021-09-02 latex/paper.tex
Typos and minor comments

@@
 In particular, we study in-depth the performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16} for their suitability to process neuroimaging pipelines.
-Our objective is to assess whether Spark or Dask has substantial performance benefits in processing data-intensive HPC clusters pipelines.
+Our objective is to assess whether Spark or Dask has substantial performance benefits in processing data-intensive HPC pipelines.
 
@@
 On the one hand, as part of the Scipy ecosystem, Dask provides almost transparent integration with APIs such as Numpy arrays, Pandas dataframe, or RAPIDS GPU accelerated code framework.
-On the other hand, Spark's Java, Scala, R, and Python APIs efficiently parallelize pipelines with minimal performance loss, thanks to the Java Virtual Machine.
+On the other hand, Spark's Java, Scala, R, and Python APIs efficiently parallelize pipelines with minimal performance loss, due to the Java Virtual Machine.
 \TG{you could add a few words to explain why the JVM is useful for parallelization ("thanks to the JVM that ...")}
@@
 In a previous study\cite{8943502} we found that there was no significant performance difference between Apache Spark and Dask for data-intensive applications.
-However, using a network file system (NFS) led to an I/O bottleneck limiting the performance of the applications.
+However, using the network file system (NFS) led to an I/O bottleneck limiting application performance.
 Another study \cite{8588652} also reported network and I/O bottlenecks when concurrency increased.
@@
 
-At the core of Lustre are the \textit{management servers} (MGSs) and \textit{metatdata server} (MDSs) in charge of the files operations and metadata,
+At the core of Lustre are the \textit{management servers} (MGSs) and \textit{metatdata server} (MDSs) in charge of the file operations and metadata,
 the \textit{object storage targets} (OSTs) to store file data, and the \textit{object sotrage server} (OSSs) to handle the OSTs.
@@
 Data locality, popularized by Map-Reduce \cite{dean2008mapreduce}, schedules tasks where the data reside.
-In-memory computing minimizes the overhead of transferring data to disk sby keeping data in memory when possible.
+In-memory computing minimizes the overhead of transferring data to disk by keeping data in memory when possible.
 Dask uses lazy evaluation to reduce unnecessary communication and computation.
 The engine builds a dynamic graph before execution, allowing it to determine which task to compute.
-Dask workflows can further reduce data transfer by leveraging multithreading whenever the Python GIL does not restrict it.
-It achieves Fault tolerance by recording data lineage, the sequence of operations used to modify the initial data.
+Dask workflows can further reduce data transfers by leveraging multithreading whenever the Python GIL does not restrict it.
+It achieves fault tolerance by recording data lineage, the sequence of operations used to modify the initial data.
 
@@
 The Dask engine is compatible with multiple distributed schedulers, including YARN and Mesos.
-It also provides a \textit{Dask Distributed scheduler}.
-We chose to use Dask Distributed scheduler to keep the environment balanced between the engines.
+It also provides its own scheduler, \textit{Dask Distributed scheduler}.
+We chose to use Dask Distributed scheduler to keep the environments similar between Dask and Spark.
 
@@
 
-Spark's primary data structure is Resilient Distributed Dataset (RDD)\cite{RDD}, a fault-tolerant, parallel collection of data elements.
-RDDs are the basis of the other Spark data structure: Datasets and DataFrames.
-Datasets are similar to RDD but benefit additional performance by leveraging Spark SQL's optimized execution engine. 
-The DataFrames are Datasets organized into named columns and are used to process tabular data. 
+Spark's primary data structure is the Resilient Distributed Dataset (RDD)\cite{RDD}, a fault-tolerant, parallel collection of data elements.
+RDDs are the basis of other Spark data structures: Datasets and DataFrames.
+Datasets are similar to RDDs but benefit additional performance by leveraging Spark SQL's optimized execution engine. 
+DataFrames are Datasets organized into named columns and are used to process tabular data. 
 While the DataFrame API is available in all supported languages, Datasets are limited to Scala and Java.
@@
 Python is a standard programming language in the scientific community, offering numerous data processing libraries.
-While serialization from Python to Java, an operation required when using Spark's Python API, creates overhead, we found it minimal \cite{8943502}.
-We focus on PySpark API to have a more balanced environment between the different engines and its suitability for neuroimaging.
+While serialization from Python to Java, an operation required when using Spark's Python API, creates overhead, we found it minimal in a previous study~\cite{8943502}.
+We focus on PySpark API to have a similar environment between the different engines and due to its suitability for neuroimaging.
 
@@
 Each worker was allocated \SI{31.5}{\gibi\byte} of memory.
-For Spark, the JVM heap space uses 10\% () of workers' memory, and the location of the log is set on an NFS.
+For Spark, the JVM heap space uses 10\% of workers' memory, and the location of the log is set on an NFS-mounted directory.
 The worker workspace for both engines is located on Lustre due to the limited fast storage space on the worker nodes.
 Other Dask configurations are left to default.
-The Spark network timeout is set to \SI{600}{\second} and the executor heartbeat to \SI{120}{\second}
+The Spark network timeout is set to \SI{600}{\second} and the executor heartbeat to \SI{120}{\second}.
 A new Dask or Spark cluster was started and tore down for each experiment.
-Dask \textit{2021.7.0} and Spark \textit{3.1.2} was used for our experiments.
+Dask \textit{2021.7.0} and Spark \textit{3.1.2} were used.
 
@@
 We use three simple synthetic applications to have a deep understanding of the 
-underlying behavior of the applications with simple I/O pattern, namely
+underlying behavior of the applications with simple I/O patterns, namely
 \textit{Increment}, \textit{Multi-Increment}, and \textit{Histogram}.
 Also, two neuroimaging applications are used to study more realistic 
-applications; namely \textit{Kmeans} and \textit{BIDS App Example}.
+applications, namely \textit{Kmeans} and \textit{BIDS App Example}.
 
@@
 This synthetic application reads blocks of the BigBrain from Lustre and
-simulates computation by sleeping for a specified period. To simulate
+simulates computation by sleeping for a specified period (Fig.~\ref{fig:graph-increment}). To simulate
 intermediate results, we repeat the sleep process for a configurable amount
@@
 As our third application, we calculate the histogram of the BigBrain image. 
-The application reads the BigBrain blocks from Lustre, calculates each intensity's frequency, and then writes the aggregated result back on Lustre.
-This map-reduce application has a very high read overwrite ratio.
+The application reads the BigBrain blocks from Lustre, calculates each intensity's frequency, and then writes the aggregated result back on Lustre (Fig.~\ref{fig:graph-histogram}).
+This map-reduce application has a very high read over write ratio.
 Moreover, this application requires shuffling, albeit of a limited amount of data. 
@@
 Our fifth application is the BIDS App example, a neuroimaging pipeline to
-measure the brain volume from MRIs. For this application, we use the CoRR
+measure brain volume from MRIs. For this application, we use the CoRR
 dataset. The application extracts the brain volume of each participant,
@@
 \subsection{Experiments}
-Table~\ref{table:parameters} shows the parameters that were varied
+Table~\ref{table:parameters} shows the parameters that we varied
 throughout the experiments. We varied (1) the number of workers to assess
@@
 number of iterations to evaluate the effect of the number of tasks, and (4) the
-sleep delay to study the effect of task duration. It should note that
+sleep delay to study the effect of task duration. \TG{these parameters don't seem to match the names used in Table I, they should be harmonized}
+ It should be noted that
 increasing the number of iterations for a given sleep delay also increases
@@
 Figure \ref{fig:increment_worker} depicts the total execution time spent in each code segment of the \textit{Increment} application.
-The bars are broken down into Idle, Read, Write, and other application-specific functions.
+The bars are broken down into Idle, Read, Write, and other application-specific functions. \TG{Add a sentence to explain what Idle is. Is it exactly the overhead mentioned previously? if
+so, I would just call it Overhead everywhere (in the figures and in the text)}
 The mean of each bar is calculated from ten repetitions, with the error bar representing the standard deviation.
@@
 
-Figure \ref{fig:increment_itr} also depicts that the total computation time increases as the total I/O time decreases.
+Figure \ref{fig:increment_itr} also shows that the total computation time increases as the total I/O time decreases.
 This is explained by an I/O bottleneck on the Lustre file system.
-When threads are busy with computations, they reduce the contention on Lustre, thus improving I/O speed for individual accesses.
+When threads are busy with computations, they reduce the contention on Lustre, thus improving I/O speed for individual accesses \TG{but then read or write time should decrease, not computing time.}.
 \begin{figure}[!t]
@@
 
-From figure \ref{fig:increment_block} we also observe that Dask has significantly higher idle time than Spark.
+From Figure \ref{fig:increment_block} we also observe that Dask has significantly higher idle time than Spark.
 However, the total execution time of the application is similar due to the tradeoff between I/O and compute/idle time.
@@
 \subsection{Histogram}
-Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} when varying the number of nodes.
+Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} application when varying the number of nodes.
 Overall, Spark has up to 5\% lower total execution time, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
@@
 It is most likely due to the use of Dask Bags, which are slower than the other Dask APIs.
-As with previous experiments, the overhead for Spark is less and scale better, although compensated by a longer total I/O time.
+As with previous experiments, the overhead for Spark is less and scales better, although compensated by a longer total I/O time.
 \begin{figure}[!t]
@@
 
-For Spark, we reduced the number of processes per node from 64 to 32 while keeping the memory the same and found similar execution times.
-However, the Spark worker did not restart with a lower process count and no longer recomputed tasks.
+% For Spark, we reduced the number of processes per node from 64 to 32 while keeping the memory the same and found similar execution times.
+% However, the Spark worker did not restart with a lower process count and no longer recomputed tasks \TG{removed this as it is discussed later more clearly}.
 
-We omit the figure when varying the number of iterations as the results are similar to varying the number of workers.
+For brevity we omitted the figure when varying the number of iterations as the results are similar to varying the number of workers.
 \begin{figure}[!t]
@@
 
+\TG{Could you add information about how the differences in total times translated into makespans, for each application? Either as sentences in the text or as an extra table.}
+
 \section{Discussion}
 \subsection{I/O bottleneck}
-We used Lustre for our share file system as it is a standard solution in HPCs.
+We used Lustre as shared file system as it is a standard solution in HPC systems.
 The performance of Lustre degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
@@
 Our Kmeans experiment shows a significant difference between Dask and Spark.
-The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks
+The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
 This is unexpected, given that Dask and Spark are configured similarly and use the same amount of resources.
@@
 (2) the serialization of Python to Java kept in memory,
-(3) Spark only using process compared to Dask using threads, thus limiting shared memory,
+(3) Spark only using procesess compared to Dask using threads, thus limiting shared memory consumption,
 and (4) different policies to persist data to disk when running out of memory.
 
-When reducing the number of processes from 64 to 32 for Spark (effectively doubling the memory of each process), the total execution time was similar.
-However, with only 32, and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
+When reducing the number of processes from 64 to 32 for Spark (effectively doubling the memory of each process), the total execution time became similar between Spark and Dask.
+However, with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
 We suppose that given more memory per thread/process, Dask and Spark would then perform similarly.
@@
 In contradiction with the experiments of \cite{Mehta:17}, we found that Dask had a larger overhead than Spark, especially when scaling the number of workers.
-However, this was compensated by the overhead and data transfer tradeoff previously.
-Since our focus is on data-intensive applications, we conclude that the performance of the engine is equivalent.
+However, this was compensated by the overhead and data transfer tradeoff mentioned previously.
+Since our focus is on data-intensive applications, we conclude that the performance of Dask and Spark is equivalent.
 
@@
 
-Our results from the Histogram experiments show that the Dask Bags was slower than Spark.
+Our results from the Histogram experiments show that the Dask Bags API was slower than Spark.
 The experiments from \cite{10.1145/3225058.3225128} also found that Dask Bags performed slower than Spark.
@@
 Two other studies support this \cite{8943502, 8588652}.
-Therefore, it is critical to assure enough bandwidth to the file system before scaling the compute resources for data-intensive applications.
+Therefore, it is critical to ensure enough bandwidth to the file system before scaling the compute resources for data-intensive applications.
 Otherwise, those extras resources might be underused.
@@
 \section{Conclusion}
-We present a detailed comparison of two Big Data engines, Apache Spark and Dask.
+We presented a detailed comparison of two Big Data engines, Apache Spark and Dask.
 We studied the engines using three synthetic data-intensive applications and a data-intensive neuroimaging pipeline.
@@
 \begin{itemize}
-	\item Limit the impact of data transfer for data-intensive applications
+	\item Limit the impact of data transfers for data-intensive applications
 	\item Reduce the footprint of application memory
@@
 % TODO
-Mathieu Dugre\'e was funded by an Alexander Graham Bell scholarship from the Natural Sciences and Engineering Research Council of Canada.
-\MD{Add other things here..?}
+Mathieu Dugre\'e was funded by an Alexander Graham Bell scholarship from
+the Natural Sciences and Engineering Research Council of Canada. This work
+was partially funded by the Canada Research Chairs program. The HPC cluster
+was funded by the Canada Foundation for Innovation.
 

### 2021-09-02 latex/paper.tex
Typos and minor comments

@@
 On the other hand, Spark's Java, Scala, R, and Python APIs efficiently parallelize pipelines with minimal performance loss, due to the Java Virtual Machine.
-\TG{you could add a few words to explain why the JVM is useful for parallelization ("thanks to the JVM that ...")}
 Our study is restrained to performance, although we recognize that other factors affect the choice of an engine in practice.

### 2021-09-02 fix_filename.py
[ADD] utils scripts.

@@ -1,0 +1,7 @@
+from glob import glob
+import os
+
+folders = glob("results/benchmarks/*")
+
+for folder in folders:
+    os.rename(folder, folder.replace("'", ""))

### 2021-09-02 src/mnc2nii.py

@@ -1,0 +1,18 @@
+# Reference: https://gist.github.com/ofgulban/46d5c51ea010611cbb53123bb5906ca9
+import glob
+import os
+import sys
+
+from nibabel import load, save, Nifti1Image
+import numpy as np
+
+input_folder = sys.argv[1]
+
+for filename in glob.glob(input_folder + "/*.mnc"):
+    minc = load(filename)
+    basename = minc.get_filename().split(os.extsep, 1)[0]
+
+    affine = np.array([[0, 0, 1, 0], [0, 1, 0, 0], [1, 0, 0, 0], [0, 0, 0, 1]])
+
+    out = Nifti1Image(minc.get_data(), affine=affine)
+    save(out, basename + ".nii")

### 2021-09-02 plot-script/plot.py
[UPDATE] the plot scripts and the final notebook figures.

@@ -239,9 +239,12 @@ def gantt(
         show(plot)
 
 
-def wasted_time(df, *, framework):
+def wasted_time(df, *, framework, func_remap=None):
     if framework == "spark":
         df["thread"] = df["process"]
+        
+    if func_remap:
+        df = func_remap(df)
 
     core_used = sum(
         [
@@ -264,9 +267,12 @@ def wasted_time(df, *, framework):
     return wasted_time, *run_time
 
 
-def idle_time(df, *, framework):
+def idle_time(df, *, framework, func_remap=None):
     if framework == "spark":
         df["thread"] = df["process"]
+        
+    if func_remap:
+        df = func_remap(df)
 
     by_thread = df.groupby(["worker", "thread"])
 
@@ -295,6 +301,7 @@ def stacked_bar(
     save_name=None,
     ylim=None,
     title=None,
+    func_remap=None,
     **kwargs,
 ):
     matplotlib.rcParams.update({"font.size": 22})
@@ -340,7 +347,7 @@ def stacked_bar(
         data = [
             np.array(
                 [
-                    idle_func(pd.read_csv(file_, names=col_names), framework=framework)
+                    idle_func(pd.read_csv(file_, names=col_names), framework=framework, func_remap=func_remap)
                     for file_ in experiments[key]
                 ]
             )
@@ -431,7 +438,9 @@ def stacked_bar(
 
     func_patches = []
     sample_df = pd.read_csv(filenames[0], names=col_names)
-    for i, function in enumerate(["idle"] + sorted(sample_df["func"].unique())):
+    if func_remap:
+        sample_df = func_remap(sample_df)
+    for i, function in enumerate(["overhead"] + sorted(sample_df["func"].unique())):
         func_patches.append(
             Patch(
                 facecolor="lightgray",
@@ -459,5 +468,100 @@ def stacked_bar(
         os.makedirs(os.path.dirname(save_name), exist_ok=True)
         plt.savefig(save_name, bbox_inches="tight")
         plt.title(title)
+
+    plt.show()
+
+def bar(
+    *,
+    col_names,
+    benchmark_dir,
+    experiment,
+    parameters,
+    xlabel,
+    save_name=None,
+    ylim=None,
+    **kwargs,
+):
+
+    fig, ax = plt.subplots(figsize=(10, 5))
+    bar_width = 0.15
+
+    benchmark_dir += f"/*{experiment}"
+    freedom = None
+    for parameter in parameters:
+        if parameter in kwargs:
[…]

### 2021-09-02 latex/paper.tex
[UPDATE] paper to latest version.

@@
 	% TODO
+	In recent years, the joint increase in image resolution, data sharing and
+	study sizes in neuroimaging demands for Big Data strategies.
+	However, while several options are available, no particular engine emerged.
+	In this paper, we compare the runtime performance of two popular Big Data engines with
+	Python APIs, Apache Spark, and Dask, in processing neuroimaging pipelines.
+	Our experiments use three synthetic data-intensive applications to process
+	the \SI{606}{\gibi\byte} BigBrain image and an actual pipeline to process data
+	from thousands of anatomical images.
+	We benchmark these applications on a dedicated HPC cluster while using varying
+	combinations of the number of nodes, file size, and duration threshold for the pipelines.
+	Our results show that although there are slight differences between Dask and Spark,
+	the performance of the engines is comparable for data-intensive applications.
+	However, Spark requires more memory than Dask to execute, which can lead to 
+	slower runtime depending on configuration and infrastructure.
+	In general, the limiting factor was the data transfer time.
+	While both engines are suitable for neuroimaging, more efforts need to be put
+	to reduce the data transfer time and the memory footprint of applications.
+
 \end{abstract}
@@
 
-\subsubsection{BIDS App example}
-Our fifth application is the BIDS App example, a neuroimaging pipeline to
+\subsubsection{BIDS App Example}
+Our fifth application is the BIDS App Example, a neuroimaging pipeline to
 measure brain volume from MRIs. For this application, we use the CoRR
@@
 then computes the average for each group of participants. Unlike the other
-applications, the BIDS App example is a command-line executed in a Docker image
+applications, the BIDS App Example is a command-line executed in a Docker image
 (bids/example on DockerHub). We converted the Docker image to a Singularity
-image for use in HPC environments, \MD{Cite paper on the reason why this is
-done.} using
+image for use in HPC environments, using
 \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
@@
 Table~\ref{table:parameters} shows the parameters that we varied
-throughout the experiments. We varied (1) the number of workers to assess
-the scalability of the engine scheduler, (2) the BigBrain
-block size in Increment, Multi-Increment, and Histogram to measure the
-effect of different I/O patterns and parallelization degrees, (3) the
-number of iterations to evaluate the effect of the number of tasks, and (4) the
-sleep delay to study the effect of task duration. \TG{these parameters don't seem to match the names used in Table I, they should be harmonized}
- It should be noted that
-increasing the number of iterations for a given sleep delay also increases
-the total compute time of an application.
+throughout the experiments. We varied (1) the number of nodes to assess
+the scalability of the engines' scheduler, (2) the number of files for the BigBrain
+in Increment, and Histogram to measure the
+effect of different I/O patterns and parallelization degrees, and (3) the
+number of iterations to evaluate the effect of the number of tasks.
+It should be noted that increasing the number of iterations for a given sleep 
+delay also increases the total compute time of an application.
 
@@
 		\hline           & Increment & Histogram & Kmeans  & BIDS App Example \\\hline
+		Dataset size [\SI{}{\gibi\byte}] &\multicolumn{2}{c|}{606} & 303 & \multicolumn{1}{c|}{26.67} \\\hline
 		\# of Nodes & \multicolumn{4}{c|}{2, 4, 8} \\\hline
-		Dataset size [\SI{}{\gibi\byte}] &\multicolumn{2}{c|}{606} & 303 & \multicolumn{1}{c|}{26.67} \\\hline
 		Number of file & \multicolumn{2}{c|}{1000, 2500, 5000} & - & \multicolumn{1}{c|}{3491} \\\hline
@@
 Figure \ref{fig:increment_worker} depicts the total execution time spent in each code segment of the \textit{Increment} application.
-The bars are broken down into Idle, Read, Write, and other application-specific functions. \TG{Add a sentence to explain what Idle is. Is it exactly the overhead mentioned previously? if
-so, I would just call it Overhead everywhere (in the figures and in the text)}
+The bars are broken down into Overhead, Read, Write, and other application-specific functions.
 The mean of each bar is calculated from ten repetitions, with the error bar representing the standard deviation.
@@
 While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
-Although Dask has a higher overhead, the total time for both engines is similar; with a maximum makespan difference of \SI{7}{\second} for two nodes.
-This is due to a tradeoff between disk bandwidth and compute/idle time.
+Although Dask has a higher overhead, the total time for both engines is similar; with a maximum makespan difference of 2\% for two nodes.
+This is due to a tradeoff between disk bandwidth and compute/overhead time.
 On the one hand, multiple threads access Lustre concurrently, thus sharing the disk bandwidth resulting in longer I/O time.
@@
 Figure \ref{fig:increment_itr} shows the detailed execution time for the \textit{Increment} application while varying the number of iterations.
-As expected, the idle time for both engines increases with the number of iterations.
-However, idle time for Dask increases significantly faster than Spark, resulting in a \SI{38}{\second} makespan difference for 25 iterations.
+As expected, the overhead for both engines increases with the number of iterations.
+However, overhead for Dask increases significantly faster than Spark, resulting in a 7\% makespan difference for 25 iterations.
 
-Figure \ref{fig:increment_itr} also shows that the total computation time increases as the total I/O time decreases.
+Figure \ref{fig:increment_itr} also shows that the the total I/O time decreases as total computation time increases.
 This is explained by an I/O bottleneck on the Lustre file system.
-When threads are busy with computations, they reduce the contention on Lustre, thus improving I/O speed for individual accesses \TG{but then read or write time should decrease, not computing time.}.
+When threads are busy with computations, it reduces the contention on Lustre, thus improving I/O speed for individual accesses.
 \begin{figure}[!t]
@@
 
-From Figure \ref{fig:increment_block} we also observe that Dask has significantly higher idle time than Spark.
-However, the total execution time of the application is similar due to the tradeoff between I/O and compute/idle time.
+From Figure \ref{fig:increment_block} we also observe that Dask has significantly higher overhead than Spark.
+However, the total execution time of the application is similar due to the tradeoff between I/O and compute/overhead time.
 \begin{figure}[!t]
@@
 The results are similar to varying the number of nodes with Spark having faster compute time for \textit{Calculate\_histogram}.
+This results in a makespan difference of up to 3\%.
 \begin{figure}[!t]
@@
 Figure \ref{fig:kmeans_worker} depicts the total execution time for the \textit{Kmeans} application when varying the number of workers.
-We observe that Dask is multifold faster than Spark.
-Specifically, Spark has much longer idle, \textit{get\_labels} and read time than Dask.
+We observe that Dask is multifold faster than Spark, resulting in a $1.79\times$ to $2.71\times$ lower makespan.
+Specifically, Spark has much longer overhead, \textit{get\_labels} and read time than Dask.
 This difference arises from Spark workers restarting due to lack of executor memory, leading to recomputation of tasks.
-
-% For Spark, we reduced the number of processes per node from 64 to 32 while keeping the memory the same and found similar execution times.
-% However, the Spark worker did not restart with a lower process count and no longer recomputed tasks \TG{removed this as it is discussed later more clearly}.
 
@@
 \subsection{BIDS App Example}
-Figure \ref{fig:bids} shows the total execution time for the \textit{BIDS App example} application when varying the number of nodes.
-We observe that Dask has a significantly lower execution time than Spark, between 10\% to 15\%.
-The main difference comes from an increased amount of stagger tasks for Spark, thus increasing the idle time of the application.
-Moreover, although minimal, a shorter cluster deployment time for Dask reinforces the difference in idle time.
+Figure \ref{fig:bids} shows the total execution time for the \textit{BIDS App Example} application when varying the number of nodes.
+We observe that Dask has slightly lower overhead, thank Spark, although not significant.
+However, the makespan for Dask is between 7.5\% to 14.5\% lower than Spark.
+The main difference comes from an increased amount of stagger tasks for Spark, thus increasing the overhead of the application.
+Moreover, although minimal, a shorter cluster deployment time for Dask reinforces the difference in overhead time.
 As with other experiments, the I/O from Lustre affects the scaling of both engines as we increase the number of nodes.
@@
 \end{figure}
-
-\TG{Could you add information about how the differences in total times translated into makespans, for each application? Either as sentences in the text or as an extra table.}
 
@@
 The experiments from \cite{Mehta:17} found that Dask has a larger startup overhead than Spark, while another study \cite{8588652} found the opposite.
-We did not find significant differences in startup overhead between the engines in this study nor our previous study \cite{8943502}.
+In this study, we found that Dask has a lower startup overhead than Spark; however, we did not in our previous study \cite{8943502}.
 Those apparent contractions most likely come from differences in infrastructure and configuration of the engines.

### 2021-09-02 latex/paper.tex

@@
 	% TODO
-	In recent years, the joint increase in image resolution, data sharing and
-	study sizes in neuroimaging demands for Big Data strategies.
-	However, while several options are available, no particular engine emerged.
+	The joint increase in data resolution, data sharing and
+	study sizes motivates the adoption of Big Data strategies in a number of scientific disciplines.
+	However, while several options are available, no particular Big Data engine emerged.
 	In this paper, we compare the runtime performance of two popular Big Data engines with
@@
 	from thousands of anatomical images.
-	We benchmark these applications on a dedicated HPC cluster while using varying
-	combinations of the number of nodes, file size, and duration threshold for the pipelines.
+	We benchmark these applications on a dedicated HPC cluster running the Lustre file system while using varying
+	combinations of number of nodes, file size, and task duration.
 	Our results show that although there are slight differences between Dask and Spark,
 	the performance of the engines is comparable for data-intensive applications.
-	However, Spark requires more memory than Dask to execute, which can lead to 
+	However, Spark requires more memory than Dask, which can lead to 
 	slower runtime depending on configuration and infrastructure.

### 2021-09-03 latex/paper.tex
Minor edits

@@
 	% TODO
-	The joint increase in data resolution, data sharing and
-	study sizes motivates the adoption of Big Data strategies in a number of scientific disciplines.
-	However, while several options are available, no particular Big Data engine emerged.
+	The general increase in data size and data sharing motivates the adoption of Big Data strategies in a number of scientific disciplines.
+	However, while several options are available, no particular guidelines exist for the selection of a Big Data engine.
 	In this paper, we compare the runtime performance of two popular Big Data engines with
@@
 We installed the Slurm scheduler and Lustre file system~\cite{braam2019lustre} on a dedicated cluster to mimic a typical HPC system.
-Lustre is a powerful and scalable file system for HPC environments offering parallel file accesses.
+Lustre is a popular file system for HPC environments offering parallel file accesses.
 
@@
 Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} application when varying the number of nodes.
-Overall, Spark has up to 5\% lower total execution time, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
+Overall, Spark's total execution time is up to 5\% lower than Dask's, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
 This is unexpected as the code base for both engines is the same, with only the calling APIs differing.
@@
 Figure \ref{fig:kmeans_worker} depicts the total execution time for the \textit{Kmeans} application when varying the number of workers.
-We observe that Dask is multifold faster than Spark, resulting in a $1.79\times$ to $2.71\times$ lower makespan.
+We observe that Dask is multifold faster than Spark, resulting in a $1.79\times$ to $2.71\times$ speed-up ratio on the makespan.
 Specifically, Spark has much longer overhead, \textit{get\_labels} and read time than Dask.
@@
 We observe that Dask has slightly lower overhead, thank Spark, although not significant.
-However, the makespan for Dask is between 7.5\% to 14.5\% lower than Spark.
+However, the makespan for Dask is between 7.5\% to 14.5\% lower than for Spark.
 The main difference comes from an increased amount of stagger tasks for Spark, thus increasing the overhead of the application.
@@
 We used Lustre as shared file system as it is a standard solution in HPC systems.
-The performance of Lustre degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
+The performance of our Lustre installation degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
 In a previous study \cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
 In this paper, Lustre shows some scaling when increasing the number of nodes, although not linear.
-This suggests that increasing the parallelism level of the file system might help alleviate the I/O bottleneck observed.
+This suggests that increasing the parallelism level of the file system, including through potential configuration optimizations, might help alleviate the I/O bottleneck observed.
 

### 2021-09-03 latex/paper.tex
[REFACTOR] use CCPE latex template.

@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
+\documentclass[AMA,STIX1COL]{WileyNJD-v2}
 % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 \hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
-\usepackage{algpseudocode}
-\usepackage{algorithm}
 \usepackage{caption}
@@
 
-%\setlength{\textfloatsep}{7pt}
+\renewcommand{\thefootnote}{\fnsymbol{footnote}}
 
-\newcommand{\TG}[1]{\color{cyan}From Tristan: #1 \color{black}}
-\newcommand{\MD}[1]{\color{magenta}From Mathieu: #1 \color{black}}
-\newcommand{\VHS}[1]{\color{green}From Valerie: #1 \color{black}}
+\articletype{WORKS19 Special Issue on Workflows in Support of Large-Scale Science GEs:  Chase Wu, Hoang Nguyen}
+
+\received{}
+\revised{}
+\accepted{}
+
+\raggedbottom
 
@@
 
-\author{Mathieu Dugr\'e, Val\'erie Hayot-Sasson, Tristan Glatard\\
-	Department of Computer Science and Software Engineering\\
-	Concordia University, Montr\'eal, Qu\'ebec, Canada\\
-	\{mathieu.dugre, valerie.hayot-sasson, tristan.glatard\}@concordia.ca
-	\vspace*{0.8cm} % to avoid weird spacing of 1st page by Latex.
-}
+\author[1]{Mathieu Dugr\'e*}
+\author[1]{Val\'erie Hayot-Sasson}
+\author[1]{Tristan Glatard}
 
-\maketitle
+\authormark{Mathieu Dugr\'e \textsc{et al}}
 
-\begin{abstract}
-	% TODO
+\address[1]{\orgdiv{Department of Computer Science and Software Engineering}, \orgname{Concordia University}, \orgaddress{\state{Quebec}, \country{Canada}}}
+
+\corres{Mathieu Dugr\'e, Department of Computer Science and Software Engineering, Concordia University, Quebec, Montreal, Canada,\email{mathieu.dugre@concordia.ca}}
+
+% \presentaddress{This is sample for present address text this is sample for present address text}
+
+\abstract[Summary]{
 	The general increase in data size and data sharing motivates the adoption of Big Data strategies in a number of scientific disciplines.
@@
 	to reduce the data transfer time and the memory footprint of applications.
+}
 
-\end{abstract}
+\keywords{Performance, Big Data, Dask, Spark, Neuroimaging, HPC}
 
-\begin{IEEEkeywords}
-	Performance, Big Data, Dask, Spark, Neuroimaging, HPC
-\end{IEEEkeywords}
+\maketitle
 
@@
 However, using the network file system (NFS) led to an I/O bottleneck limiting application performance.
-Another study \cite{8588652} also reported network and I/O bottlenecks when concurrency increased.
+Another study~\cite{8588652} also reported network and I/O bottlenecks when concurrency increased.
 This same study found that the startup overhead time for Spark was more prominent than for Dask.
@@
 The latter study processed a \SI{100}{\giga\byte} dataset using a neuroimaging pipeline and found Dask to be up to 14\% faster than Spark due to ``more efficient pipelining" and serialization time to Python.
-On the contrary, the work in \cite{10.1145/3225058.3225128} shows that with large datasets, Spark provides better speedup than Dask.
+On the contrary, the work in~\cite{10.1145/3225058.3225128} shows that with large datasets, Spark provides better speedup than Dask.
 However, that study used the Dask Bag API, which has performance limitations compared to other Dask APIs.
@@
 Dask was designed with data locality and in-memory computing in mind to mitigate the data transfer bottleneck in Big Data workflows.
-Data locality, popularized by Map-Reduce \cite{dean2008mapreduce}, schedules tasks where the data reside.
+Data locality, popularized by Map-Reduce~\cite{dean2008mapreduce}, schedules tasks where the data reside.
 In-memory computing minimizes the overhead of transferring data to disk by keeping data in memory when possible.
@@
 \subsubsection{Increment}
-We adapted the increment application used in \cite{hayot2019performance}.
+We adapted the increment application used in~\cite{hayot2019performance}.
 This synthetic application reads blocks of the BigBrain from Lustre and
@@
 mimics the processing of multiple independent subjects in parallel.
-\begin{figure}[!ht]
-	\centering
-	\includegraphics[height=\columnwidth,
-	angle=0]{figures/increment.png}
-	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
-	\label{fig:graph-increment}
-\end{figure}
+% \begin{figure}[!hb]
+% 	\centering
+% 	\includegraphics[height=0.5\columnwidth,
+% 	angle=0]{figures/increment.png}
+% 	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
+% 	\label{fig:graph-increment}
+% \end{figure}
 
@@
 The amount of inter-worker communication is in-between the increment and multi-increment applications.
-\begin{figure}[!hb]
+
+\begin{figure}[h]
 	\centering
-	\includegraphics[height=\columnwidth,
-	angle=0]{figures/histogram.png}
+	\begin{minipage}{.45\textwidth}
+		\centering
+		\includegraphics[width=0.45\linewidth]{figures/increment.png}
+		\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
+		\label{fig:graph-increment}		
+	\end{minipage}
+	\begin{minipage}{.45\textwidth}
+		\centering
+		\includegraphics[width=0.45\linewidth]{figures/histogram.png}
 		\caption{Task graph for Histogram with 3 BigBrain blocks.}
 		\label{fig:graph-histogram}			
+	\end{minipage}
 \end{figure}
@@
 On the other hand, when most threads are either idle or performing computations, it reduces the contention on the disk bandwidth resulting in shorter I/O time.
-\begin{figure}[!t]
+\begin{figure}[!h]
 	\centering
-	\includegraphics[clip,width=\columnwidth]{figures/stacked_increment_worker.jpg}
+	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_worker.jpg}
 	\caption{Increment: varying nodes -- 5000 files, 5 iterations, \SI{1}{\second} delay}
@@
 When threads are busy with computations, it reduces the contention on Lustre, thus improving I/O speed for individual accesses.
-\begin{figure}[!t]
+\begin{figure}[!h]
 	\centering
-	\includegraphics[clip,width=\columnwidth]{figures/stacked_increment_itr.jpg}
+	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_itr.jpg}
 	\caption{Increment: varying iterations -- 4 nodes, 5000 files, \SI{1}{\second} delay}
@@
 However, the total execution time of the application is similar due to the tradeoff between I/O and compute/overhead time.
-\begin{figure}[!t]
+\begin{figure}[!h]
 	\centering
-	\includegraphics[clip,width=\columnwidth]{figures/stacked_increment_block.jpg}
+	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_block.jpg}
 	\caption{Increment: varying \# of files -- 4 nodes, 5 iterations, \SI{1}{\second} delay}
@@
 As with previous experiments, the overhead for Spark is less and scales better, although compensated by a longer total I/O time.
-\begin{figure}[!t]
+\begin{figure}[!h]
 	\centering
-	\includegraphics[clip,width=\columnwidth]{figures/stacked_histogram_worker.jpg}
+	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_histogram_worker.jpg}
 	\caption{Histogram: varying nodes -- 5000 files, 5 iterations, \SI{1}{\second} delay}
@@
 This results in a makespan difference of up to 3\%.
-\begin{figure}[!t]
+\begin{figure}[!h]
 	\centering
-	\includegraphics[clip,width=\columnwidth]{figures/stacked_histogram_block.jpg}
+	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_histogram_block.jpg}
 	\caption{Histogram: varying \# of files -- 4 nodes, 5 iterations, \SI{1}{\second} delay}
@@
 For brevity we omitted the figure when varying the number of iterations as the results are similar to varying the number of workers.
-\begin{figure}[!t]
+\begin{figure}[!h]
 	\centering
-	\includegraphics[clip,width=\columnwidth]{figures/stacked_kmeans_worker.jpg}
+	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_kmeans_worker.jpg}
 	\caption{Kmeans: varying nodes -- 5000 files, 3 iterations, \SI{1}{\second} delay}
@@
 As with other experiments, the I/O from Lustre affects the scaling of both engines as we increase the number of nodes.
-\begin{figure}[!t]
+\begin{figure}[!h]
 	\centering
-	\includegraphics[clip,width=\columnwidth]{figures/stacked_bids.jpg}
+	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_bids.jpg}
 	\caption{BIDS App Example: varying nodes}
@@
 The performance of our Lustre installation degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
-In a previous study \cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
+In a previous study~\cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
 In this paper, Lustre shows some scaling when increasing the number of nodes, although not linear.
@@
 
-As in our previous study \cite{8943502}, we observe that the I/O bottleneck from the file system has a lower impact on more compute-intensive applications, such as BIDS App Example.
+As in our previous study~\cite{8943502}, we observe that the I/O bottleneck from the file system has a lower impact on more compute-intensive applications, such as BIDS App Example.
 However, it significantly affects data-intensive applications, such as Increment and Histogram.
@@
 i.e., when the bandwidth of the file system is saturated, engine overhead does not increase the makespan as long as it desynchronizes the data transfers.
-In a previous study \cite{8943502}, we also observed this tradeoff between overhead and data transfer when the file system was saturated.
+In a previous study~\cite{8943502}, we also observed this tradeoff between overhead and data transfer when the file system was saturated.
 It could be interesting to explore strategies that take advantage of this tradeoff to reduce the application makespan.
@@
 Overall, except for Kmeans, our experiments did not show substantial performance differences between Spark and Dask.
-In contradiction with the experiments of \cite{Mehta:17}, we found that Dask had a larger overhead than Spark, especially when scaling the number of workers.
+In contradiction with the experiments of~\cite{Mehta:17}, we found that Dask had a larger overhead than Spark, especially when scaling the number of workers.
 However, this was compensated by the overhead and data transfer tradeoff mentioned previously.
@@
 
-The experiments from \cite{Mehta:17} found that Dask has a larger startup overhead than Spark, while another study \cite{8588652} found the opposite.
-In this study, we found that Dask has a lower startup overhead than Spark; however, we did not in our previous study \cite{8943502}.
+The experiments from~\cite{Mehta:17} found that Dask has a larger startup overhead than Spark, while another study~\cite{8588652} found the opposite.
+In this study, we found that Dask has a lower startup overhead than Spark; however, we did not in our previous study~\cite{8943502}.
 Those apparent contractions most likely come from differences in infrastructure and configuration of the engines.
@@
 Our results from the Histogram experiments show that the Dask Bags API was slower than Spark.
-The experiments from \cite{10.1145/3225058.3225128} also found that Dask Bags performed slower than Spark.
+The experiments from~\cite{10.1145/3225058.3225128} also found that Dask Bags performed slower than Spark.
 We conclude that when using Dask, the other Dask APIs should be favored unless Dask Bags are strictly required.
@@
 Our results showed that, for data-intensive applications, increasing the number of workers should only be done if the bandwidth of the file system is not saturated.
-Two other studies support this \cite{8943502, 8588652}.
+Two other studies support this~\cite{8943502, 8588652}.
 Therefore, it is critical to ensure enough bandwidth to the file system before scaling the compute resources for data-intensive applications.
@@
 \section*{Acknowledgment}
-% TODO
 Mathieu Dugre\'e was funded by an Alexander Graham Bell scholarship from
@@
 
-\bibliographystyle{IEEEtran}
-\bibliography{IEEEabrv,reference}
+\bibliographystyle{wileyNJD-AMA}
+\bibliography{reference}
 \end{document}

### 2022-02-18 latex/paper.tex
[UPDATE] Address paper comments.

@@
 \section{Introduction}
-The rise in data sharing coupled with improved data collection technologies leads neuroimaging into the Big Data era\cite{ALFAROALMAGRO:18, van2014human, ConpPortal}.
+The rise in data sharing coupled with improved data collection technologies leads neuroimaging into the Big Data era\cite{ALFAROALMAGRO2018400, van2014human, ConpPortal}.
 While standard neuroimaging workflow engines, such as Nipype\cite{Nipype:11}, are well rounded to process standard compute-intensive pipelines,
 they lack Big Data strategies (i.e., in-memory computing, data locality, and lazy-evaluation) to improve the performance of increasingly prevalent data-intensive pipelines.
-A previous study\cite{hayot2019performance} noted the importance of in-memory computing and data locality to improve the performance of data-intensive pipelines.
+A previous study\cite{8752675} noted the importance of in-memory computing and data locality to improve the performance of data-intensive pipelines.
 In this work, we study how performance benchmarks generalize across Big Data engines in an HPC environment.
@@
 This paper focuses on the performance of Big Data engines in HPCs environments.
-We installed the Slurm scheduler and Lustre file system~\cite{braam2019lustre} on a dedicated cluster to mimic a typical HPC system.
-Lustre is a popular file system for HPC environments offering parallel file accesses.
+We installed the Slurm scheduler and Lustre file system~\cite{Braam2019TheLS} on a dedicated cluster to mimic a typical HPC system.
+Lustre is a reference file system for HPC environments, used in over 60 of the TOP100 supercomputers~\cite{OpenSFS-lustre}, offering parallel file accesses.
 
@@
 The above studies compared the performance of the engine at a high level.
-In comparison, our study provides a detailed analysis of the performance differences and similarities between Dask and Spark for data-intensive applications.
+In comparison, our study provides a detailed analysis of the performance differences
+and similarities, for HPC, between Dask and Spark for data-intensive applications.
 
@@
 \subsection{Lustre}
-The Lustre file system is a parallel file system well suited for HPC environments.
+The Lustre file system is a parallel file system of reference in HPC environments.
 Lustre offers a POSIX-compliant interface with the file system.
@@
 InfiniBand, Ethernet, or a combination of both can be used to interconnect these different components.
+
+While other filesystem such as HDFS or cloud storage could be considered, we chose Lustre to best represent an HPC environment.
+HDFS makes use of permanent local storage on compute node, which is usually not feasable in HPC.
+Morevoer, HPC node rarely have access to internet, limiting the use of cloud storage.
 
@@
 CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
+All of the compute and storage nodes are inter-connected by a \SI{10}{\giga\bit\per\second} Ethernet link through a switch.
+The login node, hosting an NFS, is also connected by a \SI{10}{\giga\bit\per\second} Ethernet link to the compute nodes.
 
@@
 Lustre \textit{2.14.0-1.el8} is installed on both the Lustre clients and servers.
+%  TODO Inode config??
+
+Our system uses similar hardware as Compute Canada Beluga, and replicates as
+close as possible it's configuration.
+This let us make a fair comparison of Spark and Dask in the condition of a realistic
+HPC system.
 
@@
 Also, two neuroimaging applications are used to study more realistic 
-applications, namely \textit{Kmeans} and \textit{BIDS App Example}.
+applications, namely \textit{Kmeans} and \textit{BIDS App Example}\cite{gorgolewski2017bids}.
 
@@
 
+The scripts used for our experiements and the results generated are available on
+GitHub at: \href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}.
+
 \subsubsection{Increment}
-We adapted the increment application used in~\cite{hayot2019performance}.
+We adapted the increment application used in~\cite{8752675}.
 This synthetic application reads blocks of the BigBrain from Lustre and
@@
 \subsection{I/O bottleneck}
-We used Lustre as shared file system as it is a standard solution in HPC systems.
+We used Lustre as shared file system as it is a defacto solution in HPC systems.
 The performance of our Lustre installation degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
@@
 The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
-This is unexpected, given that Dask and Spark are configured similarly and use the same amount of resources.
+This is unexpected, given that Dask and Spark are configured similarly and are allocated the same amount of resources.
 We think those issues arise from a combination of
@@
 When reducing the number of processes from 64 to 32 for Spark (effectively doubling the memory of each process), the total execution time became similar between Spark and Dask.
-However, with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
+That is, with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
 We suppose that given more memory per thread/process, Dask and Spark would then perform similarly.
+For our experiments, we allocated all available memory on the compute nodes.
+Therefore, it was not possible to allocate more resources to test this hypothesis.
 
@@
 \bibliographystyle{wileyNJD-AMA}
-\bibliography{reference}
+\bibliography{paper}
 \end{document}

### 2022-02-18 plot-script/plot.py

@@ -20,6 +20,7 @@ from bokeh.palettes import Colorblind8
 import matplotlib
 from matplotlib.patches import Patch
 import matplotlib.pyplot as plt
+from matplotlib.ticker import StrMethodFormatter
 import numpy as np
 import pandas as pd
 
@@ -324,7 +325,7 @@ def stacked_bar(
     print(f"{benchmark_dir=}")
     filenames = glob.glob(benchmark_dir + "/*summary-*.csv")
 
-    xticks_label = sorted(
+    xticks_labels = sorted(
         {
             float(x.replace(freedom + "=", ""))
             for k in filenames
@@ -332,7 +333,9 @@ def stacked_bar(
             if freedom in x
         }
     )
-    x_pos = lambda i: np.arange(len(xticks_label)) + bar_width * i
+    xticks_labels = atomic_array_convert(xticks_labels, int)
+    
+    x_pos = lambda i: np.arange(len(xticks_labels)) + bar_width * i
 
     results = defaultdict(lambda: defaultdict(list))
     for x in filenames:
@@ -366,7 +369,7 @@ def stacked_bar(
         ############
         # Plotting #
         ############
-        y_offset = stats["mean"].max() * 0.01
+        y_offset = stats["mean"].max() * 0.025
         totals = stats["mean"].sum(axis=0)
         color = Colorblind8[xs % len(Colorblind8)]
         # Plot idle time as a special case since it always exist.
@@ -406,15 +409,15 @@ def stacked_bar(
             ax.text(
                 x,
                 total + y_offset,
-                round(total),
+                f"{round(total):,}",
                 ha="center",
                 weight="bold",
                 fontsize=14,
                 color=color,
             )
 
-    xticks_loc = np.arange(len(xticks_label)) + bar_width * xs / 2
-    plt.xticks(xticks_loc, xticks_label)
+    xticks_locs = np.arange(len(xticks_labels)) + bar_width * xs / 2
+    plt.xticks(xticks_locs, xticks_labels)
     ax.set_xlabel(xlabel, fontweight="bold")
     ax.set_ylabel("Makespan [s]", fontweight="bold")
     ax.set_ylabel("Total time [s]", fontweight="bold")
@@ -468,7 +471,12 @@ def stacked_bar(
         os.makedirs(os.path.dirname(save_name), exist_ok=True)
         plt.savefig(save_name, bbox_inches="tight")
[…]

### 2022-02-19 latex/paper.tex
[ADD] Highligh modified parts.

@@
 \renewcommand{\thefootnote}{\fnsymbol{footnote}}
+\newcommand{\HL}[1]{\hl{#1}}
 
@@
 We installed the Slurm scheduler and Lustre file system~\cite{Braam2019TheLS} on a dedicated cluster to mimic a typical HPC system.
-Lustre is a reference file system for HPC environments, used in over 60 of the TOP100 supercomputers~\cite{OpenSFS-lustre}, offering parallel file accesses.
+\HL{Lustre is a reference file system for HPC environments, used in over 60 of the TOP100 supercomputers~{\cite{OpenSFS-lustre}}, offering parallel file accesses.}
 
@@
 In comparison, our study provides a detailed analysis of the performance differences
-and similarities, for HPC, between Dask and Spark for data-intensive applications.
+and similarities\HL{, for HPC,} between Dask and Spark for data-intensive applications.
 
@@
 \subsection{Lustre}
-The Lustre file system is a parallel file system of reference in HPC environments.
+The Lustre file system is a parallel file system \HL{of reference in} HPC environments.
 Lustre offers a POSIX-compliant interface with the file system.
@@
 
-While other filesystem such as HDFS or cloud storage could be considered, we chose Lustre to best represent an HPC environment.
+\HL{While other filesystem such as HDFS or cloud storage could be considered, we chose Lustre to best represent an HPC environment.
 	HDFS makes use of permanent local storage on compute node, which is usually not feasable in HPC.
-Morevoer, HPC node rarely have access to internet, limiting the use of cloud storage.
+Morevoer, HPC node rarely have access to internet, limiting the use of cloud storage.}
 
@@
 CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
-All of the compute and storage nodes are inter-connected by a \SI{10}{\giga\bit\per\second} Ethernet link through a switch.
-The login node, hosting an NFS, is also connected by a \SI{10}{\giga\bit\per\second} Ethernet link to the compute nodes.
+\HL{The compute and storage nodes are inter-connected by {\SI{10}{\giga\bit\per\second}} Ethernet link through a switch.
+	The login node, hosting an NFS, is also connected by a {\SI{10}{\giga\bit\per\second}} Ethernet link to the compute nodes.}
 
@@
 
-Our system uses similar hardware as Compute Canada Beluga, and replicates as
+\HL{Our system uses similar hardware as Compute Canada Beluga, and replicates as
 	close as possible it's configuration.
 	This let us make a fair comparison of Spark and Dask in the condition of a realistic
-HPC system.
+HPC system.}
 
@@
 Also, two neuroimaging applications are used to study more realistic 
-applications, namely \textit{Kmeans} and \textit{BIDS App Example}\cite{gorgolewski2017bids}.
+applications, namely \textit{Kmeans} and \textit{BIDS App Example}\HL{{\cite{gorgolewski2017bids}}}.
 
@@
 
-The scripts used for our experiements and the results generated are available on
-GitHub at: \href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}.
+\HL{The scripts used for our experiements and the results generated are available on
+	GitHub at: \mbox{\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}}.}
 
@@
 \subsection{I/O bottleneck}
-We used Lustre as shared file system as it is a defacto solution in HPC systems.
+We used Lustre as shared file system as it is a \HL{de facto} solution in HPC systems.
 The performance of our Lustre installation degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
@@
 Otherwise, the compute resource might be underutilized.
+% TODO talk about the limitation of our current setup.
 
@@
 The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
-This is unexpected, given that Dask and Spark are configured similarly and are allocated the same amount of resources.
+This is unexpected, given that Dask and Spark are configured similarly and \HL{are allocated} the same amount of resources.
 We think those issues arise from a combination of
@@
 When reducing the number of processes from 64 to 32 for Spark (effectively doubling the memory of each process), the total execution time became similar between Spark and Dask.
-That is, with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
+\HL{That is,} with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
 We suppose that given more memory per thread/process, Dask and Spark would then perform similarly.
-For our experiments, we allocated all available memory on the compute nodes.
-Therefore, it was not possible to allocate more resources to test this hypothesis.
+\HL{For our experiments, we allocated all available memory on the compute nodes.
+Therefore, it was not possible to allocate more resources to test this hypothesis.}
 

### 2022-02-22 plot-script/plot.py
[ADD] t-test + significant difference on plot.

@@ -23,6 +23,7 @@ import matplotlib.pyplot as plt
 from matplotlib.ticker import StrMethodFormatter
 import numpy as np
 import pandas as pd
+from scipy.stats import ttest_ind
 
 
 def gantt(
@@ -346,6 +347,8 @@ def stacked_bar(
         results[framework][experiment].append(x)
 
     # Application detailed time spent
+    ttest_data = {}
+    max_height = -float("inf")
     for xs, (framework, experiments) in enumerate(sorted(results.items())):
         data = [
             np.array(
@@ -359,7 +362,7 @@ def stacked_bar(
                 key=lambda k: float(re.search(f"{freedom}=(\d+\.?\d*)", k).group(1)),
             )
         ]
-
+        ttest_data[framework] = np.array([x.sum(axis=1) for x in data])
         stats = {
             "mean": np.array([x.mean(axis=0) for x in data]).T,
             "std": np.array([x.std(axis=0) for x in data]).T,
@@ -372,6 +375,7 @@ def stacked_bar(
         y_offset = stats["mean"].max() * 0.025
         totals = stats["mean"].sum(axis=0)
         color = Colorblind8[xs % len(Colorblind8)]
+        max_height = max(totals.max(), max_height)
         # Plot idle time as a special case since it always exist.
         # Thus we want to having consistent hatch and position.
         ax.bar(
@@ -415,7 +419,7 @@ def stacked_bar(
                 fontsize=14,
                 color=color,
             )
-
+            
     xticks_locs = np.arange(len(xticks_labels)) + bar_width * xs / 2
     plt.xticks(xticks_locs, xticks_labels)
     ax.set_xlabel(xlabel, fontweight="bold")
@@ -426,6 +430,37 @@ def stacked_bar(
         plt.ylim([0, ylim])
 
     plt.title(title)
+    
+    # Significance difference between frameworks.
+    y0 = max_height * 1.05
+    y1 = y0 * 1.025
+    frameworks = list(results)
+    for i in range(len(frameworks)-1):
+        a = frameworks[i]
+        b = frameworks[i+1]
+        _, pvalues = ttest_ind(ttest_data[a], ttest_data[b], axis=1, equal_var=False)
+        x0s = x_pos(i)
+        x1s = x_pos(i+1)
+        for x0, x1, pvalue in zip(x0s, x1s, pvalues):
+            if pvalue <= 0.05:
+                p_text = "*"
+                if pvalue <= 0.005: p_text = "**"
[…]

### 2022-02-22 latex/paper.tex
[UPDATE] write about I/O bottleneck limitations.

@@
 In a previous study~\cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
-In this paper, Lustre shows some scaling when increasing the number of nodes, although not linear.
-This suggests that increasing the parallelism level of the file system, including through potential configuration optimizations, might help alleviate the I/O bottleneck observed.
+\HL{In this paper, the performance scale when increasing the number of nodes, although not linear due to an I/O bottleneck.
+	The system was designed such that the filesystem has similar speed to the network.
+	Upgrading to more performant networking and storage compononents would reduce the bottleneck seen when increasing the number of workers.
 
-As in our previous study~\cite{8943502}, we observe that the I/O bottleneck from the file system has a lower impact on more compute-intensive applications, such as BIDS App Example.
+	As in our previous study~{\cite{8943502}}, we observe that the I/O bottleneck has a lower impact on more compute-intensive applications, such as Kmeans and BIDS App Example.
 	However, it significantly affects data-intensive applications, such as Increment and Histogram.
+}
 
-These results suggest that increasing the parallelism of the file system is critical when scaling the compute resources to process a data-intensive application.
+These results suggest that increasing the parallelism of the file system \HL{and speed of the network} is critical when scaling the compute resources to process a data-intensive application.
 Otherwise, the compute resource might be underutilized.
-% TODO talk about the limitation of our current setup.
 
@@
 \section*{Acknowledgment}
-Mathieu Dugre\'e was funded by an Alexander Graham Bell scholarship from
+Mathieu Dugr\'e was funded by an Alexander Graham Bell scholarship from
 the Natural Sciences and Engineering Research Council of Canada. This work

### 2022-02-22 plot-script/plot.py
[UPDATE] modify plot legend to minimize white space.

@@ -1,5 +1,6 @@
 from collections import defaultdict
 import glob
+import math
 import os
 import re
 
@@ -487,17 +488,17 @@ def stacked_bar(
             )
         )
 
-    delta = abs(len(func_patches) - len(framework_patches))
-    if len(func_patches) > len(framework_patches):
-        framework_patches += [Patch(alpha=0)] * delta
-    elif len(framework_patches) > len(func_patches):
-        func_patches = [Patch(alpha=0)] * delta + func_patches
+    # delta = abs(len(func_patches) - len(framework_patches))
+    # if len(func_patches) > len(framework_patches):
+    #     framework_patches += [Patch(alpha=0)] * delta
+    # elif len(framework_patches) > len(func_patches):
+    #     func_patches = [Patch(alpha=0)] * delta + func_patches
 
     plt.legend(
         handles=(framework_patches + func_patches[::-1]),
-        loc="upper right",
-        bbox_to_anchor=(1, -0.05),
-        ncol=2,
+        loc="upper center",
+        bbox_to_anchor=(0.5, -0.1),
+        ncol=math.ceil(len(framework_patches + func_patches)/2),
         title="Total time [s]",
     )
 

### 2022-02-22 plot-script/plot.py
[FIX] force yaxis notation for saved figures.

@@ -501,17 +501,17 @@ def stacked_bar(
         ncol=math.ceil(len(framework_patches + func_patches)/2),
         title="Total time [s]",
     )
-
-    if save_name:
-        plt.title("")  # Remove title for used with Latex Fig.
-        os.makedirs(os.path.dirname(save_name), exist_ok=True)
-        plt.savefig(save_name, bbox_inches="tight")
-        plt.title(title)
     
     yticks_locs, yticks_labels = plt.yticks()
     yticks_locs = atomic_array_convert(yticks_locs, int)
     plt.yticks(yticks_locs, yticks_labels)
     ax.yaxis.set_major_formatter(StrMethodFormatter('{x:,}'))
+    
+    if save_name:
+        plt.title("")  # Remove title for used with Latex Fig.
+        os.makedirs(os.path.dirname(save_name), exist_ok=True)
+        plt.savefig(save_name, bbox_inches="tight")
+        plt.title(title)
     
     plt.show()
 

### 2022-02-23 latex/paper.tex
[UPDATE] fix few comment and generate gantt chart for new kmeans experiments.

@@
 Lustre \textit{2.14.0-1.el8} is installed on both the Lustre clients and servers.
-%  TODO Inode config??
+\HL{We used the default inode configuration for Lustre.}
 
@@
 simulates computation by sleeping for a specified period (Fig.~\ref{fig:graph-increment}). To simulate
-intermediate results, we repeat the sleep process for a configurable amount
-of time. We prevent data caching of the blocks by incrementing their voxels
+intermediate results, we repeat \HL{a {\SI{1}{\second}} sleep delay during each incrementation}.
+We prevent data caching of the blocks by incrementing their voxels
 value by one after each sleep operation. Finally, we write the resulting
@@
 
+\HL{Due to resources limitation on our infrastructure, the Kmeans experiments
+	are performed with only a sample of half the BigBrain dataset ({\SI{303}{\gibi\byte}}).}
+
 \subsubsection{BIDS App Example}
@@
 		\# of Nodes & \multicolumn{4}{c|}{2, 4, 8} \\\hline
-		Number of file & \multicolumn{2}{c|}{1000, 2500, 5000} & - & \multicolumn{1}{c|}{3491} \\\hline
-		\# of Iterations & 1, 5, 25  & -         & 1, 3, 9 & -                \\\hline
+		Number of file & \multicolumn{2}{c|}{1000, 2500, 5000} & 5000 & \multicolumn{1}{c|}{3491} \\\hline
+		\# of Iterations & 1, 5, 25  & -         & 1, 3   & -                \\\hline
 	\end{tabular}
@@
 While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
-Although Dask has a higher overhead, the total time for both engines is similar; with a maximum makespan difference of 2\% for two nodes.
+\HL{Although Dask has a higher overhead, the total time for both engines is overall significantly similar.}
 This is due to a tradeoff between disk bandwidth and compute/overhead time.
@@
 On the other hand, when most threads are either idle or performing computations, it reduces the contention on the disk bandwidth resulting in shorter I/O time.
+\HL{The only significant difference in total time is when two nodes are used.
+	This results in an approximately 3\% total runtime difference.}
 \begin{figure}[!h]
@@
 	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_worker.jpg}
-	\caption{Increment: varying nodes -- 5000 files, 5 iterations, \SI{1}{\second} delay}
+	\caption{Increment: varying nodes -- 5000 files, 5 iterations}
 	\label{fig:increment_worker}
@@
 	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_itr.jpg}
-	\caption{Increment: varying iterations -- 4 nodes, 5000 files, \SI{1}{\second} delay}
+	\caption{Increment: varying iterations -- 4 nodes, 5000 files}
 	\label{fig:increment_itr}
@@
 From Figure \ref{fig:increment_block} we also observe that Dask has significantly higher overhead than Spark.
-However, the total execution time of the application is similar due to the tradeoff between I/O and compute/overhead time.
+However, the total execution time of the application is \HL{significantly equivalent} due to the tradeoff between I/O and compute/overhead time.
 \begin{figure}[!h]
@@
 	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_block.jpg}
-	\caption{Increment: varying \# of files -- 4 nodes, 5 iterations, \SI{1}{\second} delay}
+	\caption{Increment: varying \# of files -- 4 nodes, 5 iterations}
 	\label{fig:increment_block}
@@
 Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} application when varying the number of nodes.
-Overall, Spark's total execution time is up to 5\% lower than Dask's, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
+Overall, \HL{the execution time for Spark is significantly lower than for Dask by up to 5\%}, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
 This is unexpected as the code base for both engines is the same, with only the calling APIs differing.
@@
 	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_histogram_worker.jpg}
-	\caption{Histogram: varying nodes -- 5000 files, 5 iterations, \SI{1}{\second} delay}
+	\caption{Histogram: varying nodes -- 5000 files}
 	\label{fig:histogram_worker}
@@
 The results are similar to varying the number of nodes with Spark having faster compute time for \textit{Calculate\_histogram}.
-This results in a makespan difference of up to 3\%.
+This results in a \HL{significant} makespan difference of up to 3\%.
 \begin{figure}[!h]
@@
 	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_histogram_block.jpg}
-	\caption{Histogram: varying \# of files -- 4 nodes, 5 iterations, \SI{1}{\second} delay}
+	\caption{Histogram: varying \# of files -- 4 nodes}
 	\label{fig:histogram_block}
@@
 
-For brevity we omitted the figure when varying the number of iterations as the results are similar to varying the number of workers.
 \begin{figure}[!h]
@@
 	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_kmeans_worker.jpg}
-	\caption{Kmeans: varying nodes -- 5000 files, 3 iterations, \SI{1}{\second} delay}
+	\caption{Kmeans: varying nodes -- 5000 files, 3 iterations}
 	\label{fig:kmeans_worker}
 \end{figure}
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[clip,width=\columnwidth]{figures/dask_kmeans_gantt.png}
+	\caption{Kmeans: Dask with 8 nodes, 5000 files, and 3 iterations}
+	\label{fig:dask_kmeans_gantt}
+\end{figure}
+
+\begin{figure}[!h]
+	\centering
+	\includegraphics[clip,width=\columnwidth]{figures/spark_kmeans_gantt.png}
+	\caption{Kmeans: Spark with 8 nodes, 5000 files, and 3 iterations}
+	\label{fig:spark_kmeans_gantt}
+\end{figure}
+
+For brevity we omitted the figure\HL{s} when varying the number of iterations as the results are similar to varying the number of workers.
 
@@
 \subsection{I/O bottleneck}
+\label{subsec:io-bottleneck}
 We used Lustre as shared file system as it is a \HL{de facto} solution in HPC systems.
@@
 	The system was designed such that the filesystem has similar speed to the network.
-	Upgrading to more performant networking and storage compononents would reduce the bottleneck seen when increasing the number of workers.
+Upgrading to more performant networking and storage compononents would reduce the bottleneck seen when increasing the number of workers.}
 
+\HL{
 	As in our previous study~{\cite{8943502}}, we observe that the I/O bottleneck has a lower impact on more compute-intensive applications, such as Kmeans and BIDS App Example.
@@
 \subsection{Tradeoff between overhead and data transfer}
+\label{subsec:overhead-tradeoff}
 Overall, except for Kmeans, the overhead for Dask is larger than for Spark.
@@
 
-When reducing the number of processes from 64 to 32 for Spark (effectively doubling the memory of each process), the total execution time became similar between Spark and Dask.
+When reducing the number of processes from 64 to 32 for Spark (effectively doubling the memory of each process), \HL{the difference in execution time between Spark and Dask reduced significantly}.
 \HL{That is,} with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
@@
 Otherwise, those extras resources might be underused.
+
+\subsection{Limitations}
+\HL{
+	To make a fair comparison between Spark and Dask, we set the configurations of both
+	engine to be equivalent to the best of our capabilities, which comes with some limitations.
+	Firstly, both engines use a common implementation for the algorithms.
+	This is done to keep the theoretical runtime equivalent, however, in practice the
+	chosen implementation might not be optimal for either of the engines.
+	Secondly, we decided to use Lustre as a file system for both Spark and Dask.
+	Spark could benefit from using HDFS as a file system, however, HDFS is uncommon in HPC.
+	Finally, our infrastructure has a bottleneck from the network and Lustre.
+	As explained in section~{\ref{subsec:overhead-tradeoff}}, the data transfer time is tighly corrolated to the overhead of the application.
+	Therefore, while reducing this bottleneck should improve the performance of both engine similarly
+	it can be hard to predict the actual behavior.
+}
 

### 2022-02-23 latex/paper.tex
Minor fixes

@@
 In comparison, our study provides a detailed analysis of the performance differences
-and similarities\HL{, for HPC,} between Dask and Spark for data-intensive applications.
+and similarities between Dask and Spark for data-intensive applications \HL{executed on HPC systems}.
 
@@
 CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
-\HL{The compute and storage nodes are inter-connected by {\SI{10}{\giga\bit\per\second}} Ethernet link through a switch.
-	The login node, hosting an NFS, is also connected by a {\SI{10}{\giga\bit\per\second}} Ethernet link to the compute nodes.}
+\HL{The compute and storage nodes are inter-connected by {\SI{10}{\giga\bit\per\second}} Ethernet link through a Mellanox MSN2410 switch.
+	The login node, hosting an NFS server, is connected to the compute nodes through the same switch, with {\SI{10}{\giga\bit\per\second}} Ethernet.}
 
@@
 
-\HL{Our system uses similar hardware as Compute Canada Beluga, and replicates as
-	close as possible it's configuration.
-	This let us make a fair comparison of Spark and Dask in the condition of a realistic
+\HL{Our system uses similar hardware as Compute Canada's B\'eluga cluster --- a representative cluster available to researchers in Canada --- and replicates as
+	close as possible its configuration.
+	This let us compare Spark and Dask in the condition of a realistic
 HPC system.}
@@
 We decided to compute the nearest centroids in parallel since we use only three centroids.
-
 \HL{Due to resources limitation on our infrastructure, the Kmeans experiments
-	are performed with only a sample of half the BigBrain dataset ({\SI{303}{\gibi\byte}}).}
+	are performed with only half of the BigBrain dataset ({\SI{303}{\gibi\byte}}).}
 
@@
 While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
-\HL{Although Dask has a higher overhead, the total time for both engines is overall significantly similar.}
+\HL{Although Dask has a higher overhead, the total time for both engines is overall similar.}
 This is due to a tradeoff between disk bandwidth and compute/overhead time.
@@
 From Figure \ref{fig:increment_block} we also observe that Dask has significantly higher overhead than Spark.
-However, the total execution time of the application is \HL{significantly equivalent} due to the tradeoff between I/O and compute/overhead time.
+However, the total execution time of the application is \HL{equivalent} due to the tradeoff between I/O and compute/overhead time.
 \begin{figure}[!h]
@@
 In a previous study~\cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
-\HL{In this paper, the performance scale when increasing the number of nodes, although not linear due to an I/O bottleneck.
+\HL{In this paper, the performance scales when increasing the number of nodes, although not linearly due to an I/O bottleneck.
 	The system was designed such that the filesystem has similar speed to the network.

### 2022-02-23 latex/paper.tex
[UPDATE] Modify resutls and discussion about Kmeans + minor fixes.

--- 2022-02-23 latex/paper.tex
@@
 The results are similar to varying the number of nodes with Spark having faster compute time for \textit{Calculate\_histogram}.
-This results in a \HL{significant} makespan difference of up to 3\%.
+This results in a \HL{significant} makespan difference \HL{with Spark being up to 3\% faster}.
 \begin{figure}[!h]
@@
 \end{figure}
+
+\HL{
+	In a follow up experiement, we further optimized the Kmeans algorithm to minimize the size of data kept in memory.
+	Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}} show the gantt chart for the execution
+	of this new experiments for Dask and Spark.
+	The x-axis shows the time in second while the y-axis depicts the workers, with 
+	one worker thread ID (Dask) and process ID (Spark) per line.
+	Each task is represented on the gantt chart with the color of it's associated function.
+	We observe that Spark report 3 times more process than Dask threads.
+	However, we found out that both Dask and Spark use the same amount of maximum concurrent processes/threads.
+	This is most likely due to the difficulty to track process id within Spark.
+}
+
+\HL{
+	Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}}
+	show that Dask has a significantly shorther makespan than Spark, {\SI{338}{\second}}
+	and {\SI{409}{\second}} respectively.
+	The main differences in makespan for Spark appear from a longer initial load of the data
+	and a longer scheduling delay between the data load and the initial computation.
+}
 
@@
 \subsection{Memory management}
-Our Kmeans experiment shows a significant difference between Dask and Spark.
+Our Kmeans \HL{resutls from Figure~{\ref{fig:kmeans_worker}} show} a significant difference between Dask and Spark.
 The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
@@
 
+\subsection{Scheduling overhead}
+\HL{
+	Overall, our results show that Dask has a slightly longer startup overhead compared to Spark.
+	However, we found that this did not lead directly to makespan difference.
+	While both engines showed similar overhead for applications with low inter-worker communication,
+	Spark had significantly more scheduling overhead than Dask for the Kmeans applications.
+	Spark was considerably slower to schedule the first set of computation after reading the data on the workers.
+	Moreover, Spark read all the data before starting computation while was still reading data along side doing computation.
+	This difference in scheduling is beneficial for data-intensive applications as it reduce the risk
+	of I/O bottleneck, as explained in Section~{\ref{subsec:overhead-tradeoff}}.
+}
+
 \subsection{Comparison to other studies}
@@
 The experiments from~\cite{Mehta:17} found that Dask has a larger startup overhead than Spark, while another study~\cite{8588652} found the opposite.
-In this study, we found that Dask has a lower startup overhead than Spark; however, we did not in our previous study~\cite{8943502}.
+In this study, we found that Dask has a \HL{larger} startup overhead than Spark; however, we did not in our previous study~\cite{8943502}.
 Those apparent contractions most likely come from differences in infrastructure and configuration of the engines.
@@
 \section*{Acknowledgment}
-Mathieu Dugr\'e was funded by an Alexander Graham Bell scholarship from
+Mathieu Dugr\'e was funded by \HL{the} Alexander Graham Bell scholarship from
 the Natural Sciences and Engineering Research Council of Canada. This work

### 2022-02-24 latex/paper.tex
[UPDATE] Fix comment from Tristan.

--- 2022-02-23 latex/paper.tex
@@
 
+\HL{
+	We increase the worker memory by reducing the amount of worker process per compute nodes.
+	This was done because all available memory from the compute node is already allocated to the pool of workers.
+	When reducing the number of processes from 64 to 32 for Spark and Dask 
+	(effectively doubling the memory of each process), the difference in execution time between Spark and Dask reduced significantly.
+	That is, with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
+	We suppose that given more memory per thread/process, Dask and Spark would then perform similarly.
+}
+
 \begin{figure}[!h]
@@
 \HL{
-	In a follow up experiement, we further optimized the Kmeans algorithm to minimize the size of data kept in memory.
-	Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}} show the gantt chart for the execution
+	In a follow up experiement, we further optimized our Kmeans implementations, for both Dask and Spark, to minimize the size of data kept in memory.
+	Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}} show the Gantt chart for the execution
 	of this new experiments for Dask and Spark.
@@
 	one worker thread ID (Dask) and process ID (Spark) per line.
-	Each task is represented on the gantt chart with the color of it's associated function.
+	Each task is represented on the Gantt chart with the color of it's associated function.
 	We observe that Spark report 3 times more process than Dask threads.
-	However, we found out that both Dask and Spark use the same amount of maximum concurrent processes/threads.
-	This is most likely due to the difficulty to track process id within Spark.
+	However, we found out that both Dask and Spark use the same amount of maximum concurrent processes or threads.
 }
@@
 
-When reducing the number of processes from 64 to 32 for Spark (effectively doubling the memory of each process), \HL{the difference in execution time between Spark and Dask reduced significantly}.
-\HL{That is,} with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
-We suppose that given more memory per thread/process, Dask and Spark would then perform similarly.
-\HL{For our experiments, we allocated all available memory on the compute nodes.
-Therefore, it was not possible to allocate more resources to test this hypothesis.}
+\HL{
+	We optimized the memory usage in our kmeans implementation, for Dask and Spark,
+	and observed that with this improved implementation, Spark worker no longer failed.
+	Nevertheless, Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}} depicts that a substantial difference in makespan remained between Spark and Dask with this implementation,
+	showing that the previously-based gap was not only due to worker failures.
+}
 
@@
 Overall, our results showed no substantial difference between the engines.
-Although Spark overhead is lower than Dask, Dask performs I/O faster due to a tradeoff between overhead and I/O when the file system is saturated.
+\HL{Although Spark overhead was generally lower than Dask, it was significantly higher when applications incur inter-worker communication.}
 For one application, our results showed that Spark used more memory, causing issues with its workers resulting in a significantly faster runtime for Dask.

### 2022-08-17 latex/paper.tex
[REMOVE] highlight from previous revision.

--- 2022-02-24 latex/paper.tex
@@
 \abstract[Summary]{
-	The general increase in data size and data sharing motivates the adoption of Big Data strategies in a number of scientific disciplines.
-	However, while several options are available, no particular guidelines exist for the selection of a Big Data engine.
+	The general increase in data size and data sharing motivates the adoption of Big Data strategies in several scientific disciplines.
+	However, while several options are available, no particular guidelines exist for selecting a Big Data engine.
 	In this paper, we compare the runtime performance of two popular Big Data engines with
@@
 	We benchmark these applications on a dedicated HPC cluster running the Lustre file system while using varying
-	combinations of number of nodes, file size, and task duration.
+	combinations of the number of nodes, file size, and task duration.
 	Our results show that although there are slight differences between Dask and Spark,
@@
 A previous study\cite{8752675} noted the importance of in-memory computing and data locality to improve the performance of data-intensive pipelines.
-In this work, we study how performance benchmarks generalize across Big Data engines in an HPC environment.
+This work studies how performance benchmarks generalize across Big Data engines in an HPC environment.
 In particular, we study in-depth the performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16} for their suitability to process neuroimaging pipelines.
@@
 The scheduler for both engines operates dynamically, which benefits applications with runtime unknown ahead of time\cite{Dask:15}.
-They also provide rich high-level API and support a variety of schedulers and deployment environments, such as Mesos\cite{hindman2011mesos}, YARN\cite{vavilapalli2013apache}, Kubernetes, and HPC clusters.
+They also provide rich high-level APIs and support a variety of schedulers and deployment environments, such as Mesos\cite{hindman2011mesos}, YARN\cite{vavilapalli2013apache}, Kubernetes, and HPC clusters.
 Although sharing similarities, these engines have differences.
@@
 On the one hand, as part of the Scipy ecosystem, Dask provides almost transparent integration with APIs such as Numpy arrays, Pandas dataframe, or RAPIDS GPU accelerated code framework.
-On the other hand, Spark's Java, Scala, R, and Python APIs efficiently parallelize pipelines with minimal performance loss, due to the Java Virtual Machine.
+On the other hand, Spark's Java, Scala, R, and Python APIs efficiently parallelize pipelines with minimal performance loss due to the Java Virtual Machine.
 Our study is restrained to performance, although we recognize that other factors affect the choice of an engine in practice.
@@
 We installed the Slurm scheduler and Lustre file system~\cite{Braam2019TheLS} on a dedicated cluster to mimic a typical HPC system.
-\HL{Lustre is a reference file system for HPC environments, used in over 60 of the TOP100 supercomputers~{\cite{OpenSFS-lustre}}, offering parallel file accesses.}
+Lustre is a reference file system for HPC environments, used in over 60 TOP100 supercomputers~{\cite{OpenSFS-lustre}}, offering parallel file accesses.
 
-In a previous study\cite{8943502} we found that there was no significant performance difference between Apache Spark and Dask for data-intensive applications.
+In a previous study\cite{8943502} we found no significant performance difference between Apache Spark and Dask for data-intensive applications.
 However, using the network file system (NFS) led to an I/O bottleneck limiting application performance.
@@
 In comparison, our study provides a detailed analysis of the performance differences
-and similarities\HL{, for HPC,} between Dask and Spark for data-intensive applications.
+and similarities between Dask and Spark for data-intensive applications executed on HPC systems.
 
@@
 \subsection{Lustre}
-The Lustre file system is a parallel file system \HL{of reference in} HPC environments.
+The Lustre file system is a parallel file system of reference in HPC environments.
 Lustre offers a POSIX-compliant interface with the file system.
@@
 
-At the core of Lustre are the \textit{management servers} (MGSs) and \textit{metatdata server} (MDSs) in charge of the file operations and metadata,
-the \textit{object storage targets} (OSTs) to store file data, and the \textit{object sotrage server} (OSSs) to handle the OSTs.
+At the core of Lustre are the \textit{management servers} (MGSs) and \textit{metadata server} (MDSs) in charge of the file operations and metadata,
+the \textit{object storage targets} (OSTs) store file data, and the \textit{object storage servers} (OSSs) handle the OSTs.
 Lustre clients make requests to the MGSs to access files on the OSTs.
@@
 
-\HL{While other filesystem such as HDFS or cloud storage could be considered, we chose Lustre to best represent an HPC environment.
-	HDFS makes use of permanent local storage on compute node, which is usually not feasable in HPC.
-Morevoer, HPC node rarely have access to internet, limiting the use of cloud storage.}
+While other filesystems such as HDFS or cloud storage could be considered, we chose Lustre to represent an HPC environment best.
+HDFS uses permanent local storage on compute node, which is usually not feasible in HPC.
+Moreover, HPC nodes rarely have access to the internet, limiting cloud storage use.
 
@@
 Data locality, popularized by Map-Reduce~\cite{dean2008mapreduce}, schedules tasks where the data reside.
-In-memory computing minimizes the overhead of transferring data to disk by keeping data in memory when possible.
+In-memory computing minimizes the overhead of transferring data to disks by keeping data in memory.
 Dask uses lazy evaluation to reduce unnecessary communication and computation.
@@
 Arrays offer a clone of NumPy API for distributed processing of large arrays.
-Bags are a distributed collection of Python object that offers a programming abstraction similar to \href{https://toolz.readthedocs.io/en/latest/}{PyToolz}.
+Bags are a distributed collection of Python objects that offers a programming abstraction similar to \href{https://toolz.readthedocs.io/en/latest/}{PyToolz}.
 Dataframes are a parallel composition of \href{https://pandas.pydata.org/}{Pandas} Dataframes used to process a large amount of tabular data.
-Dask Delayed offers an API for distributing arbitrary functions that do not fit in the above frameworks.
+Dask Delayed offers an API for distributing arbitrary functions that do not fit the above frameworks.
 Lastly, Dask Futures can also execute arbitrary functions; however, it launches computation immediately rather than lazily.
@@
 The Dask engine is compatible with multiple distributed schedulers, including YARN and Mesos.
-It also provides its own scheduler, \textit{Dask Distributed scheduler}.
+It also provides its scheduler, \textit{Dask Distributed scheduler}.
 We chose to use Dask Distributed scheduler to keep the environments similar between Dask and Spark.
@@
 Jobs are divided into stages to execute onto workers.
-Each operation in a stage is represents by a high-level task in the computation graph.
+Each operation in a stage is represented by a high-level task in the computation graph.
 Like Dask, Spark Standalone scheduler uses a LIFO policy to schedule tasks.
@@
 While serialization from Python to Java, an operation required when using Spark's Python API, creates overhead, we found it minimal in a previous study~\cite{8943502}.
-We focus on PySpark API to have a similar environment between the different engines and due to its suitability for neuroimaging.
+We focus on PySpark API to have a similar environment between the different engines and  its suitability for neuroimaging.
 
@@
 CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
-\HL{The compute and storage nodes are inter-connected by {\SI{10}{\giga\bit\per\second}} Ethernet link through a switch.
-	The login node, hosting an NFS, is also connected by a {\SI{10}{\giga\bit\per\second}} Ethernet link to the compute nodes.}
+The compute and storage nodes are inter-connected by {\SI{10}{\giga\bit\per\second}} Ethernet link through a Mellanox MSN2410 switch.
+The login node, hosting an NFS server, is connected to the compute nodes through the same switch, with {\SI{10}{\giga\bit\per\second}} Ethernet.
 
@@
 Lustre \textit{2.14.0-1.el8} is installed on both the Lustre clients and servers.
-\HL{We used the default inode configuration for Lustre.}
+We used the default inode configuration for Lustre.
 
-\HL{Our system uses similar hardware as Compute Canada Beluga, and replicates as
-	close as possible it's configuration.
-	This let us make a fair comparison of Spark and Dask in the condition of a realistic
-HPC system.}
+Our system uses similar hardware as Compute Canada's B\'eluga cluster --- a representative cluster available to researchers in Canada --- and replicates as
+close as possible its configuration.
+This lets us compare Spark and Dask in the condition of a realistic
+HPC system.
 
@@
 The Spark network timeout is set to \SI{600}{\second} and the executor heartbeat to \SI{120}{\second}.
-A new Dask or Spark cluster was started and tore down for each experiment.
+A new Dask or Spark cluster was started and torn down for each experiment.
 Dask \textit{2021.7.0} and Spark \textit{3.1.2} were used.
@@
 NIfTI format, a popular format in neuroimaging. We left the NIfTI blocks
-uncompressed, resulting in total data size of \SI{606}{\gibi\byte}. To
+uncompressed, resulting in a total data size of \SI{606}{\gibi\byte}. To
 evaluate the effect of block size, we split the image into 1000, 2500,
@@
 \subsection{Applications}
-In this section, we describe the applications used to benchmark Dask and Spark.
+This section describes the applications used to benchmark Dask and Spark.
 We use three simple synthetic applications to have a deep understanding of the 
@@
 Also, two neuroimaging applications are used to study more realistic 
-applications, namely \textit{Kmeans} and \textit{BIDS App Example}\HL{{\cite{gorgolewski2017bids}}}.
+applications, namely \textit{Kmeans} and \textit{BIDS App Example}{\cite{gorgolewski2017bids}}.
 
@@
 
-\HL{The scripts used for our experiements and the results generated are available on
-	GitHub at: \mbox{\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}}.}
+The scripts used for our experiments and the results generated are available on
+GitHub at: \mbox{\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}}.
 
@@
 simulates computation by sleeping for a specified period (Fig.~\ref{fig:graph-increment}). To simulate
-intermediate results, we repeat \HL{a {\SI{1}{\second}} sleep delay during each incrementation}.
+intermediate results, we repeat a {\SI{1}{\second} sleep delay during each incrementation}.
 We prevent data caching of the blocks by incrementing their voxels
@@
 mimics the processing of multiple independent subjects in parallel.
-% \begin{figure}[!hb]
-% 	\centering
-% 	\includegraphics[height=0.5\columnwidth,
-% 	angle=0]{figures/increment.png}
-% 	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
-% 	\label{fig:graph-increment}
-% \end{figure}
 
@@
 updates each centroid with the average of its assigned voxels. It repeats
-the assignment and update steps for a configurable number of iterations.
+the assignment and updates steps for a configurable number of iterations.
 Finally, the voxels of the image blocks are classified and written back to
@@
 We decided to compute the nearest centroids in parallel since we use only three centroids.
-
-\HL{Due to resources limitation on our infrastructure, the Kmeans experiments
-	are performed with only a sample of half the BigBrain dataset ({\SI{303}{\gibi\byte}}).}
+Due to resources limitation on our infrastructure, the Kmeans experiments
+are performed with only half of the BigBrain dataset ({\SI{303}{\gibi\byte}}).
 
@@
 measure brain volume from MRIs. For this application, we use the CoRR
-dataset. The application extracts the brain volume of each participant,
+dataset. The application extracts the brain volume of each participant
 then computes the average for each group of participants. Unlike the other
@@
 the scalability of the engines' scheduler, (2) the number of files for the BigBrain
-in Increment, and Histogram to measure the
+in Increment and Histogram to measure the
 effect of different I/O patterns and parallelization degrees, and (3) the
@@
 It should be noted that increasing the number of iterations for a given sleep 
-delay also increases the total compute time of an application.
+delay also increases the total compute time.
 
@@
 As expected, the compute time, from \textit{Increment}, remains the same when we vary the number of nodes.
-However, the \textit{Read} and \textit{Write} time increase with the number of nodes for both engines.
+However, the \textit{Read} and \textit{Write} time increases with the number of nodes for both engines.
 This is explained by an I/O bottleneck on the Lustre file system.
@@
 While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
-\HL{Although Dask has a higher overhead, the total time for both engines is overall significantly similar.}
+Although Dask has a higher overhead, the total time for both engines is overall similar.
 This is due to a tradeoff between disk bandwidth and compute/overhead time.
@@
 On the other hand, when most threads are either idle or performing computations, it reduces the contention on the disk bandwidth resulting in shorter I/O time.
-\HL{The only significant difference in total time is when two nodes are used.
-	This results in an approximately 3\% total runtime difference.}
+The only significant difference in total time is when two nodes are used.
+This results in an approximately 3\% total runtime difference.
 \begin{figure}[!h]
@@
 
-Figure \ref{fig:increment_itr} also shows that the the total I/O time decreases as total computation time increases.
+Figure \ref{fig:increment_itr} also shows that the total I/O time decreases as total computation time increases.
 This is explained by an I/O bottleneck on the Lustre file system.
@@
 From Figure \ref{fig:increment_block} we also observe that Dask has significantly higher overhead than Spark.
-However, the total execution time of the application is \HL{significantly equivalent} due to the tradeoff between I/O and compute/overhead time.
+However, the total execution time of the application is equivalent due to the tradeoff between I/O and compute/overhead time.
 \begin{figure}[!h]
@@
 Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} application when varying the number of nodes.
-Overall, \HL{the execution time for Spark is significantly lower than for Dask by up to 5\%}, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
+Overall, the execution time for Spark is significantly lower than for Dask by up to 5\%, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
 This is unexpected as the code base for both engines is the same, with only the calling APIs differing.
@@
 The results are similar to varying the number of nodes with Spark having faster compute time for \textit{Calculate\_histogram}.
-This results in a \HL{significant} makespan difference \HL{with Spark being up to 3\% faster}.
+This results in a significant makespan difference with Spark being up to 3\% faster.
 \begin{figure}[!h]
@@
 
-\HL{
-	We increase the worker memory by reducing the amount of worker process per compute nodes.
-	This was done because all available memory from the compute node is already allocated to the pool of workers.
+We increase the worker memory by reducing the number of worker processes per
+compute node since all available memory from the compute node is already allocated to the pool of workers.
 When reducing the number of processes from 64 to 32 for Spark and Dask 
-	(effectively doubling the memory of each process), the difference in execution time between Spark and Dask reduced significantly.
-	That is, with only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
+(effectively doubling the memory of each process), the difference in execution time between Spark and Dask was reduced significantly.
+With only 32 processes and more memory per worker, Spark no longer recomputed tasks for the Kmeans application.
 We suppose that given more memory per thread/process, Dask and Spark would then perform similarly.
-}
 
@@
 
-\HL{
-	In a follow up experiement, we further optimized our Kmeans implementations, for both Dask and Spark, to minimize the size of data kept in memory.
+In a follow-up experiment, we further optimized our Kmeans implementations for both Dask and Spark to minimize the size of data kept in memory.
 Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}} show the Gantt chart for the execution
@@
 one worker thread ID (Dask) and process ID (Spark) per line.
-	Each task is represented on the Gantt chart with the color of it's associated function.
-	We observe that Spark report 3 times more process than Dask threads.
+We represent each task on the Gantt chart with its associated function's color.
+We observe that Spark reports three times more processes than Dask threads.
 However, we found out that both Dask and Spark use the same amount of maximum concurrent processes or threads.
-}
 
-\HL{
 Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}}
@@
 and a longer scheduling delay between the data load and the initial computation.
-}
 
@@
 
-For brevity we omitted the figure\HL{s} when varying the number of iterations as the results are similar to varying the number of workers.
+For brevity we omitted the figures when varying the number of iterations as the results are similar to varying the number of workers.
 
@@
 Figure \ref{fig:bids} shows the total execution time for the \textit{BIDS App Example} application when varying the number of nodes.
-We observe that Dask has slightly lower overhead, thank Spark, although not significant.
+We observe that Dask has a slightly lower overhead, thank Spark, although not significant.
 However, the makespan for Dask is between 7.5\% to 14.5\% lower than for Spark.
@@
 \label{subsec:io-bottleneck}
-We used Lustre as shared file system as it is a \HL{de facto} solution in HPC systems.
+We used Lustre as shared file system as it is a de facto solution in HPC systems.
 The performance of our Lustre installation degrades when a large number of concurrent tasks attempt to perform I/O to the shared file system, as shown in Figures \ref{fig:increment_worker} and \ref{fig:histogram_worker}.
 In a previous study~\cite{8943502}, we observed nearly no improvement when increasing the number of nodes to process an application while using an NFS.
-\HL{In this paper, the performance scale when increasing the number of nodes, although not linear due to an I/O bottleneck.
-	The system was designed such that the filesystem has similar speed to the network.
-Upgrading to more performant networking and storage compononents would reduce the bottleneck seen when increasing the number of workers.}
+In this paper, the performance scales when increasing the number of nodes, although not linearly due to an I/O bottleneck.
+The system was designed such that the filesystem has a similar speed to the network.
+Upgrading to more performant networking and storage components would reduce the bottleneck when increasing the number of workers.
 
-\HL{
 As in our previous study~{\cite{8943502}}, we observe that the I/O bottleneck has a lower impact on more compute-intensive applications, such as Kmeans and BIDS App Example.
 However, it significantly affects data-intensive applications, such as Increment and Histogram.
-}
 
-These results suggest that increasing the parallelism of the file system \HL{and speed of the network} is critical when scaling the compute resources to process a data-intensive application.
+These results suggest that increasing the parallelism of the file system and speed of the network is critical when scaling the compute resources to process a data-intensive application.
 Otherwise, the compute resource might be underutilized.
@@
 To take an extreme example, the makespan of $n$ concurrent file transfer of size $F$ from a file system with a bandwidth $B$ would be $nF/B$,
-as the $n$ transfer would share the bandwidth $B$, which is equivalent to sequentially transfer the $n$ files.
+as the $n$ transfer would share the bandwidth $B$, which is equivalent to sequentially transferring the $n$ files.
 i.e., when the bandwidth of the file system is saturated, engine overhead does not increase the makespan as long as it desynchronizes the data transfers.
 In a previous study~\cite{8943502}, we also observed this tradeoff between overhead and data transfer when the file system was saturated.
-It could be interesting to explore strategies that take advantage of this tradeoff to reduce the application makespan.
+Exploring strategies that take advantage of this tradeoff could be interesting to reduce the application makespan.
 
 \subsection{Memory management}
-Our Kmeans \HL{resutls from Figure~{\ref{fig:kmeans_worker}} show} a significant difference between Dask and Spark.
+Our Kmeans resutls from Figure~{\ref{fig:kmeans_worker}} show a significant difference between Dask and Spark.
 The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
-This is unexpected, given that Dask and Spark are configured similarly and \HL{are allocated} the same amount of resources.
+This is unexpected, given that Dask and Spark are configured similarly and are allocated the same amount of resources.
 We think those issues arise from a combination of
@@
 (2) the serialization of Python to Java kept in memory,
-(3) Spark only using procesess compared to Dask using threads, thus limiting shared memory consumption,
+(3) Spark using processes compared to Dask using threads, thus limiting shared memory consumption,
 and (4) different policies to persist data to disk when running out of memory.
 
-\HL{
-	We optimized the memory usage in our kmeans implementation, for Dask and Spark,
-	and observed that with this improved implementation, Spark worker no longer failed.
+We optimized the memory usage in our Kmeans implementation for Dask and Spark
+and observed that the Spark workers no longer failed with this improved implementation.
 Nevertheless, Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}} depicts that a substantial difference in makespan remained between Spark and Dask with this implementation,
 showing that the previously-based gap was not only due to worker failures.
-}
 
 \subsection{Scheduling overhead}
-\HL{
-	Overall, our results show that Dask has a slightly longer startup overhead compared to Spark.
-	However, we found that this did not lead directly to makespan difference.
+Overall, our results show that Dask has a slightly longer startup overhead than Spark.
+However, we found that this did not lead directly to the makespan difference.
 While both engines showed similar overhead for applications with low inter-worker communication,
 Spark had significantly more scheduling overhead than Dask for the Kmeans applications.
-	Spark was considerably slower to schedule the first set of computation after reading the data on the workers.
-	Moreover, Spark read all the data before starting computation while was still reading data along side doing computation.
-	This difference in scheduling is beneficial for data-intensive applications as it reduce the risk
+Spark was considerably slower to schedule the first set of computations after reading the data on the workers.
+Moreover, Spark read all the data before starting computation while was still reading data alongside doing computation.
+This difference in scheduling is beneficial for data-intensive applications as it reduces the risk
 of I/O bottleneck, as explained in Section~{\ref{subsec:overhead-tradeoff}}.
-}
 
@@
 In contradiction with the experiments of~\cite{Mehta:17}, we found that Dask had a larger overhead than Spark, especially when scaling the number of workers.
-However, this was compensated by the overhead and data transfer tradeoff mentioned previously.
+However, this was compensated by the previously mentioned overhead and data transfer tradeoff.
 Since our focus is on data-intensive applications, we conclude that the performance of Dask and Spark is equivalent.
@@
 The experiments from~\cite{Mehta:17} found that Dask has a larger startup overhead than Spark, while another study~\cite{8588652} found the opposite.
-In this study, we found that Dask has a \HL{larger} startup overhead than Spark; however, we did not in our previous study~\cite{8943502}.
+In this study, we found that Dask has a larger startup overhead than Spark; however, we did not in our previous study~\cite{8943502}.
 Those apparent contractions most likely come from differences in infrastructure and configuration of the engines.
@@
 Therefore, it is critical to ensure enough bandwidth to the file system before scaling the compute resources for data-intensive applications.
-Otherwise, those extras resources might be underused.
+Otherwise, those extra resources might be underused.
 
 \subsection{Limitations}
-\HL{
-	To make a fair comparison between Spark and Dask, we set the configurations of both
-	engine to be equivalent to the best of our capabilities, which comes with some limitations.
-	Firstly, both engines use a common implementation for the algorithms.
-	This is done to keep the theoretical runtime equivalent, however, in practice the
-	chosen implementation might not be optimal for either of the engines.
+To make a fair comparison between Spark and Dask, we set both engine configurations
+to be equivalent to the best of our capabilities, which comes with some limitations.
+Firstly, both engines use a shared implementation for the algorithms.
+This is done to keep the theoretical runtime equivalent; however, the chosen
+implementation might not be optimal for either of the engines.
 Secondly, we decided to use Lustre as a file system for both Spark and Dask.
-	Spark could benefit from using HDFS as a file system, however, HDFS is uncommon in HPC.
+Spark could benefit from using HDFS as a file system, but HDFS is uncommon in HPC.
 Finally, our infrastructure has a bottleneck from the network and Lustre.
-	As explained in section~{\ref{subsec:overhead-tradeoff}}, the data transfer time is tighly corrolated to the overhead of the application.
-	Therefore, while reducing this bottleneck should improve the performance of both engine similarly
-	it can be hard to predict the actual behavior.
-}
+As explained in section~{\ref{subsec:overhead-tradeoff}}, the data transfer 
+time is tightly correlated to the overhead of the application.
+Therefore, while reducing this bottleneck should improve the performance of 
+both engines similarly, it can be hard to predict the actual behavior.
 
@@
 Overall, our results showed no substantial difference between the engines.
-\HL{Although Spark overhead was generally lower than Dask, it was significantly higher when applications incur inter-worker communication.}
+Although Spark overhead was generally lower than Dask, it was significantly higher when applications incur inter-worker communication.
 For one application, our results showed that Spark used more memory, causing issues with its workers resulting in a significantly faster runtime for Dask.
@@
 \section*{Acknowledgment}
-Mathieu Dugr\'e was funded by \HL{the} Alexander Graham Bell scholarship from
-the Natural Sciences and Engineering Research Council of Canada. This work
-was partially funded by the Canada Research Chairs program. The HPC cluster
-was funded by the Canada Foundation for Innovation.
+Mathieu Dugr\'e was funded by the Alexander Graham Bell scholarship from
+the Natural Sciences and Engineering Research Council of Canada.
+The Canada Research Chairs program partially funded this work.
+The Canada Foundation for Innovation funded the HPC cluster.
 
-\bibliographystyle{wileyNJD-AMA}
 \bibliography{paper}

### 2022-08-25 latex/paper.tex
[FIX] first draft to address second reviewers comments.

@@
 \usepackage{subcaption}
+\usepackage{tabularx}
 
@@
 
-\title{Performance comparison of Dask and Apache Spark on HPC systems}
+\title{Performance comparison of Dask and Apache Spark on HPC systems \HL{for Neuroimaging}}
 
@@
 	Python APIs, Apache Spark, and Dask, in processing neuroimaging pipelines.
-	Our experiments use three synthetic data-intensive applications to process
+	Our experiments use three synthetic \HL{neuroimaging} applications to process
 	the \SI{606}{\gibi\byte} BigBrain image and an actual pipeline to process data
@@
 A previous study\cite{8752675} noted the importance of in-memory computing and data locality to improve the performance of data-intensive pipelines.
-This work studies how performance benchmarks generalize across Big Data engines in an HPC environment.
+This work studies how performance benchmarks \HL{of neuroimaging applications} generalize across Big Data engines in an HPC environment.
+\HL{
+	With the rising need for Big Data engines in neuroimaging and none commonly used
+	in the field, it is important to evaluate performance of different engines.
+}
 In particular, we study in-depth the performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16} for their suitability to process neuroimaging pipelines.
-Our objective is to assess whether Spark or Dask has substantial performance benefits in processing data-intensive HPC pipelines.
+\HL{
+	Our main contribution is to assess whether Spark or Dask has substantial performance benefits in processing neuroimaging pipelines in HPC systems.
+}
 
@@
 
-While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructures.
-This paper focuses on the performance of Big Data engines in HPCs environments.
+While laptops or workstations can be sufficient for some applications, 
+neuroimaging pipelines often require large\HL{r} infrastructures
+\HL{to process datasets in the scale of several hundreds of {\SI{}{\gibi\byte}} to few {\SI{}{\tebi\byte}}.}
+This paper focuses on the performance of Big Data engines in HPCs environments \HL{for neuroimaging}.
 We installed the Slurm scheduler and Lustre file system~\cite{Braam2019TheLS} on a dedicated cluster to mimic a typical HPC system.
@@
 On the contrary, the work in~\cite{10.1145/3225058.3225128} shows that with large datasets, Spark provides better speedup than Dask.
-However, that study used the Dask Bag API, which has performance limitations compared to other Dask APIs.
+However, that study used the Dask Bag API, which has performance limitations compared to other Dask APIs\HL{{~\cite{DaskBagDoc}}}.
 
@@
 Next, we present the design of our benchmarking experiments.
-We consider two types of data-intensive neuroimaging pipelines: high-resolution imaging and functional MRI studies for a cohort of subjects.
+We consider two types of neuroimaging pipelines: high-resolution imaging and functional MRI studies for a cohort of subjects.
 The pipelines we chose involve different compute and I/O patterns, e.g., map-reduce or map-only, requiring partial data in memory or the whole dataset.
@@
 
-Our system uses similar hardware as Compute Canada's B\'eluga cluster --- a representative cluster available to researchers in Canada --- and replicates as
-close as possible its configuration.
-This lets us compare Spark and Dask in the condition of a realistic
-HPC system.
+\HL{
+	While our infrastructure is not of the scale of TOP500 HPC systems,
+	it is hard to reserved multiple nodes concurrently on those infrastructure.
+	Moreover, by using our infrastructure, we can assure that no external processes
+	are influencing the benchmark for our experiements.
+	Our system uses similar hardware as Compute Canada's B\'eluga cluster
+	--- a representative cluster available to researchers in Canada --- and 
+	replicates as close as possible its configuration.
+	This lets us compare Spark and Dask in an isolated environement with the condition 
+	of a realistic  allocation of resources, for neuroimaging applications, in an HPC system.
+}
 
-Both Spark and Dask are configured to have eight worker processes, each with eight threads.
+\HL{
+	We configure Dask to have eight worker processes per node, each with eight threads.
+	This is the default recommended configuration for the Dask SLURM cluster configuration~{\cite{DaskSLURMDoc}}.
+	Similarly, we configured Spark to have eight executors per node, each with eight threads.
+	When trying different configuration for the workers, we did not observe performance difference
+	for our benchmark.
+	Therefore we chose the recommended configuration from Dask.
+}
 Each worker was allocated \SI{31.5}{\gibi\byte} of memory.
@@
 Other Dask configurations are left to default.
-The Spark network timeout is set to \SI{600}{\second} and the executor heartbeat to \SI{120}{\second}.
 A new Dask or Spark cluster was started and torn down for each experiment.
@@
 
+\HL{
+	We chose these dataset as they represent common dataset size for neuroimaging 
+	imaging studies and for their free access.
+}
+
 \subsection{Applications}
@@
 applications, namely \textit{Kmeans} and \textit{BIDS App Example}{\cite{gorgolewski2017bids}}.
+\HL{
+	We chose these neuroimaging data processing applications to study the 
+	effects from various Big Data characteristics.
+	Table~{\ref{table:applications}} depicts the Big Data characteristics used in
+	each application.
+}
 
-To have a fair comparison between the engines, the code base for the core functions of the applications is the same for both engines.
-This allows us to reduce the performance difference from different implementations.
+\begin{table*}[t]
+	\caption{Big Data characteristics of the applications}
+	\centering
+	\begin{tabular}{l|c|c|c|c|c|c|c|}
+		\cline{2-8}
+		\multicolumn{1}{c|}{} &
+		\textbf{Map} &
+		\textbf{Reduce} &
+		\textbf{\begin{tabular}[c]{@{}c@{}}Data\\ shuffling\end{tabular}} &
+		\textbf{\begin{tabular}[c]{@{}c@{}}Dataset \\ in-memory\end{tabular}} &
+		\textbf{\begin{tabular}[c]{@{}c@{}}Compute\\ intensive\end{tabular}} &
+		\textbf{\begin{tabular}[c]{@{}c@{}}Data\\ intensive\end{tabular}} &
+		\textbf{Containerized} \\ \hline
+		\multicolumn{1}{|l|}{\textbf{Increment}}        & X &   &         & Partial &   & X &   \\ \hline
+		\multicolumn{1}{|l|}{\textbf{Histogram}}        & X & X & Minimal & Partial &   & X &   \\ \hline
+		\multicolumn{1}{|l|}{\textbf{Kmeans}}           & X &   & X       & Whole   & X &   &   \\ \hline
+		\multicolumn{1}{|l|}{\textbf{BIDS App Example}} & X & X & Minimal & Partial & X &   & X \\ \hline
+	\end{tabular}
+	\label{table:applications}
+\end{table*}
+
+\HL{
+	To minimize the performance difference from implementation decisions, the 
+	code base of the applications is common for both engines.
+	Therefore, the performance observed comes directly from the engine scheduling and
+	their APIs.
+}
 
@@
 \subsubsection{Histogram}
-As our third application, we calculate the histogram of the BigBrain image. 
+As our second application, we calculate the histogram of the BigBrain image. 
 The application reads the BigBrain blocks from Lustre, calculates each intensity's frequency, and then writes the aggregated result back on Lustre (Fig.~\ref{fig:graph-histogram}).
@@
 	\centering
-	\begin{minipage}{.45\textwidth}
+	\begin{minipage}{0.4\textwidth}
 		\centering
-		\includegraphics[width=0.45\linewidth]{figures/increment.png}
+		\includegraphics[width=\linewidth]{figures/increment.png}
 		\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
@@
 	\end{minipage}
-	\begin{minipage}{.45\textwidth}
+	\begin{minipage}{0.55\textwidth}
 		\centering
-		\includegraphics[width=0.45\linewidth]{figures/histogram.png}
+		\includegraphics[width=\linewidth]{figures/histogram.png}
 		\caption{Task graph for Histogram with 3 BigBrain blocks.}
@@
 \subsubsection{Kmeans}
-For our fourth application, we apply Kmeans clustering to the voxel
+For our third application, we apply Kmeans clustering to the voxel
 intensities of the BigBrain image. We set the number of clusters to 3. The
@@
 We decided to compute the nearest centroids in parallel since we use only three centroids.
-Due to resources limitation on our infrastructure, the Kmeans experiments
+However, due to resources limitation on our infrastructure, the Kmeans experiments
 are performed with only half of the BigBrain dataset ({\SI{303}{\gibi\byte}}).
+\HL{We chose this applications for its real-life use case, large data shuffling between workers,
+and requirements to keep all data in-memory.}
 
 \subsubsection{BIDS App Example}
-Our fifth application is the BIDS App Example, a neuroimaging pipeline to
+Our fourth application is the BIDS App Example, a neuroimaging pipeline to
 measure brain volume from MRIs. For this application, we use the CoRR
@@
 \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
+\HL{We chose this application for its real-life use case and it is containerized unlike our applications.}
 
@@
 throughout the experiments. We varied (1) the number of nodes to assess
-the scalability of the engines' scheduler, (2) the number of files for the BigBrain
-in Increment and Histogram to measure the
-effect of different I/O patterns and parallelization degrees, and (3) the
+the scalability of the engines' scheduler, (2) the number of files \HL{to segment}
+BigBrain to measure the effect of different I/O patterns and parallelization degrees, and (3) the
 number of iterations to evaluate the effect of the number of tasks.
@@
 delay also increases the total compute time.
+\HL{
+	The choice of parameters for our experiements was restrained by the datasets
+	and resources available from the cluster.
+	The unrestrained parameters were chosen to have an expected linear scaling 
+	to facilitate the interpretation of the results.
+}
 
@@
 However, the \textit{Read} and \textit{Write} time increases with the number of nodes for both engines.
-This is explained by an I/O bottleneck on the Lustre file system.
+\HL{
+	With a large amount of worker threads accessing the Lustre file system
+	concurrently, the disk bandiwdth or network are saturated and it create 
+	an I/O bottleneck to access the Lustre file system.
+}
 
@@
 This is due to a tradeoff between disk bandwidth and compute/overhead time.
-On the one hand, multiple threads access Lustre concurrently, thus sharing the disk bandwidth resulting in longer I/O time.
-On the other hand, when most threads are either idle or performing computations, it reduces the contention on the disk bandwidth resulting in shorter I/O time.
+\HL{
+	On the one hand, when multiple threads access Lustre concurrently they risk 
+	of saturation the shared disk bandwidth, thus resulting in longer I/O time.
+	On the other hand, when most threads are either idle or performing computations,
+	it reduces the contention on the disk bandwidth resulting in shorter I/O time.
+}
 The only significant difference in total time is when two nodes are used.
@@
 This is unexpected as the code base for both engines is the same, with only the calling APIs differing.
-It is most likely due to the use of Dask Bags, which are slower than the other Dask APIs.
+It is most likely due to the use of Dask Bags, which are slower than the other Dask APIs\HL{~{\cite{DaskBagDoc}}}.
 As with previous experiments, the overhead for Spark is less and scales better, although compensated by a longer total I/O time.
@@
 We observe that Spark reports three times more processes than Dask threads.
-However, we found out that both Dask and Spark use the same amount of maximum concurrent processes or threads.
+\HL{
+	It is unclear what is the cause for reporting extra Spark processes.
+	However, with our analysis we found that both Dask and Spark use the same 
+	amount of maximum concurrent processes or threads.
+	Additionally, the is no significant difference in the average amount of 
+	concurrent processes or threads between Dask and Spark.
+}
 
@@
 	\centering
-	\includegraphics[clip,width=\columnwidth]{figures/dask_kmeans_gantt.png}
+	\href{https://mathdugre.me/paper-big-data-engines/ccpe/dask_kmeans_gantt.html}
+	{\includegraphics[clip,width=\columnwidth]{figures/dask_kmeans_gantt.png}}
 	\caption{Kmeans: Dask with 8 nodes, 5000 files, and 3 iterations}
@@
 	\centering
-	\includegraphics[clip,width=\columnwidth]{figures/spark_kmeans_gantt.png}
+	\href{https://mathdugre.me/paper-big-data-engines/ccpe/spark_kmeans_gantt.html}
+	{\includegraphics[clip,width=\columnwidth]{figures/spark_kmeans_gantt.png}}
 	\caption{Kmeans: Spark with 8 nodes, 5000 files, and 3 iterations}
@@
 Moreover, although minimal, a shorter cluster deployment time for Dask reinforces the difference in overhead time.
-As with other experiments, the I/O from Lustre affects the scaling of both engines as we increase the number of nodes.
 \begin{figure}[!h]
@@
 as the $n$ transfer would share the bandwidth $B$, which is equivalent to sequentially transferring the $n$ files.
-i.e., when the bandwidth of the file system is saturated, engine overhead does not increase the makespan as long as it desynchronizes the data transfers.
 In a previous study~\cite{8943502}, we also observed this tradeoff between overhead and data transfer when the file system was saturated.
@@
 \subsection{Memory management}
-Our Kmeans resutls from Figure~{\ref{fig:kmeans_worker}} show a significant difference between Dask and Spark.
+Our Kmeans results from Figure~{\ref{fig:kmeans_worker}} show a significant difference between Dask and Spark.
 The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
@@
 Spark was considerably slower to schedule the first set of computations after reading the data on the workers.
-Moreover, Spark read all the data before starting computation while was still reading data alongside doing computation.
+Moreover, Spark read all the data before starting computation while Dask was still reading data alongside doing computation.
 This difference in scheduling is beneficial for data-intensive applications as it reduces the risk
@@
 However, this was compensated by the previously mentioned overhead and data transfer tradeoff.
-Since our focus is on data-intensive applications, we conclude that the performance of Dask and Spark is equivalent.
+Since our focus is on \HL{neuroimaging} applications, we conclude that the performance of Dask and Spark is equivalent.
 
-The experiments from~\cite{Mehta:17} found that Dask has a larger startup overhead than Spark, while another study~\cite{8588652} found the opposite.
-In this study, we found that Dask has a larger startup overhead than Spark; however, we did not in our previous study~\cite{8943502}.
+\HL{
+	In this study, we observed that Dask has a larger startup overhead that Spark.
+	These results are consistent with a previous study~{\cite{Mehta:17}}.
+	However, other studies found the opposite~{\cite{8588652}} or that the difference was negligible~{\cite{8943502}}
+}
 Those apparent contractions most likely come from differences in infrastructure and configuration of the engines.
@@
 Otherwise, those extra resources might be underused.
+
+\subsection{Recommendations}
+\iffalse
+The differences in performance are minimal between both engines. However, 
+we added a “Recommendation” subsection to the discussion for qualitative 
+difference and some caveats to avoid when using the engines.
+\fi
+\HL{
+	The results from our experiments show that the performance difference between
+	Dask and Spark is minimal, in general. 
+	The main differences we observed are qualitative.
+	We found that Dask integration with the scientific Python ecosystem made it
+	easier to implement neuroimaging application.
+	Spark configuration required to set multiple environment variables.
+	In comparison, Dask is ready-to-go after an installation through PIP.
+	Dask dashboard provides more information than Spark ones, making
+	the debugging of Dask application easier and faster.
+	Therefore, for the neuroimaging field, we recommend the use of Dask over Spark.
+	However, in other fields with non-Python-based ecosystem, both Dask and spark
+	would be equal candidate as a Big Data engine.
+}
+
+\HL{
+	% TODO add caveats to avoid when using the engines.
+}
 
@@
 
+\HL{
+	Dask and Spark both support GPUs, however we did not use applications that
+	leverage them in our work.
+	Although several neuroimaging applications use GPUs, only a few preprocessing pipelines use them.
+	With preprocessing comprising the majority of compute time for neuroimaging
+	workflow, we ommitted GPUs from this study.
+	Moreover, the few preprocessing applications using GPUs are not commonly used
+	due to their recent development.
+}
+
 \section{Conclusion}
 We presented a detailed comparison of two Big Data engines, Apache Spark and Dask.
-We studied the engines using three synthetic data-intensive applications and a data-intensive neuroimaging pipeline.
+We studied the engines using three synthetic \HL{neuroimaging} applications and 
+a neuroimaging pipeline \HL{to calculate brain volume.}
 Overall, our results showed no substantial difference between the engines.

### 2022-08-25 plot-script/plot.py

@@ -27,6 +27,9 @@ import pandas as pd
 from scipy.stats import ttest_ind
 
 
+matplotlib.rc('font', size=20)
+
+
 def gantt(
     df,
     xlabel="Time [s]",
@@ -175,6 +178,11 @@ def gantt(
     plot.yaxis.major_tick_line_color = None
     plot.yaxis.minor_tick_line_color = None
     plot.yaxis.major_label_text_font_size = "0pt"
+    
+    plot.xaxis.major_label_text_font_size = "20pt"
+    plot.axis.axis_label_text_font_size = "20pt"
+    plot.legend.title_text_font_size = "20pt"
+    plot.legend.label_text_font_size = "20pt"
 
     # Hover tool
     hover = HoverTool(

### 2022-09-09 latex/paper.tex
[TMP] [WIP] Current writing.

@@
 	This is the default recommended configuration for the Dask SLURM cluster configuration~{\cite{DaskSLURMDoc}}.
-	Similarly, we configured Spark to have eight executors per node, each with eight threads.
+	Similarly, we configured Spark to have eight executors per node (hereafter called workers for consistency with Dask), each with eight threads.
 	When trying different configuration for the workers, we did not observe performance difference
@@
 \subsection{Kmeans}
-Figure \ref{fig:kmeans_worker} depicts the total execution time for the \textit{Kmeans} application when varying the number of workers.
-We observe that Dask is multifold faster than Spark, resulting in a $1.79\times$ to $2.71\times$ speed-up ratio on the makespan.
-Specifically, Spark has much longer overhead, \textit{get\_labels} and read time than Dask.
-This difference arises from Spark workers restarting due to lack of executor memory, leading to recomputation of tasks.
+\HL{
+	When varying the number of workers, we observed that Dask is multifold faster
+	than Spark, resulting in a $1.79\times$ to $2.71\times$ speed-up ratio on the makespan.
+	The problem we observed was two fold.
+	Firstly, the process ID of the workers' threads is not reported correctly and 
+	leading to falsely excessive overhead for Spark.
+	Secondly, the Spark workers were restarting due to lack of memory, leading 
+	to task recomputation and having to reload data into memory.
+}
 
@@
 
-\begin{figure}[!h]
-	\centering
-	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_kmeans_worker.jpg}
-	\caption{Kmeans: varying nodes -- 5000 files, 3 iterations}
-	\label{fig:kmeans_worker}
-\end{figure}
-
 In a follow-up experiment, we further optimized our Kmeans implementations for both Dask and Spark to minimize the size of data kept in memory.
@@
 We represent each task on the Gantt chart with its associated function's color.
-We observe that Spark reports three times more processes than Dask threads.
+We again observe that Spark reports three times more processes than Dask threads.
 \HL{
@@
 \subsection{Memory management}
-Our Kmeans results from Figure~{\ref{fig:kmeans_worker}} show a significant difference between Dask and Spark.
-The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
-This is unexpected, given that Dask and Spark are configured similarly and are allocated the same amount of resources.
+% TODO remore reference to figure.
+\HL{Our initial Kmeans results showe a significant difference between Dask and Spark.}
+The difference in performance \HL{came} from Spark worker restarting due to running out of memory; thus, Spark \HL{needed} to recompute several tasks.
+This is unexpected, given that Dask and Spark \HL{were} configured similarly and \HL{were} allocated the same amount of resources.
 We think those issues arise from a combination of
@@
 \subsection{Recommendations}
-\iffalse
-The differences in performance are minimal between both engines. However, 
-we added a “Recommendation” subsection to the discussion for qualitative 
-difference and some caveats to avoid when using the engines.
-\fi
 \HL{
@@
 	% TODO add caveats to avoid when using the engines.
+
 }
@@
 Spark could benefit from using HDFS as a file system, but HDFS is uncommon in HPC.
-Finally, our infrastructure has a bottleneck from the network and Lustre.
+\HL{Finally, our Lustre filesystem suffer from an I/O bottleneck coming from 
+either the network or disks bandwidth.}
 As explained in section~{\ref{subsec:overhead-tradeoff}}, the data transfer 
 time is tightly correlated to the overhead of the application.
-Therefore, while reducing this bottleneck should improve the performance of 
+However, while reducing this bottleneck should improve the performance of 
 both engines similarly, it can be hard to predict the actual behavior.
+
+\HL{
+	% TODO discuss UCX and other high performance networks.???
+
+}
 

### 2022-09-10 latex/paper.tex
[TMP] [WIP] fixed more comments.

@@
 \HL{
-	% TODO add caveats to avoid when using the engines.
-
+	% TODO reread this paragraph.
+	While the engines showed similar performance, they both come with their ow caveat that
+	can drastically affect their performance.
+	For Dask, the user should be aware of the Python GIL limitations.
+	For Python code multiples worker processes should be used and for code releasing the
+	GIL a balance between processes and thread is recommended.
+	For Spark, we experienced severe performance issues due to workers running out of memory.
+	Users should pay careful attetion to their Spark worker configurations to prevent similar issues.
 }
@@
 \HL{
-	% TODO discuss UCX and other high performance networks.???
-
+	% TODO reread paragraph
+	While both Dask and Spark support various high-performance network strategies
+	such as UCX, we focused on the stand point on neuroscientist that would 
+	commonly use the frameworks mostly out-of-the-box.
+	With higher network throughput, the overall performance of the engines could differ.
+	Further work would be required to compare Dask and Spark using different high-performance
+	network protocols.
 }

### 2022-09-12 latex/paper.tex
small fixes

--- 2022-08-25 latex/paper.tex
@@
 \newcommand{\HL}[1]{\hl{#1}}
+\newcommand{\TG}[1]{\color{red}\textsc{From Tristan: }#1\color{black}}
 
@@
 \HL{
-	Our main contribution is to assess whether Spark or Dask has substantial performance benefits in processing neuroimaging pipelines in HPC systems.
+	Our main contribution is to assess whether Spark or Dask has any substantial performance advantage in processing neuroimaging pipelines in HPC systems.
 }
@@
 	While our infrastructure is not of the scale of TOP500 HPC systems,
-	it is hard to reserved multiple nodes concurrently on those infrastructure.
-	Moreover, by using our infrastructure, we can assure that no external processes
-	are influencing the benchmark for our experiements.
+	we can assure that no external processes
+	impacted our measurements which would be hard to achieve on a production system.
 	Our system uses similar hardware as Compute Canada's B\'eluga cluster
 	--- a representative cluster available to researchers in Canada --- and 
-	replicates as close as possible its configuration.
+	replicates as closely as possible its configuration.
 	This lets us compare Spark and Dask in an isolated environement with the condition 
@@
 \HL{
-	We chose these dataset as they represent common dataset size for neuroimaging 
-	imaging studies and for their free access.
+	We chose these dataset as they represent common dataset sizes for neuroimaging 
+	imaging studies and are open-access.
 }
@@
 }
+\TG{Explain what each column represents}
 
@@
 	their APIs.
-}
-
-The scripts used for our experiments and the results generated are available on
+} The scripts used for our experiments and the results generated are available on
 GitHub at: \mbox{\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}}.
@@
 are performed with only half of the BigBrain dataset ({\SI{303}{\gibi\byte}}).
-\HL{We chose this applications for its real-life use case, large data shuffling between workers,
-and requirements to keep all data in-memory.}
+\HL{We chose this application as it involves substantial data shuffling between workers,
+and requires to keep all the data in memory.}
 
@@
 \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
-\HL{We chose this application for its real-life use case and it is containerized unlike our applications.}
+\HL{We chose this application for its real-life use case and the fact that it is containerized unlike our other applications.}
 
@@
 \HL{
-	With a large amount of worker threads accessing the Lustre file system
-	concurrently, the disk bandiwdth or network are saturated and it create 
-	an I/O bottleneck to access the Lustre file system.
+	Due to the large amount of worker threads accessing the Lustre file system
+	concurrently, the disk bandiwdth or network are saturated, which results in an I/O bottleneck.
 }
@@
 \HL{
-	On the one hand, when multiple threads access Lustre concurrently they risk 
-	of saturation the shared disk bandwidth, thus resulting in longer I/O time.
+	On the one hand, when multiple threads access Lustre concurrently they may saturate the shared disk bandwidth, resulting in longer I/O time.
 	On the other hand, when most threads are either idle or performing computations,
-	it reduces the contention on the disk bandwidth resulting in shorter I/O time.
+	the contention on the disk bandwidth is reduced, resulting in shorter I/O time.
 }
@@
 	It is unclear what is the cause for reporting extra Spark processes.
-	However, with our analysis we found that both Dask and Spark use the same 
+	However, we found that both Dask and Spark use the same 
 	amount of maximum concurrent processes or threads.
-	Additionally, the is no significant difference in the average amount of 
+	Additionally, there was no significant difference in the average amount of 
 	concurrent processes or threads between Dask and Spark.
@@
 	These results are consistent with a previous study~{\cite{Mehta:17}}.
-	However, other studies found the opposite~{\cite{8588652}} or that the difference was negligible~{\cite{8943502}}
+	However, other studies found the opposite~{\cite{8588652}} or that the difference was negligible~{\cite{8943502}}.
 }
@@
 	We found that Dask integration with the scientific Python ecosystem made it
-	easier to implement neuroimaging application.
+	easier to implement neuroimaging applications.
 	Spark configuration required to set multiple environment variables.
-	In comparison, Dask is ready-to-go after an installation through PIP.
-	Dask dashboard provides more information than Spark ones, making
-	the debugging of Dask application easier and faster.
+	In comparison, Dask is ready-to-go after an installation through PIP, the Python package manager.
+	The Dask dashboard provides more information than Spark's, making
+	the debugging of Dask applications easier and faster.
 	Therefore, for the neuroimaging field, we recommend the use of Dask over Spark.
-	However, in other fields with non-Python-based ecosystem, both Dask and spark
-	would be equal candidate as a Big Data engine.
+	However, in other fields with non-Python-based ecosystem, both Dask and Spark
+	may be equal candidates as a Big Data engine.
 }

### 2022-09-12 latex/paper.tex
[FIX] address comments + [FIX] grammar / syntax errors.

--- 2022-09-09 latex/paper.tex
@@
 \HL{
-	With the rising need for Big Data engines in neuroimaging and none commonly used
-	in the field, it is important to evaluate performance of different engines.
+	With the rising need for Big Data engines in neuroimaging and none commonly
+	used in the field, evaluating the performance of different frameworks is essential.
 }
@@
 \HL{
-	While our infrastructure is not of the scale of TOP500 HPC systems,
-	it is hard to reserved multiple nodes concurrently on those infrastructure.
-	Moreover, by using our infrastructure, we can assure that no external processes
-	are influencing the benchmark for our experiements.
+	While our infrastructure is not of the scale of TOP500 HPC systems, it is 
+	hard to reserve multiple nodes concurrently on those infrastructures.
+	Moreover, by using our infrastructure, we can ensure that no external
+	processes influence the benchmark for our experiments.
 	Our system uses similar hardware as Compute Canada's B\'eluga cluster
@@
 	replicates as close as possible its configuration.
-	This lets us compare Spark and Dask in an isolated environement with the condition 
-	of a realistic  allocation of resources, for neuroimaging applications, in an HPC system.
+	This lets us compare Spark and Dask in an isolated environment with the
+	condition of a realistic allocation of resources, for neuroimaging
+	applications, in an HPC system.
 }
@@
 	We configure Dask to have eight worker processes per node, each with eight threads.
-	This is the default recommended configuration for the Dask SLURM cluster configuration~{\cite{DaskSLURMDoc}}.
-	Similarly, we configured Spark to have eight executors per node (hereafter called workers for consistency with Dask), each with eight threads.
-	When trying different configuration for the workers, we did not observe performance difference
-	for our benchmark.
+	This is the default recommended configuration for the Dask SLURM cluster 
+	configuration~{\cite{DaskSLURMDoc}}.
+	Similarly, we configured Spark to have eight executors per node
+	(hereafter called workers for consistency with Dask), each with eight threads.
+	However, when trying different configurations for the workers, we did not
+	observe performance differences for our benchmark.
 	Therefore we chose the recommended configuration from Dask.
@@
 \HL{
-	We chose these dataset as they represent common dataset size for neuroimaging 
-	imaging studies and for their free access.
+	We chose these datasets as they represent typical dataset sizes for neuroimaging studies and their free access.
 }
@@
 	We chose these neuroimaging data processing applications to study the 
-	effects from various Big Data characteristics.
+	effects of various Big Data characteristics.
 	Table~{\ref{table:applications}} depicts the Big Data characteristics used in
@@
 	code base of the applications is common for both engines.
-	Therefore, the performance observed comes directly from the engine scheduling and
-	their APIs.
+	Therefore, the performance observed comes directly from the 
+	engine scheduling and APIs.
 }
@@
 are performed with only half of the BigBrain dataset ({\SI{303}{\gibi\byte}}).
-\HL{We chose this applications for its real-life use case, large data shuffling between workers,
+\HL{We chose this application for its real-life use case, large data shuffling between workers,
 and requirements to keep all data in-memory.}
@@
 \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
-\HL{We chose this application for its real-life use case and it is containerized unlike our applications.}
+\HL{We chose this application for its real-life use case and it is containerized unlike our other applications.}
 
@@
 \HL{
-	The choice of parameters for our experiements was restrained by the datasets
+	The choice of parameters for our experiments was restrained by the datasets
 	and resources available from the cluster.
@@
 \HL{
-	With a large amount of worker threads accessing the Lustre file system
-	concurrently, the disk bandiwdth or network are saturated and it create 
+	With a large number of worker threads accessing the Lustre file system
+	concurrently, the disk bandwidth or network is saturated, and it creates 
 	an I/O bottleneck to access the Lustre file system.
@@
 \HL{
-	On the one hand, when multiple threads access Lustre concurrently they risk 
-	of saturation the shared disk bandwidth, thus resulting in longer I/O time.
+	On the one hand, when multiple threads access Lustre concurrently, they risk 
+	saturation of the shared disk bandwidth, thus resulting in longer I/O time.
 	On the other hand, when most threads are either idle or performing computations,
@@
 	than Spark, resulting in a $1.79\times$ to $2.71\times$ speed-up ratio on the makespan.
-	The problem we observed was two fold.
+	The problem we observed was twofold.
 	Firstly, the process ID of the workers' threads is not reported correctly and 
 	leading to falsely excessive overhead for Spark.
-	Secondly, the Spark workers were restarting due to lack of memory, leading 
+	Secondly, the Spark workers were restarting due to a lack of memory, leading 
 	to task recomputation and having to reload data into memory.
@@
 \HL{
-	It is unclear what is the cause for reporting extra Spark processes.
-	However, with our analysis we found that both Dask and Spark use the same 
+	It is unclear what the cause is for reporting extra Spark processes.
+	However, with our analysis, we found that both Dask and Spark use the same 
 	amount of maximum concurrent processes or threads.
@@
 \subsection{Memory management}
-% TODO remore reference to figure.
-\HL{Our initial Kmeans results showe a significant difference between Dask and Spark.}
+\HL{Our initial Kmeans results showed a significant difference between Dask and Spark.}
 The difference in performance \HL{came} from Spark worker restarting due to running out of memory; thus, Spark \HL{needed} to recompute several tasks.
 This is unexpected, given that Dask and Spark \HL{were} configured similarly and \HL{were} allocated the same amount of resources.
-We think those issues arise from a combination of
+We think those issues \HL{arose} from a combination of
 (1) the reserved memory for the JVM heap space,
@@
 \HL{
-	In this study, we observed that Dask has a larger startup overhead that Spark.
+	In this study, we observed that Dask has a larger startup overhead than Spark.
 	These results are consistent with a previous study~{\cite{Mehta:17}}.
-	However, other studies found the opposite~{\cite{8588652}} or that the difference was negligible~{\cite{8943502}}
+	However, other studies found the opposite~{\cite{8588652}} or that the difference was negligible~{\cite{8943502}}.
 }
@@
 
+\HL{
+	Another study~{\cite{Slaughter2020-ql}} developed a parametrized benchmark framework
+	to evaluate the performance of different engines, including Spark and Dask.
+	On the one hand, the framework reduces the implementation effort and helps with
+	comparing frameworks fairly.
+	On the other hand, it is hard to characterize neuroimaging realistically
+	applications using their framework since the runtime of these applications can vary
+	greatly depending on the subject process, and the framework lacks kernels with
+	a variable amount of compute-intensity.
+	Due to these limitations, we decided to implement the benchmarked applications
+	directly into Spark and Dask.
+}
+
 \subsection{Recommendations}
@@
 	The results from our experiments show that the performance difference between
-	Dask and Spark is minimal, in general. 
+	Dask and Spark are minimal, in general. 
 	The main differences we observed are qualitative.
-	We found that Dask integration with the scientific Python ecosystem made it
-	easier to implement neuroimaging application.
-	Spark configuration required to set multiple environment variables.
-	In comparison, Dask is ready-to-go after an installation through PIP.
-	Dask dashboard provides more information than Spark ones, making
-	the debugging of Dask application easier and faster.
-	Therefore, for the neuroimaging field, we recommend the use of Dask over Spark.
-	However, in other fields with non-Python-based ecosystem, both Dask and spark
-	would be equal candidate as a Big Data engine.
+	First, we found that Dask integration with the scientific Python ecosystem made it
+	easier to implement neuroimaging applications.
+	Spark configuration requires setting multiple environment variables.
+	In comparison, Dask is ready to go after an installation through PIP.
+	In addition, Dask dashboard provides more information than Spark ones, making
+	the debugging of Dask applications easier and faster.
+	Therefore, for the neuroimaging field, we recommend using Dask over Spark.
+	However, in other fields with non-Python-based ecosystems, both Dask and Spark
+	would be equal candidates as Big Data engines.
 }
@@
 \HL{
-	% TODO add caveats to avoid when using the engines.
-
+	While the engines showed similar performance, they both come with their caveats that
+	can drastically affect their performance.
+	For Dask, the user should be aware of the Python GIL limitations.
+	For Python code, multiples worker processes should be used while for code releasing the
+	GIL, a balance between processes and thread is recommended.
+	For Spark, we experienced severe performance issues due to workers running out of memory.
+	Users should carefully consider their Spark worker configurations to prevent similar issues.
 }
@@
 Spark could benefit from using HDFS as a file system, but HDFS is uncommon in HPC.
-\HL{Finally, our Lustre filesystem suffer from an I/O bottleneck coming from 
-either the network or disks bandwidth.}
+\HL{Finally, our Lustre filesystem suffers from an I/O bottleneck coming from 
+either the network or disk bandwidth.}
 As explained in section~{\ref{subsec:overhead-tradeoff}}, the data transfer 
@@
 \HL{
-	% TODO discuss UCX and other high performance networks.???
-
+	While both Dask and Spark support various high-performance network strategies
+    such as UCX, we focused on the standpoint of neuroscientists that would 
+    commonly use the frameworks mostly out of the box.
+    With higher network throughput, the overall performance of the engines could differ.
+    Further work would be required to compare Dask and Spark using different high-performance
+    network protocols.
 }
@@
 \HL{
-	Dask and Spark both support GPUs, however we did not use applications that
+	Dask and Spark both support GPUs; however, we did not use applications that
     leverage them in our work.
     Although several neuroimaging applications use GPUs, only a few preprocessing pipelines use them.
-	With preprocessing comprising the majority of compute time for neuroimaging
-	workflow, we ommitted GPUs from this study.
+    Therefore, we omitted GPUs from this study because preprocessing comprises the majority of computing time for neuroimaging
+    workflow.
     Moreover, the few preprocessing applications using GPUs are not commonly used

### 2022-09-12 latex/paper.tex

@@
 \HL{
-	While our infrastructure is not of the scale of TOP500 HPC systems, it is 
-	hard to reserve multiple nodes concurrently on those infrastructures.
-	Moreover, by using our infrastructure, we can ensure that no external
-	processes influence the benchmark for our experiments.
+	While our infrastructure is not of the scale of TOP500 HPC systems,
+	we can assure that no external processes
+	impacted our measurements which would be hard to achieve on a production system.
 	Our system uses similar hardware as Compute Canada's B\'eluga cluster
 	--- a representative cluster available to researchers in Canada --- and 
-	replicates as close as possible its configuration.
-	This lets us compare Spark and Dask in an isolated environment with the
-	condition of a realistic allocation of resources, for neuroimaging
-	applications, in an HPC system.
+	replicates as closely as possible its configuration.
+	This lets us compare Spark and Dask in an isolated environement with the condition 
+	of a realistic  allocation of resources, for neuroimaging applications, in an HPC system.
 }
@@
 \HL{
-	We chose these datasets as they represent typical dataset sizes for neuroimaging studies and their free access.
+	We chose these dataset as they represent common dataset sizes for neuroimaging 
+	imaging studies and are open-access.
 }
@@
 	code base of the applications is common for both engines.
-	Therefore, the performance observed comes directly from the 
-	engine scheduling and APIs.
-}
-
-The scripts used for our experiments and the results generated are available on
+	Therefore, the performance observed comes directly from the engine scheduling and
+	their APIs.
+} The scripts used for our experiments and the results generated are available on
 GitHub at: \mbox{\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}}.
@@
 are performed with only half of the BigBrain dataset ({\SI{303}{\gibi\byte}}).
-\HL{We chose this application for its real-life use case, large data shuffling between workers,
-and requirements to keep all data in-memory.}
+\HL{We chose this application as it involves substantial data shuffling between workers,
+and requires to keep all the data in memory.}
 
@@
 \href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}
-\HL{We chose this application for its real-life use case and it is containerized unlike our other applications.}
+\HL{We chose this application for its real-life use case and the fact that it is containerized unlike our other applications.}
 
@@
 \HL{
-	With a large number of worker threads accessing the Lustre file system
-	concurrently, the disk bandwidth or network is saturated, and it creates 
-	an I/O bottleneck to access the Lustre file system.
+	Due to the large amount of worker threads accessing the Lustre file system
+	concurrently, the disk bandiwdth or network are saturated, which results in an I/O bottleneck.
 }
@@
 \HL{
-	On the one hand, when multiple threads access Lustre concurrently, they risk 
-	saturation of the shared disk bandwidth, thus resulting in longer I/O time.
+	On the one hand, when multiple threads access Lustre concurrently they may saturate the shared disk bandwidth, resulting in longer I/O time.
 	On the other hand, when most threads are either idle or performing computations,
-	it reduces the contention on the disk bandwidth resulting in shorter I/O time.
+	the contention on the disk bandwidth is reduced, resulting in shorter I/O time.
 }
@@
 \HL{
-	It is unclear what the cause is for reporting extra Spark processes.
-	However, with our analysis, we found that both Dask and Spark use the same 
+	It is unclear what is the cause for reporting extra Spark processes.
+	However, we found that both Dask and Spark use the same 
 	amount of maximum concurrent processes or threads.
-	Additionally, the is no significant difference in the average amount of 
+	Additionally, there was no significant difference in the average amount of 
 	concurrent processes or threads between Dask and Spark.
@@
 \HL{
-	Another study~{\cite{Slaughter2020-ql}} developed a parametrized benchmark framework
+	Another study~{\cite{9355308}} developed a parametrized benchmark framework
 	to evaluate the performance of different engines, including Spark and Dask.
@@
 	The main differences we observed are qualitative.
-	First, we found that Dask integration with the scientific Python ecosystem made it
+	We found that Dask integration with the scientific Python ecosystem made it
 	easier to implement neuroimaging applications.
-	Spark configuration requires setting multiple environment variables.
-	In comparison, Dask is ready to go after an installation through PIP.
-	In addition, Dask dashboard provides more information than Spark ones, making
+	Spark configuration required to set multiple environment variables.
+	In comparison, Dask is ready-to-go after an installation through PIP, the Python package manager.
+	The Dask dashboard provides more information than Spark's, making
 	the debugging of Dask applications easier and faster.
-	Therefore, for the neuroimaging field, we recommend using Dask over Spark.
-	However, in other fields with non-Python-based ecosystems, both Dask and Spark
-	would be equal candidates as Big Data engines.
+	Therefore, for the neuroimaging field, we recommend the use of Dask over Spark.
+	However, in other fields with non-Python-based ecosystem, both Dask and Spark
+	may be equal candidates as a Big Data engine.
 }

### 2022-09-12 latex/paper.tex
[FIX] address comments.

@@
 	each application. 
+	Each row represents an application, while each column represents an engine's characteristic.
+	The \textit{Map} and \textit{Reduce} represent the Map-Reduce paradigm operations,
+	and the \textit{Shuffle} column represents the amount of data shuffled between stages if any.
+	The column \textit{Dataset in-memory} depicts whether the whole dataset is required to 
+	be in memory for processing.
+	Then \textit{Compute intensive} and \textit{Data intensive} shows if the applications are
+	compute or data intensive.
+	Lastly, the \textit{Containerized} column shows applications wrapped in a container.
 }
-\TG{Explain what each column represents}
 

### 2022-09-12 latex/paper.tex
[ADD] citation for Datalad and SAM.

@@
 \SI{121.2}{\mebi\byte}, respectively.
-We used the \href{https://github.com/big-data-lab-team/sam}{sam} library to split the image.
+We used the \href{https://github.com/big-data-lab-team/sam}{sam}~{\cite{sam}} library to split the image.
 
@@
 \cite{zuo2014open}, freely available on
-\href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}.
+\href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{DataLad}~{\cite{DataladDataset}}.
 The entire dataset is \SI{379.83}{\gibi\byte}, containing anatomical, diffusion,

### 2022-09-12 latex/paper.tex
[ADD] limitation on benchmark application used.

@@
 
+\HL{
+	Due to the limited amount of benchmarks available to evaluate the performance
+    of neuroimaging applications, we use our collection of applications.
+    Multiple studies using their own applications make it harder to compare
+    their results.
+    Further efforts would be needed to create a credible and in-depth performance
+    benchmark specific to the neuroimaging field.
+}
+
 \section{Conclusion}

### 2022-09-12 latex/paper.tex
minor edits

--- 2022-09-12 latex/paper.tex
@@
 	each application. 
-	Each row represents an application, while each column represents an engine's characteristic.
-	The \textit{Map} and \textit{Reduce} represent the Map-Reduce paradigm operations,
+	In this table, \textit{Map} and \textit{Reduce} represent the Map-Reduce paradigm operations,
 	and the \textit{Shuffle} column represents the amount of data shuffled between stages if any.
@@
 
+\HL{
+	Due to the limited amount of benchmarks available to evaluate the performance
+    of neuroimaging applications, we use our own collection of applications.
+    Further efforts would be needed to create a credible and in-depth performance
+    benchmark specific to the neuroimaging field.
+}
+
 \section{Conclusion}

### 2022-10-24 latex/paper.tex
[UPDATE] Final version for proceeding.

@@
 \renewcommand{\thefootnote}{\fnsymbol{footnote}}
-\newcommand{\HL}[1]{\hl{#1}}
+\newcommand{\HL}[1]{#1}
 \newcommand{\TG}[1]{\color{red}\textsc{From Tristan: }#1\color{black}}
 
-\articletype{WORKS19 Special Issue on Workflows in Support of Large-Scale Science GEs:  Chase Wu, Hoang Nguyen}
+\articletype{WORKS19 Special Issue on Workflows in Support of Large-Scale Science GEs:  David Walker, Chase Wu}
 
@@
 \section*{Acknowledgment}
-Mathieu Dugr\'e was funded by the Alexander Graham Bell scholarship from
-the Natural Sciences and Engineering Research Council of Canada.
+Mathieu Dugr\'e was partially funded by a Alexander Graham Bell scholarship from the Natural
+Sciences and Engineering Research Council of Canada and a Master's (B1X) Research Scholarships.
 The Canada Research Chairs program partially funded this work.
@@
 
+\section*{Data Availability}
+\subsection*{BigBrain}
+The data that support the findings of this study are available from the corresponding author, M. Dugr\'e, upon reasonable request.
+A lower resolution of \SI{40}{\micro\metre} is available publicly by FTP at: 
+\newline\href{https://ftp.bigbrainproject.org/bigbrain-ftp/BigBrainRelease.2015/3D_Blocks/40um/nii/}{https://ftp.bigbrainproject.org/bigbrain-ftp/BigBrainRelease.2015/3D\_Blocks/40um/nii}
+
+\subsection*{CoRR}
+The data that support the findings of this study are openly available in DataLad at:
+\newline\href{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}{https://datasets.datalad.org/?dir=/corr/RawDataBIDS}
+
+\subsection*{Code}
+The code that support the findings of this study is openly available in GitHub at:
+\newline\href{https://github.com/big-data-lab-team/paper-big-data-engines/tree/CPE}{https://github.com/big-data-lab-team/paper-big-data-engines/tree/CPE}
+
+\section*{Conflict of Interests}
+The authors declare no conflict of interests.
+
 \bibliography{paper}

### 2022-10-24 launch_experiments.py

@@ -9,7 +9,7 @@ REPETITIONS = 10
 
 n_nodes = [2, 4, 8]
 n_iterations = [1, 5, 25]
-n_iterations_small = [1, 3, 9]
+n_iterations_small = [1, 3]
 BB_blocks = [1000, 2500, 5000]
 
 default = {
@@ -51,15 +51,15 @@ for cmd_template in cmd_templates:
                 benchmark_folder,
             )
         )
-        # cmds.append(
-        #     cmd_template.format(
-        #         BB_sample_path(5000),
-        #         "multi-increment",
-        #         f"{5000} {default['small_itr']} {default['sleep']}",
-        #         x,
-        #         benchmark_folder,
-        #     )
-        # )
+        cmds.append(
+            cmd_template.format(
+                BB_sample_path(5000),
+                "multi-increment",
+                f"{5000} {default['small_itr']} {default['sleep']}",
+                x,
+                benchmark_folder,
+            )
+        )
         cmds.append(
             cmd_template.format(
                 BB_sample_path(5000),
@@ -119,15 +119,15 @@ for cmd_template in cmd_templates:
             )
         )
     for x in n_iterations_small:
-        # cmds.append(
-        #     cmd_template.format(
-        #         BB_sample_path(5000),
-        #         "multi-increment",
-        #         f"{5000} {x} {default['sleep']}",
-        #         default["node"],
-        #         benchmark_folder,
-        #     )
-        # )
+        cmds.append(
+            cmd_template.format(
+                BB_sample_path(5000),
+                "multi-increment",
+                f"{5000} {x} {default['sleep']}",
+                default["node"],
+                benchmark_folder,
+            )
+        )
         cmds.append(
             cmd_template.format(
                 BB_sample_path(5000),

### 2022-10-24 latex/paper.tex
[UPDATE] link for github repository.

@@
 } The scripts used for our experiments and the results generated are available on
-GitHub at: \mbox{\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}}.
+GitHub at: \mbox{\href{https://github.com/big-data-lab-team/paper-big-data-engines/tree/CPE}{https://github.com/big-data-lab-team/paper-big-data-engines/tree/CPE}}.
 
